### 2020-04-15 paper_io_sim.tex
First outline version

\documentclass[11pt, a4paper]{article}

\begin{document}
\title{I/O simulation extension for SimGrid framework}
\author{Hoang-Dung Do}
\maketitle

\begin{abstract}

\end{abstract}

	\section{Introduction}

	\section{Related Work}

		\subsection{Page cache}
			\begin{itemize}
				\item What is page cache? How it works.
				\item Effects and importance of page cache
			\end{itemize}

		\subsection{ Page cache replacement policies}
			\begin{itemize}
				\item Introduce some proposed strategies.
				\item Some pros and cons of each algorithm if possible.
			\end{itemize}

		\subsection{Implementation in Linux}
			\begin{itemize}
				\item Details of the implemented strategy in Linux.
				\item Reason for this to be implemented. (implementation complexity, effectiveness, overhead, etc)
			\end{itemize}

		\subsection{Simulators}
		Compare pros and cons of some simulators: SimGrid, GridSim (accuracy, simulation time, usability).

		\subsection{Our conclusion}

	\section{Method}

		\subsection{Principle of the simulator}

			\begin{itemize}
				\item Objective: Add capability to simulate memory read/write, the impact of page cache on I/O in SimGrid.
				\item Approach: generalize dirty data, dirty ratio, cache eviction strategy implemented in Linux. 
			\end{itemize}

		\subsection{Implementation}
			\begin{itemize}
				\item Which features of memory are implemented.
				\item Level of granularity, how features are implemented.
				\item Specific implementation detail in python and SimGrid.
			\end{itemize}

		\subsection{Experiments}
			Describe data, workflow, number of tasks, task details, environment of each experiment.

			\subsubsection{Expriment 1}
				A single pipeline running one node.
			\subsubsection{Expriment 2}
				Multiple pipelines running in parallel on multiple nodes.
			\subsubsection{Expriment 3}
				Same as Experiment 2 but nodes write to a shared file system.
			\subsubsection{Expriment 4}
				A real pipeline (for example a pipeline with nighres)

	\section{Results}

		\begin{itemize}

			\item Quantized results: 
				\begin{itemize}
					\item Errors of simulation time and memory used compared to real results.
					\item Simulation time compared to baseline SimGrid.
				\end{itemize}

			\item Ability of the model to generalize trends of memory usage and disk throughput.

		\end{itemize}

	\section{Discussion}

\end{document}

### 2020-04-21 paper_io_sim.tex
Update
paper outline

@@
-\documentclass[11pt, a4paper]{article}
+\documentclass[conference]{IEEEtran}
 
@@
 \begin{abstract}
-
+	\begin{itemize}
+		\item The I/O bottleneck in HPC and the need of conducting HPC experiments.
+		\item Difficulties in HPC experiments and simulation frameworks.
+		\item Advantages of SimGrid, the missing of the ability to simulate page cache, the goal of this paper.
+	\end{itemize}
 \end{abstract}
@@
 	\section{Introduction}
-
+		\begin{itemize}
+			\item Role of high performance computing nowadays and the demand of HPC experiments.
+			\item Difficulties in conducting high performance computing experiments and the need of simulation frameworks.
+			\item Existing methods, simulators, simulation frameworks. 
+			\item The advantages of SimGrid compared to others. The missing of the ability to simulate page cache in SimGrid.
+			\item The objective of the paper: Add capability to simulate memory read/write, the impact of page cache on I/O in SimGrid.
+		\end{itemize}
 	\section{Related Work}
@@
 			\begin{itemize}
-				\item What is page cache? How it works.
-				\item Effects and importance of page cache
-			\end{itemize}
-
-		\subsection{ Page cache replacement policies}
-			\begin{itemize}
-				\item Introduce some proposed strategies.
-				\item Some pros and cons of each algorithm if possible.
-			\end{itemize}
-
-		\subsection{Implementation in Linux}
-			\begin{itemize}
-				\item Details of the implemented strategy in Linux.
-				\item Reason for this to be implemented. (implementation complexity, effectiveness, overhead, etc)
+				\item What is page cache? How it works. Effects and importance of page cache
+				\item Introduce some existing strategies.
+				\item Current implementation in Linux and some reasons for this to be implemented (implementation complexity, effectiveness, overhead, etc)
 			\end{itemize}
@@
 		\subsection{Simulators}
-		Compare pros and cons of some simulators: SimGrid, GridSim (accuracy, simulation time, usability).
-
-		\subsection{Our conclusion}
+		Discuss some ecisting methods, simulation frameworks to conduct HPC experiments. Compare pros and cons (accuracy, simulation time, usability) of some simulators (SimGrid, GridSim) .
 
@@
 			\begin{itemize}
-				\item Objective: Add capability to simulate memory read/write, the impact of page cache on I/O in SimGrid.
 				\item Approach: generalize dirty data, dirty ratio, cache eviction strategy implemented in Linux. 
+				\item Some implemented details of the simulator in SimGrid.
 			\end{itemize}
@@
 
-			\item Ability of the model to generalize trends of memory usage and disk throughput.
+			\item Ability of the model to generalize trends seen in memory (amount of dirty data, page cache) and disk throughput.
 
@@
 
-	\section{Discussion}
+	\section{Discussion and Future Work}
 

### 2020-04-22 paper_io_sim.tex
update outline

@@
 		\begin{itemize}
-		\item The I/O bottleneck in HPC and the need of conducting HPC experiments.
-		\item Difficulties in HPC experiments and simulation frameworks.
-		\item Advantages of SimGrid, the missing of the ability to simulate page cache, the goal of this paper.
+			\item The I/O bottleneck in HPC and the need of experiments.
+			\item HPC experiment frameworks and advantages of SimGrid.
+			\item The missing of the ability to simulate page cache, the goal of this paper.
+			\item Principle of the simulator, experiment scenarios and comparisons.
+			\item Brief discussion on results and future work.
 		\end{itemize}
@@
 		\begin{itemize}
-			\item Role of high performance computing nowadays and the demand of HPC experiments.
+			\item HPC, the bottleneck in I/O and the demand of HPC experiments. 
 			\item Difficulties in conducting high performance computing experiments and the need of simulation frameworks.
-			\item Existing methods, simulators, simulation frameworks. 
-			\item The advantages of SimGrid compared to others. The missing of the ability to simulate page cache in SimGrid.
-			\item The objective of the paper: Add capability to simulate memory read/write, the impact of page cache on I/O in SimGrid.
+			\item Existing experiment methods, simulators, simulation frameworks. The advantages of SimGrid compared to others. The missing of the ability to simulate page cache in SimGrid.
+			\item The objective of the paper: Add capability to simulate I/O with page cache in SimGrid.
 		\end{itemize}
@@
 			\begin{itemize}
-				\item What is page cache? How it works. Effects and importance of page cache
-				\item Introduce some existing strategies.
-				\item Current implementation in Linux and some reasons for this to be implemented (implementation complexity, effectiveness, overhead, etc)
+				\item What is page cache? How it works. Effects and importance of page cache.
+				\item Introduce some existing strategies with some highlighted pros and cons.
+				\item Current implementation in Linux and some reasons why it is chosen to be implemented (implementation complexity, effectiveness, overhead, etc)
 			\end{itemize}
@@
 		\subsection{Simulators}
-		Discuss some ecisting methods, simulation frameworks to conduct HPC experiments. Compare pros and cons (accuracy, simulation time, usability) of some simulators (SimGrid, GridSim) .
+			\begin{itemize}
+				\item Discuss some existing methods, simulation frameworks to conduct HPC experiments. Compare pros and cons (accuracy, simulation time, usability) of some simulators (SimGrid, GridSim).
+				\item Discuss the pros of SimGrid and the reasons why we chose it to extend.
+			\end{itemize}
 
@@
 				\item Approach: generalize dirty data, dirty ratio, cache eviction strategy implemented in Linux. 
-				\item Some implemented details of the simulator in SimGrid.
+				\item Some implemented details of the simulator.
 			\end{itemize}
@@
 				\item Level of granularity, how features are implemented.
-				\item Specific implementation detail in python and SimGrid.
+				\item Specific implementation details in python and SimGrid.
 			\end{itemize}

### 2020-04-22 paper_io_sim.tex
Add references

@@
 			\item Difficulties in conducting high performance computing experiments and the need of simulation frameworks.
-			\item Existing experiment methods, simulators, simulation frameworks. The advantages of SimGrid compared to others. The missing of the ability to simulate page cache in SimGrid.
+			\item Existing experiment methods, simulators, simulation frameworks.
+The advantages of SimGrid compared to others \cite{casanova2008, lebre2015}.
+The missing of the ability to simulate page cache in SimGrid \cite{lebre2015}.
 			\item The objective of the paper: Add capability to simulate I/O with page cache in SimGrid.
@@
 			\begin{itemize}
-				\item What is page cache? How it works. Effects and importance of page cache.
+				\item What is page cache? How it works \cite{linuxdev3rd2010}. Effects and importance of page cache.
 				\item Introduce some existing strategies with some highlighted pros and cons.
-				\item Current implementation in Linux and some reasons why it is chosen to be implemented (implementation complexity, effectiveness, overhead, etc)
+				\item Current implementation in Linux and some reasons why it is chosen to be implemented (implementation complexity, effectiveness, overhead, etc) \cite{linuxdev3rd2010}
 			\end{itemize}
@@
 
+\bibliographystyle{plain}
+\bibliography{citation}
+
 \end{document}

### 2020-04-29 paper_io_sim.tex
Update discussion, add refs

@@
 				\item Discuss some existing methods, simulation frameworks to conduct HPC experiments. Compare pros and cons (accuracy, simulation time, usability) of some simulators (SimGrid, GridSim).
+				\item Related development: RAM energy consumption \cite{gill2019} \cite{ouarnoughi2017} 
 				\item Discuss the pros of SimGrid and the reasons why we chose it to extend.
@@
 		\subsection{Experiments}
-			Describe data, workflow, number of tasks, task details, environment of each experiment.
+			For each experiment: 
+			\begin{itemize}
+				\item Describe data, workflow, number of tasks, task details, environment of each experiment.
+				\item Compare the results from real pipeline, baseline SimGrid  and SimGrid with our extension.
+			\end{itemize}
 
@@
 
-			\item Quantized results: 
+			\item Quantitative results: 
 				\begin{itemize}
@@
 
-			\item Ability of the model to generalize trends seen in memory (amount of dirty data, page cache) and disk throughput.
+			\item Ability of the model to generalize memory trends (dirty data, cache used) and disk throughput.
 
@@
 	\section{Discussion and Future Work}
-
+		\begin{itemize}
+			\item Sensitivity of the simulator on the variation of memory and disk bandwidth. 
+		\end{itemize}
 \bibliographystyle{plain}

### 2020-05-10 paper_io_sim.tex
Section 3.A: principle of the simulator

@@
 				\item Related development: RAM energy consumption \cite{gill2019} \cite{ouarnoughi2017} 
-				\item Discuss the pros of SimGrid and the reasons why we chose it to extend.
+				\item Discuss the pros of SimGrid and the reasons why we chose it to extend. (Section 2.2.2 in \cite{casanova2014})
 			\end{itemize}
@@
 	\section{Method}
-
+		In this section, we discuss our approach to model page cache and the cache eviction strategy implemented in Linux.
+We also detail the design the simulators in order to generalize page cache as well as the implementation in Python and SimGrid framework.
+Finally, we describe some experimental scenarios to compare the results of Python simulator, baseline SimGrid simulator and the results from the real pipelines.
 		\subsection{Principle of the simulator}
 
-			\begin{itemize}
-				\item Approach: generalize dirty data, dirty ratio, cache eviction strategy implemented in Linux. 
-				\item Some implemented details of the simulator.
-			\end{itemize}
+			In order to make the simulator accurate, we try to simulate simulate page cache and mimic the 2-LRU lists cache eviction algorithm implemented in Linux.
+Our design also keeps track two LRU lists ans specific system configurations as in Linux kernel.
+However, maintaining and handling pages, which is in packet level, requires significant implementation overhead and hinders the performance of the simulator.
+To simplify the model whilst preserving accuracy, in our approach, we consider a sequence of continuously accessed pages of the same file from the same device as a single unit, a \textit{block} of data.
+The information of a single page in a sequence is representative for other pages the a sequence since they have the same properties including file name, last accessed, dirty or not.
+Thus, instead of dealing with individual pages, we view a sequence of continuously accessed pages as a block of data.
+A data block has the information about file name, size, last access and a dirty flag.
+As a block represents a set of pages in real systems, it can be split if some data needs to be moved between the LRU lists or evicted from cache, but blocks can not be merged as we want to preserve the access times of blocks to maintain the accuracy.
+
+			With this data block concept, when a file is accessed by an I/O operation, blocks are added to the top of inactive list if they are not existing in cache, or they are moved to the top of the list that they are kept in with last accesses updated.
+The lists are balanced after every time the simulator finishes reading or writing a portion of data from/to cache or disk.
+
+			To simulate flushing and periodical flushing mechanisms, in which dirty data in cache is written to disk, we simply calculate the amount of flushed dirty data.
+Data flushing is called when the memory required by the application is more than available memory, while periodical flushing is called during CPU time.
+In flushing, the amount of dirty data to be flushed is calculated based on the amount of required memory, while the flushed amount in periodical flushing is estimated with CPU time.
+Initially, the blocks in inactive list are flushed first, and the next are dirty blocks in the active list, until the total flushed amount is reached.
+The flushed blocks are split if needed and become not dirty.
+The blocks in the inactive list which are not dirty are available for cache eviction.
+
+			Similar to flushing and periodical flushing, the amount of evicted data is calculated based on the the amount of memory required by application and the amount of free memory.
+Then, evicted blocks are deleted from inactive list, and the lists are balanced.
 
@@
 				\item Describe data, workflow, number of tasks, task details, environment of each experiment.
-				\item Compare the results from real pipeline, baseline SimGrid  and SimGrid with our extension.
+				\item How we compare the results from real pipeline, baseline SimGrid  and SimGrid with our extension.
 			\end{itemize}

### 2020-05-27 paper_io_sim.tex
Section 3 v1

--- 2020-04-29 paper_io_sim.tex
@@
 \documentclass[conference]{IEEEtran}
+\usepackage{graphicx}
+\usepackage{algorithm}
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
+\usepackage[dvipsnames]{xcolor}
+\usepackage{amsmath}
+
+\algblock{Input}{EndInput}
+\algnotext{EndInput}
+\algblock{Output}{EndOutput}
+\algnotext{EndOutput}
+\newcommand{\Desc}[2]{\State \makebox[5em][l]{#1}#2}
+
+\newenvironment{conditions}
+  {\par\vspace{\abovedisplayskip}\noindent\begin{tabular}{>{$}l<{$} @{${}={}$} l}}
+  {\end{tabular}\par\vspace{\belowdisplayskip}}
 
@@
 	\section{Method}
-
+		In this section, we discuss our approach to model page cache and the cache eviction strategy implemented in Linux.
+We also detail the design the simulators in order to generalize page cache as well as the implementation in Python and SimGrid framework.
+Finally, we describe some experimental scenarios to compare the results of Python simulator, baseline SimGrid simulator and the results from the real pipelines.
 		\subsection{Principle of the simulator}
 
+			When modeling the I/O time, we consider the key factors of the I/O mechanism having the biggest impacts on data read/write.
+The first factor taken into account is the use of the page cache in Linux kernel.
+In addition, the kernel controls I/O activities based on the status of the memory including the information about the cache size, the amount of available memory and dirty data.
+As these are managed and changed with the flushing and cache eviction mechanisms implemented in the kernel, data flushing and cache eviction must be generalized.
+The simulator should also make system parameters such as dirty{\_}ratio and dirty{\_}background{\_}ratio configurable since they can vary in different systems.
+
+			The idea of our simulator is to generalize the memory, the page cache, the LRU lists, storage device and mimic the cache eviction algorithm implemented in the kernel.
+As for storage devices, we can leverage the storage model with disk capacity and disk bandwidth implemented in SimGrid \cite{lebre2015}.
+Because memory capacity and bandwidth are fixed, they can be kept as constants, while the amount of cache is dynamically allocated based on the amount of available memory.
+The simulator also keeps track of the LRU lists as in the kernel since they are the key component in cache eviction algorithm.
+However, maintaining and handling pages, which are in packet level, are non-trivial when they require significant implementation overhead and hinder the performance of the simulator.
+To simplify cache eviction while preserving accuracy, in our approach, we consider a sequence of continuously accessed pages of a file as a single unit, a block of data.
+The information of a single page in a sequence can be used as the representative for other pages in a sequence since they have the same properties including file name, last access, the page is dirty or not.
+Hence, instead of dealing with individual pages, we consider a sequence of continuously accessed pages as a block of data.
+A data block keeps the information about file name, block size, last access and a dirty flag describing whether the data is dirty or not.
+Since a block represents a set of pages in real systems, it can be split if some data needs to be moved between the LRU lists or evicted from cache, but blocks can not be merged as we want to preserve this level of granularity to maintain the accuracy.
+Having these components, we can generalize I/O activities in our simulator.
+When the kernel reads and writes data, it simply reads/writes data from/to disk or cache with predefined disk bandwidth or memory bandwidth.
+In flushing and periodical flushing, the kernel traverses through the inactive and active lists to reduce the amount of dirty data, while cache eviction algorithm deletes blocks from the inactive list.
+The LRU lists are updated after every I/O action.
+
+			We create an extra layer with kernel-like functionalities between user application and SimGrid as is shown in Figure~\ref{fig:interaction}.
+Our new layer including two main additional classes, the \textbf{\textit{IOController}} class, which mimics the kernel I/O controlling logic, and the \textbf{\textit{MemoryManager}} class, which stores all necessary information about memory including capacity, the amount of cache, amount of memory used by applications, and manages LRU lists.
+Because every node has its own memory, storage devices and system configurations, there is one IOController object and one MemoryManager object created for each node.
+The IOController class keeps system parameters (dirty{\_}ratio, dirty{\_}background{\_}ratio), implements necessary I/O control functionalities such as read, write, flushing, and cache evicting.
+With these new components, instead of sending read/write requests directly to storage devices as in simulators with the current SimGrid version, user simulated applications send requests to an IOController of the node where requested data is stored.
+Based on the memory status in MemoryManager, the IOController orchestrates cache read/write, disk read/write, flushing and cache eviction to fulfill application requests.
+The time for each action is estimated and the total time of a request is returned to user applications.
+
+			\begin{figure}
+   				\centering
+   				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
+   				\caption{The additional layer between user application layer and original SimGrid API}\label{fig:interaction}
+			\end{figure}
+
+			\begin{algorithm}\caption{Read}\label{alg:read}
+				\small
+				\begin{algorithmic}[1]
+					\Input
+        				\Desc{$file\_name$}{file name}
+			        	\Desc{$run\_time$}{current simulated run time}
+   					\EndInput
+					\State $mem\_required \gets max(0, 2* $file size$-cached\_data -$ available memory)
+					\State $mem\_avail \gets max(0, 2 *$ file size$ - cached\_data -$ dirty data)
+					\If {$mem\_required > mem\_avail$}
+    					\State $run\_time \gets run\_time + $  flushing $(mem\_required - mem\_avail)	$
+					\EndIf
+
+					\State $free\_mem\_required \gets max(0, 2 *$ file size $- cached\_data -$ free memory)
+					\If {$free\_mem\_required > 0$}
+						\State evict from cache $(free\_mem\_required)$
+					\EndIf
+
+					\If {$cached\_data > 0$}
+    					\State update cached blocks in LRU lists with the last access
+						\If {all file data is cached}
+    						\State $cache\_read\_time \gets cached\_data /$ memory read bandwidth
+    						\State periodically flush $($flushing time $= cache\_read\_time)$
+
+							\Return $run\_time \gets run\_time + cache\_read\_time$
+						\EndIf
+					\EndIf
+
+					\If {$cached\_data <$ file size}
+						\If {dirty amount $> 0$} 
+							\State $run\_time \gets run\_time +$ dirty amount flushing time
+						\EndIf
+						\State read from disk (amount from disk, $file\_name$), add accessed data to inactive list
+    					\State $run\_time \gets run\_time +$ (amount from disk / disk read bandwidth
+					\EndIf
+
+					\Return run{\_}time
+
+				\end{algorithmic}
+			\end{algorithm}
+
+			The algorithm to generalize file read is described in \textbf{\textit{Algorithm~\ref{alg:read}}}.
+In Linux systems, data can be read in two ways, read from cache with memory bandwidth and read from disk with disk bandwidth.
+Because total read time can be seen as a linear function of the read time in each of these two modes, we can generalize reading process as two separate and sequential phases, one is cache read, and the other is disk read.
+When a user application requests a file read, it calls the read method of the IOController of which a parameter is the file name.
+Then, IOController calculates how much memory is required to read the file based on the amount of available memory, file size, and the amount of file data cached.
+If there is not enough available memory, dirty data is flushed to disk to make more memory available by calling flush function with the amount of memory needed as the parameter.
+IOController then evicts old  blocks from the inactive list to accommodate the requested file.
+After that, IOController starts estimating file read time by checking if a some file data is cached.
+If there is cached data, the blocks containing this data are  updated in the LRU lists with the last access time.
+If all file data is cached, time to read from cache is added to total run time, and periodical flushing is triggered to flush dirty data during the cache read.
+If only a part of the file is in cache, disk read is required, this does not allow flushing and makes the cache read time neglectable.
+At the beginning of a disk read, dirty data is flushed \textcolor{red}{(Is this correct?)}, then data is read and added to the inactive list, disk read time is estimated with disk read bandwidth and added to total run time.
+
+			\textbf{\textit{Algorithm~\ref{alg:write}}} describes our generalization of file write.
+Similar to the file read process, a file can also be written with two different bandwidths, memory bandwidth and disk bandwidth, and file write can be generalized with two modes.
+File can be written with memory bandwidth before the page cache is saturated, or the amount of dirty data reaches dirty\_ratio.
+In our simulator, dirty data can be flushed concurrently during cache write, and the maximum amount of data can be written with memory bandwidth is calculated with the formula:
+			\begin{equation}
+				M = \frac{(dirty\_ratio*C - d)*mw}{mw - dw}
+			\end{equation}
+
+			where:
 			\begin{itemize}
-				\item Approach: generalize dirty data, dirty ratio, cache eviction strategy implemented in Linux. 
-				\item Some implemented details of the simulator.
+				\item $M$ is the maximum amount of data can be written with memory write bandwidth
+				\item $A$ is the total available memory
+				\item $d$ is the amount of dirty data
+				\item $mw$ is the memory write bandwidth
+				\item $dw$ is the disk write bandwidth			
 			\end{itemize}
 
+			If file size is smaller than $M$, the whole file is written to cache as dirty data with memory bandwidth.
+Otherwise, the remaining file data is written with disk bandwidth.
+Because cache write and data flushing are concurrent, the actual amount of written dirty data reduces by the amount of flushed data.
+Before simulating cache write, IOController checks if the page cache needs to be evicted to accommodate the file.
+If yes, evictable data in the inactive list is evicted.
+In the second phase, in which data is written with disk bandwidth, our simulator can generalize not only the case that data is written directly to disk, but also the case that applications have to wait for dirty data being flushed.
+Before writing, if the amount of free memory is not enough for the remaining file data, evictable memory is evicted.
+IOController then simulates the time to write this amount of data, adds written data to cache, but the amount of dirty data remains the same.
+The total simulated time of these two phases is added and returned to user applications.
+
+			\begin{algorithm}\caption{Write}\label{alg:write}
+				\small
+				\begin{algorithmic}[1]
+					\Input
+        				\Desc{$file\_name$}{file name}
+			        	\Desc{$run\_time$}{current simulated run time}
+   					\EndInput
+					\State $avail\_dirty\_amt \gets dirty\_ratio *$ available memory - current dirty amount
+					\If {$avail\_dirty\_amt > 0$}
+   						\State calculate maximum amount $M$
+   						\State $free\_written\_amt \gets min($file size, $M)$
+    					\If {$free\_written\_amt >$ free memory}
+    						\State evict(amount = $free\_written\_amt -$ free memory)
+    					\EndIf
+    					\State $free\_write\_time \gets free\_written\_amt /$ memory write bandwidth
+    					\State write(file name, amount =$free\_written\_amt$, dirty amount=$free\_written\_amt - free\_write\_time *$ disk write bandwidth)
+   						\State $run\_time \gets run_time +free\_write\_time$
+    				\EndIf
+
+					\If {$free\_written\_amt >=$ file size}
+						\Return run\_time
+					\EndIf
+
+					\State $throttled\_amt \gets $ file size - $free\_written\_amt$
+					\State $cache\_required \gets max(0, throttled\_amt  -$ free memory)
+					\State $evicted\_amt \gets min(cache\_required,$ total evictable memory)
+					\If {$evicted\_amt > 0$} 
+						\State  evict($evicted\_amt$)
+					\EndIf
+					\State write(file name, amount to cache = $min($free memory, $throttled\_amt$), dirty data = 0)
+					\State $run\_time \gets = run\_time + throttled\_amt /$ storage write bandwidth
+
+					\Return $run{\_}time$
+
+				\end{algorithmic}
+			\end{algorithm}
+
+			Data flushing is another key function in our simulator.
+The input of the function is the amount of flushed data and the returned value is the simulated flushing time.
+To flush dirty data, IOController simply traverses the sorted inactive and active list, reduces the dirty data amount by unmarking dirty blocks until the flushed amount reached the amount passed in or there is no dirty data left in cache.
+Similarly, periodical flushing is the function that simulates data flushing during CPU time and cache read.
+However, the flushed data amount is calculated with the input which is flushing time.
+In both flushing functions, the last block being flushed is split if its data is not entirely flushed.
+
+			Equally important, cache eviction function is called from both read and write to free up the page cache when free memory is not enough.
+The amount of evicted data is passed in as in flushing, however, this function does not add up the simulated time as cache eviction time is neglectable in real systems .
+When eviction is called, IOController deletes the last accessed blocks from the inactive list until the evicted amount reaches the amount passed in or there is no evictable memory left.
+If the block is not entirely deleted, the block is split.
+
+			Last but not leats, we need to modified LRU lists after every I/O activity.
+When IOController reads the data that is not cache, the data is read and added to the inactive list.
+If the data is already in cache, the cached blocks are moved to the active list with the last access time updated.
+Finally, MemoryManager makes the LRU lists balanced by moving blocks from the longer list to the shorter and sorting the lists.
+
 		\subsection{Implementation}
-			\begin{itemize}
-				\item Which features of memory are implemented.
-				\item Level of granularity, how features are implemented.
-				\item Specific implementation details in python and SimGrid.
-			\end{itemize}
+
+			Firstly, to validate our hypothesis, we create a simple simulator independent to simulation frameworks and libraries.
+This enables us to evaluate the accuracy and correctness of our model in a simple use case before integrating it in SimGrid.
+This simulator simulates a single thread pipeline running on a system with a single core CPU, one local storage device and all input and output files are stored locally.
+The results from the simulator is compared and validated with the results from a real pipeline running on a real system.
+
+			Having our model validated, we create other different simulators of different use cases using the current version of SimGrid and the SimGrid version that is extended with our model.
+Then, we compare the results of the simulators with original SimGrid and extended SimGrid with the results of real pipelines on a real system.
+The results are compared in two different aspects, task completion time and I/O time, as well as the ability of the simulator in generalization of the page cache in terms of dirty data and cache used.
+
+			In this work, we use Python 3.7 to implement the simple simulator, SimGrid 3.25 and Wrench 1.6 for SimGrid simulators.
 
 		\subsection{Experiments}
-			For each experiment: 
-			\begin{itemize}
-				\item Describe data, workflow, number of tasks, task details, environment of each experiment.
-				\item Compare the results from real pipeline, baseline SimGrid  and SimGrid with our extension.
-			\end{itemize}
+
+			When designing experiments, we consider a range of scenarios that have different impacts on the results.
+The experiments are a number of pipelines of sequential tasks.
+Each task reads an input file, performs some computation, generates an output file and writes it to disk.
+The output file of the previous task is the input file of the next task.
+Thus, a pipeline has only one actual input file.
+The size input and output files should vary have significant impacts on system I/O behaviors.
+The number of pipelines should also vary in different experiments since the shared bandwidth is affected by the number of concurrent I/O processes.
+Different storage types, which include local and shared storage, are also taken into account to evaluate the performance of our simulator on shared file systems.
+Finally, we adopt a real neuroimaging pipeline to evaluate the applicability of our simulator in simulating I/O time in real applications.
+Because our work focuses on I/O time, we assume that CPU time is correctly modeled and use the CPU time measured in the real pipelines to setup our simulated workflows.
+
+			We use the cluster of \textit{Big Data Infrastructure for Neuroimaging Lab} at \textit{Concordia University} to conduct our experiments.
+The cluster consists of 1 login node which is the control node, 8 compute nodes and 4 storage nodes connected with 2 network switches.
+The login node has 128 GB of RAM, 1.8 TB of storage and 385 TB of Lustre shared file system.
+Each compute node has 6 SSDs with the size of 450GB each, 32 cores CPU and 256 GB of RAM, while each storage node has 12 HDDs and 2 SSDs.
+The cluster is run on CentOS 8.1 with the Slurm Workload Manager installed.
+We use \textbf{\textit{atop}} and \textbf{\textit{collectl}} as tools to monitor and collect data of memory, page cache status and disk throughput on the cluster.
 
 			\subsubsection{Expriment 1}
-				A single pipeline running one node.
+
+				In this experiment, our goal is to validate our model with a simple simulator implemented in Python.
+Thus, we only simulate one single thread pipeline of three tasks with the Python simulator, SimGrid simulator and run on our real system.
+All files are read and written locally.
+We use the input sizes of 20 GB, 50 GB, 75 GB and 100 GB to run the pipeline on the cluster, measure the CPU time of the pipeline with each input to simulate the pipeline with the simulators.
+
 			\subsubsection{Expriment 2}
-				Multiple pipelines running in parallel on multiple nodes.
+
 			\subsubsection{Expriment 3}

### 2020-06-01 paper_io_sim.tex
Some comments

@@
 
+\newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:}#1\color{black}}
+
 \begin{document}
-\title{I/O simulation extension for SimGrid framework}
+\title{I/O simulation extension \tristan{``extension'' is too general} for SimGrid framework \tristan{``SimGrid framework'' is too specific}}
 \author{Hoang-Dung Do}
@@
 	\section{Method}
-		In this section, we discuss our approach to model page cache and the cache eviction strategy implemented in Linux.
-We also detail the design the simulators in order to generalize page cache as well as the implementation in Python and SimGrid framework.
-Finally, we describe some experimental scenarios to compare the results of Python simulator, baseline SimGrid simulator and the results from the real pipelines.
+		In this section, we discuss our approach to model page cache and
+		the cache eviction strategy implemented in Linux. We also detail
+		the design the simulators in order to generalize page cache as well
+		as the implementation in Python and SimGrid framework. Finally, we
+		describe some experimental scenarios to compare the results of
+		Python simulator, baseline SimGrid simulator and the results from
+		the real pipelines. \tristan{It would help if you could make lines
+		shorter than 80 chars, as is the common practice with code.}
+
 		\subsection{Principle of the simulator}
 
-			When modeling the I/O time, we consider the key factors of the I/O mechanism having the biggest impacts on data read/write.
-The first factor taken into account is the use of the page cache in Linux kernel.
-In addition, the kernel controls I/O activities based on the status of the memory including the information about the cache size, the amount of available memory and dirty data.
-As these are managed and changed with the flushing and cache eviction mechanisms implemented in the kernel, data flushing and cache eviction must be generalized.
-The simulator should also make system parameters such as dirty{\_}ratio and dirty{\_}background{\_}ratio configurable since they can vary in different systems.
+			When modeling the I/O time, we consider the key factors of the
+			I/O mechanism having the biggest impacts on data read/write.
+			The first factor taken into account is the use of the page
+			cache in Linux kernel. \tristan{Writing is unclear, there should be a sentence starting with ``The second factor is ...''} In addition, the kernel controls I/O
+			activities based on the status of the memory including the
+			information about the cache size, the amount of available
+			memory and dirty data. As these are managed and changed with
+			the flushing and cache eviction mechanisms implemented in the
+			kernel, data flushing and cache eviction must be generalized \tristan{not sure what this sentence means, it should be rewritten}.
+			The simulator should also make system parameters such as
+			dirty{\_}ratio and dirty{\_}background{\_}ratio configurable
+			since they can vary in different systems \tristan{is it important to know here?}.
 
-			The idea of our simulator is to generalize the memory, the page cache, the LRU lists, storage device and mimic the cache eviction algorithm implemented in the kernel.
-As for storage devices, we can leverage the storage model with disk capacity and disk bandwidth implemented in SimGrid \cite{lebre2015}.
-Because memory capacity and bandwidth are fixed, they can be kept as constants, while the amount of cache is dynamically allocated based on the amount of available memory.
-The simulator also keeps track of the LRU lists as in the kernel since they are the key component in cache eviction algorithm.
-However, maintaining and handling pages, which are in packet level, are non-trivial when they require significant implementation overhead and hinder the performance of the simulator.
-To simplify cache eviction while preserving accuracy, in our approach, we consider a sequence of continuously accessed pages of a file as a single unit, a block of data.
-The information of a single page in a sequence can be used as the representative for other pages in a sequence since they have the same properties including file name, last access, the page is dirty or not.
-Hence, instead of dealing with individual pages, we consider a sequence of continuously accessed pages as a block of data.
-A data block keeps the information about file name, block size, last access and a dirty flag describing whether the data is dirty or not.
-Since a block represents a set of pages in real systems, it can be split if some data needs to be moved between the LRU lists or evicted from cache, but blocks can not be merged as we want to preserve this level of granularity to maintain the accuracy.
-Having these components, we can generalize I/O activities in our simulator.
-When the kernel reads and writes data, it simply reads/writes data from/to disk or cache with predefined disk bandwidth or memory bandwidth.
-In flushing and periodical flushing, the kernel traverses through the inactive and active lists to reduce the amount of dirty data, while cache eviction algorithm deletes blocks from the inactive list.
-The LRU lists are updated after every I/O action.
+			The idea of our simulator is to generalize \tristan{A simulator doesn't ``generalize'', it ``models'' or simply ``simulate''}
+			 the memory, the page
+			cache, the LRU lists, storage device and mimic \tristan{why a different verb for that?} the cache
+			eviction algorithm implemented in the kernel. As for storage
+			devices, we can leverage the storage model with disk capacity
+			and disk bandwidth implemented in SimGrid~\cite{lebre2015}.
+			Because memory capacity and bandwidth are fixed, they can be
+			kept as constants, while the amount of cache is dynamically
+			allocated based on the amount of available memory \tristan{what is the important message in this sentence?}. The
+			simulator also keeps track of the LRU lists as in the kernel
+			since they are the key component in cache eviction algorithm.
+			\tristan{This paragraph probably needs to be broken down in several ones, 
+			it seems to be mixing multiple ideas. Remember: 1 paragraph = 1 idea. The first sentence of the
+			paragraph should summarize this idea.}
+			However, maintaining and handling pages, which are in packet
+			level, are non-trivial when they require significant
+			implementation overhead and hinder the performance of the
+			simulator \tristan{how do you know? if it's just a guess, you should write it as such: ``and would likely hinder ...''}.
+			 To simplify cache eviction while preserving
+			accuracy, in our approach, we consider a sequence of
+			continuously accessed pages of a file as a single unit, a block
+			of data \tristan{In general it's clearer to explain first what you did, make sure that the reader understands, and 
+			contrast it next with what you didn't do and why}.
+			 The information of a single page in a sequence can be
+			used as the representative for other pages in a sequence since
+			they have the same properties including file name, last access,
+			the page is dirty or not. Hence, instead of dealing with
+			individual pages, we consider a sequence of continuously
+			accessed pages as a block of data. A data block keeps the
+			information about file name, block size, last access and a
+			dirty flag describing whether the data is dirty or not. Since a
+			block represents a set of pages in real systems, it can be
+			split if some data needs to be moved between the LRU lists or
+			evicted from cache, but blocks can not be merged as we want to
+			preserve this level of granularity to maintain the accuracy.
+			Having these components, we can generalize I/O activities in
+			our simulator. When the kernel reads and writes data, it simply
+			reads/writes data from/to disk or cache with predefined disk
+			bandwidth or memory bandwidth. In flushing and periodical
+			flushing, the kernel traverses through the inactive and active
+			lists to reduce the amount of dirty data, while cache eviction
+			algorithm deletes blocks from the inactive list. The LRU lists
+			are updated after every I/O action.
 
-			We create an extra layer with kernel-like functionalities between user application and SimGrid as is shown in Figure~\ref{fig:interaction}.
-Our new layer including two main additional classes, the \textbf{\textit{IOController}} class, which mimics the kernel I/O controlling logic, and the \textbf{\textit{MemoryManager}} class, which stores all necessary information about memory including capacity, the amount of cache, amount of memory used by applications, and manages LRU lists.
-Because every node has its own memory, storage devices and system configurations, there is one IOController object and one MemoryManager object created for each node.
-The IOController class keeps system parameters (dirty{\_}ratio, dirty{\_}background{\_}ratio), implements necessary I/O control functionalities such as read, write, flushing, and cache evicting.
-With these new components, instead of sending read/write requests directly to storage devices as in simulators with the current SimGrid version, user simulated applications send requests to an IOController of the node where requested data is stored.
-Based on the memory status in MemoryManager, the IOController orchestrates cache read/write, disk read/write, flushing and cache eviction to fulfill application requests.
-The time for each action is estimated and the total time of a request is returned to user applications.
+			We create an extra layer with kernel-like functionalities
+			between user application and SimGrid (Figure~\ref{fig:interaction}). \tristan{This is already implementation, as it is specific to Simgrid}
+			Our new layer includes two main
+			classes, the \textbf{\textit{IOController}},
+			which models the kernel I/O controlling logic, and the
+			\textbf{\textit{MemoryManager}}, which manages LRU lists \tristan{an adjective is missing to refer unambiguously to these lists: ``page cache LRU lists?'' (here and 
+			for other occurrences)} and stores all
+			necessary information about memory including capacity, the
+			amount of cache, amount of memory used by applications. Because every node has its own memory,
+			storage devices and system configurations, there is one
+			IOController object and one MemoryManager object created for
+			each node. The IOController class keeps system parameters
+			(dirty{\_}ratio, dirty{\_}background{\_}ratio), implements
+			necessary I/O control functionalities such as read, write,
+			flushing, and cache evicting. With these new components,
+			instead of sending read/write requests directly to storage
+			devices as in simulators with the current SimGrid version, user
+			simulated applications send requests to an IOController of the
+			node where requested data is stored. Based on the memory status
+			in MemoryManager, the IOController orchestrates cache
+			read/write, disk read/write, flushing and cache eviction to
+			fulfill application requests. The time for each action is
+			estimated and the total time of a request is returned to user
+			applications.
 
@@
 
-			\begin{algorithm}\caption{Read}\label{alg:read}
+			\begin{algorithm}\caption{Read \tristan{Avoid using math mode for variable names}}\label{alg:read}
 				\small
@@
 						\Desc{$run\_time$}{current simulated run time}
+					\tristan{Add an output}
    					\EndInput
-					\State $mem\_required \gets max(0, 2* $file size$-cached\_data -$ available memory)
+					\State $mem\_required \gets max(0, 2* $file size$-cached\_data -$ available memory) \tristan{you should explain where available memory and cached\_data are coming from}
 					\State $mem\_avail \gets max(0, 2 *$ file size$ - cached\_data -$ dirty data)
@@
 					\If {$free\_mem\_required > 0$}
-						\State evict from cache $(free\_mem\_required)$
+						\State evict from cache $(free\_mem\_required)$ \tristan{call a function from MemoryManager}
 					\EndIf
@@
 					\If {$cached\_data > 0$}
-    					\State update cached blocks in LRU lists with the last access
+    					\State update cached blocks in LRU lists with the last access \tristan{call a function from MemoryManager}
 						\If {all file data is cached}
-    						\State $cache\_read\_time \gets cached\_data /$ memory read bandwidth
-    						\State periodically flush $($flushing time $= cache\_read\_time)$
+    						\State $cache\_read\_time \gets cached\_data /$ memory read bandwidth \tristan{memory read bandwidth should be an input}
+    						\State periodically flush $($flushing time $= cache\_read\_time)$ \tristan{where does this function come from?}
 
@@
 
-			The algorithm to generalize file read is described in \textbf{\textit{Algorithm~\ref{alg:read}}}.
-In Linux systems, data can be read in two ways, read from cache with memory bandwidth and read from disk with disk bandwidth.
-Because total read time can be seen as a linear function of the read time in each of these two modes, we can generalize reading process as two separate and sequential phases, one is cache read, and the other is disk read.
-When a user application requests a file read, it calls the read method of the IOController of which a parameter is the file name.
-Then, IOController calculates how much memory is required to read the file based on the amount of available memory, file size, and the amount of file data cached.
-If there is not enough available memory, dirty data is flushed to disk to make more memory available by calling flush function with the amount of memory needed as the parameter.
-IOController then evicts old  blocks from the inactive list to accommodate the requested file.
-After that, IOController starts estimating file read time by checking if a some file data is cached.
-If there is cached data, the blocks containing this data are  updated in the LRU lists with the last access time.
-If all file data is cached, time to read from cache is added to total run time, and periodical flushing is triggered to flush dirty data during the cache read.
-If only a part of the file is in cache, disk read is required, this does not allow flushing and makes the cache read time neglectable.
-At the beginning of a disk read, dirty data is flushed \textcolor{red}{(Is this correct?)}, then data is read and added to the inactive list, disk read time is estimated with disk read bandwidth and added to total run time.
+			\tristan{In the algo, all variables and functions should be defined:
+			\begin{itemize}
+				\item As inputs or class variables
+				\item As class variables or functions of another class
+			\end{itemize}
+			Try to:
+			\begin{itemize}
+			\item reduce the number of lines, for instance you could always call evict\_from\_cache (please add \_s btw)
+			and avoid an if statement there.
+			\item streamline the logic: avoid a return statement in a second level of 
+			if statements.
+			\end{itemize}
+			Refer to the two classes when you define and call functions. 
+			}
+			The algorithm to generalize file read is described in
+			\textit{Algorithm~\ref{alg:read}} \tristan{(I removed the bold: avoid using bold in general.
+			 You can use $\backslash$emph if you want to put emphasis on 
+			something.)}. \tristan{refer to lines in Algo 1 throughout this paragraph, in brackets at the end of a sentence (line 2-4).} In Linux systems,
+			data can be read in two ways, read from cache with memory
+			bandwidth and read from disk with disk bandwidth. Because total
+			read time can be seen as a linear function of the read time in
+			each of these two modes, we can generalize reading process as
+			two separate and sequential phases, one is cache read, and the
+			other is disk read. When a user application requests a file
+			read, it calls the read method of the IOController of which a
+			parameter is the file name. Then, IOController calculates how
+			much memory is required to read the file based on the amount of
+			available memory, file size, and the amount of file data
+			cached. If there is not enough available memory, dirty data is
+			flushed to disk to make more memory available by calling flush
+			function with the amount of memory needed as the parameter.
+			IOController then evicts old  blocks from the inactive list to
+			accommodate the requested file. After that, IOController starts
+			estimating file read time by checking if a some file data is
+			cached. If there is cached data, the blocks containing this
+			data are  updated in the LRU lists with the last access time.
+			If all file data is cached, time to read from cache is added to
+			total run time, and periodical flushing is triggered to flush
+			dirty data during the cache read. If only a part of the file is
+			in cache, disk read is required, this does not allow flushing
+			and makes the cache read time neglectable. At the beginning of
+			a disk read, dirty data is flushed \textcolor{red}{(Is this
+			correct?)}, then data is read and added to the inactive list,
+			disk read time is estimated with disk read bandwidth and added
+			to total run time.
 
-			\textbf{\textit{Algorithm~\ref{alg:write}}} describes our generalization of file write.
-Similar to the file read process, a file can also be written with two different bandwidths, memory bandwidth and disk bandwidth, and file write can be generalized with two modes.
-File can be written with memory bandwidth before the page cache is saturated, or the amount of dirty data reaches dirty\_ratio.
-In our simulator, dirty data can be flushed concurrently during cache write, and the maximum amount of data can be written with memory bandwidth is calculated with the formula:
+			\textbf{\textit{Algorithm~\ref{alg:write}}} \tristan{fix consistently with Algo 1.} describes our
+			generalization of file write. Similar to the file read process,
+			a file can also be written with two different bandwidths,
+			memory bandwidth and disk bandwidth, and file write can be
+			generalized with two modes. File can be written with memory
+			bandwidth before the page cache is saturated, or the amount of
+			dirty data reaches dirty\_ratio. In our simulator, dirty data
+			can be flushed concurrently during cache write, and the maximum
+			amount of data can be written with memory bandwidth is
+			calculated with the formula:
 			\begin{equation}
@@
 				\item $M$ is the maximum amount of data can be written with memory write bandwidth
-				\item $A$ is the total available memory
+				\item $A$ is the total available memory \tristan{Should it be C?}
 				\item $d$ is the amount of dirty data
-				\item $mw$ is the memory write bandwidth
+				\item $mw$ is the memory write bandwidth \tristan{Make notations consistent in the algorithms}
 				\item $dw$ is the disk write bandwidth			
 			\end{itemize}
+			\tristan{M can be negative, add a max} \tristan{Add a sentence to explain this equation.	}
+			\tristan{In math mode, variable names should be 1 char max. Avoid $dirty\_ratio$ and even $mw$. Use $m_w$ instead.}
 
@@
 			Data flushing is another key function in our simulator.
-The input of the function is the amount of flushed data and the returned value is the simulated flushing time.
+The input of the function \tristan{which function?
+Be explicit} is the amount of flushed data and the returned value is the simulated flushing time.
 To flush dirty data, IOController simply traverses the sorted inactive and active list, reduces the dirty data amount by unmarking dirty blocks until the flushed amount reached the amount passed in or there is no dirty data left in cache.
@@
 
-			Last but not leats, we need to modified LRU lists after every I/O activity.
+			Last but not least, we need to modified LRU lists after every I/O activity.
 When IOController reads the data that is not cache, the data is read and added to the inactive list.
@@
 
-			Firstly, to validate our hypothesis, we create a simple simulator independent to simulation frameworks and libraries.
-This enables us to evaluate the accuracy and correctness of our model in a simple use case before integrating it in SimGrid.
-This simulator simulates a single thread pipeline running on a system with a single core CPU, one local storage device and all input and output files are stored locally.
-The results from the simulator is compared and validated with the results from a real pipeline running on a real system.
+			Firstly, to validate our hypothesis \tristan{hypothesis has a
+			strong meaning in science. ``simulation model'' would be
+			better}, we create a simple simulator independent to simulation
+			frameworks and libraries. This enables us to evaluate the
+			accuracy and correctness of our model in a simple use case
+			before integrating it in SimGrid. This simulator simulates a
+			single-threaded pipeline running on a system with a single core
+			CPU, one local storage device and all input and output files
+			are stored locally. The results from the simulator is compared
+			and validated with the results from a real pipeline running on
+			a real system.
 
-			Having our model validated, we create other different simulators of different use cases using the current version of SimGrid and the SimGrid version that is extended with our model.
-Then, we compare the results of the simulators with original SimGrid and extended SimGrid with the results of real pipelines on a real system.
-The results are compared in two different aspects, task completion time and I/O time, as well as the ability of the simulator in generalization of the page cache in terms of dirty data and cache used.
+			Having our model validated, we create other different
+			simulators of different use cases using the current version of
+			SimGrid and the SimGrid version that is extended with our
+			model. Then, we compare the results of the simulators with
+			original SimGrid and extended SimGrid with the results of real
+			pipelines on a real system. The results are compared in two
+			different aspects, task completion time and I/O time, as well
+			as the ability of the simulator in generalization of the page
+			cache in terms of dirty data and cache used.
 
-			In this work, we use Python 3.7 to implement the simple simulator, SimGrid 3.25 and Wrench 1.6 for SimGrid simulators.
+			In this work, we use Python 3.7 to implement the simple
+			simulator, SimGrid 3.25 and Wrench 1.6 for SimGrid simulators. \tristan{make sure to refer to the GitHub repo somewhere}
 
@@
 
-			When designing experiments, we consider a range of scenarios that have different impacts on the results.
-The experiments are a number of pipelines of sequential tasks.
-Each task reads an input file, performs some computation, generates an output file and writes it to disk.
-The output file of the previous task is the input file of the next task.
-Thus, a pipeline has only one actual input file.
-The size input and output files should vary have significant impacts on system I/O behaviors.
-The number of pipelines should also vary in different experiments since the shared bandwidth is affected by the number of concurrent I/O processes.
-Different storage types, which include local and shared storage, are also taken into account to evaluate the performance of our simulator on shared file systems.
-Finally, we adopt a real neuroimaging pipeline to evaluate the applicability of our simulator in simulating I/O time in real applications.
-Because our work focuses on I/O time, we assume that CPU time is correctly modeled and use the CPU time measured in the real pipelines to setup our simulated workflows.
+			When designing experiments, we consider a range of scenarios
+			that have different impacts on the results. The experiments are
+			a number of pipelines of sequential tasks. Each task reads an
+			input file, performs some computation, generates an output file
+			and writes it to disk. The output file of the previous task is
+			the input file of the next task. Thus, a pipeline has only one
+			actual input file. The size input and output files should vary
+			have significant impacts on system I/O behaviors. The number of
+			pipelines should also vary in different experiments since the
+			shared bandwidth is affected by the number of concurrent I/O
+			processes. Different storage types, which include local and
+			shared storage, are also taken into account to evaluate the
+			performance of our simulator on shared file systems. Finally,
+			we adopt a real neuroimaging pipeline to evaluate the
+			applicability of our simulator in simulating I/O time in real
+			applications. Because our work focuses on I/O time, we assume
+			that CPU time is correctly modeled and use the CPU time
+			measured in the real pipelines to setup our simulated
+			workflows. \tristan{This paragraph mixes too many things: you
+			should make one on pipelines, one on multi-threaded vs
+			single-treaded, one on data (file creation and size), a separate one on infrastructure (storage),
+			etc}
 
-			We use the cluster of \textit{Big Data Infrastructure for Neuroimaging Lab} at \textit{Concordia University} to conduct our experiments.
-The cluster consists of 1 login node which is the control node, 8 compute nodes and 4 storage nodes connected with 2 network switches.
-The login node has 128 GB of RAM, 1.8 TB of storage and 385 TB of Lustre shared file system.
-Each compute node has 6 SSDs with the size of 450GB each, 32 cores CPU and 256 GB of RAM, while each storage node has 12 HDDs and 2 SSDs.
-The cluster is run on CentOS 8.1 with the Slurm Workload Manager installed.
-We use \textbf{\textit{atop}} and \textbf{\textit{collectl}} as tools to monitor and collect data of memory, page cache status and disk throughput on the cluster.
+			We use the cluster of \textit{Big Data Infrastructure for
+			Neuroimaging Lab} at \textit{Concordia University} to conduct
+			our experiments. The cluster consists of 1 login node which is
+			the control node, 8 compute nodes and 4 storage nodes connected
+			with 2 network switches. The login node has 128 GB of RAM, 1.8
+			TB of storage and 385 TB of Lustre shared file system. Each
+			compute node has 6 SSDs with the size of 450GB each, 32 cores
+			CPU and 256 GB of RAM, while each storage node has 12 HDDs and
+			2 SSDs. The cluster is run on CentOS 8.1 with the Slurm
+			Workload Manager installed. We use \textbf{\textit{atop}} and
+			\textbf{\textit{collectl}} as tools to monitor and collect data
+			of memory, page cache status and disk throughput on the
+			cluster. \tristan{You should report file system types and versions, CPU and disk types, network type and bandwidth, Lustre configuration.}
 
-			\subsubsection{Expriment 1}
+			\subsubsection{Experiment 1 \tristan{Give experiments a more meaningful name than a number. For instance: ``Single-threaded evaluation''}}
 
-				In this experiment, our goal is to validate our model with a simple simulator implemented in Python.
-Thus, we only simulate one single thread pipeline of three tasks with the Python simulator, SimGrid simulator and run on our real system.
-All files are read and written locally.
-We use the input sizes of 20 GB, 50 GB, 75 GB and 100 GB to run the pipeline on the cluster, measure the CPU time of the pipeline with each input to simulate the pipeline with the simulators.
+				In this experiment, our goal is to validate our model with
+				a simple simulator implemented in Python \tristan{If that was your only goal you wouldn't need simgrid at all in this experiment}. Thus, we only
+				simulate one single thread pipeline of three tasks with the
+				Python simulator, SimGrid simulator and run on our real
+				system. All files are read and written locally. We use the
+				input sizes of 20 GB, 50 GB, 75 GB and 100 GB to run the
+				pipeline on the cluster, measure the CPU time of the
+				pipeline with each input to simulate the pipeline with the
+				simulators.
 

### 2020-06-04 paper_io_sim.tex
Method section revised

--- 2020-05-10 paper_io_sim.tex
@@
 \documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+\usepackage{graphicx}
+\usepackage{algorithm}
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
+\usepackage[dvipsnames]{xcolor}
+\usepackage{amsmath}
+\usepackage{hyperref}
+
+\algblock{Input}{EndInput}
+\algnotext{EndInput}
+\algblock{Output}{EndOutput}
+\algnotext{EndOutput}
+\newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
+
+\newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:}#1\color{black}}
 
 \begin{document}
-\title{I/O simulation extension for SimGrid framework}
-\author{Hoang-Dung Do}
+\title{I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework}
+
+\author{\IEEEauthorblockN{Hoang-Dung Do, Tristan Glatard and Val\'erie Hayot-Sasson
+  }\\
+  \IEEEauthorblockA{
+    Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada
+  }
+}
+
 \maketitle
@@
 	\section{Method}
-		In this section, we discuss our approach to model page cache and the cache eviction strategy implemented in Linux.
-We also detail the design the simulators in order to generalize page cache as well as the implementation in Python and SimGrid framework.
-Finally, we describe some experimental scenarios to compare the results of Python simulator, baseline SimGrid simulator and the results from the real pipelines.
+		In this section, we discuss our approach to model page cache and
+		the cache eviction strategy implemented in Linux. We also detail
+		the design the simulators in order to generalize page cache as well
+		as the implementation in Python and SimGrid framework. Finally, we
+		describe some experimental scenarios to compare the results of
+		Python simulator, baseline SimGrid simulator and the results from
+		the real pipelines.
+
 		\subsection{Principle of the simulator}
 
-			In order to make the simulator accurate, we try to simulate simulate page cache and mimic the 2-LRU lists cache eviction algorithm implemented in Linux.
-Our design also keeps track two LRU lists ans specific system configurations as in Linux kernel.
-However, maintaining and handling pages, which is in packet level, requires significant implementation overhead and hinders the performance of the simulator.
-To simplify the model whilst preserving accuracy, in our approach, we consider a sequence of continuously accessed pages of the same file from the same device as a single unit, a \textit{block} of data.
-The information of a single page in a sequence is representative for other pages the a sequence since they have the same properties including file name, last accessed, dirty or not.
-Thus, instead of dealing with individual pages, we view a sequence of continuously accessed pages as a block of data.
-A data block has the information about file name, size, last access and a dirty flag.
-As a block represents a set of pages in real systems, it can be split if some data needs to be moved between the LRU lists or evicted from cache, but blocks can not be merged as we want to preserve the access times of blocks to maintain the accuracy.
+			When modeling the I/O time, we consider the key factors of the
+			I/O mechanism having the biggest impacts on data read/write.
+			The first factor taken into account is the use of the page
+			cache in Linux kernel. The second factor  is the cache eviction, 
+			data flushing and cache eviction mechanisms implemented in the 
+			Linux kernel, as they manage and change the status of the memory 
+			including cache size, amount of available memory and 
+			dirty data, based on which the kernel controls I/O activities. Therefore, 
+			the idea of our simulator is to simulate the memory, the Linux 
+			page cache with its LRU lists, storage device and imitate the data flushing 
+			and cache eviction algorithm implemented in the kernel.
 
-			With this data block concept, when a file is accessed by an I/O operation, blocks are added to the top of inactive list if they are not existing in cache, or they are moved to the top of the list that they are kept in with last accesses updated.
-The lists are balanced after every time the simulator finishes reading or writing a portion of data from/to cache or disk.
+			At the level of detail in our simulator, the simulation models of memory 
+			and storage devices are straightforward when they can be modeled with 
+			predefined capacity and bandwidth. To simulate the cache eviction mechanism 
+			implemented in Linux, our simulator also keeps tracks of page cache 
+			LRU lists as in the kernel. However, instead of dealing with lists pages, we 
+			consider a sequence of continuously accessed pages of a file as a single 
+			unit, a block of data. A data block keeps the information about file name, 
+			block size, last access and a dirty flag describing whether the data is 
+			dirty or not.
 
-			To simulate flushing and periodical flushing mechanisms, in which dirty data in cache is written to disk, we simply calculate the amount of flushed dirty data.
-Data flushing is called when the memory required by the application is more than available memory, while periodical flushing is called during CPU time.
-In flushing, the amount of dirty data to be flushed is calculated based on the amount of required memory, while the flushed amount in periodical flushing is estimated with CPU time.
-Initially, the blocks in inactive list are flushed first, and the next are dirty blocks in the active list, until the total flushed amount is reached.
-The flushed blocks are split if needed and become not dirty.
-The blocks in the inactive list which are not dirty are available for cache eviction.
+			The idea of data block offers our simulator some obvious advantages. 
+			In cache eviction algorithm, which is one of the key components 
+			in our model, maintaining and handling pages in packet level are 
+			non-trivial when they require significant implementation overhead and 
+			would likely to hinder the performance of the simulator. The concept of data 
+			block not only simplifies our cache eviction simulation but also allows to 
+			preserve the accuracy of the simulator by using a single page as the 
+			representative for a sequence of pages with similar properties including 
+			file name, last access, dirty or not. Because a block represents a set of 
+			pages in real systems, it can be split if some data needs to be moved 
+			between the page cache LRU lists or evicted from cache. However, blocks can 
+			not be merged as we want to preserve this level of granularity to maintain 
+			high accuracy.
 
-			Similar to flushing and periodical flushing, the amount of evicted data is calculated based on the the amount of memory required by application and the amount of free memory.
-Then, evicted blocks are deleted from inactive list, and the lists are balanced.
+			Having these simulation components, we can generalize I/O activities in
+			our simulator. When the kernel reads and writes data, it simply 
+			reads/writes data from/to disk or cache with predefined disk bandwidth 
+			or memory bandwidth. The data blocks are added or updated in page cache 
+			LRU lists accordingly. In flushing and periodical flushing, the kernel 
+			traverses through the inactive and active lists to reduce the amount of 
+			dirty data, while cache eviction algorithm deletes blocks from the 
+			inactive list. The page cache LRU lists are updated after every I/O action.
+
+			\begin{figure}
+   				\centering
+   				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
+   				\caption{The additional layer between user application layer and 
+   				original SimGrid API}\label{fig:interaction}
+			\end{figure}
+
+			We encapsulate our simulation components with kernel-like functionalities 
+			into a layer between user simulated application and storage simulation 
+			component (Figure~\ref{fig:interaction}). Our new layer includes two main
+			classes, the \textbf{\textit{IOController}}, which models the kernel 
+			I/O controlling logic, and the \textbf{\textit{MemoryManager}}, which 
+			manages page cache LRU lists and stores all necessary information about 
+			memory including capacity, amount of cache, amount of memory used 
+			by applications. Because every node has its own memory, storage devices 
+			and system configurations, there is one IOController object and one 
+			MemoryManager object created for each node. The IOController class 
+			keeps system parameters (dirty{\_}ratio, dirty{\_}background{\_}ratio), 
+			implements necessary I/O control functionalities such as read, write,
+			flushing, and cache evicting.
+
+			With this layer, user simulated applications send requests to an 
+			IOController of the node where requested data is stored, instead of 
+			sending read/write requests directly to storage model as in simulators 
+			with the current SimGrid version. Based on the memory status in 
+			MemoryManager, the IOController orchestrates cache read/write, 
+			disk read/write, flushing and cache eviction to fulfill application 
+			requests. The time for each action is estimated and the total time of 
+			a request is returned to user applications.
+
+			\begin{algorithm}\caption{Read}\label{alg:read}
+				\small
+				\begin{algorithmic}[1]
+					\Input
+        				\Desc{fn}{file name}
+        				\Desc{fs}{file size}
+						\Desc{rt}{current simulated run time}
+						\Desc{mm}{MemoryManager object}
+						\Desc{$r_m$}{memory read bandwidth}
+						\Desc{$r_d$}{disk read bandwidth}
+   					\EndInput
+   					\Output
+						\Desc{rt}{simulated run time after the file is read}
+   					\EndOutput
+					\State mem\_required = max(0, 2*fs - mm.cached\_data(fn) - 
+					mm.available\_memory()) 
+					\State mem\_avail = max(0, 2 * fs - mm.cached\_data(fn) - 
+					mm.dirty\_data())
+					\State rt = rt +  mm.flush(mem\_required - 
+					mem\_avail)	
+					\State free\_mem\_required = max(0, 2 * fs - mm.cached\_data(fn) - 
+					mm.free\_memory())
+					\State mm.evict(free\_mem\_required) 
+					\If {mm.cached\_data(fn) $>$ 0}
+    					\State mm.cache\_read(fn) 
+						\If {mm.cached\_data(fn) == fs}
+    						\State time = mm.cache\_data(fn) / $r_m$
+    						\State IOController.periodic\_flush(time) 
+							\State rt = rt + time
+						\EndIf
+					\EndIf
+					\If {mm.cached\_data(fn) $<$ fs}
+						\State rt = rt + IOController.flush(mm.dirty\_data())
+						\State mm.disk\_read(fs - mm.cache\_data(fn), fn)
+    					\State rt = rt + (fs - mm.cache\_data(fn)) / $r_d$
+					\EndIf
+
+					\Return rt
+
+				\end{algorithmic}
+			\end{algorithm}
+
+			The algorithm to generalize file read is described in 
+			\textit{Algorithm~\ref{alg:read}}. In Linux systems, data can be read in 
+			two ways, read from cache with memory bandwidth and read from disk 
+			with disk bandwidth. Because total read time can be seen as a 
+			linear function of the read time in each of these two modes, 
+			we can generalize reading process as two separate and sequential phases, 
+			one is cache read, and the other is disk read. When a user application 
+			requests a file read, it calls the read method of the IOController. Then, 
+			IOController calculates how much memory is required to read the file 
+			based on the amount of available memory, file size, and the amount of 
+			file data cached (line 10). If there is not enough available memory, 
+			dirty data is flushed to disk to make more memory available by calling flush
+			function with the amount of memory needed as the parameter (line 11-12).
+			IOController then evicts old blocks from the inactive list to
+			accommodate the requested file and flushing time is simulated (line 13-14). 
+			After that, IOController starts estimating file read time by first checking 
+			if a some file data is cached (line 15). If there is cached data, 
+			it is read from cache and the blocks containing this data are updated in 
+			the page cache LRU lists with the last access time (line 16). If all file data 
+			is cached, time to read from cache is added to total run time, and 
+			periodical flushing is triggered to flush dirty data during the cache read 
+			(line 18-19). If only a part of the file is in cache, disk read is required, 
+			this does not allow flushing and makes the cache read time neglectable. 
+			At the beginning of a disk read, dirty data is flushed, then data is read 
+			and added to the inactive list, disk read time is estimated with disk read 
+			bandwidth and added to total run time (line 23-27).
+
+			\textit{Algorithm~\ref{alg:write}} describes our
+			generalization of file write. Similar to the file read process,
+			a file can also be written with two different bandwidths,
+			memory bandwidth and disk bandwidth, and thus file write can be
+			generalized with two modes. File can be written with memory
+			bandwidth before the page cache is saturated, or the amount of
+			dirty data reaches dirty\_ratio. Firstly, if dirty\_ratio is not reached 
+			the amount of data can be written with memory bandwidth is
+			calculated with the formula:
+			\begin{equation}
+				M = max(0, \frac{(d*C - D)*w_m}{w_m - w_d})\label{equa:freeamt}
+			\end{equation}
+
+			where:
+			\begin{itemize}
+				\item $M$ is the maximum amount of data that can be written with memory 
+				write bandwidth
+				\item $d$ is dirty ratio
+				\item $C$ is the total available memory
+				\item $D$ is the amount of dirty data
+				\item $w_m$ is the memory write bandwidth 
+				\item $w_d$ is the disk write bandwidth			
+			\end{itemize}
+
+			By this formular, we assume that data is written with with memory 
+			bandwidth until all file data is written or dirty\_ratio is reached, 
+			the kernel concurrently writes data to cache and flushes dirty data 
+			to disk. If file size is smaller than or equal to $M$, the whole file is 
+			written to cache as dirty data with memory bandwidth (line 12). Otherwise, 
+			the remaining file data is written with disk bandwidth. Before simulating 
+			cache write, IOController evicts the page cache to accommodate new data 
+			(line 13). Then, data is written to cache as dirty data with memory 
+			bandwidth write time is added to simulated time (line 14-16). Because 
+			cache write and data flushing are concurrent, the actual amount of written 
+			dirty data reduces by the amount of flushed data (line 15). In the 
+			second phase, in which data is written with disk bandwidth, our simulator 
+			can generalize not only the case that data is written directly to disk, 
+			but also the case that applications have to wait for dirty data 
+			being flushed. Before writing, if the amount of free memory is not enough 
+			for the remaining file data, evictable memory is evicted (line 20-23). 
+			Here, the kernel does not need to guarantee that the amount of free 
+			memory after eviction can accommodate the remaining data. In the case 
+			free memory is less than remaining data, we the kernel can flush, 
+			evict and write new data concurrently. Finally, IOController simulates 
+			the time to write the remaining amount data, adds written data to 
+			page cache LRU lists, and the amount of dirty data remains the same 
+			(line 24). The total simulated time of the second phases is added and 
+			returned to user applications.
+
+			\begin{algorithm}\caption{Write}\label{alg:write}
+				\small
+				\begin{algorithmic}[1]
+					\Input
+        				\Desc{fn}{file name}
+        				\Desc{fs}{file size}
+						\Desc{rt}{current simulated run time}
+						\Desc{mm}{MemoryManager object}
+						\Desc{$w_m$}{memory write bandwidth}
+						\Desc{$w_d$}{disk write bandwidth}
+   					\EndInput
+   					\Output
+						\Desc{rt}{simulated run time after the file is read}
+   					\EndOutput
+					\State dirty\_left = $d$ * mm.available\_memory() - mm.dirty\_data()
+					\If {dirty\_left $>$ 0}
+   						\State free\_amt = min(fs, $M$) \Comment{$M$ in Equation
+   						~\ref{equa:freeamt}}
+    					\State IOController.evict(free\_amt - mm.free\_memory())
+    					\State free\_time = free\_amt / $w_m$
+    					\State mm.write(fn, free\_amt, dirty = free\_amt - free\_time * $w_d$)
+   						\State rt = rt + free\_time
+    				\EndIf
+
+					\If {free\_amt $\geq$ fs}
+						\Return rt
+					\EndIf
+
+					\State throttled\_amt = fs - free\_amt
+					\State cache\_required = max(0, throttled\_amt  - mm.free\_memory())
+					\State evicted\_amt = min(cache\_required, mm.evictable\_memory())
+					\State mm.evict(evicted\_amt)
+					\State mm.write(fn, cached = min(mm.free\_memory(), 
+					throttled\_amt), dirty = 0)
+					\State rt = rt + throttled\_amt /$w_d$
+
+					\Return rt
+
+				\end{algorithmic}
+			\end{algorithm}
+
+			Data flushing is another key function in our simulator. It takes 
+			the input as the amount of flushed data and the returned value is the 
+			simulated flushing time. To flush dirty data, IOController simply 
+			traverses the sorted inactive and active list, reduces the dirty data 
+			amount by unmarking dirty blocks until the flushed amount reached 
+			the amount passed in or there is no dirty data left in cache. 
+			Similarly, periodical flushing is the function that simulates data flushing 
+			during CPU time and cache read. However, the flushed data amount is 
+			calculated with the input which is flushing time. In both flushing 
+			functions, the last block being flushed is split if its data is not 
+			entirely flushed.
+
+			Equally important, cache eviction function is called from both 
+			read and write to free up the page cache when free memory 
+			is not enough. The amount of evicted data is passed in as in flushing, 
+			however, this function does not add up the simulated time as 
+			cache eviction time is neglectable in real systems . When eviction 
+			is called, IOController deletes the last accessed blocks from the 
+			inactive list until the evicted amount reaches the amount passed in or 
+			there is no evictable memory left. If the block is not entirely deleted, 
+			the block is split.
+
+			Last but not least, we need to modified page cache LRU lists after 
+			every I/O activity. When IOController reads the data that is not cache, 
+			the data is read and added to the inactive list. If the data is already 
+			in cache, the cached blocks are moved to the active list with the 
+			last access time updated. Finally, MemoryManager makes the page cache 
+			LRU lists balanced by moving blocks from the longer list to the shorter 
+			and sorting the lists.
 
 		\subsection{Implementation}
-			\begin{itemize}
-				\item Which features of memory are implemented.
-				\item Level of granularity, how features are implemented.
-				\item Specific implementation details in python and SimGrid.
-			\end{itemize}
+
+			Firstly, to validate our simulation model, we create a simple 
+			simulator independent to simulation frameworks and libraries. 
+			This enables us to evaluate the accuracy and correctness of our 
+			model in a simple use case before integrating it in SimGrid. 
+			This simulator simulates a single-threaded pipeline running on a system 
+			with a single core CPU, one local storage device and all input and 
+			output files are stored locally. The results from the simulator is compared 
+			and validated with the results from a real pipeline running on
+			a real system.
+
+			Having our model validated, we create other different
+			simulators of different use cases using the current version of
+			SimGrid and the SimGrid version that is extended with our
+			model. Then, we compare the results of the simulators with
+			original SimGrid and extended SimGrid with the results of real
+			pipelines on a real system. The results are compared in two
+			different aspects, task completion time and I/O time, as well
+			as the ability of the simulator in generalization of the page
+			cache in terms of dirty data and cache used.
+
+			In this work, we use Python 3.7 to implement the simple
+			simulator, SimGrid 3.25 and Wrench 1.6 for SimGrid simulators. 
+			SimGrid source code is available at 
+			\url{https://framagit.org/simgrid/simgrid}, and Wrench is provided on
+			github at \url{https://github.com/wrench-project/wrench}.
 
 		\subsection{Experiments}
-			For each experiment: 
-			\begin{itemize}
-				\item Describe data, workflow, number of tasks, task details, environment of each experiment.
-				\item How we compare the results from real pipeline, baseline SimGrid  and SimGrid with our extension.
-			\end{itemize}
 
-			\subsubsection{Expriment 1}
-				A single pipeline running one node.
-			\subsubsection{Expriment 2}
-				Multiple pipelines running in parallel on multiple nodes.
-			\subsubsection{Expriment 3}
-				Same as Experiment 2 but nodes write to a shared file system.
-			\subsubsection{Expriment 4}
+			When designing experiments, we consider a range of scenarios
+			that have different impacts on the results. The experiments are
+			a number of pipelines of sequential tasks. Each task reads an
+			input file, performs some computation, generates an output file
+			and writes it to disk. The output file of the previous task is
+			the input file of the next task. Thus, each pipeline has only one
+			actual input file.
+
+			Initially, we want to have a detailed view of the effects of data size 
+			on I/O behaviors and validate our simulator in modeling and 
+			capturing those impacts. Thus, there should be only one pipeline and 
+			the size of input and output files should vary in a range that 
+			the change in file size can cause significant changes on I/O behaviors.
+
+			Furthermore, the number of pipelines should also vary. Having our 
+			simulator validated on a single threaded experiment, the next goal 
+		    is to evaluate our simulation model with a multi-threaded experiment, 
+		    with a varying number of pipelines run on a multi-core host and write 
+		    to the same local storage device. As a result, the shared bandwidth 
+		    the pipelines is different with different numbers of concurrent processes. 
+		    By doing so, we can evaluate the performance of our model on the simulation 
+		    of multi-threaded, shared bandwidth applications.
+
+		    Moreover, different storage types, which include local and shared 
+		    storage, should also be taken into account when they are commonly used 
+		    in computer systems. The scenario of the next experiment is the same as 
+		    the multi-threaded experiment, but should replace local storage device 
+		    with a shared file systems, so that we can not only access our 
+		    simulator on I/O with shared file system, but also compare the 
+		    performance of the simulator on different storage types.
+
+		    Finally, we adopt a real neuroimaging pipeline to evaluate the 
+		    applicability of our simulator in simulating I/O time in real
+			applications. We simulate the pipeline with a simulated workflow running 
+			on a simulated plaform and compare the simulation results with the real 
+			results. Because our work focuses on I/O time, we assume
+			that CPU time is correctly modeled and use the CPU time measured 
+			in the real pipelines to setup our simulated workflows.
+
+			We use the cluster of \textit{Big Data Infrastructure for Neuroimaging Lab} 
+			at \textit{Concordia University} to conduct our experiments. The cluster 
+			consists of one login node which is the control node, 8 compute nodes 
+			and 4 storage nodes connected with two network switches. The login node 
+			has one Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 128GiB of RAM, 1.8TB 
+			of storage of XFS file system, 13GB of tmpfs file system and 385 TB of 
+			Lustre shared file system. Each compute node has two 16 cores Intel(R) 
+			Xeon(R) Gold 6130 CPU @ 2.10GHz, 256 GB of RAM and 6 SSDs, 450GB each 
+			with XFS file system, 378GB of tmpfs and 126GB of devtmpfs file system.
+			Lustre file system is configured with one metadata target with 854GB 
+			of storage, 44 object storage targets with 8.7TB of storage each. 
+			The cluster is run on CentOS 8.1 with the Slurm Workload Manager installed. We use 
+			\textbf{\textit{atop}} and \textbf{\textit{collectl}} as tools to monitor 
+			and collect data of memory, page cache status and disk throughput on 
+			the cluster. The cluster nodes are connected with ... 
+			\textcolor{red}{[network description here]}
+
+			\subsubsection{Single-threaded evaluation}
+
+				In this experiment, our goal is to validate our model with a simple 
+				simulator implemented in Python and compare the simulation precision 
+				of our simulation model and the existing model in SimGrid. 
+				Thus, we simulate one single thread pipeline of three tasks with the
+				Python simulator, SimGrid simulator and run on our real system. 
+				All files are read and written locally. We use the input sizes of 
+				20 GB, 50 GB, 75 GB and 100 GB to run the pipeline on the cluster, 
+				measure the CPU time of the pipeline with each input to simulate 
+				the pipeline with the simulators.
+
+			\subsubsection{Multi-threaded evaluation}
+
+				In the second experiment, we evaluate the simulator in concurrent I/O 
+				simulation with a multi-threaded application. As we run the pipelines on 
+				only one compute node of the cluster, there are upto 32 cores per node 
+				and all input and output files are read and written on the same disk,  
+				we create 32 input files with the size of 3GB each and vary the number of 
+				concurrent pipelines from 1 to 32. Because our goal in this experiment is 
+				to evaluate concurrent I/O in our simulation model, and multi-threaded 
+				simulator implementation of the Python is expensive, we only compare the 
+				performance of simulators with original SimGrid, SimGrid integrated with 
+				our model and and real pipelines. The results are compared in total 
+				makespan of the pipelines, cumulative read time and cumulative write time.
+
+			\subsubsection{Evaluation on storage types}
+
+			\subsubsection{Simulation of a real application}
 				A real pipeline (for example a pipeline with nighres)

### 2020-06-04 paper_io_sim.tex
Minor fix

@@
    					\EndOutput
-					\State mem\_required = max(0, 2*fs - mm.cached\_data(fn) - 
+					\State mem\_required = max(0, 2 * fs - mm.cached\_data(fn) - 
 					mm.available\_memory()) 
@@
 			traverses the sorted inactive and active list, reduces the dirty data 
-			amount by unmarking dirty blocks until the flushed amount reached 
+			amount by turning off dirty flags until the flushed amount reached 
 			the amount passed in or there is no dirty data left in cache. 

### 2020-06-08 paper_io_sim.tex
Comments up to Algo 1

@@
 	\section{Method}
-		In this section, we discuss our approach to model page cache and
+		In this section, we present our approach to model page cache and
 		the cache eviction strategy implemented in Linux. We also detail
@@
 			page cache with its LRU lists, storage device and imitate the data flushing 
-			and cache eviction algorithm implemented in the kernel.
+			and cache eviction algorithm implemented in the kernel. \tristan{This paragraph is convoluted, it needs to be rewritten with a straightforward logic.}
 
@@
 			and storage devices are straightforward when they can be modeled with 
-			predefined capacity and bandwidth. To simulate the cache eviction mechanism 
-			implemented in Linux, our simulator also keeps tracks of page cache 
-			LRU lists as in the kernel. However, instead of dealing with lists pages, we 
+			predefined capacity and bandwidth \tristan{Do you mean ``We
+			model memory as a storage device, with a capacity and a
+			bandwidth''? If that's the case, write this sentence like this,
+			it's more direct and less ambiguous. You should also explain
+			how you calculate read/write time from the bandwidth. And does
+			your model distinguish read and write bandwidths? I think the
+			memory model is one paragraph, not just one sentence in a
+			paragraph that then talks about other things}. To simulate the
+			cache eviction mechanism implemented in Linux, our simulator
+			also keeps tracks of page cache LRU lists as in the kernel.
+			However, instead of dealing with lists pages \tristan{as done by the kernel?}, we 
 			consider a sequence of continuously accessed pages of a file as a single 
-			unit, a block of data. A data block keeps the information about file name, 
+			unit, a block of data \tristan{so is a data block a set of pages? The definition of a data block is unclear.}. A data block keeps the information about file name, 
 			block size, last access and a dirty flag describing whether the data is 
@@
 			our simulator. When the kernel reads and writes data, it simply 
-			reads/writes data from/to disk or cache with predefined disk bandwidth 
+			reads/writes data \tristan{do you mean data blocks?} from/to disk or cache with predefined disk bandwidth 
 			or memory bandwidth. The data blocks are added or updated in page cache 
@@
 			We encapsulate our simulation components with kernel-like functionalities 
-			into a layer between user simulated application and storage simulation 
-			component (Figure~\ref{fig:interaction}). Our new layer includes two main
-			classes, the \textbf{\textit{IOController}}, which models the kernel 
-			I/O controlling logic, and the \textbf{\textit{MemoryManager}}, which 
+			into a layer between the simulated application and the simulated storage (Figure~\ref{fig:interaction}). Our new layer includes two main
+			classes, the \emph{IOController}, which models the kernel 
+			I/O controlling logic, and the \emph{MemoryManager}, which 
 			manages page cache LRU lists and stores all necessary information about 
-			memory including capacity, amount of cache, amount of memory used 
-			by applications. Because every node has its own memory, storage devices 
-			and system configurations, there is one IOController object and one 
-			MemoryManager object created for each node. The IOController class 
-			keeps system parameters (dirty{\_}ratio, dirty{\_}background{\_}ratio), 
+			memory including total capacity, current cache size, and current amount of memory used 
+			by applications \tristan{is it exhaustive?}. Because every node has its own memory, storage devices 
+			and system configurations, one IOController object and one 
+			MemoryManager object are created for each node. The IOController class 
+			stores system parameters (dirty{\_}ratio, dirty{\_}background{\_}ratio), 
 			implements necessary I/O control functionalities such as read, write,
-			flushing, and cache evicting.
+			flushing, and cache eviction.
 
-			With this layer, user simulated applications send requests to an 
-			IOController of the node where requested data is stored, instead of 
-			sending read/write requests directly to storage model as in simulators 
-			with the current SimGrid version. Based on the memory status in 
+			Simulated applications send requests to an IOController of the
+			node where the requested data is stored, instead of sending
+			read/write requests directly to storage model as in the current
+			SimGrid version. Based on the memory status in 
 			MemoryManager, the IOController orchestrates cache read/write, 
@@
 			requests. The time for each action is estimated and the total time of 
-			a request is returned to user applications.
+			a request is returned to application.
 
-			\begin{algorithm}\caption{Read}\label{alg:read}
+			\begin{algorithm}\caption{Read \tristan{Find a more descriptive title. Lines 10, 11 and 13 should fit on a single line. Suggestion: replace cached\_data with cached, available\_memory with avail.
+				Add comments in the code. Line 23: should it be in an else clause? } }\label{alg:read}
 				\small
@@
 
-					\Return rt
+					\State \Return rt
 
@@
 
-			The algorithm to generalize file read is described in 
-			\textit{Algorithm~\ref{alg:read}}. In Linux systems, data can be read in 
+			The algorithm to generalize \tristan{``generalize'' isn't a great word here, I'd simply say ``simulate''} file reads is described in 
+			Algorithm~\ref{alg:read}. In Linux systems, data can be read in 
 			two ways, read from cache with memory bandwidth and read from disk 
@@
 			function with the amount of memory needed as the parameter (line 11-12).
-			IOController then evicts old blocks from the inactive list to
+			IOController then evicts old blocks \tristan{I don't see blocks being handled in this algo. It adds to the confusion related to the 
+			lack of a proper definition of blocks.} from the inactive list to
 			accommodate the requested file and flushing time is simulated (line 13-14). 
@@
 
-			\textit{Algorithm~\ref{alg:write}} describes our
+			Algorithm~\ref{alg:write} describes our
 			generalization of file write. Similar to the file read process,

### 2020-06-14 paper_io_sim.tex

--- 2020-06-04 paper_io_sim.tex
@@
 	\section{Method}
-		In this section, we discuss our approach to model page cache and
-		the cache eviction strategy implemented in Linux. We also detail
-		the design the simulators in order to generalize page cache as well
-		as the implementation in Python and SimGrid framework. Finally, we
-		describe some experimental scenarios to compare the results of
-		Python simulator, baseline SimGrid simulator and the results from
-		the real pipelines.
+		In this section, we present our approach to model file read/write, 
+		I/O in memory, data flushing and cache eviction mechanisms 
+		implemented in Linux. 
+		We also detail the design the simulators in order to simulate I/O activities 
+		and the implementation in Python and SimGrid framework. 
+		Finally, we describe some experimental scenarios to evaluate our model 
+		with a Python simulator, a baseline SimGrid simulator, a simulator 
+		with SimGrid integrated with our model, and the results from real pipelines.
 
@@
 
-			When modeling the I/O time, we consider the key factors of the
-			I/O mechanism having the biggest impacts on data read/write.
-			The first factor taken into account is the use of the page
-			cache in Linux kernel. The second factor  is the cache eviction, 
-			data flushing and cache eviction mechanisms implemented in the 
-			Linux kernel, as they manage and change the status of the memory 
-			including cache size, amount of available memory and 
-			dirty data, based on which the kernel controls I/O activities. Therefore, 
-			the idea of our simulator is to simulate the memory, the Linux 
-			page cache with its LRU lists, storage device and imitate the data flushing 
-			and cache eviction algorithm implemented in the kernel.
+			In our I/O simulation model, we try to simulate the write back page cache 
+			mechanisms including memory I/O, cache eviction and data flushing as we 
+			believe that they are the most important factors that affect the I/O time.
 
-			At the level of detail in our simulator, the simulation models of memory 
-			and storage devices are straightforward when they can be modeled with 
-			predefined capacity and bandwidth. To simulate the cache eviction mechanism 
-			implemented in Linux, our simulator also keeps tracks of page cache 
-			LRU lists as in the kernel. However, instead of dealing with lists pages, we 
-			consider a sequence of continuously accessed pages of a file as a single 
-			unit, a block of data. A data block keeps the information about file name, 
-			block size, last access and a dirty flag describing whether the data is 
-			dirty or not.
+			Memory is modeled with capacity, read bandwidth and write 
+			bandwidth. When data is read from or written to memory, the actual 
+			read/write time is calculated with the amount of data and memory 
+			read/write bandwidth:
 
-			The idea of data block offers our simulator some obvious advantages. 
-			In cache eviction algorithm, which is one of the key components 
-			in our model, maintaining and handling pages in packet level are 
-			non-trivial when they require significant implementation overhead and 
-			would likely to hinder the performance of the simulator. The concept of data 
-			block not only simplifies our cache eviction simulation but also allows to 
-			preserve the accuracy of the simulator by using a single page as the 
-			representative for a sequence of pages with similar properties including 
-			file name, last access, dirty or not. Because a block represents a set of 
-			pages in real systems, it can be split if some data needs to be moved 
-			between the page cache LRU lists or evicted from cache. However, blocks can 
-			not be merged as we want to preserve this level of granularity to maintain 
-			high accuracy.
+			\begin{align*}
+				& t_{mr} = D / r_m \\ 
+				& t_{mw} = D / w_m\
+			\end{align*}
 
-			Having these simulation components, we can generalize I/O activities in
-			our simulator. When the kernel reads and writes data, it simply 
-			reads/writes data from/to disk or cache with predefined disk bandwidth 
-			or memory bandwidth. The data blocks are added or updated in page cache 
-			LRU lists accordingly. In flushing and periodical flushing, the kernel 
-			traverses through the inactive and active lists to reduce the amount of 
-			dirty data, while cache eviction algorithm deletes blocks from the 
-			inactive list. The page cache LRU lists are updated after every I/O action.
+			where:
+			\begin{itemize}
+				\item $t_{mr}$ is time to read data from memory
+				\item $t_{mw}$ is time to write data to memory
+				\item $D$ is the amount of data to read/write
+				\item $r_m$ is memory read bandwidth
+				\item $w_m$ is memory write bandwidth
+			\end{itemize}
+
+			We introduce the concept of data block as a unit to represent cached data. 
+			A data block is a subset of pages of a file in page cache that is accessed 
+			at a specific time. Therefore, a block has the information about file name, 
+			block size, last access and a dirty flag with values of 0 and 1, 
+			0 represents block of lean data, while 1 indicates that the data block 
+			is dirty. 
+			Cached data of a file can be represented by multiple blocks. 
+			Because a data block represents a subset of cached pages, 
+			it can be split into smaller blocks, and multiple blocks can also 
+			be merged into a larger one.
+
+			In Linux kernel, the page cache LRU lists are lists of pages of 
+			files in cache. In our simulator, instead of using lists of pages, 
+			we simulate these lists with two lists of data blocks, an active list and 
+			an inactive list, ordered by block last access as is shown in 
+			Figure~\ref{fig:lrulist}.
+			As in the kernel, in our simulator, we keep the size of the active list not 
+			larger than twice the size of the inactive list by moving least recently 
+			used blocks from the active list to the inactive list.
+
+			\begin{figure}
+   				\centering
+   				\includegraphics[width=\columnwidth]{figures/lru_lists.pdf}
+   				\caption{Simulation of page cache LRU lists with data blocks}\label{fig:lrulist}
+			\end{figure}
+
+			Modeling page cache LRU lists as lists of blocks offers our simulator 
+			some obvious advantages. 
+			Maintaining and handling lists of pages in the simulator, which is in 
+			packet level, are non-trivial when they require significant implementation 
+			overhead and would likely to hinder the performance of the simulator. 
+			Therefore, using page cache lists of data blocks can simplify the 
+			simulation of cache eviction and data flushing while preserving
+			the accuracy in i/o time simulation.
+
+			With the simulation model of page cache LRU lists of blocks, now we can 
+			model I/O in memory.
+			At the time when a file is accessed, some data of the file can be in 
+			cache or not. 
+			If accessed data is already cached, there are existing data blocks 
+			representing cached data in two page cache LRU lists. 
+			As cached data is accessed again, these blocks are moved to 
+			the active list, dirty blocks are merged into a bigger dirty block, 
+			while clean data blocks are merged into another clean block. 
+			If accessed data is not in cache, new blocks representing uncached data 
+			are created and put into the top of the inactive list. 
+			For more detail when uncached data is accessed, in the case of 
+			reading file, as all uncached data is clean data, only one clean data 
+			block is added to the inactive list.
+			But in the case of writing uncached file data though cache, 
+			this written data can be either dirty or clean. 
+			Thus, at most two data blocks can be added to the inactive lists, 
+			one is clean and one is dirty. 
+			We also keep the size of the active list not larger than twice the size of 
+			the inactive list as in the kernel. This is done by moving least recently 
+			used blocks in the active list to the inactive list when the active list 
+			grows. If only a part of a data block needs to be moved, 
+			the block is split.
+
+			Next, we simulate data flushing mechanism, which writes back 
+			dirty data to disk. 
+			The data flushing simulation function traverses the sorted inactive first, 
+			and then the active list, looks for least recently used dirty blocks and 
+			sets dirty flags of these blocks to 0 until the total flushed amount 
+			reaches the amount to be flushed or there is no dirty data left in cache. 
+			If the last flushed block is not entirely flushed, it is split into 
+			two blocks, one is flushed and one remains dirty.
+			The flushing time is calculated as:
+
+			\begin{align*}
+				t_f = A / w_d
+			\end{align*}
+
+			where:
+			\begin{itemize}
+				\item $t_f$ is the flushing time
+				\item $A$ is the amount of data to be flushed
+				\item $w_d$ is the disk write bandwidth
+			\end{itemize}
+
+			For periodical flushing, because dirty data can be flushed during CPU time 
+			or cache read with disk write bandwidth, the amount of flushed data is 
+			calculated with disk bandwidth and flushing duration.
+
+			Similarly, our cache eviction simulation function frees up the page cache 
+			by traversing and deleting least recently used clean data blocks 
+			in the inactive list.
+			The amount of evicted data is passed in and data blocks are deleted 
+			from the inactive list until total evicted data reaches the amount 
+			passed in, or there is no clean block left in the list.
+			If the last evicted block is not entirely evicted, the block is split, 
+			and only one block is deleted.
+			Our cache eviction simulation does not add up to the simulated time 
+			since cache eviction time is negligible in real systems.
+
+			We encapsulate the models of memory and page cache LRU lists in 
+			\textit{MemoryManager} class. 
+			The functionalities of this class are exposed to the simulation  
+			of I/O activities including file read and file write. 
+			We create \textit{IOController} class which simulates file read and 
+			file write by interacting with the MemoryManager class, 
+			exposing APIs to user simulated applications. 
+			These two class act as an extra layer with kernel-like functionalities as 
+			illustrated in Figure~\ref{fig:interaction}. 
+			Simulated applications can simply request for a file read 
+			or file write by calling the APIs of this layer instead of 
+			calling directly to storage simulation models.
+			Because every computer has its own memory and system configurations, 
+			one IOController object and one MemoryManager object are created 
+			for each simulated machine.
 
@@
    				\caption{The additional layer between user application layer and 
-   				original SimGrid API}\label{fig:interaction}
+   				storage simulation models}\label{fig:interaction}
 			\end{figure}
 
-			We encapsulate our simulation components with kernel-like functionalities 
-			into a layer between user simulated application and storage simulation 
-			component (Figure~\ref{fig:interaction}). Our new layer includes two main
-			classes, the \textbf{\textit{IOController}}, which models the kernel 
-			I/O controlling logic, and the \textbf{\textit{MemoryManager}}, which 
-			manages page cache LRU lists and stores all necessary information about 
-			memory including capacity, amount of cache, amount of memory used 
-			by applications. Because every node has its own memory, storage devices 
-			and system configurations, there is one IOController object and one 
-			MemoryManager object created for each node. The IOController class 
-			keeps system parameters (dirty{\_}ratio, dirty{\_}background{\_}ratio), 
-			implements necessary I/O control functionalities such as read, write,
-			flushing, and cache evicting.
+			To read or write a file, a simulated application sends a request to the 
+			IOController instance on the node where the requested file is stored.
+			Based on the memory status in MemoryManager, the IOController 
+			orchestrates cache read/write, disk read/write, flushing and cache eviction 
+			to fulfill I/O requests, calculates total simulated time and returns to 
+			simulated applications.
 
-			With this layer, user simulated applications send requests to an 
-			IOController of the node where requested data is stored, instead of 
-			sending read/write requests directly to storage model as in simulators 
-			with the current SimGrid version. Based on the memory status in 
-			MemoryManager, the IOController orchestrates cache read/write, 
-			disk read/write, flushing and cache eviction to fulfill application 
-			requests. The time for each action is estimated and the total time of 
-			a request is returned to user applications.
-
-			\begin{algorithm}\caption{Read}\label{alg:read}
+			\begin{algorithm}\caption{File read simulation}\label{alg:read}
 				\small
@@
    					\EndOutput
-					\State mem\_required = max(0, 2*fs - mm.cached\_data(fn) - 
-					mm.available\_memory()) 
-					\State mem\_avail = max(0, 2 * fs - mm.cached\_data(fn) - 
-					mm.dirty\_data())
-					\State rt = rt +  mm.flush(mem\_required - 
-					mem\_avail)	
-					\State free\_mem\_required = max(0, 2 * fs - mm.cached\_data(fn) - 
-					mm.free\_memory())
-					\State mm.evict(free\_mem\_required) 
-					\If {mm.cached\_data(fn) $>$ 0}
+					\State req = max(0, 2 * fs - mm.cached(fn) - mm.avail\_mem) 
+					\State avail = max(0, 2 * fs - mm.cached(fn) - mm.dirty)
+					\State rt = rt + mm.flush(req - avail)	
+					\State req = max(0, 2 * fs - mm.cached(fn) - mm.free\_mem)
+					\State mm.evict(req) 
+					\If {mm.cached(fn) $>$ 0}  \Comment{Read file cached data}
     					\State mm.cache\_read(fn) 
-						\If {mm.cached\_data(fn) == fs}
-    						\State time = mm.cache\_data(fn) / $r_m$
-    						\State IOController.periodic\_flush(time) 
+						\If {mm.cached(fn) == fs} \Comment{If all file data is cached}
+    						\State time = mm.cached(fn) / $r_m$
+    						\State mm.periodic\_flush(time) 
 							\State rt = rt + time
@@
 					\EndIf
-					\If {mm.cached\_data(fn) $<$ fs}
-						\State rt = rt + IOController.flush(mm.dirty\_data())
-						\State mm.disk\_read(fs - mm.cache\_data(fn), fn)
-    					\State rt = rt + (fs - mm.cache\_data(fn)) / $r_d$
+					\If {mm.cached(fn) $<$ fs} \Comment{Read uncached data from disk}
+						\State mm.disk\_read(fs - mm.cached(fn), fn)
+    					\State rt = rt + (fs - mm.cached(fn)) / $r_d$
 					\EndIf
+
+					\State \Return rt
+
+				\end{algorithmic}
+			\end{algorithm}
+
+			The algorithm used to simulate file read is described in 
+			Algorithm~\ref{alg:read}. 
+			In Linux systems with write back storage devices, data can be either read 
+			from cache with memory bandwidth or read from disk with disk bandwidth. 
+			Therefore, total read time can be considered as a linear function 
+			of the read time with each of these bandwidths, 
+			enabling us to generalize reading process as two separate phases, 
+			cache read and disk read. 
+			In our simulator, when a user simulated application reads a file, 
+			it sends a file read request to IOController. 
+			Then, IOController calculates the amount of required memory to read 
+			the file based on the amount of available memory, file size, and 
+			the amount of cached data of that file (line 10). 
+			If there is not enough available memory, dirty data is flushed to disk 
+			to increase the amount of memory available by calling flush function 
+			of MemoryManager with the amount of flushed data, flushing time is 
+			simulated and added to total run time (line 11-12).
+			After data is flushed, MemoryManager is asked to evict data 
+			from cache to accommodate the requested file (line 13-14). 
+			Next, if file is cached either partially or entirely, MemoryManager is 
+			called to re-access and update this cached data (line 16). 
+			If all file data is cached, periodical flushing is called is to flush 
+			dirty data during the cache read, and time to read from cache is added to 
+			total run time, (line 18-20). 
+			If file is not cached or partially cached, a disk read is required (line 23) 
+			This blocks flushing and makes the cache read time neglectable. 
+			Finally, data is read from disk to cache and managed 
+			by MemoryManager, disk read time is estimated with disk read 
+			bandwidth and added to total run time (line 24-25).
+
+			\begin{algorithm}\caption{File write simulation}\label{alg:write}
+				\small
+				\begin{algorithmic}[1]
+					\Input
+        				\Desc{fn}{file name}
+        				\Desc{fs}{file size}
+						\Desc{rt}{current simulated run time}
+						\Desc{mm}{MemoryManager object}
+						\Desc{$w_m$}{memory write bandwidth}
+						\Desc{$w_d$}{disk write bandwidth}
+   					\EndInput
+   					\Output
+						\Desc{rt}{simulated run time after the file is read}
+   					\EndOutput
+					\State dirty\_left = $d$ * mm.avail\_mem - mm.dirty
+					\If {dirty\_left $>$ 0}
+   						\State free\_amt = min(fs, $M$) \Comment{$M$ in Equation
+   						~\ref{equa:freeamt}}
+    					\State mm.evict(free\_amt - mm.free\_mem)
+    					\State free\_time = free\_amt / $w_m$
+    					\State mm.write(fn, free\_amt, dirty = free\_amt - free\_time * $w_d$)
+   						\State rt = rt + free\_time
+    				\EndIf
+
+					\If {free\_amt $\geq$ fs}
+						\Return rt
+					\EndIf
+
+					\State throttled\_amt = fs - free\_amt \Comment{Written with disk bandwidth}
+					\State req = max(0, throttled\_amt  - mm.free\_mem)
+					\State evicted\_amt = min(req, mm.evictable)
+					\State mm.evict(evicted\_amt)
+					\State cached\_amt = min(mm.free\_mem, throttled\_amt)
+					\State mm.write(fn, cached = cached\_amt, dirty = 0)
+					\State rt = rt + throttled\_amt /$w_d$
 
@@
 
-			The algorithm to generalize file read is described in 
-			\textit{Algorithm~\ref{alg:read}}. In Linux systems, data can be read in 
-			two ways, read from cache with memory bandwidth and read from disk 
-			with disk bandwidth. Because total read time can be seen as a 
-			linear function of the read time in each of these two modes, 
-			we can generalize reading process as two separate and sequential phases, 
-			one is cache read, and the other is disk read. When a user application 
-			requests a file read, it calls the read method of the IOController. Then, 
-			IOController calculates how much memory is required to read the file 
-			based on the amount of available memory, file size, and the amount of 
-			file data cached (line 10). If there is not enough available memory, 
-			dirty data is flushed to disk to make more memory available by calling flush
-			function with the amount of memory needed as the parameter (line 11-12).
-			IOController then evicts old blocks from the inactive list to
-			accommodate the requested file and flushing time is simulated (line 13-14). 
-			After that, IOController starts estimating file read time by first checking 
-			if a some file data is cached (line 15). If there is cached data, 
-			it is read from cache and the blocks containing this data are updated in 
-			the page cache LRU lists with the last access time (line 16). If all file data 
-			is cached, time to read from cache is added to total run time, and 
-			periodical flushing is triggered to flush dirty data during the cache read 
-			(line 18-19). If only a part of the file is in cache, disk read is required, 
-			this does not allow flushing and makes the cache read time neglectable. 
-			At the beginning of a disk read, dirty data is flushed, then data is read 
-			and added to the inactive list, disk read time is estimated with disk read 
-			bandwidth and added to total run time (line 23-27).
+			Algorithm~\ref{alg:write} describes our simulation of file write. 
+			Similar to the file read simulation, a file can also be written with 
+			two different bandwidths, memory bandwidth and disk bandwidth. 
+			Files can be written with memory bandwidth before the amount 
+			of dirty data reaches dirty\_ratio. Thus, our algorithm initially checks 
+			this condition (line 10), and estimates the maximum amount 
+			of data can be written with memory bandwidth with the following 
+			formula:
 
-			\textit{Algorithm~\ref{alg:write}} describes our
-			generalization of file write. Similar to the file read process,
-			a file can also be written with two different bandwidths,
-			memory bandwidth and disk bandwidth, and thus file write can be
-			generalized with two modes. File can be written with memory
-			bandwidth before the page cache is saturated, or the amount of
-			dirty data reaches dirty\_ratio. Firstly, if dirty\_ratio is not reached 
-			the amount of data can be written with memory bandwidth is
-			calculated with the formula:
 			\begin{equation}
@@
 
-			By this formular, we assume that data is written with with memory 
-			bandwidth until all file data is written or dirty\_ratio is reached, 
-			the kernel concurrently writes data to cache and flushes dirty data 
-			to disk. If file size is smaller than or equal to $M$, the whole file is 
-			written to cache as dirty data with memory bandwidth (line 12). Otherwise, 
-			the remaining file data is written with disk bandwidth. Before simulating 
-			cache write, IOController evicts the page cache to accommodate new data 
-			(line 13). Then, data is written to cache as dirty data with memory 
-			bandwidth write time is added to simulated time (line 14-16). Because 
-			cache write and data flushing are concurrent, the actual amount of written 
-			dirty data reduces by the amount of flushed data (line 15). In the 
-			second phase, in which data is written with disk bandwidth, our simulator 
-			can generalize not only the case that data is written directly to disk, 
-			but also the case that applications have to wait for dirty data 
-			being flushed. Before writing, if the amount of free memory is not enough 
-			for the remaining file data, evictable memory is evicted (line 20-23). 
-			Here, the kernel does not need to guarantee that the amount of free 
-			memory after eviction can accommodate the remaining data. In the case 
-			free memory is less than remaining data, we the kernel can flush, 
-			evict and write new data concurrently. Finally, IOController simulates 
-			the time to write the remaining amount data, adds written data to 
-			page cache LRU lists, and the amount of dirty data remains the same 
-			(line 24). The total simulated time of the second phases is added and 
-			returned to user applications.
-
-			\begin{algorithm}\caption{Write}\label{alg:write}
-				\small
-				\begin{algorithmic}[1]
-					\Input
-        				\Desc{fn}{file name}
-        				\Desc{fs}{file size}
-						\Desc{rt}{current simulated run time}
-						\Desc{mm}{MemoryManager object}
-						\Desc{$w_m$}{memory write bandwidth}
-						\Desc{$w_d$}{disk write bandwidth}
-   					\EndInput
-   					\Output
-						\Desc{rt}{simulated run time after the file is read}
-   					\EndOutput
-					\State dirty\_left = $d$ * mm.available\_memory() - mm.dirty\_data()
-					\If {dirty\_left $>$ 0}
-   						\State free\_amt = min(fs, $M$) \Comment{$M$ in Equation
-   						~\ref{equa:freeamt}}
-    					\State IOController.evict(free\_amt - mm.free\_memory())
-    					\State free\_time = free\_amt / $w_m$
-    					\State mm.write(fn, free\_amt, dirty = free\_amt - free\_time * $w_d$)
-   						\State rt = rt + free\_time
-    				\EndIf
-
-					\If {free\_amt $\geq$ fs}
-						\Return rt
-					\EndIf
-
-					\State throttled\_amt = fs - free\_amt
-					\State cache\_required = max(0, throttled\_amt  - mm.free\_memory())
-					\State evicted\_amt = min(cache\_required, mm.evictable\_memory())
-					\State mm.evict(evicted\_amt)
-					\State mm.write(fn, cached = min(mm.free\_memory(), 
-					throttled\_amt), dirty = 0)
-					\State rt = rt + throttled\_amt /$w_d$
-
-					\Return rt
-
-				\end{algorithmic}
-			\end{algorithm}
-
-			Data flushing is another key function in our simulator. It takes 
-			the input as the amount of flushed data and the returned value is the 
-			simulated flushing time. To flush dirty data, IOController simply 
-			traverses the sorted inactive and active list, reduces the dirty data 
-			amount by unmarking dirty blocks until the flushed amount reached 
-			the amount passed in or there is no dirty data left in cache. 
-			Similarly, periodical flushing is the function that simulates data flushing 
-			during CPU time and cache read. However, the flushed data amount is 
-			calculated with the input which is flushing time. In both flushing 
-			functions, the last block being flushed is split if its data is not 
-			entirely flushed.
-
-			Equally important, cache eviction function is called from both 
-			read and write to free up the page cache when free memory 
-			is not enough. The amount of evicted data is passed in as in flushing, 
-			however, this function does not add up the simulated time as 
-			cache eviction time is neglectable in real systems . When eviction 
-			is called, IOController deletes the last accessed blocks from the 
-			inactive list until the evicted amount reaches the amount passed in or 
-			there is no evictable memory left. If the block is not entirely deleted, 
-			the block is split.
-
-			Last but not least, we need to modified page cache LRU lists after 
-			every I/O activity. When IOController reads the data that is not cache, 
-			the data is read and added to the inactive list. If the data is already 
-			in cache, the cached blocks are moved to the active list with the 
-			last access time updated. Finally, MemoryManager makes the page cache 
-			LRU lists balanced by moving blocks from the longer list to the shorter 
-			and sorting the lists.
+			In this formula, we assume that data is written with with memory 
+			bandwidth until dirty\_ratio is reached, the kernel writes data 
+			to cache and flushes dirty data to disk concurrently. 
+			If file size is smaller than or equal to $M$, the whole file is 
+			written to cache as dirty data with memory bandwidth (line 12). 
+			Otherwise, the remaining file data is written with disk bandwidth (line 20). 
+			Before simulating cache write, MemoryManager is asked to evict 
+			the page cache to accommodate written data (line 13). 
+			Then, data is written to cache as dirty data with memory 
+			bandwidth, and write time is added to total simulated time (line 14-16). 
+			Because cache write and data flushing are concurrent in our algorithm, 
+			the actual amount of written dirty data reduces by the amount 
+			of data during cache write time (line 15). 
+			If not all data is written with memory write bandwidth, the remaining data 
+			is written with disk write bandwidth. This can simulate both cases that 
+			data is written directly to disk, or applications wait for dirty data 
+			to be flushed, and evicted to write new data. 
+			If the amount of free memory is not enough for the remaining file data, 
+			evictable memory is evicted (line 20-23). 
+			Here, the simulator does not guarantee that the amount of free memory 
+			after eviction can accommodate the remaining data. 
+			In the case that free memory is less than remaining data, the amount to be 
+			cached is limited by the amount of free memory (line 24). 
+			However, the kernel still flushes, evicts and writes new data concurrently, 
+			and remaining data is still written to cache with disk write bandwidth.
+			Thus, we can simulate the time to write the remaining data with 
+			disk bandwidth and ask MemoryManager to add clean cached data (line 25). 
+			The simulated time to write width disk bandwidth is added to total 
+			simulated run time and returned to simulated applications (line 26).
 
@@
 
-			Firstly, to validate our simulation model, we create a simple 
-			simulator independent to simulation frameworks and libraries. 
+			To validate our simulation model, we create a simple 
+			simulator which is independent to simulation frameworks and libraries. 
 			This enables us to evaluate the accuracy and correctness of our 
-			model in a simple use case before integrating it in SimGrid. 
+			model in a simple use case before integrating it to SimGrid. 
 			This simulator simulates a single-threaded pipeline running on a system 
@@
 
-			Having our model validated, we create other different
-			simulators of different use cases using the current version of
-			SimGrid and the SimGrid version that is extended with our
-			model. Then, we compare the results of the simulators with
+			Having our model validated, we create simulators for different use cases 
+			using the current version of SimGrid and the SimGrid version that is 
+			extended with our model. We leverage Wrench framework, which is 
+			built on top of SimGrid to simulate our pipelines and system.
+			Then, we compare the results of the simulators with
 			original SimGrid and extended SimGrid with the results of real
-			pipelines on a real system. The results are compared in two
-			different aspects, task completion time and I/O time, as well
-			as the ability of the simulator in generalization of the page
-			cache in terms of dirty data and cache used.
+			pipelines on a real system.
 
@@
 
-			When designing experiments, we consider a range of scenarios
-			that have different impacts on the results. The experiments are
-			a number of pipelines of sequential tasks. Each task reads an
-			input file, performs some computation, generates an output file
-			and writes it to disk. The output file of the previous task is
-			the input file of the next task. Thus, each pipeline has only one
-			actual input file.
+			When designing experiments, our goal is to evaluate our simulation model 
+			in single-threaded and multi-threaded applications which access data on 
+			on different file system types. 
+			To do that, we create a pipeline of sequential tasks, each task 
+			reads an input file, increases every byte of the input, and writes the 
+			output to disk. The output of the previous task is the input of 
+			the next task. We also adopt an use case in real applications and 
+			to evaluate the applicability of our simulation model.
 
-			Initially, we want to have a detailed view of the effects of data size 
-			on I/O behaviors and validate our simulator in modeling and 
-			capturing those impacts. Thus, there should be only one pipeline and 
-			the size of input and output files should vary in a range that 
-			the change in file size can cause significant changes on I/O behaviors.
+			The simplest scenario is a single-threaded application, where memory 
+			bandwidth and disk bandwidth are not shared with another thread. 
+			This allows us to compare the detailed results in terms of read/write time, 
+			memory states after every single task and ,more importantly, validate our 
+			simulation model.
 
-			Furthermore, the number of pipelines should also vary. Having our 
-			simulator validated on a single threaded experiment, the next goal 
-		    is to evaluate our simulation model with a multi-threaded experiment, 
-		    with a varying number of pipelines run on a multi-core host and write 
-		    to the same local storage device. As a result, the shared bandwidth 
-		    the pipelines is different with different numbers of concurrent processes. 
-		    By doing so, we can evaluate the performance of our model on the simulation 
-		    of multi-threaded, shared bandwidth applications.
+			For multi-threaded applications, we do the experiment with a varying 
+			number of concurrent pipelines running on a multi-core host 
+			and writing to the same local storage device. 
+		    By doing so, we can assess the performance of our model in the case  
+		    of concurrent I/O with shared bandwidth on a local storage device.
 
-		    Moreover, different storage types, which include local and shared 
-		    storage, should also be taken into account when they are commonly used 
-		    in computer systems. The scenario of the next experiment is the same as 
-		    the multi-threaded experiment, but should replace local storage device 
-		    with a shared file systems, so that we can not only access our 
-		    simulator on I/O with shared file system, but also compare the 
+		    Since different storage types should also be taken into account, 
+		    we conduct the next experiment with the same scenario as 
+		    in the multi-threaded experiment, except that files are stored on a 
+		    shared file system instead of local disk. 
+		    In this way, we can not only evaluate our simulation model on I/O 
+		    with shared file system, but also compare the 
 		    performance of the simulator on different storage types.
 
-		    Finally, we adopt a real neuroimaging pipeline to evaluate the 
-		    applicability of our simulator in simulating I/O time in real
-			applications. We simulate the pipeline with a simulated workflow running 
-			on a simulated plaform and compare the simulation results with the real 
-			results. Because our work focuses on I/O time, we assume
-			that CPU time is correctly modeled and use the CPU time measured 
-			in the real pipelines to setup our simulated workflows.
+		    Finally, we adopt a real pipeline in a real neuroimaging applications. 
+		    We simulate the pipeline with a simulated workflow in SimGrid 
+			and compare the simulation results with the real results. 
+			Because our work focuses on I/O time, we assume that CPU time is 
+			correctly modeled and use the CPU time measured in the real pipelines 
+			to setup our simulated workflows.
 
@@
 
-				In this experiment, our goal is to validate our model with a simple 
-				simulator implemented in Python and compare the simulation precision 
-				of our simulation model and the existing model in SimGrid. 
-				Thus, we simulate one single thread pipeline of three tasks with the
-				Python simulator, SimGrid simulator and run on our real system. 
-				All files are read and written locally. We use the input sizes of 
-				20 GB, 50 GB, 75 GB and 100 GB to run the pipeline on the cluster, 
-				measure the CPU time of the pipeline with each input to simulate 
-				the pipeline with the simulators.
+				In this experiment, our goal is to validate our model by comparing 
+				the simulation results of the simple Python simulator with results of 
+				a simulator implemented with original SimGrid, a simulator 
+				implemented with SimGrid integrated with our model, and a real pipeline 
+				running of the cluster. 
+				The pipeline consists of three tasks, each task reads a file, 
+				increases every byte and writes output file to local disk. 
+				We use the input sizes of 20 GB, 50 GB, 75 GB and 100 GB, run 
+				the pipeline on the cluster, measure the CPU time of the pipeline 
+				with each input to simulate the pipeline with the simulators.
 
@@
 				In the second experiment, we evaluate the simulator in concurrent I/O 
-				simulation with a multi-threaded application. As we run the pipelines on 
-				only one compute node of the cluster, there are upto 32 cores per node 
-				and all input and output files are read and written on the same disk,  
+				with a multi-threaded application. As we run the pipelines on 
+				only one compute node of the cluster, which has upto 32 cores per node,  
 				we create 32 input files with the size of 3GB each and vary the number of 
-				concurrent pipelines from 1 to 32. Because our goal in this experiment is 
-				to evaluate concurrent I/O in our simulation model, and multi-threaded 
-				simulator implementation of the Python is expensive, we only compare the 
-				performance of simulators with original SimGrid, SimGrid integrated with 
-				our model and and real pipelines. The results are compared in total 
-				makespan of the pipelines, cumulative read time and cumulative write time.
+				concurrent pipelines from 1 to 32. 
+				Because multi-threaded simulator implementation in Python is 
+				expensive, and we have our model integrated in SimGrid, 
+				we only compare the results of simulators with original SimGrid, 
+				SimGrid integrated with our model and and real pipelines. 
+				The results are compared in total makespan of the pipelines, 
+				cumulative read time and cumulative write time.
 

### 2020-06-15 paper_io_sim.tex
Edits in Algo 1

@@
 
-\author{\IEEEauthorblockN{Hoang-Dung Do, Tristan Glatard and Val\'erie Hayot-Sasson
+\author{\IEEEauthorblockN{Hoang-Dung Do, Val\'erie Hayot-Sasson, Tristan Glatard
   }\\
@@
         				\Desc{fn}{file name}
-        				\Desc{fs}{file size}
+        				\Desc{fs}{file size (assumed to fit in memory)}
 						\Desc{rt}{current simulated run time}
@@
    					\EndOutput
-					\State req = max(0, 2 * fs - mm.cached(fn) - mm.avail\_mem) 
-					\State avail = max(0, 2 * fs - mm.cached(fn) - mm.dirty)
-					\State rt = rt + mm.flush(req - avail)	
-					\State req = max(0, 2 * fs - mm.cached(fn) - mm.free\_mem)
-					\State mm.evict(req) 
+					\State rt = rt + mm.flush(2*fs - mm.cached(fn) - mm.avail\_mem)
+					\State mm.evict(2*fs - mm.cached(fn) - mm.free\_mem) 
 					\If {mm.cached(fn) $>$ 0}  \Comment{Read file cached data}
     					\State mm.cache\_read(fn) 
-						\If {mm.cached(fn) == fs} \Comment{If all file data is cached}
-    						\State time = mm.cached(fn) / $r_m$
-    						\State mm.periodic\_flush(time) 
-							\State rt = rt + time
-						\EndIf
+						\State rt = rt + mm.cached(fn) / $r_m$
 					\EndIf

### 2020-06-21 paper_io_sim.tex
Section 3 revised

@@
 			believe that they are the most important factors that affect the I/O time. 
+			We separate our model into components and encapsulate them in a 
+			kernel-like layer described in Figure~\ref{fig:interaction}. 
+			This layer consists of two main classes, MemoryManager and IOManager.
+			MemoryManager is responsible for memory simulation, which includes 
+			memory I/O, flushing, cache eviction, and exposed its functions 
+			to IOManager. The IOController class models higher abstraction with the 
+			simulation of file read and file write.
+			With this layer, simulated applications can simply request for 
+			file reads and file writes by calling functions of IOManager class 
+			instead of calling directly to storage simulation models.
 
-			Memory is modeled with capacity, read bandwidth and write 
-			bandwidth. When data is read from or written to memory, the actual 
+			\begin{figure}
+   				\centering
+   				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
+   				\caption{The additional layer between user application layer and 
+   				storage simulation models}\label{fig:interaction}
+			\end{figure}
+
+			\subsubsection{Memory simulation model}
+
+			We model memory with capacity, read bandwidth and write bandwidth. 
+			When data is read from or written to memory, the actual 
 			read/write time is calculated with the amount of data and memory 
@@
 
-			For periodical flushing, because dirty data can be flushed during CPU time 
-			or cache read with disk write bandwidth, the amount of flushed data is 
-			calculated with disk bandwidth and flushing duration.
+			For periodical flushing, our simulator searches in two page cache 
+			LRU lists and flushes all the expired dirty data block. 
+			A dirty data block is considered expired if the time from its last access 
+			to current simulated time is longer than the predefined expire time. 
+			Then the flushing time is calculated with total amount of flushed data.
 
-			Similarly, our cache eviction simulation function frees up the page cache 
-			by traversing and deleting least recently used clean data blocks 
-			in the inactive list.
+			Similarly to data flushing, our cache eviction simulation function frees up 
+			the page cache by traversing and deleting least recently used clean 
+			data blocks in the inactive list.
 			The amount of evicted data is passed in and data blocks are deleted 
@@
 
-			We encapsulate the models of memory and page cache LRU lists in 
-			\textit{MemoryManager} class. 
-			The functionalities of this class are exposed to the simulation  
-			of I/O activities including file read and file write. 
-			We create \textit{IOController} class which simulates file read and 
-			file write by interacting with the MemoryManager class, 
-			exposing APIs to user simulated applications. 
-			These two class act as an extra layer with kernel-like functionalities as 
-			illustrated in Figure~\ref{fig:interaction}. 
+			\subsubsection{File I/O simulation model}
+
 			Simulated applications can simply request for a file read 
@@
 			for each simulated machine.
-
-			\begin{figure}
-   				\centering
-   				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
-   				\caption{The additional layer between user application layer and 
-   				storage simulation models}\label{fig:interaction}
-			\end{figure}
 
@@
 						\Desc{mm}{MemoryManager object}
-						\Desc{$r_m$}{memory read bandwidth}
-						\Desc{$r_d$}{disk read bandwidth}
+						\Desc{rm}{memory read bandwidth}
+						\Desc{sm}{storage simulation model}
    					\EndInput
@@
    					\EndOutput
-					\State rt = rt + mm.flush(2*fs - mm.cached(fn) - mm.avail\_mem)
+					\State rt = rt + mm.flush(2*fs - mm.cached(fn) - mm.evictable\_mem) 
 					\State mm.evict(2*fs - mm.cached(fn) - mm.free\_mem) 
 					\If {mm.cached(fn) $>$ 0}  \Comment{Read file cached data}
-    					\State mm.cache\_read(fn) 
-						\State rt = rt + mm.cached(fn) / $r_m$
+    					\State mm.cache\_read(fn)  \Comment{Update blocks in LRU lists}
+    					\State pft = mm.periodic\_flush(rt) \Comment{Periodical flushing time}
+						\State rt = rt + max(mm.cached(fn) / rm, pft) \Comment{Add the longer}
 					\EndIf
 					\If {mm.cached(fn) $<$ fs} \Comment{Read uncached data from disk}
+						\State rt = rt + mm.periodic\_flush(rt) \Comment{Another flushing interval}
 						\State mm.disk\_read(fs - mm.cached(fn), fn)
-    					\State rt = rt + (fs - mm.cached(fn)) / $r_d$
+    					\State rt = rt + sm.read(fs - mm.cached(fn))
 					\EndIf
@@
 			it sends a file read request to IOController. 
-			Then, IOController calculates the amount of required memory to read 
-			the file based on the amount of available memory, file size, and 
-			the amount of cached data of that file (line 10). 
-			If there is not enough available memory, dirty data is flushed to disk 
-			to increase the amount of memory available by calling flush function 
-			of MemoryManager with the amount of flushed data, flushing time is 
-			simulated and added to total run time (line 11-12).
+			First, dirty data is flushed to disk if needed to make the amount of 
+			available memory enough to read the file.
+			Data flushing function in MemoryManager is called, flushing time is 
+			simulated and added to total run time (line 10).
 			After data is flushed, MemoryManager is asked to evict data 
-			from cache to accommodate the requested file (line 13-14). 
+			from cache to accommodate the requested file (line 11). 
 			Next, if file is cached either partially or entirely, MemoryManager is 
-			called to re-access and update this cached data (line 16). 
-			If all file data is cached, periodical flushing is called is to flush 
-			dirty data during the cache read, and time to read from cache is added to 
-			total run time, (line 18-20). 
-			If file is not cached or partially cached, a disk read is required (line 23) 
-			This blocks flushing and makes the cache read time neglectable. 
-			Finally, data is read from disk to cache and managed 
-			by MemoryManager, disk read time is estimated with disk read 
-			bandwidth and added to total run time (line 24-25).
+			called to re-access and update this cached data (line 13). 
+			Then, periodical flushing is called to flushed expired dirty data 
+			from cache (line 14).
+			Since periodical flushing and cache read can be concurrent, the longer 
+			time is added to simulated run time (line 15).
+			If file is not cached or partially cached, a disk read is required (line 17). 
+			Periodical flushing is called once again, but this time the flushing time is 
+			added to total simulated time since flushing and disk read are time-shared, 
+			or sequential in other words (line 18).
+			Finally, data is read from disk to cache and managed by MemoryManager, 
+			disk read time estimated by the storage model is added to 
+			total simulated run time (line 19-20).
 
@@
 						\Desc{mm}{MemoryManager object}
-						\Desc{$w_m$}{memory write bandwidth}
-						\Desc{$w_d$}{disk write bandwidth}
+						\Desc{wm}{memory write bandwidth}
+						\Desc{sm}{storage simulation model}
    					\EndInput
@@
    					\EndOutput
-					\State dirty\_left = $d$ * mm.avail\_mem - mm.dirty
-					\If {dirty\_left $>$ 0}
-   						\State free\_amt = min(fs, $M$) \Comment{$M$ in Equation
-   						~\ref{equa:freeamt}}
-    					\State mm.evict(free\_amt - mm.free\_mem)
-    					\State free\_time = free\_amt / $w_m$
-    					\State mm.write(fn, free\_amt, dirty = free\_amt - free\_time * $w_d$)
-   						\State rt = rt + free\_time
+   					\State pft = mm.periodic\_flush(rt)
+					\State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
+					\State mem\_amt = min(fs, remain\_dirty)
+					\If {remain\_dirty $>$ 0} \Comment{Write with memory bandwidth}
+    					\State mm.evict(mem\_amt - mm.free\_mem)
+    					\State mm.write(fn, mem\_amt)
     				\EndIf
-
-					\If {free\_amt $\geq$ fs}
-						\Return rt
+    				\State rt = rt + max(mem\_amt / wm, pft) \Comment{Concurrent processes}
+					\If {mem\_amt $<$ fs}  \Comment{Write with disk bandwidth}
+						\State mm.flushed(fs - mem\_amt)  \Comment{Concurrent flushing}
+						\State mm.evict(fs - mem\_amt  - mm.free\_mem) 
+						\State mm.write(fn, min(fs - mem\_amt, mm.free\_mem))
+						\State rt = rt + sm.write(fs - mem\_amt)
 					\EndIf
-
-					\State throttled\_amt = fs - free\_amt \Comment{Written with disk bandwidth}
-					\State req = max(0, throttled\_amt  - mm.free\_mem)
-					\State evicted\_amt = min(req, mm.evictable)
-					\State mm.evict(evicted\_amt)
-					\State cached\_amt = min(mm.free\_mem, throttled\_amt)
-					\State mm.write(fn, cached = cached\_amt, dirty = 0)
-					\State rt = rt + throttled\_amt /$w_d$
 
@@
 			Algorithm~\ref{alg:write} describes our simulation of file write. 
+			Initially, periodical flushing is called to flush expired dirty data 
+			at the beginning of the write (line 10).
 			Similar to the file read simulation, a file can also be written with 
@@
 			of dirty data reaches dirty\_ratio. Thus, our algorithm initially checks 
-			this condition (line 10), and estimates the maximum amount 
-			of data can be written with memory bandwidth with the following 
-			formula:
-
-			\begin{equation}
-				M = max(0, \frac{(d*C - D)*w_m}{w_m - w_d})\label{equa:freeamt}
-			\end{equation}
-
-			where:
-			\begin{itemize}
-				\item $M$ is the maximum amount of data that can be written with memory 
-				write bandwidth
-				\item $d$ is dirty ratio
-				\item $C$ is the total available memory
-				\item $D$ is the amount of dirty data
-				\item $w_m$ is the memory write bandwidth 
-				\item $w_d$ is the disk write bandwidth			
-			\end{itemize}
-
-			In this formula, we assume that data is written with with memory 
-			bandwidth until dirty\_ratio is reached, the kernel writes data 
-			to cache and flushes dirty data to disk concurrently. 
-			If file size is smaller than or equal to $M$, the whole file is 
-			written to cache as dirty data with memory bandwidth (line 12). 
-			Otherwise, the remaining file data is written with disk bandwidth (line 20). 
-			Before simulating cache write, MemoryManager is asked to evict 
-			the page cache to accommodate written data (line 13). 
-			Then, data is written to cache as dirty data with memory 
-			bandwidth, and write time is added to total simulated time (line 14-16). 
-			Because cache write and data flushing are concurrent in our algorithm, 
-			the actual amount of written dirty data reduces by the amount 
-			of data during cache write time (line 15). 
-			If not all data is written with memory write bandwidth, the remaining data 
-			is written with disk write bandwidth. This can simulate both cases that 
-			data is written directly to disk, or applications wait for dirty data 
-			to be flushed, and evicted to write new data. 
-			If the amount of free memory is not enough for the remaining file data, 
-			evictable memory is evicted (line 20-23). 
-			Here, the simulator does not guarantee that the amount of free memory 
-			after eviction can accommodate the remaining data. 
-			In the case that free memory is less than remaining data, the amount to be 
-			cached is limited by the amount of free memory (line 24). 
-			However, the kernel still flushes, evicts and writes new data concurrently, 
-			and remaining data is still written to cache with disk write bandwidth.
-			Thus, we can simulate the time to write the remaining data with 
-			disk bandwidth and ask MemoryManager to add clean cached data (line 25). 
-			The simulated time to write width disk bandwidth is added to total 
-			simulated run time and returned to simulated applications (line 26).
+			the amount of data that can be written with memory write 
+			bandwidth (line 11-12).
+			If this amount is greater than 0, MemoryManager is asked to evict 
+			data from cache if needed and write this amount of data to cache 
+			(line 14-15).
+			The simulated run time can be added with the longer time of cache write 
+			time and periodical flushing time because these can happen 
+			concurrently (line 17).
+			The amount of data that is not written with memory bandwidth is written 
+			with disk bandwidth (line 18). 
+			Now, the dirty ratio threshold is reached, we have to flush and evict 
+			data from cache as much as possible so that 
+			the remaining data can be written to cache (line 19-20). 
+			Here, MemoryManager does not guarantee that the amount of flushed 
+			and evicted data is enough for the remaining data. In the case it is 
+			less than the required amount, the remaining data is written to cache 
+			but then can be flushed and evicted right away. Thus, the amount 
+			of remaining data that is written and kept in cache after the write 
+			is limited by the amount free memory (line 21). 
+			Finally, the write time is simulated, added to the total time and 
+			returned to simulated applications (line 22).
 

### 2020-06-21 paper_io_sim.tex
Minor fix

@@
 			For periodical flushing, our simulator searches in two page cache 
-			LRU lists and flushes all the expired dirty data block. 
+			LRU lists and flushes all the expired dirty data blocks. 
 			A dirty data block is considered expired if the time from its last access 
 			to current simulated time is longer than the predefined expire time. 
-			Then the flushing time is calculated with total amount of flushed data.
+			The flushing time is calculated by storage simulation model with 
+			total amount of flushed data.
 
@@
 			cache read and disk read. 
-			In our simulator, when a user simulated application reads a file, 
+			When a user simulated application reads a file, 
 			it sends a file read request to IOController. 
-			First, dirty data is flushed to disk if needed to make the amount of 
-			available memory enough to read the file.
+			First, dirty data is flushed to disk to make more available memory 
+			to read the file if needed.
 			Data flushing function in MemoryManager is called, flushing time is 
@@
 			from cache (line 14).
-			Since periodical flushing and cache read can be concurrent, the longer 
-			time is added to simulated run time (line 15).
+			Since periodical flushing and cache read can be concurrent, the time of 
+			the longer activity between them is added to simulated run time (line 15). 
 			If file is not cached or partially cached, a disk read is required (line 17). 
@@
 			of dirty data reaches dirty\_ratio. Thus, our algorithm initially checks 
-			the amount of data that can be written with memory write 
+			the amount of file data that can be written with memory write 
 			bandwidth (line 11-12).
 			If this amount is greater than 0, MemoryManager is asked to evict 
-			data from cache if needed and write this amount of data to cache 
+			data from cache if needed and to write this amount of data to cache 
 			(line 14-15).

### 2020-06-30 paper_io_sim.tex
Some comments in methods

@@
 		I/O in memory, data flushing and cache eviction mechanisms 
-		implemented in Linux. 
+		implemented in the Linux kernel. 
 		We also detail the design the simulators in order to simulate I/O activities 
-		and the implementation in Python and SimGrid framework. 
-		Finally, we describe some experimental scenarios to evaluate our model 
-		with a Python simulator, a baseline SimGrid simulator, a simulator 
-		with SimGrid integrated with our model, and the results from real pipelines.
+		and the implementation in Python and SimGrid framework \tristan{check grammar}. 
+		Finally, we describe experimental scenarios to evaluate our model using real applications and
+		a Python simulator prototype, the current SimGrid simulator, our extension of SimGrid, and the results from real pipelines.
 
@@
 
-			In our I/O simulation model, we try to simulate the write back page cache 
-			mechanisms including memory I/O, cache eviction and data flushing as we 
-			believe that they are the most important factors that affect the I/O time. 
-			We separate our model into components and encapsulate them in a 
-			kernel-like layer described in Figure~\ref{fig:interaction}. 
-			This layer consists of two main classes, MemoryManager and IOManager.
-			MemoryManager is responsible for memory simulation, which includes 
-			memory I/O, flushing, cache eviction, and exposed its functions 
-			to IOManager. The IOController class models higher abstraction with the 
-			simulation of file read and file write.
-			With this layer, simulated applications can simply request for 
-			file reads and file writes by calling functions of IOManager class 
-			instead of calling directly to storage simulation models.
+			Our simulation model simulates the main mechanisms of the
+			write-back page cache including memory I/O, cache eviction and
+			data flushing. We separate our model in two components, the IO
+			Controller and the Memory Manager, which together simulate I/O
+			in the Linux kernel (Figure~\ref{fig:interaction}). The Memory
+			Manager simulates memory I/O, flushing, and cache eviction, whereas the
+			IO Controller simulates file read and file write.
+			Simulated applications can simply request
+			file reads and file writes from the IO Manager
+			instead of calling simulated storage directly.
 
@@
    				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
-   				\caption{The additional layer between user application layer and 
-   				storage simulation models}\label{fig:interaction}
+   				\caption{Overview of the simulated write-back page cache}\label{fig:interaction}
 			\end{figure}
@@
 
-			We model memory with capacity, read bandwidth and write bandwidth. 
+			We model memory with a capacity, a read bandwidth and a write bandwidth. 
 			When data is read from or written to memory, the actual 
@@
 			\begin{itemize}
-				\item $t_{mr}$ is time to read data from memory
-				\item $t_{mw}$ is time to write data to memory
+				\item $t_{mr}$ is the time to read data from memory
+				\item $t_{mw}$ is the time to write data to memory
 				\item $D$ is the amount of data to read/write
-				\item $r_m$ is memory read bandwidth
-				\item $w_m$ is memory write bandwidth
+				\item $r_m$ is the memory read bandwidth
+				\item $w_m$ is the  memory write bandwidth
 			\end{itemize}
+
+			\tristan{how about parallel accesses to memory? I think it would be a motivation to simualte memory as a storage device, 
+			and leverage existing bandwidth sharing mechanisms in SimGrid. You could updated Fig 1 to include memory in the storage devices, and 
+			explain in your Python simulator implementation that only a simplistic model of storage device is used (for both disk and memory).}
 

### 2020-07-01 paper_io_sim.tex
Some comments in methods
Methods revised

@@
 		implemented in the Linux kernel. 
-		We also detail the design the simulators in order to simulate I/O activities 
-		and the implementation in Python and SimGrid framework \tristan{check grammar}. 
-		Finally, we describe experimental scenarios to evaluate our model using real applications and
-		a Python simulator prototype, the current SimGrid simulator, our extension of SimGrid, and the results from real pipelines.
+		We also detail the design of the simulators in order to simulate 
+		I/O activities and the implementation in Python and SimGrid framework. 
+		Finally, we describe experimental scenarios to evaluate our model using 
+		real applications and a Python simulator prototype, the current SimGrid 
+		simulator, our extension of SimGrid, and the results from real pipelines.
 
@@
 
-			We model memory with a capacity, a read bandwidth and a write bandwidth. 
-			When data is read from or written to memory, the actual 
-			read/write time is calculated with the amount of data and memory 
-			read/write bandwidth:
-
-			\begin{align*}
-				& t_{mr} = D / r_m \\ 
-				& t_{mw} = D / w_m\
-			\end{align*}
-
-			where:
-			\begin{itemize}
-				\item $t_{mr}$ is the time to read data from memory
-				\item $t_{mw}$ is the time to write data to memory
-				\item $D$ is the amount of data to read/write
-				\item $r_m$ is the memory read bandwidth
-				\item $w_m$ is the  memory write bandwidth
-			\end{itemize}
-
-			\tristan{how about parallel accesses to memory? I think it would be a motivation to simualte memory as a storage device, 
-			and leverage existing bandwidth sharing mechanisms in SimGrid. You could updated Fig 1 to include memory in the storage devices, and 
-			explain in your Python simulator implementation that only a simplistic model of storage device is used (for both disk and memory).}
+			Memory can be characterized with a capacity, read bandwidth and write 
+			bandwidth. However, memory bandwidth is always shared between 
+			concurrent processes and implementing a simulation model of concurrent 
+			accesses to memory is non-trivial. Thus, we leverage existing storage 
+			models, which handle parallel storage accesses, in some simulation 
+			frameworks to simulate I/O time in memory.
 
@@
 			two blocks, one is flushed and one remains dirty.
-			The flushing time is calculated as:
-
-			\begin{align*}
-				t_f = A / w_d
-			\end{align*}
-
-			where:
-			\begin{itemize}
-				\item $t_f$ is the flushing time
-				\item $A$ is the amount of data to be flushed
-				\item $w_d$ is the disk write bandwidth
-			\end{itemize}
+			The flushing time is simulated with storage model.
 
@@
 			to current simulated time is longer than the predefined expire time. 
-			The flushing time is calculated by storage simulation model with 
-			total amount of flushed data.
+			Periodical flushing can be concurrent with cache read/write, 
+			and sequential with disk read/write. When cache accesses and 
+			periodical flushing are concurrent, flushing time is not added to simulated 
+			time but the amount of flushed data is limited by flushing time. 
+			When data is flushed in parallel with disk read/write, flushing time is 
+			calculated by storage simulation model and total amount of data 
+			to be flushed.
 
@@
 					\If {mm.cached(fn) $>$ 0}  \Comment{Read file cached data}
-    					\State mm.cache\_read(fn)  \Comment{Update blocks in LRU lists}
-    					\State pft = mm.periodic\_flush(rt) \Comment{Periodical flushing time}
-						\State rt = rt + max(mm.cached(fn) / rm, pft) \Comment{Add the longer}
+    					\State crt = mm.cache\_read(fn)  \Comment{Read data in cache}
+    					\State mm.periodic\_flush(rt, crt) \Comment{Flush in duration = crt}
+						\State rt = rt + crt
 					\EndIf
 					\If {mm.cached(fn) $<$ fs} \Comment{Read uncached data from disk}
-						\State rt = rt + mm.periodic\_flush(rt) \Comment{Another flushing interval}
+						\State rt = rt + mm.periodic\_flush(rt) \Comment{Sequential flushing}
 						\State mm.disk\_read(fs - mm.cached(fn), fn)
@@
 			called to re-access and update this cached data (line 13). 
-			Then, periodical flushing is called to flushed expired dirty data 
-			from cache (line 14).
-			Since periodical flushing and cache read can be concurrent, the time of 
-			the longer activity between them is added to simulated run time (line 15). 
+			During cache read, periodical flushing is called to flush dirty data if needed 
+			and limited to cache read time (line 14).
+			Cache read time is then added to total simulated time (line 15). 
 			If file is not cached or partially cached, a disk read is required (line 17). 
-			Periodical flushing is called once again, but this time the flushing time is 
-			added to total simulated time since flushing and disk read are time-shared, 
-			or sequential in other words (line 18).
+			Periodical flushing is called once again, but not time-limited since 
+			data flushing and disk read are sequential, and the flushing time is 
+			added to simulated time (line 18).
 			Finally, data is read from disk to cache and managed by MemoryManager, 
@@
     					\State mm.evict(mem\_amt - mm.free\_mem)
-    					\State mm.write(fn, mem\_amt)
+    					\State mm.periodic\_flush(rt, mm.write\_time(mem\_amt)) 
+    					\State rt = rt + mm.write(fn, mem\_amt)
     				\EndIf
@@
 			If this amount is greater than 0, MemoryManager is asked to evict 
-			data from cache if needed and to write this amount of data to cache 
-			(line 14-15).
-			The simulated run time can be added with the longer time of cache write 
-			time and periodical flushing time because these can happen 
-			concurrently (line 17).
+			data from cache if needed (line 14).
+			Next, periodical flushing flushes data during cache write because they 
+			can be concurrent (line 15). Then, data is written to cache with memory 
+			bandwidth, write time is simulated and added to total simulated 
+			time (line 16).
 			The amount of data that is not written with memory bandwidth is written 
@@
 			data from cache as much as possible so that 
-			the remaining data can be written to cache (line 19-20). 
+			the remaining data can be written to cache (line 20-21). 
 			Here, MemoryManager does not guarantee that the amount of flushed 
@@
 			of remaining data that is written and kept in cache after the write 
-			is limited by the amount free memory (line 21). 
+			is limited by the amount free memory (line 22). 
 			Finally, the write time is simulated, added to the total time and 
-			returned to simulated applications (line 22).
+			returned to simulated applications (line 23).
 
@@
 			model in a simple use case before integrating it to SimGrid. 
+			Since the simulator does not use any storage model, we simulate the 
+			I/O time in memory as follow:
+
+			\begin{align*}
+				& t_{mr} = D / r_m \\ 
+				& t_{mw} = D / w_m\
+			\end{align*}
+
+			where:
+			\begin{itemize}
+				\item $t_{mr}$ is the time to read data from memory
+				\item $t_{mw}$ is the time to write data to memory
+				\item $D$ is the amount of data to read/write
+				\item $r_m$ is the memory read bandwidth
+				\item $w_m$ is the  memory write bandwidth
+			\end{itemize}
+
 			This simulator simulates a single-threaded pipeline running on a system 

### 2020-07-01 exp/pysim/app.py
add python simulator

@@ -1,0 +1,75 @@
+# i/o simulator in python
+import plot
+import csv
+from components import IOManager
+from components import File
+from components import Storage
+from components import MemoryManager
+
+
+def export_mem(mem_log, filename):
+    with open(filename, 'w', newline='') as csvfile:
+        writer = csv.writer(csvfile)
+        writer.writerow(["time", "total_mem", "dirty", "cache", "used_mem"])
+        for i in range(len(mm.get_log()["time"])):
+            writer.writerow([mem_log["time"][i], mem_log["total"][i], mem_log["dirty"][i],
+                             mem_log["cache"][i], mem_log["used"][i]])
+
+
+def export_time(task_list, filename):
+    with open(filename, 'w', newline='') as csvfile:
+        writer = csv.writer(csvfile)
+        writer.writerow(["type", "start", "end"])
+        for i in range(len(task_list)):
+            writer.writerow([task_list[i][0], task_list[i][1], task_list[i][2]])
+
+
+mm = MemoryManager(268600, 268600, read_bw=7100, write_bw=3300)
+storage = Storage(450000, read_bw=465, write_bw=465)
+kernel = IOManager(mm, storage, dirty_ratio=0.4)
+
+input_size = 20000
+compute_time = 28
+
+file1 = File("file1", input_size, input_size)
+file2 = File("file2", input_size, input_size)
+file3 = File("file3", input_size, input_size)
+file4 = File("file4", input_size, input_size)
+
+start_time = 0
+
+task1_read_end = kernel.read(file1, start_time)
+task1_compute_end = kernel.compute(task1_read_end, compute_time)
+task1_write_end = kernel.write(file2, task1_compute_end)
+kernel.release(file2)
+
+task2_read_end = kernel.read(file2, task1_write_end)
+task2_compute_end = kernel.compute(task2_read_end, compute_time)
+task2_write_end = kernel.write(file3, task2_compute_end)
+kernel.release(file3)
+
+task3_read_end = kernel.read(file3, task2_write_end)
+task3_compute_end = kernel.compute(task3_read_end, compute_time)
+task3_write_end = kernel.write(file4, task3_compute_end)
+kernel.release(file4)
+
+task_time = {
+    "read_start": [start_time, task1_write_end, task2_write_end],
+    "read_end": [task1_read_end, task2_read_end, task3_read_end],
[…]

### 2020-07-01 exp/pysim/components.py

@@ -1,0 +1,537 @@
+class File:
+    """
+    File class.
+    """
+
+    def __init__(self, name, size=0, disk=0, cache=0, dirty=0, active=0, inactive=0):
+        """
+        :param name: filename
+        :param size: size in MB
+        :param disk: data on disk in MB
+        :param cache: data in cache in MB
+        :param dirty: dirty data in MB
+        :param active: data in active list in MB
+        :param inactive: data in inactive list in MB
+        """
+        self.name = name
+        self.size = size
+        # self.disk = disk
+        # self.dirty = dirty
+        # self.active = active
+        # self.inactive = inactive
+
+
+class Block:
+    """Data block of a File"""
+
+    def __init__(self, filename, size=0, dirty=False, last_access=0.0):
+        self.filename = filename
+        self.size = size
+        self.dirty = dirty
+        self.last_access = last_access
+
+
+class MemoryManager:
+    def __init__(self, size=0, free=0, cache=0, dirty=0, read_bw=0, write_bw=0, dirty_expire=30):
+        """
+        LRU list: list of tuples, the first value is filename, the 2nd value is timestamp, the 3rd value amount of data
+
+        :param size: total memory in MB
+        :param free: free memory in MB
+        :param cache: cache used in MB
+        :param dirty: dirty data in MB
+        :param read_bw: read bandwidth in MBps
+        :param write_bw: write bandwidth in MBps
+        """
+        self.size = size
+        self.free = free
+        self.cache = cache
+        self.dirty = dirty
+        self.read_bw = read_bw
+        self.active = []
+        self.inactive = []
+        self.write_bw = write_bw
+        self.dirty_expire = dirty_expire
+        # self.active_list = []
+        # self.inactive_list = []
+        self.log = {
+            "total": [self.size],
+            "free": [self.free],
+            "cache": [self.cache],
+            "dirty": [self.dirty],
+            "used": [self.size - self.free],
+            "time": [0]
+        }
+
+    def get_data_in_cache(self, filename):
+        """
+        Return the amount of data cached
+        :param filename:
+        :return:
+        """
+        amount = 0
+        for block in self.inactive:
+            if block.filename == filename:
+                amount += block.size
+
+        for block in self.active:
+            if block.filename == filename:
+                amount += block.size
+
+        return amount
+
+    def get_available_memory(self):
+        return self.free + self.cache - self.dirty
+
+    def get_evictable_memory(self):
[…]

### 2020-07-01 exp/pysim/evaluate.py

@@ -1,0 +1,141 @@
+import log_parse
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+def get_atop_mem_prop(time_log, mem_log, key):
+    """
+    :param time_log: real time log tuple
+    :param mem_log: real mem log dictionary
+    :param key: key of the memory property to be returned
+    :return: list of values of mem properties corresponding to task end time in log
+    """
+
+    task_ends = [task[2] for task in time_log]
+    start = time_log[0][1]
+
+    arr = mem_log[key]
+    result = []
+    for i in range(len(task_ends)):
+        time = task_ends[i] - start
+        prop = arr[int(time)] + (arr[int(time) + 1] - arr[int(time)]) * (time - int(time))
+        result.append(prop)
+
+    return result
+
+
+def get_sim_mem_prop(time_log, mem_log, key):
+    """
+    :param time_log: simulation time log tuple
+    :param mem_log: simulation mem log dictionary
+    :param key: key of the memory property to be returned
+    :return: list of values of mem properties corresponding to task end time in log
+    """
+
+    task_ends = [task[2] for task in time_log]
+    result = []
+    for i in range(len(task_ends)):
+        idx = mem_log["time"].index(task_ends[i])
+        if idx >= 0:
+            result.append(mem_log[key][idx])
+        else:
+            result.append(-1000)
+
+    return result
+
+
+def task_time_error(realtime_logfile, simtime_logfile):
+    real_time = log_parse.read_timelog(realtime_logfile, skip_header=False)
+    sim_time = log_parse.read_timelog(simtime_logfile)
+
+    time_acc = []
+
+    for i in range(len(sim_time)):
+        sim = sim_time[i][2] - sim_time[i][1]
+        real = real_time[i][2] - real_time[i][1]
+        time_acc.append(abs(sim - real) / real)
+
+    return time_acc
+
+
+def mem_error(real_time_logfile, sim_time_logfile, real_mem_logfile, sim_mem_logfile):
+    real_time_log = log_parse.read_timelog(real_time_logfile, skip_header=False)
+    real_mem_log = log_parse.read_atop_log(real_mem_logfile)
+
[…]

### 2020-07-01 exp/pysim/log_parse.py

@@ -1,0 +1,70 @@
+import csv
+
+
+def read_timelog(filename, skip_header=True):
+    result = []
+
+    with open(filename) as csv_file:
+        csv_reader = csv.reader(csv_file, delimiter=',')
+        if skip_header:
+            next(csv_reader)
+
+        for line in csv_reader:
+            result.append((line[0], float(line[1]), float(line[2])))
+
+    return result
+
+
+def read_atop_log(filename):
+    sys_mem = []
+    free_mem = []
+    used_mem = []
+    cache_used = []
+    dirty_data = []
+
+    f = open(filename)
+    lines = f.readlines()
+    for i in range(len(lines)):
+        line = lines[i]
+        if line.startswith("MEM"):
+            values = line.split(" ")
+
+            sys_mem_mb = int(values[7]) * 4096 / 1000 ** 2
+            sys_mem.append(sys_mem_mb)
+
+            cache_in_mb = int(values[9]) * 4096 / 1000 ** 2
+            cache_used.append(cache_in_mb)
+
+            dirty_amt_mb = int(values[12]) * 4096 / 1000 ** 2
+            dirty_data.append(dirty_amt_mb)
+
+    return {
+        "total": sys_mem,
+        "used_mem": used_mem,
+        "cache": cache_used,
+        "dirty_data": dirty_data
+    }
+
+
+def read_sim_log(filename):
+    time = []
+    sys_mem = []
+    cache_used = []
+    dirty_data = []
+
+    with open(filename) as csv_file:
+        csv_reader = csv.reader(csv_file, delimiter=',')
+        next(csv_reader)
+
+        for line in csv_reader:
+            time.append(float(line[0]))
+            sys_mem.append(float(line[1]))
+            dirty_data.append(float(line[2]))
+            cache_used.append(float(line[3]))
+
+    return {
+        "time": time,
+        "total": sys_mem,
+        "dirty_data": dirty_data,
+        "cache": cache_used
+    }

### 2020-07-01 exp/pysim/plot.py

@@ -1,0 +1,49 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+def plot_mem_log(mem_log, time_stamps, text, xmin, xmax, ymin, ymax):
+    time = mem_log["time"]
+    total = mem_log["total"]
+    free = mem_log["free"]
+    used = mem_log["used"]
+    cache = mem_log["cache"]
+    dirty = mem_log["dirty"]
+    available = list(np.array(free) + np.array(cache) - np.array(dirty))
+    dirty_ratio = list(np.array(available) * 0.4)
+    dirty_bg_ratio = list(np.array(available) * 0.1)
+
+    read_start = time_stamps["read_start"]
+    read_end = time_stamps["read_end"]
+    write_start = time_stamps["write_start"]
+    write_end = time_stamps["write_end"]
+    start = read_start[0]
+
+    plt.figure()
+    plt.title("python simulator result")
+
+    for idx in range(len(read_start)):
+        if idx == 0:
+            plt.axvspan(xmin=read_end[idx] - start, xmax=write_start[idx] - start, color="k",
+                        alpha=0.2, label="computation")
+            plt.axvspan(xmin=0, xmax=read_end[idx] - start, color="g", alpha=0.2, label="read")
+            plt.axvspan(xmin=write_start[idx] - start, xmax=write_end[idx] - start, color="b", alpha=0.2, label="write")
+        else:
+            plt.axvspan(xmin=read_end[idx] - start, xmax=write_start[idx] - start, color="k", alpha=0.2)
+            plt.axvspan(xmin=read_start[idx] - start, xmax=read_end[idx] - start, color="g", alpha=0.2)
+            plt.axvspan(xmin=write_start[idx] - start, xmax=write_end[idx] - start, color="b", alpha=0.2)
+
+    plt.plot(time, total, color='k', linewidth=1, linestyle="-.", label="total mem")
+    plt.plot(time, used, color='g', linewidth=1, label="used mem")
+    plt.plot(time, cache, color='m', linewidth=1, label="cache")
+    plt.plot(time, dirty, color='r', linewidth=1, label="dirty")
+    plt.plot(time, available, color='b', linewidth=1, linestyle="-.", label="available mem")
+    plt.plot(time, dirty_ratio, color='k', linewidth=1, linestyle="-.", label="dirty_ratio")
[…]

### 2020-07-01 exp/pysim/log_parse.py
add code to single thread sim errors

@@ -1,70 +1,0 @@
-import csv
-
-
-def read_timelog(filename, skip_header=True):
-    result = []
-
-    with open(filename) as csv_file:
-        csv_reader = csv.reader(csv_file, delimiter=',')
-        if skip_header:
-            next(csv_reader)
-
-        for line in csv_reader:
-            result.append((line[0], float(line[1]), float(line[2])))
-
-    return result
-
-
-def read_atop_log(filename):
-    sys_mem = []
-    free_mem = []
-    used_mem = []
-    cache_used = []
-    dirty_data = []
-
-    f = open(filename)
-    lines = f.readlines()
-    for i in range(len(lines)):
-        line = lines[i]
-        if line.startswith("MEM"):
-            values = line.split(" ")
-
-            sys_mem_mb = int(values[7]) * 4096 / 1000 ** 2
-            sys_mem.append(sys_mem_mb)
-
-            cache_in_mb = int(values[9]) * 4096 / 1000 ** 2
-            cache_used.append(cache_in_mb)
-
-            dirty_amt_mb = int(values[12]) * 4096 / 1000 ** 2
-            dirty_data.append(dirty_amt_mb)
-
-    return {
-        "total": sys_mem,
-        "used_mem": used_mem,
-        "cache": cache_used,
-        "dirty_data": dirty_data
-    }
-
-
-def read_sim_log(filename):
-    time = []
-    sys_mem = []
-    cache_used = []
-    dirty_data = []
-
-    with open(filename) as csv_file:
-        csv_reader = csv.reader(csv_file, delimiter=',')
-        next(csv_reader)
-
-        for line in csv_reader:
-            time.append(float(line[0]))
-            sys_mem.append(float(line[1]))
-            dirty_data.append(float(line[2]))
-            cache_used.append(float(line[3]))
-
-    return {
-        "time": time,
-        "total": sys_mem,
-        "dirty_data": dirty_data,
-        "cache": cache_used
-    }

### 2020-07-01 result/single/parse_result.py

@@ -1,6 +1,72 @@
-import log_parse
-import matplotlib.pyplot as plt
-import numpy as np
+import csv
+
+def read_timelog(filename, skip_header=True):
+    result = []
+
+    with open(filename) as csv_file:
+        csv_reader = csv.reader(csv_file, delimiter=',')
+        if skip_header:
+            next(csv_reader)
+
+        for line in csv_reader:
+            result.append((line[0], float(line[1]), float(line[2])))
+
+    return result
+
+
+def read_atop_log(filename):
+    sys_mem = []
+    free_mem = []
+    used_mem = []
+    cache_used = []
+    dirty_data = []
+
+    f = open(filename)
+    lines = f.readlines()
+    for i in range(len(lines)):
+        line = lines[i]
+        if line.startswith("MEM"):
+            values = line.split(" ")
+
+            sys_mem_mb = int(values[7]) * 4096 / 1000 ** 2
+            sys_mem.append(sys_mem_mb)
+
+            cache_in_mb = int(values[9]) * 4096 / 1000 ** 2
+            cache_used.append(cache_in_mb)
+
+            dirty_amt_mb = int(values[12]) * 4096 / 1000 ** 2
+            dirty_data.append(dirty_amt_mb)
+
+    return {
+        "total": sys_mem,
+        "used_mem": used_mem,
+        "cache": cache_used,
+        "dirty_data": dirty_data
+    }
+
+
+def read_sim_log(filename):
+    time = []
+    sys_mem = []
+    cache_used = []
+    dirty_data = []
+
+    with open(filename) as csv_file:
+        csv_reader = csv.reader(csv_file, delimiter=',')
+        next(csv_reader)
+
+        for line in csv_reader:
+            time.append(float(line[0]))
+            sys_mem.append(float(line[1]))
+            dirty_data.append(float(line[2]))
+            cache_used.append(float(line[3]))
+
+    return {
+        "time": time,
+        "total": sys_mem,
+        "dirty_data": dirty_data,
+        "cache": cache_used
+    }
 
 
 def get_atop_mem_prop(time_log, mem_log, key):
@@ -45,8 +111,8 @@ def get_sim_mem_prop(time_log, mem_log, key):
 
 
 def task_time_error(realtime_logfile, simtime_logfile):
[…]

### 2020-07-01 result/single/plot.py

@@ -1,0 +1,44 @@
+import parse_result
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+def grouped_bar_chart(labels, title, xlabel, ylabel, *argv):
+    x = np.arange(len(labels))  # the label locations
+    width = 0.2  # the width of the bars
+    bars = len(argv)
+
+    fig, ax = plt.subplots()
+    for i in range(len(argv)):
+        arg = argv[i]
+        ax.bar(x + (2 * i + 1 - bars) * width / 2, arg[1], width, label=arg[0])
+
+    ax.set_xlabel(xlabel)
+    ax.set_ylabel(ylabel)
+    ax.set_title(title)
+    ax.set_xticks(x)
+    ax.set_xticklabels(labels)
+    ax.set_ylim(bottom=0, top=8)
+    ax.legend()
+
+    plt.show()
+
+
+def plot_task_error(size):
+    labels = ["read1", "write1", "read2", "write2", "read3", "write3"]
+
+    # atop_file =     "log/cluster/100gb/atop_mem.log"
+    real_time_log = "real/%dgb/timestamps_pipeline.csv" % size
+    sim_py_log = "pysim/%dgb_sim_time.csv" % size
+    simgrid_log = "simgrid_org/timestamp_sim_exp1_%dgb.csv" % size
+
+    py_error = parse_result.task_time_error(real_time_log, sim_py_log)
+    simgrid_error = parse_result.task_time_error(real_time_log, simgrid_log)
+
+    grouped_bar_chart(labels, "Simulation error of tasks with %dGB" % size, "tasks", "error",
+                      ("Python", py_error), ("Original SimGrid", simgrid_error))
+
+
+sizes = [20, 50, 75, 100]
+for size in sizes:
+    plot_task_error(size)

### 2020-07-02 exp/real/gen_data_multi.sh
Add real pipeline source and scripts

@@ -1,0 +1,12 @@
+#! /bin/bash
+
+rm -f /disk0/dzung/data/file*
+
+max_pipe=32
+size=3000 # File size, in MB
+data_dir="/disk0/dzung/data/"
+
+for i in $(seq 1 ${max_pipe})
+do
+    dd if=/dev/urandom of=${data_dir}file${i} bs=1MB count=${size}
+done

### 2020-07-02 exp/real/read_write.c

@@ -1,0 +1,75 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <sys/time.h>
+#include <string.h>
+
+int main(int argc, char ** argv)
+{
+
+    if(argc != 4)
+    {
+        printf("usage: read_write <file_in> <file_out> <log_file>\n");
+        return 1;
+    }
+
+    // Parse arguments
+    char * in_file_name = argv[1];
+    char * out_file_name = argv[2];
+    char * log_file_name = argv[3];
+
+    FILE * log_file = fopen(log_file_name, "a");
+
+    struct timeval  start;
+    struct timeval  end;
+
+    gettimeofday(&start, NULL);
+    // Get file size
+    FILE * in_file = fopen(in_file_name, "rb");
+    fseek(in_file, 0, SEEK_END);
+    long fsize = ftell(in_file);
+    fseek(in_file, 0, SEEK_SET);  
+
+    // Read file in buffer
+    char * buff = malloc(fsize + 1);
+    fread(buff, fsize, 1, in_file);
+    fclose(in_file);
+
+    gettimeofday(&end, NULL);
+
+    double read_start_in_mil = (start.tv_sec) * 1000 + (start.tv_usec) / 1000 ;
+    double read_end_in_mil = (end.tv_sec) * 1000 + (end.tv_usec) / 1000 ;
+
+    printf("Read %s in: %lf\n", in_file_name, (read_end_in_mil - read_start_in_mil) / 1000);
+    printf("Avg read bw: %4.2lf MBps\n\n", fsize / ((read_end_in_mil - read_start_in_mil) * 1000));
+
+    // Increment buffer
+    gettimeofday(&start, NULL);
+    long int i;
+    for(i = 0 ; i < fsize ; i++)
+        buff[i]++;
+    gettimeofday(&end, NULL);
+
+    double cpu_start_in_mil = (start.tv_sec) * 1000 + (start.tv_usec) / 1000 ;
+    double cpu_end_in_mil = (end.tv_sec) * 1000 + (end.tv_usec) / 1000 ;
+
+    gettimeofday(&start, NULL);
+//     Write buffer
+    FILE * out_file = fopen(out_file_name, "wb");
+    fwrite(buff, fsize, 1, out_file);
+    fclose(out_file);
+    gettimeofday(&end, NULL);
+
[…]

### 2020-07-02 exp/real/run_multi.sh

@@ -1,0 +1,30 @@
+#! /bin/bash
+
+rm -f export/*.*
+
+#collectl -sCDnfM -omT --dskopts z --cpuopts z -i 1 --sep , -P -f export/collectl --procfilt P p &
+#sleep 1
+#atop -P MEM 1 500 > export/pipeline_mem_c.log &
+
+max_pipe=32
+data_dir="/disk0/dzung/data/"
+log_dir="export/"
+
+for no_pipe in $(seq 1 ${max_pipe})
+do
+    rm /disk0/dzung/data/file*_*
+    echo "No of concurrent pipelines: ${no_pipe}."
+    echo "Wait for dirty data to be flushed..."
+    sleep 300
+    echo 3 | sudo tee /proc/sys/vm/drop_caches
+    echo "Cache cleared..."
+    for pipe_idx in $(seq 1 ${no_pipe})
+    do
+        echo "read_start,read_end,cpu_start,cpu_end,write_start,write_end" \
+            > "export/time_pipeline_${no_pipe}_${pipe_idx}.csv"
+        echo "Start pipeline #${pipe_idx} ..."
+        ./pipeline.sh "${data_dir}file${pipe_idx}" "${log_dir}time_pipeline_${no_pipe}_${pipe_idx}.csv" &
+    done
+    wait
+done
+

### 2020-07-02 exp/real/run_single.sh

@@ -1,0 +1,20 @@
+#! /bin/bash
+
+rm -f export/*.*
+rm -f /disk0/dzung/input/*.*
+rm -f /disk0/dzung/output/*.*
+
+input="data/file1"
+logfile="export/timestamps_pipeline.csv"
+size=20000 #File size is in MB
+logperiod=500 #in seconds
+
+dd if=/dev/urandom of=${input} bs=1MB count=${size}
+
+echo 3 | sudo tee /proc/sys/vm/drop_caches
+
+collectl -sCDnfM -omT --dskopts z --cpuopts z -i 1 --sep , -P -f export/collectl --procfilt P p &
+sleep 1
+atop -P MEM 1 ${logperiod} > export/atop.log &
+
+./pipeline.sh "${input}" "${logfile}"

### 2020-07-02 result/multi/plot_exp2.py

@@ -1,0 +1,94 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import csv
+import json
+
+
+def parse_single_pipeline(filename):
+    """
+    :param filename:
+    :return: tuple (start, end, readtime, writetime)
+    """
+    df = pd.read_csv(filename)
+    readtime = sum(df["read_end"] - df["read_start"])
+    writetime = sum(df["write_end"] - df["write_start"])
+
+    return min(df["read_start"]), max(df["write_end"]), readtime, writetime
+
+
+def aggregate_result(no_pipeline):
+    """
+
+    :param no_pipeline:
+    :return: makespan, total_readtime, total_writetime
+    """
+
+    makespan = 0
+    readtime = 0
+    writetime = 0
+    for i in range(no_pipeline):
+        filename = "export/cluster/multi/time/time_pipeline_%d_%d.csv" % (no_pipeline, i + 1)
+        start, end, read, write = parse_single_pipeline(filename)
+        makespan += end - start
+        readtime += read
+        writetime += write
+
+    return no_pipeline, makespan, readtime, writetime
+
+
+def export_real_results(filename):
+    with open(filename, 'w', newline='') as csvfile:
+        writer = csv.writer(csvfile)
+        writer.writerow(["no_pipeline", "makespan", "readtime", "writetime"])
+        for i in range(32):
+            writer.writerow(list(aggregate_result(i + 1)))
+
+
+def export_simgrid_result(filename):
+    with open(filename, 'w', newline='') as csvfile:
+        writer = csv.writer(csvfile)
+        writer.writerow(["no_pipeline", "makespan", "readtime", "writetime"])
+        for i in range(32):
+            writer.writerow(list(parse_simgrid_result(i + 1)))
+
+
+def parse_simgrid_result(no_pipeline):
+    filename = "simgrid/multi/unified_%d.json" % no_pipeline
+    with open(filename) as json_file:
+        res = json.load(json_file)
+        tasks = res["workflow_execution"]["tasks"]
+
+        makespan = sum([task["whole_task"]["end"] - task["whole_task"]["start"] for task in tasks])
+        read = sum([sum([read["end"] - read["start"] for read in task["read"]]) for task in tasks])
+        write = sum([sum([write["end"] - write["start"] for write in task["write"]]) for task in tasks])
+
+    return no_pipeline, makespan, read, write
+
+
[…]

### 2020-07-02 result/single/evaluate.py

@@ -1,72 +1,4 @@
-import csv
-
-def read_timelog(filename, skip_header=True):
-    result = []
-
-    with open(filename) as csv_file:
-        csv_reader = csv.reader(csv_file, delimiter=',')
-        if skip_header:
-            next(csv_reader)
-
-        for line in csv_reader:
-            result.append((line[0], float(line[1]), float(line[2])))
-
-    return result
-
-
-def read_atop_log(filename):
-    sys_mem = []
-    free_mem = []
-    used_mem = []
-    cache_used = []
-    dirty_data = []
-
-    f = open(filename)
-    lines = f.readlines()
-    for i in range(len(lines)):
-        line = lines[i]
-        if line.startswith("MEM"):
-            values = line.split(" ")
-
-            sys_mem_mb = int(values[7]) * 4096 / 1000 ** 2
-            sys_mem.append(sys_mem_mb)
-
-            cache_in_mb = int(values[9]) * 4096 / 1000 ** 2
-            cache_used.append(cache_in_mb)
-
-            dirty_amt_mb = int(values[12]) * 4096 / 1000 ** 2
-            dirty_data.append(dirty_amt_mb)
-
-    return {
-        "total": sys_mem,
-        "used_mem": used_mem,
-        "cache": cache_used,
-        "dirty_data": dirty_data
-    }
-
-
-def read_sim_log(filename):
-    time = []
-    sys_mem = []
-    cache_used = []
-    dirty_data = []
-
-    with open(filename) as csv_file:
-        csv_reader = csv.reader(csv_file, delimiter=',')
-        next(csv_reader)
-
-        for line in csv_reader:
-            time.append(float(line[0]))
-            sys_mem.append(float(line[1]))
-            dirty_data.append(float(line[2]))
-            cache_used.append(float(line[3]))
-
-    return {
-        "time": time,
-        "total": sys_mem,
-        "dirty_data": dirty_data,
-        "cache": cache_used
-    }
+import log_parse
 
 
 def get_atop_mem_prop(time_log, mem_log, key):
@@ -111,8 +43,8 @@ def get_sim_mem_prop(time_log, mem_log, key):
 
 
 def task_time_error(realtime_logfile, simtime_logfile):
-    real_time = read_timelog(realtime_logfile, skip_header=False)
[…]

### 2020-07-02 result/single/log_parse.py

@@ -1,0 +1,88 @@
+import csv
+
+
+def read_timelog(filename, skip_header=True):
+    result = []
+
+    with open(filename) as csv_file:
+        csv_reader = csv.reader(csv_file, delimiter=',')
+        if skip_header:
+            next(csv_reader)
+
+        for line in csv_reader:
+            result.append((line[0], float(line[1]), float(line[2])))
+
+    return result
+
+
+def read_atop_log(filename, dirty_ratio=0.2, dirty_bg_ratio=0.1):
+    sys_mem = []
+    free_mem = []
+    used_mem = []
+    cache_used = []
+    dirty_data = []
+    avai_mem = []
+    dirty_threshold = []
+    dirty_bg_threshold = []
+
+    f = open(filename)
+    lines = f.readlines()
+    for i in range(len(lines)):
+        line = lines[i]
+        if line.startswith("MEM"):
+            values = line.split(" ")
+
+            sys_mem_mb = int(values[7]) * 4096 / 1000 ** 2
+            sys_mem.append(sys_mem_mb)
+
+            cache_in_mb = int(values[9]) * 4096 / 1000 ** 2
+            cache_used.append(cache_in_mb)
+
+            dirty_amt_mb = int(values[12]) * 4096 / 1000 ** 2
+            dirty_data.append(dirty_amt_mb)
+
+            free_mem_mb = int(values[8]) * 4096 / 1000 ** 2
+            free_mem.append(free_mem_mb)
+
+            used_mem.append(sys_mem_mb - free_mem_mb)
+
+            available_mb = free_mem_mb + cache_in_mb - dirty_amt_mb
+            avai_mem.append(available_mb)
+
+            dirty_threshold.append(dirty_ratio * available_mb)
+            dirty_bg_threshold.append(dirty_bg_ratio * available_mb)
+
+    return {
+        "total": sys_mem,
+        "used_mem": used_mem,
+        "cache": cache_used,
+        "dirty_data": dirty_data,
+        "avai_mem": avai_mem,
+        "free_mem": free_mem,
+        "dirty_ratio": dirty_threshold,
+        "dirty_bg_ratio": dirty_bg_threshold
+    }
+
+
+def read_sim_log(filename):
+    time = []
+    sys_mem = []
+    cache_used = []
+    dirty_data = []
+
+    with open(filename) as csv_file:
+        csv_reader = csv.reader(csv_file, delimiter=',')
+        next(csv_reader)
+
+        for line in csv_reader:
[…]

### 2020-07-02 result/single/plot_error.py

@@ -1,4 +1,4 @@
-import parse_result
+import evaluate
 import matplotlib.pyplot as plt
 import numpy as np
 
@@ -32,8 +32,8 @@ def plot_task_error(size):
     sim_py_log = "pysim/%dgb_sim_time.csv" % size
     simgrid_log = "simgrid_org/timestamp_sim_exp1_%dgb.csv" % size
 
-    py_error = parse_result.task_time_error(real_time_log, sim_py_log)
-    simgrid_error = parse_result.task_time_error(real_time_log, simgrid_log)
+    py_error = evaluate.task_time_error(real_time_log, sim_py_log)
+    simgrid_error = evaluate.task_time_error(real_time_log, simgrid_log)
 
     grouped_bar_chart(labels, "Simulation error of tasks with %dGB" % size, "tasks", "error",
                       ("Python", py_error), ("Original SimGrid", simgrid_error))

### 2020-07-02 result/single/plot_real.py

@@ -1,0 +1,116 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import csv
+import log_parse
+
+
+def timestamp_plot(fig, tasks_time):
+    start = tasks_time[0][1]
+
+    for i in range(len(tasks_time)):
+        if tasks_time[i][0] == "read":
+            fig.axvspan(xmin=tasks_time[i][1] - start, xmax=tasks_time[i][2] - start, color="g", alpha=0.2,
+                        label="read" if i == 0 else "")
+        else:
+            fig.axvspan(xmin=tasks_time[i - 1][2] - start, xmax=tasks_time[i][1] - start, color="k", alpha=0.2,
+                        label="computation" if i == 1 else "")
+            fig.axvspan(xmin=tasks_time[i][1] - start, xmax=tasks_time[i][2] - start, color="b", alpha=0.2,
+                        label="write" if i == 1 else "")
+
+
+def timestamp_readonly_plot(fig, time_stamps):
+    read_start = time_stamps["read_start"]
+    read_end = time_stamps["read_end"]
+    start = read_start[0]
+
+    for idx in range(len(read_start)):
+        fig.axvspan(xmin=read_start[idx] - start, xmax=read_end[idx] - start, color="g",
+                    alpha=0.2)
+        if idx < len(read_start) - 1:
+            fig.axvspan(xmin=read_end[idx] - start, xmax=read_start[idx + 1] - start, color="k", alpha=0.2)
+
+
+def mem_plot(fig, atoplog, time_stamp, input_size, readonly=False):
+    dirty_data = np.array(atoplog["total"])
+    intervals = len(dirty_data)
+    time = np.arange(0, intervals)
+
+    fig.minorticks_on()
+    fig.set_title("memory profiling (input size = %s MB)" % input_size)
+
+    if time_stamp is not None:
+        if readonly:
+            timestamp_readonly_plot(fig, time_stamp)
+        else:
+            timestamp_plot(fig, time_stamp)
+
+    # app_cache = list(np.array(app_mem) + np.array(cache_used))
+
+    fig.plot(time, atoplog["total"], color='k', linewidth=1, linestyle=":", label="total mem")
+    # ax1.plot(time, free_mem, color='g', linewidth=1.5, linestyle="-.", label="free memory")
+    fig.plot(time, atoplog["used_mem"], color='g', linewidth=1.5, label="used mem")
[…]

### 2020-07-03 generate_figures.sh
Fixed inkscape params

@@ -12,7 +12,7 @@ type inkscape &>/dev/null || die "Cannot find inkscape"
 for i in figures/*.svg
 do
     name=$(basename $i .svg)
-    command="inkscape $i --export-pdf=figures/${name}.pdf"
+    command="inkscape $i -o figures/${name}.pdf"
     echo ${command} 
     ${command} || die "Failed"
 done

### 2020-07-03 paper_io_sim.tex
edits and comments in methods

@@
 				\item Current implementation in Linux and some reasons why it is chosen to be implemented (implementation complexity, effectiveness, overhead, etc) \cite{linuxdev3rd2010}
+				\item \tristan{LRU lists are mentioned in the next section, they should be described here.}
 			\end{itemize}
@@
 				\item Discuss the pros of SimGrid and the reasons why we chose it to extend. (Section 2.2.2 in \cite{casanova2014})
+				\item \tristan{Mention the status of storage simulation in simulators, so that we understand from Fig 1 that storage devices are available}
 			\end{itemize}
@@
 		In this section, we present our approach to model file read/write,
-		I/O in memory, data flushing and cache eviction mechanisms 
-		implemented in the Linux kernel. 
-		We also detail the design of the simulators in order to simulate 
-		I/O activities and the implementation in Python and SimGrid framework. 
-		Finally, we describe experimental scenarios to evaluate our model using 
-		real applications and a Python simulator prototype, the current SimGrid 
-		simulator, our extension of SimGrid, and the results from real pipelines.
+		memory I/O, cache eviction and data flushing mechanisms implemented in
+		the Linux kernel. We also detail the design of our simulators and their
+		implementation. Finally, we describe experimental scenarios to evaluate
+		our simulation model on real applications using a standalone Python
+		prototype, the current SimGrid simulator, and our extension of SimGrid.
 
@@
 
-			Our simulation model simulates the main mechanisms of the
-			write-back page cache including memory I/O, cache eviction and
-			data flushing. We separate our model in two components, the IO
-			Controller and the Memory Manager, which together simulate I/O
-			in the Linux kernel (Figure~\ref{fig:interaction}). The Memory
-			Manager simulates memory I/O, flushing, and cache eviction, whereas the
-			IO Controller simulates file read and file write.
-			Simulated applications can simply request
-			file reads and file writes from the IO Manager
-			instead of calling simulated storage directly.
+            We separate our simulation model in two components, the IO
+			Controller and the Memory Manager, which together simulate Linux
+			kernel I/Os (Figure~\ref{fig:interaction}). The Memory Manager
+			simulates memory I/O, flushing, and cache eviction, whereas the IO
+			Controller simulates file reads and writes. Simulated applications
+			can simply interact with the IO Manager instead of calling simulated
+			storage directly as is the case with current simulators.
 
@@
 
-			Memory can be characterized with a capacity, read bandwidth and write 
-			bandwidth. However, memory bandwidth is always shared between 
-			concurrent processes and implementing a simulation model of concurrent 
-			accesses to memory is non-trivial. Thus, we leverage existing storage 
-			models, which handle parallel storage accesses, in some simulation 
-			frameworks to simulate I/O time in memory.
+			We model memory as a storage device, characterized with a capacity,
+			a read bandwidth and a write bandwidth in most simulators
+			\tristan{are the read and write bandwidths distinguished in simgrid?
+			we should also talk about latency}. We refer to existing simulation
+			models for storage devices, to simulate bandwidth sharing for
+			parallel accesses \tristan{we should refer to related work, see
+			comment there}.
 
-			We introduce the concept of data block as a unit to represent cached data. 
-			A data block is a subset of pages of a file in page cache that is accessed 
-			at a specific time. Therefore, a block has the information about file name, 
-			block size, last access and a dirty flag with values of 0 and 1, 
-			0 represents block of lean data, while 1 indicates that the data block 
-			is dirty. 
-			Cached data of a file can be represented by multiple blocks. 
-			Because a data block represents a subset of cached pages, 
-			it can be split into smaller blocks, and multiple blocks can also 
-			be merged into a larger one.
+			We introduce the concept of a data block as a unit to represent data
+			cached in memory. A data block is a subset of file pages stored in
+			page cache that were accessed in the same read or write
+			operation. A data block
+			has information about file name, block size, last access time, and a
+			dirty flag that represents whether the data is clean (0) or dirty
+			(1). A given file can have multiple data blocks in page cache. In
+			addition, a data block can be split into an arbitrary number of
+			smaller blocks, and data blocks can be merged together.
+			\tristan{I reformulated this paragraph, please check}
 
-			In Linux kernel, the page cache LRU lists are lists of pages of 
-			files in cache. In our simulator, instead of using lists of pages, 
-			we simulate these lists with two lists of data blocks, an active list and 
-			an inactive list, ordered by block last access as is shown in 
-			Figure~\ref{fig:lrulist}.
-			As in the kernel, in our simulator, we keep the size of the active list not 
-			larger than twice the size of the inactive list by moving least recently 
-			used blocks from the active list to the inactive list.
+			We simulate the kernel LRU lists of file pages
+			with two lists of data blocks: an active list and 
+			an inactive list, ordered by last access time (earliest first, see Figure~\ref{fig:lrulist}).
+			As in the kernel, our simulator keeps the size of the active list under
+			twice the size of the inactive list by moving least recently 
+			used data blocks from the active list to the inactive list. 
+			Modeling page cache LRU lists as lists of data blocks reduces the
+			overhead of the simulator, as data blocks are obviously less
+			numerous than file pages, while preserving the accuracy in I/O time
+			simulation \tristan{how do you know that, can you provide an argument?}.
 
@@
 
-			Modeling page cache LRU lists as lists of blocks offers our simulator 
-			some obvious advantages. 
-			Maintaining and handling lists of pages in the simulator, which is in 
-			packet level, are non-trivial when they require significant implementation 
-			overhead and would likely to hinder the performance of the simulator. 
-			Therefore, using page cache lists of data blocks can simplify the 
-			simulation of cache eviction and data flushing while preserving
-			the accuracy in i/o time simulation.
-
-			With the simulation model of page cache LRU lists of blocks, now we can 
-			model I/O in memory.
-			At the time when a file is accessed, some data of the file can be in 
-			cache or not. 
-			If accessed data is already cached, there are existing data blocks 
+			\tristan{is this paragraph about read accesses or write accesses? I
+			think you should make one for reads, and one for writes. The
+			paragraph could start as follows: ``The simulated LRU lists are used
+			as follows. When a file is accessed, \ldots ''} At the time when a
+			file is accessed, some data of the file can be in cache or not. If
+			accessed data is already cached, there are existing data blocks
 			representing cached data in two page cache LRU lists. 
@@
 			The data flushing simulation function traverses the sorted inactive first, 
-			and then the active list, looks for least recently used dirty blocks and 
+			and then the active list, looks for least recently used dirty blocks \tristan{, writes them to disk through the storage device model,} and 
 			sets dirty flags of these blocks to 0 until the total flushed amount 
@@
 			If the last flushed block is not entirely flushed, it is split into 
-			two blocks, one is flushed and one remains dirty.
+			two blocks, one that is flushed and one that remains dirty.
 			The flushing time is simulated with storage model.
@@
 			calculated by storage simulation model and total amount of data 
-			to be flushed.
+			to be flushed \tristan{This concurrency should be handled by the storage model, not the memory manager described here. 
+			Rephrase the last sentences to explain this.}.
 
@@
 
+            \tristan{very good writing Dzung, it reads very well now.}
+
 			\subsubsection{File I/O simulation model}
@@
 			Simulated applications can simply request for a file read 
-			or file write by calling the APIs of this layer instead of 
-			calling directly to storage simulation models.
-			Because every computer has its own memory and system configurations, 
-			one IOController object and one MemoryManager object are created 
-			for each simulated machine.
+			or a file write by calling the IO Controller instead of 
+			the storage simulation models. 
+			\tristan{this should move, not sure where for now: Because every computer has its own memory and system configurations, 
+			one IO Controller object and one Memory Manager object are created 
+			for each simulated machine.}
 
 			To read or write a file, a simulated application sends a request to the 
-			IOController instance on the node where the requested file is stored.
-			Based on the memory status in MemoryManager, the IOController 
-			orchestrates cache read/write, disk read/write, flushing and cache eviction 
-			to fulfill I/O requests, calculates total simulated time and returns to 
-			simulated applications.
+			IO Controller instance on the node where the requested file is stored.
+			Based on the memory status in Memory Manager, the IOController 
+			orchestrates cache reads and writes, disk reads and writes, flushing, cache eviction 
+			 and returns to the simulated application.
 
@@
 
-			The algorithm used to simulate file read is described in 
-			Algorithm~\ref{alg:read}. 
-			In Linux systems with write back storage devices, data can be either read 
+			The algorithm to simulate file reads is described in 
+			Algorithm~\ref{alg:read} \tristan{you should now remove the incrementation of simulation time, as the storage models will do that.
+			Essentially, we neglect kernel time and assume that time is entirely spent in IO. It will simplify the algorithms further, which is also elegant. If you 
+			agree, please revise the algorithm and its description accordingly, and do the same for algorithm 2.
+			}.  
+			In Linux systems with write-back storage devices, data can be either read 
 			from cache with memory bandwidth or read from disk with disk bandwidth. 
@@
 			cache read and disk read. 
-			When a user simulated application reads a file, 
-			it sends a file read request to IOController. 
-			First, dirty data is flushed to disk to make more available memory 
-			to read the file if needed.
-			Data flushing function in MemoryManager is called, flushing time is 
-			simulated and added to total run time (line 10).
-			After data is flushed, MemoryManager is asked to evict data 
-			from cache to accommodate the requested file (line 11). 
-			Next, if file is cached either partially or entirely, MemoryManager is 
-			called to re-access and update this cached data (line 13). 
+			When a simulated application reads a file, 
+			it sends a file read request to IO Controller. 
+			First, if there  is not enough memory available to store two
+			copies of the file (one in cache, one in anonymous memory), the
+			Memory Manager is called to flush dirty data to disk (line 10).
+			This flushing is complemented by eviction by the Memory Manager (line 11). 
+			Next, if the file is cached, either partially or entirely, the Memory Manager is 
+			called to read and update the corresponding data blocks in cache (line 13). 
 			During cache read, periodical flushing is called to flush dirty data if needed 
-			and limited to cache read time (line 14).
-			Cache read time is then added to total simulated time (line 15). 
+			and limited to cache read time \tristan{unclear, reword} (line 14).
+			Cache read time is then added to total simulated time (line 15) \tristan{this should be part of the storage model}. 
 			If file is not cached or partially cached, a disk read is required (line 17). 
@@
 
-			To validate our simulation model, we create a simple 
-			simulator which is independent to simulation frameworks and libraries. 
+			To validate our simulation model, we create a simple prototype
+			simulator independently of existing simulation frameworks and libraries. 
 			This enables us to evaluate the accuracy and correctness of our 
-			model in a simple use case before integrating it to SimGrid. 
-			Since the simulator does not use any storage model, we simulate the 
-			I/O time in memory as follow:
-
+			model in a simple scenario before integrating it in the more complex SimGrid environment. 
+			In this prototype we use the following basic storage model, used for both memory and disk: 
 			\begin{align*}
@@
 			\end{itemize}			
+			\tristan{if this applies equally to memory and disk, use ``device'' instead of ``memory''}
 
-			This simulator simulates a single-threaded pipeline running on a system 
-			with a single core CPU, one local storage device and all input and 
-			output files are stored locally. The results from the simulator is compared 
-			and validated with the results from a real pipeline running on
+			Since bandwidth sharing is not simulated, this prototype doesn't support concurrency: it is limited 
+			to single-threaded pipelines running on systems 
+			with a single-core CPU. \tristan{the following aren't resulting from the lack of bandwidth sharing simulation, these are more general limitations, I think we 
+			should remove them: , a single local storage device, and all input and 
+			output files are stored locally.} We use it in a first validation of our simulation model,
+			against a real sequential pipeline running on
 			a real system.
@@
 			using the current version of SimGrid and the SimGrid version that is 
-			extended with our model. We leverage Wrench framework, which is 
-			built on top of SimGrid to simulate our pipelines and system.
-			Then, we compare the results of the simulators with
+			extended with our model. To simulate our pipelines and system, we leverage the Workflow Management System Simulation Workbench (WRENCH~\cite{wrench}), a framework  
+			built over SimGrid.
+		    We compare the results of the simulators with
 			original SimGrid and extended SimGrid with the results of real
-			pipelines on a real system.
+			pipelines on a real system \tristan{this will have to be updated with a description of extended SimGrid, 
+			you should add a placeholder for this description.}.
 
-			In this work, we use Python 3.7 to implement the simple
-			simulator, SimGrid 3.25 and Wrench 1.6 for SimGrid simulators. 
-			SimGrid source code is available at 
-			\url{https://framagit.org/simgrid/simgrid}, and Wrench is provided on
-			github at \url{https://github.com/wrench-project/wrench}.
+			We use Python 3.7 to implement the simple simulator, SimGrid 3.25
+			and Wrench 1.6 for SimGrid simulators. SimGrid source code is
+			available at \url{https://framagit.org/simgrid/simgrid}, and WRENCH
+			is at \url{https://github.com/wrench-project/wrench}.
+			\tristan{You should also add the URL of your Python prototype}
 
@@
 
-			When designing experiments, our goal is to evaluate our simulation model 
-			in single-threaded and multi-threaded applications which access data on 
-			on different file system types. 
-			To do that, we create a pipeline of sequential tasks, each task 
-			reads an input file, increases every byte of the input, and writes the 
-			output to disk. The output of the previous task is the input of 
-			the next task. We also adopt an use case in real applications and 
-			to evaluate the applicability of our simulation model.
+			The goal of our experiments is to evaluate our write-back cache
+			simulation in single-threaded and multi-threaded applications
+			accessing data on different types of file system. We use a pipeline
+			of sequential tasks where each task reads an input file, increments
+			every byte of the input to emulate real processing and avoid caching
+			effects, and writes the result to disk. The output of the previous
+			task is the input of the next task. We also use a real application pipeline, to evaluate the applicability of our
+			simulation model \tristan{add a placeholder to describe the real pipeline here}.
 
-			The simplest scenario is a single-threaded application, where memory 
-			bandwidth and disk bandwidth are not shared with another thread. 
+			The simplest execution scenario is single-threaded, where memory 
+			bandwidth and disk bandwidth are dedicated. 
 			This allows us to compare the detailed results in terms of read/write time, 
-			memory states after every single task and ,more importantly, validate our 
-			simulation model.
+			memory states after every single task and, more importantly, validate our 
+			simulation model \tristan{last sentence is convoluted, reword}.
 
-			For multi-threaded applications, we do the experiment with a varying 
-			number of concurrent pipelines running on a multi-core host 
-			and writing to the same local storage device. 
-		    By doing so, we can assess the performance of our model in the case  
-		    of concurrent I/O with shared bandwidth on a local storage device.
+			For multi-threaded applications, we vary the number of concurrent
+			pipelines running on a multi-core host and writing to the same local
+			disk. In this way, we evaluate our model with concurrent I/Os, when
+			the device bandwidth is shared.
 
@@
 		    shared file system instead of local disk.
-		    In this way, we can not only evaluate our simulation model on I/O 
-		    with shared file system, but also compare the 
-		    performance of the simulator on different storage types.
 
-		    Finally, we adopt a real pipeline in a real neuroimaging applications. 
+		    Finally, we consider a real pipeline from the neuroimaging domain. 
 		    We simulate the pipeline with a simulated workflow in SimGrid 
@@
 			correctly modeled and use the CPU time measured in the real pipelines 
-			to setup our simulated workflows.
+			to setup our simulated workflow \tristan{the pipeline should be described, including the workflow engine if it's Dask.}.
 
-			We use the cluster of \textit{Big Data Infrastructure for Neuroimaging Lab} 
-			at \textit{Concordia University} to conduct our experiments. The cluster 
-			consists of one login node which is the control node, 8 compute nodes 
+			We use our dedicated cluster hosted at Concordia University to conduct the experiments. The cluster 
+			consists of one login node, 8 compute nodes 
 			and 4 storage nodes connected with two network switches. The login node 
@@
 			\textcolor{red}{[network description here]}
+			\tristan{mention the number of disks on all nodes. Maybe add a diagram.}
 
+			\tristan{I don't think you need an intro paragraph for each experiment, as you did above, AND then subsections. I'd just make a single paragraph or sub-section.}
 			\subsubsection{Single-threaded evaluation}
@@
 		\end{itemize}
+	\section{CFI for cluster}
 \bibliographystyle{plain}

### 2020-07-09 paper_io_sim.tex

@@
 
-			We model memory as a storage device, characterized with a capacity,
-			a read bandwidth and a write bandwidth in most simulators
-			\tristan{are the read and write bandwidths distinguished in simgrid?
-			we should also talk about latency}. We refer to existing simulation
-			models for storage devices, to simulate bandwidth sharing for
-			parallel accesses \tristan{we should refer to related work, see
-			comment there}.
+			We leverage existing storage simulation models to simulate the memory,
+			which is characterized with capacity, read and write bandwidths. 
+			This can be done when disks are modeled with distinguished read/write 
+			bandwidths and negligible disk latency, which is applicable to memory 
+			\cite{lebre2015}. In addition, leveraging existing disk models allows us 
+			to take advantage of bandwidth sharing between concurrent disk I/O to 
+			simulate parallel memory accesses.\tristan{we should refer to related work, 
+			see comment there}.
 
@@
 			cached in memory. A data block is a subset of file pages stored in
-			page cache that were accessed in the same read or write
-			operation. A data block
-			has information about file name, block size, last access time, and a
-			dirty flag that represents whether the data is clean (0) or dirty
-			(1). A given file can have multiple data blocks in page cache. In
+			page cache that were accessed in the same read or write operation. 
+			A data block has information about file name, block size, last access 
+			time, and a dirty flag that represents whether the data is clean (0) 
+			or dirty (1). 
+			A given file can have multiple data blocks in page cache. In
 			addition, a data block can be split into an arbitrary number of
 			smaller blocks, and data blocks can be merged together.
-			\tristan{I reformulated this paragraph, please check}
 
@@
 			with two lists of data blocks: an active list and 
-			an inactive list, ordered by last access time (earliest first, see Figure~\ref{fig:lrulist}).
+			an inactive list, ordered by last access time (earliest first, see 
+			Figure~\ref{fig:lrulist}).
 			As in the kernel, our simulator keeps the size of the active list under
@@
 
-			\tristan{is this paragraph about read accesses or write accesses? I
-			think you should make one for reads, and one for writes. The
-			paragraph could start as follows: ``The simulated LRU lists are used
-			as follows. When a file is accessed, \ldots ''} At the time when a
-			file is accessed, some data of the file can be in cache or not. If
-			accessed data is already cached, there are existing data blocks
-			representing cached data in two page cache LRU lists. 
-			As cached data is accessed again, these blocks are moved to 
+			The simulated LRU lists are used as follows. When a file is read, 
+			file data can be partially cached, completely cached, or uncached. 
+			For cached data, there are existing data blocks representing 
+			this data in two page cache LRU lists. 
+			As this cached data is accessed again, these blocks are moved to 
 			the active list, dirty blocks are merged into a bigger dirty block, 
 			while clean data blocks are merged into another clean block. 
-			If accessed data is not in cache, new blocks representing uncached data 
-			are created and put into the top of the inactive list. 
-			For more detail when uncached data is accessed, in the case of 
-			reading file, as all uncached data is clean data, only one clean data 
-			block is added to the inactive list.
-			But in the case of writing uncached file data though cache, 
-			this written data can be either dirty or clean. 
-			Thus, at most two data blocks can be added to the inactive lists, 
-			one is clean and one is dirty. 
+			If data read is not in cache, a clean block representing uncached data 
+			is created and put into the top of the inactive list.
+
+			In the case of writing data, we can assume that all written data 
+			is uncached. After data is written, some can remain dirty, some is 
+			clean after being flushed. 
+			Thus, at most two data blocks can be added, one is clean and one 
+			is dirty, to the inactive list.
+
 			We also keep the size of the active list not larger than twice the size of 
-			the inactive list as in the kernel. This is done by moving least recently 
-			used blocks in the active list to the inactive list when the active list 
-			grows. If only a part of a data block needs to be moved, 
-			the block is split.
+			the inactive list as in the kernel 
+			\cite{gorman2004understanding, linuxdev3rd2010}. 
+			This is done by moving least recently used blocks in the active list 
+			to the inactive list when the active list grows. If only a part of a data 
+			block needs to be moved, the block is split.
 
@@
 			The data flushing simulation function traverses the sorted inactive first, 
-			and then the active list, looks for least recently used dirty blocks \tristan{, writes them to disk through the storage device model,} and 
+			and then the active list, looks for least recently used dirty blocks, 
+			writes them to disk through the storage device model, and 
 			sets dirty flags of these blocks to 0 until the total flushed amount 
@@
 			time but the amount of flushed data is limited by flushing time. 
-			When data is flushed in parallel with disk read/write, flushing time is 
-			calculated by storage simulation model and total amount of data 
-			to be flushed \tristan{This concurrency should be handled by the storage model, not the memory manager described here. 
-			Rephrase the last sentences to explain this.}.
+			When data is flushed in parallel with disk accesses, the concurrency 
+			is handled by the storage model used as memory, and flushing time 
+			is simulated with the amount of flushed data by this model.
 
@@
 			since cache eviction time is negligible in real systems.
-
-            \tristan{very good writing Dzung, it reads very well now.}
 
@@
         				\Desc{fs}{file size (assumed to fit in memory)}
-						\Desc{rt}{current simulated run time}
 						\Desc{mm}{MemoryManager object}
-						\Desc{rm}{memory read bandwidth}
 						\Desc{sm}{storage simulation model}
    					\EndInput
-   					\Output
-						\Desc{rt}{simulated run time after the file is read}
-   					\EndOutput
-					\State rt = rt + mm.flush(2*fs - mm.cached(fn) - mm.evictable\_mem) 
+					\State mm.flush(2*fs - mm.cached(fn) - mm.evictable\_mem) 
 					\State mm.evict(2*fs - mm.cached(fn) - mm.free\_mem) 
 					\If {mm.cached(fn) $>$ 0}  \Comment{Read file cached data}
-    					\State crt = mm.cache\_read(fn)  \Comment{Read data in cache}
-    					\State mm.periodic\_flush(rt, crt) \Comment{Flush in duration = crt}
-						\State rt = rt + crt
+					    \State mm.periodic\_flush() \Comment{Flush during cache read}
+    					\State mm.cache\_read(fn) 
 					\EndIf
 					\If {mm.cached(fn) $<$ fs} \Comment{Read uncached data from disk}
-						\State rt = rt + mm.periodic\_flush(rt) \Comment{Sequential flushing}
-						\State mm.disk\_read(fs - mm.cached(fn), fn)
-    					\State rt = rt + sm.read(fs - mm.cached(fn))
+						\State mm.periodic\_flush() \Comment{Sequential with disk read}
+						\State mm.add\_to\_cache(fs - mm.cached(fn), fn)
+    					\State sm.read(fs - mm.cached(fn))
 					\EndIf
-
-					\State \Return rt
 
@@
 			The algorithm to simulate file reads is described in 
-			Algorithm~\ref{alg:read} \tristan{you should now remove the incrementation of simulation time, as the storage models will do that.
-			Essentially, we neglect kernel time and assume that time is entirely spent in IO. It will simplify the algorithms further, which is also elegant. If you 
-			agree, please revise the algorithm and its description accordingly, and do the same for algorithm 2.
-			}.  
-			In Linux systems with write-back storage devices, data can be either read 
-			from cache with memory bandwidth or read from disk with disk bandwidth. 
-			Therefore, total read time can be considered as a linear function 
-			of the read time with each of these bandwidths, 
-			enabling us to generalize reading process as two separate phases, 
-			cache read and disk read. 
-			When a simulated application reads a file, 
-			it sends a file read request to IO Controller. 
+			Algorithm~\ref{alg:read}.   
+			In Linux systems, data can be either read from cache with memory bandwidth 
+			or read from disk with disk bandwidth. 
+			Therefore, total read time can be considered as sequential reads with 
+			these different bandwidths, enabling us to generalize reading process as 
+			two separate phases, cache read and disk read. 
+			When a simulated application reads a file, it sends a file read request 
+			to IO Controller. 
 			First, if there is not enough memory available to store two
 			copies of the file (one in cache, one in anonymous memory), the
-			Memory Manager is called to flush dirty data to disk (line 10).
-			This flushing is complemented by eviction by the Memory Manager (line 11). 
-			Next, if the file is cached, either partially or entirely, the Memory Manager is 
-			called to read and update the corresponding data blocks in cache (line 13). 
-			During cache read, periodical flushing is called to flush dirty data if needed 
-			and limited to cache read time \tristan{unclear, reword} (line 14).
-			Cache read time is then added to total simulated time (line 15) \tristan{this should be part of the storage model}. 
-			If file is not cached or partially cached, a disk read is required (line 17). 
-			Periodical flushing is called once again, but not time-limited since 
-			data flushing and disk read are sequential, and the flushing time is 
-			added to simulated time (line 18).
-			Finally, data is read from disk to cache and managed by MemoryManager, 
-			disk read time estimated by the storage model is added to 
-			total simulated run time (line 19-20).
+			Memory Manager is called to flush dirty data to disk and simulate 
+			flushing time with memory model (line 6).
+			This flushing is complemented by eviction by the Memory Manager (line 7). 
+			Next, if the file is cached, either partially or entirely, data is read from 
+			cache first. Periodical flushing is triggered at the beginning of the 
+			cache read to flush dirty data if some dirty data is expired (line 9).
+			At the same time, Memory Manager is called to read data from cache, 
+			and update the corresponding data blocks in cache (line 10). 
+			The concurrency of periodical flushing and cache read is handled and 
+			time is simulated by the memory and disk models.
+
+			If file is not cached or partially cached, a disk read is required (line 12). 
+			Periodical flushing is called again at the beginning of the disk read 
+			(line 13), the storage model simulates the disk write with the remaining 
+			amount of data (line 14), and data read from disk is added to cache 
+			(line 15). Similar to cache read period, disk bandwidth sharing is handled 
+			and time is simulated by storage models.
 
@@
         				\Desc{fs}{file size}
-						\Desc{rt}{current simulated run time}
 						\Desc{mm}{MemoryManager object}
-						\Desc{wm}{memory write bandwidth}
 						\Desc{sm}{storage simulation model}
    					\EndInput
-   					\Output
-						\Desc{rt}{simulated run time after the file is read}
-   					\EndOutput
-   					\State pft = mm.periodic\_flush(rt)
+   					\State mm.periodic\_flush()
 					\State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
@@
     					\State mm.evict(mem\_amt - mm.free\_mem)
-    					\State mm.periodic\_flush(rt, mm.write\_time(mem\_amt)) 
-    					\State rt = rt + mm.write(fn, mem\_amt)
+    					\State mm.write(fn, mem\_amt) 
     				\EndIf
-    				\State rt = rt + max(mem\_amt / wm, pft) \Comment{Concurrent processes}
 					\If {mem\_amt $<$ fs}  \Comment{Write with disk bandwidth}
-						\State mm.flushed(fs - mem\_amt)  \Comment{Concurrent flushing}
+						\State mm.flush(fs - mem\_amt)  \Comment{Concurrent with disk write}
 						\State mm.evict(fs - mem\_amt  - mm.free\_mem) 
-						\State mm.write(fn, min(fs - mem\_amt, mm.free\_mem))
-						\State rt = rt + sm.write(fs - mem\_amt)
+						\State mm.add\_to\_cache(fn, min(fs - mem\_amt, mm.free\_mem))
+						\State sm.write(fs - mem\_amt)
 					\EndIf
-
-					\Return rt
 
@@
 			Algorithm~\ref{alg:write} describes our simulation of file write. 
-			Initially, periodical flushing is called to flush expired dirty data 
-			at the beginning of the write (line 10).
+			Initially, periodical flushing is called to asynchronously flush expired 
+			dirty data at the beginning of the write (line 6).
 			Similar to the file read simulation, a file can also be written with 
 			two different bandwidths, memory bandwidth and disk bandwidth. 
-			Files can be written with memory bandwidth before the amount 
+			Data can be written with memory bandwidth before the amount 
 			of dirty data reaches dirty\_ratio. Thus, our algorithm initially checks 
-			the amount of file data that can be written with memory write 
-			bandwidth (line 11-12).
+			the amount of data that can be written with memory write 
+			bandwidth (line 7-8).
 			If this amount is greater than 0, MemoryManager is asked to evict 
-			data from cache if needed (line 14).
-			Next, periodical flushing flushes data during cache write because they 
-			can be concurrent (line 15). Then, data is written to cache with memory 
-			bandwidth, write time is simulated and added to total simulated 
-			time (line 16).
+			data from cache to accommodate the written data if needed (line 10). 
+			Then, cache write is simulated with memory model and data written is 
+			added to page cache LRU lists (11).
 			The amount of data that is not written with memory bandwidth is written 
-			with disk bandwidth (line 18). 
+			with disk bandwidth (line 13). 
 			Now, the dirty ratio threshold is reached, we have to flush and evict 
 			data from cache as much as possible so that 
-			the remaining data can be written to cache (line 20-21). 
+			the remaining data can be written to cache (line 14-15). 
 			Here, MemoryManager does not guarantee that the amount of flushed 
@@
 			of remaining data that is written and kept in cache after the write 
-			is limited by the amount free memory (line 22). 
-			Finally, the write time is simulated, added to the total time and 
-			returned to simulated applications (line 23).
+			is limited by the amount free memory (line 16). 
+			Finally, the disk write is simulated with the remaining amount of 
+			data (line 17). As in the simulation of file read, all the concurrent 
+			cache write and disk write are simulated by memory and disk models.
 
@@
 			\begin{align*}
-				& t_{mr} = D / r_m \\ 
-				& t_{mw} = D / w_m\
+				& t_{r} = D / b_r \\ 
+				& t_{w} = D / b_w\
 			\end{align*}
@@
 			\begin{itemize}
-				\item $t_{mr}$ is the time to read data from memory
-				\item $t_{mw}$ is the time to write data to memory
+				\item $t_{r}$ is the data read time
+				\item $t_{w}$ is the data write time
 				\item $D$ is the amount of data to read/write
-				\item $r_m$ is the memory read bandwidth
-				\item $w_m$ is the  memory write bandwidth
+				\item $b_r$ is the read bandwidth of the device
+				\item $b_w$ is the write bandwidth of the device
 			\end{itemize}
-			\tristan{if this applies equally to memory and disk, use ``device'' instead of ``memory''}
 
-			Since bandwidth sharing is not simulated, this prototype doesn't support concurrency: it is limited 
-			to single-threaded pipelines running on systems 
-			with a single-core CPU. \tristan{the following aren't resulting from the lack of bandwidth sharing simulation, these are more general limitations, I think we 
-			should remove them: , a single local storage device, and all input and 
-			output files are stored locally.} We use it in a first validation of our simulation model,
-			against a real sequential pipeline running on
-			a real system.
+			Since bandwidth sharing is not simulated, this prototype doesn't support 
+			concurrency: it is limited to single-threaded pipelines running on systems 
+			with a single-core CPU. We use it in a first validation of our simulation 
+			model, against a real sequential pipeline running on a real system.
+			Source code of our Python prototype is available at 
+			https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim
 
@@
 			using the current version of SimGrid and the SimGrid version that is 
-			extended with our model. To simulate our pipelines and system, we leverage the Workflow Management System Simulation Workbench (WRENCH~\cite{wrench}), a framework  
-			built over SimGrid.
+			extended with our model. To simulate our pipelines and system, we leverage 
+			the Workflow Management System Simulation Workbench 
+			(WRENCH~\cite{wrench}), a framework  built over SimGrid.
 		    We compare the results of the simulators with
 			original SimGrid and extended SimGrid with the results of real
-			pipelines on a real system \tristan{this will have to be updated with a description of extended SimGrid, 
-			you should add a placeholder for this description.}.
+			pipelines on a real system.
+
+			\textcolor{red}{[Description, implementation of our SimGrid extension]}.
 
@@
 			is at \url{https://github.com/wrench-project/wrench}.
-			\tristan{You should also add the URL of your Python prototype}
 
@@
 			effects, and writes the result to disk. The output of the previous
-			task is the input of the next task. We also use a real application pipeline, to evaluate the applicability of our
-			simulation model \tristan{add a placeholder to describe the real pipeline here}.
+			task is the input of the next task. We also use a real application 
+			pipeline, to evaluate the applicability of oursimulation model.
 
-			The simplest execution scenario is single-threaded, where memory 
-			bandwidth and disk bandwidth are dedicated. 
-			This allows us to compare the detailed results in terms of read/write time, 
-			memory states after every single task and, more importantly, validate our 
-			simulation model \tristan{last sentence is convoluted, reword}.
-
-			For multi-threaded applications, we vary the number of concurrent
-			pipelines running on a multi-core host and writing to the same local
-			disk. In this way, we evaluate our model with concurrent I/Os, when
-			the device bandwidth is shared.
-
-		    Since different storage types should also be taken into account, 
-		    we conduct the next experiment with the same scenario as 
-		    in the multi-threaded experiment, except that files are stored on a 
-		    shared file system instead of local disk.
-
-		    Finally, we consider a real pipeline from the neuroimaging domain. 
-		    We simulate the pipeline with a simulated workflow in SimGrid 
-			and compare the simulation results with the real results. 
-			Because our work focuses on I/O time, we assume that CPU time is 
-			correctly modeled and use the CPU time measured in the real pipelines 
-			to setup our simulated workflow \tristan{the pipeline should be described, including the workflow engine if it's Dask.}.
-
-			We use our dedicated cluster hosted at Concordia University to conduct the experiments. The cluster 
-			consists of one login node, 8 compute nodes 
+			We use our dedicated cluster hosted at Concordia University to conduct 
+			the experiments. The cluster consists of one login node, 8 compute nodes 
 			and 4 storage nodes connected with two network switches. The login node 
@@
 			of storage, 44 object storage targets with 8.7TB of storage each. 
-			The cluster is run on CentOS 8.1 with the Slurm Workload Manager installed. We use 
-			\textbf{\textit{atop}} and \textbf{\textit{collectl}} as tools to monitor 
-			and collect data of memory, page cache status and disk throughput on 
-			the cluster. The cluster nodes are connected with ... 
-			\textcolor{red}{[network description here]}
-			\tristan{mention the number of disks on all nodes. Maybe add a diagram.}
+			The cluster is run on CentOS 8.1 with the Slurm Workload Manager installed. 
+			We use \textbf{\textit{atop}} and \textbf{\textit{collectl}} as tools to 
+			monitor and collect data of memory, page cache status and disk throughput 
+			on the cluster. 
+			\textcolor{red}{[Add network description here, include number of disks 
+			on all nodes ,maybe add a diagram]}
 
-			\tristan{I don't think you need an intro paragraph for each experiment, as you did above, AND then subsections. I'd just make a single paragraph or sub-section.}
 			\subsubsection{Single-threaded evaluation}
 
-				In this experiment, our goal is to validate our model by comparing 
-				the simulation results of the simple Python simulator with results of 
-				a simulator implemented with original SimGrid, a simulator 
-				implemented with SimGrid integrated with our model, and a real pipeline 
-				running of the cluster. 
-				The pipeline consists of three tasks, each task reads a file, 
-				increases every byte and writes output file to local disk. 
+			The simplest execution scenario is single-threaded, where memory 
+			bandwidth and disk bandwidth are dedicated. Our goal is to 
+			compare the detailed simulation results of each task in the pipeline 
+			including read/write time, memory status and then validate the correctness 
+			of the simulation model.
+
+			The pipeline used in this experiment consists of three tasks, 
+			each task reads a file, increases every byte and writes output file 
+			to local disk. 
 			We use the input sizes of 20 GB, 50 GB, 75 GB and 100 GB, run 
-				the pipeline on the cluster, measure the CPU time of the pipeline 
-				with each input to simulate the pipeline with the simulators.
+			the pipeline on the cluster, measure the CPU time of the tasks  
+			with each input and use it to simulate CPU time in the simulators .
+
+			We simulate the pipeline with the Python prototype simulator, 
+			a simulator implemented with original SimGrid and a simulator implemented 
+			with SimGrid extended with our model.
 
@@
 
-				In the second experiment, we evaluate the simulator in concurrent I/O 
-				with a multi-threaded application. As we run the pipelines on 
-				only one compute node of the cluster, which has upto 32 cores per node,  
+			In the second experiment, we evaluate the simulator in a concurrent I/O, 
+			bandwidth-sharing scenario with a multi-threaded application. 			
+			We vary the number of concurrent pipelines running on a multi-core host.
+
+			All the task in the pipelines read input files from and write output files 
+			to a local disk. 
+			As a compute node of the cluster has 32 cores CPU and 256 GB of RAM,  
 			we create 32 input files with the size of 3GB each and vary the number of 
 			concurrent pipelines from 1 to 32.
-				Because multi-threaded simulator implementation in Python is 
+
+			Because implementation of a multi-threaded simulator  in Python is 
 			expensive, and we have our model integrated in SimGrid, 
 			we only compare the results of simulators with original SimGrid, 
-				SimGrid integrated with our model and and real pipelines. 
-				The results are compared in total makespan of the pipelines, 
+			SimGrid integrated with our model and and real pipelines in this 
+			experiment. The results are compared in total makespan of the pipelines, 
 			cumulative read time and cumulative write time.
@@
 
+		    Since different storage types should also be taken into account, 
+		    we conduct another experiment with the same scenario as 
+		    in the multi-threaded experiment, except that files are stored on a 
+		    shared file system instead of local disk.
+
+		    \textcolor{red}{[Description of Lustre on the cluster and how it's 
+		    simulated]}.
+
 		    \subsubsection{Simulation of a real application}
-				A real pipeline (for example a pipeline with nighres)
+
+		    Finally, we consider a real pipeline from the neuroimaging domain. 			
+			\textcolor{red}{[Description of the real pipeline with nighres 
+			(including the workflow engine if it's Dask)]}.
+
+			We simulate this pipeline with a simulator with original SimGrid and 
+			extended SimGrid. 
+			The simulation results are then compared with the real results. 
+			Because our work focuses on I/O time, we assume that CPU time is 
+			correctly modeled and use the CPU time measured in the real pipelines 
+			to setup our simulated workflow. The results are compared in terms of 
+			I/O time and total makespan.
 

### 2020-07-17 paper_io_sim.tex

@@
 \newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:}#1\color{black}}
+\newcommand{\valerie}[1]{\color{blue}\textbf{From Valerie:}#1\color{black}}
 
@@
 	\section{Method}
-		In this section, we present our approach to model file read/write,
+
+    \valerie{ In this section, you use terms like, ``file read and write'', ``memory I/O''
+    and ``Kernel I/O''. I think it's important that you define them in your introduction
+and make sure that these terms are actually distinct. By ``file read and write'', i'm guessing you essentially
+mean disk+memory I/O, by memory I/O, you must mean I/O to page cache (and not anonymous memory management). I'm not sure
+what you mean by ``kernel I/O'' unless it's the whole ensemble of i/o componenets?}
+		In this section, we present our approach to model file read and write,
 		memory I/O, cache eviction and data flushing mechanisms implemented in
@@
 			We leverage existing storage simulation models to simulate the memory,
-			which is characterized with capacity, read and write bandwidths. 
-			This can be done when disks are modeled with distinguished read/write 
-			bandwidths and negligible disk latency, which is applicable to memory 
+			which is characterized by storage capacity and read and write bandwidths. 
+			This can be done when disks are modeled with distinct read/write 
+            bandwidths and negligible disk latency \valerie{Does this mean we assume that all file i/o within a pipeline occurs on relatively large files?}, which is applicable to memory 
 			\cite{lebre2015}. In addition, leveraging existing disk models allows us 
@@
 			simulate parallel memory accesses.\tristan{we should refer to related work, 
-			see comment there}.
+            see comment there} \valerie{Agreed. It's unclear how existing storage models enable this at the moment}.
 
@@
 			twice the size of the inactive list by moving least recently 
-			used data blocks from the active list to the inactive list. 
+            used data blocks from the active list to the inactive list.\valerie{should maybe explain this in the datablock section}
 			Modeling page cache LRU lists as lists of data blocks reduces the
@@
 
+            \valerie{I think you might need to reorder the paragraphs. First one would be
+            to explain what and dirty/clean blocks, 2. what happens when data is read, 3. what happens
+            when data is written, and maybe mention file accesses that are not read/write}
             The simulated LRU lists are used as follows. When a file is read, 
-			file data can be partially cached, completely cached, or uncached. 
+            file data \valerie{are you talking about the file you just read or file data already in cache?} can be partially cached, completely cached, or uncached. 
+            \valerie{rewrite: Cached data can be composed of data blocks existing in either or both
+            of the page cache LRU lists. Cached blocks reside in the inactive list
+            unless they have been recently been reaccessed. Cached blocks located on the inactive
+            list that are subsequently accessed, will be moved to the top of the active list.}
             For cached data, there are existing data blocks representing
@@
 
-			We also keep the size of the active list not larger than twice the size of 
-			the inactive list as in the kernel 
+            As done in the kernel, we ensure the size of the active list does not exceed twice the size of 
+			the inactive list 
 			\cite{gorman2004understanding, linuxdev3rd2010}. 
@@
 			Therefore, total read time can be considered as sequential reads with 
-			these different bandwidths, enabling us to generalize reading process as 
+            these different bandwidths\valerie{i understand what you're saying,
+            but maybe it'd be better to say: ``Therefore read time can be considered as the sum of the time required to write to all devices''.
+            Can you clarify if disk+memory writes are not concurrent }, enabling us to generalize reading process as 
 			two separate phases, cache read and disk read. 
@@
 			(line 13), the storage model simulates the disk write with the remaining 
-			amount of data (line 14), and data read from disk is added to cache 
+            amount of data (line 14) \valerie{Has this line been removed?}, and data read from disk is added to cache 
 			(line 15). Similar to cache read period, disk bandwidth sharing is handled 
@@
 
+            \valerie{I think we need to take periodic flush out as it shouldn't be added to the write time}
+
 			Algorithm~\ref{alg:write} describes our simulation of file write. 
@@
 			Similar to the file read simulation, a file can also be written with 
-			two different bandwidths, memory bandwidth and disk bandwidth. 
+            two different bandwidths, memory bandwidth and disk bandwidth \valerie{can't we just say cached to memory and written to disk and assume the bandwidths are implied?}. 
 			Data can be written with memory bandwidth before the amount 
-			of dirty data reaches dirty\_ratio. Thus, our algorithm initially checks 
+			of dirty data reaches the dirty\_ratio. Thus, our algorithm initially checks 
 			the amount of data that can be written with memory write 
@@
 			To validate our simulation model, we create a simple prototype
-			simulator independently of existing simulation frameworks and libraries. 
+			simulator independent of existing simulation frameworks and libraries. 
 			This enables us to evaluate the accuracy and correctness of our 
@@
 			Source code of our Python prototype is available at 
-			https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim
+            \url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}
 
@@
 			using the current version of SimGrid and the SimGrid version that is 
-			extended with our model. To simulate our pipelines and system, we leverage 
+            extended with our model \valerie{I think you should give the extended simgrid a new name ``SimGridExt'' or something just so you can use that name in you figures}.
+To simulate our pipelines and system, we leverage
 			the Workflow Management System Simulation Workbench 
@@
 
-			The goal of our experiments is to evaluate our write-back cache
+        The goal of our experiments is to evaluate our write-back cache \valerie{This is the first time the write-back cache is mentioned apart from in a figure caption.
+definitely should be described in your intro alongside write-through and direct I/O}
 			simulation in single-threaded and multi-threaded applications
@@
 			task is the input of the next task. We also use a real application 
-			pipeline, to evaluate the applicability of oursimulation model.
+            pipeline \valerie{Maybe briefly describe this application's purpose (e.g.
+a standard fMRI preprocessing piepeline) or something of the sort}, to evaluate the applicability of our simulation model.
 
@@
 			and 4 storage nodes connected with two network switches. The login node 
-			has one Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 128GiB of RAM, 1.8TB 
-			of storage of XFS file system, 13GB of tmpfs file system and 385 TB of 
+			has one Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 128~GiB of RAM, 1.8TB 
+			of storage of XFS file system, 13~GB of tmpfs file system and 385 TB of 
 			Lustre shared file system. Each compute node has two 16 cores Intel(R) 
-			Xeon(R) Gold 6130 CPU @ 2.10GHz, 256 GB of RAM and 6 SSDs, 450GB each 
-			with XFS file system, 378GB of tmpfs and 126GB of devtmpfs file system.
-			Lustre file system is configured with one metadata target with 854GB 
-			of storage, 44 object storage targets with 8.7TB of storage each. 
+			Xeon(R) Gold 6130 CPU @ 2.10GHz, 256~GB of RAM and 6 SSDs, 450~GB each 
+			with XFS file system, 378~GB of tmpfs and 126~GB of devtmpfs file system.
+			Lustre file system is configured with one metadata target with 854~GB 
+			of storage, 44 object storage targets with 8.7~TB of storage each. 
 			The cluster is run on CentOS 8.1 with the Slurm Workload Manager installed. 
@@
 			to local disk. 
-			We use the input sizes of 20 GB, 50 GB, 75 GB and 100 GB, run 
+            \valerie{Be careful because you've been switching between GiB and GB. Make sure your input sizes are not GiB}
+			We use the input sizes of 20~GB, 50~GB, 75~GB and 100~GB, run 
 			the pipeline on the cluster, measure the CPU time of the tasks  
@@
 			to a local disk. 
-			As a compute node of the cluster has 32 cores CPU and 256 GB of RAM,  
-			we create 32 input files with the size of 3GB each and vary the number of 
+			As a compute node of the cluster has 32 cores CPU and 256~GB of RAM,  
+			we create 32 input files with the size of 3~GB each and vary the number of 
 			concurrent pipelines from 1 to 32.

### 2020-07-20 paper_io_sim.tex
Revision

--- 2020-07-09 paper_io_sim.tex
@@
 			\item The objective of the paper: Add capability to simulate I/O with page cache in SimGrid.
+			\item Clarify, distinguish between memory I/O (i/o to page cache) and disk I/O.
 		\end{itemize}
@@
 			\begin{itemize}
-				\item What is page cache? How it works \cite{linuxdev3rd2010}. Effects and importance of page cache.
+				\item What is page cache? How it works \cite{linuxdev3rd2010}. 
+				Write-back vs writethrough and direct i/o. Effects and importance of page cache.
 				\item Introduce some existing strategies with some highlighted pros and cons.
@@
 				\item Related development: RAM energy consumption \cite{gill2019} \cite{ouarnoughi2017} 
+				\item What is implemented in simulators that useful for our model:
+				distinguished and shared bandwidths, concurrency, device latency. 
 				\item Discuss the pros of SimGrid and the reasons why we chose it to extend. (Section 2.2.2 in \cite{casanova2014})
@@
 	\section{Method}
-		In this section, we present our approach to model file read/write,
-		memory I/O, cache eviction and data flushing mechanisms implemented in
-		the Linux kernel. We also detail the design of our simulators and their
+
+		In this section, we present our approach to model file read and write with
+		I/O to page cache, cache eviction and data flushing mechanisms implemented 
+		in the Linux kernel. We also detail the design of our simulators and their
 		implementation. Finally, we describe experimental scenarios to evaluate
 		our simulation model on real applications using a standalone Python
-		prototype, the current SimGrid simulator, and our extension of SimGrid.
+		prototype, the current SimGrid simulator, and the SimGrid version that is 
+		extended with our simulation model (SimGridExt).
 
@@
             We separate our simulation model in two components, the IO
-			Controller and the Memory Manager, which together simulate Linux
-			kernel I/Os (Figure~\ref{fig:interaction}). The Memory Manager
-			simulates memory I/O, flushing, and cache eviction, whereas the IO
-			Controller simulates file reads and writes. Simulated applications
-			can simply interact with the IO Manager instead of calling simulated
-			storage directly as is the case with current simulators.
+			Controller and the Memory Manager, which together simulate 
+			file read/write mechanisms in Linux (Figure~\ref{fig:interaction}). 
+			The Memory Manager simulates I/Os to cache, flushing, and cache eviction, 
+			whereas the IO Controller simulates file reads and writes. 
+			Simulated applications can simply interact with the IO Manager instead of 
+			calling simulated storage directly as is the case with current simulators.
 
@@
 			We leverage existing storage simulation models to simulate the memory,
-			which is characterized with capacity, read and write bandwidths. 
-			This can be done when disks are modeled with distinguished read/write 
-			bandwidths and negligible disk latency, which is applicable to memory 
-			\cite{lebre2015}. In addition, leveraging existing disk models allows us 
+			which is characterized by storage capacity, distinguished read/write 
+			bandwidths and device latency. 
+			All of these features are dealt within models in simulation 
+			frameworks, which makes storage models applicable to memory \cite{lebre2015}. 
+			In addition, leveraging existing disk models allows us 
 			to take advantage of bandwidth sharing between concurrent disk I/O to 
 			simulate parallel memory accesses.\tristan{we should refer to related work, 
-			see comment there}.
+            see comment there} \valerie{Agreed. It's unclear how existing storage models enable this at the moment}.
 
-			We introduce the concept of a data block as a unit to represent data
+			In Linux kernel, the page cache LRU lists contain file pages. However, 
+			due to the large number of file pages, managing lists of pages may 
+			require significant computation overhead. 
+			Thus, we introduce the concept of a data block as a unit to represent data 
 			cached in memory. A data block is a subset of file pages stored in
@@
 
-			We simulate the kernel LRU lists of file pages
-			with two lists of data blocks: an active list and 
-			an inactive list, ordered by last access time (earliest first, see 
-			Figure~\ref{fig:lrulist}).
+			\begin{figure}
+   				\centering
+   				\includegraphics[width=\columnwidth]{figures/lru_lists.pdf}
+   				\caption{Simulation of page cache LRU lists with data blocks}	\label{fig:lrulist}
+			\end{figure}
+
+			Now, page cache LRU lists can be modeled with two lists of data blocks: 
+			an active list and an inactive list, ordered by last access time 
+			(earliest first, see Figure~\ref{fig:lrulist}).
 			As in the kernel, our simulator keeps the size of the active list under
@@
 
-			\begin{figure}
-   				\centering
-   				\includegraphics[width=\columnwidth]{figures/lru_lists.pdf}
-   				\caption{Simulation of page cache LRU lists with data blocks}\label{fig:lrulist}
-			\end{figure}
+			Given an arbitrary file, at a specific time, its data can be partially 
+			cached, completely cached, or uncached.
+			Cached data can be composed of data blocks existing in either or both
+            of the page cache LRU lists. Cached blocks reside in the inactive list
+            unless they have been recently been reaccessed. Cached blocks 
+            located on the inactive list that are subsequently accessed, 
+            will be moved to the top of the active list. Cached blocks 
+            recently written to cache are dirty until they are flushed to disk.
 
             The simulated LRU lists are used as follows. When a file is read, 
-			file data can be partially cached, completely cached, or uncached. 
+            as mentioned, its data can be partially cached, completely cached, 
+            or uncached. 
             For cached data, there are existing data blocks representing
@@
 
-			We also keep the size of the active list not larger than twice the size of 
-			the inactive list as in the kernel 
+			\textcolor{red}{[Is there any kind of accesses other than read and write? 
+			Should we mention them?]}.
+
+            As done in the kernel, we ensure the size of the active list does not exceed twice the size of 
+			the inactive list 
 			\cite{gorman2004understanding, linuxdev3rd2010}. 
@@
 					\State mm.evict(2*fs - mm.cached(fn) - mm.free\_mem) 
+					\State mm.periodic\_flush() \Comment{Periodically flush expired dirty data}
 					\If {mm.cached(fn) $>$ 0}  \Comment{Read file cached data}
-					    \State mm.periodic\_flush() \Comment{Flush during cache read}
-    					\State mm.cache\_read(fn) 
+    					\State mm.cache\_read(fn)  \Comment{Update cached data}
 					\EndIf
 					\If {mm.cached(fn) $<$ fs} \Comment{Read uncached data from disk}
-						\State mm.periodic\_flush() \Comment{Sequential with disk read}
+						\State sm.read(fs - mm.cached(fn)) \Comment{Simulated disk read}
 						\State mm.add\_to\_cache(fs - mm.cached(fn), fn)
-    					\State sm.read(fs - mm.cached(fn))
 					\EndIf
@@
 
-			The algorithm to simulate file reads is described in 
-			Algorithm~\ref{alg:read}.   
-			In Linux systems, data can be either read from cache with memory bandwidth 
-			or read from disk with disk bandwidth. 
-			Therefore, total read time can be considered as sequential reads with 
-			these different bandwidths, enabling us to generalize reading process as 
-			two separate phases, cache read and disk read. 
+			The algorithm to simulate file reads is described in Algorithm~\ref{alg:read}.   
+%			In Linux systems with page cache, data can be either read from cache 
+%			or read from disk with disk bandwidth. 
+%			Therefore read time can be considered as the sum of the time required to 
+%			read from all devices\valerie{Can you clarify if disk+memory writes are not concurrent }, enabling us to generalize reading process as 
+%			two separate phases, cache read and disk read. 
+			In Linux systems with page cache, data can be either read from cache 
+			or read from disk concurrently. As this concurrency is modeled in simulation 
+			frameworks, we simulate a file read as follows.
 			When a simulated application reads a file, it sends a file read request 
@@
 			This flushing is complemented by eviction by the Memory Manager (line 7). 
+			In lines 6-7, we assume that user applications require the amount of memory 
+			equal to the file size, making total required memory twice the size of 
+			the file. Apart from flushing a required amount of memory, periodical 
+			flushing is then triggered to flush expired dirty data (line 8).
 			Next, if the file is cached, either partially or entirely, data is read from 
-			cache first. Periodical flushing is triggered at the beginning of the 
-			cache read to flush dirty data if some dirty data is expired (line 9).
-			At the same time, Memory Manager is called to read data from cache, 
-			and update the corresponding data blocks in cache (line 10). 
-			The concurrency of periodical flushing and cache read is handled and 
-			time is simulated by the memory and disk models.
+			cache first (line 9). Memory Manager is called to read data from cache, 
+			update the corresponding data blocks in cache (line 10).
 
 			If file is not cached or partially cached, a disk read is required (line 12). 
-			Periodical flushing is called again at the beginning of the disk read 
-			(line 13), the storage model simulates the disk write with the remaining 
-			amount of data (line 14), and data read from disk is added to cache 
-			(line 15). Similar to cache read period, disk bandwidth sharing is handled 
-			and time is simulated by storage models.
+			The storage model simulates the disk read with the remaining 
+            amount of data (line 13), and data read from disk is added to 
+            cache (line 14). The total file read time is simulated at the end of the algorithm. 
+            During the entire file read, the concurrency between periodical flushing, 
+            cache read and disk read is handled by storage models.
 
@@
 			dirty data at the beginning of the write (line 6).
-			Similar to the file read simulation, a file can also be written with 
-			two different bandwidths, memory bandwidth and disk bandwidth. 
-			Data can be written with memory bandwidth before the amount 
-			of dirty data reaches dirty\_ratio. Thus, our algorithm initially checks 
-			the amount of data that can be written with memory write 
-			bandwidth (line 7-8).
+			During file write simulation, a file is cached in memory and also written 
+			to disk. Data can be written to cache without waiting for flushing or 
+			being throttled before the amount of dirty data reaches dirty\_ratio. 
+			Thus, our algorithm initially checks the amount of data that can be written 
+			with memory write bandwidth (line 7-8).
 			If this amount is greater than 0, MemoryManager is asked to evict 
@@
 			added to page cache LRU lists (11).
-			The amount of data that is not written with memory bandwidth is written 
-			with disk bandwidth (line 13). 
-			Now, the dirty ratio threshold is reached, we have to flush and evict 
-			data from cache as much as possible so that 
-			the remaining data can be written to cache (line 14-15). 
+			After dirty\_ratio is reached, dirty data has to be flushed before new data 
+			is written to cache again. 
+			We flush and evict data from cache as much as possible so that the 
+			remaining data can be written to cache (line 14-15). 
 			Here, MemoryManager does not guarantee that the amount of flushed 
 			and evicted data is enough for the remaining data. In the case it is 
-			less than the required amount, the remaining data is written to cache 
-			but then can be flushed and evicted right away. Thus, the amount 
-			of remaining data that is written and kept in cache after the write 
-			is limited by the amount free memory (line 16). 
+			less than the required amount and existing dirty data can not be flushed, 
+			the remaining data being written to cache can be flushed and evicted right away. 
+			Thus, the amount of remaining data written and kept in cache 
+			after the write is limited by the amount free memory, and this amount 
+			is added to cache (line 16). 
 			Finally, the disk write is simulated with the remaining amount of 
-			data (line 17). As in the simulation of file read, all the concurrent 
-			cache write and disk write are simulated by memory and disk models.
+			data (line 17). As in the simulation of file read, all the concurrency  
+			of cache write and disk write are simulated by memory and disk models.
 
@@
 			Source code of our Python prototype is available at 
-			https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim
+            \url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}
 
 			Having our model validated, we create simulators for different use cases 
-			using the current version of SimGrid and the SimGrid version that is 
-			extended with our model. To simulate our pipelines and system, we leverage 
+			using original SimGrid and our SimGridExt version. 
+			To simulate our pipelines and system, we leverage 
 			the Workflow Management System Simulation Workbench 
@@
 
-			\textcolor{red}{[Description, implementation of our SimGrid extension]}.
+			\textcolor{red}{[Description, implementation of our SimGridExt]}.
 
@@
 			task is the input of the next task. We also use a real application 
-			pipeline, to evaluate the applicability of oursimulation model.
+            pipeline \textcolor{red}{[Maybe briefly describe this application's purpose (e.g. a standard fMRI preprocessing piepeline)]}, to evaluate the applicability of our simulation model.
 
@@
 			and 4 storage nodes connected with two network switches. The login node 
-			has one Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 128GiB of RAM, 1.8TB 
-			of storage of XFS file system, 13GB of tmpfs file system and 385 TB of 
+			has one Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 137~GB (128~GiB) of RAM, 
+			1.8TB of storage of XFS file system, 13~GB of tmpfs file system and 385 TB of 
 			Lustre shared file system. Each compute node has two 16 cores Intel(R) 
-			Xeon(R) Gold 6130 CPU @ 2.10GHz, 256 GB of RAM and 6 SSDs, 450GB each 
-			with XFS file system, 378GB of tmpfs and 126GB of devtmpfs file system.
-			Lustre file system is configured with one metadata target with 854GB 
-			of storage, 44 object storage targets with 8.7TB of storage each. 
+			Xeon(R) Gold 6130 CPU @ 2.10GHz, 275~GB (256~GiB) of RAM and 6 SSDs, 
+			450~GB each with XFS file system, 378~GB of tmpfs and 126~GB of devtmpfs file system.
+			Lustre file system is configured with one metadata target with 854~GB 
+			of storage, 44 object storage targets with 8.7~TB of storage each. 
 			The cluster is run on CentOS 8.1 with the Slurm Workload Manager installed. 
@@
 			to local disk. 
-			We use the input sizes of 20 GB, 50 GB, 75 GB and 100 GB, run 
+			We use the input sizes of 20~GB, 50~GB, 75~GB and 100~GB, run 
 			the pipeline on the cluster, measure the CPU time of the tasks  
@@
 			to setup our simulated workflow. The results are compared in terms of 
-			I/O time and total makespan.
+			read/write time and total makespan.
 
@@
 				\begin{itemize}
-					\item Errors of simulation time and memory used compared to real results.
+					\item Relative errors of simulation time and memory used compared to real results.
 					\item Simulation time compared to baseline SimGrid.
@@
 			\item Sensitivity of the simulator on the variation of memory and disk bandwidth. 
+			\item Our model does not support random I/O and readahead.
 		\end{itemize}

### 2020-07-22 paper_io_sim.tex
Methods updated

@@
 			\item Existing experiment methods, simulators, simulation frameworks.
-The advantages of SimGrid compared to others \cite{casanova2008, lebre2015}.
+The advantages of SimGrid compared to others \cite{simgrid2008, lebre2015}.
 The missing of the ability to simulate page cache in SimGrid \cite{lebre2015}.
@@
 				distinguished and shared bandwidths, concurrency, device latency. 
-				\item Discuss the pros of SimGrid and the reasons why we chose it to extend. (Section 2.2.2 in \cite{casanova2014})
+				\item Discuss the pros of SimGrid and the reasons why we chose it to extend. (Section 2.2.2 in \cite{simgrid2014})
 				\item \tristan{Mention the status of storage simulation in simulators, so that we understand from Fig 1 that storage devices are available}
@@
 			file read/write mechanisms in Linux (Figure~\ref{fig:interaction}). 
-			The Memory Manager simulates I/Os to cache, flushing, and cache eviction, 
-			whereas the IO Controller simulates file reads and writes. 
+			The Memory Manager simulates I/Os to cache, flushing, periodical flushing 
+			and cache eviction, whereas the IO Controller simulates file reads and writes. 
 			Simulated applications can simply interact with the IO Manager instead of 
@@
             see comment there} \valerie{Agreed. It's unclear how existing storage models enable this at the moment}.
+
+			Memory Manager simulates two parallel threads. The first thread is the 
+			main thread where activities of applications are simulated, and responsible 
+			for the simulation of data flushing, cache eviction and I/Os to cache. 
+			The second thread is a simulated background thread, in which Memory Manager 
+			periodically searches and flushes expired dirty data in page cache LRU lists 
+			to disks.
 
@@
 
-            As done in the kernel, we ensure the size of the active list does not exceed twice the size of 
-			the inactive list 
+            As done in the kernel, we ensure the size of the active list does not 
+            exceed twice the size of the inactive list 
 			\cite{gorman2004understanding, linuxdev3rd2010}. 
@@
 
-			Next, we simulate data flushing mechanism, which writes back 
-			dirty data to disk. 
+			Next, we simulate data flushing, which runs on the main simulated 
+			thread and writes back dirty data to disk. 
 			The data flushing simulation function traverses the sorted inactive first, 
@@
 
-			For periodical flushing, our simulator searches in two page cache 
-			LRU lists and flushes all the expired dirty data blocks. 
-			A dirty data block is considered expired if the time from its last access 
-			to current simulated time is longer than the predefined expire time. 
-			Periodical flushing can be concurrent with cache read/write, 
-			and sequential with disk read/write. When cache accesses and 
-			periodical flushing are concurrent, flushing time is not added to simulated 
-			time but the amount of flushed data is limited by flushing time. 
-			When data is flushed in parallel with disk accesses, the concurrency 
-			is handled by the storage model used as memory, and flushing time 
-			is simulated with the amount of flushed data by this model.
-
-			Similarly to data flushing, our cache eviction simulation function frees up 
-			the page cache by traversing and deleting least recently used clean 
-			data blocks in the inactive list.
+			Similarly to data flushing, our cache eviction simulation runs on 
+			the main thread, frees up the page cache by traversing and deleting 
+			least recently used clean data blocks in the inactive list.
 			The amount of evicted data is passed in and data blocks are deleted 
@@
 			since cache eviction time is negligible in real systems.
+
+			\begin{algorithm}\caption{Periodical flushing simulation}\label{alg:pdflush}
+				\small
+				\begin{algorithmic}[1]
+					\Input
+        				\Desc{in}{page cache inactive list}
+        				\Desc{ac}{page cache active list}
+						\Desc{t}{predefined flushing time interval}
+						\Desc{sm}{storage simulation model}
+   					\EndInput
+   					\While{has unfinished tasks or flushing}
+						\State blocks = expired dirty blocks in active and inactive list
+						\State flushing\_time = sm.write(blocks)
+						\State Flag blocks as clean data
+						\If{flushing\_time $<$ t}
+							\State sleep(t - flushing\_time)
+						\EndIf
+					\EndWhile
+				\end{algorithmic}
+			\end{algorithm}
+
+			The periodical flushing simulation is run on a simulated background thread 
+			to flush expired dirty data blocks from cache to disk.
+			Similar to expired dirty pages in real Linux systems, a dirty block in our model 
+			is considered expired if the time from its last access to current simulated time 
+			is longer than the predefined expire time.
+			The simulation algorithm of periodical flushing is an infinite loop in which 
+			Memory Manager searches and flushes these expired dirty blocks to disk, 
+			as described in Algorithm~\ref{alg:pdflush}. 
+			The loop ends when the host does not has any remaining incomplete task 
+			and all dirty data has been flushed (line 6). 
+			In each repetition, Memory Manager collects expired dirty blocks in two 
+			page cache LRU lists (line 7), simulates a write of this data to disk (line 8), 
+			and mark these blocks as clean (line 9).
+			If the time to flush data does not exceed our time interval, the thread is put 
+			to sleep for the remaining time (lines 10-12). 
+			Then simulation time is advanced, the algorithm finishes or continues the loop 
+			based on the conditions at current simulated time.
+			Because periodical flushing is simulated as a background thread, it can be 
+			concurrent with disk read/write. This parallelism is handled by the 
+			storage model and reflected in simulated I/O time.
 
@@
 					\State mm.evict(2*fs - mm.cached(fn) - mm.free\_mem) 
-					\State mm.periodic\_flush() \Comment{Periodically flush expired dirty data}
 					\If {mm.cached(fn) $>$ 0}  \Comment{Read file cached data}
-    					\State mm.cache\_read(fn)  \Comment{Update cached data}
+    					\State cache\_read = mm.cache\_read(fn)  \Comment{Read async}
 					\EndIf
 					\If {mm.cached(fn) $<$ fs} \Comment{Read uncached data from disk}
-						\State sm.read(fs - mm.cached(fn)) \Comment{Simulated disk read}
+						\State disk\_read = sm.read(fs - mm.cached(fn)) \Comment{Async disk read}
 						\State mm.add\_to\_cache(fs - mm.cached(fn), fn)
 					\EndIf					
-
+					\State Wait for completion of cache\_read and disk\_read 
 				\end{algorithmic}
@@
 			This flushing is complemented by eviction by the Memory Manager (line 7). 
-			In lines 6-7, we assume that user applications require the amount of memory 
-			equal to the file size, making total required memory twice the size of 
-			the file. Apart from flushing a required amount of memory, periodical 
-			flushing is then triggered to flush expired dirty data (line 8).
-			Next, if the file is cached, either partially or entirely, data is read from 
-			cache first (line 9). Memory Manager is called to read data from cache, 
-			update the corresponding data blocks in cache (line 10).
-
-			If file is not cached or partially cached, a disk read is required (line 12). 
-			The storage model simulates the disk read with the remaining 
-            amount of data (line 13), and data read from disk is added to 
-            cache (line 14). The total file read time is simulated at the end of the algorithm. 
-            During the entire file read, the concurrency between periodical flushing, 
-            cache read and disk read is handled by storage models.
+			Here, we assume that user applications use the amount of memory 
+			equal to the file size. 
+			Next, if the file is cached, either partially or entirely, data is read from cache (line 8). 
+			Memory Manager is called to simulate a asynchronous cache read, 
+			update the corresponding data blocks in cache (line 9). 
+			If file is not cached or partially cached, a disk read is required (line 11). 
+			The storage model simulates a asynchronous disk read with the remaining 
+            amount of data (line 12), and data read from disk is added to 
+            cache (line 13). 
+            At the end of the algorithm, we wait for the completion of two 
+            asynchronous activities, the cache read in line 9 and the disk read in line 12, 
+            to finish our file read and get the total simulated read time (line 15).
 
@@
    					\EndInput
-   					\State mm.periodic\_flush()
 					\State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
@@
 
-            \valerie{I think we need to take periodic flush out as it shouldn't be added to the write time}
-
 			Algorithm~\ref{alg:write} describes our simulation of file write. 
-			Initially, periodical flushing is called to asynchronously flush expired 
-			dirty data at the beginning of the write (line 6).
 			During file write simulation, a file is cached in memory and also written 
@@
 			Thus, our algorithm initially checks the amount of data that can be written 
-			with memory write bandwidth (line 7-8).
+			with memory write bandwidth (line 6-7).
 			If this amount is greater than 0, MemoryManager is asked to evict 
@@
 			Then, cache write is simulated with memory model and data written is 
-			added to page cache LRU lists (11).
+			added to page cache LRU lists (10).
 			After dirty\_ratio is reached, dirty data has to be flushed before new data 
@@
 			We flush and evict data from cache as much as possible so that the 
-			remaining data can be written to cache (line 14-15). 
+			remaining data can be written to cache (line 13-14). 
 			Here, MemoryManager does not guarantee that the amount of flushed 
@@
 			after the write is limited by the amount free memory, and this amount 
-			is added to cache (line 16). 
+			is added to cache (line 15). 
 			Finally, the disk write is simulated with the remaining amount of 
-			data (line 17). As in the simulation of file read, all the concurrency  
-			of cache write and disk write are simulated by memory and disk models.
+			data (line 16), and the algorithm finishes with the total file write simulation time.
 
@@
 			the Workflow Management System Simulation Workbench 
-			(WRENCH~\cite{wrench}), a framework  built over SimGrid.
+			(WRENCH~\cite{wrench2018}), a framework built over SimGrid.
 		    We compare the results of the simulators with

### 2020-07-24 paper_io_sim.tex
minor comment

@@
 			file read/write mechanisms in Linux (Figure~\ref{fig:interaction}). 
-			The Memory Manager simulates I/Os to cache, flushing, periodical flushing 
-			and cache eviction, whereas the IO Controller simulates file reads and writes. 
-			Simulated applications can simply interact with the IO Manager instead of 
+			The Memory Manager simulates I/Os to cache, flushing, periodical flushing, 
+			and cache eviction \tristan{how about periodical eviction?}, whereas the IO Controller simulates file reads and writes. 
+			Simulated applications can simply interact with the IO Controller instead of 
 			calling simulated storage directly as is the case with current simulators.
@@
 			\item Our model does not support random I/O and readahead.
+			\item Disk state save/restore when hosts are turned on/off.
 		\end{itemize}

### 2020-09-21 paper_io_sim.tex
Update with chunk i/o

@@
 
-\author{\IEEEauthorblockN{Hoang-Dung Do, Val\'erie Hayot-Sasson, Tristan Glatard
+\author{\IEEEauthorblockN{Hoang-Dung Do, Tristan Glatard, Val\'erie Hayot-Sasson
   }\\
@@
 				distinguished and shared bandwidths, concurrency, device latency. 
-				\item Discuss the pros of SimGrid and the reasons why we chose it to extend. (Section 2.2.2 in \cite{simgrid2014})
+				\item Discuss the pros of SimGrid and the reasons why we chose it to conduct experiments. (Section 2.2.2 in \cite{simgrid2014})
+				\item Introduce WRENCH and it's advantage.
 				\item \tristan{Mention the status of storage simulation in simulators, so that we understand from Fig 1 that storage devices are available}
@@
 		our simulation model on real applications using a standalone Python
-		prototype, the current SimGrid simulator, and the SimGrid version that is 
-		extended with our simulation model (SimGridExt).
+		prototype, SimGrid framework using original WRENCH and the WRENCH version 
+		that is integrated with our simulation model (WRENCH-Ext).
 
@@
 			The Memory Manager simulates I/Os to cache, flushing, periodical flushing, 
-			and cache eviction \tristan{how about periodical eviction?}, whereas the IO Controller simulates file reads and writes. 
+			and cache eviction, whereas the IO Controller simulates file reads and writes. 
 			Simulated applications can simply interact with the IO Controller instead of 
@@
 			We leverage existing storage simulation models to simulate the memory,
-			which is characterized by storage capacity, distinguished read/write 
+			which is characterized with storage capacity, distinguished read/write 
 			bandwidths and device latency. 
@@
 			cached in memory. A data block is a subset of file pages stored in
-            page cache that were accessed in the same read or write operation. 
+            page cache that were accessed in the same I/O operation. 
 			A data block has information about file name, block size, last access 
-			time, and a dirty flag that represents whether the data is clean (0) 
-			or dirty (1). 
-			A given file can have multiple data blocks in page cache. In
-			addition, a data block can be split into an arbitrary number of
-			smaller blocks, and data blocks can be merged together.
+			time, a dirty flag that represents whether the data is clean (0) 
+			or dirty (1), and the time it is written to cache as dirty data. 
+			Blocks can have different sizes and a given file can have multiple 
+			data blocks in page cache. In addition, a data block can be split into an 
+			arbitrary number of smaller blocks.
 
@@
 			overhead of the simulator, as data blocks are obviously less
-			numerous than file pages, while preserving the accuracy in I/O time
-			simulation \tristan{how do you know that, can you provide an argument?}.
+			numerous than file pages, while achieving the accuracy in I/O time
+			simulation that can be seen in the later section.
 
-			Given an arbitrary file, at a specific time, its data can be partially 
-			cached, completely cached, or uncached.
-			Cached data can be composed of data blocks existing in either or both
+			Given an arbitrary file, at a specific time, it can be partially cached, 
+			completely cached, or uncached.
+			Cached data is composed of data blocks existing in either or both
             of the page cache LRU lists. Cached blocks reside in the inactive list
-            unless they have been recently been reaccessed. Cached blocks 
+            unless they have been recently reaccessed. Cached blocks 
             located on the inactive list that are subsequently accessed, 
             will be moved to the top of the active list. Cached blocks 
-            recently written to cache are dirty until they are flushed to disk.
+            recently written to cache are dirty blocks until they are flushed to disk.
 
-            The simulated LRU lists are used as follows. When a file is read, 
-            as mentioned, its data can be partially cached, completely cached, 
+            Our model supports chunk-by-chunk file read/write simulation
+            with an user-defined chunk size, in which chunks are read/written 
+            until file is entirely read/written. 
+            The simulated LRU lists are used as follows. 
+            When a file is read, its data can be partially cached, completely cached, 
             or uncached. 
-            For cached data, there are existing data blocks representing
-			this data in two page cache LRU lists. 
-			As this cached data is accessed again, these blocks are moved to 
-			the active list, dirty blocks are merged into a bigger dirty block, 
-			while clean data blocks are merged into another clean block. 
-			If data read is not in cache, a clean block representing uncached data 
-			is created and put into the top of the inactive list.
+            Uncached data is read from disk first, and cached data is only read 
+            after all uncached data is read (left to right in Figure~\ref{fig:read_order}).
+            When a chunk of uncached data is read, a new clean block is created 
+            and put to the top of the inactive list. 
+            For a chunk of cached data, because one or more existing data blocks 
+            in two page cache LRU lists are accessed again, read dirty blocks are 
+            moved to the active list, while clean data blocks are merged into 
+            a single clean block and then also put in the active list. 
+			The reason we do not merge dirty blocks is that different blocks 
+			are written to cache at different time, and this time is used later for 
+			periodical flushing process.
+			In case a block is not entirely read, it is split into two smaller blocks, 
+			one is reaccessed and the other remains the same.
 
-			In the case of writing data, we can assume that all written data 
-			is uncached. After data is written, some can remain dirty, some is 
-			clean after being flushed. 
-			Thus, at most two data blocks can be added, one is clean and one 
-			is dirty, to the inactive list.
+			\begin{figure}
+   				\centering
+   				\includegraphics[width=\columnwidth]{figures/read_order.pdf}
+   				\caption{File data read order}	\label{fig:read_order}
+			\end{figure}
 
-			\textcolor{red}{[Is there any kind of accesses other than read and write? 
-			Should we mention them?]}.
+			In the case of writing data, we can assume that all data to be written is 
+			uncached and it is written to cache as dirty data. 
+			Thus, each time a chunk is written, we create a block of dirty data and 
+			put it in the inactive list.
 
-            As done in the kernel, we ensure the size of the active list does not 
+			\textcolor{red}{[Is there any kind of accesses other than read and write?]}.
+
+            As is done in the kernel, we ensure the size of the active list does not 
             exceed twice the size of the inactive list 
@@
 			thread and writes back dirty data to disk. 
-			The data flushing simulation function traverses the sorted inactive first, 
-			and then the active list, looks for least recently used dirty blocks, 
-			writes them to disk through the storage device model, and 
-			sets dirty flags of these blocks to 0 until the total flushed amount 
+			The data flushing simulation function traverses the sorted 
+			inactive list first, and then the active list, looks for least recently used 
+			dirty blocks, writes them to disk through the storage device model, 
+			and sets dirty flags of these blocks to 0 until the total flushed amount 
 			reaches the amount to be flushed or there is no dirty data left in cache. 
@@
 			and only one block is deleted.
-			Our cache eviction simulation does not add up to the simulated time 
+			The cache eviction simulation does not add up to the simulated time 
 			since cache eviction time is negligible in real systems.
@@
    					\EndInput
-   					\While{has unfinished tasks or flushing}
+   					\While{host is on}
 						\State blocks = expired dirty blocks in active and inactive list
@@
 			Similar to expired dirty pages in real Linux systems, a dirty block in our model 
-			is considered expired if the time from its last access to current simulated time 
-			is longer than the predefined expire time.
+			is considered expired if the time it remains in cache as dirty is longer than 
+			the predefined expire time. This time is counted from the time it is written 
+			to cache to current simulated time. 
 			The simulation algorithm of periodical flushing is an infinite loop in which 
@@
 			as described in Algorithm~\ref{alg:pdflush}. 
-			The loop ends when the host does not has any remaining incomplete task 
-			and all dirty data has been flushed (line 6). 
-			In each repetition, Memory Manager collects expired dirty blocks in two 
-			page cache LRU lists (line 7), simulates a write of this data to disk (line 8), 
-			and mark these blocks as clean (line 9).
-			If the time to flush data does not exceed our time interval, the thread is put 
+			The loop ends when the host the host is off (line 6). 
+			In each repetition, Memory Manager finds expired dirty blocks in two 
+			page cache LRU lists (line 7), simulates writes of data of these blocks 
+			to disk (line 8), and mark them as clean (line 9).
+			If the flushing time does not exceed our time interval, the thread is put 
 			to sleep for the remaining time (lines 10-12). 
-			Then simulation time is advanced, the algorithm finishes or continues the loop 
-			based on the conditions at current simulated time.
+			Then based on the state of the host at current simulated time, 
+			the algorithm continues or finishes the loop .
 			Because periodical flushing is simulated as a background thread, it can be 
-			concurrent with disk read/write. This parallelism is handled by the 
+			concurrent with disk I/O on main thread. This parallelism is handled by the 
 			storage model and reflected in simulated I/O time.
@@
 
-			Simulated applications can simply request for a file read 
-			or a file write by calling the IO Controller instead of 
-			the storage simulation models. 
-			\tristan{this should move, not sure where for now: Because every computer has its own memory and system configurations, 
+			When simulated applications need to read/write a file, they can simply call 
+			the IO Controller instead of the storage simulation models. 
+			\tristan{this should move, not sure where for now: 
+			Because every computer has its own memory and system configurations, 
 			one IO Controller object and one Memory Manager object are created 
-			for each simulated machine.}
+			for each simulated host.}
 
-			To read or write a file, a simulated application sends a request to the 
-			IO Controller instance on the node where the requested file is stored.
+			As mentioned, in our model, a file read/write can be complete by read/write 
+			chunk-by-chunk until entire file is read/written.
+			To read or write a chunk of file, a simulated application sends a request to the 
+			IOController instance on the node where the requested file is stored.
 			Based on the memory status in Memory Manager, the IOController 
-			orchestrates cache reads and writes, disk reads and writes, flushing, cache eviction 
-			 and returns to the simulated application.
+			orchestrates cache and disk reads/writes, flushing, cache eviction 
+			to simulate I/O time.
 
-			\begin{algorithm}\caption{File read simulation}\label{alg:read}
+			\begin{algorithm}\caption{File chunk read simulation}\label{alg:read}
 				\small
@@
 					\Input
+        				\Desc{cs}{chunk size}
         				\Desc{fn}{file name}
-        				\Desc{fs}{file size (assumed to fit in memory)}
+        				\Desc{fs}{file size (assume to fit in memory)}
 						\Desc{mm}{MemoryManager object}
@@
    					\EndInput
-					\State mm.flush(2*fs - mm.cached(fn) - mm.evictable\_mem) 
-					\State mm.evict(2*fs - mm.cached(fn) - mm.free\_mem) 
-					\If {mm.cached(fn) $>$ 0}  \Comment{Read file cached data}
-    					\State cache\_read = mm.cache\_read(fn)  \Comment{Read async}
+   					\State disk\_read = min(cs, fs - mm.cached(fn))
+   					\State cache\_read = cs - disk\_read
+					\State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
+					\State mm.evict(cs + disk\_read - mm.free\_mem, fn) 
+					\If {disk\_read $>$ 0}  \Comment{Read uncached data}	
+						\State sm.read(disk\_read)  
+						\State mm.add\_to\_cache(disk\_read, fn) 	
 					\EndIf
-					\If {mm.cached(fn) $<$ fs} \Comment{Read uncached data from disk}
-						\State disk\_read = sm.read(fs - mm.cached(fn)) \Comment{Async disk read}
-						\State mm.add\_to\_cache(fs - mm.cached(fn), fn)
+					\If {cache\_read $>$ 0} \Comment{Read cached}
+						\State mm.cache\_read(cache\_read)  
 					\EndIf
-					\State Wait for completion of cache\_read and disk\_read 
 				\end{algorithmic}
@@
 
-			The algorithm to simulate file reads is described in Algorithm~\ref{alg:read}.   
-%			In Linux systems with page cache, data can be either read from cache 
-%			or read from disk with disk bandwidth. 
-%			Therefore read time can be considered as the sum of the time required to 
-%			read from all devices\valerie{Can you clarify if disk+memory writes are not concurrent }, enabling us to generalize reading process as 
-%			two separate phases, cache read and disk read. 
-			In Linux systems with page cache, data can be either read from cache 
-			or read from disk concurrently. As this concurrency is modeled in simulation 
-			frameworks, we simulate a file read as follows.
-			When a simulated application reads a file, it sends a file read request 
-			to IO Controller. 
-			First, if there is not enough memory available to store two
-			copies of the file (one in cache, one in anonymous memory), the
-			Memory Manager is called to flush dirty data to disk and simulate 
-			flushing time with memory model (line 6).
-			This flushing is complemented by eviction by the Memory Manager (line 7). 
-			Here, we assume that user applications use the amount of memory 
-			equal to the file size. 
-			Next, if the file is cached, either partially or entirely, data is read from cache (line 8). 
-			Memory Manager is called to simulate a asynchronous cache read, 
-			update the corresponding data blocks in cache (line 9). 
-			If file is not cached or partially cached, a disk read is required (line 11). 
-			The storage model simulates a asynchronous disk read with the remaining 
-            amount of data (line 12), and data read from disk is added to 
-            cache (line 13). 
-            At the end of the algorithm, we wait for the completion of two 
-            asynchronous activities, the cache read in line 9 and the disk read in line 12, 
-            to finish our file read and get the total simulated read time (line 15).
+			The algorithm to simulate a chunk read is described in Algorithm~\ref{alg:read}.   
+			When a simulated application read a chunk of a file, it sends a 
+			chunk read request to IO Controller. 
+			First, we calculate the amount of uncached data that needs to be read 
+			from disk, and the remaining amount is read from cache (line 7-8).
+			We assume that after a file is read, there are two copies of the file in memory 
+			(one in cache, one in anonymous memory). 
+			Thus, when a chunk is read, one copy of the chunk occupies anonymous memory, 
+			and the uncached part of that chunk is read into cache.	
+			The amount of required memory to read the chunk is calculate and 
+			if there is not enough memory available, Memory Manager is called to 
+			flush dirty data to disk and simulate flushing time with memory model (line 9).
+			This flushing is complemented by eviction in the Memory Manager (line 10). 
+			Then, if there is uncached data, a disk read is simulated with disk model (line 12), 
+			and this data of the file is added to cache (line 13)
+			Finally, if cached data is read, Memory Manager is called to simulate a cache read, 
+			update the corresponding data blocks in cache (line 16).
 
-			\begin{algorithm}\caption{File write simulation}\label{alg:write}
+			\begin{algorithm}\caption{File chunk write simulation}\label{alg:write}
 				\small
@@
 					\Input
+        				\Desc{cs}{chunk size}
         				\Desc{fn}{file name}
-        				\Desc{fs}{file size}
 						\Desc{mm}{MemoryManager object}
@@
 					\State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
-					\State mem\_amt = min(fs, remain\_dirty)
 					\If {remain\_dirty $>$ 0} \Comment{Write with memory bandwidth}
-    					\State mm.evict(mem\_amt - mm.free\_mem)
+    					\State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
+    					\State mem\_amt = min(cs, mm.free\_mem)
     					\State mm.write(fn, mem\_amt) 
     				\EndIf
-					\If {mem\_amt $<$ fs}  \Comment{Write with disk bandwidth}
-						\State mm.flush(fs - mem\_amt)  \Comment{Concurrent with disk write}
-						\State mm.evict(fs - mem\_amt  - mm.free\_mem) 
-						\State mm.add\_to\_cache(fn, min(fs - mem\_amt, mm.free\_mem))
-						\State sm.write(fs - mem\_amt)
+					\If {mem\_amt $<$ cs}  \Comment{Write with disk bandwidth}
+						\State mm.flush(cs - mem\_amt)  
+						\State mm.evict(cs - mem\_amt  - mm.free\_mem) 
+						\State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem))
+						\State sm.write(cs - mem\_amt)
 					\EndIf
@@
 
-			Algorithm~\ref{alg:write} describes our simulation of file write. 
-			During file write simulation, a file is cached in memory and also written 
-			to disk. Data can be written to cache without waiting for flushing or 
-			being throttled before the amount of dirty data reaches dirty\_ratio. 
-			Thus, our algorithm initially checks the amount of data that can be written 
-			with memory write bandwidth (line 6-7).
+			Algorithm~\ref{alg:write} describes our simulation of chunk write. 
+			In our chunk write simulation, data can be written with either memory bandwidth 
+			or disk bandwidth. 
+			Data can be written to cache with memory bandwidth without waiting for 
+			flushing or being throttled before the amount of dirty data reaches dirty\_ratio. 
+			Thus, our algorithm initially checks the remaining amount of data that can be written 
+			as dirty data (line 6).
 			If this amount is greater than 0, MemoryManager is asked to evict 
-			data from cache to accommodate the written data if needed (line 10). 
-			Then, cache write is simulated with memory model and data written is 
-			added to page cache LRU lists (10).
+			data from cache to accommodate the written data if needed (line 8). 
+			After evicting cache, the actual amount of data that can be written with 
+			memory bandwidth is calculated (line 9), a cache write is then simulated 
+			with this amount through memory model and data written is added to 
+			page cache LRU lists (10).
 			After dirty\_ratio is reached, dirty data has to be flushed before new data 
-			is written to cache again. 
-			We flush and evict data from cache as much as possible so that the 
-			remaining data can be written to cache (line 13-14). 
+			can be written to cache again. 
+			If there is remaining chunk data that has not been written, we flush and 
+			evict data from cache as much as possible so that the this data can be 
+			written to cache (line 13-14). 
 			Here, MemoryManager does not guarantee that the amount of flushed 
-			and evicted data is enough for the remaining data. In the case it is 
-			less than the required amount and existing dirty data can not be flushed, 
-			the remaining data being written to cache can be flushed and evicted right away. 
+			and evicted data is enough for the write of remaining data. 
+			In the case it is less than the required amount, the remaining data 
+			being written to cache can be flushed and evicted right away. 
 			Thus, the amount of remaining data written and kept in cache 
@@
 			is added to cache (line 15). 
-			Finally, the disk write is simulated with the remaining amount of 
-			data (line 16), and the algorithm finishes with the total file write simulation time.
+			Finally, the disk write is simulated with the remaining data amount (line 16), 
+			and the chunk write algorithm finishes.
 
@@
 			This enables us to evaluate the accuracy and correctness of our 
-			model in a simple scenario before integrating it in the more complex SimGrid environment. 
+			model in a simple scenario before integrating it to the more complex WRENCH framework. 
 			In this prototype we use the following basic storage model, used for both memory and disk: 
@@
 			Having our model validated, we create simulators for different use cases 
-			using original SimGrid and our SimGridExt version. 
-			To simulate our pipelines and system, we leverage 
-			the Workflow Management System Simulation Workbench 
-			(WRENCH~\cite{wrench2018}), a framework built over SimGrid.
-		    We compare the results of the simulators with
-			original SimGrid and extended SimGrid with the results of real
+			using WRENCH.
+			We integrate our I/O model to the original WRENCH by implementing 
+			the core of the model separately from existing WRENCH components, 
+			then providing an API for users to call to our simulation model. 
+			We also add an optional command line argument so that users will be able 
+			to choose use or not to use the model. Doing this, we can avoid 
+			unexpected impacts on original WRENCH and can independently update 
+			the model in the future. 
+			We then compare the results of the simulators with
+			original WRENCH and extended WRENCH with the results of real
 			pipelines on a real system.
 
-			\textcolor{red}{[Description, implementation of our SimGridExt]}.
-
-			We use Python 3.7 to implement the simple simulator, SimGrid 3.25
-			and Wrench 1.6 for SimGrid simulators. SimGrid source code is
-			available at \url{https://framagit.org/simgrid/simgrid}, and WRENCH
-			is at \url{https://github.com/wrench-project/wrench}.
+			In this work, we use Python 3.7 to implement the single-threaded simulator, 
+			SimGrid 3.25 and WRENCH 1.6 for more multi-threaded simulators. 
+			SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid}, 
+			and WRENCH is at \url{https://github.com/wrench-project/wrench}.
 
@@
 			We simulate the pipeline with the Python prototype simulator, 
-			a simulator implemented with original SimGrid and a simulator implemented 
-			with SimGrid extended with our model.
+			a simulator with original WRENCH and another with WRENCH-Ext.
 
@@
 			Because implementation of a multi-threaded simulator in Python is 
-			expensive, and we have our model integrated in SimGrid, 
-			we only compare the results of simulators with original SimGrid, 
-			SimGrid integrated with our model and and real pipelines in this 
-			experiment. The results are compared in total makespan of the pipelines, 
+			expensive, and we have our model implemented in WRENCH, 
+			we only compare the results of simulators with original WRENCH, 
+			WRENCH-Ext and real pipelines in this experiment. 
+			The results are compared in total makespan of the pipelines, 
 			cumulative read time and cumulative write time.
@@
 
-			We simulate this pipeline with a simulator with original SimGrid and 
-			extended SimGrid. 
+			We simulate this pipeline with simulators using original WRENCH and 
+			WRENCH-Ext. 
 			The simulation results are then compared with the real results. 
@@
 					\item Relative errors of simulation time and memory used compared to real results.
-					\item Simulation time compared to baseline SimGrid.
+					\item Simulation time compared to SimGrid and original WRENCH.
 				\end{itemize}
@@
 		\begin{itemize}
+			\item Read order: disk, inactive list, then active list. 
+			\item Cache eviction mechanism not evicting active list, leads to evicting file being written.
 			\item Sensitivity of the simulator on the variation of memory and disk bandwidth. 

### 2020-10-08 result/multi/process_result.py
Export svg figures

@@ -89,9 +89,12 @@ def plot_prop(propname, title, average=False):
 
     plt.legend()
     if average:
-        plt.savefig("figures/%s_avg.pdf" % propname)
+        plt.savefig("figures/%s_avg.pdf" % propname, format="pdf")
+        plt.savefig("figures/%s_avg.svg" % propname, format="svg")
     else:
-        plt.savefig("figures/%s.pdf" % propname)
+        plt.savefig("figures/%s.pdf" % propname, format="pdf")
+        plt.savefig("figures/%s.svg" % propname, format="svg")
+
     plt.show()
 
 

### 2020-10-08 result/single/evaluate.py

@@ -56,26 +56,26 @@ def task_time_error(realtime_logfile, simtime_logfile):
     return time_acc
 
 
-def mem_error(real_time_logfile, sim_time_logfile, real_mem_logfile, sim_mem_logfile):
-    real_time_log = log_parse.read_timelog(real_time_logfile, skip_header=False)
-    real_mem_log = log_parse.read_atop_log(real_mem_logfile)
-
-    real_dirty_amt = [amt for amt in get_atop_mem_prop(real_time_log, real_mem_log, "dirty_data")]
-    real_cache_amt = [amt for amt in get_atop_mem_prop(real_time_log, real_mem_log, "cache")]
-
-    print("real:")
-    print(get_atop_mem_prop(real_time_log, real_mem_log, "dirty_data"))
-
-    sim_time_log = log_parse.read_timelog(sim_time_logfile)
-    sim_mem_log = log_parse.read_sim_log(sim_mem_logfile)
-
-    sim_dirty_amt = get_sim_mem_prop(sim_time_log, sim_mem_log, "dirty")
-    sim_cache_amt = get_sim_mem_prop(sim_time_log, sim_mem_log, "cache")
-
-    print("sim:")
-    print(sim_dirty_amt)
-
-    dirty_err = [abs(sim - real) / real for real, sim in zip(real_dirty_amt, sim_dirty_amt)]
-    cache_err = [abs(sim - real) / real for real, sim in zip(real_cache_amt, sim_cache_amt)]
-
-    return dirty_err, cache_err
+# def mem_error(real_time_logfile, sim_time_logfile, real_mem_logfile, sim_mem_logfile):
+#     real_time_log = log_parse.read_timelog(real_time_logfile, skip_header=False)
+#     real_mem_log = log_parse.read_atop_log(real_mem_logfile)
+#
+#     real_dirty_amt = [amt for amt in get_atop_mem_prop(real_time_log, real_mem_log, "dirty_data")]
+#     real_cache_amt = [amt for amt in get_atop_mem_prop(real_time_log, real_mem_log, "cache")]
+#
+#     print("real:")
+#     print(get_atop_mem_prop(real_time_log, real_mem_log, "dirty_data"))
+#
+#     sim_time_log = log_parse.read_timelog(sim_time_logfile)
+#     sim_mem_log = log_parse.read_sim_log(sim_mem_logfile)
+#
[…]

### 2020-10-08 result/single/log_parse.py

@@ -32,16 +32,16 @@ def read_atop_log(filename, dirty_ratio=0.2, dirty_bg_ratio=0.1):
         if line.startswith("MEM"):
             values = line.split(" ")
 
-            sys_mem_mb = int(values[7]) * 4096 / 1000 ** 2
+            sys_mem_mb = int(values[7]) * 4096 / 1000 ** 3
             sys_mem.append(sys_mem_mb)
 
-            cache_in_mb = int(values[9]) * 4096 / 1000 ** 2
+            cache_in_mb = int(values[9]) * 4096 / 1000 ** 3
             cache_used.append(cache_in_mb)
 
-            dirty_amt_mb = int(values[12]) * 4096 / 1000 ** 2
+            dirty_amt_mb = int(values[12]) * 4096 / 1000 ** 3
             dirty_data.append(dirty_amt_mb)
 
-            free_mem_mb = int(values[8]) * 4096 / 1000 ** 2
+            free_mem_mb = int(values[8]) * 4096 / 1000 ** 3
             free_mem.append(free_mem_mb)
 
             used_mem.append(sys_mem_mb - free_mem_mb)
@@ -77,10 +77,10 @@ def read_sim_log(filename):
 
         for line in csv_reader:
             time.append(float(line[0]))
-            total.append(float(line[1]))
-            dirty.append(float(line[2]))
-            cache.append(float(line[3]))
-            used.append(float(line[4]))
+            total.append(float(line[1]) / 1000)
+            dirty.append(float(line[2]) / 1000)
+            cache.append(float(line[3]) / 1000)
+            used.append(float(line[4]) / 1000)
 
     return {
         "time": time,

### 2020-10-08 result/single/plot_sim.py

@@ -92,8 +92,7 @@ def compare_sep(real_time_file, real_mem_file,
                 pysim_time_file, pysim_mem_file,
                 simgrid_time_file, simgrid_mem_file,
                 size, title, xmin, xmax, ymin, ymax):
-    figure = plt.figure()
-    plt.tight_layout()
+    figure = plt.figure(num=None, figsize=(8, 8), dpi=100, facecolor='w', edgecolor='k')
 
     ax1 = figure.add_subplot(3, 1, 1)
     ax2 = figure.add_subplot(3, 1, 2, sharex=ax1)
@@ -106,10 +105,15 @@ def compare_sep(real_time_file, real_mem_file,
     sim_subplot(ax2, pysim_time_file, pysim_mem_file, "python simulator", bar_alpha=0.2)
     sim_subplot(ax3, simgrid_time_file, simgrid_mem_file, "SimGrid simulator", bar_alpha=0.2)
 
-    plt.ylim(top=ymax, bottom=ymin)
+    # plt.ylim(top=ymax, bottom=ymin)
     plt.xlim(right=xmax, left=xmin)
-    # plt.text(1, 200000, text, fontsize=9)
-    plt.subplots_adjust(left=0.1, bottom=0.05, right=0.95, top=0.95, wspace=0, hspace=0.25)
+
+    figure.text(0.5, 0.02, "time (sec)", ha="center", va="center")
+    figure.text(0.04, 0.5, "memory (GB)", ha="center", va="center", rotation=90)
+
+    plt.subplots_adjust(left=0.12, bottom=0.07, right=0.95, top=0.95, wspace=0, hspace=0.25)
+    plt.savefig("figures/single_comp_sep_%dgb.svg" % size, format="svg")
+    plt.savefig("figures/single_comp_sep_%dgb.pdf" % size, format="pdf")
     plt.show()
 
 
@@ -123,6 +127,9 @@ def compare_overlap(real_time_file, real_mem_file,
 
     real_subplot(ax1, real_time_file, real_mem_file, xmin, xmax, ymin, ymax, alpha=0.2)
     sim_subplot(ax1, sim_time_file, sim_mem_file, "", alpha=0.3)
+
+    figure.xlabel("time (sec)")
+    figure.ylabel("Memory (GB)")
 
     plt.ylim(top=ymax, bottom=ymin)
     plt.xlim(right=xmax, left=xmin)
[…]

### 2020-10-12 paper_io_sim.tex
Export svg figures
Write results of experiment 1, 2

@@
 \usepackage{hyperref}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 			We use \textbf{\textit{atop}} and \textbf{\textit{collectl}} as tools to 
-			monitor and collect data of memory, page cache status and disk throughput 
+			monitor and collect memory status and disk throughput 
 			on the cluster. 
@@
 
+		We use the disk and memory benchmark results and CPU time measured from 
+		the cluster to configure the simulated platform. The benchmark results of 
+		bandwidths and configured values in the simulators are shown in \ref{table:benchmark}.
+
+		\begin{table}
+		\centering
+		\begin{tabular}{|l|c|c|c|}
+		\hline
+			Bandwidths (MBps)  & Cluster & Python & WRENCH\\
+		\hline
+			Memory read  & 6860	& 6860	 & 2765\\
+			Memory write & 2765	& 2765 & 2765\\
+			Disk read & ~510 & 465 & 465\\
+			Disk write & ~420 & 465	 & 465\\
+		\hline
+		\end{tabular}
+		\caption{Bandwidth benchmark results and simulator configurations}
+		\label{table:benchmark}
+		\end{table}
+
+		In the Python simulator, we use different read and write bandwidths for both 
+		memory and disk as in cluster. However, as read and write bandwidths are 
+		symmetrical in SimGrid, we configure memory with write bandwidth 
+		and disk with average bandwidth.
+
+		\subsection{Single-threaded experiment}
+
+		\subsubsection{Simulated time}
+
+			\begin{figure*}
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/error_20.pdf}
+   				\caption{20GB input file}	\label{fig:single_error_20}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/error_50.pdf}
+   				\caption{50GB input file}	\label{fig:single_error_50}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/error_75.pdf}
+   				\caption{75GB input file}	\label{fig:single_error_75}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/error_100.pdf}
+   				\caption{100GB input file}	\label{fig:single_error_100}
+			\end{subfigure}	
+			\caption{Relative error of individual simulated tasks}
+			\label{fig:single_error}
+			\end{figure*}
+
+			The I/O time of tasks in both simulators is tracked, and then compared 
+			to the I/O time in the real pipeline execution. 
+			Figure~\ref{fig:single_error} shows the relative error of simulated I/O time of 
+			individual I/O activities in simulators with different input sizes. 
+			There are 3 tasks in the pipeline and each task reads an input file, 
+			performs some computation, and writes output file. 
+			Thus, we have 6 sequential I/O activities, \textit{read1} and \textit{write1} 
+			from the first task, \textit{read2} and \textit{write2} the second task, 
+			and \textit{read3}, \textit{write3} from the last task.
+
+			In general, with all input sizes used in the experiment, the simulation errors 
+			of our simulators, both Python and WRENCH-Ext are remarkably lower 
+			in comparison with that of the original WRENCH simulator . 
+			The reason for this is, in original WRENCH, because page cache is not simulated, 
+			all data is only read/written with disk bandwidth, which makes simulated 
+			I/O time longer than in reality.
+			Especially, in our simulators, the errors of the Python simulator are significantly 
+			higher than that of the WRENCH-Ext simulator. 
+			This difference may attribute to the limitation in the periodical flushing 
+			mechanism simulation in Python. 
+			In WRENCH, it can be easily implemented with a simulated background thread, 
+			while there is only one thread simulated in our Python simulator, which 
+			hinders the correctness of this mechanism. 
+			Noticeably, the simulation time errors of \textit{read3} in the Python and 
+			WRENCH-Ext simulators are just under the error in original WRENCH. 
+			We will give our explanation for this later in this section.
+
+		\subsubsection{Memory behavior}
+
+			\begin{figure*}
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_20gb.pdf}
+   				\caption{20GB input file}	\label{fig:single_memprof_20gb}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_50gb.pdf}
+   				\caption{50GB input file}	\label{fig:single_memprof_50gb}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_75gb.pdf}
+   				\caption{75GB input file}	\label{fig:single_memprof_75gb}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_100gb.pdf}
+   				\caption{100GB input file}	\label{fig:single_memprof_100gb}
+			\end{subfigure}	
+			\caption{Memory profiling results with different input file sizes}
+			\label{fig:single_memprof}
+			\end{figure*}
+
+			Memory behaviors of real pipeline execution and simulators are also recorded 
+			and the memory profiling results are illustrated in Figure~\ref{fig:single_memprof}, 
+			with the bars presenting activities in tasks (read, compute and write) 
+			and the lines describing memory status throughout pipeline run time.
+
+			As shown in the figures, the makespans of the simulated pipeline are 
+			close to the makespan of the real pipeline. 
+			But more importantly, with all input file sizes, the trends of the lines describing 
+			memory in simulators are very similar to observations in the real execution. 
+			The similar trends in the amount of cache used (purple lines) in both reality 
+			and simulation shows that the use of page cache is well reproduced in the simulators. 
+			Besides, the red lines, which represent dirty data in cache, also 
+			follow similar trends. That means not only the amount dirty data written to 
+			cache is simulated, but also the flushing and periodical flushing 
+			mechanisms are accurately modeled in our simulators. 
+			This can be seen in the decrease in the amount of dirty data during computation 
+			time as well as read time. 
+			In addition, the effect of \textit{dirty\_data}  threshold is also simulated, 
+			which is reflected the interactions between dirty data (red line) and dirty\_ratio 
+			(black dashed line).
+
+			\begin{figure*}
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/fincore_real.pdf}
+   				\caption{Real pipeline execution}	
+   				\label{fig:fincore_real}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/fincore_wrench.pdf}
+   				\caption{WRENCH Simulator}	
+   				\label{fig:fincore_wrench}
+			\end{subfigure}	
+			\caption{Amount of file data in cache after each I/O activity with 100GB of input}
+			\label{fig:fincore}
+			\end{figure*}
+
+			Again, with 100GB input file, memory behavior in the third read in the 
+			simulators is different from reality, causing unexpected long read time, 
+			which we have seen in Figure~\ref{fig:single_error_100}. 
+			This can be explained with the amount of cached data of files at the beginning 
+			of the read shown in Figure~\ref{fig:fincore}. 
+			In real pipeline execution, after \textit{write2}, which writes \textit{file3}, 
+			all data of \textit{file3} is in cache (Figure~\ref{fig:fincore_real}). 
+			Thus, in \textit{read3}, \textit{file3} is read with memory bandwidth. 
+			But in our simulator, after \textit{write2}, only a part of \textit{file3} 
+			remains in cache (Figure~\ref{fig:fincore_wrench}). 
+			The reason is while \textit{file3}  is being written, the page cache is saturated, 
+			and eviction is evoked to evict data from page cache, make space 
+			available for \textit{file3} data. 
+			However, because the amount of evictable memory is less than required amount, 
+			some \textit{file3} data is evicted after being written, make \textit{file3} 
+			partially cached. As a result, in \textit{read3}, only a part of \textit{file3} is read 
+			from cache with memory bandwidth, a part of it is read from disk with disk read 
+			bandwidth, which is many times slower, leading to longer read time.
+
+			In summary, in single-threaded applications, our model can accurately 
+			simulate I/O time as well as the internal memory behavior. 
+			The model not only proves its applicability, but also proves that it is on 
+			a right track and can be further improve to achieve higher accuracy.
+
+		\subsection{Multi-threaded experiment}
+
+			\begin{figure*}
+			\begin{subfigure}{\linewidth}
+				\centering
+   				\includegraphics[width=0.5\columnwidth]{result/multi/figures/makespan.pdf}
+   				\caption{Total makespan}	
+   				\label{fig:multi_makespan}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/multi/figures/readtime.pdf}
+   				\caption{Cumulative read time}	
+   				\label{fig:multi_readtime}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/multi/figures/readtime_avg.pdf}
+   				\caption{Average read time}	
+   				\label{fig:multi_readtime_avg}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/multi/figures/writetime.pdf}
+   				\caption{Cumulative write time}	
+   				\label{fig:multi_writetime}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/multi/figures/writetime_avg.pdf}
+   				\caption{Average write time}	
+   				\label{fig:multi_writetime_avg}
+			\end{subfigure}	
+			\caption{Total makespan and I/O time with different number of concurrent pipeline}
+			\label{fig:multi_result}
+			\end{figure*}
+
+			In this experiment, we vary the number of 	concurrent pipeline from 1 to 32 
+			and record total makespan, cumulative and average read/write time 
+			of the pipelines on the cluster as well as from original WRENCH and 
+			WRENCH-Ext simulators. 
+			The results are presented in Figure~\ref{fig:multi_result}.
+
+			The Figure~\ref{fig:multi_makespan} shows that as the number of pipelines 
+			increases, the total makespan raises accordingly because the amount of data 
+			increases but the memory and disk bandwidths are constant and shared between 
+			pipelines. 
+			In comparison with original WRENCH, the result from WRENCH-Ext are closer to 
+			reality because it is implemented with our I/O with page cache model.
+
+			If we focus on read time only, both cumulative read time (Figure~\ref{fig:multi_readtime}) 
+			and average read time (Figure~\ref{fig:multi_readtime_avg}) 
+			follow the same trend in total makespan, and the simulated read time 
+			of WRENCH-Ext simulator is closer to reality than the read time in original WRENCH. 
+			In addition, the average read time seems to increase linearly in all cases when 
+			the bandwidth is shared by pipelines. Again, the result from WRENCH-Ext is more 
+			accurate than that from original WRENCH.
+
+			Similarly, the cumulative write time (shown in Figure~\ref{fig:multi_writetime}) 
+			and the average write time (shown in Figure~\ref{fig:multi_writetime_avg}) have 
+			the same increasing trend and the result from WRENCH-Ext are more realistic than 
+			the result from original WRENCH. 
+			Especially, the average write time in both reality and WRENCH-Ext have a similar 
+			pattern, when they gradually increase before a surge linearly with the same slope 
+			after the number of pipelines reaches 10. 
+			This can be explained that with less than 10 concurrent pipelines, the page cache 
+			is still not saturated, so all files can be written entirely to cache with memory 
+			bandwidth in a short time. 
+			After the page cache is saturated with dirty data (after around 10 concurrent pipelines 
+			in this experiment), this dirty data needs to be flushed in order to make space 
+			available for new data being written to cache. This dirty data is flushed to 
+			disk with disk bandwidth, which is many times slower than memory bandwidth. 
+			The more pipelines we have, the more data needs to be written, the more data 
+			needs to be flushed, leading to a sharper increase in the average write time.
+
+			To conclude, the results from WRENCH-Ext shows that our model can simulate I/O 
+			with page cache accurately in a multi-threaded application.
+
+		\subsection{Remote storage}
+
+		\subsection{Real application}
+
 	\section{Discussion and Future Work}

### 2020-10-12 exp/pysim/app.py

@@ -26,7 +26,7 @@ def export_time(task_list, filename):
 
 
 mm = MemoryManager(268600, 268600, read_bw=6860, write_bw=2765)
-storage = Storage(450000, read_bw=465, write_bw=465)
+storage = Storage(450000, read_bw=510, write_bw=420)
 kernel = IOManager(mm, storage, dirty_ratio=0.4)
 
 input_size = 100000

### 2020-10-12 result/multi/process_result.py

@@ -76,13 +76,13 @@ def plot_prop(propname, title, average=False):
     plt.title(title)
 
     if average:
-        plt.plot(real_df["no_pipeline"], real_df[propname] / real_df["no_pipeline"], label="real")
+        plt.plot(real_df["no_pipeline"], real_df[propname] / real_df["no_pipeline"], label="reality")
         plt.plot(simg_org_df["no_pipeline"], simg_org_df[propname] / simg_org_df["no_pipeline"], label="original WRENCH")
-        plt.plot(simg_ext_df["no_pipeline"], simg_ext_df[propname] / simg_ext_df["no_pipeline"], label="extended WRENCH")
+        plt.plot(simg_ext_df["no_pipeline"], simg_ext_df[propname] / simg_ext_df["no_pipeline"], label="WRENCH-Ext")
     else:
-        plt.plot(real_df["no_pipeline"], real_df[propname], label="real")
+        plt.plot(real_df["no_pipeline"], real_df[propname], label="reality")
         plt.plot(simg_org_df["no_pipeline"], simg_org_df[propname], label="original WRENCH")
-        plt.plot(simg_ext_df["no_pipeline"], simg_ext_df[propname], label="extended WRENCH")
+        plt.plot(simg_ext_df["no_pipeline"], simg_ext_df[propname], label="WRENCH-Ext")
 
     plt.xlabel("number of pipelines")
     plt.ylabel("time (s)")
@@ -98,11 +98,11 @@ def plot_prop(propname, title, average=False):
     plt.show()
 
 
-# export_simgrid_result("wrench_org", "aggregated.csv")
-# export_simgrid_result("wrench_ext", "aggregated.csv")
-#
-# plot_prop("makespan", "total makespan")
-# plot_prop("readtime", "cumulative read time")
-# plot_prop("readtime", "average read time", average=True)
-# plot_prop("writetime", "cumulative write time")
-# plot_prop("writetime", "average write time", average=True)
+export_simgrid_result("wrench_org", "aggregated.csv")
+export_simgrid_result("wrench_ext", "aggregated.csv")
+
+plot_prop("makespan", "Total makespan")
+plot_prop("readtime", "Cumulative read time")
+plot_prop("readtime", "Average read time", average=True)
+plot_prop("writetime", "Cumulative write time")
+plot_prop("writetime", "Average write time", average=True)

### 2020-10-12 result/single/plot_error.py

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 
-def grouped_bar_chart(labels, title, xlabel, ylabel, *argv):
+def grouped_bar_chart(labels, size, xlabel, ylabel, *argv):
     x = np.arange(len(labels))  # the label locations
     width = 0.2  # the width of the bars
     bars = len(argv)
@@ -15,12 +15,13 @@ def grouped_bar_chart(labels, title, xlabel, ylabel, *argv):
 
     ax.set_xlabel(xlabel)
     ax.set_ylabel(ylabel)
-    ax.set_title(title)
+    ax.set_title("Simulation time error with %dGB input file" % size)
     ax.set_xticks(x)
     ax.set_xticklabels(labels)
-    ax.set_ylim(bottom=0, top=8)
+    ax.set_ylim(bottom=0, top=800)
     ax.legend()
-
+    plt.savefig("figures/error_%d.svg" % size, format="svg")
+    plt.savefig("figures/error_%d.pdf" % size, format="pdf")
     plt.show()
 
 
@@ -33,15 +34,15 @@ def plot_task_error(size):
     simgrid_org_log = "wrench_org/%dgb_sim_time.csv" % size
     simgrid_ext_log = "wrench_ext/%dgb_sim_time.csv" % size
 
-    py_error = evaluate.task_time_error(real_time_log, sim_py_log)
-    simgrid_org_error = evaluate.task_time_error(real_time_log, simgrid_org_log)
-    simgrid_ext_error = evaluate.task_time_error(real_time_log, simgrid_ext_log)
+    py_error = [item * 100 for item in evaluate.task_time_error(real_time_log, sim_py_log)]
+    simgrid_org_error = [item * 100 for item in evaluate.task_time_error(real_time_log, simgrid_org_log)]
+    simgrid_ext_error = [item * 100 for item in evaluate.task_time_error(real_time_log, simgrid_ext_log)]
 
-    grouped_bar_chart(labels, "Simulation error with %dGB" % size, "tasks", "error",
+    grouped_bar_chart(labels, size, "", "error (%)",
                       ("Python", py_error), ("Original WRENCH", simgrid_org_error),
-                      ("Extended WRENCH", simgrid_ext_error))
+                      ("WRENCH-Ext", simgrid_ext_error))
 
 
-# sizes = [20, 50, 75, 100]
-# for size in sizes:
[…]

### 2020-10-12 result/single/plot_sim.py

@@ -102,8 +102,8 @@ def compare_sep(real_time_file, real_mem_file,
     real_subplot(ax1, real_time_file, real_mem_file, xmin, xmax, ymin, ymax, bar_alpha=0.2, linestyle="-", linewidth=1)
 
     # PYSIM RESULTS
-    sim_subplot(ax2, pysim_time_file, pysim_mem_file, "python simulator", bar_alpha=0.2)
-    sim_subplot(ax3, simgrid_time_file, simgrid_mem_file, "SimGrid simulator", bar_alpha=0.2)
+    sim_subplot(ax2, pysim_time_file, pysim_mem_file, "Python", bar_alpha=0.2)
+    sim_subplot(ax3, simgrid_time_file, simgrid_mem_file, "WRENCH-Ext", bar_alpha=0.2)
 
     # plt.ylim(top=ymax, bottom=ymin)
     plt.xlim(right=xmax, left=xmin)

### 2020-10-12 paper_io_sim.tex
Figure explanation

@@
 
-\author{\IEEEauthorblockN{Hoang-Dung Do, Tristan Glatard, Val\'erie Hayot-Sasson
+\author{\IEEEauthorblockN{Hoang-Dung Do, Val\'erie Hayot-Sasson, Tristan Glatard
   }\\
@@
 
-			\subsubsection{Evaluation on storage types}
+			\subsubsection{Evaluation on remote storage}
 
-		    Since different storage types should also be taken into account, 
-		    we conduct another experiment with the same scenario as 
-		    in the multi-threaded experiment, except that files are stored on a 
-		    shared file system instead of local disk.
-
-		    \textcolor{red}{[Description of Lustre on the cluster and how it's 
-		    simulated]}.
+		    Since remote storage is common in HPC, we conduct another experiment with 
+		    the same scenario as in the multi-threaded experiment, except that files 
+		    are stored on a node and read by another node.
 
@@
 			of the read shown in Figure~\ref{fig:fincore}. 
+			In this figure, \textit{file\_i} is the input of \textit{read\_i} and output of \textit{write\_i-1}. 
 			In real pipeline execution, after \textit{write2}, which writes \textit{file3}, 

### 2020-10-19 paper_io_sim.tex
Update paper

--- 2020-09-21 paper_io_sim.tex
@@
 
+		We use the disk and memory benchmark results and CPU time measured from 
+		the cluster to configure the simulated platform. The benchmark results of 
+		bandwidths and configured values in the simulators are shown in \ref{table:benchmark}.
+
+		\begin{table}
+		\centering
+		\begin{tabular}{|l|c|c|c|}
+		\hline
+			Bandwidths (MBps)  & Cluster (real) & Python simulator & WRENCH simulator\\
+		\hline
+			Memory read  & 6860	& 4812	 & 4812\\
+			Memory write & 2764	& 4812 & 4812\\
+			Disk read & ~510 & 465 & 465\\
+			Disk write & ~420 & 465	 & 465\\
+		\hline
+		\end{tabular}
+		\caption{Bandwidth benchmark results and simulator configurations}
+		\label{table:benchmark}
+		\end{table}
+
+		As read and write bandwidths are symmetrical in SimGrid, we take the average values 
+		to configure our simulators.
+
+		\subsection{Single-threaded experiment}
+
+			Memory behaviors of real pipeline execution and simulators are recorded 
+			and the memory profiling results with 20GB and 100GB input files 
+			are illustrated in Figure~\ref{fig:single_memprof_20gb} and Figure~\ref{fig:single_memprof_100gb}. 
+			The bars in the figures present activities in tasks (read, compute and write), 
+			while the lines describe memory status throughout run time of the pipeline. 
+			Because the results with 20Gb, 50GB, 100GB are analogous, we only include 
+			the results with 20GB and 100GB in this paper.
+
+			As shown in the figures, the makespans of the simulated pipeline are 
+			close to the makespan of the real execution with similar patterns of I/O time 
+			real execution and simulation. 
+			But more importantly, with all input sizes, similar trends can also be found 
+			in the lines describing memory status in simulators and the real execution. 
+			The akin movements of amount of cache used (purple lines) in reality 
+			and simulation show that the use of page cache is well reproduced by the simulators. 
+			The amount of dirty data ( red lines) also follow similar trends, 
+			which means not only the amount dirty data, but also the flushing and 
+			periodical flushing mechanisms are accurately modeled in our simulators. 
+			This can be seen in in the increase of the amount of dirty data during the writes 
+			and the decrease during the reads and computations. 
+			According to the results achieved with all input sizes in the experiment, 
+			we can observe identical trends in reality and simulation with 20GB, 
+			50GB and 75GB input files. However, with 100GB input file, 
+			the third read is not correctly simulated, leading to longer simulated time 
+			than in real execution. This will be explained later in this section.
+
+			The I/O time of tasks in simulators and real execution is also tracked and compared. 
+			Figure~\ref{fig:single_error_20} and Figure~\ref{fig:single_error_100} show 
+			the relative error of simulated I/O time of individual I/O activities. 
+			Our pipeline consists of 3 tasks, and each task reads an input file, 
+			performs some computation, and writes output file. 
+			Thus, there are 6 sequential I/O activities, \textit{read1} and \textit{write1} 
+			from the first task, \textit{read2} and \textit{write2} the second task, 
+			and \textit{read3}, \textit{write3} from the last task.
+
+			In general, with all input sizes in the experiment, the simulation errors 
+			of simulators with page cache model (Python and WRENCH with page cache simulators) 
+			are remarkably lower than the result with original WRENCH . 
+			The reason is, in original WRENCH, because page cache is not simulated, 
+			data is only read/written with disk bandwidth, which is much slower 
+			than memory bandwidths, leading to simulated I/O time longer than in reality.
+			Especially, the simulation time errors of \textit{read3} in the simulators 
+			with page cache are just under the error in original WRENCH.
+
+			\begin{figure*}
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_20gb.pdf}
+   				\caption{Memory profiling with 20GB input file}	\label{fig:single_memprof_20gb}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_100gb.pdf}
+   				\caption{Memory profiling with 100GB input file}	\label{fig:single_memprof_100gb}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/error_20.pdf}
+   				\caption{Simulation error with 20GB input file}	\label{fig:single_error_20}
+			\end{subfigure}		
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/error_100.pdf}
+   				\caption{Simulation error with 100GB input file}	\label{fig:single_error_100}
+			\end{subfigure}	
+			\caption{Simulation results with different input file sizes}
+			\label{fig:single_results}
+			\end{figure*}
+
+			\begin{figure*}
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/fincore_real.pdf}
+   				\caption{Real pipeline execution}	
+   				\label{fig:fincore_real}
+			\end{subfigure}	
+			\begin{subfigure}{\columnwidth}
+				\centering
+   				\includegraphics[width=\columnwidth]{result/single/figures/fincore_wrench.pdf}
+   				\caption{WRENCH with page cache simulator}	
+   				\label{fig:fincore_wrench}
+			\end{subfigure}	
+			\caption{Amount of file data in cache after each I/O activity with 100GB of input}
+			\label{fig:fincore}
+			\end{figure*}
+
+			This longer read can be explained with the amount of cached data of files 
+			at the beginning of the read in reality and simulators shown in Figure~\ref{fig:fincore}. 			
+			In this figure, \textit{file$<$i$>$} is the input of \textit{read$<$i$>$} and 
+			output of \textit{write$<$i-1$>$}. 
+			In the real execution, after \textit{write2}, which writes \textit{file3}, 
+			all data of \textit{file3} is in cache (Figure~\ref{fig:fincore_real}). 
+			Thus, in \textit{read3}, \textit{file3} is read with memory bandwidth. 
+			But in our simulator, after \textit{write2}, only a part of \textit{file3} 
+			remains in cache (Figure~\ref{fig:fincore_wrench}). 
+			The reason is while \textit{file3}  is being written, the page cache is saturated, 
+			eviction is then evoked to evict data from page cache, make space 
+			available for \textit{file3}. 
+			However, because the evicted amount is less than required, 
+			some \textit{file3} data is evicted after being written, make \textit{file3} 
+			partially cached. 
+			As a result, in \textit{read3}, only a part of \textit{file3} is read from cache 
+			with memory bandwidth, a part of it is read from disk with disk read bandwidth, 
+			which is many times slower, leading to longer read time.
+
+			In summary, in this single-threaded experiment, our model can accurately 
+			simulate I/O time as well as the internal memory behavior. 
+			This can help confirm that our model is on a right track and 
+			it can be further improved to achieve higher accuracy.
+
+		\subsection{Multi-threaded experiment}
+
+			\begin{figure*}
+			\begin{subfigure}{0.33\linewidth}
+				\centering
+   				\includegraphics[width=\linewidth]{result/multi/figures/makespan.pdf}
+   				\caption{Total makespan}	
+   				\label{fig:multi_makespan}
+			\end{subfigure}	
+			\begin{subfigure}{0.33\linewidth}
+				\centering
+   				\includegraphics[width=\linewidth]{result/multi/figures/readtime.pdf}
+   				\caption{Average read time}	
+   				\label{fig:multi_readtime}
+			\end{subfigure}	
+			\begin{subfigure}{0.33\linewidth}
+				\centering
+   				\includegraphics[width=\linewidth]{result/multi/figures/writetime.pdf}
+   				\caption{Average write time}	
+   				\label{fig:multi_writetime}
+			\end{subfigure}		
+			\caption{Total makespan and I/O time with different number of concurrent pipeline}
+			\label{fig:multi_result}
+			\end{figure*}
+
+			In this experiment, the results is analyzed in terms of the total makespan  
+			and average I/O time concurrent pipelines.
+			The results from real execution and simulation are presented in Figure~\ref{fig:multi_result}.
+
+			The Figure~\ref{fig:multi_makespan} shows that as the number of pipelines 
+			increases, the total makespan raises accordingly because the amount of data 
+			increases but the memory and disk bandwidths are fixed and shared between pipelines. 
+			In comparison with original WRENCH, the result from WRENCH with page cache are closer 
+			to reality. 
+			This proximity can also be seen in average read time shown in 
+			Figure~\ref{fig:multi_readtime}. 
+			Especially, when it comes to the average write time 
+			(shown in Figure~\ref{fig:multi_writetime}), 
+			although the trend in reality is more complex, our WRENCH with page cache 
+			simulator can still capture it and share the same pattern with 
+			the real execution, when they gradually increase before surging 
+			with the same slope after the number of pipelines reaches 10. 
+			This can be explained that with less than 10 concurrent pipelines, 
+			the page cache is not saturated, so all files can be written entirely to 
+			cache with memory bandwidth in a short time. 
+			After the page cache is saturated with dirty data (at around 
+			10 concurrent pipelines in this experiment), this dirty data needs 
+			to be flushed in order to make space available for writing new data to cache. 
+			This dirty data is flushed to disk with disk bandwidth, which is much  
+			slower than memory bandwidth. 
+			The more pipelines we have, the more data needs to be written, the more data 
+			needs to be flushed, leading to a sharper increase in the average write time.
+
+			To conclude, the results from WRENCH with page cache simulator show that 
+			our model can also simulate I/O with page cache accurately in a 
+			multi-threaded application.
+
+		\subsection{Remote storage}
+
+		\subsection{Real application}
+
 	\section{Discussion and Future Work}

### 2020-10-19 result/single/plot_fincore.py

@@ -5,17 +5,34 @@ columns = ["read1", "write1", "read2", "write2", "read3", "write3"]
 files = ["file1", "file2", "file3", "file4"]
 df_real = pd.read_csv("fincore/real.csv", index_col="task")
 df_sim = pd.read_csv("fincore/sim.csv", index_col="task")
+patterns = ["", "...", "///", "\\\\\\",]
 
-# fig, axes = plt.subplots(nrows=1, ncols=2, figsize=[10,5])
+# axe = df_real[files].plot.bar(stacked=True, rot=0, color="tab:purple", edgecolor="k")
+axe = df_sim[files].plot.bar(stacked=True, rot=0, color="tab:purple", edgecolor="k")
 
-df_real[files].plot.bar(stacked=True, rot=0, title="Real pipeline")
-# df_sim[files].plot.bar(stacked=True, rot=0, title="WRENCH simulator")
+bars = axe.patches
+hatches = []  # list for hatches in the order of the bars
+for h in patterns:  # loop over patterns to create bar-ordered hatches
+    for i in range(int(len(bars) / len(patterns))):
+        hatches.append(h)
+for bar, hatch in zip(bars, hatches):  # loop over bars and hatches to set hatches in correct order
+    bar.set_hatch(hatch)
 
 plt.subplots_adjust(left=0.12, bottom=0.15, right=0.92, top=0.9)
 
 plt.xlabel("")
 plt.ylabel("amount (GB)")
-plt.savefig("figures/fincore_real.svg", format="svg")
-plt.savefig("figures/fincore_real.pdf", format="pdf")
+
+legend = plt.legend()
+l_hatches = []
+legends = legend.axes.patches
+for h in patterns:  # loop over patterns to create bar-ordered hatches
+    for i in range(int(len(legends) / len(patterns))):
+        l_hatches.append(h)
+for bar, hatch in zip(legends, l_hatches):  # loop over bars and hatches to set hatches in correct order
+    bar.set_hatch(hatch)
+
+plt.savefig("figures/fincore_wrench.svg", format="svg")
+plt.savefig("figures/fincore_wrench.pdf", format="pdf")
 
 plt.show()

### 2020-10-19 result/single/plot_sim.py

@@ -101,9 +101,9 @@ def compare_sep(real_time_file, real_mem_file,
     # REAL RESULTS
     real_subplot(ax1, real_time_file, real_mem_file, xmin, xmax, ymin, ymax, bar_alpha=0.2, linestyle="-", linewidth=1)
 
-    # PYSIM RESULTS
-    sim_subplot(ax2, pysim_time_file, pysim_mem_file, "Python", bar_alpha=0.2)
-    sim_subplot(ax3, simgrid_time_file, simgrid_mem_file, "WRENCH-Ext", bar_alpha=0.2)
+    # SIMULATION RESULTS
+    sim_subplot(ax2, pysim_time_file, pysim_mem_file, "Python simulator", bar_alpha=0.2)
+    sim_subplot(ax3, simgrid_time_file, simgrid_mem_file, "WRENCH with page cache simulator", bar_alpha=0.2)
 
     # plt.ylim(top=ymax, bottom=ymin)
     plt.xlim(right=xmax, left=xmin)
@@ -148,12 +148,12 @@ def real_subplot(subplot_ax, real_time_file, real_mem_file, xmin, xmax, ymin, ym
 
     for i in range(len(timestamps)):
         if timestamps[i][0] == "read":
-            subplot_ax.axvspan(xmin=timestamps[i][1] - start, xmax=timestamps[i][2] - start, color="g", alpha=bar_alpha,
+            subplot_ax.axvspan(xmin=timestamps[i][1] - start, xmax=timestamps[i][2] - start, color="k", alpha=0.1,
                                label="read" if i == 0 else "")
         else:
-            subplot_ax.axvspan(xmin=timestamps[i - 1][2] - start, xmax=timestamps[i][1] - start, color="k", alpha=bar_alpha,
+            subplot_ax.axvspan(xmin=timestamps[i - 1][2] - start, xmax=timestamps[i][1] - start, color="k", alpha=0.25,
                                label="computation" if i == 1 else "")
-            subplot_ax.axvspan(xmin=timestamps[i][1] - start, xmax=timestamps[i][2] - start, color="b", alpha=bar_alpha,
+            subplot_ax.axvspan(xmin=timestamps[i][1] - start, xmax=timestamps[i][2] - start, color="k", alpha=0.4,
                                label="write" if i == 1 else "")
 
     # app_cache = list(np.array(app_mem) + np.array(cache_used))
[…]

### 2020-10-19 result/single/plot_fincore.py
Update cache plot and add notebook

@@ -1,38 +1,48 @@
 import pandas as pd
 import matplotlib.pyplot as plt
 
-columns = ["read1", "write1", "read2", "write2", "read3", "write3"]
-files = ["file1", "file2", "file3", "file4"]
-df_real = pd.read_csv("fincore/real.csv", index_col="task")
-df_sim = pd.read_csv("fincore/sim.csv", index_col="task")
-patterns = ["", "...", "///", "\\\\\\",]
 
-# axe = df_real[files].plot.bar(stacked=True, rot=0, color="tab:purple", edgecolor="k")
-axe = df_sim[files].plot.bar(stacked=True, rot=0, color="tab:purple", edgecolor="k")
+def plot_cache(file, source):
+    columns = ["read1", "write1", "read2", "write2", "read3", "write3"]
+    files = ["file1", "file2", "file3", "file4"]
 
-bars = axe.patches
-hatches = []  # list for hatches in the order of the bars
-for h in patterns:  # loop over patterns to create bar-ordered hatches
-    for i in range(int(len(bars) / len(patterns))):
-        hatches.append(h)
-for bar, hatch in zip(bars, hatches):  # loop over bars and hatches to set hatches in correct order
-    bar.set_hatch(hatch)
+    if source == "real":
+        df = pd.read_csv("fincore/real.csv", index_col="task")
+    if source == "wrench":
+        df = pd.read_csv("fincore/sim.csv", index_col="task")
 
-plt.subplots_adjust(left=0.12, bottom=0.15, right=0.92, top=0.9)
+    patterns = ["", "...", "///", "\\\\\\"]
 
-plt.xlabel("")
-plt.ylabel("amount (GB)")
+    # axe = df_real[files].plot.bar(stacked=True, rot=0, color="tab:purple", edgecolor="k")
+    axe = df[files].plot.bar(stacked=True, rot=0, color="tab:purple", edgecolor="k")
 
-legend = plt.legend()
-l_hatches = []
-legends = legend.axes.patches
-for h in patterns:  # loop over patterns to create bar-ordered hatches
-    for i in range(int(len(legends) / len(patterns))):
-        l_hatches.append(h)
-for bar, hatch in zip(legends, l_hatches):  # loop over bars and hatches to set hatches in correct order
-    bar.set_hatch(hatch)
+    bars = axe.patches
+    hatches = []  # list for hatches in the order of the bars
[…]

### 2020-10-20 paper_io_sim.tex
Add simulated platform descriptions

@@
 
-			Having our model validated, we create simulators for different use cases 
+			Having the model validated, we create simulators for different use cases 
 			using WRENCH.
@@
 
-        	The goal of our experiments is to evaluate our write-back cache 
+        	The goal of our experiments is to evaluate our page cache 
 			simulation in single-threaded and multi-threaded applications
@@
 
+			In simulated platform description, because the disk bandwidths in SimGrid 
+			are symmetrical, we measure the read/write bandwidths on our cluster and 
+			take the average bandwidths. 
+			As the pipelines are single-core, we create simulated tasks, 
+			each of them runs on only one core. 
+			While in Python, computation time can be directly configured, 
+			this time is injected to the WRENCH simulators with number of flops, 
+			which is calculated with CPU speed and compute duration. 
+			In our experiments, the CPU speed is set to 1 Gflops. 
+			Finally, because WRENCH simulates applications base on network-communication, 
+			we use an infinite bandwidth to eliminate network latency for local I/Os.
+			The details of simulated platform is available here: 
+			https://github.com/dohoangdzung/wrench/blob/io_chunk_test/examples/basic-examples/io-pagecache/single_host.xml
+
 			\subsubsection{Single-threaded evaluation}
 
-			The simplest execution scenario is single-threaded, where memory 
-			bandwidth and disk bandwidth are dedicated. Our goal is to 
-			compare the detailed simulation results of each task in the pipeline 
-			including read/write time, memory status and then validate the correctness 
-			of the simulation model.
+			The first experiment is single-threaded, where memory bandwidth 
+			and disk bandwidth are dedicated. 
+			Our goal is to compare the detailed simulation results of each task 
+			in the pipeline including read/write time, memory behavior and 
+			then validate the correctness of the simulation model.
 
@@
 			to local disk. 
-			We use the input sizes of 20~GB, 50~GB, 75~GB and 100~GB, run 
-			the pipeline on the cluster, measure the CPU time of the tasks  
-			with each input and use it to simulate CPU time in the simulators .
-
-			We simulate the pipeline with the Python prototype simulator, 
-			a simulator with original WRENCH and another with WRENCH-Ext.
+			We use the input sizes of 20~GB, 50~GB, 75~GB, 100~GB and run 
+			the pipeline on the cluster.
+			The pipeline is simulated with the Python prototype simulator, 
+			original WRENCH and WRENCH with page cache simulator.
 
@@
 
-			In the second experiment, we evaluate the simulator in a concurrent I/O, 
+			In the second scenario, we evaluate the simulator in a concurrent I/O, 
 			bandwidth-sharing scenario with a multi-threaded application. 			
@@
 
-		We use the disk and memory benchmark results and CPU time measured from 
-		the cluster to configure the simulated platform. The benchmark results of 
-		bandwidths and configured values in the simulators are shown in \ref{table:benchmark}.
+		The benchmark results of bandwidths and the values used in the simulators are shown in \ref{table:benchmark}. These configurations are used in all experiments.
 
@@
 		\end{table}
-
-		As read and write bandwidths are symmetrical in SimGrid, we take the average values 
-		to configure our simulators.
 

### 2020-10-23 paper_io_sim.tex
Update figures

@@
 			The details of simulated platform is available here: 
-			https://github.com/dohoangdzung/wrench/blob/io_chunk_test/examples/basic-examples/io-pagecache/single_host.xml
+			\url{https://github.com/dohoangdzung/wrench/blob/io_chunk_test/examples/basic-examples/io-pagecache/single_host.xml}
 
@@
 			and the memory profiling results with 20GB and 100GB input files 
-			are illustrated in Figure~\ref{fig:single_memprof_20gb} and Figure~\ref{fig:single_memprof_100gb}. 
+			are illustrated in Figure~\ref{fig:single_memprof}. 
 			The bars in the figures present activities in tasks (read, compute and write), 
@@
 			The I/O time of tasks in simulators and real execution is also tracked and compared. 
-			Figure~\ref{fig:single_error_20} and Figure~\ref{fig:single_error_100} show 
-			the relative error of simulated I/O time of individual I/O activities. 
+			Figure~\ref{fig:single_errors} shows the relative error of simulated I/O time 
+			of individual I/O activities. 
 			Our pipeline consists of 3 tasks, and each task reads an input file, 
@@
 			\begin{figure*}
+			\centering
 			\begin{subfigure}{\columnwidth}
@@
    				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_20gb.pdf}
-   				\caption{Memory profiling with 20GB input file}	\label{fig:single_memprof_20gb}
+   				\caption{20GB input file}	\label{fig:single_memprof_20gb}
 			\end{subfigure}	
@@
    				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_100gb.pdf}
-   				\caption{Memory profiling with 100GB input file}	\label{fig:single_memprof_100gb}
+   				\caption{100GB input file}	\label{fig:single_memprof_100gb}
 			\end{subfigure}	
-			\begin{subfigure}{\columnwidth}
+			\caption{Memory profiling results with different input file sizes}
+			\label{fig:single_memprof}	
+			\begin{subfigure}{\linewidth}
 				\centering
-   				\includegraphics[width=\columnwidth]{result/single/figures/error_20.pdf}
-   				\caption{Simulation error with 20GB input file}	\label{fig:single_error_20}
+   				\includegraphics[width=\linewidth]{result/single/figures/errors.pdf}
 			\end{subfigure}
-			\begin{subfigure}{\columnwidth}
+			\caption{Simulation errors with different input file sizes}
+			\label{fig:single_error}	
+			\begin{subfigure}{\linewidth}
 				\centering
-   				\includegraphics[width=\columnwidth]{result/single/figures/error_100.pdf}
-   				\caption{Simulation error with 100GB input file}	\label{fig:single_error_100}
+   				\includegraphics[width=\linewidth]{result/single/figures/cached_files.pdf}
 			\end{subfigure}	
-			\caption{Simulation results with different input file sizes}
-			\label{fig:single_results}
-			\end{figure*}
-
-			\begin{figure*}
-			\begin{subfigure}{\columnwidth}
-				\centering
-   				\includegraphics[width=\columnwidth]{result/single/figures/fincore_real.pdf}
-   				\caption{Real pipeline execution}	
-   				\label{fig:fincore_real}
-			\end{subfigure}	
-			\begin{subfigure}{\columnwidth}
-				\centering
-   				\includegraphics[width=\columnwidth]{result/single/figures/fincore_wrench.pdf}
-   				\caption{WRENCH with page cache simulator}	
-   				\label{fig:fincore_wrench}
-			\end{subfigure}	
+			\label{fig:single_cache}
    			\caption{Amount of file data in cache after each I/O activity with 100GB of input}
-			\label{fig:fincore}
 			\end{figure*}
@@
 			This longer read can be explained with the amount of cached data of files 
-			at the beginning of the read in reality and simulators shown in Figure~\ref{fig:fincore}. 			
+			at the beginning of the read in reality and simulators shown in 
+			Figure~\ref{fig:single_cache}. 			
 			In this figure, \textit{file$<$i$>$} is the input of \textit{read$<$i$>$} and 
@@
 			\begin{figure*}		
-			\begin{subfigure}{0.33\linewidth}
+			\begin{subfigure}{\linewidth}
 				\centering
-   				\includegraphics[width=\linewidth]{result/multi/figures/makespan.pdf}
-   				\caption{Total makespan}	
-   				\label{fig:multi_makespan}
+   				\includegraphics[width=\linewidth]{result/multi/figures/multi_local.pdf}
 			\end{subfigure}		
-			\begin{subfigure}{0.33\linewidth}
-				\centering
-   				\includegraphics[width=\linewidth]{result/multi/figures/readtime.pdf}
-   				\caption{Average read time}	
-   				\label{fig:multi_readtime}
-			\end{subfigure}	
-			\begin{subfigure}{0.33\linewidth}
-				\centering
-   				\includegraphics[width=\linewidth]{result/multi/figures/writetime.pdf}
-   				\caption{Average write time}	
-   				\label{fig:multi_writetime}
-			\end{subfigure}		
-			\caption{Total makespan and I/O time with different number of concurrent pipeline}
-			\label{fig:multi_result}
+			\caption{I/O time with different numbers of concurrent pipeline}
+			\label{fig:multi_local}
 			\end{figure*}
 
-			In this experiment, the results is analyzed in terms of the total makespan  
-			and average I/O time concurrent pipelines.
+			In this experiment, the results is analyzed with the average I/O time of concurrent pipelines.
 			The results from real execution and simulation are presented in Figure~\ref{fig:multi_result}.
 
-			The Figure~\ref{fig:multi_makespan} shows that as the number of pipelines 
-			increases, the total makespan raises accordingly because the amount of data 
-			increases but the memory and disk bandwidths are fixed and shared between pipelines. 
+			The Figure~\ref{fig:multi_readtime} shows that as the number of pipelines 
+			increases, the average read time raises accordingly because the amount of data 
+			increases but memory and disk bandwidths are fixed and shared between pipelines. 
 			In comparison with original WRENCH, the result from WRENCH with page cache are closer 
 			to reality. 
-			This proximity can also be seen in average read time shown in 
-			Figure~\ref{fig:multi_readtime}. 
 			Especially, when it comes to the average write time 

### 2020-10-23 result/multi/process_result.py

@@ -17,9 +17,10 @@ def parse_single_pipeline(filename):
     return min(df["read_start"]), max(df["write_end"]), readtime, writetime
 
 
-def aggregate_result(dir, no_pipeline):
+def aggregate_result(folder, no_pipeline):
     """
 
+    :param folder:
     :param no_pipeline:
     :return: makespan, total_readtime, total_writetime
     """
@@ -28,7 +29,7 @@ def aggregate_result(dir, no_pipeline):
     readtime = 0
     writetime = 0
     for i in range(no_pipeline):
-        filename = "%s/indv_logs/time_pipeline_%d_%d.csv" % (dir, no_pipeline, i + 1)
+        filename = "%s/time_pipeline_%d_%d.csv" % (folder, no_pipeline, i + 1)
         start, end, read, write = parse_single_pipeline(filename)
         makespan += end - start
         readtime += read
@@ -37,21 +38,22 @@ def aggregate_result(dir, no_pipeline):
     return no_pipeline, makespan, readtime, writetime
 
 
-def export_real_results(filename):
-    with open(filename, 'w', newline='') as csvfile:
+def export_real_results(folder, filename):
+    file = "%s%s" % (folder, filename)
+    with open(file, 'w', newline='') as csvfile:
         writer = csv.writer(csvfile)
         writer.writerow(["no_pipeline", "makespan", "readtime", "writetime"])
         for i in range(32):
-            writer.writerow(list(aggregate_result(i + 1)))
+            writer.writerow(list(aggregate_result(folder, i + 1)))
 
 
-def export_simgrid_result(dir, filename):
-    exported_file = "%s/%s" % (dir, filename)
+def export_simgrid_result(folder, filename):
+    exported_file = "%s%s" % (folder, filename)
     with open(exported_file, 'w', newline='') as csvfile:
         writer = csv.writer(csvfile)
         writer.writerow(["no_pipeline", "makespan", "readtime", "writetime"])
         for i in range(32):
-            dump_file = "%s/dump_%d.json" % (dir, i + 1)
+            dump_file = "%s/dump_%d.json" % (folder, i + 1)
             writer.writerow(list(parse_simgrid_result(dump_file, i + 1)))
 
 
@@ -67,38 +69,35 @@ def parse_simgrid_result(filename, no_pipeline):
[…]

### 2020-10-23 result/single/plot_cache.py

@@ -1,0 +1,59 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+
+
+def plot_cache(real_file, sim_file):
+    columns = ["read_1", "write_1", "read#2", "write_2", "read_3", "write_3"]
+    files = ["file_1", "file_2", "file_3", "file_4"]
+
+    df_real = pd.read_csv(real_file, index_col="task")
+    df_sim = pd.read_csv(sim_file, index_col="task")
+
+    patterns = ["", "...", "///", "\\\\\\"]
+
+    plt.rcParams.update({'font.size': 8})
+    fig, (ax1, ax2) = plt.subplots(figsize=(11, 4), ncols=2, nrows=1)
+    plt.subplots_adjust(left=0.12, bottom=0.1, right=0.92, top=0.8)
+
+    ax1 = df_real[files].plot.bar(stacked=True, rot=0, color="tab:purple", edgecolor="k", ax=ax1,
+                                  title="Real execution")
+    ax2 = df_sim[files].plot.bar(stacked=True, rot=0, color="tab:purple", edgecolor="k", ax=ax2,
+                                 title="WRENCH with page cache", legend=False)
+
+    ax1.set_xlabel("tasks")
+    ax2.set_xlabel("tasks")
+    ax1.set_ylabel("amount (GB)")
+    ax2.set_ylabel("amount (GB)")
+
+    bars = ax1.patches
+    hatches = []  # list for hatches in the order of the bars
+    for h in patterns:  # loop over patterns to create bar-ordered hatches
+        for i in range(int(len(bars) / len(patterns))):
+            hatches.append(h)
+    for bar, hatch in zip(bars, hatches):  # loop over bars and hatches to set hatches in correct order
+        bar.set_hatch(hatch)
+
+    bars = ax2.patches
+    hatches = []  # list for hatches in the order of the bars
+    for h in patterns:  # loop over patterns to create bar-ordered hatches
+        for i in range(int(len(bars) / len(patterns))):
+            hatches.append(h)
+    for bar, hatch in zip(bars, hatches):  # loop over bars and hatches to set hatches in correct order
+        bar.set_hatch(hatch)
+
+    legend = ax1.legend(loc='upper center', bbox_to_anchor=(1, 1.2), ncol=4)
+    l_hatches = []
+    legends = legend.axes.patches
+    for h in patterns:  # loop over patterns to create bar-ordered hatches
[…]

### 2020-10-23 result/single/plot_error.py

@@ -3,12 +3,11 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 
-def grouped_bar_chart(labels, size, xlabel, ylabel, *argv):
+def grouped_bar_chart(ax, labels, size, xlabel, ylabel, *argv):
     x = np.arange(len(labels))  # the label locations
     width = 0.2  # the width of the bars
     bars = len(argv)
 
-    fig, ax = plt.subplots()
     for i in range(len(argv)):
         arg = argv[i]
         ax.bar(x + (2 * i + 1 - bars) * width / 2, arg[1], width, label=arg[0], color=arg[2])
@@ -17,15 +16,11 @@ def grouped_bar_chart(labels, size, xlabel, ylabel, *argv):
     ax.set_ylabel(ylabel)
     ax.set_xticks(x)
     ax.set_xticklabels(labels)
-    ax.set_ylim(bottom=0, top=800)
-    if size == 20: ax.legend()
-    plt.savefig("figures/error_%d.svg" % size, format="svg")
-    plt.savefig("figures/error_%d.pdf" % size, format="pdf")
-    plt.show()
+    ax.set_ylim(bottom=0, top=700)
 
 
-def plot_task_error(size):
-    labels = ["read1", "write1", "read2", "write2", "read3", "write3"]
+def plot_task_error(ax, size):
+    labels = ["read_1", "write_1", "read_2", "write_2", "read_3", "write_3"]
 
     # atop_file =     "log/cluster/100gb/atop_mem.log"
     real_time_log = "real/%dgb/timestamps.csv" % size
@@ -37,11 +32,20 @@ def plot_task_error(size):
     simgrid_org_error = [item * 100 for item in evaluate.task_time_error(real_time_log, simgrid_org_log)]
     simgrid_ext_error = [item * 100 for item in evaluate.task_time_error(real_time_log, simgrid_ext_log)]
 
-    grouped_bar_chart(labels, size, "", "error (%)",
-                      ("Python simulator", py_error, "tab:blue"), ("Original WRENCH", simgrid_org_error, 'tab:orange'),
-                      ("WRENCH simulator with page cache", simgrid_ext_error, 'tab:green'))
+    grouped_bar_chart(ax, labels, size, "", "error (%)",
[…]

### 2020-10-23 result/single/plot_fincore.py

@@ -1,48 +1,0 @@
-import pandas as pd
-import matplotlib.pyplot as plt
-
-
-def plot_cache(file, source):
-    columns = ["read1", "write1", "read2", "write2", "read3", "write3"]
-    files = ["file1", "file2", "file3", "file4"]
-
-    if source == "real":
-        df = pd.read_csv("fincore/real.csv", index_col="task")
-    if source == "wrench":
-        df = pd.read_csv("fincore/sim.csv", index_col="task")
-
-    patterns = ["", "...", "///", "\\\\\\"]
-
-    # axe = df_real[files].plot.bar(stacked=True, rot=0, color="tab:purple", edgecolor="k")
-    axe = df[files].plot.bar(stacked=True, rot=0, color="tab:purple", edgecolor="k")
-
-    bars = axe.patches
-    hatches = []  # list for hatches in the order of the bars
-    for h in patterns:  # loop over patterns to create bar-ordered hatches
-        for i in range(int(len(bars) / len(patterns))):
-            hatches.append(h)
-    for bar, hatch in zip(bars, hatches):  # loop over bars and hatches to set hatches in correct order
-        bar.set_hatch(hatch)
-
-    plt.subplots_adjust(left=0.12, bottom=0.15, right=0.92, top=0.9)
-
-    plt.xlabel("")
-    plt.ylabel("amount (GB)")
-
-    legend = plt.legend()
-    l_hatches = []
-    legends = legend.axes.patches
-    for h in patterns:  # loop over patterns to create bar-ordered hatches
-        for i in range(int(len(legends) / len(patterns))):
-            l_hatches.append(h)
-    for bar, hatch in zip(legends, l_hatches):  # loop over bars and hatches to set hatches in correct order
-        bar.set_hatch(hatch)
-
-    plt.savefig("figures/fincore_%s.svg" % source, format="svg")
-    plt.savefig("figures/fincore_%s.pdf" % source, format="pdf")
-
-    plt.show()
-
-
-plot_cache("fincore/real.csv", "real")
-plot_cache("fincore/wrench.csv", "wrench")

### 2020-10-23 result/single/plot_memprof.py

@@ -1,0 +1,247 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import log_parse
+
+
+def plot_pysim_log(mem_log, time_stamps, text, xmin, xmax, ymin, ymax):
+    time = mem_log["time"]
+    total = mem_log["total"]
+    dirty = mem_log["dirty"]
+    cache = mem_log["cache"]
+    used = mem_log["used"]
+    # free = mem_log["free"]
+    # available = list(np.array(free) + np.array(cache) - np.array(dirty))
+    # dirty_ratio = list(np.array(available) * 0.4)
+    # dirty_bg_ratio = list(np.array(available) * 0.1)
+
+    read_starts = time_stamps["read_start"]
+    read_ends = time_stamps["read_end"]
+    write_starts = time_stamps["write_start"]
+    write_ends = time_stamps["write_end"]
+
+    plot_log(time, read_starts, read_ends, write_starts, write_ends, total, cache, dirty, used,
+             "python simulator result", xmax, xmin, ymax, ymin)
+
+
+def plot_sim_result(time_log_file, mem_log_file, title, xmin, xmax, ymin, ymax):
+    sim_time_log = log_parse.read_timelog(time_log_file)
+    sim_mem_log = log_parse.read_sim_log(mem_log_file)
+
+    read_starts = []
+    read_ends = []
+    write_starts = []
+    write_ends = []
+    for i in range(len(sim_time_log)):
+        if sim_time_log[i][0] == "read":
+            read_starts.append(sim_time_log[i][1])
+            read_ends.append(sim_time_log[i][2])
+        if sim_time_log[i][0] == "write":
+            write_starts.append(sim_time_log[i][1])
+            write_ends.append(sim_time_log[i][2])
+
+    time = sim_mem_log["time"]
+    total = sim_mem_log["total"]
+    cache = sim_mem_log["cache"]
+    dirty = sim_mem_log["dirty"]
+    used = sim_mem_log["used"]
+
+    plot_log(time, read_starts, read_ends, write_starts, write_ends, total, cache, dirty, used,
+             title, xmax, xmin, ymax, ymin)
+
+
+def plot_log(time, read_starts, read_ends, write_starts, write_ends, total, cache, dirty, used,
+             title, xmax, xmin, ymax, ymin):
+    free = list(np.array(total) - np.array(used))
[…]

### 2020-10-23 result/single/plot_cache.py
Update single-threaded exps notebooks

@@ -56,4 +56,4 @@ def plot_cache(real_file, sim_file):
     plt.show()
 
 
-plot_cache("fincore/real.csv", "fincore/sim.csv")
+# plot_cache("fincore/real.csv", "fincore/sim.csv")

### 2020-10-23 result/single/plot_error.py

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 
-def grouped_bar_chart(ax, labels, size, xlabel, ylabel, *argv):
+def grouped_bar_chart(ax, labels, xlabel, ylabel, *argv):
     x = np.arange(len(labels))  # the label locations
     width = 0.2  # the width of the bars
     bars = len(argv)
@@ -32,20 +32,20 @@ def plot_task_error(ax, size):
     simgrid_org_error = [item * 100 for item in evaluate.task_time_error(real_time_log, simgrid_org_log)]
     simgrid_ext_error = [item * 100 for item in evaluate.task_time_error(real_time_log, simgrid_ext_log)]
 
-    grouped_bar_chart(ax, labels, size, "", "error (%)",
+    grouped_bar_chart(ax, labels, "", "error (%)",
                       ("Python simulator", py_error, "tab:pink"), ("Original WRENCH", simgrid_org_error, 'tab:orange'),
                       ("WRENCH simulator with page cache", simgrid_ext_error, 'tab:cyan'))
     ax.set_title("%d GB" % size)
 
 
-plt.rcParams.update({'font.size': 8})
-fig, (ax1, ax2) = plt.subplots(figsize=(11, 3), ncols=2, nrows=1)
-plot_task_error(ax1, 20)
-plot_task_error(ax2, 100)
-
-lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.3), ncol=3)
-
-plt.savefig("figures/errors.svg", format="svg", bbox_extra_artists=(lgd,), bbox_inches='tight')
-plt.savefig("figures/errors.pdf", format="pdf", bbox_extra_artists=(lgd,), bbox_inches='tight')
-plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.3)
-plt.show()
+# plt.rcParams.update({'font.size': 8})
+# fig, (ax1, ax2) = plt.subplots(figsize=(11, 3), ncols=2, nrows=1)
+# plot_task_error(ax1, 20)
+# plot_task_error(ax2, 100)
+#
+# lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.3), ncol=3)
+#
+# plt.savefig("figures/single_errors.svg", format="svg", bbox_extra_artists=(lgd,), bbox_inches='tight')
[…]

### 2020-10-23 result/multi/process_result.py
Update multi-threaded exp notebooks

@@ -84,20 +84,20 @@ def plot_prop(ax, exp_folder, propname, title):
     ax.set_ylabel("time (s)")
 
 
-export_real_results("local/real/", "aggregated.csv")
-export_simgrid_result("local/wrench/original/", "aggregated.csv")
-export_simgrid_result("local/wrench/pagecache/", "aggregated.csv")
-
-plt.rcParams.update({'font.size': 8})
-fig, (ax1, ax2) = plt.subplots(figsize=(10, 5), ncols=2, nrows=1)
-
-plot_prop(ax1, "local", "readtime", "average read time")
-plot_prop(ax2, "local", "writetime", "average write time")
-
-plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.3), ncol=3)
-plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.3)
-
-plt.savefig("figures/multi_local.pdf", format="pdf")
-plt.savefig("figures/multi_local.svg", format="svg")
-
-plt.show()
+# export_real_results("local/real/", "aggregated.csv")
+# export_simgrid_result("local/wrench/original/", "aggregated.csv")
+# export_simgrid_result("local/wrench/pagecache/", "aggregated.csv")
+#
+# plt.rcParams.update({'font.size': 8})
+# fig, (ax1, ax2) = plt.subplots(figsize=(10, 5), ncols=2, nrows=1)
+#
+# plot_prop(ax1, "local", "readtime", "average read time")
+# plot_prop(ax2, "local", "writetime", "average write time")
+#
+# plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.3), ncol=3)
+# plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.3)
+#
+# plt.savefig("figures/multi_local.pdf", format="pdf")
+# plt.savefig("figures/multi_local.svg", format="svg")
+#
+# plt.show()

### 2020-10-26 paper_io_sim.tex
Update single-threaded exps notebooks
Update memprof figures and README

@@
 			\centering
-			\begin{subfigure}{\columnwidth}
+			\begin{subfigure}{\linewidth}
 				\centering
-   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_20gb.pdf}
-   				\caption{20GB input file}	\label{fig:single_memprof_20gb}
-			\end{subfigure}	
-			\begin{subfigure}{\columnwidth}
-				\centering
-   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_100gb.pdf}
-   				\caption{100GB input file}	\label{fig:single_memprof_100gb}
+   				\includegraphics[width=\linewidth]{result/single/figures/single_memprof.pdf}
 			\end{subfigure}
@@
 				\centering
-   				\includegraphics[width=\linewidth]{result/single/figures/errors.pdf}
+   				\includegraphics[width=\linewidth]{result/single/figures/single_errors.pdf}
 			\end{subfigure}
@@
 
-			In this experiment, the results is analyzed with the average I/O time of concurrent pipelines.
-			The results from real execution and simulation are presented in Figure~\ref{fig:multi_result}.
+			In this experiment, the results are analyzed in terms of average I/O time 
+			of concurrent pipelines shown in Figure~\ref{fig:multi_local}.
 
-			The Figure~\ref{fig:multi_readtime} shows that as the number of pipelines 
-			increases, the average read time raises accordingly because the amount of data 
+			As is shown in the figure, when the number of pipelines increases, 
+			the average read time raises accordingly because the amount of data 
 			increases but memory and disk bandwidths are fixed and shared between pipelines. 
@@
 			to reality. 
-			Especially, when it comes to the average write time 
-			(shown in Figure~\ref{fig:multi_writetime}), 
+			Especially, when it comes to the average write time, 
 			although the trend in reality is more complex, our WRENCH with page cache 
-			simulator can still capture it and share the same pattern with 
-			the real execution, when they gradually increase before surging 
-			with the same slope after the number of pipelines reaches 10. 
+			simulator can still capture the pattern in real execution.
+			In both, the average write time gradually increases before surging with the 
+			same slopes after the number of pipelines reaches 10. 
 			This can be explained that with less than 10 concurrent pipelines, 

### 2020-10-26 result/single/plot_error.py

@@ -38,12 +38,12 @@ def plot_task_error(ax, size):
     ax.set_title("%d GB" % size)
 
 
-# plt.rcParams.update({'font.size': 8})
-# fig, (ax1, ax2) = plt.subplots(figsize=(11, 3), ncols=2, nrows=1)
-# plot_task_error(ax1, 20)
-# plot_task_error(ax2, 100)
-#
-# lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.3), ncol=3)
+plt.rcParams.update({'font.size': 8})
+fig, (ax1, ax2) = plt.subplots(figsize=(11, 3), ncols=2, nrows=1)
+plot_task_error(ax1, 20)
+plot_task_error(ax2, 100)
+
+lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.3), ncol=3)
 #
 # plt.savefig("figures/single_errors.svg", format="svg", bbox_extra_artists=(lgd,), bbox_inches='tight')
 # plt.savefig("figures/single_errors.pdf", format="pdf", bbox_extra_artists=(lgd,), bbox_inches='tight')

### 2020-10-26 result/single/plot_memprof.py

@@ -3,137 +3,22 @@ import numpy as np
 import log_parse
 
 
-def plot_pysim_log(mem_log, time_stamps, text, xmin, xmax, ymin, ymax):
-    time = mem_log["time"]
-    total = mem_log["total"]
-    dirty = mem_log["dirty"]
-    cache = mem_log["cache"]
-    used = mem_log["used"]
-    # free = mem_log["free"]
-    # available = list(np.array(free) + np.array(cache) - np.array(dirty))
-    # dirty_ratio = list(np.array(available) * 0.4)
-    # dirty_bg_ratio = list(np.array(available) * 0.1)
-
-    read_starts = time_stamps["read_start"]
-    read_ends = time_stamps["read_end"]
-    write_starts = time_stamps["write_start"]
-    write_ends = time_stamps["write_end"]
-
-    plot_log(time, read_starts, read_ends, write_starts, write_ends, total, cache, dirty, used,
-             "python simulator result", xmax, xmin, ymax, ymin)
-
-
-def plot_sim_result(time_log_file, mem_log_file, title, xmin, xmax, ymin, ymax):
-    sim_time_log = log_parse.read_timelog(time_log_file)
-    sim_mem_log = log_parse.read_sim_log(mem_log_file)
-
-    read_starts = []
-    read_ends = []
-    write_starts = []
-    write_ends = []
-    for i in range(len(sim_time_log)):
-        if sim_time_log[i][0] == "read":
-            read_starts.append(sim_time_log[i][1])
-            read_ends.append(sim_time_log[i][2])
-        if sim_time_log[i][0] == "write":
-            write_starts.append(sim_time_log[i][1])
-            write_ends.append(sim_time_log[i][2])
-
-    time = sim_mem_log["time"]
-    total = sim_mem_log["total"]
-    cache = sim_mem_log["cache"]
-    dirty = sim_mem_log["dirty"]
-    used = sim_mem_log["used"]
-
-    plot_log(time, read_starts, read_ends, write_starts, write_ends, total, cache, dirty, used,
-             title, xmax, xmin, ymax, ymin)
-
-
-def plot_log(time, read_starts, read_ends, write_starts, write_ends, total, cache, dirty, used,
-             title, xmax, xmin, ymax, ymin):
-    free = list(np.array(total) - np.array(used))
[…]

### 2020-10-27 gen_fig.py
Add one notebook to generate all figures

@@ -1,0 +1,16 @@
+import os
+import sys
+sys.path.append(os.path.abspath("result/single/"))
+sys.path.append(os.path.abspath("result/multi/"))
+
+os.chdir("result/single")
+import result.single.plot_memprof as plot_memprof
+import result.single.plot_error as plot_error
+import result.single.plot_cache as plot_cache
+plot_memprof.compare_plot([20, 100], [200, 1300])
+plot_error.plot_error()
+plot_cache.plot_cache("fincore/real.csv", "fincore/sim.csv")
+
+os.chdir("../multi")
+import result.multi.process_result as multi_result
+multi_result.aggrerate_and_plot()

### 2020-10-27 result/multi/process_result.py

@@ -69,7 +69,7 @@ def parse_simgrid_result(filename, no_pipeline):
     return no_pipeline, makespan, read, write
 
 
-def plot_prop(ax, exp_folder, propname, title):
+def suplot_prop(ax, exp_folder, propname, title):
     real_df = pd.read_csv("%s/real/aggregated.csv" % exp_folder)
     simg_org_df = pd.read_csv("%s/wrench/original/aggregated.csv" % exp_folder)
     simg_ext_df = pd.read_csv("%s/wrench/pagecache/aggregated.csv" % exp_folder)
@@ -77,27 +77,29 @@ def plot_prop(ax, exp_folder, propname, title):
     ax.set_title(title)
 
     ax.plot(real_df["no_pipeline"], real_df[propname] / real_df["no_pipeline"], color="k", label="Real execution")
-    ax.plot(simg_org_df["no_pipeline"], simg_org_df[propname] / simg_org_df["no_pipeline"], color="tab:orange", label="Original WRENCH")
-    ax.plot(simg_ext_df["no_pipeline"], simg_ext_df[propname] / simg_ext_df["no_pipeline"], color="tab:cyan",label="WRENCH with page cache")
+    ax.plot(simg_org_df["no_pipeline"], simg_org_df[propname] / simg_org_df["no_pipeline"], color="tab:orange",
+            label="Original WRENCH")
+    ax.plot(simg_ext_df["no_pipeline"], simg_ext_df[propname] / simg_ext_df["no_pipeline"], color="tab:cyan",
+            label="WRENCH with page cache")
 
     ax.set_xlabel("number of pipelines")
     ax.set_ylabel("time (s)")
 
 
-# export_real_results("local/real/", "aggregated.csv")
-# export_simgrid_result("local/wrench/original/", "aggregated.csv")
-# export_simgrid_result("local/wrench/pagecache/", "aggregated.csv")
-#
-# plt.rcParams.update({'font.size': 8})
-# fig, (ax1, ax2) = plt.subplots(figsize=(10, 5), ncols=2, nrows=1)
-#
-# plot_prop(ax1, "local", "readtime", "average read time")
-# plot_prop(ax2, "local", "writetime", "average write time")
-#
-# plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.3), ncol=3)
-# plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.3)
-#
[…]

### 2020-10-27 result/single/plot_error.py

@@ -38,14 +38,15 @@ def plot_task_error(ax, size):
     ax.set_title("%d GB" % size)
 
 
-plt.rcParams.update({'font.size': 8})
-fig, (ax1, ax2) = plt.subplots(figsize=(11, 3), ncols=2, nrows=1)
-plot_task_error(ax1, 20)
-plot_task_error(ax2, 100)
+def plot_error():
+    plt.rcParams.update({'font.size': 8})
+    fig, (ax1, ax2) = plt.subplots(figsize=(11, 3), ncols=2, nrows=1)
+    plot_task_error(ax1, 20)
+    plot_task_error(ax2, 100)
 
-lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.3), ncol=3)
-#
-# plt.savefig("figures/single_errors.svg", format="svg", bbox_extra_artists=(lgd,), bbox_inches='tight')
-# plt.savefig("figures/single_errors.pdf", format="pdf", bbox_extra_artists=(lgd,), bbox_inches='tight')
-# plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.3)
-# plt.show()
+    lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.3), ncol=3)
+
+    plt.savefig("figures/single_errors.svg", format="svg", bbox_extra_artists=(lgd,), bbox_inches='tight')
+    plt.savefig("figures/single_errors.pdf", format="pdf", bbox_extra_artists=(lgd,), bbox_inches='tight')
+    plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.3)
+    plt.show()

### 2020-10-27 paper_io_sim.tex
Update figures

@@
 
-\author{\IEEEauthorblockN{Hoang-Dung Do, Val\'erie Hayot-Sasson, Tristan Glatard
+\author{\IEEEauthorblockN{Hoang-Dung Do, Val\'erie Hayot-Sasson, Rafael Ferreira da Silva, Henri Casanova, Tristan Glatard
   }\\
@@
 			Figure~\ref{fig:single_cache}. 			
-			In this figure, \textit{file$<$i$>$} is the input of \textit{read$<$i$>$} and 
-			output of \textit{write$<$i-1$>$}. 
-			In the real execution, after \textit{write2}, which writes \textit{file3}, 
-			all data of \textit{file3} is in cache (Figure~\ref{fig:fincore_real}). 
-			Thus, in \textit{read3}, \textit{file3} is read with memory bandwidth. 
-			But in our simulator, after \textit{write2}, only a part of \textit{file3} 
-			remains in cache (Figure~\ref{fig:fincore_wrench}). 
-			The reason is while \textit{file3}  is being written, the page cache is saturated, 
+			In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and 
+			output of \textit{Write $<$i-1$>$}. 
+			In the real execution, after \textit{Write 2}, which writes \textit{file 3}, 
+			all data of \textit{file 3} is in cache. 
+			Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth. 
+			But in our simulator, after \textit{Write 2}, only a part of \textit{file 3} 
+			remains in cache. 
+			The reason is while \textit{file 3}  is being written, the page cache is saturated, 
 			eviction is then evoked to evict data from page cache, make space 
-			available for \textit{file3}. 
+			available for \textit{file 3}. 
 			However, because the evicted amount is less than required, 
-			some \textit{file3} data is evicted after being written, make \textit{file3} 
+			some \textit{file 3} data is evicted after being written, make \textit{file 3} 
 			partially cached. 
-			As a result, in \textit{read3}, only a part of \textit{file3} is read from cache 
+			As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache 
 			with memory bandwidth, a part of it is read from disk with disk read bandwidth, 

### 2020-10-27 result/single/plot_cache.py

@@ -3,8 +3,9 @@ import matplotlib.pyplot as plt
 
 
 def plot_cache(real_file, sim_file):
-    columns = ["read_1", "write_1", "read#2", "write_2", "read_3", "write_3"]
-    files = ["file_1", "file_2", "file_3", "file_4"]
+    index = ["file_1", "file_2", "file_3", "file_4"]
+    columns = ["Read 1", "Write 1", "Read 2", "Write 2", "Read 3", "Write 3"]
+    files = ["file 1", "file 2", "file 3", "file 4"]
 
     df_real = pd.read_csv(real_file, index_col="task")
     df_sim = pd.read_csv(sim_file, index_col="task")
@@ -15,15 +16,17 @@ def plot_cache(real_file, sim_file):
     fig, (ax1, ax2) = plt.subplots(figsize=(11, 4), ncols=2, nrows=1)
     plt.subplots_adjust(left=0.12, bottom=0.1, right=0.92, top=0.8)
 
-    ax1 = df_real[files].plot.bar(stacked=True, rot=0, color="tab:purple", edgecolor="k", ax=ax1,
+    ax1 = df_real[index].plot.bar(stacked=True, rot=0, color="tab:cyan", edgecolor="k", ax=ax1,
                                   title="Real execution")
-    ax2 = df_sim[files].plot.bar(stacked=True, rot=0, color="tab:purple", edgecolor="k", ax=ax2,
+    ax2 = df_sim[index].plot.bar(stacked=True, rot=0, color="tab:cyan", edgecolor="k", ax=ax2,
                                  title="WRENCH with page cache", legend=False)
 
     ax1.set_xlabel("tasks")
     ax2.set_xlabel("tasks")
     ax1.set_ylabel("amount (GB)")
     ax2.set_ylabel("amount (GB)")
+    ax1.set_xticklabels(columns)
+    ax2.set_xticklabels(columns)
 
     bars = ax1.patches
     hatches = []  # list for hatches in the order of the bars
@@ -41,7 +44,7 @@ def plot_cache(real_file, sim_file):
     for bar, hatch in zip(bars, hatches):  # loop over bars and hatches to set hatches in correct order
         bar.set_hatch(hatch)
 
-    legend = ax1.legend(loc='upper center', bbox_to_anchor=(1, 1.2), ncol=4)
+    legend = ax1.legend(files, loc='upper center', bbox_to_anchor=(1, 1.2), ncol=4)
     l_hatches = []
[…]

### 2020-10-27 result/single/plot_error.py

@@ -20,7 +20,7 @@ def grouped_bar_chart(ax, labels, xlabel, ylabel, *argv):
 
 
 def plot_task_error(ax, size):
-    labels = ["read_1", "write_1", "read_2", "write_2", "read_3", "write_3"]
+    labels = ["Read 1", "Write 1", "Read 2", "Write 2", "Read 3", "Write 3"]
 
     # atop_file =     "log/cluster/100gb/atop_mem.log"
     real_time_log = "real/%dgb/timestamps.csv" % size

### 2020-10-27 result/single/plot_memprof.py

@@ -34,24 +34,24 @@ def real_subplot(subplot_ax, real_time_file, real_mem_file, xmin, xmax, ymin, ym
     for i in range(len(timestamps)):
         if timestamps[i][0] == "read":
             subplot_ax.axvspan(xmin=timestamps[i][1] - start, xmax=timestamps[i][2] - start, color="k", alpha=0.1,
-                               label="read" if i == 0 else "")
+                               label="Read" if i == 0 else "")
         else:
             subplot_ax.axvspan(xmin=timestamps[i - 1][2] - start, xmax=timestamps[i][1] - start, color="k", alpha=0.25,
-                               label="computation" if i == 1 else "")
+                               label="Computation" if i == 1 else "")
             subplot_ax.axvspan(xmin=timestamps[i][1] - start, xmax=timestamps[i][2] - start, color="k", alpha=0.4,
                                label="write" if i == 1 else "")
 
     # app_cache = list(np.array(app_mem) + np.array(cache_used))
-    subplot_ax.plot(time, atop_log["total"], color='k', linewidth=1, linestyle=":", label="total memory",
+    subplot_ax.plot(time, atop_log["total"], color='k', linewidth=1, linestyle=":", label="Total memory",
                     alpha=line_alpha)
     subplot_ax.plot(time, atop_log["used_mem"], color='g', linewidth=linewidth, linestyle=linestyle,
-                    label="used memory", alpha=line_alpha)
-    subplot_ax.plot(time, atop_log["cache"], color='m', linewidth=linewidth, linestyle=linestyle,
-                    label="cache", alpha=line_alpha)
+                    label="Used memory", alpha=line_alpha)
+    subplot_ax.plot(time, atop_log["cache"], color='purple', linewidth=linewidth, linestyle=linestyle,
+                    label="Cache", alpha=line_alpha)
     subplot_ax.plot(time, atop_log["dirty_data"], color='r', linewidth=linewidth, linestyle=linestyle,
-                    label="dirty data", alpha=line_alpha)
+                    label="Dirty data", alpha=line_alpha)
     subplot_ax.plot(time, atop_log["avai_mem"], color='b', linewidth=linewidth, linestyle=linestyle,
-                    label="available memory", alpha=line_alpha)
+                    label="Available memory", alpha=line_alpha)
[…]

### 2020-11-16 paper_io_sim.tex
Draft version of the paper

@@
 
-\author{\IEEEauthorblockN{Hoang-Dung Do, Val\'erie Hayot-Sasson, Rafael Ferreira da Silva, Henri Casanova, Tristan Glatard
+\author{\IEEEauthorblockN{Hoang-Dung Do$^1$, Val\'erie Hayot-Sasson$^1$, Rafael Ferreira da Silva$^3$, Henri Casanova$^2$, Tristan Glatard$^1$
   }\\
   \IEEEauthorblockA{
-    Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada
+    $^1$Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada \\
+     $^2$Department of Information and Computer Sciences, University of Hawai‘i at Manoa, U.S.A \\
+     $^3$Department of Computer Science, University of Southern California, Los Angeles, CA, USA
   }
@@
 	\begin{abstract}
-		\begin{itemize}
-			\item The I/O bottleneck in HPC and the need of experiments.
-			\item HPC experiment frameworks and advantages of SimGrid.
-			\item The missing of the ability to simulate page cache, the goal of this paper.
-			\item Principle of the simulator, experiment scenarios and comparisons.
-			\item Brief discussion on results and future work.
-		\end{itemize}
+
+		The data explosion era seen in recent years results in growing needs in 
+		data storing, data analysis. 
+		Large computing systems including grids, clusters, clouds, peer-to-peer, 
+		volunteer computers have been increasingly used to solve the problem. 
+		Despite of their powerful computing capability and technological improvements, 
+		there are still bottlenecks in these systems. 
+		When it comes to I/O, a bottleneck can be seen in the disk I/O due to hardware 
+		limitations. Data caching using page cache is one of the solutions applied 
+		to minimize disk accesses and improve the I/O performance. 
+		In computing system and application optimization studies, one of the 
+		effective approaches is using simulation tools to conduct experiments instead 
+		of running applications on real infrastructures. 
+		Although simulators are built with individual simulation model for each computing 
+		resource, the majority does not take page cache into consideration, 
+		which can lead to low simulation accuracy in many cases.
+
+		In this paper, we propose a simulation model page cache to simulate I/O 
+		with key features of page cache. 
+		The model is build with the consideration to allow it to be integrated to 
+		existing simulation tools. 
+		We also implement the model with SimGrid using WRENCH. 
+		Then we create some simulators for different use cases. 
+		Finally, we analyze the results, compare them to results of the existing 
+		tool as well as real execution.
+
 	\end{abstract}
@@
 	\section{Introduction}
-		\begin{itemize}
-			\item HPC, the bottleneck in I/O and the demand of HPC experiments. 
-			\item Difficulties in conducting high performance computing experiments and the need of simulation frameworks.
-			\item Existing experiment methods, simulators, simulation frameworks.
-The advantages of SimGrid compared to others \cite{simgrid2008, lebre2015}.
-The missing of the ability to simulate page cache in SimGrid \cite{lebre2015}.
-			\item The objective of the paper: Add capability to simulate I/O with page cache in SimGrid.
-			\item Clarify, distinguish between memory I/O (i/o to page cache) and disk I/O.
-		\end{itemize}
+
+		In recent years, as a result of the growth of data generated from a variety 
+		of data sources, we have seen increasingly growing demands in data storage, 
+		processing and analysis. 
+		Storing and processing such a massive amount of data require large 
+		computing infrastructures. 
+		Data processing is a complex process whose the performance is affected 
+		by many factors and hindered by bottlenecks in the systems.
+		One of the bottlenecks that should be taken into consideration is disk access 
+		time due to hardware limitations. 
+		Beside hardware technical improvements, architectural solutions have been 
+		proposed and implemented in computing systems in order to tackle the 
+		disk I/O problem. 
+		Caching, a successful technique applied into memory to maximize the 
+		I/O performance by keeping data in memory for future accesses,
+		has been proving the effectiveness when employed in computer systems.
+
+		Studying in optimizing not only the utilization of page cache but also other 
+		computing resources is a need that has never been overlooked.
+		The research on computing infrastructures to maximize performance, 
+		shorten processing time, and minimize costs gets involved with experiments. 
+		However, real platforms are complex and not built for the experimental 
+		purposes due to limited and shared resources, operational costs, unstable 
+		environments and the result reproducibility difficulties. 
+		An alternative approach is to used simulation tools developed for research 
+		and experimental purposes.
+		The obvious advantages of using simulation are the freedom of choosing 
+		simulated infrastructures, reproducible results and low costs.
+		These simulators are developed and integrated with simulation models 
+		of computing resources including CPU, network and storage.
+		However, some reviews over many simulators report that there has been 
+		a trade-off between accuracy and scalability due to the level of details in 
+		simulation models \cite{casanova2014simgrid} \cite{byrne2017review}.
+		Moreover, different simulators are developed by and for a specific community,
+		which hinders them from being used in cross-domain.
+		SimGrid is a simulation framework developed with the objectives of  
+		balancing between accuracy and scalability, as well as versatility.
+
+		In spite of the importance of page cache and the impacts it has on I/O 
+		operations, its simulation model is absent in many simulators. 
+		Some simulation tools try to model the page cache with cached data, 
+		but their models lack of some key features \cite{nunez2012simcan} 
+		\cite{nunez2012icancloud}. 
+		Moreover, these simulators have a problem in scalability with 
+		long simulation time regarding to their very detailed level of simulation.
+		Some other works are able to capture some important features of page cache, 
+		but they focus on very specific domains \cite{xu2018saving}.
+
+		In this paper, we propose a simulation model of I/O with page cache.
+		The model is integrated in WRENCH, an open-source library built 
+		on top of SimGrid with more abstract API. 
+		Then we evaluate our model with four simulators in four 
+		experiment scenarios. 
+		This paper is organized as follows. 
+		In Section \ref{relatedwork} we provide details of page cache 
+		existing simulation tools.
+		In Section \ref{relatedwork} we propose our I/O with page 
+		cache simulation model and how we implement it and our experiments 
+		to evaluate the model.  
+		Section \ref{results} compares the results of our model with the results 
+		of original WRENCH and real execution results. 
+		Finally, Section \ref{discussion} concludes and presents future 
+		work.
+
 	\section{Related Work}		
+	\label{relatedwork}
 
 		\subsection{Page cache}
-			\begin{itemize}
-				\item What is page cache? How it works \cite{linuxdev3rd2010}. 
-				Write-back vs writethrough and direct i/o. Effects and importance of page cache.
-				\item Introduce some existing strategies with some highlighted pros and cons.
-				\item Current implementation in Linux and some reasons why it is chosen to be implemented (implementation complexity, effectiveness, overhead, etc) \cite{linuxdev3rd2010}
-				\item \tristan{LRU lists are mentioned in the next section, they should be described here.}
-			\end{itemize}
+
+			In Linux, page cache is a part of memory, which includes pages referring to 
+			physical pages on disk \cite{linuxdev3rd2010}. 
+			The goal of the page cache is to avoid disk I/O by keeping data in memory 
+			for later accesses, minimize I/O time since disk accesses are considerably 
+			slower than memory accesses.
+			When the kernel starts a read operation, it checks if required data is in page cache.
+			If yes, data is read from memory. This is called a \textit{cache hit}. 
+			Otherwise, data is read from disk and placed in page cache.
+			For writing, page cache is implemented with three different strategies. 
+			The first is \textit{direct~I/O} or \textit{no-write}, which means the page cache 
+			is not involved in write operations. 
+			The second is \textit{write-through}, in which data goes \textit{through} 
+			page cache to disk is. In this case, data is written to both page cache and disk. 
+			In the third strategy, which is \textit{write-back}, data are written directly 
+			to disk as \textit{dirty} pages and these pages will be periodically flushed 
+			to disk later by a \textit{writeback} process and then marked as not dirty. 
+			The write-back strategy is considered to outperform write-through as well as 
+			direct I/O since it delays disk writes and perform in batch later 
+			\cite{linuxdev3rd2010}.
+
+			In write-back, one of the keys is the cache eviction mechanism, which 
+			removes pages from the page cache when available memory is insufficient. 
+			The kernel chooses clean pages from page cache to evict. 
+			If the clean pages are inadequate, dirty pages are flushed to disk, 
+			marked as clean and available for eviction. 
+			Different cache eviction algorithms have been developed and proposed
+			\cite{owda2014comparison}.
+			In Linux kernel, a \textit{two lists strategy (LRU/2)} is implemented based on 
+			the least-recently-used (LRU) scheme, with an \textit{active list} only containing 
+			pages accesses more than once, and an \textit{inactive list} containing 
+			pages available for eviction. 
+			Pages are moved from inactive list to active list if they are 
+			accessed when residing in the inactive list. 
+			The lists work on LRU manner, pages are added to the tail and removed 
+			from the top. 
+			The lists are also kept balanced by moving pages from the active list
+			to the inactive list when the active list grows too large.
 
 		\subsection{Simulators}
-			\begin{itemize}
-				\item Discuss some existing methods, simulation frameworks to conduct HPC experiments. Compare pros and cons (accuracy, simulation time, usability) of some simulators (SimGrid, GridSim).
-				\item Related development: RAM energy consumption \cite{gill2019} \cite{ouarnoughi2017} 
-				\item What is implemented in simulators that useful for our model:
-				distinguished and shared bandwidths, concurrency, device latency. 
-				\item Discuss the pros of SimGrid and the reasons why we chose it to conduct experiments. (Section 2.2.2 in \cite{simgrid2014})
-				\item Introduce WRENCH and it's advantage.
-				\item \tristan{Mention the status of storage simulation in simulators, so that we understand from Fig 1 that storage devices are available}
-			\end{itemize}
+
+			There has been the research and development of many simulation tools 
+			and frameworks for cloud computing with the goal to simulate cloud, grid, 
+			HPC platforms, etc. 
+			A number of the most famous existing simulators have been analyzed 
+			and compared in some reviews (e.g., \cite{byrne2017review}, \cite{casanova2014simgrid}). 
+			These simulators provide simulation models of CPU, network, storage 
+			with different approaches and levels of details.
+			Some simulators are at packet-level, which allows to achieve 
+			high simulation accuracy with fine-grain details such as storage simulation 
+			in \textit{DiskSim} or network simulation in \textit{CloudSim} 
+			and \textit{iCanCloud} \cite{casanova2014simgrid}. 
+			However, the issues with the packet-level simulation are scalability due to 
+			long simulation time when the simulated platforms and applications grow 
+			to hundreds and thousands nodes. 
+			SimGrid is an open source simulation platform that has been developed 
+			not only to balance the trade-off between accuracy and scalability, 
+			but also to achieve versatility, which allows the use of the framework 
+			to simulate applications in multiple domains \cite{casanova2014simgrid}.
+
+			However, I/O with page cache model has not been provided in SimGrid. 
+			It only simulates storage with capacity and bandwidths.
+			Lebre \textit{et al.} adds a storage simulation extension to SimGrid 
+			to simulate file system,  but the page cache is not taken into account 
+			\cite{lebre2015}. 
+			SIMCAN is the simulation framework that models I/O with page cache 
+			with data on disk can be kept in page cache \cite{nunez2012simcan}.
+			Memory is also modeled in iCanCloud simulator with a component in charge 
+			of managing memory accesses and cached data \cite{nunez2012icancloud}, 
+			but one of the issues is scalability due to its sophisticated model.
+			Ultimately, non of these simulators provide any cache write strategy, 
+			cache eviction mechanism with dirty/clean data and the page cache LRU lists. 
+			Xu \textit{et al.} applies four cache replacement policies to simulate in-memory 
+			caching, but for the study in energy consumption of multi-tier heterogenous 
+			network \cite{xu2018saving}.
+
+			In this study, we chose SimGrid to implement our I/O with page cache model
+			due to the balancing between accuracy and scalibility, the versatility allowing it 
+			to be adopted in different domains, and the design enabling this 
+			framework to be extended with new models. 
+			The main issue users have to face when using SimGrid is everything has to be 
+			implemented from scratch due to its low-level API. 
+			To ease this challenge, WRENCH, an open-source library, is built on top 
+			of SimGrid to abstract the low-level logic and expose a higher abstraction 
+			level API \cite{casanova2018wrench}. 
+			We integrate our model in WRENCH and use the SimGrid API to interact with 
+			simulated storage resource.
+			This approach allows us to leverage existing features in WRENCH and SimGrid.
+			The first is the process control, synchronization of resource consumption by 
+			concurrent processes handled by the simulation core in SimGrid. 
+			The second is the existing SimGrid storage model with capacity and 
+			bandwidth properties, which are similar to the characteristics of memory.
+			The third is dropping the burden of implementation with the high level 
+			of abstraction in WRENCH.
 
 	\section{Method}
+	\label{method}
 
@@
 			to take advantage of bandwidth sharing between concurrent disk I/O to 
-			simulate parallel memory accesses.\tristan{we should refer to related work, 
-            see comment there} \valerie{Agreed. It's unclear how existing storage models enable this at the moment}.
+			simulate parallel memory accesses.
 
@@
 			the IO Controller instead of the storage simulation models.
-			\tristan{this should move, not sure where for now: 
-			Because every computer has its own memory and system configurations, 
-			one IO Controller object and one Memory Manager object are created 
-			for each simulated host.}
 
-			As mentioned, in our model, a file read/write can be complete by read/write 
+			As mentioned, in our model, a file read/write can be completed by read/write 
 			chunk-by-chunk until entire file is read/written.
@@
 			the experiments. The cluster consists of one login node, 8 compute nodes 
-			and 4 storage nodes connected with two network switches. The login node 
-			has one Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 137~GB (128~GiB) of RAM, 
-			1.8TB of storage of XFS file system, 13~GB of tmpfs file system and 385 TB of 
-			Lustre shared file system. Each compute node has two 16 cores Intel(R) 
-			Xeon(R) Gold 6130 CPU @ 2.10GHz, 275~GB (256~GiB) of RAM and 6 SSDs, 
-			450~GB each with XFS file system, 378~GB of tmpfs and 126~GB of devtmpfs file system.
+			and 4 storage nodes connected with a network bandwidth of 3GBps. 
+			The login node has one Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 
+			137~GB (128~GiB) of RAM, 1.8TB of storage of XFS file system, 
+			13~GB of tmpfs file system and 385 TB of Lustre shared file system. 
+			Each compute node has two 16 cores Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 
+			275~GB (256~GiB) of RAM and 6 SSDs, 450~GB each with XFS file system, 
+			378~GB of tmpfs and 126~GB of devtmpfs file system.
 			Lustre file system is configured with one metadata target with 854~GB 
 			of storage, 44 object storage targets with 8.7~TB of storage each. 
-			The cluster is run on CentOS 8.1 with the Slurm Workload Manager installed. 
+			The cluster is run on CentOS 8.1 with NFS version 4. 
 			We use \textbf{\textit{atop}} and \textbf{\textit{collectl}} as tools to 
@@
 			on the cluster.
-			\textcolor{red}{[Add network description here, include number of disks 
-			on all nodes ,maybe add a diagram]}
 
@@
 
-			\subsubsection{Evaluation on remote storage}
+			\subsubsection{Evaluation on NFS}
 
-		    Since remote storage is common in HPC, we conduct another experiment with 
-		    the same scenario as in the multi-threaded experiment, except that files 
-		    are stored on a node and read by another node.
+		    Since network file systems are very common in HPC, we conduct 
+		    the next experiment with the same scenario as in the multi-threaded experiment, 
+		    except that files are stored on a node, read and written by another node. 
+		    The compute node is the node used in the multi-threaded experiment, 
+		    and the capacity of the disk used on the storage node is 400 GB.
+
+		    In NFS, client cache and server cache are enabled in reading, but only 
+		    server write-through cache is involved in writing. 
+		    Although write-through does not have any impacts writing performance, 
+		    caching data on server side allows server to avoid disk reads, which are 
+		    significantly longer than cache reads.
+		    Thus, we simulate page cache for reading on both client and server node.
+		    In writing, write-though is simulated with data being written directly to 
+		    server storage and then added to server cache.
 
@@
 	\section{Results}
-
-		\begin{itemize}
-
-			\item Quantitative results: 
-				\begin{itemize}
-					\item Relative errors of simulation time and memory used compared to real results.
-					\item Simulation time compared to SimGrid and original WRENCH.
-				\end{itemize}
-
-			\item Ability of the model to generalize memory trends (dirty data, cache used) and disk throughput.
-
-			\end{itemize}
-
-		The benchmark results of bandwidths and the values used in the simulators are shown in \ref{table:benchmark}. These configurations are used in all experiments.
+	\label{results}			
+		The benchmark results of bandwidths and the values used in the simulators 
+		in our all experiments are shown in \ref{table:benchmark}.
 
@@
 			Memory write & 2764	& 4812 & 4812\\
-			Disk read & ~510 & 465 & 465\\
-			Disk write & ~420 & 465	 & 465\\
+			Local disk read & 510 & 465 & 465\\
+			Local disk write & 420 & 465	 & 465\\
+			Remote disk read & 515 & - & 445\\
+			Remote disk write & 375 & - & 445\\
+			Network bandwidth & 3000 & - & 3000\\
 		\hline
@@
 			performs some computation, and writes output file. 
-			Thus, there are 6 sequential I/O activities, \textit{read1} and \textit{write1} 
-			from the first task, \textit{read2} and \textit{write2} the second task, 
-			and \textit{read3}, \textit{write3} from the last task.
+			Thus, there are 6 sequential I/O activities, \textit{Read 1} and \textit{Write 1} 
+			from the first task, \textit{Read 2} and \textit{Write 2} the second task, 
+			and \textit{Read 3}, \textit{Write3} from the last task.
 
@@
 			than memory bandwidths, leading to simulated I/O time longer than in reality.
-			Especially, the simulation time errors of \textit{read3} in the simulators 
+			Especially, the simulation time errors of \textit{Read 3} in the simulators 
 			with page cache are just under the error in original WRENCH.
@@
 			\end{subfigure}		
-			\caption{I/O time with different numbers of concurrent pipeline}
+			\caption{I/O time of concurrent pipelines with local storage}
 			\label{fig:multi_local}
@@
 
+			\begin{figure*}		
+			\begin{subfigure}{\linewidth}
+				\centering
+   				\includegraphics[width=\linewidth]{result/multi/figures/multi_nfs.pdf}
+			\end{subfigure}		
+			\caption{I/O time of concurrent pipelines with NFS}
+			\label{fig:multi_nfs}
+			\end{figure*}
+
+			Similarly, the results with from the experiment with NFS are described 
+			by the average I/O time shown in Figure~\ref{fig:multi_nfs}.
+			In general, the average I/O time reflect a very similar level of accuracy 
+			to the previous experiment results when the I/O time of pipelines 
+			is very close to the real execution. 
+			The simulated read time if a bit off when the number of pipelines 
+			surpasses 22, with 264 GB of data in total. 
+			This is due to the impact of simulated cache eviction when 
+			the page cache is saturated, which has been seen in 
+			Figure~\ref{fig:single_error} with 100 GB of input. 
+			When it comes to write time, original WRENCH and WRENCH with page cache 
+			results are identical when both are writing directly to disk.
+
 		\subsection{Real application}
@@
 	\section{Discussion and Future Work}
-		\begin{itemize}
-			\item Read order: disk, inactive list, then active list. 
-			\item Cache eviction mechanism not evicting active list, leads to evicting file being written.
-			\item Sensitivity of the simulator on the variation of memory and disk bandwidth. 
-			\item Our model does not support random I/O and readahead.
-			\item Disk state save/restore when hosts are turned on/off.
-		\end{itemize}
-	\section{CFI for cluster}
+	\label{discussion}			
+			In computing infrastructures, especially those use for data-intensive applications, 
+			page cache undoubtedly has certain impacts on performance. 
+			Conducting simulation experiments is an effective approach to predict, 
+			evaluate the performance of not only application, scheduler but also 
+			infrastructure settings. 
+			Unfortunately, page cache is missing in many simulation tools for HPC systems. 
+			In this study, we proposed a simulation model of I/O with page cache, 
+			implement it with SimGrid using WRENCH, and evaluate the results 
+			with different experiments. 
+			The results shows our achievements in terms of simulation accuracy 
+			and the application potential in practice. 
+			On the other hands, the model exposed some limitations that can be 
+			further improved.
+
+			Our short term future work will be re-evaluate the model with asymmetrical 
+			disk bandwidths that will be released in the next SimGrid version.
+			The next step is to study the detail in cache eviction mechanism to achieve 
+			better simulation results. 
+			We also plan to support random I/O and file readahead, as well as 
+			saving and restoring disk state when hosts are turned on/off.
+
 \bibliographystyle{plain}

### 2020-11-16 gen_fig.py

@@ -13,4 +13,5 @@ plot_cache.plot_cache("fincore/real.csv", "fincore/sim.csv")
 
 os.chdir("../multi")
 import result.multi.process_result as multi_result
-multi_result.aggrerate_and_plot()
+multi_result.result_local()
+multi_result.result_nfs()

### 2020-11-16 result/multi/process_result.py

@@ -53,7 +53,7 @@ def export_simgrid_result(folder, filename):
         writer = csv.writer(csvfile)
         writer.writerow(["no_pipeline", "makespan", "readtime", "writetime"])
         for i in range(32):
-            dump_file = "%s/dump_%d.json" % (folder, i + 1)
+            dump_file = "%sdump_%d.json" % (folder, i + 1)
             writer.writerow(list(parse_simgrid_result(dump_file, i + 1)))
 
 
@@ -86,7 +86,7 @@ def suplot_prop(ax, exp_folder, propname, title):
     ax.set_ylabel("time (s)")
 
 
-def aggrerate_and_plot():
+def result_local():
     export_real_results("local/real/", "aggregated.csv")
     export_simgrid_result("local/wrench/original/", "aggregated.csv")
     export_simgrid_result("local/wrench/pagecache/", "aggregated.csv")
@@ -96,10 +96,35 @@ def aggrerate_and_plot():
     suplot_prop(ax1, "local", "readtime", "average read time")
     suplot_prop(ax2, "local", "writetime", "average write time")
 
+    ax1.set_ylim(bottom=0, top=1500)
+    ax2.set_ylim(bottom=0, top=1500)
+
     plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.3), ncol=3)
     plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.3)
-
     plt.savefig("figures/multi_local.pdf", format="pdf")
     plt.savefig("figures/multi_local.svg", format="svg")
 
     plt.show()
+
+
+def result_nfs():
+    export_real_results("remote/real/", "aggregated.csv")
+    export_simgrid_result("remote/wrench/original/", "aggregated.csv")
+    export_simgrid_result("remote/wrench/pagecache/", "aggregated.csv")
+    plt.rcParams.update({'font.size': 8})
+    fig, (ax1, ax2) = plt.subplots(figsize=(10, 5), ncols=2, nrows=1)
+
+    suplot_prop(ax1, "remote", "readtime", "average read time")
+    suplot_prop(ax2, "remote", "writetime", "average write time")
+
+    ax1.set_ylim(bottom=0, top=1500)
+    ax2.set_ylim(bottom=0, top=1500)
+
+    plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.3), ncol=3)
[…]

### 2020-11-16 result/single/plot_memprof.py

@@ -139,3 +139,6 @@ def compare_plot(input_sizes=[20, 100], makespan=[200, 1300]):
     plt.savefig("figures/single_memprof.svg", format="svg")
     plt.savefig("figures/single_memprof.pdf", format="pdf")
     plt.show()
+
+
+compare_plot()

### 2020-11-16 result/single/plot_real.py

@@ -1,5 +1,6 @@
 import matplotlib.pyplot as plt
 import numpy as np
+import pandas as pd
 import csv
 import log_parse
 
@@ -36,7 +37,7 @@ def mem_plot(fig, atoplog, time_stamp, input_size, readonly=False):
     time = np.arange(0, intervals)
 
     fig.minorticks_on()
-    fig.set_title("memory profiling (input size = %s MB)" % input_size)
+    fig.set_title("memory profiling (input size = %s GB)" % input_size)
 
     if time_stamp is not None:
         if readonly:
@@ -48,11 +49,11 @@ def mem_plot(fig, atoplog, time_stamp, input_size, readonly=False):
 
     fig.plot(time, atoplog["total"], color='k', linewidth=1, linestyle=":", label="total mem")
     # ax1.plot(time, free_mem, color='g', linewidth=1.5, linestyle="-.", label="free memory")
-    fig.plot(time, atoplog["used_mem"], color='g', linewidth=1.5, label="used mem")
+    fig.plot(time, atoplog["used_mem"], color='g', linewidth=1, label="used mem")
     # ax1.plot(time, app_mem, color='c', linewidth=1.5, label="app memory")
-    fig.plot(time, atoplog["cache"], color='m', linewidth=1.5, label="cache used")
+    fig.plot(time, atoplog["cache"], color='m', linewidth=1, label="cache used")
     # ax1.plot(time, app_cache, color='c', linewidth=1.5, label="cache + app")
-    fig.plot(time, atoplog["dirty_data"], color='r', linewidth=1.5, label="dirty data")
+    fig.plot(time, atoplog["dirty_data"], color='r', linewidth=1, label="dirty data")
     fig.plot(time, atoplog["avai_mem"], color='b', linewidth=1, linestyle="-.", label="available mem")
     fig.plot(time, atoplog["dirty_ratio"], color='k', linewidth=1, linestyle="-.", label="dirty_ratio")
     fig.plot(time, atoplog["dirty_bg_ratio"], color='r', linewidth=1, linestyle="-.", label="dirty_bg_ratio")
@@ -67,32 +68,36 @@ def collectl_plot(fig, collectl_log_file, time_stamp, readonly=False):
     read = []
     write = []
 
-    with open(collectl_log_file) as csv_file:
+    #         # Read Disk Stats
+    # with open(collectl_log_file) as csv_file:
+    #
[…]

### 2020-11-16 result/multi/process_result.py
Refine scripts

@@ -126,5 +126,3 @@ def result_nfs():
     plt.savefig("figures/multi_nfs.svg", format="svg")
 
     plt.show()
-
-result_nfs()

### 2020-11-16 result/single/plot_memprof.py

@@ -139,6 +139,3 @@ def compare_plot(input_sizes=[20, 100], makespan=[200, 1300]):
     plt.savefig("figures/single_memprof.svg", format="svg")
     plt.savefig("figures/single_memprof.pdf", format="pdf")
     plt.show()
-
-
-compare_plot()

### 2020-11-18 paper_io_sim.tex
some comments"

@@
 \usepackage{subcaption}
+\usepackage{ulem}
 
@@
 \newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:}#1\color{black}}
+\newcommand{\english}[1]{\uwave{#1}}
+
 \newcommand{\valerie}[1]{\color{blue}\textbf{From Valerie:}#1\color{black}}
@@
   \IEEEauthorblockA{
-    $^1$Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada \\
-     $^2$Department of Information and Computer Sciences, University of Hawai‘i at Manoa, U.S.A \\
-     $^3$Department of Computer Science, University of Southern California, Los Angeles, CA, USA
+    $^1$Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada\\
+     $^2$Department of Information and Computer Sciences, University of Hawai‘i at Manoa, USA\\
+     $^3$Department of Computer Science, University of Southern California, Los Angeles, USA
   }
@@
 		resource, the majority does not take page cache into consideration, 
-		which can lead to low simulation accuracy in many cases.
+		which can lead to low simulation accuracy in many cases. \tristan{The
+		 ideas are fine, but I don't find them well ordered. I think your intro 
+		could adopt this logic (1) Computing systems are key to the processing of Big Data, (2) page 
+		cache is critical to I/O performance, (3) page cache is not simulated so how 
+		can we expect Big Data applications on computing systems to be well simulated? 
+		This problem definition should take 2-3 sentences, the bulk of the abstract should be on the paper, not the context.
+		}
 
 		In this paper, we propose a simulation model page cache to simulate I/O 
-		with key features of page cache. 
+		with key features of page cache. \tristan{The model consists of ...}
 		The model is build with the consideration to allow it to be integrated to 
 		existing simulation tools. 
-		We also implement the model with SimGrid using WRENCH. 
-		Then we create some simulators for different use cases. 
+		We also implement the model with SimGrid using WRENCH \tristan{people
+		 may not know what these things are, characterize them. For instance:
+		  "using WRENCH, a SimGrid-based workflow simulator."}. 
+		Then we create some simulators for different use cases \tristan{be more specific: characterize the use cases.}. 
 		Finally, we analyze the results, compare them to results of the existing 
-		tool as well as real execution.
+		tool as well as real execution \tristan{This is too general, 
+		you should explain specifically what you found: how is your simulator better than plain simgrid?}.
+		\tristan{A conclusion sentence is missing on the impact of your work: how 
+		could it be used? what doors does it open? why is it important work?}
 
@@
 		Data processing is a complex process whose the performance is affected 
-		by many factors and hindered by bottlenecks in the systems.
+		by many factors and hindered by bottlenecks in the systems. \tristan{The previous
+		sentences are true but a 
+		bit general, everybody knows that, it's a boring way to open a paper. 
+		Your first sentences could be more specific to the simulation of big data applications.}
 		One of the bottlenecks that should be taken into consideration is disk access 
@@
 
-		Studying in optimizing not only the utilization of page cache but also other 
-		computing resources is a need that has never been overlooked.
+		\english{Studying in optimizing not only the utilization of page cache but also other 
+		computing resources is a need that has never been overlooked.}
 		The research on computing infrastructures to maximize performance, 
@@
 		The obvious advantages of using simulation are the freedom of choosing 
-		simulated infrastructures, reproducible results and low costs.
+		simulated infrastructures, reproducible results and low costs \tristan{The message of the previous 5 sentences could be summarized in 1 or 2.}.
 		These simulators are developed and integrated with simulation models 
 		of computing resources including CPU, network and storage.
-		However, some reviews over many simulators report that there has been 
+		However \tristan{why "however" and why is this trade-off important?}, some reviews over many simulators report that there has been 
 		a trade-off between accuracy and scalability due to the level of details in 
 		simulation models \cite{casanova2014simgrid} \cite{byrne2017review}.
-		Moreover, different simulators are developed by and for a specific community,
+		Moreover, different simulators are developed by and for a specific community \tristan{such as? not sure it's important for our purpose},
 		which hinders them from being used in cross-domain.
@@
 		but their models lack of some key features \cite{nunez2012simcan} 
-		\cite{nunez2012icancloud}. 
+		\cite{nunez2012icancloud} \tristan{which ones?}. 
 		Moreover, these simulators have a problem in scalability with 
@@
 		Some other works are able to capture some important features of page cache, 
-		but they focus on very specific domains \cite{xu2018saving}.
+		but they focus on very specific domains \cite{xu2018saving} \tristan{such as?}.
 
-		In this paper, we propose a simulation model of I/O with page cache.
+		In this paper, we propose a simulation model of I/O with page cache \tristan{you should mention what the paper is about earlier in the intro}.
 		The model is integrated in WRENCH, an open-source library built 
@@
 
-			In Linux, page cache is a part of memory, which includes pages referring to 
-			physical pages on disk \cite{linuxdev3rd2010}. 
+			In Linux, page cache is a part of memory that includes pages referring to 
+			physical pages on disk \cite{linuxdev3rd2010} \tristan{This sounds like a definition but is too vague.}. 
 			The goal of the page cache is to avoid disk I/O by keeping data in memory 
@@
 			The second is \textit{write-through}, in which data goes \textit{through} 
-			page cache to disk is. In this case, data is written to both page cache and disk. 
-			In the third strategy, which is \textit{write-back}, data are written directly 
+			page cache to disk is \tristan{is what?}. In this case, data is written to both page cache and disk. 
+			In the third strategy, called \textit{write-back}, data are written directly 
 			to disk as \textit{dirty} pages and these pages will be periodically flushed 
-			to disk later by a \textit{writeback} process and then marked as not dirty. 
+			to disk later by a \textit{writeback} \tristan{spell writeback consistently (dash or no dash?)} process and then marked as not dirty. 
 			The write-back strategy is considered to outperform write-through as well as 
-			direct I/O since it delays disk writes and perform in batch later 
+			direct I/O since it delays disk writes \english{ and perform in batch later }
 			\cite{linuxdev3rd2010}.
@@
 			The kernel chooses clean pages from page cache to evict. 
-			If the clean pages are inadequate, dirty pages are flushed to disk, 
+			If the clean pages are inadequate \tristan{why would they be so?}, dirty pages are flushed to disk, 
 			marked as clean and available for eviction. 
@@
 			pages accesses more than once, and an \textit{inactive list} containing 
-			pages available for eviction. 
+			pages available for eviction \tristan{and accessed how many times?}. 
 			Pages are moved from inactive list to active list if they are 
 			accessed when residing in the inactive list. 
-			The lists work on LRU manner, pages are added to the tail and removed 
+			The lists work in LRU manner, pages are added to the tail and removed 
 			from the top. 
@@
 
-			There has been the research and development of many simulation tools 
+			\english{There has been the research and development of many simulation tools 
 			and frameworks for cloud computing with the goal to simulate cloud, grid, 
-			HPC platforms, etc. 
-			A number of the most famous existing simulators have been analyzed 
+			HPC platforms, etc.} 
+			\english{A number of the most famous} existing simulators have been analyzed 
 			and compared in some reviews (e.g., \cite{byrne2017review}, \cite{casanova2014simgrid}). 
 			These simulators provide simulation models of CPU, network, storage 
-			with different approaches and levels of details.
-			Some simulators are at packet-level, which allows to achieve 
+			with different approaches and levels of details \tristan{what do the reviews conclude?}.
+			Some simulators are at packet-level \tristan{for the network, not a general statement}, which allows to achieve 
 			high simulation accuracy with fine-grain details such as storage simulation 
@@
 
-			However, I/O with page cache model has not been provided in SimGrid. 
+			However, I/O with page cache model has not been provided in SimGrid \tristan{has it been in other simulators?}. 
 			It only simulates storage with capacity and bandwidths.
-			Lebre \textit{et al.} adds a storage simulation extension to SimGrid 
+			Lebre \textit{et al.} \tristan{it looks like you're targeting this author, just use the reference as for the other works} adds a storage simulation extension to SimGrid 
 			to simulate file system,  but the page cache is not taken into account 
 			\cite{lebre2015}. 
-			SIMCAN is the simulation framework that models I/O with page cache 
-			with data on disk can be kept in page cache \cite{nunez2012simcan}.
+			SIMCAN is the \tristan{the or a?} simulation framework that models I/O with page cache 
+			\english{with data on disk can be kept in page cache} \cite{nunez2012simcan} \tristan{This paragraph started with Simgrid, so it gives the impression that 
+			all these simulators are in simgrid but I don't think they are. Revise.}.
 			Memory is also modeled in iCanCloud simulator with a component in charge 
@@
 			to be adopted in different domains, and the design enabling this 
-			framework to be extended with new models. 
+			framework to be extended with new models. \tristan{you could also mention that it's open source and has been maintained over the past years}
 			The main issue users have to face when using SimGrid is everything has to be 
@@
 			This approach allows us to leverage existing features in WRENCH and SimGrid.
-			The first is the process control, synchronization of resource consumption by 
+			\english{The first} is the process control, synchronization of resource consumption by 
 			concurrent processes handled by the simulation core in SimGrid. 
-			The second is the existing SimGrid storage model with capacity and 
+			\english{The second} is the existing SimGrid storage model with capacity and 
 			bandwidth properties, which are similar to the characteristics of memory.
-			The third is dropping the burden of implementation with the high level 
+			\english{The third} is dropping the burden of implementation with the high level 
 			of abstraction in WRENCH.
@@
 		In this section, we present our approach to model file read and write with
-		I/O to page cache, cache eviction and data flushing mechanisms implemented 
-		in the Linux kernel. We also detail the design of our simulators and their
+		I/O \tristan{"read and write with I/O" seems redundant} to page cache, cache eviction and data flushing mechanisms implemented 
+		in the Linux kernel. We also detail the design of our simulators \tristan{are there multiple simulators?} and their
 		implementation. Finally, we describe experimental scenarios to evaluate
@@
 		prototype, SimGrid framework using original WRENCH and the WRENCH version 
-		that is integrated with our simulation model (WRENCH-Ext).
+		that is integrated with our simulation model (WRENCH-Ext) \tristan{Introduce the simulator names first, then talk about experiments.}.
 
@@
 			file read/write mechanisms in Linux (Figure~\ref{fig:interaction}). 
-			The Memory Manager simulates I/Os to cache, flushing, periodical flushing, 
+			The Memory Manager simulates I/Os to cache, flushing, periodical flushing \tristan{why is periodical flushing different from flushing?}, 
 			and cache eviction, whereas the IO Controller simulates file reads and writes. 
@@
 			We leverage existing storage simulation models to simulate the memory,
-			which is characterized with storage capacity, distinguished read/write 
+			which is characterized with storage capacity, \english{distinguished} read/write 
 			bandwidths and device latency. 
-			All of these features are dealt within models in simulation 
-			frameworks, which makes storage models applicable to memory \cite{lebre2015}. 
+			\english{All of these features are dealt within models in simulation 
+			frameworks, which makes storage models applicable to memory} \cite{lebre2015}. 
 			In addition, leveraging existing disk models allows us 
@@
 			Memory Manager simulates two parallel threads. The first thread is the 
-			main thread where activities of applications are simulated, and responsible 
+			main thread where activities of applications are simulated \tristan{unclear, what are "activities of applications"?}, and responsible 
 			for the simulation of data flushing, cache eviction and I/Os to cache. 
@@
 			periodically searches and flushes expired dirty data in page cache LRU lists 
-			to disks.
+			to disks \tristan{how about periodical eviction?}.
 
@@
 
-			Now, page cache LRU lists can be modeled with two lists of data blocks: 
+			\english{Now,} page cache LRU lists can be modeled with two lists of data blocks: 
 			an active list and an inactive list, ordered by last access time 
@@
 			overhead of the simulator, as data blocks are obviously less
-			numerous than file pages, while achieving the accuracy in I/O time
-			simulation that can be seen in the later section.
+			numerous than file pages, \english{while achieving the accuracy in I/O time
+			simulation that can be seen in the later section}.
 

### 2020-11-18 paper_io_sim.tex
more comments

@@
    				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
-   				\caption{Overview of the simulated write-back page cache}\label{fig:interaction}
+				   \caption{Overview of the simulated write-back page cache \tristan{as discussed before, it would be useful to represent 
+				   which functions are exposed by each object. To understand the interface of each object. Also, add a line of explanation in the caption so 
+				   that we understand the main concept better by just looking at the figure.}}\label{fig:interaction}
 			\end{figure}
@@
 			\subsubsection{Memory simulation model}
+			\tristan{in this section there needs to be a clearer mapping between functions shown in Fig 1 and the paragraphs describing them.
+			Maybe add paragraph headers that match function names in Fig 1.}
 
@@
 
-			Memory Manager simulates two parallel threads. The first thread is the 
+			Memory Manager simulates two parallel threads \tristan{refer to figure 1}. The first thread is the 
 			main thread where activities of applications are simulated \tristan{unclear, what are "activities of applications"?}, and responsible 
@@
 
-			Given an arbitrary file, at a specific time, it can be partially cached, 
-			completely cached, or uncached.
+			\english{Given an arbitrary file, at a specific time, it can be partially cached, 
+			completely cached, or uncached}.
 			Cached data is composed of data blocks existing in either or both
@@
             unless they have been recently reaccessed. Cached blocks 
-            located on the inactive list that are subsequently accessed, 
-            will be moved to the top of the active list. Cached blocks 
-            recently written to cache are dirty blocks until they are flushed to disk.
+            located in the inactive list that are subsequently accessed, 
+            will be moved to the top of the active list. \tristan{The previous 2 sentences could be made clearer} Cached blocks 
+            written to cache are marked as dirty blocks until they are flushed to disk.
 
             Our model supports chunk-by-chunk file read/write simulation
-            with an user-defined chunk size, in which chunks are read/written 
+            with a user-defined chunk size, in which chunks are read/written 
             until file is entirely read/written. 
@@
             Uncached data is read from disk first, and cached data is only read 
-            after all uncached data is read (left to right in Figure~\ref{fig:read_order}).
+            after all uncached data is read \tristan{why did you use this order?} (left to right in Figure~\ref{fig:read_order}).
             When a chunk of uncached data is read, a new clean block is created 
-            and put to the top of the inactive list. 
+            and put at the top of the inactive list. 
             For a chunk of cached data, because one or more existing data blocks 
@@
             moved to the active list, while clean data blocks are merged into 
-            a single clean block and then also put in the active list. 
+			a single clean block and then also put in the active list \tristan{The
+			construction of the sentence insists on the difference between dirty and clean blocks, it would be clearer to first explain what is common, then mention what is different.}. 
 			The reason we do not merge dirty blocks is that different blocks 
@@
 			periodical flushing process.
-			In case a block is not entirely read, it is split into two smaller blocks, 
+			In case a block is not entirely read \tristan{why would this happen?}, it is split into two smaller blocks, 
 			one is reaccessed and the other remains the same.
@@
    				\includegraphics[width=\columnwidth]{figures/read_order.pdf}
-   				\caption{File data read order}	\label{fig:read_order}
+   				\caption{File data read order \tristan{add a sentence in the caption to summarize what we should see in this Figure}}	\label{fig:read_order}
 			\end{figure}
 
-			In the case of writing data, we can assume that all data to be written is 
+			For writes, we assume that all data to be written is 
 			uncached and it is written to cache as dirty data. 
@@
 
-			Next, we simulate data flushing, which runs on the main simulated 
+			Next, we simulate data flushing, which runs in the main simulated 
 			thread and writes back dirty data to disk. 
@@
 			reaches the amount to be flushed or there is no dirty data left in cache. 
-			If the last flushed block is not entirely flushed, it is split into 
+			If the last flushed block is not entirely flushed \tristan{why would this be the case?}, it is split into 
 			two blocks, one that is flushed and one that remains dirty.
@@
 
-			\begin{algorithm}\caption{Periodical flushing simulation}\label{alg:pdflush}
+			\begin{algorithm}\caption{Periodical flushing simulation \tristan{expiration time could be aded as a parameter}}\label{alg:pdflush}
 				\small
@@
 						\State flushing\_time = sm.write(blocks)
-						\State Flag blocks as clean data
+						\State flag blocks as clean data
 						\If{flushing\_time $<$ t}
@@
 			Similar to expired dirty pages in real Linux systems, a dirty block in our model 
-			is considered expired if the time it remains in cache as dirty is longer than 
-			the predefined expire time. This time is counted from the time it is written 
+			is considered expired if \english{the time it remains in cache as dirty} is longer than 
+			the predefined expiration time. This time is counted from the time it is written 
 			to cache to current simulated time. 
@@
 			as described in Algorithm~\ref{alg:pdflush}. 
-			The loop ends when the host the host is off (line 6). 
+			The loop ends when the host is off (line 6). 
 			In each repetition, Memory Manager finds expired dirty blocks in two 
@@
 			Then based on the state of the host at current simulated time, 
-			the algorithm continues or finishes the loop .
+			the algorithm continues or finishes the loop.
 			Because periodical flushing is simulated as a background thread, it can be 
@@
 
-			When simulated applications need to read/write a file, they can simply call 
-			the IO Controller instead of the storage simulation models.
+			When simulated applications need to read/write \tristan{avoid the use of /, it's ok from time to time but it's a bit informal}
+			 a file, they can simply call 
+			the IO Controller instead of the storage simulation models \tristan{as is the case for what?}.
 
@@
 			orchestrates cache and disk reads/writes, flushing, cache eviction 
-			to simulate I/O time.
+			to simulate I/O time \tristan{some of this could be moved to the top of II.A. It would be important to understand early on 
+			what are the relations between the memory manager and the io controller.}.
 
@@
 			The algorithm to simulate a chunk read is described in Algorithm~\ref{alg:read}.   
-			When a simulated application read a chunk of a file, it sends a 
-			chunk read request to IO Controller. 
+			When a simulated application reads a file chunk, it sends a 
+			chunk read request to IO Controller \tristan{say that Alg 2 implements the IOcontroller read method, here and in the title and in Fig 1.}. 
 			First, we calculate the amount of uncached data that needs to be read 
@@
 			and the uncached part of that chunk is read into cache.	
-			The amount of required memory to read the chunk is calculate and 
+			The amount of required memory to read the chunk is calculated and 
 			if there is not enough memory available, Memory Manager is called to 
-			flush dirty data to disk and simulate flushing time with memory model (line 9).
+			flush dirty data to disk and simulate flushing time with memory model (line 9). \tristan{do you assume that nothing is stored in anonymous memory when file read is called? 
+			it's ok but you should explain that this is due to the lack of application memory management in simgrid.}
 			This flushing is complemented by eviction in the Memory Manager (line 10). 
 			Then, if there is uncached data, a disk read is simulated with disk model (line 12), 
-			and this data of the file is added to cache (line 13)
+			and this data of the file is added to cache (line 13).
 			Finally, if cached data is read, Memory Manager is called to simulate a cache read, 
-			update the corresponding data blocks in cache (line 16).
+			\english{update the corresponding data blocks in cache (line 16). }
 
@@
 			Data can be written to cache with memory bandwidth without waiting for 
-			flushing or being throttled before the amount of dirty data reaches dirty\_ratio. 
+			flushing or being throttled before the amount of dirty data reaches dirty\_ratio. \tristan{the previous 2 sentences can be merged.}
 			Thus, our algorithm initially checks the remaining amount of data that can be written 
@@
 			If this amount is greater than 0, MemoryManager is asked to evict 
-			data from cache to accommodate the written data if needed (line 8). 
+			data from cache to accommodate the written data if needed (line 8). \tristan{say that evict doesn't do anything if it's passed a negative amount}
 			After evicting cache, the actual amount of data that can be written with 
-			memory bandwidth is calculated (line 9), a cache write is then simulated 
+			memory bandwidth \tristan{does "with memory bandwidth" mean "to memory"? if so, I think "to memory" is clearer.} is calculated (line 9), a cache write is then simulated 
 			with this amount through memory model and data written is added to 
-			page cache LRU lists (10).
+			page cache LRU lists (10 \tristan{line 10?}).
 			After dirty\_ratio is reached, dirty data has to be flushed before new data 
@@
 
-			To validate our simulation model, we create a simple prototype
+			To validate our simulation model, we create \tristan{Using past tense in this section would read better} a simple prototype
 			simulator independent of existing simulation frameworks and libraries. 
@@
 			using WRENCH.
-			We integrate our I/O model to the original WRENCH by implementing 
+			We integrate our I/O model in the original WRENCH by implementing 
 			the core of the model separately from existing WRENCH components, 
@@
 			We also add an optional command line argument so that users will be able 
-			to choose use or not to use the model. Doing this, we can avoid 
+			to choose use or not to use the model \tristan{this sentence is too verbose, check verbosity overall, many 
+			sentences can be summarized to say the same thing with less words.}. Doing this, we can avoid 
 			unexpected impacts on original WRENCH and can independently update 
@@
 			SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid}, 
-			and WRENCH is at \url{https://github.com/wrench-project/wrench}.
+			and WRENCH is at \url{https://github.com/wrench-project/wrench} \tristan{mention the WRENCH commit used for your simulator}.
 

### 2020-11-23 paper_io_sim.tex
Paper revised

--- 2020-11-16 paper_io_sim.tex
@@
 
-		The data explosion era seen in recent years results in growing needs in 
-		data storing, data analysis. 
-		Large computing systems including grids, clusters, clouds, peer-to-peer, 
-		volunteer computers have been increasingly used to solve the problem. 
-		Despite of their powerful computing capability and technological improvements, 
-		there are still bottlenecks in these systems. 
-		When it comes to I/O, a bottleneck can be seen in the disk I/O due to hardware 
-		limitations. Data caching using page cache is one of the solutions applied 
-		to minimize disk accesses and improve the I/O performance. 
-		In computing system and application optimization studies, one of the 
-		effective approaches is using simulation tools to conduct experiments instead 
-		of running applications on real infrastructures. 
-		Although simulators are built with individual simulation model for each computing 
-		resource, the majority does not take page cache into consideration, 
-		which can lead to low simulation accuracy in many cases.
+		The explosive development of Big Data era seen in recent years results in 
+		growing needs in data storing, data analysis. 
+		Although large computing systems with powerful computing capability 
+		have been increasingly used to process big datasets, there is still a bottleneck 
+		in disk I/O. 
+		Memory page cache is a solution applied to minimize disk accesses and 
+		improve the I/O performance. 
+		Despite of the fact that simulation tools for HPC have been developed to 
+		support studies in computing systems optimization, page cache simulation 
+		has not been fully captured, or absent in these simulators. 
+		The lack of ability to simulate I/O with page cache leads to inaccurate 
+		simulation results in many situations, especially in data intensive applications.
 
-		In this paper, we propose a simulation model page cache to simulate I/O 
-		with key features of page cache. 
-		The model is build with the consideration to allow it to be integrated to 
-		existing simulation tools. 
-		We also implement the model with SimGrid using WRENCH. 
-		Then we create some simulators for different use cases. 
-		Finally, we analyze the results, compare them to results of the existing 
-		tool as well as real execution.
+		In this paper, we propose a I/O simulation model with key features of the page cache.
+		The model consists of simulation algorithms of I/O, main components 
+		and mechanisms of page cache.
+		We also implement the model with SimGrid, a simulation framework for 
+		distributed systems, using WRENCH, a SimGrid-based workflow simulator. 
+		Then we create simulators for applications in single-threaded, multi-threaded,
+		network file system and neuroimaging usecases.
+		Finally, we analyze the results, compare to the results of base-line WRENCH 
+		simulator as well as real execution.
+		The results show that our simulator not only achieves higher accuracy than 
+		original WRENCH, but also simulates well the status of memory and 
+		page cache during I/O operations. 
+		Through this study, our effort is to contribute an analytical model 
+		of the page cache, which can also be integrated to other simulators 
+		to improve I/O simulation accuracy, with its practical results.
 
@@
 
-		In recent years, as a result of the growth of data generated from a variety 
-		of data sources, we have seen increasingly growing demands in data storage, 
-		processing and analysis. 
-		Storing and processing such a massive amount of data require large 
-		computing infrastructures. 
-		Data processing is a complex process whose the performance is affected 
-		by many factors and hindered by bottlenecks in the systems.
-		One of the bottlenecks that should be taken into consideration is disk access 
-		time due to hardware limitations. 
-		Beside hardware technical improvements, architectural solutions have been 
-		proposed and implemented in computing systems in order to tackle the 
-		disk I/O problem. 
-		Caching, a successful technique applied into memory to maximize the 
-		I/O performance by keeping data in memory for future accesses,
-		has been proving the effectiveness when employed in computer systems.
-
-		Studying in optimizing not only the utilization of page cache but also other 
-		computing resources is a need that has never been overlooked.
-		The research on computing infrastructures to maximize performance, 
-		shorten processing time, and minimize costs gets involved with experiments. 
-		However, real platforms are complex and not built for the experimental 
-		purposes due to limited and shared resources, operational costs, unstable 
-		environments and the result reproducibility difficulties. 
-		An alternative approach is to used simulation tools developed for research 
-		and experimental purposes.
-		The obvious advantages of using simulation are the freedom of choosing 
+		In recent years, as a result of the growing demand for Big Data processing 
+		on large computing infrastructures, there have been the research and development 
+		of simulation applications and frameworks for distributed systems. 
+		These simulation tools have shown their advantages in evaluation 
+		of algorithms, applications and solutions. 
+		Experimenting on real platforms exposes shortcomings due to limited and 
+		shared resources, operational costs, unstable environments and the result 
+		reproducibility difficulties.
+		On the contrary, simulators provide to researchers the freedom of choosing 
 		simulated infrastructures, reproducible results and low costs.
@@
 		of computing resources including CPU, network and storage.
-		However, some reviews over many simulators report that there has been 
-		a trade-off between accuracy and scalability due to the level of details in 
-		simulation models \cite{casanova2014simgrid} \cite{byrne2017review}.
-		Moreover, different simulators are developed by and for a specific community,
-		which hinders them from being used in cross-domain.
-		SimGrid is a simulation framework developed with the objectives of  
-		balancing between accuracy and scalability, as well as versatility.
+		The ultimate goal that all simulators try to achieve is to accurately simulate 
+		applications. 
+		However, there is a trade-off between accuracy and scalability, which has been 
+		indicated in some reviews \cite{casanova2014simgrid} \cite{byrne2017review}. 
+		This means that in most ciscumstances, users have to, more or less, sacrifice 
+		accuracy for the scalability and performance when their platforms grow to 
+		hundreds or thousands nodes.
 
-		In spite of the importance of page cache and the impacts it has on I/O 
-		operations, its simulation model is absent in many simulators. 
-		Some simulation tools try to model the page cache with cached data, 
-		but their models lack of some key features \cite{nunez2012simcan} 
-		\cite{nunez2012icancloud}. 
-		Moreover, these simulators have a problem in scalability with 
+		Beside computing resources such as CPU, network, storage, another 
+		important factor that has significant impacts on I/O performance is 
+		page cache. 
+		Page cache is an architectural solution employed in systems to tackle 
+		the problem of disk I/O bottleneck in data processing. 
+		Although this is not a radical solution for disk performance, it contributes 
+		to minimizing disk accesses and maximizing I/O performance. 
+		Surprisingly, page cache has not been fully modeled, if not completely absent, 
+		in simulation tools. 
+		Some simulators try to simulate I/O with page cache, but fail to capture the 
+		key features such as dirty data and cache eviction mechanism 
+		\cite{nunez2012simcan} \cite{nunez2012icancloud}.
+		Again, these simulators also have a problem in scalability with 
 		long simulation time regarding to their very detailed level of simulation. 
-		Some other works are able to capture some important features of page cache, 
-		but they focus on very specific domains \cite{xu2018saving}.
+		Other works are able to capture important features of page cache, 
+		but they focus on very specific domains, such as energy consumption 
+		in \cite{xu2018saving}.
 
-		In this paper, we propose a simulation model of I/O with page cache.
-		The model is integrated in WRENCH, an open-source library built 
-		on top of SimGrid with more abstract API. 
-		Then we evaluate our model with four simulators in four 
-		experiment scenarios. 
+		In this work, we propose a simulation model of I/O with page cache, 
+		implement it with SimGrid, a simulation framework for distributed computing, 
+		and WRENCH, an open-source library for developing workflow simulator 
+		built on top of SimGrid. 
+		One of the reasons for us to choose SimGrid is that it has a balance 
+		between accuracy and scalability. Moreover, the versatility of SimGrid 
+		enables it to be used in cross-domain \cite{casanova2014simgrid}. 
+		In addition, SimGrid has been actively maintained with updated documents. 
+		Then we evaluate our model with four simulators in four experiment scenarios. 
 		This paper is organized as follows. 
@@
 
-			In Linux, page cache is a part of memory, which includes pages referring to 
-			physical pages on disk \cite{linuxdev3rd2010}. 
+			Linux implements a disk cache called \textit{page cache}. 
+			It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
 			The goal of the page cache is to avoid disk I/O by keeping data in memory 
@@
 			The second is \textit{write-through}, in which data goes \textit{through} 
-			page cache to disk is. In this case, data is written to both page cache and disk. 
-			In the third strategy, which is \textit{write-back}, data are written directly 
+			page cache to disk. In this case, data is written to both page cache and disk. 
+			In the third strategy, called \textit{writeback}, data are written directly 
 			to disk as \textit{dirty} pages and these pages will be periodically flushed 
 			to disk later by a \textit{writeback} process and then marked as not dirty. 
-			The write-back strategy is considered to outperform write-through as well as 
-			direct I/O since it delays disk writes and perform in batch later 
+			The writeback strategy is considered to outperform write-through as well as 
+			direct I/O as it delays disk writes and perform a bulk write later.
 			\cite{linuxdev3rd2010}.
 
-			In write-back, one of the keys is the cache eviction mechanism, which 
-			removes pages from the page cache when available memory is insufficient. 
-			The kernel chooses clean pages from page cache to evict. 
+			In writeback, one of the keys is the cache eviction mechanism, which is
+			triggered when available memory is insufficient to removes pages 
+			from page cache. 
+			Because dirty pages are data written to page cache but yet to be 
+			written to disk, dirty pages must be kept in page cache and only clean pages 
+			are available for eviction.
 			If the clean pages are inadequate, dirty pages are flushed to disk, 
 			marked as clean and available for eviction. 
-			Different cache eviction algorithms have been developed and proposed
+			There is another type of flushing mechanism, which is 
+			\textit{periodical flushing}, called periodically in predefined intervals. 
+			Periodical flushing only flushes expired dirty pages, which remain dirty in 
+			page cache longer than an expiration time configured in the kernel. 
+			Different cache eviction algorithms have also been developed and proposed
 			\cite{owda2014comparison}.
 			In Linux kernel, a \textit{two lists strategy (LRU/2)} is implemented based on 
-			the least-recently-used (LRU) scheme, with an \textit{active list} only containing 
-			pages accesses more than once, and an \textit{inactive list} containing 
-			pages available for eviction. 
-			Pages are moved from inactive list to active list if they are 
-			accessed when residing in the inactive list. 
-			The lists work on LRU manner, pages are added to the tail and removed 
-			from the top. 
+			the least-recently-used (LRU) scheme, with an \textit{active list} and 
+			an \textit{inactive list}. 
+			If accessed pages are not in the page cache, they are added to the inactive list. 
+			If pages are accessed when residing in the inactive list, they will be moved from 
+			inactive list to active list. 
 			The lists are also kept balanced by moving pages from the active list
 			to the inactive list when the active list grows too large.
+			Thus, the active list only contains pages which are accessed more than once 
+			and not evictable, while the inactive list includes pages accessed once only, 
+			or pages accessed more than once but moved from the active list. 
+			The lists work in LRU manner, pages are added to the tail and removed 
+			from the top.
 
@@
 
-			There has been the research and development of many simulation tools 
-			and frameworks for cloud computing with the goal to simulate cloud, grid, 
-			HPC platforms, etc. 
-			A number of the most famous existing simulators have been analyzed 
-			and compared in some reviews (e.g., \cite{byrne2017review}, \cite{casanova2014simgrid}). 
-			These simulators provide simulation models of CPU, network, storage 
-			with different approaches and levels of details.
-			Some simulators are at packet-level, which allows to achieve 
-			high simulation accuracy with fine-grain details such as storage simulation 
-			in \textit{DiskSim} or network simulation in \textit{CloudSim} 
-			and \textit{iCanCloud} \cite{casanova2014simgrid}. 
-			However, the issues with the packet-level simulation are scalability due to 
-			long simulation time when the simulated platforms and applications grow 
-			to hundreds and thousands nodes. 
+			Many simulation frameworks and applications have been developed 
+			in recent years to support studies in grids, cloud, HPC, peer-to-peer 
+			networks, volunteer computers. 					
+			These simulators are built with models of resources such as CPU, network, 
+			storage with different approaches and levels of details in different simulators. 
+			The detail level of the model causes a conflict between accuracy and scalability, 
+			and these simulators usually trade of one for the other 
+			\cite{casanova2014simgrid}. 
+			For example, some simulators are built with packet-level network models 
+			such as \textit{CloudSim} and \textit{iCanCloud}, or storage model that 
+			simulates the operation of storage hardware in \textit{DiskSim} 
+			\cite{casanova2014simgrid}. 
+			Fine-grained models allow simulators to achieve high accuracy, 
+			but at the same time lead to scalability issues due to long simulation time 
+			when simulated platforms and applications grow. 
 			SimGrid is an open source simulation platform that has been developed 
@@
 
-			However, I/O with page cache model has not been provided in SimGrid. 
-			It only simulates storage with capacity and bandwidths.
-			Lebre \textit{et al.} adds a storage simulation extension to SimGrid 
-			to simulate file system,  but the page cache is not taken into account 
-			\cite{lebre2015}. 
-			SIMCAN is the simulation framework that models I/O with page cache 
-			with data on disk can be kept in page cache \cite{nunez2012simcan}.
-			Memory is also modeled in iCanCloud simulator with a component in charge 
+			Despite the fact that many sophisticated models for resources have been 
+			integrated in simulators, and the page cache shows apparent impacts on I/O
+			performance, the simulation model of page cache is still missing in most 
+			of simulators.
+			For example, SimGrid only simulates I/O operations with storage bandwidths 
+			and capacity. 
+			SIMCAN is a simulation framework that models page cache by storing data 
+			accessed on disk in a block cache \cite{nunez2012simcan}. 
+			Page cache is also modeled in iCanCloud simulator with a component in charge 
 			of managing memory accesses and cached data \cite{nunez2012icancloud}, 
@@
 			cache eviction mechanism with dirty/clean data and the page cache LRU lists. 
-			Xu \textit{et al.} applies four cache replacement policies to simulate in-memory 
-			caching, but for the study in energy consumption of multi-tier heterogenous 
-			network \cite{xu2018saving}.
+			Cache replacement policies is applied in \cite{xu2018saving} to simulate 
+			in-memory caching, but for the study in energy consumption of multi-tier 
+			heterogenous network.
 
-			In this study, we chose SimGrid to implement our I/O with page cache model
+			In this study, we chose SimGrid to implement our model of I/O with page cache 
 			due to the balancing between accuracy and scalibility, the versatility allowing it 
-			to be adopted in different domains, and the design enabling this 
-			framework to be extended with new models. 
+			to be adopted in different domains. 
+			In addition, it is an actively maintained open-scource framework with 
+			modular design, which enables the framework to be extended with new models. 
 			The main issue users have to face when using SimGrid is everything has to be 
 			implemented from scratch due to its low-level API. 
-			To ease this challenge, WRENCH, an open-source library, is built on top 
+			To ease this challenge, WRENCH, another open-source library, is built on top 
 			of SimGrid to abstract the low-level logic and expose a higher abstraction 
@@
 			This approach allows us to leverage existing features in WRENCH and SimGrid.
-			The first is the process control, synchronization of resource consumption by 
-			concurrent processes handled by the simulation core in SimGrid. 
-			The second is the existing SimGrid storage model with capacity and 
-			bandwidth properties, which are similar to the characteristics of memory.
-			The third is dropping the burden of implementation with the high level 
-			of abstraction in WRENCH.
+			Firstly, we can take advantage of the process control, synchronization of 
+			resource consumption by concurrent processes provided by the simulation core 
+			in SimGrid. 
+			Secondly, we can reuse the existing SimGrid storage model with capacity 
+			and bandwidth properties, which are similar to the characteristics of memory.
+			And finally, WRENCH provides a high level abstraction, which help us drop 
+			the burden of implementation.
 
@@
 		In this section, we present our approach to model file read and write with 
-		I/O to page cache, cache eviction and data flushing mechanisms implemented 
-		in the Linux kernel. We also detail the design of our simulators and their
-		implementation. Finally, we describe experimental scenarios to evaluate
-		our simulation model on real applications using a standalone Python
-		prototype, SimGrid framework using original WRENCH and the WRENCH version 
-		that is integrated with our simulation model (WRENCH-Ext).
+		page cache, cache eviction and data flushing mechanisms implemented 
+		in the Linux kernel. 
+		We also detail the design of our simulator and its implementation. 
+		Because the simulator is implemented with SimGrid using WRENCH, we 
+		name the simulator as WRENCH-Ext, along with the WRENCH simulator,
+		which is created with original WRENCH.
+		Finally, we describe experimental scenarios to evaluate our simulation model 
+		using a standalone Python prototype, WRENCH and WRENCH-Ext simulator.
 
@@
 			and cache eviction, whereas the IO Controller simulates file reads and writes. 			
-			Simulated applications can simply interact with the IO Controller instead of 
-			calling simulated storage directly as is the case with current simulators.
+			To read or write a chunk of file, a simulated application sends a request 
+			to the IOController.
+			Based on the memory status in Memory Manager, the IOController 
+			orchestrates cache and disk reads/writes, flushing, cache eviction 
+			to simulate I/O time.
 
@@
    				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
-   				\caption{Overview of the simulated write-back page cache}\label{fig:interaction}
+				   \caption{Overview of the simulated writeback page cache.
+				   Application sends file read and file write requests to IOController, 
+				   then IOController manages orchestrate flushing, eviction, cache and 
+				   disk accesses to fulfill the requests with the MemoryManager object. 
+				   Memory and Disk are responsible for simulating I/O time.}
+				   \label{fig:interaction}
 			\end{figure}
@@
 			We leverage existing storage simulation models to simulate the memory,
-			which is characterized with storage capacity, distinguished read/write 
-			bandwidths and device latency. 
-			All of these features are dealt within models in simulation 
-			frameworks, which makes storage models applicable to memory \cite{lebre2015}. 
+			which is characterized with storage capacity, read and write bandwidths 
+			and device latency. 
+			These features are captured in simulation frameworks and applicable 
+			to memory simulation \cite{lebre2015}. 
 			In addition, leveraging existing disk models allows us 
@@
 
-			Memory Manager simulates two parallel threads. The first thread is the 
-			main thread where activities of applications are simulated, and responsible 
-			for the simulation of data flushing, cache eviction and I/Os to cache. 
+			\paragraph*{Memory Manager}
+			Memory Manager simulates two parallel threads presented in \ref{fig:interaction}.
+			The first thread is the main thread, which is responsible for the simulation 
+			of data flushing, cache eviction and I/Os to cache. 
 			The second thread is a simulated background thread, in which Memory Manager 
@@
 
+			\paragraph*{Page cache LRU lists}
 			In Linux kernel, the page cache LRU lists contain file pages. However, 
@@
 
-			Now, page cache LRU lists can be modeled with two lists of data blocks: 
-			an active list and an inactive list, ordered by last access time 
-			(earliest first, see Figure~\ref{fig:lrulist}).
+			With the data block definition, page cache LRU lists can be modeled with 
+			two lists of data blocks: an active list and an inactive list, ordered by 
+			last access time (earliest first, see Figure~\ref{fig:lrulist}).
 			As in the kernel, our simulator keeps the size of the active list under
@@
 			overhead of the simulator, as data blocks are obviously less
-			numerous than file pages, while achieving the accuracy in I/O time
-			simulation that can be seen in the later section.
+			numerous than file pages, and at the same time achieves the accuracy in 
+			I/O time simulation that can be seen in the later section.
 
-			Given an arbitrary file, at a specific time, it can be partially cached, 
-			completely cached, or uncached.
-			Cached data is composed of data blocks existing in either or both
-            of the page cache LRU lists. Cached blocks reside in the inactive list
-            unless they have been recently reaccessed. Cached blocks 
-            located on the inactive list that are subsequently accessed, 
-            will be moved to the top of the active list. Cached blocks 
-            recently written to cache are dirty blocks until they are flushed to disk.
+			At a specific time, an arbitrary file can be partially cached, completely cached, 
+			or uncached.
+			Cached data is composed of data blocks existing in one or both
+            of the page cache LRU lists. 
+            Initially, blocks accessed for the first time are added to the inactive list in 
+            page cache. 
+            If the blocks residing the inactive list are subsequently accessed, 
+            they will be moved to the top of the active list
+            Cached blocks written to cache are marked as dirty blocks until 
+            they are flushed to disk.
 
-            Our model supports chunk-by-chunk file read/write simulation
-            with an user-defined chunk size, in which chunks are read/written 
+			\paragraph*{File read and write}
+            Our model supports chunk-by-chunk file read and write simulation
+            with a user-defined chunk size, in which chunks are read/written
             until file is entirely read/written.
@@
             or uncached. 
-            Uncached data is read from disk first, and cached data is only read 
-            after all uncached data is read (left to right in Figure~\ref{fig:read_order}).
+            In the scope of this study, our model support file accesses in chunk 
+            but not file random access.
+            Thus, we simplify the data read order, in which uncached data is read 
+            from disk first, and cached data is only read after all uncached data is read 
+            (left to right in Figure~\ref{fig:read_order}).
             When a chunk of uncached data is read, a new clean block is created 
-            and put to the top of the inactive list. 
-            For a chunk of cached data, because one or more existing data blocks 
-            in two page cache LRU lists are accessed again, read dirty blocks are 
-            moved to the active list, while clean data blocks are merged into 
-            a single clean block and then also put in the active list. 
-			The reason we do not merge dirty blocks is that different blocks 
-			are written to cache at different time, and this time is used later for 
-			periodical flushing process.
-			In case a block is not entirely read, it is split into two smaller blocks, 
-			one is reaccessed and the other remains the same.
+            and put at the top of the inactive list. 
+            For a chunk of cached data, one or more existing data blocks in two 
+            page cache LRU lists are accessed again, and moved to the active list. 
+			If the blocks are clean, we only need to keep the information about the 
+			amount of data as they are available for eviction. 
+			Thus, these re-accessed clean blocks can be merge into a bigger block 
+			to reduce the number of block objects and put it in the active list. 
+            If the blocks are dirty, since the blocks need to keep the information 
+            of entry time for periodical flushing, they are only moved to the active list. 
+			Because the chunk size and block size can be different, there are circumstances 
+			in which a block is not entirely read. 
+			In this case, the block is split into two smaller blocks, one is re-accessed 
+			and the other remains the same.
 
@@
    				\includegraphics[width=\columnwidth]{figures/read_order.pdf}
-   				\caption{File data read order}	\label{fig:read_order}
+   				\caption{File data read order from left to right, in which uncached data 
+   				is read first, then cached data from inactive list and active list is read later.}	
+   				\label{fig:read_order}
 			\end{figure}
 
-			In the case of writing data, we can assume that all data to be written is 
+			For writes, we assume that all data to be written is 
 			uncached and it is written to cache as dirty data. 
-			Thus, each time a chunk is written, we create a block of dirty data and 
-			put it in the inactive list.
-
-			\textcolor{red}{[Is there any kind of accesses other than read and write?]}.
+			Thus, each time a chunk is written, we create a block of dirty data 
+			and put it in the inactive list.
 
@@
 
-			Next, we simulate data flushing, which runs on the main simulated 
+			\paragraph*{Flushing}
+			Next, we simulate data flushing, which runs in the main simulated 
 			thread and writes back dirty data to disk. 
@@
 			reaches the amount to be flushed or there is no dirty data left in cache. 
-			If the last flushed block is not entirely flushed, it is split into 
-			two blocks, one that is flushed and one that remains dirty.
+			Similar to file read and write, the size of flushed data and block size 
+			can be different, there are cases that the last block is not entirely flushed.  
+			In that case, the last block is split into two blocks, one that is flushed 
+			and one that remains dirty.
 			The flushing time is simulated with storage model.
 
+			\paragraph*{Cache eviction}	
 			Similarly to data flushing, our cache eviction simulation runs on 
@@
 						\Desc{t}{predefined flushing time interval}
+						\Desc{exp}{predefined expiration time}						
 						\Desc{sm}{storage simulation model}
@@
    					\While{host is on}
-						\State blocks = expired dirty blocks in active and inactive list
+						\State blocks = expired\_blocks(exp, in) + expired\_blocks(exp, ac) 
 						\State flushing\_time = sm.write(blocks)
-						\State Flag blocks as clean data
+						\State flag blocks as clean data
 						\If{flushing\_time $<$ t}
@@
 
+			\paragraph*{Periodical flushing}	
 			The periodical flushing simulation is run on a simulated background thread 
@@
 			Similar to expired dirty pages in real Linux systems, a dirty block in our model 
-			is considered expired if the time it remains in cache as dirty is longer than 
-			the predefined expire time. This time is counted from the time it is written 
-			to cache to current simulated time. 
+			is considered expired if: 
+			\begin{enumerate}
+				\item It is dirty and 
+				\item The time from its entry time to current time is longer than the 
+				predefined expiration time. 
+			\end{enumerate}
 			The simulation algorithm of periodical flushing is an infinite loop in which 
@@
 			as described in Algorithm~\ref{alg:pdflush}. 
-			The loop ends when the host the host is off (line 6). 
+			The loop ends when the host is off (line 7). 
 			In each repetition, Memory Manager finds expired dirty blocks in two 
-			page cache LRU lists (line 7), simulates writes of data of these blocks 
-			to disk (line 8), and mark them as clean (line 9).
+			page cache LRU lists (line 8), simulates writes of data of these blocks 
+			to disk (line 9), and mark them as clean (line 10).
 			If the flushing time does not exceed our time interval, the thread is put 
-			to sleep for the remaining time (lines 10-12). 
+			to sleep for the remaining time (lines 11-13). 
 			Then based on the state of the host at current simulated time, 
-			the algorithm continues or finishes the loop .
+			the algorithm continues or finishes the loop.
 			Because periodical flushing is simulated as a background thread, it can be 
@@
 
-			When simulated applications need to read/write a file, they can simply call 
-			the IO Controller instead of the storage simulation models.
+			As mentioned, in our model, file reads and writes can be completed by 
+			reading or writing chunk-by-chunk until entire file is read or written.
 
-			As mentioned, in our model, a file read/write can be completed by read/write 
-			chunk-by-chunk until entire file is read/written.
-			To read or write a chunk of file, a simulated application sends a request to the 
-			IOController instance on the node where the requested file is stored.
-			Based on the memory status in Memory Manager, the IOController 
-			orchestrates cache and disk reads/writes, flushing, cache eviction 
-			to simulate I/O time.
-
-			\begin{algorithm}\caption{File chunk read simulation}\label{alg:read}
+			\begin{algorithm}\caption{File chunk read simulation of IOController}
+			\label{alg:read}
 				\small
@@
 
-			The algorithm to simulate a chunk read is described in Algorithm~\ref{alg:read}.   
-			When a simulated application read a chunk of a file, it sends a 
-			chunk read request to IO Controller. 
+			When a simulated application reads a file chunk, it sends a chunk read request 
+			to IOController. 			
+			The algorithm to simulate a chunk read implemented in IOController 
+			is described in Algorithm~\ref{alg:read}.   
 			First, we calculate the amount of uncached data that needs to be read 
@@
 			(one in cache, one in anonymous memory). 
-			Thus, when a chunk is read, one copy of the chunk occupies anonymous memory, 
-			and the uncached part of that chunk is read into cache.	
-			The amount of required memory to read the chunk is calculate and 
+			Unfortunately, SimGrid lacks of application memory management and WRENCH 
+			only use the memory capacity to check if the applications can be run. 
+			Thus, in our model, when a chunk is read, one copy of the chunk 
+			occupies anonymous memory, and the uncached part of that chunk is 
+			read into cache.	
+			The amount of required memory to read the chunk is calculated and 
 			if there is not enough memory available, Memory Manager is called to 
@@
 			Then, if there is uncached data, a disk read is simulated with disk model (line 12), 
-			and this data of the file is added to cache (line 13)
-			Finally, if cached data is read, Memory Manager is called to simulate a cache read, 
-			update the corresponding data blocks in cache (line 16).
+			and this data of the file is added to cache (line 13).
+			Finally, if cached data is read, Memory Manager is called to simulate a cache read  
+			and update the corresponding data blocks in cache (line 16).
 
-			\begin{algorithm}\caption{File chunk write simulation}\label{alg:write}
+			\begin{algorithm}\caption{File chunk write simulation of IOController}
+			\label{alg:write}
 				\small
@@
 
-			Algorithm~\ref{alg:write} describes our simulation of chunk write. 
-			In our chunk write simulation, data can be written with either memory bandwidth 
-			or disk bandwidth. 
-			Data can be written to cache with memory bandwidth without waiting for 
-			flushing or being throttled before the amount of dirty data reaches dirty\_ratio. 
-			Thus, our algorithm initially checks the remaining amount of data that can be written 
-			as dirty data (line 6).
+			Algorithm~\ref{alg:write} describes our simulation of chunk write in 
+			the IOController. 
+			In the chunk write, data can be written to cache with memory bandwidth 
+			without waiting for flushing or being throttled before the amount of 
+			dirty data reaches dirty\_ratio, or otherwise written with disk bandwidth.
+			Thus, our algorithm initially checks the remaining amount of data that 
+			can be written as dirty data (line 6).
 			If this amount is greater than 0, MemoryManager is asked to evict 
-			data from cache to accommodate the written data if needed (line 8). 
-			After evicting cache, the actual amount of data that can be written with 
-			memory bandwidth is calculated (line 9), a cache write is then simulated 
+			data from cache to accommodate the written data (line 8). 
+			If the amount to evict calculated and passed in the evict function is less than 0, 
+			the eviction does nothing.
+			After evicting cache, the actual amount of data that can be written to 
+			page cache is calculated (line 9), a cache write is then simulated 
 			with this amount through memory model and data written is added to 
-			page cache LRU lists (10).
+			page cache LRU lists (line 10).
 			After dirty\_ratio is reached, dirty data has to be flushed before new data 
@@
 
-			To validate our simulation model, we create a simple prototype
+			To validate our simulation model, we created a simple prototype
 			simulator independent of existing simulation frameworks and libraries. 
-			This enables us to evaluate the accuracy and correctness of our 
-			model in a simple scenario before integrating it to the more complex WRENCH framework. 
-			In this prototype we use the following basic storage model, used for both memory and disk: 
+			This allowed us to evaluate the accuracy and correctness of our 
+			model in a simple scenario before integrating it to the more complex 
+			WRENCH framework. 
+			In this prototype we used the following basic storage model for 
+			both memory and disk: 
 			\begin{align*}
@@
 
-			Since bandwidth sharing is not simulated, this prototype doesn't support 
-			concurrency: it is limited to single-threaded pipelines running on systems 
-			with a single-core CPU. We use it in a first validation of our simulation 
+			Since bandwidth sharing was not simulated, this prototype did not support 
+			concurrency: it was limited to single-threaded pipelines running on systems 
+			with a single-core CPU. We used it in a first validation of our simulation 
 			model, against a real sequential pipeline running on a real system.
@@
 
-			Having the model validated, we create simulators for different use cases 
+			Having the model validated, we created simulators for different use cases 
 			using WRENCH.
-			We integrate our I/O model to the original WRENCH by implementing 
-			the core of the model separately from existing WRENCH components, 
-			then providing an API for users to call to our simulation model. 
-			We also add an optional command line argument so that users will be able 
-			to choose use or not to use the model. Doing this, we can avoid 
-			unexpected impacts on original WRENCH and can independently update 
-			the model in the future. 
-			We then compare the results of the simulators with
-			original WRENCH and extended WRENCH with the results of real
-			pipelines on a real system.
+			We implemented our model separately from existing WRENCH components, 
+			exposed an API for users, and added an command line argument to enable 
+			the model in WRENCH. 
+			This helps us to avoid unexpected impacts on WRENCH and 
+			independently update the model in the future. 
+			We then compared the results of the WRENCH, WRENCH-Ext simulators 
+			the results of real execution on a real system.
 
@@
 			SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid}, 
-			and WRENCH is at \url{https://github.com/wrench-project/wrench}.
+			and WRENCH is at \url{https://github.com/dohoangdzung/wrench/tree/io_nfs} 	
+			\textcolor{red}{Update later with the commit used in the simulator}
 
@@
 			The pipeline is simulated with the Python prototype simulator, 
-			original WRENCH and WRENCH with page cache simulator.
+			original WRENCH and WRENCH-Ext simulator.
 
@@
 			In general, with all input sizes in the experiment, the simulation errors 
-			of simulators with page cache model (Python and WRENCH with page cache simulators) 
+			of simulators with page cache model (Python and WRENCH-Ext simulators) 
 			are remarkably lower than the result with original WRENCH . 
@@
 			increases but memory and disk bandwidths are fixed and shared between pipelines. 
-			In comparison with original WRENCH, the result from WRENCH with page cache are closer 
+			In comparison with original WRENCH, the results from WRENCH-Ext are closer 
 			to reality. 
 			Especially, when it comes to the average write time, 
-			although the trend in reality is more complex, our WRENCH with page cache 
+			although the trend in reality is more complex, our WRENCH-Ext 
 			simulator can still capture the pattern in real execution.
@@
 
-			To conclude, the results from WRENCH with page cache simulator show that 
+			To conclude, the results from WRENCH-Ext simulator show that 
 			our model can also simulate I/O with page cache accurately in a 
@@
 			Figure~\ref{fig:single_error} with 100 GB of input. 
-			When it comes to write time, original WRENCH and WRENCH with page cache 
+			When it comes to write time, original WRENCH and WRENCH-Ext 
 			results are identical when both are writing directly to disk.

### 2020-11-23 result/multi/process_result.py

@@ -78,9 +78,9 @@ def suplot_prop(ax, exp_folder, propname, title):
 
     ax.plot(real_df["no_pipeline"], real_df[propname] / real_df["no_pipeline"], color="k", label="Real execution")
     ax.plot(simg_org_df["no_pipeline"], simg_org_df[propname] / simg_org_df["no_pipeline"], color="tab:orange",
-            label="Original WRENCH")
+            label="WRENCH")
     ax.plot(simg_ext_df["no_pipeline"], simg_ext_df[propname] / simg_ext_df["no_pipeline"], color="tab:cyan",
-            label="WRENCH with page cache")
+            label="WRENCH-Ext")
 
     ax.set_xlabel("number of pipelines")
     ax.set_ylabel("time (s)")

### 2020-11-23 result/single/plot_memprof.py

@@ -13,7 +13,7 @@ def compare_size(axes, real_time_file, real_mem_file,
 
     # SIMULATION RESULTS
     sim_subplot(axes[1], pysim_time_file, pysim_mem_file, "Python simulator", bar_alpha=0.2)
-    sim_subplot(axes[2], simgrid_time_file, simgrid_mem_file, "WRENCH with page cache simulator", bar_alpha=0.2)
+    sim_subplot(axes[2], simgrid_time_file, simgrid_mem_file, "WRENCH-Ext simulator", bar_alpha=0.2)
 
     for ax in axes:
         ax.set_xlim(right=xmax, left=xmin)

### 2020-11-24 paper_io_sim.tex
more comments

--- 2020-11-18 paper_io_sim.tex
@@
 				\item $t_{w}$ is the data write time
-				\item $D$ is the amount of data to read/write
+				\item $D$ is the amount of data to read or write
 				\item $b_r$ is the read bandwidth of the device
@@
 
-			Since bandwidth sharing is not simulated, this prototype doesn't support 
+			Since bandwidth sharing is not simulated, this prototype does not support 
 			concurrency: it is limited to single-threaded pipelines running on systems 
@@
 			model, against a real sequential pipeline running on a real system.
-			Source code of our Python prototype is available at 
+			The source code of our Python prototype is available at 
             \url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}
@@
 			We integrate our I/O model in the original WRENCH by implementing 
-			the core of the model separately from existing WRENCH components, 
+			the core of the model separately from existing WRENCH components \tristan{why?}, 
 			then providing an API for users to call to our simulation model. 
@@
 			simulation in single-threaded and multi-threaded applications
-			accessing data on different types of file system. We use a pipeline
+			accessing data on different types of file systems \tristan{"different types" is vague, first thing I thought of was zfs vs ext4 that you don't discuss. Use "local and network" fs instead.}. 
+			We use a pipeline
 			of sequential tasks where each task reads an input file, increments
 			every byte of the input to emulate real processing and avoid caching
-			effects, and writes the result to disk. The output of the previous
+			effects \tristan{ambiguous: you're simulating caching effects, you don't want to avoid them. I'd just remove that note and leave "emulate real processing". You could comment on 
+			the close-to-zero execution time though, since modeling execution time is out of your scope.}, and writes the result to disk. The output of the previous
 			task is the input of the next task. We also use a real application 
@@
 
-			We use our dedicated cluster hosted at Concordia University to conduct 
-			the experiments. The cluster consists of one login node, 8 compute nodes 
-			and 4 storage nodes connected with a network bandwidth of 3GBps. 
+			We use a dedicated cluster hosted at Concordia University to conduct 
+			the experiments. The cluster consists of one login node, 8 compute nodes \tristan{it now has 9 if you add the GPU node}
+			and 4 storage nodes connected with a network bandwidth of 3GBps \tristan{put the Gbps value, 3GBps is unusual as a network bandwidth.}. 
 			The login node has one Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 
 			137~GB (128~GiB) of RAM, 1.8TB of storage of XFS file system, 
-			13~GB of tmpfs file system and 385 TB of Lustre shared file system. 
+			13~GB of tmpfs file system and 385 TB of Lustre shared file system \tristan{why do you need to describe the login node in such details?}. 
 			Each compute node has two 16 cores Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 
@@
 			Lustre file system is configured with one metadata target with 854~GB 
-			of storage, 44 object storage targets with 8.7~TB of storage each. 
+			of storage, 44 object storage targets with 8.7~TB of storage each \tristan{why do we care about Lustre?}. 
 			The cluster is run on CentOS 8.1 with NFS version 4. 

### 2020-11-30 paper_io_sim.tex
Update results with repetitions

--- 2020-11-23 paper_io_sim.tex
@@
 			disk bandwidths that will be released in the next SimGrid version.
-			The next step is to study the detail in cache eviction mechanism to achieve 
-			better simulation results. 
+			The next step is to study the detail in cache eviction mechanism to make 
+			the our simulation results closer to reality. 
+			Another point that needs to be further improved is the ability to configure  
+			different anonymous memory usage levels of applications since it affects  
+			the amount of cache used.
 			We also plan to support random I/O and file readahead, as well as 

### 2020-11-30 result/multi/process_result.py

@@ -3,6 +3,7 @@ import numpy as np
 import pandas as pd
 import csv
 import json
+import seaborn as sns
 
 
 def parse_single_pipeline(filename):
@@ -69,37 +70,51 @@ def parse_simgrid_result(filename, no_pipeline):
     return no_pipeline, makespan, read, write
 
 
-def suplot_prop(ax, exp_folder, propname, title):
-    real_df = pd.read_csv("%s/real/aggregated.csv" % exp_folder)
-    simg_org_df = pd.read_csv("%s/wrench/original/aggregated.csv" % exp_folder)
-    simg_ext_df = pd.read_csv("%s/wrench/pagecache/aggregated.csv" % exp_folder)
+def suplot_prop(ax, real_dir, wrench_dir, propname, title, rep_no=1):
+    simg_org_df = pd.read_csv("%s/original/aggregated.csv" % wrench_dir)
+    simg_ext_df = pd.read_csv("%s/pagecache/aggregated.csv" % wrench_dir)
+    no_pipeline_df = simg_org_df["no_pipeline"]
+
+    real_df = pd.DataFrame()
+    for i in range(rep_no):
+        df = pd.read_csv("%s/%d/aggregated.csv" % (real_dir, i + 1))
+        real_df[i + 1] = df[propname]
+
+    mean_df = real_df.mean(axis=1)
+    max_df = real_df.max(axis=1)
+    min_df = real_df.min(axis=1)
 
     ax.set_title(title)
 
-    ax.plot(real_df["no_pipeline"], real_df[propname] / real_df["no_pipeline"], color="k", label="Real execution")
-    ax.plot(simg_org_df["no_pipeline"], simg_org_df[propname] / simg_org_df["no_pipeline"], color="tab:orange",
-            label="WRENCH")
-    ax.plot(simg_ext_df["no_pipeline"], simg_ext_df[propname] / simg_ext_df["no_pipeline"], color="tab:cyan",
-            label="WRENCH-Ext")
+    ax.plot(no_pipeline_df, mean_df / no_pipeline_df, color="k", linewidth=1, label="Real execution mean")
+    ax.plot(no_pipeline_df, max_df / no_pipeline_df, color="k", linewidth=1, alpha=0.5)
+    ax.plot(no_pipeline_df, min_df / no_pipeline_df, color="k", linewidth=1, alpha=0.5)
[…]

### 2020-11-30 paper_io_sim.tex
Update README, source code url

@@
 			SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid}, 
-			and WRENCH is at \url{https://github.com/dohoangdzung/wrench/tree/io_nfs} 	
-			\textcolor{red}{Update later with the commit used in the simulator}
+			and WRENCH is at 
+			\url{https://github.com/wrench-project/wrench/tree/master/src/wrench/services/memory}.
 
@@
 			we use an infinite bandwidth to eliminate network latency for local I/Os.
-			The details of simulated platform is available here: 
-			\url{https://github.com/dohoangdzung/wrench/blob/io_chunk_test/examples/basic-examples/io-pagecache/single_host.xml}
+			The details of simulated platform and the source code of the simulator 
+			are available here: 
+			\url{https://github.com/wrench-project/wrench/tree/master/examples/basic-examples/io-pagecache}
 

### 2020-12-02 paper_io_sim.tex

--- 2020-11-30 paper_io_sim.tex
@@
 
-			Since bandwidth sharing was not simulated, this prototype did not support 
-			concurrency: it was limited to single-threaded pipelines running on systems 
+			Since bandwidth sharing was not simulated, this prototype does not support 
+			concurrency: it is limited to single-threaded pipelines running on systems 
 			with a single-core CPU. We used it in a first validation of our simulation 
 			model, against a real sequential pipeline running on a real system.
-			Source code of our Python prototype is available at 
+			The source code of our Python prototype is available at 
 			\url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}
@@
 			The I/O time of tasks in simulators and real execution is also tracked and compared. 
-			Figure~\ref{fig:single_errors} shows the relative error of simulated I/O time 
+			Figure~\ref{fig:single_errors} \tristan{wrong reference} shows the relative error of simulated I/O time 
 			of individual I/O activities. 
@@
 			disk bandwidths that will be released in the next SimGrid version.
-			The next step is to study the detail in cache eviction mechanism to make 
-			the our simulation results closer to reality. 
-			Another point that needs to be further improved is the ability to configure  
-			different anonymous memory usage levels of applications since it affects  
-			the amount of cache used.
+			The next step is to study the detail in cache eviction mechanism to achieve 
+			better simulation results. 
 			We also plan to support random I/O and file readahead, as well as 

### 2020-12-02 paper_io_sim.tex
comments on results

@@
 			\url{https://github.com/wrench-project/wrench/tree/master/examples/basic-examples/io-pagecache}
+			\tristan{Our experiments use commit abcd}
 
@@
 			The pipeline is simulated with the Python prototype simulator, 
-			original WRENCH and WRENCH-Ext simulator.
+			original WRENCH and WRENCH-Ext simulator \tristan{name your simulator something 
+			that refers to page cache, maybe WRENCH-cache}.
 
@@
 		The benchmark results of bandwidths and the values used in the simulators 
-		in our all experiments are shown in \ref{table:benchmark}.
+		in our all experiments are shown in Table~\ref{table:benchmark} \tristan{This isn't results, it should go in II.C, 
+		next to the paragraph where you talk about symmetrical bandwidths}.
 
@@
 		\end{tabular}
-		\caption{Bandwidth benchmark results and simulator configurations}
+		\caption{Bandwidth benchmark results and simulator configurations \tristan{add a sentence of explanation. Make the table fit in 1 column}}
 		\label{table:benchmark}
@@
 			and the memory profiling results with 20GB and 100GB input files 
-			are illustrated in Figure~\ref{fig:single_memprof}. 
+			are illustrated in Figure~\ref{fig:single_memprof} \tristan{convoluted writing}. 
 			The bars in the figures present activities in tasks (read, compute and write), 
-			while the lines describe memory status throughout run time of the pipeline. 
+			while the lines describe memory status throughout run time of the pipeline \tristan{move any description of graphical features to the figure caption}. 
 			Because the results with 20Gb, 50GB, 100GB are analogous, we only include 
-			the results with 20GB and 100GB in this paper.
+			the results with 20GB and 100GB in this paper \tristan{analogous is not the right word}.
 
@@
 			in the lines describing memory status in simulators and the real execution. 
-			The akin movements of amount of cache used (purple lines) in reality 
+			The similar evolutions of amount of cache used (purple lines) in reality 
 			and simulation show that the use of page cache is well reproduced by the simulators. 
@@
 			\end{subfigure}
-			\caption{Memory profiling results with different input file sizes}
+			\caption{Memory profiling results with different input file sizes \tristan{put the simulation/real conditions on the side, make them larger and use the same color code as below (pink for real, etc).
+Put the memory values on top, similar to how they are in the graphs below.
+}}
 			\label{fig:single_memprof}	
@@
 			\label{fig:single_cache}
-   			\caption{Amount of file data in cache after each I/O activity with 100GB of input}
+   			\caption{Amount of file data in cache after each I/O activity with 100GB of input \tristan{Make a grouped histogram (1 histogram with two bars, one for real, one for simulation).
+Add a grouped histogram for 20GB.
+Use the same color code everywhere: real should be pink.}}
 			\end{figure*}
@@
 			The results shows our achievements in terms of simulation accuracy 
-			and the application potential in practice. 
+			and the application potential in practice \tristan{what do you conclude? can we use your simulator to study big data applications on hpc clusters? under what conditions?}. 
 			On the other hands, the model exposed some limitations that can be 

### 2020-12-03 paper_io_sim.tex
comments on results
comments and edits in method

@@
 		page cache, cache eviction and data flushing mechanisms implemented 
-		in the Linux kernel. 
-		We also detail the design of our simulator and its implementation. 
-		Because the simulator is implemented with SimGrid using WRENCH, we 
-		name the simulator as WRENCH-Ext, along with the WRENCH simulator,
-		which is created with original WRENCH.
-		Finally, we describe experimental scenarios to evaluate our simulation model 
-		using a standalone Python prototype, WRENCH and WRENCH-Ext simulator.
+		in the Linux kernel. We also detail the design and implementation of two simulators:
+		(1) a standalone Python prototype used to evaluate the model independently of any other 
+		library, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
+		Finally, we describe experimental scenarios to evaluate our simulations in various configurations.
 
@@
 			Controller and the Memory Manager, which together simulate 
-			file read/write mechanisms in Linux (Figure~\ref{fig:interaction}). 
-			The Memory Manager simulates I/Os to cache, flushing, periodical flushing, 
-			and cache eviction, whereas the IO Controller simulates file reads and writes. 			
-			To read or write a chunk of file, a simulated application sends a request 
-			to the IOController.
-			Based on the memory status in Memory Manager, the IOController 
-			orchestrates cache and disk reads/writes, flushing, cache eviction 
-			to simulate I/O time.
+			file read and write mechanisms in Linux (Figure~\ref{fig:interaction}). 
+			To read or write a file chunk, a simulated application sends a request 
+			to the IOController. The IOController determines the amount of memory required by the operation,
+			interacts with the Memory Manager accordingly, reads or write to cache through the Memory Manager, 
+			and finally reads or writes on disk if necessary.
+			The Memory Manager simulates flushing, eviction, and I/O to cache.
 
@@
 				   \caption{Overview of the simulated writeback page cache.
-				   Application sends file read and file write requests to IOController, 
-				   then IOController manages orchestrate flushing, eviction, cache and 
-				   disk accesses to fulfill the requests with the MemoryManager object. 
-				   Memory and Disk are responsible for simulating I/O time.}
+				   Applications send file read or write requests to the
+				   IOController that orchestrates flushing, eviction, cache
+				   and disk accesses with the MemoryManager. Storage
+				   devices (memory and disk) simulate file transfers with
+				   bandwidth sharing. \tristan{Could you annotate arrows to
+				   make them more explicit?}.}
 				   \label{fig:interaction}
@@
 
-			We leverage existing storage simulation models to simulate the memory,
-			which is characterized with storage capacity, read and write bandwidths 
+			We leverage existing storage simulation models~\cite{lebre2015} to simulate the memory,
+			characterized by its storage capacity, read and write bandwidths,
 			and device latency.
-			These features are captured in simulation frameworks and applicable 
-			to memory simulation \cite{lebre2015}. 
-			In addition, leveraging existing disk models allows us 
-			to take advantage of bandwidth sharing between concurrent disk I/O to 
+			Leveraging existing storage models allows us 
+			to take advantage of bandwidth sharing between concurrent I/Os to 
 			simulate parallel memory accesses.
@@
 			\paragraph*{Memory Manager}
-			Memory Manager simulates two parallel threads presented in \ref{fig:interaction}.
-			The first thread is the main thread, which is responsible for the simulation 
-			of data flushing, cache eviction and I/Os to cache. 
-			The second thread is a simulated background thread, in which Memory Manager 
-			periodically searches and flushes expired dirty data in page cache LRU lists 
-			to disks.
+			The Memory Manager simulates  two parallel threads: the first and main one simulates 
+			flushing, eviction, and cache I/Os synchronously; the second one is a background thread that  
+			periodically searches for expired dirty data in page cache LRU lists 
+			and flushes them to disk.
 
 			\paragraph*{Page cache LRU lists}
-			In Linux kernel, the page cache LRU lists contain file pages. However, 
+			In the Linux kernel, the page cache LRU lists contain file pages. However, 
 			due to the large number of file pages, managing lists of pages may 
@@
 			\end{subfigure}		
-			\caption{I/O time of concurrent pipelines with local storage}
+			\caption{I/O time of concurrent pipelines with local storage. \tristan{revise caption for real execution variability: "real execution" doesnt tell what the gray shade is.}}
 			\label{fig:multi_local}

### 2020-12-03 paper_io_sim.tex

@@
 			Controller and the Memory Manager, which together simulate 
-			file read and write mechanisms in Linux (Figure~\ref{fig:interaction}). 
+			file read and write mechanisms in Linux (Fig.~\ref{fig:interaction}). 
 			To read or write a file chunk, a simulated application sends a request 
@@
 			In the Linux kernel, the page cache LRU lists contain file pages. However, 
-			due to the large number of file pages, managing lists of pages may 
-			require significant computation overhead. 
+			due to the large number of file pages, managing lists of pages 
+			induces substantial overheads.
 			Thus, we introduce the concept of a data block as a unit to represent data 
@@
    				\includegraphics[width=\columnwidth]{figures/lru_lists.pdf}
-   				\caption{Simulation of page cache LRU lists with data blocks}	\label{fig:lrulist}
+   				\caption{Model of page cache LRU lists with data blocks.}	\label{fig:lrulist}
 			\end{figure}
 
-			With the data block definition, page cache LRU lists can be modeled with 
-			two lists of data blocks: an active list and an inactive list, ordered by 
-			last access time (earliest first, see Figure~\ref{fig:lrulist}).
-			As in the kernel, our simulator keeps the size of the active list under
-			twice the size of the inactive list by moving least recently 
+			We model page cache LRU lists as 
+			two lists of data blocks, an active list and an inactive list, both ordered by 
+			last access time (earliest first, Fig.~\ref{fig:lrulist}).
+			As in the kernel, our simulator limits the size of the active list to
+			twice the size of the inactive list, by moving least recently 
             used data blocks from the active list to the inactive list.
@@
             from disk first, and cached data is only read after all uncached data is read 
-            (left to right in Figure~\ref{fig:read_order}).
+            (left to right in Fig.~\ref{fig:read_order}).
             When a chunk of uncached data is read, a new clean block is created 

### 2020-12-05 paper_io_sim.tex

@@
 
-		The explosive development of Big Data era seen in recent years results in 
-		growing needs in data storing, data analysis. 
-		Although large computing systems with powerful computing capability 
-		have been increasingly used to process big datasets, there is still a bottleneck 
-		in disk I/O. 
-		Memory page cache is a solution applied to minimize disk accesses and 
-		improve the I/O performance. 
-		Despite of the fact that simulation tools for HPC have been developed to 
+		The rapid emergence of Big Data seen in recent years has resulted in a 
+		growing needs for efficient data storing and optimized data analysis. 
+		Despite relying on large computing infrastructures with powerful computing capability 
+		to process these large datasets, an I/O bottleneck remains. 
+        One widely-used approach to minimize the costs of data transfers during
+        regular use is through the employment of the page cache.
+		While many simulation tools have been developed to 
 		support studies in computing systems optimization, page cache simulation 
-		has not been fully captured, or absent in these simulators. 
+		has not been fully captured or remains absent from these simulators. 
 		The lack of ability to simulate I/O with page cache leads to inaccurate 
-		simulation results in many situations, especially in data intensive applications.
+		simulation results in many situations, especially when dealing with data intensive applications.
 
-		In this paper, we propose a I/O simulation model with key features of the page cache.
-		The model consists of simulation algorithms of I/O, main components 
-		and mechanisms of page cache.
-		We also implement the model with SimGrid, a simulation framework for 
+		In this paper, we propose a I/O simulation model that characterizes the key features of the page cache.
+		We have implemented the model with SimGrid, a simulation framework for 
 		distributed systems, using WRENCH, a SimGrid-based workflow simulator. 
-		Then we create simulators for applications in single-threaded, multi-threaded,
-		network file system and neuroimaging usecases.
-		Finally, we analyze the results, compare to the results of base-line WRENCH 
-		simulator as well as real execution.
+        Our simulator is compatible with both single-threaded and multithreaded applications,
+        and can simulate both writeback and writethrough cache that can factor into local
+        and network-based filesystems.
 		The results show that our simulator not only achieves higher accuracy than 
-		original WRENCH, but also simulates well the status of memory and 
+		original WRENCH, but also effectively simulates memory contents and 
 		page cache during I/O operations. 

### 2020-12-07 paper_io_sim.tex
Replaced tabs by spaces

@@
 
-		In this section, we present our approach to model file read and write with 
-		page cache, cache eviction and data flushing mechanisms implemented 
-		in the Linux kernel. We also detail the design and implementation of two simulators:
-		(1) a standalone Python prototype used to evaluate the model independently of any other 
-		library, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
-		Finally, we describe experimental scenarios to evaluate our simulations in various configurations.
+        In this section, we present our page cache simulation model for file I/O,
+        flusing, and eviction. We detail the design and implementation of two simulators:
+        (1) a standalone Python prototype to evaluate the simulation model independently of any other 
+        tool, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
+        We also describe experimental scenarios to evaluate our simulators in various configurations.
 
@@
             Controller and the Memory Manager, which together simulate 
-			file read and write mechanisms in Linux (Fig.~\ref{fig:interaction}). 
-			To read or write a file chunk, a simulated application sends a request 
-			to the IOController. The IOController determines the amount of memory required by the operation,
-			interacts with the Memory Manager accordingly, reads or write to cache through the Memory Manager, 
-			and finally reads or writes on disk if necessary.
-			The Memory Manager simulates flushing, eviction, and I/O to cache.
+            file reads and writes in Linux (Fig.~\ref{fig:interaction}). 
+            To read or write a file chunk, a simulated application sends a
+            request to the IOController. The IOController interacts as needed with
+            the Memory Manager to free memory through flushing or eviction,
+            and to read or write to cache or to disk. The Memory Manager
+            implements these operations, and simulates periodical flushing
+            and eviction from cache.
 
@@
 
-			We leverage existing storage simulation models~\cite{lebre2015} to simulate the memory,
+            We use existing storage simulation models~\cite{lebre2015} to simulate memory,
             characterized by its storage capacity, read and write bandwidths,
-			and device latency.
-			Leveraging existing storage models allows us 
-			to take advantage of bandwidth sharing between concurrent I/Os to 
-			simulate parallel memory accesses.
+            and latency. Reusing existing storage models allows us 
+            to simulate bandwidth sharing between concurrent memory accesses.
 
             \paragraph*{Memory Manager}
-			The Memory Manager simulates  two parallel threads: the first and main one simulates 
-			flushing, eviction, and cache I/Os synchronously; the second one is a background thread that  
-			periodically searches for expired dirty data in page cache LRU lists 
-			and flushes them to disk.
+            The Memory Manager simulates two parallel threads: the main one
+            implements flushing, eviction, and cache I/Os synchronously,
+            whereas the second one, operating in the background,
+            periodically searches for expired dirty data in page cache LRU
+            lists and flushes them to disk.
 
             \paragraph*{Page cache LRU lists}
-			In the Linux kernel, the page cache LRU lists contain file pages. However, 
+            In the Linux kernel, page cache LRU lists contain file pages. However, 
             due to the large number of file pages, managing lists of pages 
@@
             simulation in single-threaded and multi-threaded applications
-			accessing data on different types of file systems \tristan{"different types" is vague, first thing I thought of was zfs vs ext4 that you don't discuss. Use "local and network" fs instead.}. 
+            accessing data on different types of file systems \tristan{"different types" is vague, first thing I thought of was zfs vs ext4 that you don't discuss.
+Use "local and network" fs instead.}.
             We use a pipeline

### 2020-12-07 paper_io_sim.tex
Update paper text, plotting scripts

@@
 \usepackage{ulem}
+\usepackage{tabularx}
 
@@
                    \caption{Overview of the simulated writeback page cache.
-                   Applications send file read or write requests to the
-                   IOController that orchestrates flushing, eviction, cache
-                   and disk accesses with the MemoryManager. Storage
-                   devices (memory and disk) simulate file transfers with
-                   bandwidth sharing. \tristan{Could you annotate arrows to
-                   make them more explicit?}.}
+                   Application sends file read and file write requests to IOController,
+                   then IOController manages orchestrate flushing, eviction, cache and
+                   disk accesses to fulfill the requests with the MemoryManager object.
+                   Memory and Disk are responsible for simulating I/O time.}
                    \label{fig:interaction}
@@
             independently update the model in the future.
-            We then compared the results of the WRENCH, WRENCH-Ext simulators 
+            We then compared the results of the WRENCH, WRENCH-cache simulators
             the results of real execution on a real system.
@@
             simulation in single-threaded and multi-threaded applications
-            accessing data on different types of file systems \tristan{"different types" is vague, first thing I thought of was zfs vs ext4 that you don't discuss.
-Use "local and network" fs instead.}.
-            We use a pipeline
-            of sequential tasks where each task reads an input file, increments
-            every byte of the input to emulate real processing and avoid caching
-            effects \tristan{ambiguous: you're simulating caching effects, you don't want to avoid them. I'd just remove that note and leave "emulate real processing". You could comment on 
-            the close-to-zero execution time though, since modeling execution time is out of your scope.}, and writes the result to disk. The output of the previous
-            task is the input of the next task. We also use a real application 
-            pipeline \textcolor{red}{[Maybe briefly describe this application's purpose (e.g. a standard fMRI preprocessing piepeline)]}, to evaluate the applicability of our simulation model.
+            accessing data on local and network file systems.
+            We use a pipeline of sequential tasks where each task reads an input file,
+            increments every byte of the input to emulate real processing, and writes
+            the result to disk.
+            The output of the previous task is the input of the next task.
+            We also use a real application pipeline \textcolor{red}{[Maybe briefly describe this application's purpose (e.g. a standard fMRI preprocessing piepeline)]},
+            to evaluate the applicability of our simulation model.
 
             We use a dedicated cluster hosted at Concordia University to conduct
-            the experiments. The cluster consists of one login node, 8 compute nodes \tristan{it now has 9 if you add the GPU node}
-            and 4 storage nodes connected with a network bandwidth of 3GBps \tristan{put the Gbps value, 3GBps is unusual as a network bandwidth.}. 
-            The login node has one Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 
-            137~GB (128~GiB) of RAM, 1.8TB of storage of XFS file system, 
-            13~GB of tmpfs file system and 385 TB of Lustre shared file system \tristan{why do you need to describe the login node in such details?}. 
+            the experiments. The cluster consists of one login node, 9 compute nodes
+            and 4 storage nodes connected with a network bandwidth of 25 Gbps.
             Each compute node has two 16 cores Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz,
@@
             378~GB of tmpfs and 126~GB of devtmpfs file system.
-            Lustre file system is configured with one metadata target with 854~GB 
-            of storage, 44 object storage targets with 8.7~TB of storage each \tristan{why do we care about Lustre?}. 
             The cluster is run on CentOS 8.1 with NFS version 4.
@@
             take the average bandwidths.
+            The benchmark results of bandwidths and the values used in the simulators
+            in our all experiments are shown in Table~\ref{table:benchmark}.
+
+            \begin{table}[htbp]
+            \centering
+            \begin{tabularx}{\columnwidth}{|l
+            |>{\centering\arraybackslash}X
+            |>{\centering\arraybackslash}X
+            |>{\centering\arraybackslash}X|}
+            \hline
+                Bandwidths (MBps)  & Cluster (real) & Python simulator & WRENCH simulator\\
+            \hline
+                Memory read  & 6860	& 4812	 & 4812\\
+                Memory write & 2764	& 4812 & 4812\\
+                Local disk read & 510 & 465 & 465\\
+                Local disk write & 420 & 465	 & 465\\
+                Remote disk read & 515 & - & 445\\
+                Remote disk write & 375 & - & 445\\
+                Network bandwidth & 3000 & - & 3000\\
+            \hline
+            \end{tabularx}
+            \caption{Bandwidth benchmark results and simulator configurations.
+            The bandwidths used in simulators are average of read and write bandwidths
+            measured on the cluster.
+            Remote disk and network are not simulated in Python simulator.}
+            \label{table:benchmark}
+            \end{table}
+
             As the pipelines are single-core, we create simulated tasks,
@@
             The details of simulated platform and the source code of the simulator
-            are available here: 
-            \url{https://github.com/wrench-project/wrench/tree/master/examples/basic-examples/io-pagecache}
-            \tristan{Our experiments use commit abcd}
+            used in our experiments are available in this commit:
+            \url{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}.
 
@@
             The pipeline is simulated with the Python prototype simulator,
-            original WRENCH and WRENCH-Ext simulator \tristan{name your simulator something 
-            that refers to page cache, maybe WRENCH-cache}.
+            original WRENCH and WRENCH-cache simulator.
 
@@
             we only compare the results of simulators with original WRENCH, 
-            WRENCH-Ext and real pipelines in this experiment. 
+            WRENCH-cache and real pipelines in this experiment. 
             The results are compared in total makespan of the pipelines, 
@@
             We simulate this pipeline with simulators using original WRENCH and 
-            WRENCH-Ext. 
+            WRENCH-cache. 
             The simulation results are then compared with the real results. 
@@
     \label{results}
-        The benchmark results of bandwidths and the values used in the simulators 
-        in our all experiments are shown in Table~\ref{table:benchmark} \tristan{This isn't results, it should go in II.C, 
-        next to the paragraph where you talk about symmetrical bandwidths}.
-
-        \begin{table}
-        \centering
-        \begin{tabular}{|l|c|c|c|}
-        \hline
-            Bandwidths (MBps)  & Cluster (real) & Python simulator & WRENCH simulator\\
-        \hline
-            Memory read  & 6860    & 4812     & 4812\\
-            Memory write & 2764    & 4812 & 4812\\
-            Local disk read & 510 & 465 & 465\\
-            Local disk write & 420 & 465     & 465\\
-            Remote disk read & 515 & - & 445\\
-            Remote disk write & 375 & - & 445\\
-            Network bandwidth & 3000 & - & 3000\\
-        \hline
-        \end{tabular}
-        \caption{Bandwidth benchmark results and simulator configurations \tristan{add a sentence of explanation. Make the table fit in 1 column}}
-        \label{table:benchmark}
-        \end{table}
 
@@
 
-            Memory behaviors of real pipeline execution and simulators are recorded 
-            and the memory profiling results with 20GB and 100GB input files 
-            are illustrated in Figure~\ref{fig:single_memprof} \tristan{convoluted writing}. 
-            The bars in the figures present activities in tasks (read, compute and write), 
-            while the lines describe memory status throughout run time of the pipeline \tristan{move any description of graphical features to the figure caption}. 
-            Because the results with 20Gb, 50GB, 100GB are analogous, we only include 
-            the results with 20GB and 100GB in this paper \tristan{analogous is not the right word}.
+            Memory profiling results of real pipeline execution and simulators 
+			with 20GB and 100GB input files are illustrated in Fig.~\ref{fig:single_memprof}.
+			Because the results with 20Gb, 50GB, 100GB shows a similar pattern,
+			we only include the results with 20GB and 100GB in this paper.
 
@@
 			and simulation show that the use of page cache is well reproduced by the simulators.
-            The amount of dirty data ( red lines) also follow similar trends, 
+			The amount of dirty data (red lines) also follow similar trends,
 			which means not only the amount dirty data, but also the flushing and
@@
 			The I/O time of tasks in simulators and real execution is also tracked and compared.
-            Figure~\ref{fig:single_errors} \tristan{wrong reference} shows the relative error of simulated I/O time 
+			Fig.~\ref{fig:single_error} shows the relative error of simulated I/O time
 			of individual I/O activities.
@@
 
-            In general, with all input sizes in the experiment, the simulation errors 
-            of simulators with page cache model (Python and WRENCH-Ext simulators) 
-            are remarkably lower than the result with original WRENCH . 
-            The reason is, in original WRENCH, because page cache is not simulated, 
-            data is only read/written with disk bandwidth, which is much slower 
-            than memory bandwidths, leading to simulated I/O time longer than in reality.
-            Especially, the simulation time errors of \textit{Read 3} in the simulators 
-            with page cache are just under the error in original WRENCH.
-
 			\begin{figure*}
@@
 			\end{subfigure}
-            \caption{Memory profiling results with different input file sizes \tristan{put the simulation/real conditions on the side, make them larger and use the same color code as below (pink for real, etc).
-Put the memory values on top, similar to how they are in the graphs below.
-}}
+			\caption{Memory profiling results with different input file sizes.
+			The bars in the figures present activities in tasks (read, compute and write),
+			the lines describe memory status throughout run time of the pipeline.}
 			\label{fig:single_memprof}
@@
 			\label{fig:single_cache}
-               \caption{Amount of file data in cache after each I/O activity with 100GB of input \tristan{Make a grouped histogram (1 histogram with two bars, one for real, one for simulation).
-Add a grouped histogram for 20GB.
-Use the same color code everywhere: real should be pink.}}
+   			\caption{Amount of file data in cache after each I/O operation. 
+   			\textcolor{red}{Update real results of 20GB}}
 			\end{figure*}
 
+			In general, with all input sizes in the experiment, the simulation errors
+			of simulators with page cache model (Python and WRENCH-cache simulators)
+			are remarkably lower than the result with original WRENCH .
+			The reason is, in original WRENCH, because page cache is not simulated,
+			data is only read and written with disk bandwidth, which is much slower
+			than memory bandwidths, leading to simulated I/O time longer than in reality.
+			Especially, the simulation time errors of \textit{Read 3} with 100 GB of input
+			in the simulators with page cache are just under the error in original WRENCH.
 			This longer read can be explained with the amount of cached data of files
 			at the beginning of the read in reality and simulators shown in
-            Figure~\ref{fig:single_cache}.             
+			Fig.~\ref{fig:single_cache}.
 			In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
 			output of \textit{Write $<$i-1$>$}.
-            In the real execution, after \textit{Write 2}, which writes \textit{file 3}, 
-            all data of \textit{file 3} is in cache. 
+			With the input size of 20 GB, in both real execution and WRENCH-cache
+			simulator, there is no difference in the page cache since all files in the
+			pipeline can be entirely cached.
+			With the input of 100 GB, the difference between real execution and WRENCH-cache
+			simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
+			In real execution, after being written by \textit{Write 2},
+			\textit{file 3} is entirely in cache.
 			Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
-            But in our simulator, after \textit{Write 2}, only a part of \textit{file 3} 
-            remains in cache. 
-            The reason is while \textit{file 3}  is being written, the page cache is saturated, 
-            eviction is then evoked to evict data from page cache, make space 
-            available for \textit{file 3}. 
+			However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
+			cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
+			\textit{file 3} from disk.
+			The reason for the difference is in our WRENCH-cache simulator,
+			with 100 GB of input, while \textit{file 3} is being written,
+			the page cache is saturated, eviction is then evoked to evict data
+			from page cache and make space available for \textit{file 3}.
 			However, because the evicted amount is less than required,
@@
             \end{subfigure}
-            \caption{I/O time of concurrent pipelines with local storage. \tristan{revise caption for real execution variability: "real execution" doesnt tell what the gray shade is.}}
+            \caption{I/O time of concurrent pipelines with local storage.}
             \label{fig:multi_local}
@@
 
-            In this experiment, the results are analyzed in terms of average I/O time 
-            of concurrent pipelines shown in Figure~\ref{fig:multi_local}.
+            In this experiment, we run the real execution in 5 repetitions to
+            capture the variability on the cluster.
+            The results are analyzed in terms of average I/O time
+            of concurrent pipelines shown in Fig.~\ref{fig:multi_local}.
 
@@
             increases but memory and disk bandwidths are fixed and shared between pipelines.
-            In comparison with original WRENCH, the results from WRENCH-Ext are closer 
+            In comparison with original WRENCH, the results from WRENCH-cache are closer
             to reality.
             Especially, when it comes to the average write time,
-            although the trend in reality is more complex, our WRENCH-Ext 
+            although the trend in reality is more complex, our WRENCH-cache
             simulator can still capture the pattern in real execution.
@@
 
-            To conclude, the results from WRENCH-Ext simulator show that 
+            To conclude, the results from WRENCH-cache simulator show that
             our model can also simulate I/O with page cache accurately in a
@@
 
-            Similarly, the results with from the experiment with NFS are described 
-            by the average I/O time shown in Figure~\ref{fig:multi_nfs}.
+            Similar to the previous experiment, the real execution with NFS is also
+            repeated for 5 times and the results are described with the average I/O time
+            shown in Fig.~\ref{fig:multi_nfs}.
             In general, the average I/O time reflect a very similar level of accuracy
@@
             the page cache is saturated, which has been seen in
-            Figure~\ref{fig:single_error} with 100 GB of input. 
-            When it comes to write time, original WRENCH and WRENCH-Ext 
+            Fig.~\ref{fig:single_error} with 100 GB of input.
+            When it comes to write time, original WRENCH and WRENCH-cache
             results are identical when both are writing directly to disk.
@@
         with different experiments.
-            The results shows our achievements in terms of simulation accuracy 
-            and the application potential in practice \tristan{what do you conclude? can we use your simulator to study big data applications on hpc clusters? under what conditions?}. 
-            On the other hands, the model exposed some limitations that can be 
+        The results shows that our model can result in better simulation accuracy than
+        the simulators in which page cache is not modeled.
+        Furthermore, this level of accuracy can be achieved without a fine-grained
+        level of details.
+        This may create opportunities for the model to be applied in simulation of
+        big data applications on HPC infrastructures, which leverage memory to
+        employ page cache as a disk cache.
+        Nevertheless, the model exposed some limitations that can be
         further improved.

### 2020-12-07 result/single/plot_cache.py

@@ -1,58 +1,63 @@
 import pandas as pd
 import matplotlib.pyplot as plt
+from matplotlib.patches import Patch
 
 
-def plot_cache(real_file, sim_file):
+def plot_cache(ax, real_file, sim_file, patterns, color_real="gainsboro", color_wrench="tab:cyan"):
     index = ["file_1", "file_2", "file_3", "file_4"]
     columns = ["Read 1", "Write 1", "Read 2", "Write 2", "Read 3", "Write 3"]
-    files = ["file 1", "file 2", "file 3", "file 4"]
 
     df_real = pd.read_csv(real_file, index_col="task")
     df_sim = pd.read_csv(sim_file, index_col="task")
 
+    df_real[index].plot.bar(stacked=True, rot=0, color=color_real, edgecolor="k", ax=ax, position=1, width=0.3, legend=False, linewidth=0.5)
+    df_sim[index].plot.bar(stacked=True, rot=0, color=color_wrench, edgecolor="k", ax=ax, position=0, width=0.3, legend=False, linewidth=0.5)
+
+    ax.set_xlabel("tasks")
+    ax.set_ylabel("amount (GB)")
+    ax.set_xticklabels(columns)
+
+    bars = ax.patches
+    hatches = []  # list for hatches in the order of the bars
+    for h in patterns:  # loop over patterns to create bar-ordered hatches
+        for i in range(len(columns)):
+            hatches.append(h)
+    for h in patterns:  # loop over patterns to create bar-ordered hatches
+        for i in range(len(columns)):
+            hatches.append(h)
+    for bar, hatch in zip(bars, hatches):  # loop over bars and hatches to set hatches in correct order
+        bar.set_hatch(hatch)
+
+    ax.set_xlim(left=-0.5)
+    ax.set_ylim(top=180)
+
+
+def plot_cache_v2():
+
+    files = ["file 1", "file 2", "file 3", "file 4"]
+    envs = ["Real execution", "WRENCH-cache"]
     patterns = ["", "...", "///", "\\\\\\"]
+    color_real = "gainsboro"
+    color_wrench = "tab:cyan"
 
-    plt.rcParams.update({'font.size': 8})
     fig, (ax1, ax2) = plt.subplots(figsize=(11, 4), ncols=2, nrows=1)
-    plt.subplots_adjust(left=0.12, bottom=0.1, right=0.92, top=0.8)
[…]

### 2020-12-07 paper_io_sim.tex
edits in methods

--- 2020-12-07 paper_io_sim.tex
@@
     \label{method}
-
     In this section, we present our page cache simulation model for file I/O,
@@
 
-        \subsection{Principle of the simulator}
-
     We separate our simulation model in two components, the IO
     Controller and the Memory Manager, which together simulate 
-            file reads and writes in Linux (Fig.~\ref{fig:interaction}). 
+    file reads and writes in Linux (Figure~\ref{fig:interaction}). 
     To read or write a file chunk, a simulated application sends a
@@
     the Memory Manager to free memory through flushing or eviction,
-            and to read or write to cache or to disk. The Memory Manager
-            implements these operations, and simulates periodical flushing
-            and eviction from cache.
-
+    and to read or write to cache. The Memory Manager
+    implements these operations, simulates periodical flushing
+    and eviction from cache, and reads or writes to disk when necessary.
     \begin{figure}
@@
 
-            \subsubsection{Memory simulation model}
+    \subsection{Memory Manager}
 
-            We use existing storage simulation models~\cite{lebre2015} to simulate memory,
-            characterized by its storage capacity, read and write bandwidths,
-            and latency. Reusing existing storage models allows us 
-            to simulate bandwidth sharing between concurrent memory accesses.
+    The Memory Manager simulates two parallel threads: the main one
+    implements flushing, eviction, and cache I/Os synchronously, whereas
+    the second one, operating in the background, periodically searches for
+    expired dirty data in page cache LRU lists and flushes them to disk. We
+    use existing storage simulation models~\cite{lebre2015} to simulate
+    memory, characterized by its storage capacity, read and write
+    bandwidths, and latency. Reusing existing storage models allows us to
+    simulate bandwidth sharing between concurrent memory accesses.
 
-            \paragraph*{Memory Manager}
-            The Memory Manager simulates two parallel threads: the main one
-            implements flushing, eviction, and cache I/Os synchronously,
-            whereas the second one, operating in the background,
-            periodically searches for expired dirty data in page cache LRU
-            lists and flushes them to disk.
+    \subsubsection{Page cache LRU lists}
 
-            \paragraph*{Page cache LRU lists}
     In the Linux kernel, page cache LRU lists contain file pages. However, 
@@
     induces substantial overheads.
-            Thus, we introduce the concept of a data block as a unit to represent data 
+    Therefore, we introduce the concept of a data block as a unit to represent data 
     cached in memory. A data block is a subset of file pages stored in
@@
     time, a dirty flag that represents whether the data is clean (0) 
-            or dirty (1), and the time it is written to cache as dirty data. 
+    or dirty (1), and an entry (creation) time.
     Blocks can have different sizes and a given file can have multiple 
@@
     arbitrary number of smaller blocks.
-
     \begin{figure}
@@
     two lists of data blocks, an active list and an inactive list, both ordered by 
-            last access time (earliest first, Fig.~\ref{fig:lrulist}).
+    last access time (earliest first, Figure~\ref{fig:lrulist}).
     As in the kernel, our simulator limits the size of the active list to
     twice the size of the inactive list, by moving least recently 
-            used data blocks from the active list to the inactive list.
-            Modeling page cache LRU lists as lists of data blocks reduces the
-            overhead of the simulator, as data blocks are obviously less
-            numerous than file pages, and at the same time achieves the accuracy in 
-            I/O time simulation that can be seen in the later section.
+    used data blocks from the active list to the inactive list~\cite{gorman2004understanding, linuxdev3rd2010}.
+    % Modeling page cache LRU lists as lists of data blocks reduces the
+    % overhead of the simulator, as data blocks are obviously less
+    % numerous than file pages. At the same time, it achieves the accuracy in 
+    % I/O time simulation that can be seen in the later section.
+    % From Tristan: I don't think the sentences above are necessary,
+    % they are quite repetitive.
 
-            At a specific time, an arbitrary file can be partially cached, completely cached, 
-            or uncached.
-            Cached data is composed of data blocks existing in one or both
-            of the page cache LRU lists. 
-            Initially, blocks accessed for the first time are added to the inactive list in 
-            page cache. 
-            If the blocks residing the inactive list are subsequently accessed, 
-            they will be moved to the top of the active list
-            Cached blocks written to cache are marked as dirty blocks until 
-            they are flushed to disk.
+    At any given time, a file can be partially cached, completely cached,
+    or uncached at all. Cached data blocks may reside in one or both of the
+    page cache LRU lists. The first time they are accessed, blocks are
+    added to the inactive list. On subsequent accesses, blocks of the
+    inactive list are moved to \tristan{if they are moved then they can't
+    be in both lists} the top of the active list. Cached blocks written to
+    cache are marked as dirty until they are flushed to disk.
 
-            \paragraph*{File read and write}
-            Our model supports chunk-by-chunk file read and write simulation
-            with a user-defined chunk size, in which chunks are read/written
-            until file is entirely read/written.
-            The simulated LRU lists are used as follows. 
-            When a file is read, its data can be partially cached, completely cached, 
-            or uncached. 
-            In the scope of this study, our model support file accesses in chunk 
-            but not file random access.
-            Thus, we simplify the data read order, in which uncached data is read 
-            from disk first, and cached data is only read after all uncached data is read 
-            (left to right in Fig.~\ref{fig:read_order}).
-            When a chunk of uncached data is read, a new clean block is created 
+    \subsubsection{Reads and Writes}
+
+    Our simulation model supports chunk-by-chunk file accesses
+    with a user-defined chunk size. However, for simplicity, we assume that file pages are 
+    accessed in a round-robin fashion rather than fully randomly. 
+    Therefore, when a file is read, cached data is read only after all uncached data was read, and data from the inactive list is read
+    before data from the active list
+    (left to right in Figure~\ref{fig:read_order}).
+    When a chunk of \emph{uncached} data is read, a new clean block is created 
     and put at the top of the inactive list. 
-            For a chunk of cached data, one or more existing data blocks in two 
-            page cache LRU lists are accessed again, and moved to the active list. 
-            If the blocks are clean, we only need to keep the information about the 
-            amount of data as they are available for eviction. 
-            Thus, these re-accessed clean blocks can be merge into a bigger block 
-            to reduce the number of block objects and put it in the active list. 
-            If the blocks are dirty, since the blocks need to keep the information 
-            of entry time for periodical flushing, they are only moved to the active list. 
-            Because the chunk size and block size can be different, there are circumstances 
-            in which a block is not entirely read. 
-            In this case, the block is split into two smaller blocks, one is re-accessed 
-            and the other remains the same.
-
+    When a chunk of \emph{cached} data is read, one or more existing data blocks in the LRU lists are accessed.
+    If these blocks are clean, we merge them together, update the access time and size of the resulting block, 
+    and put it in the active list. 
+    If the blocks are dirty, we move them to the active list to preserve their entry time. 
+    Because the chunk size and block size can be different, there are situations
+    where a block is not entirely read. 
+    In this case, the block is split in two smaller blocks and one of them is re-accessed.
+    % From Tristan: not sure where to put this nor if it's necessary:
+    % , in which chunks are read/written
+    % until file is entirely read/written.
     \begin{figure}
@@
            \includegraphics[width=\columnwidth]{figures/read_order.pdf}
-                   \caption{File data read order from left to right, in which uncached data 
-                   is read first, then cached data from inactive list and active list is read later.}    
+           \caption{File data read order. Data is read from left to right: uncached data 
+           is read first, followed by data from the inactive list, and finally data from the active list.}
            \label{fig:read_order}
@@
 
-            For writes, we assume that all data to be written is 
-            uncached and it is written to cache as dirty data. 
-            Thus, each time a chunk is written, we create a block of dirty data 
-            and put it in the inactive list.
+    For file writes, we assume that all data to be written is 
+    uncached. Thus, each time a chunk is written, we create a block of dirty data 
+    and add it to the inactive list.
 
-            As is done in the kernel, we ensure the size of the active list does not 
-            exceed twice the size of the inactive list 
-            \cite{gorman2004understanding, linuxdev3rd2010}. 
-            This is done by moving least recently used blocks in the active list 
-            to the inactive list when the active list grows. If only a part of a data 
-            block needs to be moved, the block is split.
+    \subsubsection{Flushing and eviction}
 
     \paragraph*{Flushing}
-            Next, we simulate data flushing, which runs in the main simulated 
-            thread and writes back dirty data to disk. 
-            The data flushing simulation function traverses the sorted 
-            inactive list first, and then the active list, looks for least recently used 
-            dirty blocks, writes them to disk through the storage device model, 
-            and sets dirty flags of these blocks to 0 until the total flushed amount 
-            reaches the amount to be flushed or there is no dirty data left in cache. 
-            Similar to file read and write, the size of flushed data and block size 
-            can be different, there are cases that the last block is not entirely flushed.  
-            In that case, the last block is split into two blocks, one that is flushed 
-            and one that remains dirty.
-            The flushing time is simulated with storage model.
+    The main simulated thread in the Memory Manager can flush the
+    cache to write dirty data to disk. The data flushing simulation
+    function takes the amount of data to flush as parameter. While
+    the amount of data to flush is not reached and there are dirty
+    blocks remaining in cache, this function traverses the sorted
+    inactive list, then the sorted active list, and writes the
+    least recently used dirty block to disk having set its dirty
+    flag to 0. In case the amount of data to flush requires that a
+    block be partially flushed, the block is split in two blocks,
+    one that is flushed and one that remains dirty. Flushing time
+    to disk is simulated with the storage model.
 
     \paragraph*{Cache eviction}
-            Similarly to data flushing, our cache eviction simulation runs on 
-            the main thread, frees up the page cache by traversing and deleting 
+    Similarly to data flushing, our cache eviction simulation runs in
+    the main thread. It frees up the page cache by traversing and deleting 
     least recently used clean data blocks in the inactive list.
-            The amount of evicted data is passed in and data blocks are deleted 
-            from the inactive list until total evicted data reaches the amount 
-            passed in, or there is no clean block left in the list.
-            If the last evicted block is not entirely evicted, the block is split, 
-            and only one block is deleted.
-            The cache eviction simulation does not add up to the simulated time 
-            since cache eviction time is negligible in real systems.
+    The amount of data to evict is passed as a parameter and data blocks are deleted 
+    from the inactive list until total evicted data reaches the required amount,
+    or until there is no clean block left in the list.
+    If the last evicted block does not have to be entirely evicted, the block is split in two blocks,
+    and only one of them is evicted.
+    The cache eviction simulation does not contribute to the simulated time 
+    since cache eviction time is negligible in real systems \tristan{a reference would be welcome}.
 
@@
                 \State flushing\_time = sm.write(blocks)
-                        \State flag blocks as clean data
+                \State flag blocks as clean data \tristan{a bit informal}
                 \If{flushing\_time $<$ t}
                     \State sleep(t - flushing\_time)
-                        \EndIf
+                \EndIf \tristan{else?}
             \EndWhile
@@
     \paragraph*{Periodical flushing}
-            The periodical flushing simulation is run on a simulated background thread 
-            to flush expired dirty data blocks from cache to disk.
-            Similar to expired dirty pages in real Linux systems, a dirty block in our model 
-            is considered expired if: 
-            \begin{enumerate}
-                \item It is dirty and 
-                \item The time from its entry time to current time is longer than the 
+    Periodical flushing is simulated in the Memory Manager
+    background thread. As in the Linux kernel, a dirty block
+    in our model is considered expired if (1) it is dirty, and (2)
+    the duration since its entry time is longer than a
     predefined expiration time. 
-            \end{enumerate}
-            The simulation algorithm of periodical flushing is an infinite loop in which 
-            Memory Manager searches and flushes these expired dirty blocks to disk, 
-            as described in Algorithm~\ref{alg:pdflush}. 
-            The loop ends when the host is off (line 7). 
-            In each repetition, Memory Manager finds expired dirty blocks in two 
-            page cache LRU lists (line 8), simulates writes of data of these blocks 
-            to disk (line 9), and mark them as clean (line 10).
-            If the flushing time does not exceed our time interval, the thread is put 
-            to sleep for the remaining time (lines 11-13). 
-            Then based on the state of the host at current simulated time, 
-            the algorithm continues or finishes the loop.
-            Because periodical flushing is simulated as a background thread, it can be 
-            concurrent with disk I/O on main thread. This parallelism is handled by the 
+    Periodical flushing is simulated as an infinite loop in which 
+    the Memory Manager searches and flushes these expired dirty blocks to disk (Algorithm~\ref{alg:pdflush}). 
+    % From Tristan: the algorithm is quite straightforward, I don't think this is 
+    % necessary
+    % In each repetition, Memory Manager finds expired dirty blocks in two 
+    % page cache LRU lists (line 8), simulates writes of data of these blocks 
+    % to disk (line 9), and mark them as clean (line 10).
+    % If the flushing time does not exceed our time interval, the thread is put 
+    % to sleep for the remaining time (lines 11-13). 
+    % Then based on the state of the host at current simulated time, 
+    % the algorithm continues or finishes the loop.
+    Because periodical flushing is simulated as a background thread, it can happen concurrently
+    with disk I/O initiated by the main thread. This parallelism is handled by the 
     storage model and reflected in simulated I/O time.
 
-            \subsubsection{File I/O simulation model}
-
-            As mentioned, in our model, file reads and writes can be completed by 
-            reading or writing chunk-by-chunk until entire file is read or written.
+    \subsection{I/O Controller}
 
@@
     \end{algorithm}            
-
+    As mentioned previously, our model reads and writes file chunks in a round-robin 
+    fashion.
     When a simulated application reads a file chunk, it sends a chunk read request 
@@
     and update the corresponding data blocks in cache (line 16).
-
     \begin{algorithm}\caption{File chunk write simulation of IOController}
@@
     \end{algorithm}
-
     Algorithm~\ref{alg:write} describes our simulation of chunk write in 

### 2020-12-07 paper_io_sim.tex

@@
            devices (memory and disk) simulate file transfers with
-           bandwidth sharing. \tristan{Could you annotate arrows to
-           make them more explicit?}.}
+           bandwidth sharing.}
            \label{fig:interaction}

### 2020-12-07 paper_io_sim.tex
edits in methods

@@
 \usepackage{tabularx}
+
+% Fix link colors
+\hypersetup{
+    colorlinks = true,
+    linkcolor=red,
+    citecolor=red,
+    urlcolor=blue,
+    linktocpage % so that page numbers are clickable in toc
+}
 
@@
 
-    \paragraph*{Flushing}
     The main simulated thread in the Memory Manager can flush the
@@
 
-    \paragraph*{Cache eviction}
     Similarly to data flushing, our cache eviction simulation runs in
@@
 
-    \paragraph*{Periodical flushing}
     Periodical flushing is simulated in the Memory Manager
@@
                 \Desc{fn}{file name}
-                \Desc{fs}{file size (assume to fit in memory)}
+                \Desc{fs}{file size (assumed to fit in memory)}
                 \Desc{mm}{MemoryManager object}
@@
     \end{algorithm}
-    As mentioned previously, our model reads and writes file chunks in a round-robin 
-    fashion.
-    When a simulated application reads a file chunk, it sends a chunk read request 
-    to IOController.             
-    The algorithm to simulate a chunk read implemented in IOController 
-    is described in Algorithm~\ref{alg:read}.   
-    First, we calculate the amount of uncached data that needs to be read 
-    from disk, and the remaining amount is read from cache (line 7-8).
-    We assume that after a file is read, there are two copies of the file in memory 
-    (one in cache, one in anonymous memory). 
-    Unfortunately, SimGrid lacks of application memory management and WRENCH 
-    only use the memory capacity to check if the applications can be run. 
-    Thus, in our model, when a chunk is read, one copy of the chunk 
-    occupies anonymous memory, and the uncached part of that chunk is 
-    read into cache.    
-    The amount of required memory to read the chunk is calculated and 
-    if there is not enough memory available, Memory Manager is called to 
-    flush dirty data to disk and simulate flushing time with memory model (line 9). 
-    This flushing is complemented by eviction in the Memory Manager (line 10). 
-    Then, if there is uncached data, a disk read is simulated with disk model (line 12), 
-    and this data of the file is added to cache (line 13).
-    Finally, if cached data is read, Memory Manager is called to simulate a cache read  
-    and update the corresponding data blocks in cache (line 16).
+    As mentioned previously, our model reads and writes file chunks in a
+    round-robin fashion. To read a file chunk, simulated applications send
+    chunk read requests to the IOController which processes them using
+    Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
+    data that needs to be read from disk, and the remaining amount is read
+    from cache (line 7-8). The amount of memory required to read the chunk
+    is calculated, corresponding to a copy of the chunk in anonymous memory
+    and a copy of the chunk in cache \tristan{Create a variable
+    "required\_mem" in the algorithm to clarify}.
+    \tristan{add a call to mm in the algorithm to increment anonymous memory. Mention in the experiments that anonymous memory is flushed after each task.}
+    If there is not enough available memory, the Memory Manager is called
+    to flush dirty data (line 9). If necessary, flushing is complemented by
+    eviction (line 10). When called with negative arguments, functions
+    flush and evict simply return and do not do anything. Then, if there is
+    uncached data, a disk read is simulated with disk model (line 12)
+    \tristan{this is not consistent with the new Figure 1 where disk reads
+    are done by the mm}, and this data of the file is added to cache (line
+    13).
+    Finally, if cached data needs to be read, the Memory Manager is called
+    to simulate a cache read  
+    and update the corresponding data blocks accordingly (line 16).
+
     \begin{algorithm}\caption{File chunk write simulation of IOController}
@@
                 \State mm.evict(cs - mem\_amt  - mm.free\_mem) 
-                \State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem))
-                \State sm.write(cs - mem\_amt)
+                \State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem)) \tristan{why is data written to cache with "addtocache" here and with "write" at line 10?}
+                \State sm.write(cs - mem\_amt) \tristan{According to Fig 1 the IOController is not supposed to interact directly with sm}
             \EndIf
@@
     \end{algorithm}
-    Algorithm~\ref{alg:write} describes our simulation of chunk write in 
+    Algorithm~\ref{alg:write} describes our simulation of chunk writes in 
     the IOController. 
-    In the chunk write, data can be written to cache with memory bandwidth 
-    without waiting for flushing or being throttled before the amount of 
-    dirty data reaches dirty\_ratio, or otherwise written with disk bandwidth.
-    Thus, our algorithm initially checks the remaining amount of data that 
-    can be written as dirty data (line 6).
-    If this amount is greater than 0, MemoryManager is asked to evict 
-    data from cache to accommodate the written data (line 8). 
-    If the amount to evict calculated and passed in the evict function is less than 0, 
-    the eviction does nothing.
-    After evicting cache, the actual amount of data that can be written to 
-    page cache is calculated (line 9), a cache write is then simulated 
-    with this amount through memory model and data written is added to 
-    page cache LRU lists (line 10).
-    After dirty\_ratio is reached, dirty data has to be flushed before new data 
-    can be written to cache again. 
-    If there is remaining chunk data that has not been written, we flush and 
-    evict data from cache as much as possible so that the this data can be 
+    % Data is written to cache until the amount of dirty data in the cache 
+    % exceeds the dirty ratio times the amount of available memory. Data is then written to disk.
+    Our algorithm initially checks the  amount of data that 
+    can be written as dirty data given the dirty ratio (line 6).
+    If this amount is greater than 0, the Memory Manager is requested to evict 
+    data from cache if necessary (line 8) \tristan{the difference between available and free memory should be explained somewhere}.
+    After eviction, the amount of data that can be written to 
+    page cache is calculated (line 9), and a cache write is simulated (line 10).
+    To write the remaining data, we flush and 
+    evict data as possible so that this data can be 
     written to cache (line 13-14). 
-    Here, MemoryManager does not guarantee that the amount of flushed 
-    and evicted data is enough for the write of remaining data. 
+    The Memory Manager does not guarantee that the amount of flushed 
+    and evicted data is enough to write the remaining data. 
     In the case it is less than the required amount, the remaining data 
@@
     after the write is limited by the amount free memory, and this amount 
-    is added to cache (line 15). 
+    is added to cache (line 15) \tristan{the previous two sentences are very unclear, I don't understand them so I can't reformulate.}. 
     Finally, the disk write is simulated with the remaining data amount (line 16), 
-    and the chunk write algorithm finishes.
+    and the chunk write algorithm finishes \tristan{see comment in algorithm}.
 
@@
             Since bandwidth sharing was not simulated, this prototype does not support
-            concurrency: it is limited to single-threaded pipelines running on systems
-            with a single-core CPU. We used it in a first validation of our simulation
-            model, against a real sequential pipeline running on a real system.
-            The source code of our Python prototype is available at
-            \url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}
+            concurrency: it is limited to single-threaded applications running on systems
+            with a single-core CPU. We used this prototype for a first validation of our simulation
+            model against a real sequential application running on a real system.
+            The source code this prototype is available at
+            \url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}. We used Python 3.7.
 
-            Having the model validated, we created simulators for different use cases
-            using WRENCH.
-            We implemented our model separately from existing WRENCH components,
-            exposed an API for users, and added an command line argument to enable
-            the model in WRENCH.
-            This helps us to avoid unexpected impacts on WRENCH and
-            independently update the model in the future.
-            We then compared the results of the WRENCH, WRENCH-cache simulators
-            the results of real execution on a real system.
+            We also implemented our model in WRENCH, separately from existing WRENCH components. 
+            Page cache simulation is available through a user API and can be activated using a command-line argument.
+            We used SimGrid 3.25 and WRENCH 1.6. SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid},
+            and WRENCH is at \url{https://github.com/wrench-project/wrench/tree/master/src/wrench/services/memory}.
+            \tristan{which commit did you use?}
 
-            In this work, we use Python 3.7 to implement the single-threaded simulator,
-            SimGrid 3.25 and WRENCH 1.6 for more multi-threaded simulators.
-            SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid},
-            and WRENCH is at
-            \url{https://github.com/wrench-project/wrench/tree/master/src/wrench/services/memory}.
+\tristan{anything else interesting to mention about the implementation?}
 
@@
 
-            The goal of our experiments is to evaluate our page cache
-            simulation in single-threaded and multi-threaded applications
-            accessing data on local and network file systems.
-            We use a pipeline of sequential tasks where each task reads an input file,
-            increments every byte of the input to emulate real processing, and writes
-            the result to disk.
-            The output of the previous task is the input of the next task.
-            We also use a real application pipeline \textcolor{red}{[Maybe briefly describe this application's purpose (e.g. a standard fMRI preprocessing piepeline)]},
-            to evaluate the applicability of our simulation model.
+        Our experiments compared real executions with the original WRENCH,
+        our WRENCH-cache extension, and our prototype Python simulator. We
+        evaluated our page cache simulation in single-threaded and
+        multi-threaded applications accessing data on local and network
+        file systems.
+
+            We used a pipeline of three sequential tasks where each task reads the output of the previous task,
+            increments every byte of the input to emulate real
+            processing, and writes the resulting data to disk. The first file was provided as input. We also used
+            a real application pipeline \textcolor{red}{[Maybe briefly
+            describe this application's purpose (e.g. a standard fMRI
+            preprocessing piepeline)]}, to evaluate the applicability of
+            our simulation model.
 

### 2020-12-07 paper_io_sim.tex
Comments in III.D

@@
 
-        Our experiments compared real executions with the original WRENCH,
-        our WRENCH-cache extension, and our prototype Python simulator. We
-        evaluated our page cache simulation in single-threaded and
-        multi-threaded applications accessing data on local and network
+        Our experiments compare real executions with the original WRENCH,
+        with our WRENCH-cache extension, and with our prototype Python simulator. We
+        evaluate our page cache simulation in single-threaded and
+        multi-threaded applications, accessing data on local and remote
         file systems.
 
-            We used a pipeline of three sequential tasks where each task reads the output of the previous task,
-            increments every byte of the input to emulate real
-            processing, and writes the resulting data to disk. The first file was provided as input. We also used
-            a real application pipeline \textcolor{red}{[Maybe briefly
-            describe this application's purpose (e.g. a standard fMRI
-            preprocessing piepeline)]}, to evaluate the applicability of
-            our simulation model.
+        We simulate an application with three single-core, sequential tasks
+        where each task reads the file produced by the previous task,
+        increments every byte of this file to emulate real processing, and
+        writes the resulting data to disk. We also use a real application
+        \textcolor{red}{[Maybe briefly describe this application's purpose
+        (e.g. a standard fMRI preprocessing piepeline)]}, to evaluate the
+        applicability of our simulation model.
 
-            We use a dedicated cluster hosted at Concordia University to conduct
-            the experiments. The cluster consists of one login node, 9 compute nodes
-            and 4 storage nodes connected with a network bandwidth of 25 Gbps.
-            Each compute node has two 16 cores Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz,
-            275~GB (256~GiB) of RAM and 6 SSDs, 450~GB each with XFS file system,
-            378~GB of tmpfs and 126~GB of devtmpfs file system.
-            The cluster is run on CentOS 8.1 with NFS version 4.
-            We use \textbf{\textit{atop}} and \textbf{\textit{collectl}} as tools to
-            monitor and collect memory status and disk throughput
-            on the cluster.
+        The real application runs on a dedicated cluster hosted at
+        Concordia University, with one login node, 9 compute nodes, and 4
+        storage nodes connected with a network bandwidth of 25 Gbps. Each
+        compute node has 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
+        @ 2.10GHz, 275~GB (256~GiB) \tristan{are you sure it's not 256GB?
+        pls check.} of RAM, 6 $\times$ SSDs of 450~GB each with the XFS
+        file system, 378~GB of tmpfs, and 126~GB of devtmpfs file system.
+        Nodes run CentOS~8.1 with NFS version 4. We use the \texttt{atop}
+        and \texttt{collectl} tools to monitor and collect memory status
+        and disk throughput.
 
-            In simulated platform description, because the disk bandwidths in SimGrid
-            are symmetrical, we measure the read/write bandwidths on our cluster and
-            take the average bandwidths.
-            The benchmark results of bandwidths and the values used in the simulators
-            in our all experiments are shown in Table~\ref{table:benchmark}.
-
+        To parametrize our simulators, we benchmarked the bandwidths of
+        memory, local disk, remote disk (NFS), and network
+        (Table~\ref{table:benchmark}). Since SimGrid only supports
+        symmetrical bandwidths, we use the mean of the read and write
+        bandwidth values in our experiments.
             \begin{table}[htbp]
             \centering
-            \begin{tabularx}{\columnwidth}{|l
+            \begin{tabularx}{\columnwidth}{|c
             |>{\centering\arraybackslash}X
@@
             \hline
-                Bandwidths (MBps)  & Cluster (real) & Python simulator & WRENCH simulator\\
+                Device  & Cluster (real) & Python simulator & WRENCH simulator\\
             \hline
@@
             \end{tabularx}
-            \caption{Bandwidth benchmark results and simulator configurations.
-            The bandwidths used in simulators are average of read and write bandwidths
-            measured on the cluster.
-            Remote disk and network are not simulated in Python simulator.}
+            \caption{Bandwidth benchmarks (MBps) and simulator configurations.
+            The bandwidths used in the simulations are the average of the read and write bandwidths
+            measured in the cluster.
+            Remote disk and network accesses are not simulated in the Python simulator. \tristan{table could be prettified}}
             \label{table:benchmark}
@@
 
-            As the pipelines are single-core, we create simulated tasks,
-            each of them runs on only one core.
-            While in Python, computation time can be directly configured,
-            this time is injected to the WRENCH simulators with number of flops,
-            which is calculated with CPU speed and compute duration.
-            In our experiments, the CPU speed is set to 1 Gflops.
-            Finally, because WRENCH simulates applications base on network-communication,
-            we use an infinite bandwidth to eliminate network latency for local I/Os.
-            The details of simulated platform and the source code of the simulator
-            used in our experiments are available in this commit:
+            As our focus is on I/O rather than compute, we measured
+            application task durations on a cluster node and used these
+            durations in our simulation. For the Python simulator,
+            execution times were directly used in the simulation. For
+            WRENCH and WRENCH-cache, we determined the corresponding number
+            of flops on a 1~Gflops CPU and used these values in the
+            simulation (see Table~\ref{table:tasks} \tristan{add a table to describe execution times}). The final simulated
+            platform and application used in our experiment are avaiable at
             \url{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}.
 
-            \subsubsection{Single-threaded evaluation}
+            % Finally, because WRENCH simulates applications base on network-communication,
+            % we use an infinite bandwidth to eliminate network latency for local I/Os.
 
-            The first experiment is single-threaded, where memory bandwidth
-            and disk bandwidth are dedicated.
-            Our goal is to compare the detailed simulation results of each task
-            in the pipeline including read/write time, memory behavior and
-            then validate the correctness of the simulation model.
+            Our first experiment simply runs one instance of the
+            application on a single cluster node, with different input file
+            sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed
+            to the same local disk.
 
-            The pipeline used in this experiment consists of three tasks,
-            each task reads a file, increases every byte and writes output file
-            to local disk.
-            We use the input sizes of 20~GB, 50~GB, 75~GB, 100~GB and run
-            the pipeline on the cluster.
-            The pipeline is simulated with the Python prototype simulator,
-            original WRENCH and WRENCH-cache simulator.
+            Our second experiment runs concurrent instances of the
+            application on a single node, all application instances
+            operating on different files of size 3~GB stored in the same
+            local disk, sharing the disk bandwidth. We vary the number of
+            concurrent application instances from 1 to 32 since cluster
+            nodes have 32 CPU cores.
 
-            \subsubsection{Multi-threaded evaluation}
+            Our third experiment occurs in the same configuration as the
+            previous one, albeit on a 400-GB NFS-mounted remote disk
+            mounted from another compute node. \tristan{As is commonly
+            configured in HPC environments to avoid data loss,?} NFS client
+            and server read caches are enabled, but server writes are configured with
+            write-through cache instead of write back \tristan{how about client write
+            cache?}. Therefore, all the writes happen at disk bandwidth, but
+            reads can benefit from cache hits. We simulate write-through by
+            writing data directly to disk and adding them to the server
+            cache \tristan{it it part of the IOController? it would deserve
+            a few sentences in the methods.}.
 
-            In the second scenario, we evaluate the simulator in a concurrent I/O, 
-            bandwidth-sharing scenario with a multi-threaded application.             
-            We vary the number of concurrent pipelines running on a multi-core host.
-
-            All the task in the pipelines read input files from and write output files 
-            to a local disk. 
-            As a compute node of the cluster has 32 cores CPU and 256~GB of RAM,  
-            we create 32 input files with the size of 3~GB each and vary the number of 
-            concurrent pipelines from 1 to 32.
-
-            Because implementation of a multi-threaded simulator in Python is 
-            expensive, and we have our model implemented in WRENCH, 
-            we only compare the results of simulators with original WRENCH, 
-            WRENCH-cache and real pipelines in this experiment. 
-            The results are compared in total makespan of the pipelines, 
-            cumulative read time and cumulative write time.
-
-            \subsubsection{Evaluation on NFS}
-
-            Since network file systems are very common in HPC, we conduct 
-            the next experiment with the same scenario as in the multi-threaded experiment, 
-            except that files are stored on a node, read and written by another node. 
-            The compute node is the node used in the multi-threaded experiment, 
-            and the capacity of the disk used on the storage node is 400 GB.
-
-            In NFS, client cache and server cache are enabled in reading, but only 
-            server write-through cache is involved in writing. 
-            Although write-through does not have any impacts writing performance, 
-            caching data on server side allows server to avoid disk reads, which are 
-            significantly longer than cache reads.
-            Thus, we simulate page cache for reading on both client and server node.
-            In writing, write-though is simulated with data being written directly to 
-            server storage and then added to server cache.
-
-            \subsubsection{Simulation of a real application}
-
-            Finally, we consider a real pipeline from the neuroimaging domain.             
-            \textcolor{red}{[Description of the real pipeline with nighres 
-            (including the workflow engine if it's Dask)]}.
-
-            We simulate this pipeline with simulators using original WRENCH and 
-            WRENCH-cache. 
-            The simulation results are then compared with the real results. 
-            Because our work focuses on I/O time, we assume that CPU time is 
-            correctly modeled and use the CPU time measured in the real pipelines 
-            to setup our simulated workflow. The results are compared in terms of 
-            read/write time and total makespan.
+            Finally, we simulate a real application from the neuroimaging
+            domain. \textcolor{red}{[Description of the real pipeline with
+            nighres (including the workflow engine if it's Dask)]}.
+            We simulate this pipeline with simulators using original WRENCH
+            and WRENCH-cache. Because our work focuses on I/O time, we
+            assume that CPU time is correctly modeled and use the CPU time
+            measured in the real pipelines to setup our simulated
+            application.
 

### 2020-12-07 paper_io_sim.tex
minor edit in abstract

@@
 \begin{document}
-\title{I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework}
+
+\title{Simulating the Linux page cache in distributed applications with the WRENCH library}
 
@@
         One widely-used approach to minimize the costs of data transfers during
-        regular use is through the employment of the page cache.
+        regular use is through the employment of the Linux page cache.
         While many simulation tools have been developed to 

### 2020-12-07 paper_io_sim.tex
[WIP] intro fixes

--- 2020-12-07 paper_io_sim.tex
@@
 \usepackage{ulem}
-\usepackage{tabularx}
-
-% Fix link colors
-\hypersetup{
-    colorlinks = true,
-    linkcolor=red,
-    citecolor=red,
-    urlcolor=blue,
-    linktocpage % so that page numbers are clickable in toc
-}
 
@@
 		The rapid emergence of Big Data seen in recent years has resulted in a 
-        growing needs for efficient data storing and optimized data analysis. 
+		growing need for efficient data storing and optimized data analysis techniques. 
 		Despite relying on large computing infrastructures with powerful computing capability 
@@
 		has not been fully captured or remains absent from these simulators. 
-        The lack of ability to simulate I/O with page cache leads to inaccurate 
-        simulation results in many situations, especially when dealing with data intensive applications.
+		The inability to simulate I/O with page cache leads to inaccurate 
+		simulations in many situations, especially when dealing with data intensive applications.
 
@@
 		of the page cache, which can also be integrated to other simulators 
-        to improve I/O simulation accuracy, with its practical results.
+		to improve I/O simulation accuracy.
 
@@
 
-        In recent years, as a result of the growing demand for Big Data processing 
-        on large computing infrastructures, there have been the research and development 
-        of simulation applications and frameworks for distributed systems. 
-        These simulation tools have shown their advantages in evaluation 
-        of algorithms, applications and solutions. 
-        Experimenting on real platforms exposes shortcomings due to limited and 
-        shared resources, operational costs, unstable environments and the result 
-        reproducibility difficulties.
-        On the contrary, simulators provide to researchers the freedom of choosing 
-        simulated infrastructures, reproducible results and low costs.
-        These simulators are developed and integrated with simulation models 
-        of computing resources including CPU, network and storage.
-        The ultimate goal that all simulators try to achieve is to accurately simulate 
-        applications. 
-        However, there is a trade-off between accuracy and scalability, which has been 
-        indicated in some reviews \cite{casanova2014simgrid} \cite{byrne2017review}. 
-        This means that in most ciscumstances, users have to, more or less, sacrifice 
-        accuracy for the scalability and performance when their platforms grow to 
-        hundreds or thousands nodes.
+        The Linux Page Cache, introduced in kernel version 2.4, plays an important role in
+        the reduction of filesystem data transfer costs~\cite{pagecache}. With the page cache, previously
+        read data can be reread directly from memory and written data can be written to
+        memory before being asynchronously flushed to disk, resulting in improved I/O performance
+        on slower storage devices. The performance benefits that can be obtained from using the page
+        cache are determined by many factors including the total amount of memory,
+        the amount of data being written (i.e. dirty data), and the percentage of available memory available for
+        written data. All these factors are important when determining the impact of data transfers on
+        application performance, particularly in data intensive applications.
 
-        Beside computing resources such as CPU, network, storage, another 
-        important factor that has significant impacts on I/O performance is 
-        page cache. 
-        Page cache is an architectural solution employed in systems to tackle 
-        the problem of disk I/O bottleneck in data processing. 
-        Although this is not a radical solution for disk performance, it contributes 
-        to minimizing disk accesses and maximizing I/O performance. 
-        Surprisingly, page cache has not been fully modeled, if not completely absent, 
-        in simulation tools. 
-        Some simulators try to simulate I/O with page cache, but fail to capture the 
-        key features such as dirty data and cache eviction mechanism 
-        \cite{nunez2012simcan} \cite{nunez2012icancloud}.
-        Again, these simulators also have a problem in scalability with 
-        long simulation time regarding to their very detailed level of simulation. 
-        Other works are able to capture important features of page cache, 
-        but they focus on very specific domains, such as energy consumption 
-        in \cite{xu2018saving}.
+        Characterizing the performance of applications directly on computing infrastructure
+        has various shortcomings, which include, limited and shared resources, operational
+        costs, unstable environments, and the difficulties in producing reproducible results.
+        Simulations address these concerns by providing abstractions to the performance of computer
+        hardware, such as CPU, network and storage. As a result, simulations provide and
+        cost-effective, fast, easy and reproducible  way evaluate application performance
+        on different infrastructures without having to execute the code.
 
-        In this work, we propose a simulation model of I/O with page cache, 
-        implement it with SimGrid, a simulation framework for distributed computing, 
-        and WRENCH, an open-source library for developing workflow simulator 
-        built on top of SimGrid. 
-        One of the reasons for us to choose SimGrid is that it has a balance 
-        between accuracy and scalability. Moreover, the versatility of SimGrid 
-        enables it to be used in cross-domain \cite{casanova2014simgrid}. 
-        In addition, SimGrid has been actively maintained with updated documents. 
-        Then we evaluate our model with four simulators in four experiment scenarios. 
-        This paper is organized as follows. 
-        In Section \ref{relatedwork} we provide details of page cache 
-        existing simulation tools.
-        In Section \ref{relatedwork} we propose our I/O with page 
-        cache simulation model and how we implement it and our experiments 
-        to evaluate the model.  
-        Section \ref{results} compares the results of our model with the results 
-        of original WRENCH and real execution results. 
-        Finally, Section \ref{discussion} concludes and presents future 
-        work.
+        Page cache simulation is crucial when simulating data intensive applications,
+        which would rely on cache hits and writing to memory for additional performance.
+        While existing simulators can adequately characterize many different computing environments,
+        they lack much of the ability to adequately describe the page cache. Some
+        simulators try to simulator I/O with page cache, but fail to capture the key features such
+        as dirty data and cache eviction policies~\cite{nunez2012simcan,nunez2012icancloud}. 
+        Other simulators, such as~\cite{xu2018}, capture the important feature of the
+        page cache, but are domain-specific.
+
+		%However, there is a trade-off between accuracy and scalability, which has been 
+		%indicated in some reviews \cite{casanova2014simgrid} \cite{byrne2017review}. 
+		%This means that in most cicumstances, users have to, more or less, sacrifice 
+		%accuracy for the scalability and performance when their platforms grow to 
+		%hundreds or thousands nodes.
+
+        In this paper, we present WRENCH-Cache, a simulation model of I/O with page
+        cache incorporated into WRENCH, a workflow management system simulator based of
+        the popular SimGrid framework. The goal of WRENCH-Cache is to provide users
+        with an accurate representation of a workflow execution, when file system page
+        cache is used. WRENCH-Cache is able to simulate both writeback and writethrough
+        cache which would normally be available to local and NFS-based filesystems.
 
@@
 	\label{method}
-    In this section, we present our page cache simulation model for file I/O,
-    flusing, and eviction. We detail the design and implementation of two simulators:
-    (1) a standalone Python prototype to evaluate the simulation model independently of any other 
-    tool, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
-    We also describe experimental scenarios to evaluate our simulators in various configurations.
+
+		In this section, we present our approach to model file read and write with 
+		page cache, cache eviction and data flushing mechanisms implemented 
+		in the Linux kernel. We also detail the design and implementation of two simulators:
+		(1) a standalone Python prototype used to evaluate the model independently of any other 
+		library, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
+		Finally, we describe experimental scenarios to evaluate our simulations in various configurations.
+
+		\subsection{Principle of the simulator}
 
@@
 			Controller and the Memory Manager, which together simulate 
-    file reads and writes in Linux (Figure~\ref{fig:interaction}). 
-    To read or write a file chunk, a simulated application sends a
-    request to the IOController. The IOController interacts as needed with
-    the Memory Manager to free memory through flushing or eviction,
-    and to read or write to cache. The Memory Manager
-    implements these operations, simulates periodical flushing
-    and eviction from cache, and reads or writes to disk when necessary.
+			file read and write mechanisms in Linux (Fig.~\ref{fig:interaction}). 
+			To read or write a file chunk, a simulated application sends a request 
+			to the IOController. The IOController determines the amount of memory required by the operation,
+			interacts with the Memory Manager accordingly, reads or write to cache through the Memory Manager, 
+			and finally reads or writes on disk if necessary.
+			The Memory Manager simulates flushing, eviction, and I/O to cache.
+
 			\begin{figure}
@@
 				   devices (memory and disk) simulate file transfers with
-           bandwidth sharing.}
+				   bandwidth sharing. \tristan{Could you annotate arrows to
+				   make them more explicit?}.}
 				   \label{fig:interaction}
@@
 
-    \subsection{Memory Manager}
+			\subsubsection{Memory simulation model}
 
-    The Memory Manager simulates two parallel threads: the main one
-    implements flushing, eviction, and cache I/Os synchronously, whereas
-    the second one, operating in the background, periodically searches for
-    expired dirty data in page cache LRU lists and flushes them to disk. We
-    use existing storage simulation models~\cite{lebre2015} to simulate
-    memory, characterized by its storage capacity, read and write
-    bandwidths, and latency. Reusing existing storage models allows us to
-    simulate bandwidth sharing between concurrent memory accesses.
+			We leverage existing storage simulation models~\cite{lebre2015} to simulate the memory,
+			characterized by its storage capacity, read and write bandwidths,
+			and device latency.
+			Leveraging existing storage models allows us 
+			to take advantage of bandwidth sharing between concurrent I/Os to 
+			simulate parallel memory accesses.
 
-    \subsubsection{Page cache LRU lists}
+			\paragraph*{Memory Manager}
+			The Memory Manager simulates  two parallel threads: the first and main one simulates 
+			flushing, eviction, and cache I/Os synchronously; the second one is a background thread that  
+			periodically searches for expired dirty data in page cache LRU lists 
+			and flushes them to disk.
 
-    In the Linux kernel, page cache LRU lists contain file pages. However, 
+			\paragraph*{Page cache LRU lists}
+			In the Linux kernel, the page cache LRU lists contain file pages. However, 
 			due to the large number of file pages, managing lists of pages 
 			induces substantial overheads.
-    Therefore, we introduce the concept of a data block as a unit to represent data 
+			Thus, we introduce the concept of a data block as a unit to represent data 
 			cached in memory. A data block is a subset of file pages stored in
@@
 			time, a dirty flag that represents whether the data is clean (0) 
-    or dirty (1), and an entry (creation) time.
+			or dirty (1), and the time it is written to cache as dirty data. 
 			Blocks can have different sizes and a given file can have multiple 
@@
 			arbitrary number of smaller blocks.
+
 			\begin{figure}
@@
 			two lists of data blocks, an active list and an inactive list, both ordered by 
-    last access time (earliest first, Figure~\ref{fig:lrulist}).
+			last access time (earliest first, Fig.~\ref{fig:lrulist}).
 			As in the kernel, our simulator limits the size of the active list to
 			twice the size of the inactive list, by moving least recently 
-    used data blocks from the active list to the inactive list~\cite{gorman2004understanding, linuxdev3rd2010}.
-    % Modeling page cache LRU lists as lists of data blocks reduces the
-    % overhead of the simulator, as data blocks are obviously less
-    % numerous than file pages. At the same time, it achieves the accuracy in 
-    % I/O time simulation that can be seen in the later section.
-    % From Tristan: I don't think the sentences above are necessary,
-    % they are quite repetitive.
+            used data blocks from the active list to the inactive list.
+			Modeling page cache LRU lists as lists of data blocks reduces the
+			overhead of the simulator, as data blocks are obviously less
+			numerous than file pages, and at the same time achieves the accuracy in 
+			I/O time simulation that can be seen in the later section.
 
-    At any given time, a file can be partially cached, completely cached,
-    or uncached at all. Cached data blocks may reside in one or both of the
-    page cache LRU lists. The first time they are accessed, blocks are
-    added to the inactive list. On subsequent accesses, blocks of the
-    inactive list are moved to \tristan{if they are moved then they can't
-    be in both lists} the top of the active list. Cached blocks written to
-    cache are marked as dirty until they are flushed to disk.
+			At a specific time, an arbitrary file can be partially cached, completely cached, 
+			or uncached.
+			Cached data is composed of data blocks existing in one or both
+            of the page cache LRU lists. 
+            Initially, blocks accessed for the first time are added to the inactive list in 
+            page cache. 
+            If the blocks residing the inactive list are subsequently accessed, 
+            they will be moved to the top of the active list
+            Cached blocks written to cache are marked as dirty blocks until 
+            they are flushed to disk.
 
-    \subsubsection{Reads and Writes}
+			\paragraph*{File read and write}
+            Our model supports chunk-by-chunk file read and write simulation
+            with a user-defined chunk size, in which chunks are read/written
+            until file is entirely read/written.
+            The simulated LRU lists are used as follows. 
+            When a file is read, its data can be partially cached, completely cached, 
+            or uncached. 
+            In the scope of this study, our model support file accesses in chunk 
+            but not file random access.
+            Thus, we simplify the data read order, in which uncached data is read 
+            from disk first, and cached data is only read after all uncached data is read 
+            (left to right in Fig.~\ref{fig:read_order}).
+            When a chunk of uncached data is read, a new clean block is created 
+            and put at the top of the inactive list. 
+            For a chunk of cached data, one or more existing data blocks in two 
+            page cache LRU lists are accessed again, and moved to the active list. 
+			If the blocks are clean, we only need to keep the information about the 
+			amount of data as they are available for eviction. 
+			Thus, these re-accessed clean blocks can be merge into a bigger block 
+			to reduce the number of block objects and put it in the active list. 
+            If the blocks are dirty, since the blocks need to keep the information 
+            of entry time for periodical flushing, they are only moved to the active list. 
+			Because the chunk size and block size can be different, there are circumstances 
+			in which a block is not entirely read. 
+			In this case, the block is split into two smaller blocks, one is re-accessed 
+			and the other remains the same.
 
-    Our simulation model supports chunk-by-chunk file accesses
-    with a user-defined chunk size. However, for simplicity, we assume that file pages are 
-    accessed in a round-robin fashion rather than fully randomly. 
-    Therefore, when a file is read, cached data is read only after all uncached data was read, and data from the inactive list is read
-    before data from the active list
-    (left to right in Figure~\ref{fig:read_order}).
-    When a chunk of \emph{uncached} data is read, a new clean block is created 
-    and put at the top of the inactive list. 
-    When a chunk of \emph{cached} data is read, one or more existing data blocks in the LRU lists are accessed.
-    If these blocks are clean, we merge them together, update the access time and size of the resulting block, 
-    and put it in the active list. 
-    If the blocks are dirty, we move them to the active list to preserve their entry time. 
-    Because the chunk size and block size can be different, there are situations
-    where a block is not entirely read. 
-    In this case, the block is split in two smaller blocks and one of them is re-accessed.
-    % From Tristan: not sure where to put this nor if it's necessary:
-    % , in which chunks are read/written
-    % until file is entirely read/written.
 			\begin{figure}
@@
    				\includegraphics[width=\columnwidth]{figures/read_order.pdf}
-           \caption{File data read order. Data is read from left to right: uncached data 
-           is read first, followed by data from the inactive list, and finally data from the active list.}
+   				\caption{File data read order from left to right, in which uncached data 
+   				is read first, then cached data from inactive list and active list is read later.}	
    				\label{fig:read_order}
@@
 
-    For file writes, we assume that all data to be written is 
-    uncached. Thus, each time a chunk is written, we create a block of dirty data 
-    and add it to the inactive list.
+			For writes, we assume that all data to be written is 
+			uncached and it is written to cache as dirty data. 
+			Thus, each time a chunk is written, we create a block of dirty data 
+			and put it in the inactive list.
 
-    \subsubsection{Flushing and eviction}
+            As is done in the kernel, we ensure the size of the active list does not 
+            exceed twice the size of the inactive list 
+			\cite{gorman2004understanding, linuxdev3rd2010}. 
+			This is done by moving least recently used blocks in the active list 
+			to the inactive list when the active list grows. If only a part of a data 
+			block needs to be moved, the block is split.
 
-    The main simulated thread in the Memory Manager can flush the
-    cache to write dirty data to disk. The data flushing simulation
-    function takes the amount of data to flush as parameter. While
-    the amount of data to flush is not reached and there are dirty
-    blocks remaining in cache, this function traverses the sorted
-    inactive list, then the sorted active list, and writes the
-    least recently used dirty block to disk having set its dirty
-    flag to 0. In case the amount of data to flush requires that a
-    block be partially flushed, the block is split in two blocks,
-    one that is flushed and one that remains dirty. Flushing time
-    to disk is simulated with the storage model.
+			\paragraph*{Flushing}
+			Next, we simulate data flushing, which runs in the main simulated 
+			thread and writes back dirty data to disk. 
+			The data flushing simulation function traverses the sorted 
+			inactive list first, and then the active list, looks for least recently used 
+			dirty blocks, writes them to disk through the storage device model, 
+			and sets dirty flags of these blocks to 0 until the total flushed amount 
+			reaches the amount to be flushed or there is no dirty data left in cache. 
+			Similar to file read and write, the size of flushed data and block size 
+			can be different, there are cases that the last block is not entirely flushed.  
+			In that case, the last block is split into two blocks, one that is flushed 
+			and one that remains dirty.
+			The flushing time is simulated with storage model.
 
-    Similarly to data flushing, our cache eviction simulation runs in
-    the main thread. It frees up the page cache by traversing and deleting 
+			\paragraph*{Cache eviction}	
+			Similarly to data flushing, our cache eviction simulation runs on 
+			the main thread, frees up the page cache by traversing and deleting 
 			least recently used clean data blocks in the inactive list.
-    The amount of data to evict is passed as a parameter and data blocks are deleted 
-    from the inactive list until total evicted data reaches the required amount,
-    or until there is no clean block left in the list.
-    If the last evicted block does not have to be entirely evicted, the block is split in two blocks,
-    and only one of them is evicted.
-    The cache eviction simulation does not contribute to the simulated time 
-    since cache eviction time is negligible in real systems \tristan{a reference would be welcome}.
+			The amount of evicted data is passed in and data blocks are deleted 
+			from the inactive list until total evicted data reaches the amount 
+			passed in, or there is no clean block left in the list.
+			If the last evicted block is not entirely evicted, the block is split, 
+			and only one block is deleted.
+			The cache eviction simulation does not add up to the simulated time 
+			since cache eviction time is negligible in real systems.
 
@@
 						\State flushing\_time = sm.write(blocks)
-                \State flag blocks as clean data \tristan{a bit informal}
+						\State flag blocks as clean data
 						\If{flushing\_time $<$ t}
 							\State sleep(t - flushing\_time)
-                \EndIf \tristan{else?}
+						\EndIf
 					\EndWhile
@@
 
-    Periodical flushing is simulated in the Memory Manager
-    background thread. As in the Linux kernel, a dirty block
-    in our model is considered expired if (1) it is dirty, and (2)
-    the duration since its entry time is longer than a
+			\paragraph*{Periodical flushing}	
+			The periodical flushing simulation is run on a simulated background thread 
+			to flush expired dirty data blocks from cache to disk.
+			Similar to expired dirty pages in real Linux systems, a dirty block in our model 
+			is considered expired if: 
+			\begin{enumerate}
+				\item It is dirty and 
+				\item The time from its entry time to current time is longer than the 
 				predefined expiration time. 
-    Periodical flushing is simulated as an infinite loop in which 
-    the Memory Manager searches and flushes these expired dirty blocks to disk (Algorithm~\ref{alg:pdflush}). 
-    % From Tristan: the algorithm is quite straightforward, I don't think this is 
-    % necessary
-    % In each repetition, Memory Manager finds expired dirty blocks in two 
-    % page cache LRU lists (line 8), simulates writes of data of these blocks 
-    % to disk (line 9), and mark them as clean (line 10).
-    % If the flushing time does not exceed our time interval, the thread is put 
-    % to sleep for the remaining time (lines 11-13). 
-    % Then based on the state of the host at current simulated time, 
-    % the algorithm continues or finishes the loop.
-    Because periodical flushing is simulated as a background thread, it can happen concurrently
-    with disk I/O initiated by the main thread. This parallelism is handled by the 
+			\end{enumerate}
+			The simulation algorithm of periodical flushing is an infinite loop in which 
+			Memory Manager searches and flushes these expired dirty blocks to disk, 
+			as described in Algorithm~\ref{alg:pdflush}. 
+			The loop ends when the host is off (line 7). 
+			In each repetition, Memory Manager finds expired dirty blocks in two 
+			page cache LRU lists (line 8), simulates writes of data of these blocks 
+			to disk (line 9), and mark them as clean (line 10).
+			If the flushing time does not exceed our time interval, the thread is put 
+			to sleep for the remaining time (lines 11-13). 
+			Then based on the state of the host at current simulated time, 
+			the algorithm continues or finishes the loop.
+			Because periodical flushing is simulated as a background thread, it can be 
+			concurrent with disk I/O on main thread. This parallelism is handled by the 
 			storage model and reflected in simulated I/O time.
 
-    \subsection{I/O Controller}
+			\subsubsection{File I/O simulation model}
+
+			As mentioned, in our model, file reads and writes can be completed by 
+			reading or writing chunk-by-chunk until entire file is read or written.
 
@@
         				\Desc{fn}{file name}
-                \Desc{fs}{file size (assumed to fit in memory)}
+        				\Desc{fs}{file size (assume to fit in memory)}
 						\Desc{mm}{MemoryManager object}
@@
 			\end{algorithm}
-    As mentioned previously, our model reads and writes file chunks in a
-    round-robin fashion. To read a file chunk, simulated applications send
-    chunk read requests to the IOController which processes them using
-    Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
-    data that needs to be read from disk, and the remaining amount is read
-    from cache (line 7-8). The amount of memory required to read the chunk
-    is calculated, corresponding to a copy of the chunk in anonymous memory
-    and a copy of the chunk in cache \tristan{Create a variable
-    "required\_mem" in the algorithm to clarify}.
-    \tristan{add a call to mm in the algorithm to increment anonymous memory. Mention in the experiments that anonymous memory is flushed after each task.}
-    If there is not enough available memory, the Memory Manager is called
-    to flush dirty data (line 9). If necessary, flushing is complemented by
-    eviction (line 10). When called with negative arguments, functions
-    flush and evict simply return and do not do anything. Then, if there is
-    uncached data, a disk read is simulated with disk model (line 12)
-    \tristan{this is not consistent with the new Figure 1 where disk reads
-    are done by the mm}, and this data of the file is added to cache (line
-    13).
-    Finally, if cached data needs to be read, the Memory Manager is called
-    to simulate a cache read  
-    and update the corresponding data blocks accordingly (line 16).
+
+			When a simulated application reads a file chunk, it sends a chunk read request 
+			to IOController. 			
+			The algorithm to simulate a chunk read implemented in IOController 
+			is described in Algorithm~\ref{alg:read}.   
+			First, we calculate the amount of uncached data that needs to be read 
+			from disk, and the remaining amount is read from cache (line 7-8).
+			We assume that after a file is read, there are two copies of the file in memory 
+			(one in cache, one in anonymous memory). 
+			Unfortunately, SimGrid lacks of application memory management and WRENCH 
+			only use the memory capacity to check if the applications can be run. 
+			Thus, in our model, when a chunk is read, one copy of the chunk 
+			occupies anonymous memory, and the uncached part of that chunk is 
+			read into cache.	
+			The amount of required memory to read the chunk is calculated and 
+			if there is not enough memory available, Memory Manager is called to 
+			flush dirty data to disk and simulate flushing time with memory model (line 9). 
+			This flushing is complemented by eviction in the Memory Manager (line 10). 
+			Then, if there is uncached data, a disk read is simulated with disk model (line 12), 
+			and this data of the file is added to cache (line 13).
+			Finally, if cached data is read, Memory Manager is called to simulate a cache read  
+			and update the corresponding data blocks in cache (line 16).
 
@@
 						\State mm.evict(cs - mem\_amt  - mm.free\_mem) 
-                \State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem)) \tristan{why is data written to cache with "addtocache" here and with "write" at line 10?}
-                \State sm.write(cs - mem\_amt) \tristan{According to Fig 1 the IOController is not supposed to interact directly with sm}
+						\State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem))
+						\State sm.write(cs - mem\_amt)
 					\EndIf
@@
 			\end{algorithm}
-    Algorithm~\ref{alg:write} describes our simulation of chunk writes in 
+
+			Algorithm~\ref{alg:write} describes our simulation of chunk write in 
 			the IOController. 
-    % Data is written to cache until the amount of dirty data in the cache 
-    % exceeds the dirty ratio times the amount of available memory. Data is then written to disk.
-    Our algorithm initially checks the  amount of data that 
-    can be written as dirty data given the dirty ratio (line 6).
-    If this amount is greater than 0, the Memory Manager is requested to evict 
-    data from cache if necessary (line 8) \tristan{the difference between available and free memory should be explained somewhere}.
-    After eviction, the amount of data that can be written to 
-    page cache is calculated (line 9), and a cache write is simulated (line 10).
-    To write the remaining data, we flush and 
-    evict data as possible so that this data can be 
+			In the chunk write, data can be written to cache with memory bandwidth 
+			without waiting for flushing or being throttled before the amount of 
+			dirty data reaches dirty\_ratio, or otherwise written with disk bandwidth.
+			Thus, our algorithm initially checks the remaining amount of data that 
+			can be written as dirty data (line 6).
+			If this amount is greater than 0, MemoryManager is asked to evict 
+			data from cache to accommodate the written data (line 8). 
+			If the amount to evict calculated and passed in the evict function is less than 0, 
+			the eviction does nothing.
+			After evicting cache, the actual amount of data that can be written to 
+			page cache is calculated (line 9), a cache write is then simulated 
+			with this amount through memory model and data written is added to 
+			page cache LRU lists (line 10).
+			After dirty\_ratio is reached, dirty data has to be flushed before new data 
+			can be written to cache again. 
+			If there is remaining chunk data that has not been written, we flush and 
+			evict data from cache as much as possible so that the this data can be 
 			written to cache (line 13-14). 
-    The Memory Manager does not guarantee that the amount of flushed 
-    and evicted data is enough to write the remaining data. 
+			Here, MemoryManager does not guarantee that the amount of flushed 
+			and evicted data is enough for the write of remaining data. 
 			In the case it is less than the required amount, the remaining data 
@@
 			after the write is limited by the amount free memory, and this amount 
-    is added to cache (line 15) \tristan{the previous two sentences are very unclear, I don't understand them so I can't reformulate.}. 
+			is added to cache (line 15). 
 			Finally, the disk write is simulated with the remaining data amount (line 16), 
-    and the chunk write algorithm finishes \tristan{see comment in algorithm}.
+			and the chunk write algorithm finishes.
 
@@
 			Since bandwidth sharing was not simulated, this prototype does not support 
-            concurrency: it is limited to single-threaded applications running on systems
-            with a single-core CPU. We used this prototype for a first validation of our simulation
-            model against a real sequential application running on a real system.
-            The source code this prototype is available at
-            \url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}. We used Python 3.7.
+			concurrency: it is limited to single-threaded pipelines running on systems 
+			with a single-core CPU. We used it in a first validation of our simulation 
+			model, against a real sequential pipeline running on a real system.
+			The source code of our Python prototype is available at 
+			\url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}
 
-            We also implemented our model in WRENCH, separately from existing WRENCH components. 
-            Page cache simulation is available through a user API and can be activated using a command-line argument.
-            We used SimGrid 3.25 and WRENCH 1.6. SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid},
-            and WRENCH is at \url{https://github.com/wrench-project/wrench/tree/master/src/wrench/services/memory}.
-            \tristan{which commit did you use?}
+			Having the model validated, we created simulators for different use cases 
+			using WRENCH.
+			We implemented our model separately from existing WRENCH components, 
+			exposed an API for users, and added an command line argument to enable 
+			the model in WRENCH. 
+			This helps us to avoid unexpected impacts on WRENCH and 
+			independently update the model in the future. 
+			We then compared the results of the WRENCH, WRENCH-Ext simulators 
+			the results of real execution on a real system.
 
-\tristan{anything else interesting to mention about the implementation?}
+			In this work, we use Python 3.7 to implement the single-threaded simulator, 
+			SimGrid 3.25 and WRENCH 1.6 for more multi-threaded simulators. 
+			SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid}, 
+			and WRENCH is at 
+			\url{https://github.com/wrench-project/wrench/tree/master/src/wrench/services/memory}.
 
@@
 
-        Our experiments compared real executions with the original WRENCH,
-        our WRENCH-cache extension, and our prototype Python simulator. We
-        evaluated our page cache simulation in single-threaded and
-        multi-threaded applications accessing data on local and network
-        file systems.
-
-            We used a pipeline of three sequential tasks where each task reads the output of the previous task,
-            increments every byte of the input to emulate real
-            processing, and writes the resulting data to disk. The first file was provided as input. We also used
-            a real application pipeline \textcolor{red}{[Maybe briefly
-            describe this application's purpose (e.g. a standard fMRI
-            preprocessing piepeline)]}, to evaluate the applicability of
-            our simulation model.
+        	The goal of our experiments is to evaluate our page cache 
+			simulation in single-threaded and multi-threaded applications
+			accessing data on different types of file systems \tristan{"different types" is vague, first thing I thought of was zfs vs ext4 that you don't discuss. Use "local and network" fs instead.}. 
+			We use a pipeline
+			of sequential tasks where each task reads an input file, increments
+			every byte of the input to emulate real processing and avoid caching
+			effects \tristan{ambiguous: you're simulating caching effects, you don't want to avoid them. I'd just remove that note and leave "emulate real processing". You could comment on 
+			the close-to-zero execution time though, since modeling execution time is out of your scope.}, and writes the result to disk. The output of the previous
+			task is the input of the next task. We also use a real application 
+            pipeline \textcolor{red}{[Maybe briefly describe this application's purpose (e.g. a standard fMRI preprocessing piepeline)]}, to evaluate the applicability of our simulation model.
 
 			We use a dedicated cluster hosted at Concordia University to conduct 
-            the experiments. The cluster consists of one login node, 9 compute nodes
-            and 4 storage nodes connected with a network bandwidth of 25 Gbps.
+			the experiments. The cluster consists of one login node, 8 compute nodes \tristan{it now has 9 if you add the GPU node}
+			and 4 storage nodes connected with a network bandwidth of 3GBps \tristan{put the Gbps value, 3GBps is unusual as a network bandwidth.}. 
+			The login node has one Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 
+			137~GB (128~GiB) of RAM, 1.8TB of storage of XFS file system, 
+			13~GB of tmpfs file system and 385 TB of Lustre shared file system \tristan{why do you need to describe the login node in such details?}. 
 			Each compute node has two 16 cores Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 
@@
 			378~GB of tmpfs and 126~GB of devtmpfs file system.
+			Lustre file system is configured with one metadata target with 854~GB 
+			of storage, 44 object storage targets with 8.7~TB of storage each \tristan{why do we care about Lustre?}. 
 			The cluster is run on CentOS 8.1 with NFS version 4. 
@@
 			take the average bandwidths. 
-            The benchmark results of bandwidths and the values used in the simulators
-            in our all experiments are shown in Table~\ref{table:benchmark}.
-
-            \begin{table}[htbp]
-            \centering
-            \begin{tabularx}{\columnwidth}{|l
-            |>{\centering\arraybackslash}X
-            |>{\centering\arraybackslash}X
-            |>{\centering\arraybackslash}X|}
-            \hline
-                Bandwidths (MBps)  & Cluster (real) & Python simulator & WRENCH simulator\\
-            \hline
-                Memory read  & 6860	& 4812	 & 4812\\
-                Memory write & 2764	& 4812 & 4812\\
-                Local disk read & 510 & 465 & 465\\
-                Local disk write & 420 & 465	 & 465\\
-                Remote disk read & 515 & - & 445\\
-                Remote disk write & 375 & - & 445\\
-                Network bandwidth & 3000 & - & 3000\\
-            \hline
-            \end{tabularx}
-            \caption{Bandwidth benchmark results and simulator configurations.
-            The bandwidths used in simulators are average of read and write bandwidths
-            measured on the cluster.
-            Remote disk and network are not simulated in Python simulator.}
-            \label{table:benchmark}
-            \end{table}
-
 			As the pipelines are single-core, we create simulated tasks, 
@@
 			The details of simulated platform and the source code of the simulator 
-            used in our experiments are available in this commit:
-            \url{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}.
+			are available here: 
+			\url{https://github.com/wrench-project/wrench/tree/master/examples/basic-examples/io-pagecache}
+			\tristan{Our experiments use commit abcd}
 
@@
 			The pipeline is simulated with the Python prototype simulator, 
-            original WRENCH and WRENCH-cache simulator.
+			original WRENCH and WRENCH-Ext simulator \tristan{name your simulator something 
+			that refers to page cache, maybe WRENCH-cache}.
 
@@
 			we only compare the results of simulators with original WRENCH, 
-            WRENCH-cache and real pipelines in this experiment. 
+			WRENCH-Ext and real pipelines in this experiment. 
 			The results are compared in total makespan of the pipelines, 
@@
 			We simulate this pipeline with simulators using original WRENCH and 
-            WRENCH-cache. 
+			WRENCH-Ext. 
 			The simulation results are then compared with the real results. 
@@
 	\label{results}			
+		The benchmark results of bandwidths and the values used in the simulators 
+		in our all experiments are shown in Table~\ref{table:benchmark} \tristan{This isn't results, it should go in II.C, 
+		next to the paragraph where you talk about symmetrical bandwidths}.
+
+		\begin{table}
+		\centering
+		\begin{tabular}{|l|c|c|c|}
+		\hline
+			Bandwidths (MBps)  & Cluster (real) & Python simulator & WRENCH simulator\\
+		\hline
+			Memory read  & 6860	& 4812	 & 4812\\
+			Memory write & 2764	& 4812 & 4812\\
+			Local disk read & 510 & 465 & 465\\
+			Local disk write & 420 & 465	 & 465\\
+			Remote disk read & 515 & - & 445\\
+			Remote disk write & 375 & - & 445\\
+			Network bandwidth & 3000 & - & 3000\\
+		\hline
+		\end{tabular}
+		\caption{Bandwidth benchmark results and simulator configurations \tristan{add a sentence of explanation. Make the table fit in 1 column}}
+		\label{table:benchmark}
+		\end{table}
 
@@
 
-            Memory profiling results of real pipeline execution and simulators 
-			with 20GB and 100GB input files are illustrated in Fig.~\ref{fig:single_memprof}.
-			Because the results with 20Gb, 50GB, 100GB shows a similar pattern,
-			we only include the results with 20GB and 100GB in this paper.
+			Memory behaviors of real pipeline execution and simulators are recorded 
+			and the memory profiling results with 20GB and 100GB input files 
+			are illustrated in Figure~\ref{fig:single_memprof} \tristan{convoluted writing}. 
+			The bars in the figures present activities in tasks (read, compute and write), 
+			while the lines describe memory status throughout run time of the pipeline \tristan{move any description of graphical features to the figure caption}. 
+			Because the results with 20Gb, 50GB, 100GB are analogous, we only include 
+			the results with 20GB and 100GB in this paper \tristan{analogous is not the right word}.
 
@@
 			and simulation show that the use of page cache is well reproduced by the simulators. 
-			The amount of dirty data (red lines) also follow similar trends,
+			The amount of dirty data ( red lines) also follow similar trends, 
 			which means not only the amount dirty data, but also the flushing and 
@@
 			The I/O time of tasks in simulators and real execution is also tracked and compared. 
-			Fig.~\ref{fig:single_error} shows the relative error of simulated I/O time
+			Figure~\ref{fig:single_errors} \tristan{wrong reference} shows the relative error of simulated I/O time 
 			of individual I/O activities. 
@@
 
+			In general, with all input sizes in the experiment, the simulation errors 
+			of simulators with page cache model (Python and WRENCH-Ext simulators) 
+			are remarkably lower than the result with original WRENCH . 
+			The reason is, in original WRENCH, because page cache is not simulated, 
+			data is only read/written with disk bandwidth, which is much slower 
+			than memory bandwidths, leading to simulated I/O time longer than in reality.
+			Especially, the simulation time errors of \textit{Read 3} in the simulators 
+			with page cache are just under the error in original WRENCH.
+
 			\begin{figure*}
@@
 			\end{subfigure}
-			\caption{Memory profiling results with different input file sizes.
-			The bars in the figures present activities in tasks (read, compute and write),
-			the lines describe memory status throughout run time of the pipeline.}
+			\caption{Memory profiling results with different input file sizes \tristan{put the simulation/real conditions on the side, make them larger and use the same color code as below (pink for real, etc).
+Put the memory values on top, similar to how they are in the graphs below.
+}}
 			\label{fig:single_memprof}	
@@
 			\label{fig:single_cache}
-   			\caption{Amount of file data in cache after each I/O operation. 
-   			\textcolor{red}{Update real results of 20GB}}
+   			\caption{Amount of file data in cache after each I/O activity with 100GB of input \tristan{Make a grouped histogram (1 histogram with two bars, one for real, one for simulation).
+Add a grouped histogram for 20GB.
+Use the same color code everywhere: real should be pink.}}
 			\end{figure*}
 
-			In general, with all input sizes in the experiment, the simulation errors
-			of simulators with page cache model (Python and WRENCH-cache simulators)
-			are remarkably lower than the result with original WRENCH .
-			The reason is, in original WRENCH, because page cache is not simulated,
-			data is only read and written with disk bandwidth, which is much slower
-			than memory bandwidths, leading to simulated I/O time longer than in reality.
-			Especially, the simulation time errors of \textit{Read 3} with 100 GB of input
-			in the simulators with page cache are just under the error in original WRENCH.
 			This longer read can be explained with the amount of cached data of files 
 			at the beginning of the read in reality and simulators shown in 
-			Fig.~\ref{fig:single_cache}.
+			Figure~\ref{fig:single_cache}. 			
 			In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and 
 			output of \textit{Write $<$i-1$>$}. 
-			With the input size of 20 GB, in both real execution and WRENCH-cache
-			simulator, there is no difference in the page cache since all files in the
-			pipeline can be entirely cached.
-			With the input of 100 GB, the difference between real execution and WRENCH-cache
-			simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
-			In real execution, after being written by \textit{Write 2},
-			\textit{file 3} is entirely in cache.
+			In the real execution, after \textit{Write 2}, which writes \textit{file 3}, 
+			all data of \textit{file 3} is in cache. 
 			Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth. 
-			However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
-			cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
-			\textit{file 3} from disk.
-			The reason for the difference is in our WRENCH-cache simulator,
-			with 100 GB of input, while \textit{file 3} is being written,
-			the page cache is saturated, eviction is then evoked to evict data
-			from page cache and make space available for \textit{file 3}.
+			But in our simulator, after \textit{Write 2}, only a part of \textit{file 3} 
+			remains in cache. 
+			The reason is while \textit{file 3}  is being written, the page cache is saturated, 
+			eviction is then evoked to evict data from page cache, make space 
+			available for \textit{file 3}. 
 			However, because the evicted amount is less than required, 
@@
 			\end{subfigure}		
-            \caption{I/O time of concurrent pipelines with local storage.}
+			\caption{I/O time of concurrent pipelines with local storage. \tristan{revise caption for real execution variability: "real execution" doesnt tell what the gray shade is.}}
 			\label{fig:multi_local}
@@
 
-            In this experiment, we run the real execution in 5 repetitions to
-            capture the variability on the cluster.
-            The results are analyzed in terms of average I/O time
-            of concurrent pipelines shown in Fig.~\ref{fig:multi_local}.
+			In this experiment, the results are analyzed in terms of average I/O time 
+			of concurrent pipelines shown in Figure~\ref{fig:multi_local}.
 
@@
 			increases but memory and disk bandwidths are fixed and shared between pipelines. 
-            In comparison with original WRENCH, the results from WRENCH-cache are closer
+			In comparison with original WRENCH, the results from WRENCH-Ext are closer 
 			to reality. 
 			Especially, when it comes to the average write time, 
-            although the trend in reality is more complex, our WRENCH-cache
+			although the trend in reality is more complex, our WRENCH-Ext 
 			simulator can still capture the pattern in real execution.
@@
 
-            To conclude, the results from WRENCH-cache simulator show that
+			To conclude, the results from WRENCH-Ext simulator show that 
 			our model can also simulate I/O with page cache accurately in a 
@@
 
-            Similar to the previous experiment, the real execution with NFS is also
-            repeated for 5 times and the results are described with the average I/O time
-            shown in Fig.~\ref{fig:multi_nfs}.
+			Similarly, the results with from the experiment with NFS are described 
+			by the average I/O time shown in Figure~\ref{fig:multi_nfs}.
 			In general, the average I/O time reflect a very similar level of accuracy 
@@
 			the page cache is saturated, which has been seen in 
-            Fig.~\ref{fig:single_error} with 100 GB of input.
-            When it comes to write time, original WRENCH and WRENCH-cache
+			Figure~\ref{fig:single_error} with 100 GB of input. 
+			When it comes to write time, original WRENCH and WRENCH-Ext 
 			results are identical when both are writing directly to disk.
@@
 			with different experiments. 
-        The results shows that our model can result in better simulation accuracy than
-        the simulators in which page cache is not modeled.
-        Furthermore, this level of accuracy can be achieved without a fine-grained
-        level of details.
-        This may create opportunities for the model to be applied in simulation of
-        big data applications on HPC infrastructures, which leverage memory to
-        employ page cache as a disk cache.
-        Nevertheless, the model exposed some limitations that can be
+			The results shows our achievements in terms of simulation accuracy 
+			and the application potential in practice \tristan{what do you conclude? can we use your simulator to study big data applications on hpc clusters? under what conditions?}. 
+			On the other hands, the model exposed some limitations that can be 
 			further improved.

### 2020-12-08 paper_io_sim.tex
edits in results

--- 2020-12-07 paper_io_sim.tex
@@
 
+            \tristan{mention that cache was flushed before each experiment}
+
     \section{Results}
@@
         \subsection{Single-threaded experiment}
+
+        Figure~\ref{fig:single_error} shows the relative error of simulated
+        read and write times throughout the three application tasks. The
+        page cache simulation model drastically reduces simulation errors
+        for all I/O operations except the first read that deals with
+        uncached data. Errors are reduced from an average of \texttt{xxx\%}
+        in WRENCH to \texttt{yyy \%} in the Python prototype and
+        \texttt{zzz \%} in WRENCH-cache. The original WRENCH simulator significantly overestimates
+        read and write times, due to its lack of page cache simulation
+        model.
+
+        % In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
+        % output of \textit{Write $<$i-1$>$}.
+        With the input size of 20 GB, in both real execution and WRENCH-cache
+        simulator, there is no difference in the page cache since all files in the
+        pipeline can be entirely cached \tristan{what does this refer to? 
+        For 20 GB there is still some error in WRENCH-cache, what do you mean ``there is no difference in the page cache''?}.
+
+        With the input of 100 GB, the difference between real execution and WRENCH-cache
+        simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
+        In real execution, after being written by \textit{Write 2},
+        \textit{file 3} is entirely in cache.
+        Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
+        However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
+        cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
+        \textit{file 3} from disk.
+        The reason for the difference is in our WRENCH-cache simulator,
+        with 100 GB of input, while \textit{file 3} is being written,
+        the page cache is saturated, eviction is then evoked to evict data
+        from page cache and make space available for \textit{file 3}.
+        However, because the evicted amount is less than required,
+        some \textit{file 3} data is evicted after being written, make \textit{file 3}
+        partially cached.
+        As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
+        with memory bandwidth, a part of it is read from disk with disk read bandwidth,
+        which is many times slower, leading to longer read time.
 
@@
 
-			The I/O time of tasks in simulators and real execution is also tracked and compared.
-			Fig.~\ref{fig:single_error} shows the relative error of simulated I/O time
-			of individual I/O activities.
-			Our pipeline consists of 3 tasks, and each task reads an input file,
-			performs some computation, and writes output file.
-			Thus, there are 6 sequential I/O activities, \textit{Read 1} and \textit{Write 1}
-			from the first task, \textit{Read 2} and \textit{Write 2} the second task,
-			and \textit{Read 3}, \textit{Write3} from the last task.
-
             \begin{figure*}
             \centering
+            \begin{subfigure}{\linewidth}
+                \centering
+                   \includegraphics[width=\linewidth]{result/single/figures/single_errors.pdf}
+            \end{subfigure}
+            \caption{Simulation errors with different input file sizes}
+            \label{fig:single_error}
             \begin{subfigure}{\linewidth}
@@
                 \centering
-   				\includegraphics[width=\linewidth]{result/single/figures/single_errors.pdf}
-			\end{subfigure}
-			\caption{Simulation errors with different input file sizes}
-			\label{fig:single_error}
-			\begin{subfigure}{\linewidth}
-				\centering
                    \includegraphics[width=\linewidth]{result/single/figures/cached_files.pdf}
@@
             \end{figure*}
-
-			In general, with all input sizes in the experiment, the simulation errors
-			of simulators with page cache model (Python and WRENCH-cache simulators)
-			are remarkably lower than the result with original WRENCH .
-			The reason is, in original WRENCH, because page cache is not simulated,
-			data is only read and written with disk bandwidth, which is much slower
-			than memory bandwidths, leading to simulated I/O time longer than in reality.
-			Especially, the simulation time errors of \textit{Read 3} with 100 GB of input
-			in the simulators with page cache are just under the error in original WRENCH.
-			This longer read can be explained with the amount of cached data of files
-			at the beginning of the read in reality and simulators shown in
-			Fig.~\ref{fig:single_cache}.
-			In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
-			output of \textit{Write $<$i-1$>$}.
-			With the input size of 20 GB, in both real execution and WRENCH-cache
-			simulator, there is no difference in the page cache since all files in the
-			pipeline can be entirely cached.
-			With the input of 100 GB, the difference between real execution and WRENCH-cache
-			simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
-			In real execution, after being written by \textit{Write 2},
-			\textit{file 3} is entirely in cache.
-			Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
-			However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
-			cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
-			\textit{file 3} from disk.
-			The reason for the difference is in our WRENCH-cache simulator,
-			with 100 GB of input, while \textit{file 3} is being written,
-			the page cache is saturated, eviction is then evoked to evict data
-			from page cache and make space available for \textit{file 3}.
-			However, because the evicted amount is less than required,
-			some \textit{file 3} data is evicted after being written, make \textit{file 3}
-			partially cached.
-			As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
-			with memory bandwidth, a part of it is read from disk with disk read bandwidth,
-			which is many times slower, leading to longer read time.
 

### 2020-12-08 paper_io_sim.tex
replaced tabs with spaces (again)
Edits in method

@@
 
-        In this paper, we present WRENCH-Cache, a simulation model of I/O with page
+        In this paper, we present WRENCH-cache, a simulation model of I/O with page
         cache incorporated into WRENCH, a workflow management system simulator based of
-        the popular SimGrid framework. The goal of WRENCH-Cache is to provide users
+        the popular SimGrid framework. The goal of WRENCH-cache is to provide users
         with an accurate representation of a workflow execution, when file system page
-        cache is used. WRENCH-Cache is able to simulate both writeback and writethrough
+        cache is used. WRENCH-cache is able to simulate both writeback and writethrough
         cache which would normally be available to local and NFS-based filesystems.
@@
     In this section, we present our page cache simulation model for file I/O,
-    flusing, and eviction. We detail the design and implementation of two simulators:
-    (1) a standalone Python prototype to evaluate the simulation model independently of any other 
-    tool, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
+    flusing, and eviction. We implement this model in two simulators:
+    (1) a standalone Python prototype, and (2) WRENCH-cache, an extension of the WRENCH library.
     We also describe experimental scenarios to evaluate our simulators in various configurations.
@@
     To read or write a file chunk, a simulated application sends a
-    request to the IOController. The IOController interacts as needed with
+    request to the IO Controller. The IO Controller interacts as needed with
     the Memory Manager to free memory through flushing or eviction,
@@
            \includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
-           \caption{Overview of the simulated writeback page cache.
+           \caption{Overview of the simulated writeback page cache \tristan{does it simulate only writeback? See comment in NFS section.}.
            Applications send file read or write requests to the
-           IOController that orchestrates flushing, eviction, cache
-           and disk accesses with the MemoryManager. Storage
-           devices (memory and disk) simulate file transfers with
-           bandwidth sharing.}
+           IO Controller that orchestrates flushing, eviction, cache
+           and disk accesses with the Memory Manager. Concurrent accesses to storage
+           devices (memory and disk) are simulated using existing models.}
            \label{fig:interaction}
@@
     expired dirty data in page cache LRU lists and flushes them to disk. We
-    use existing storage simulation models~\cite{lebre2015} to simulate
+    use existing storage simulation models~\cite{lebre2015} to simulate disk and
     memory, characterized by its storage capacity, read and write
     bandwidths, and latency. Reusing existing storage models allows us to
-    simulate bandwidth sharing between concurrent memory accesses.
+    simulate bandwidth sharing between concurrent memory accesses and concurrent disk accesses.
 
@@
     two lists of data blocks, an active list and an inactive list, both ordered by 
-    last access time (earliest first, Figure~\ref{fig:lrulist}).
+    last access time (earliest first, see Figure~\ref{fig:lrulist}).
     As in the kernel, our simulator limits the size of the active list to
@@
 
-    \subsubsection{Reads and Writes}
+    \subsubsection{Reads and writes}
 
@@
     before data from the active list
-    (left to right in Figure~\ref{fig:read_order}).
+    (data reads occur from left to right in Figure~\ref{fig:read_order}).
     When a chunk of \emph{uncached} data is read, a new clean block is created 
-    and put at the top of the inactive list. 
+    and appended to the inactive list. 
     When a chunk of \emph{cached} data is read, one or more existing data blocks in the LRU lists are accessed.
     If these blocks are clean, we merge them together, update the access time and size of the resulting block, 
-    and put it in the active list. 
-    If the blocks are dirty, we move them to the active list to preserve their entry time. 
-    Because the chunk size and block size can be different, there are situations
+    and append it to the active list. 
+    If the blocks are dirty, we move them independently to the active list, to preserve their entry time. 
+    Because the chunk and block sizes may be different, there are situations
     where a block is not entirely read. 
@@
            \caption{File data read order. Data is read from left to right: uncached data 
-           is read first, followed by data from the inactive list, and finally data from the active list.}
+           is read first, followed by data from the inactive list, and finally data from the active list. \tristan{Colors 
+           shouldn't match that in Fig 1 as they dont represent the same thing.}}
            \label{fig:read_order}
@@
     uncached. Thus, each time a chunk is written, we create a block of dirty data 
-    and add it to the inactive list.
+    and append it to the inactive list.
 
@@
 
-    The main simulated thread in the Memory Manager can flush the
-    cache to write dirty data to disk. The data flushing simulation
+    The main simulated thread in the Memory Manager can flush or evict data from the
+    memory cache. The data flushing simulation
     function takes the amount of data to flush as parameter. While
@@
     inactive list, then the sorted active list, and writes the
-    least recently used dirty block to disk having set its dirty
+    least recently used dirty block to disk, having set its dirty
     flag to 0. In case the amount of data to flush requires that a
@@
     one that is flushed and one that remains dirty. Flushing time
-    to disk is simulated with the storage model.
+    to disk is simulated by the storage model.
 
-    Similarly to data flushing, our cache eviction simulation runs in
+    The cache eviction simulation also runs in
     the main thread. It frees up the page cache by traversing and deleting 
@@
     The amount of data to evict is passed as a parameter and data blocks are deleted 
-    from the inactive list until total evicted data reaches the required amount,
+    from the inactive list until the evicted data reaches the required amount,
     or until there is no clean block left in the list.
@@
     Because periodical flushing is simulated as a background thread, it can happen concurrently
-    with disk I/O initiated by the main thread. This parallelism is handled by the 
+    with disk I/O initiated by the main thread. This is taken into account by the 
     storage model and reflected in simulated I/O time.
@@
 
-    \begin{algorithm}\caption{File chunk read simulation of IOController}
+    \begin{algorithm}\caption{File chunk read simulation of IO Controller}
     \label{alg:read}
@@
                \EndInput
-               \State disk\_read = min(cs, fs - mm.cached(fn))
-               \State cache\_read = cs - disk\_read
+               \State disk\_read = min(cs, fs - mm.cached(fn)) \Comment{To be read from disk} 
+               \State cache\_read = cs - disk\_read \Comment{To be read from cache} 
             \State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
@@
     round-robin fashion. To read a file chunk, simulated applications send
-    chunk read requests to the IOController which processes them using
+    chunk read requests to the IO Controller which processes them using
     Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
@@
     to flush dirty data (line 9). If necessary, flushing is complemented by
-    eviction (line 10). When called with negative arguments, functions
-    flush and evict simply return and do not do anything. Then, if there is
-    uncached data, a disk read is simulated with disk model (line 12)
+    eviction (line 10). Note that, when called with negative arguments, functions
+    \texttt{flush} and \texttt{evict} simply return and do not do anything. Then, if the block requires
+    uncached data, a disk read is simulated with the disk model (line 12)
     \tristan{this is not consistent with the new Figure 1 where disk reads
-    are done by the mm}, and this data of the file is added to cache (line
+    are done by the mm}, and the data read is added to cache (line
     13).
@@
 
-    \begin{algorithm}\caption{File chunk write simulation of IOController}
+    \begin{algorithm}\caption{File chunk write simulation of IO Controller}
     \label{alg:write}
@@
                 \State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem)) \tristan{why is data written to cache with "addtocache" here and with "write" at line 10?}
-                \State sm.write(cs - mem\_amt) \tristan{According to Fig 1 the IOController is not supposed to interact directly with sm}
+                \State sm.write(cs - mem\_amt) \tristan{According to Fig 1 the IO Controller is not supposed to interact directly with sm}
             \EndIf
@@
     Algorithm~\ref{alg:write} describes our simulation of chunk writes in 
-    the IOController. 
+    the IO Controller. 
     % Data is written to cache until the amount of dirty data in the cache 
@@
     If this amount is greater than 0, the Memory Manager is requested to evict 
-    data from cache if necessary (line 8) \tristan{the difference between available and free memory should be explained somewhere}.
+    data from cache if necessary (line 8) \tristan{the difference between available and free memory should be explained the first time it is referred}.
     After eviction, the amount of data that can be written to 
@@
             We also implemented our model in WRENCH, separately from existing WRENCH components. 
-            Page cache simulation is available through a user API and can be activated using a command-line argument.
+            We exposed page cache simulation through a user API, activated with a command-line argument.
             We used SimGrid 3.25 and WRENCH 1.6. SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid},
@@
 
-        Our experiments compare real executions with the original WRENCH,
-        with our WRENCH-cache extension, and with our prototype Python simulator. We
+        Our experiments compare real executions with our Python prototype,
+        with the original WRENCH, and with our WRENCH-cache extension. We
         evaluate our page cache simulation in single-threaded and
@@
 
-        We simulate an application with three single-core, sequential tasks
+        We simulate an application that consists of three single-core, sequential tasks
         where each task reads the file produced by the previous task,
@@
         memory, local disk, remote disk (NFS), and network
-        (Table~\ref{table:benchmark}). Since SimGrid only supports
+        (Table~\ref{table:benchmark}). Since WRENCH currently only supports
         symmetrical bandwidths, we use the mean of the read and write
@@
             sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed
-            to the same local disk.
+            to the same local disk. The goal is to validate the simulation model 
+            in absence of any concurrent disk access.
 
@@
             Our third experiment occurs in the same configuration as the
-            previous one, albeit on a 400-GB NFS-mounted remote disk
+            previous one, albeit reading and writing on a 400-GB remote NFS disk
             mounted from another compute node. \tristan{As is commonly
@@
             writing data directly to disk and adding them to the server
-            cache \tristan{it it part of the IOController? it would deserve
+            cache \tristan{it it part of the IO Controller? it would deserve
             a few sentences in the methods.}.

### 2020-12-08 paper_io_sim.tex
Update section III, IV, figures

@@
             In writeback, one of the keys is the cache eviction mechanism, which is
-            triggered when available memory is insufficient to removes pages 
+            triggered when available memory is insufficient to remove pages 
             from page cache. 
+            The term \textit{available memory} here includes both free memory and 
+            the memory used but can be reclaimed by the kernel with flushing 
+            and eviction. 
             Because dirty pages are data written to page cache but yet to be 
@@
     and eviction from cache, and reads or writes to disk when necessary.  
+    In case write-through strategy is used, IO Controller can also interact directly 
+    with disk model to simulate disk writes.
+
     \begin{figure}
@@
            \includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
-           \caption{Overview of the simulated writeback page cache \tristan{does it simulate only writeback? See comment in NFS section.}.
+           \caption{Overview of the page cache simulator.
            Applications send file read or write requests to the
@@
     used data blocks from the active list to the inactive list~\cite{gorman2004understanding, linuxdev3rd2010}.
-    % Modeling page cache LRU lists as lists of data blocks reduces the
-    % overhead of the simulator, as data blocks are obviously less
-    % numerous than file pages. At the same time, it achieves the accuracy in 
-    % I/O time simulation that can be seen in the later section.
-    % From Tristan: I don't think the sentences above are necessary,
-    % they are quite repetitive.
 
     At any given time, a file can be partially cached, completely cached,
-    or uncached at all. Cached data blocks may reside in one or both of the
+    or uncached at all. A cached data block can only reside in one of two 
     page cache LRU lists. The first time they are accessed, blocks are
     added to the inactive list. On subsequent accesses, blocks of the
-    inactive list are moved to \tristan{if they are moved then they can't
-    be in both lists} the top of the active list. Cached blocks written to
-    cache are marked as dirty until they are flushed to disk.
+    inactive list are moved to the top of the active list. Cached blocks 
+    written to cache are marked as dirty until they are flushed to disk.
 
@@
            \caption{File data read order. Data is read from left to right: uncached data 
-           is read first, followed by data from the inactive list, and finally data from the active list. \tristan{Colors 
-           shouldn't match that in Fig 1 as they dont represent the same thing.}}
+           is read first, followed by data from the inactive list, and finally data from the active list. }
            \label{fig:read_order}
@@
                 \State flushing\_time = sm.write(blocks)
-                \State flag blocks as clean data \tristan{a bit informal}
+                \State for blk in blocks: blk.dirty = 0
                 \If{flushing\_time $<$ t}
                     \State sleep(t - flushing\_time)
-                \EndIf \tristan{else?}
+                \EndIf  % Else: do nothing, continue the loop
             \EndWhile
@@
                \State cache\_read = cs - disk\_read \Comment{To be read from cache}
+               \State required\_mem = cs + disk\_read
                \State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
@@
             \EndIf
+            \State mm.use\_anonymous\_mem(cs)
         \end{algorithmic}
@@
     data that needs to be read from disk, and the remaining amount is read
-    from cache (line 7-8). The amount of memory required to read the chunk
+    from cache (line 6-7). The amount of memory required to read the chunk
     is calculated, corresponding to a copy of the chunk in anonymous memory
-    and a copy of the chunk in cache \tristan{Create a variable
-    "required\_mem" in the algorithm to clarify}.
-    \tristan{add a call to mm in the algorithm to increment anonymous memory. Mention in the experiments that anonymous memory is flushed after each task.}
+    and a copy of the chunk in cache (line 8). 
     If there is not enough available memory, the Memory Manager is called
@@
     eviction (line 10). Note that, when called with negative arguments, functions
-    \texttt{flush} and \texttt{evict} simply return and do not do anything. Then, if the block requires
-    uncached data, a disk read is simulated with the disk model (line 12)
-    \tristan{this is not consistent with the new Figure 1 where disk reads
-    are done by the mm}, and the data read is added to cache (line
-    13).
-    Finally, if cached data needs to be read, the Memory Manager is called
-    to simulate a cache read  
-    and update the corresponding data blocks accordingly (line 16).
+    \texttt{flush} and \texttt{evict} simply return and do not do anything. Then, if there is
+    uncached data, memory manager is called to read data from disk and add this 
+    data to cache (line 12).
+    If cached data needs to be read, the Memory Manager is called to simulate 
+    a cache read  and update the corresponding data blocks accordingly (line 15).
+    Finally, the memory manager is called to reduce the amount of free memory 
+    corresponding to the amount of anonymous memory used by application (line 17).
 
@@
                 \Desc{mm}{MemoryManager object}
-                \Desc{sm}{storage simulation model}
                \EndInput
@@
                 \State mem\_amt = min(cs, mm.free\_mem)
-                \State mm.write(fn, mem\_amt) 
+                \State mm.write\_to\_cache(fn, mem\_amt) 
             \EndIf
-            \If {mem\_amt $<$ cs}  \Comment{Write with disk bandwidth}
+            \State remaining = cs - mem\_amt
+            \While {remaining $>$ 0}  \Comment{Write with disk bandwidth}
                 \State mm.flush(cs - mem\_amt)
                 \State mm.evict(cs - mem\_amt  - mm.free\_mem) 
-                \State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem)) \tristan{why is data written to cache with "addtocache" here and with "write" at line 10?}
-                \State sm.write(cs - mem\_amt) \tristan{According to Fig 1 the IO Controller is not supposed to interact directly with sm}
-            \EndIf
+                \State to\_cache = min(remaining, mm.free\_mem)
+                \State mm.write\_to\_cache(fn, to\_cache) 
+                \State remaining = remaining - to\_cache
+            \EndWhile
 
@@
     Our algorithm initially checks the  amount of data that 
-    can be written as dirty data given the dirty ratio (line 6).
+    can be written as dirty data given the dirty ratio (line 5).
     If this amount is greater than 0, the Memory Manager is requested to evict 
-    data from cache if necessary (line 8) \tristan{the difference between available and free memory should be explained the first time it is referred}.
+    data from cache if necessary (line 7).
     After eviction, the amount of data that can be written to 
-    page cache is calculated (line 9), and a cache write is simulated (line 10).
-    To write the remaining data, we flush and 
-    evict data as possible so that this data can be 
-    written to cache (line 13-14). 
-    The Memory Manager does not guarantee that the amount of flushed 
-    and evicted data is enough to write the remaining data. 
-    In the case it is less than the required amount, the remaining data 
-    being written to cache can be flushed and evicted right away. 
-    Thus, the amount of remaining data written and kept in cache 
-    after the write is limited by the amount free memory, and this amount 
-    is added to cache (line 15) \tristan{the previous two sentences are very unclear, I don't understand them so I can't reformulate.}. 
-    Finally, the disk write is simulated with the remaining data amount (line 16), 
-    and the chunk write algorithm finishes \tristan{see comment in algorithm}.
+    page cache is calculated (line 8), and a cache write is simulated (line 9). 
+    If there is remaining data, this data is written to cache in a loop (line 12 - line 18).
+    In this loop, we flush and evict data as possible so that there will be 
+    available space for writing (line 13-14).  
+    After flushing and eviction, the amount of data that can be written 
+    is calculated (line 15) and a cache write is simulated (line 16). 
+    We update the remaining amount (line 17) and continue the loop until all data 
+    is written.
+
+    In addition to the chunk write simulation with writeback page cache above, 
+    we also add a chunk write function with write-through strategy to the IOController.
+    This function simply simulates a disk write with the amount of data passed in, 
+    then evicts cache if needed and adds this data to cache.
 
@@
             We also implemented our model in WRENCH, separately from existing WRENCH components. 
-            We exposed page cache simulation through a user API, activated with a command-line argument.
-            We used SimGrid 3.25 and WRENCH 1.6. SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid},
-            and WRENCH is at \url{https://github.com/wrench-project/wrench/tree/master/src/wrench/services/memory}.
-            \tristan{which commit did you use?}
+            We exposed page cache simulation through a user API, activated using a command-line argument.
+            We used SimGrid 3.25 which is available at \url{https://framagit.org/simgrid/simgrid}.
+            To implement the simulator, we used the WRENCH 1.6 version at this commit \url{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}.
 
@@
         increments every byte of this file to emulate real processing, and
-        writes the resulting data to disk. We also use a real application
+        writes the resulting data to disk. 
+        In real execution, we clear cache before starting the application to make sure 
+        the conditions are consistent, the cache is not in use and it is only used by 
+        our application.
+        In addition, the anonymous memory used by application 
+        is released after each task in the pipeline. 
+        Thus, we also release the memory used at the end of each task in our simulator.
+        We also use a real application
         \textcolor{red}{[Maybe briefly describe this application's purpose
@@
             \centering
-            \begin{tabularx}{\columnwidth}{|c
+            \begin{tabularx}{\columnwidth}{|l
             |>{\centering\arraybackslash}X
@@
             \hline
-                Device  & Cluster (real) & Python simulator & WRENCH simulator\\
+                Bandwidths  & Cluster (real) & Python simulator & WRENCH simulator\\
             \hline
@@
                 Remote disk write & 375 & - & 445\\
-                Network bandwidth & 3000 & - & 3000\\
+                Network & 3000 & - & 3000\\
             \hline
@@
             of flops on a 1~Gflops CPU and used these values in the
-            simulation (see Table~\ref{table:tasks} \tristan{add a table to describe execution times}). The final simulated
+            simulation (see Table~\ref{table:cputime}). The final simulated
             platform and application used in our experiment are avaiable at
             \url{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}.
+
+            \begin{table}[htbp]
+            \centering
+            \begin{tabularx}{0.8\columnwidth}{|l|>{\centering\arraybackslash}X|}
+            \hline
+                Input size (GB)  & CPU time (second)\\
+            \hline
+                3 	 & 4.4 \\
+                20  & 28 \\
+                50  & 75 \\
+                75  & 110 \\
+                100  & 155 \\
+            \hline
+            \end{tabularx}
+            \caption{Task CPU time corresponding to input file size.}
+            \label{table:cputime}
+            \end{table}
 
@@
             Our third experiment occurs in the same configuration as the
-            previous one, albeit reading and writing on a 400-GB remote NFS disk
-            mounted from another compute node. \tristan{As is commonly
-            configured in HPC environments to avoid data loss,?} NFS client
-            and server read caches are enabled, but server writes are configured with
-            write-through cache instead of write back \tristan{how about client write
-            cache?}. Therefore, all the writes happen at disk bandwidth, but
+            previous one, albeit reading and writing on a 400-GB NFS-mounted remote disk
+            mounted from another compute node. As is commonly
+            configured in HPC environments to avoid data loss, in our simulator, 
+            NFS client and server read caches are enabled, but there is no cache 
+            for writing on client side and server writes are configured with
+            write-through cache instead of writeback. 
+            Therefore, all the writes happen at disk bandwidth, but
             reads can benefit from cache hits. We simulate write-through by
-            writing data directly to disk and adding them to the server
-            cache \tristan{it it part of the IO Controller? it would deserve
-            a few sentences in the methods.}.
+            writing data directly to disk and adding that data to the server
+            cache.
 
@@
 
-            \tristan{mention that cache was flushed before each experiment}
-
     \section{Results}
@@
         \subsection{Single-threaded experiment}
-
-        Figure~\ref{fig:single_error} shows the relative error of simulated
-        read and write times throughout the three application tasks. The
-        page cache simulation model drastically reduces simulation errors
-        for all I/O operations except the first read that deals with
-        uncached data. Errors are reduced from an average of \texttt{xxx\%}
-        in WRENCH to \texttt{yyy \%} in the Python prototype and
-        \texttt{zzz \%} in WRENCH-cache. The original WRENCH simulator significantly overestimates
-        read and write times, due to its lack of page cache simulation
-        model.
-
-        % In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
-        % output of \textit{Write $<$i-1$>$}.
-        With the input size of 20 GB, in both real execution and WRENCH-cache
-        simulator, there is no difference in the page cache since all files in the
-        pipeline can be entirely cached \tristan{what does this refer to? 
-        For 20 GB there is still some error in WRENCH-cache, what do you mean ``there is no difference in the page cache''?}.
-
-        With the input of 100 GB, the difference between real execution and WRENCH-cache
-        simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
-        In real execution, after being written by \textit{Write 2},
-        \textit{file 3} is entirely in cache.
-        Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
-        However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
-        cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
-        \textit{file 3} from disk.
-        The reason for the difference is in our WRENCH-cache simulator,
-        with 100 GB of input, while \textit{file 3} is being written,
-        the page cache is saturated, eviction is then evoked to evict data
-        from page cache and make space available for \textit{file 3}.
-        However, because the evicted amount is less than required,
-        some \textit{file 3} data is evicted after being written, make \textit{file 3}
-        partially cached.
-        As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
-        with memory bandwidth, a part of it is read from disk with disk read bandwidth,
-        which is many times slower, leading to longer read time.
 
@@
             than in real execution. This will be explained later in this section.
+
+             Figure~\ref{fig:single_error} shows the relative error of simulated
+             read and write times throughout the three application tasks. The
+             page cache simulation model drastically reduces simulation errors
+             for all I/O operations except the first read that deals with
+             uncached data. Errors are reduced from an average of 345\%
+             in WRENCH to 46\% in the Python prototype and 39\% in WRENCH-cache.
+             The original WRENCH simulator significantly overestimates
+             read and write times, due to its lack of page cache simulation
+             model.
+             Especially, the simulation time errors of \textit{Read 3} with 100 GB of input
+             in the simulators with page cache are just under the error in original WRENCH.
+             This longer read can be explained with the amount of cached data of files
+             at the beginning of the read in reality and simulators shown in
+             Fig.~\ref{fig:single_cache}.
+             % In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
+             % output of \textit{Write $<$i-1$>$}.
+             With the input size of 20 GB, the amount of cached data of each files after 
+             every I/O operation is similar since the page cache can keep all these files.
+%             \tristan{what does this refer to?
+%             For 20 GB there is still some error in WRENCH-cache, what do you mean 
+%``there is no difference in the page cache''?}.
+%   Dzung: Some text was lost in a merge, I added it again.
+             With the input of 100 GB, the difference between real execution and WRENCH-cache
+             simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
+             In real execution, after being written by \textit{Write 2},
+             \textit{file 3} is entirely in cache.
+             Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
+             However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
+             cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
+             \textit{file 3} from disk.
+             The reason for the difference is in our WRENCH-cache simulator,
+             with 100 GB of input, while \textit{file 3} is being written,
+             the page cache is saturated, eviction is then evoked to evict data
+             from page cache and make space available for \textit{file 3}.
+             However, because the evicted amount is less than required,
+             some \textit{file 3} data is evicted after being written, make \textit{file 3}
+             partially cached.
+             As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
+             with memory bandwidth, a part of it is read from disk with disk read bandwidth,
+             which is many times slower, leading to longer read time.
 

### 2020-12-08 paper_io_sim.tex
edits in results

--- 2020-12-08 paper_io_sim.tex
@@
         page cache simulation model drastically reduces simulation errors
-        for all I/O operations except the first read that deals with
+        for all I/O operations except the first read as it only involves
         uncached data. Errors are reduced from an average of \texttt{xxx\%}
         in WRENCH to \texttt{yyy \%} in the Python prototype and
-        \texttt{zzz \%} in WRENCH-cache. The original WRENCH simulator significantly overestimates
+        \texttt{zzz \%} in WRENCH-cache \tristan{Dzung, update average percentages}.
+        Unsurprisingly, the original WRENCH simulator significantly overestimates
         read and write times, due to its lack of page cache simulation

### 2020-12-08 gen_fig.py
Update generate figure scripts

@@ -3,15 +3,25 @@ import sys
 sys.path.append(os.path.abspath("result/single/"))
 sys.path.append(os.path.abspath("result/multi/"))
 
+
+# ======== Single-threaded experiment results ==========
 os.chdir("result/single")
 import result.single.plot_memprof as plot_memprof
 import result.single.plot_error as plot_error
 import result.single.plot_cache as plot_cache
+
+# Generate memory profiling comparision between 20GB and 100GB inputs.
+# xmax is set to 200 and 1300
 plot_memprof.compare_plot([20, 100], [200, 1300])
+# Generate simulation errors with 20GB and 100GB inputs
 plot_error.plot_error()
+# Generate amount of cached data with 20GB and 100GB inputs
 plot_cache.plot_cache_v2()
 
+# ======== Multi-threaded experiment results ==========
 os.chdir("../multi")
 import result.multi.process_result as multi_result
+# Generate figures of experiment with local file system
 multi_result.result_local(rep_no=5)
+# Generate figures of experiment with NFS
 multi_result.result_nfs(rep_no=5)

### 2020-12-08 result/multi/process_result.py
Not show figure after generating

@@ -119,7 +119,7 @@ def result_local(rep_no=1):
     plt.savefig("figures/multi_local.pdf", format="pdf")
     plt.savefig("figures/multi_local.svg", format="svg")
 
-    plt.show()
+    # plt.show()
 
 
 def result_nfs(rep_no=1):
@@ -141,5 +141,30 @@ def result_nfs(rep_no=1):
     plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.3)
     plt.savefig("figures/multi_nfs.pdf", format="pdf")
     plt.savefig("figures/multi_nfs.svg", format="svg")
+
+    # plt.show()
+
+
+def run_time():
+    local_org_df = pd.read_csv("local/wrench/run_time_original.csv")
+    local_cache_df = pd.read_csv("local/wrench/run_time_pagecache.csv")
+    nfs_org_df = pd.read_csv("nfs/wrench/run_time_original.csv")
+    nfs_cache_df = pd.read_csv("nfs/wrench/run_time_pagecache.csv")
+
+    fig, (ax1, ax2) = plt.subplots(figsize=(10,4), ncols=2, nrows=1)
+
+    ax1.set_ylim(top=2)
+    ax1.plot(local_org_df["no_pipeline"], local_org_df["run_time"], label = "WRENCH")
+    ax1.plot(local_cache_df["no_pipeline"], local_cache_df["run_time"],label = "WRENCH-cache")
+    ax1.set_title("Local file system")
+
+    ax2.set_ylim(top=2)
+    ax2.plot(nfs_org_df["no_pipeline"], nfs_org_df["run_time"], label = "WRENCH")
+    ax2.plot(nfs_cache_df["no_pipeline"], nfs_cache_df["run_time"],label = "WRENCH-cache")
+    ax2.set_title("NFS")
+
+    plt.xlabel("number of pipelines")
+    plt.ylabel("time (seconds)")
+    plt.legend()
 
     plt.show()

### 2020-12-08 result/single/plot_cache.py

@@ -61,4 +61,4 @@ def plot_cache_v2():
     plt.savefig("figures/cached_files.svg", format="svg")
     plt.savefig("figures/cached_files.pdf", format="pdf")
 
-    plt.show()
+    # plt.show()

### 2020-12-08 result/single/plot_error.py

@@ -48,4 +48,4 @@ def plot_error():
     plt.savefig("figures/single_errors.svg", format="svg", bbox_extra_artists=(lgd,), bbox_inches='tight')
     plt.savefig("figures/single_errors.pdf", format="pdf", bbox_extra_artists=(lgd,), bbox_inches='tight')
     plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.3)
-    plt.show()
+    # plt.show()

### 2020-12-08 result/single/plot_memprof.py

@@ -156,4 +156,4 @@ def compare_plot(input_sizes=[20, 100], makespan=[200, 1300]):
 
     plt.savefig("figures/single_memprof.svg", format="svg")
     plt.savefig("figures/single_memprof.pdf", format="pdf")
-    plt.show()
+    # plt.show()

### 2020-12-09 paper_io_sim.tex
fixed figure 4

--- 2020-12-08 paper_io_sim.tex
@@
 
-            Our first experiment simply runs one instance of the
+            Our first experiment (\textbf{Exp single-threaded}) simply runs one instance of the
             application on a single cluster node, with different input file
@@
 
-            Our second experiment runs concurrent instances of the
+            Our second experiment (\textbf{Exp concurrent}) runs concurrent instances of the
             application on a single node, all application instances
@@
 
-            Our third experiment occurs in the same configuration as the
+            Our third experiment (\textbf{Exp NFS}) occurs in the same configuration as the
             previous one, albeit reading and writing on a 400-GB NFS-mounted remote disk
@@
         \subsection{Single-threaded experiment}
+
+        Figure~\ref{fig:single_error} shows the relative error of simulated
+        read and write times for the three successive application tasks in \textbf{Exp single-threaded}. The
+        page cache simulation model drastically reduces simulation errors
+        for all I/O operations except the first read as it involves
+        uncached data. Errors are reduced from an average of 345\%
+        in the original WRENCH to 46\% in the Python prototype and 39\% in WRENCH-cache.
+        Unsurprisingly, the original WRENCH simulator significantly overestimates
+        read and write times, due to its lack of page cache simulation
+        model.
+
+        Especially, the simulation time errors of \textit{Read 3} with 100 GB of input
+        in the simulators with page cache are just under the error in original WRENCH.
+        This longer read can be explained with the amount of cached data of files
+        at the beginning of the read in reality and simulators shown in
+        Fig.~\ref{fig:single_cache}.
+        % In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
+        % output of \textit{Write $<$i-1$>$}.
+        With the input size of 20 GB, the amount of cached data of each files after 
+        every I/O operation is similar since the page cache can keep all these files.
+%             \tristan{what does this refer to?
+%             For 20 GB there is still some error in WRENCH-cache, what do you mean 
+%``there is no difference in the page cache''?}.
+%   Dzung: Some text was lost in a merge, I added it again.
+        With the input of 100 GB, the difference between real execution and WRENCH-cache
+        simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
+        In real execution, after being written by \textit{Write 2},
+        \textit{file 3} is entirely in cache.
+        Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
+        However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
+        cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
+        \textit{file 3} from disk.
+        The reason for the difference is in our WRENCH-cache simulator,
+        with 100 GB of input, while \textit{file 3} is being written,
+        the page cache is saturated, eviction is then evoked to evict data
+        from page cache and make space available for \textit{file 3}.
+        However, because the evicted amount is less than required,
+        some \textit{file 3} data is evicted after being written, make \textit{file 3}
+        partially cached.
+        As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
+        with memory bandwidth, a part of it is read from disk with disk read bandwidth,
+        which is many times slower, leading to longer read time.
 
@@
                    \includegraphics[width=\linewidth]{result/single/figures/single_errors.pdf}
+                   \vspace*{-0.7cm}
+                   \caption{Simulation errors}
+                   \vspace*{0.5cm}
+                   \label{fig:single_error}
                 \end{subfigure}
-            \caption{Simulation errors with different input file sizes}
-            \label{fig:single_error}
             \begin{subfigure}{\linewidth}
                 \centering
+                %    Gray shades represent task phases (read, compute and write). 
+                %    Lines represent memory usage along pipeline execution time.}
                    \includegraphics[width=\linewidth]{result/single/figures/single_memprof.pdf}
+                   \vspace*{-0.7cm}
+                   \caption{Memory profiles}
+                   \vspace*{0.5cm}
+                   \label{fig:single_memprof}
             \end{subfigure}
-            \caption{Memory profiling results with different input file sizes.
-            The bars in the figures present activities in tasks (read, compute and write),
-            the lines describe memory status throughout run time of the pipeline.}
-            \label{fig:single_memprof}
             \begin{subfigure}{\linewidth}
                 \centering
-                   \includegraphics[width=\linewidth]{result/single/figures/cached_files.pdf}
-            \end{subfigure}
+                   \includegraphics[width=1.05\linewidth]{result/single/figures/cached_files.pdf}
+                   \caption{Cache contents}
+                %    \textcolor{red}{Update real results of 20GB}}
                    \label{fig:single_cache}
-               \caption{Amount of file data in cache after each I/O operation. 
-               \textcolor{red}{Update real results of 20GB}}
+            \end{subfigure}\\
+            \caption{Single-threaded results}
             \end{figure*}

### 2020-12-09 result/single/plot_cache.py

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 from matplotlib.patches import Patch
 
 
-def plot_cache(ax, real_file, sim_file, patterns, color_real="gainsboro", color_wrench="tab:cyan"):
+def plot_cache(ax, real_file, sim_file, patterns, ylabel, color_real="#90C987", color_wrench="#1965B0"):
     index = ["file_1", "file_2", "file_3", "file_4"]
     columns = ["Read 1", "Write 1", "Read 2", "Write 2", "Read 3", "Write 3"]
 
@@ -13,8 +13,8 @@ def plot_cache(ax, real_file, sim_file, patterns, color_real="gainsboro", color_
     df_real[index].plot.bar(stacked=True, rot=0, color=color_real, edgecolor="k", ax=ax, position=1, width=0.3, legend=False, linewidth=0.5)
     df_sim[index].plot.bar(stacked=True, rot=0, color=color_wrench, edgecolor="k", ax=ax, position=0, width=0.3, legend=False, linewidth=0.5)
 
-    ax.set_xlabel("tasks")
-    ax.set_ylabel("amount (GB)")
+    if ylabel:
+        ax.set_ylabel("amount (GB)")
     ax.set_xticklabels(columns)
 
     bars = ax.patches
@@ -28,6 +28,8 @@ def plot_cache(ax, real_file, sim_file, patterns, color_real="gainsboro", color_
     for bar, hatch in zip(bars, hatches):  # loop over bars and hatches to set hatches in correct order
         bar.set_hatch(hatch)
 
+    x_axis = ax.xaxis
+    x_axis.label.set_visible(False)
     ax.set_xlim(left=-0.5)
     ax.set_ylim(top=180)
 
@@ -37,15 +39,15 @@ def plot_cache_v2():
     files = ["file 1", "file 2", "file 3", "file 4"]
     envs = ["Real execution", "WRENCH-cache"]
     patterns = ["", "...", "///", "\\\\\\"]
-    color_real = "gainsboro"
-    color_wrench = "tab:cyan"
+    color_real = "#90C987"
+    color_wrench = "#1965B0"
 
-    fig, (ax1, ax2) = plt.subplots(figsize=(11, 4), ncols=2, nrows=1)
-    plt.subplots_adjust(left=0.1, bottom=0.15, right=0.95, top=0.8, wspace=0.25)
[…]

### 2020-12-09 result/single/plot_error.py

@@ -19,7 +19,7 @@ def grouped_bar_chart(ax, labels, xlabel, ylabel, *argv):
     ax.set_ylim(bottom=0, top=700)
 
 
-def plot_task_error(ax, size):
+def plot_task_error(ax, size, ylabel):
     labels = ["Read 1", "Write 1", "Read 2", "Write 2", "Read 3", "Write 3"]
 
     # atop_file =     "log/cluster/100gb/atop_mem.log"
@@ -32,20 +32,25 @@ def plot_task_error(ax, size):
     wrench_org_error = [item * 100 for item in evaluate.task_time_error(real_time_log, wrench_org_log)]
     wrench_ext_error = [item * 100 for item in evaluate.task_time_error(real_time_log, wrench_ext_log)]
 
-    grouped_bar_chart(ax, labels, "", "error (%)",
-                      ("Python simulator", py_error, "tab:pink"), ("WRENCH", wrench_org_error, 'tab:orange'),
-                      ("WRENCH-cache", wrench_ext_error, 'tab:cyan'))
+    if ylabel:
+        ylabel = "error (%)"
+    else:
+        ylabel = ""
+    grouped_bar_chart(ax, labels, "", ylabel,
+                      ("Python prototype", py_error, "#994F88"), ("WRENCH", wrench_org_error, '#7BAFDE'),
+                      ("WRENCH-cache", wrench_ext_error, '#1965B0'))
     ax.set_title("%d GB" % size)
 
 def plot_error():
-    plt.rcParams.update({'font.size': 7})
-    fig, (ax1, ax2) = plt.subplots(figsize=(10, 2.2), ncols=2, nrows=1)
-    plot_task_error(ax1, 20)
-    plot_task_error(ax2, 100)
+    fig, (ax1, ax2) = plt.subplots(figsize=(11, 3), ncols=2, nrows=1)
+    plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.8, wspace=0.1)
+    plt.rcParams.update({'font.size': 9})
 
-    lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.35), ncol=3)
+    plot_task_error(ax1, 20, ylabel=True)
+    plot_task_error(ax2, 100, ylabel=False)
+
+    lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.12, 1.25), ncol=3)
 
[…]

### 2020-12-09 result/single/plot_memprof.py

@@ -10,21 +10,21 @@ def compare_size(axes, real_time_file, real_mem_file,
                  real_cond=None, py_cond=None, wrench_cond=None):
     # REAL RESULTS
     real_subplot(axes[0], real_time_file, real_mem_file, xmin, xmax, ymin, ymax, bar_alpha=0.2, linestyle="-",
-                 linewidth=1, cond_text=real_cond)
+                 linewidth=3, cond_text=real_cond)
 
     # SIMULATION RESULTS
-    sim_subplot(axes[1], pysim_time_file, pysim_mem_file, "Python simulator", bar_alpha=0.2, cond_text=py_cond, color='deeppink')
-    sim_subplot(axes[2], simgrid_time_file, simgrid_mem_file, "WRENCH-cache simulator", bar_alpha=0.2,
+    sim_subplot(axes[1], pysim_time_file, pysim_mem_file, "Python prototype", bar_alpha=0.2, cond_text=py_cond, color='deeppink')
+    sim_subplot(axes[2], simgrid_time_file, simgrid_mem_file, "WRENCH-cache", bar_alpha=0.2,
                 cond_text=wrench_cond, color='darkcyan')
 
     for ax in axes:
         ax.set_xlim(right=xmax, left=xmin)
 
-    axes[0].text(0.45 * xmax, 330, "%d GB" % size, fontsize=16)
+    axes[0].text(0.45 * xmax, 300, "%d GB" % size, fontsize=14)
 
 
 def real_subplot(subplot_ax, real_time_file, real_mem_file, xmin, xmax, ymin, ymax,
-                 bar_alpha=0.2, line_alpha=1, linestyle=".-", linewidth=1.5, cond_text=None):
+                 bar_alpha=0.2, line_alpha=1, linestyle=".-", linewidth=2, cond_text=None):
     timestamps = log_parse.read_timelog(real_time_file, skip_header=False)
     atop_log = log_parse.read_atop_log(real_mem_file, dirty_ratio=0.4, dirty_bg_ratio=0.1)
     dirty_data = np.array(atop_log["total"])
@@ -44,25 +44,26 @@ def real_subplot(subplot_ax, real_time_file, real_mem_file, xmin, xmax, ymin, ym
                                label="write" if i == 1 else "")
 
     # app_cache = list(np.array(app_mem) + np.array(cache_used))
-    subplot_ax.plot(time, atop_log["total"], color='k', linewidth=1, linestyle=":", label="Total memory",
[…]

### 2020-12-09 result/single/plot_cache.py
fixed edge color

@@ -10,8 +10,10 @@ def plot_cache(ax, real_file, sim_file, patterns, ylabel, color_real="#90C987", 
     df_real = pd.read_csv(real_file, index_col="task")
     df_sim = pd.read_csv(sim_file, index_col="task")
 
-    df_real[index].plot.bar(stacked=True, rot=0, color=color_real, edgecolor="k", ax=ax, position=1, width=0.3, legend=False, linewidth=0.5)
-    df_sim[index].plot.bar(stacked=True, rot=0, color=color_wrench, edgecolor="k", ax=ax, position=0, width=0.3, legend=False, linewidth=0.5)
+    # plt.rcParams.update({'hatch.color': 'green'})
+    df_real[index].plot.bar(stacked=True, rot=0, edgecolor='green', color=color_real, ax=ax, position=1, width=0.3, legend=False, linewidth=0.5)
+    # plt.rcParams.update({'hatch.color': 'blue'})
+    df_sim[index].plot.bar(stacked=True, rot=0, edgecolor='blue', color=color_wrench, ax=ax, position=0, width=0.3, legend=False, linewidth=0.5)
 
     if ylabel:
         ax.set_ylabel("amount (GB)")

### 2020-12-09 paper_io_sim.tex
fixes to intro and related work

--- 2020-12-08 paper_io_sim.tex
@@
 
-        The Linux Page Cache, introduced in kernel version 2.4, plays an important role in
-        the reduction of filesystem data transfer costs~\cite{pagecache}. With the page cache, previously
+        The Linux Page Cache plays an important role in
+        the reduction of filesystem data transfer costs. With the page cache, previously
         read data can be reread directly from memory and written data can be written to
@@
         costs, unstable environments, and the difficulties in producing reproducible results.
-        Simulations address these concerns by providing abstractions to the performance of computer
-        hardware, such as CPU, network and storage. As a result, simulations provide and
+        Simulations address these concerns by providing abstractions for the performance of computer
+        hardware, such as CPU, network and storage. Consequently, simulations provide and
         cost-effective, fast, easy and reproducible  way evaluate application performance
@@
 
+        The amount of data-intensive applications has been on the rise as a result of
+        open-data and data sharing initiatives. Due to the sheer size of the data being
+        processed, these applications must be executed on large scale infrastructure
+        such as High Performance Computing (HPC) clusters or the cloud. Determining
+        which type of infrastructure is the most cost-effective for the particular
+        application can be tricky without the help of a simulator.
+
         Page cache simulation is crucial when simulating data intensive applications,
-        which would rely on cache hits and writing to memory for additional performance.
+        which rely on cache hits and writing to memory for additional performance.
+        Furthermore, simulating the page cache enables us to study important application-I/O
+        bottlenecks which can arise over the course of execution, such as page cache
+        starvation~\cite{zhuang2017}.
         While existing simulators can adequately characterize many different computing environments,
@@
 
-        In this paper, we present WRENCH-cache, a simulation model of I/O with page
-        cache incorporated into WRENCH, a workflow management system simulator based of
-        the popular SimGrid framework. The goal of WRENCH-cache is to provide users
+        In this paper, we present WRENCH-Cache, a simulation model of I/O with page
+        cache incorporated into WRENCH~\cite{casanova2018wrench}, a workflow management system simulator based on
+        the popular SimGrid framework. The goal of WRENCH-Cache is to provide users
         with an accurate representation of a workflow execution, when file system page
-        cache is used. WRENCH-cache is able to simulate both writeback and writethrough
+        cache is used, while providing a good balance between accuracy and scalability.
+        WRENCH-Cache is able to simulate both writeback and writethrough
         cache which would normally be available to local and NFS-based filesystems.
+        WRENCH-Cache can provide accurate simulations for large-scale parallel
+        applications using the page cache.
 
@@
 
-            Linux implements a disk cache called \textit{page cache}. 
-            It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
-            The goal of the page cache is to avoid disk I/O by keeping data in memory 
-            for later accesses, minimize I/O time since disk accesses are considerably 
-            slower than memory accesses.
-            When the kernel starts a read operation, it checks if required data is in page cache.
-            If yes, data is read from memory. This is called a \textit{cache hit}. 
-            Otherwise, data is read from disk and placed in page cache.
-            For writing, page cache is implemented with three different strategies. 
-            The first is \textit{direct~I/O} or \textit{no-write}, which means the page cache 
-            is not involved in write operations. 
-            The second is \textit{write-through}, in which data goes \textit{through} 
-            page cache to disk. In this case, data is written to both page cache and disk. 
-            In the third strategy, called \textit{writeback}, data are written directly 
-            to disk as \textit{dirty} pages and these pages will be periodically flushed 
-            to disk later by a \textit{writeback} process and then marked as not dirty. 
+        To improve performance of disk accesses, the page cache was introduced to the Linux kernel. %It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
+            The goal of the page cache is to try ensure that most I/O operations occur entirely
+            within memory space rather than synchronously to and from the storage device.
+            All disk-backed data located in the page cache are partitioned into many small pages.
+            When a read operation is initiated, the kernel checks whether the required pages are in memory.
+            If it is the case, data is read entirely from memory (i.e. \textit{cache hit}). Otherwise,
+            the necessary pages must be loaded to memory from the disk (i.e. \textit{cache miss})
+            Conversely, for writing, one of three strategies can be applied: direct~I/O (or, ``no-write''),
+            writethrough and writeback.
+            If direct~I/O is enabled for a given file system, all I/O operations bypass the
+            page cache for writes, writing data directly to disk. In contrast, write-through enables the page cache
+            for file system writes, but the data is written synchronously from memory to disk. Writeback extends write-through
+            by enabling asynchronous writes to disk, meaning that a write call will return once
+            the data has been written to memory.
             The writeback strategy is considered to outperform write-through as well as 
-            direct I/O as it delays disk writes and perform a bulk write later.
+            direct I/O as it delays disk writes to perform a bulk write at a later time.
             \cite{linuxdev3rd2010}.
 
-            In writeback, one of the keys is the cache eviction mechanism, which is
-            triggered when available memory is insufficient to removes pages 
-            from page cache. 
-            Because dirty pages are data written to page cache but yet to be 
-            written to disk, dirty pages must be kept in page cache and only clean pages 
-            are available for eviction.
-            If the clean pages are inadequate, dirty pages are flushed to disk, 
-            marked as clean and available for eviction. 
-            There is another type of flushing mechanism, which is 
-            \textit{periodical flushing}, called periodically in predefined intervals. 
+            Cache eviction and flushing strategies are integral to proper page cache functioning.
+            Whenever space in memory becomes limited, either as a result of application memory
+            or page cache use, page cache data may be evicted. Only data that 
+            has been persisted to storage (clean pages) can be flagged for eviction and removed from
+            memory. Written data that has not yet been persisted to disk (dirty data) must first
+            be copied (flushed) to storage prior to eviction. When sufficient memory is
+            being occupied, the flushing process is synchronous. However, even when 
+            there is sufficient available memory, written data will be flushed to disk
+            at a predefined interval through a process known as \textit{periodical flushing}.
             Periodical flushing only flushes expired dirty pages, which remain dirty in 
@@
             \cite{owda2014comparison}.
-            In Linux kernel, a \textit{two lists strategy (LRU/2)} is implemented based on 
-            the least-recently-used (LRU) scheme, with an \textit{active list} and 
-            an \textit{inactive list}. 
+            The Linux kernel uses a two-list strategy to flag pages for eviction.
+            The two-list strategy is based on a least recently used (LRU) policy
+            and uses an active and inactive list in its implementation.
             If accessed pages are not in the page cache, they are added to the inactive list. 
-            If pages are accessed when residing in the inactive list, they will be moved from 
-            inactive list to active list. 
+            Should pages located on the inactive list be accessed, they will be moved from 
+            the inactive to the active list. 
             The lists are also kept balanced by moving pages from the active list
@@
             and not evictable, while the inactive list includes pages accessed once only, 
-            or pages accessed more than once but moved from the active list. 
-            The lists work in LRU manner, pages are added to the tail and removed 
-            from the top.
+            or pages that have be accessed more than once but moved from the active list. 
+            Both lists operate using LRU eviction policies meaning the data that has
+            not be accessed recently will be moved first.
 
@@
 
-            Many simulation frameworks and applications have been developed 
-            in recent years to support studies in grids, cloud, HPC, peer-to-peer 
-            networks, volunteer computers.                     
-            These simulators are built with models of resources such as CPU, network, 
-            storage with different approaches and levels of details in different simulators. 
-            The detail level of the model causes a conflict between accuracy and scalability, 
-            and these simulators usually trade of one for the other 
-            \cite{casanova2014simgrid}. 
-            For example, some simulators are built with packet-level network models 
-            such as \textit{CloudSim} and \textit{iCanCloud}, or storage model that 
-            simulates the operation of storage hardware in \textit{DiskSim} 
-            \cite{casanova2014simgrid}. 
+            Many simulation frameworks exist to describe the behaviours of applications
+            running on the various infrastructures available to users, such as personal computers,
+            clouds and clusters.
+            These simulators are built with resource models such CPU, network and 
+            storage, using various different approaches to describe the hardware.
+            For example, there is \textit{CloudSim} and \textit{iCanCloud} for packet-level network models 
+            and \textit{DiskSim} the simulates storage hardware functioning.
+
+            Simulators must choose between scalability and accuracy when describing the
+            hardware.\cite{casanova2014simgrid}. 
             Fine-grained models allow simulators to achieve high accuracy, 
-            but at the same time lead to scalability issues due to long simulation time 
+            but lead to scalability issues due to long simulation time 
             when simulated platforms and applications grow. 
@@
             not only to balance the trade-off between accuracy and scalability, 
-            but also to achieve versatility, which allows the use of the framework 
-            to simulate applications in multiple domains \cite{casanova2014simgrid}.
+            but also to achieve versatility, allowing the use of the framework 
+            to simulate applications in multiple domains. WRENCH, a workflow management
+            system simulator based on the SimGrid framework, provides a high-level API
+            for simulating workflows, thus facilitating the integration of new models.
 
@@
             accessed on disk in a block cache \cite{nunez2012simcan}. 
-            Page cache is also modeled in iCanCloud simulator with a component in charge 
-            of managing memory accesses and cached data \cite{nunez2012icancloud}, 
-            but one of the issues is scalability due to its sophisticated model.
-            Ultimately, non of these simulators provide any cache write strategy, 
-            cache eviction mechanism with dirty/clean data and the page cache LRU lists. 
-            Cache replacement policies is applied in \cite{xu2018saving} to simulate 
-            in-memory caching, but for the study in energy consumption of multi-tier 
-            heterogenous network.
+            Page cache is also modeled in the iCanCloud simulator through a component which
+            manages memory accesses and cached data \cite{nunez2012icancloud}. However,
+            the scalability of the iCanCloud simulator is limited by its sophisticated model.
+            Ultimately, none of these simulators provide any writeback cache simulator nor
+            cache eviction policys through LRU lists.
+            Although cache replacement policies are applied in \cite{xu2018saving} to simulate 
+            in-memory caching, their simulator is specific to energy consumption of multi-tier 
+            heterogenous networks.
 
-            In this study, we chose SimGrid to implement our model of I/O with page cache 
-            due to the balancing between accuracy and scalibility, the versatility allowing it 
-            to be adopted in different domains. 
-            In addition, it is an actively maintained open-scource framework with 
-            modular design, which enables the framework to be extended with new models. 
-            The main issue users have to face when using SimGrid is everything has to be 
-            implemented from scratch due to its low-level API. 
-            To ease this challenge, WRENCH, another open-source library, is built on top 
-            of SimGrid to abstract the low-level logic and expose a higher abstraction 
-            level API \cite{casanova2018wrench}. 
-            We integrate our model in WRENCH and use the SimGrid API to interact with 
-            simulated storage resource.
-            This approach allows us to leverage existing features in WRENCH and SimGrid.
-            Firstly, we can take advantage of the process control, synchronization of 
-            resource consumption by concurrent processes provided by the simulation core 
-            in SimGrid. 
-            Secondly, we can reuse the existing SimGrid storage model with capacity 
+            In this study, we use WRENCH 
+            to implement our page cache model. We chose WRENCH due to accurate yet scalable implementation, resulting
+            from its SimGrid backend, as well as its high-level API.
+            We have integrated our model in WRENCH and use the SimGrid API to interact with 
+            simulated storage resources.
+            This approach allows us to leverage existing features in WRENCH and SimGrid such as process control and synchronization of 
+            resource consumption.
+            We also leverage the existing SimGrid storage model with capacity 
             and bandwidth properties, which are similar to the characteristics of memory.
-            And finally, WRENCH provides a high level abstraction, which help us drop 
-            the burden of implementation.
 
@@
     In this section, we present our page cache simulation model for file I/O,
-    flusing, and eviction. We implement this model in two simulators:
-    (1) a standalone Python prototype, and (2) WRENCH-cache, an extension of the WRENCH library.
+    flusing, and eviction. We detail the design and implementation of two simulators:
+    (1) a standalone Python prototype to evaluate the simulation model independently of any other 
+    tool, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
     We also describe experimental scenarios to evaluate our simulators in various configurations.
@@
     To read or write a file chunk, a simulated application sends a
-    request to the IO Controller. The IO Controller interacts as needed with
+    request to the IOController. The IOController interacts as needed with
     the Memory Manager to free memory through flushing or eviction,
@@
            \includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
-           \caption{Overview of the simulated writeback page cache \tristan{does it simulate only writeback? See comment in NFS section.}.
+           \caption{Overview of the simulated writeback page cache.
            Applications send file read or write requests to the
-           IO Controller that orchestrates flushing, eviction, cache
-           and disk accesses with the Memory Manager. Concurrent accesses to storage
-           devices (memory and disk) are simulated using existing models.}
+           IOController that orchestrates flushing, eviction, cache
+           and disk accesses with the MemoryManager. Storage
+           devices (memory and disk) simulate file transfers with
+           bandwidth sharing.}
            \label{fig:interaction}
@@
     expired dirty data in page cache LRU lists and flushes them to disk. We
-    use existing storage simulation models~\cite{lebre2015} to simulate disk and
+    use existing storage simulation models~\cite{lebre2015} to simulate
     memory, characterized by its storage capacity, read and write
     bandwidths, and latency. Reusing existing storage models allows us to
-    simulate bandwidth sharing between concurrent memory accesses and concurrent disk accesses.
+    simulate bandwidth sharing between concurrent memory accesses.
 
@@
     two lists of data blocks, an active list and an inactive list, both ordered by 
-    last access time (earliest first, see Figure~\ref{fig:lrulist}).
+    last access time (earliest first, Figure~\ref{fig:lrulist}).
     As in the kernel, our simulator limits the size of the active list to
@@
 
-    \subsubsection{Reads and writes}
+    \subsubsection{Reads and Writes}
 
@@
     before data from the active list
-    (data reads occur from left to right in Figure~\ref{fig:read_order}).
+    (left to right in Figure~\ref{fig:read_order}).
     When a chunk of \emph{uncached} data is read, a new clean block is created 
-    and appended to the inactive list. 
+    and put at the top of the inactive list. 
     When a chunk of \emph{cached} data is read, one or more existing data blocks in the LRU lists are accessed.
     If these blocks are clean, we merge them together, update the access time and size of the resulting block, 
-    and append it to the active list. 
-    If the blocks are dirty, we move them independently to the active list, to preserve their entry time. 
-    Because the chunk and block sizes may be different, there are situations
+    and put it in the active list. 
+    If the blocks are dirty, we move them to the active list to preserve their entry time. 
+    Because the chunk size and block size can be different, there are situations
     where a block is not entirely read. 
@@
            \caption{File data read order. Data is read from left to right: uncached data 
-           is read first, followed by data from the inactive list, and finally data from the active list. \tristan{Colors 
-           shouldn't match that in Fig 1 as they dont represent the same thing.}}
+           is read first, followed by data from the inactive list, and finally data from the active list.}
            \label{fig:read_order}
@@
     uncached. Thus, each time a chunk is written, we create a block of dirty data 
-    and append it to the inactive list.
+    and add it to the inactive list.
 
@@
 
-    The main simulated thread in the Memory Manager can flush or evict data from the
-    memory cache. The data flushing simulation
+    The main simulated thread in the Memory Manager can flush the
+    cache to write dirty data to disk. The data flushing simulation
     function takes the amount of data to flush as parameter. While
@@
     inactive list, then the sorted active list, and writes the
-    least recently used dirty block to disk, having set its dirty
+    least recently used dirty block to disk having set its dirty
     flag to 0. In case the amount of data to flush requires that a
@@
     one that is flushed and one that remains dirty. Flushing time
-    to disk is simulated by the storage model.
+    to disk is simulated with the storage model.
 
-    The cache eviction simulation also runs in
+    Similarly to data flushing, our cache eviction simulation runs in
     the main thread. It frees up the page cache by traversing and deleting 
@@
     The amount of data to evict is passed as a parameter and data blocks are deleted 
-    from the inactive list until the evicted data reaches the required amount,
+    from the inactive list until total evicted data reaches the required amount,
     or until there is no clean block left in the list.
@@
     Because periodical flushing is simulated as a background thread, it can happen concurrently
-    with disk I/O initiated by the main thread. This is taken into account by the 
+    with disk I/O initiated by the main thread. This parallelism is handled by the 
     storage model and reflected in simulated I/O time.
@@
 
-    \begin{algorithm}\caption{File chunk read simulation of IO Controller}
+    \begin{algorithm}\caption{File chunk read simulation of IOController}
     \label{alg:read}
@@
                \EndInput
-               \State disk\_read = min(cs, fs - mm.cached(fn)) \Comment{To be read from disk} 
-               \State cache\_read = cs - disk\_read \Comment{To be read from cache} 
+               \State disk\_read = min(cs, fs - mm.cached(fn))
+               \State cache\_read = cs - disk\_read
             \State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
@@
     round-robin fashion. To read a file chunk, simulated applications send
-    chunk read requests to the IO Controller which processes them using
+    chunk read requests to the IOController which processes them using
     Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
@@
     to flush dirty data (line 9). If necessary, flushing is complemented by
-    eviction (line 10). Note that, when called with negative arguments, functions
-    \texttt{flush} and \texttt{evict} simply return and do not do anything. Then, if the block requires
-    uncached data, a disk read is simulated with the disk model (line 12)
+    eviction (line 10). When called with negative arguments, functions
+    flush and evict simply return and do not do anything. Then, if there is
+    uncached data, a disk read is simulated with disk model (line 12)
     \tristan{this is not consistent with the new Figure 1 where disk reads
-    are done by the mm}, and the data read is added to cache (line
+    are done by the mm}, and this data of the file is added to cache (line
     13).
@@
 
-    \begin{algorithm}\caption{File chunk write simulation of IO Controller}
+    \begin{algorithm}\caption{File chunk write simulation of IOController}
     \label{alg:write}
@@
                 \State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem)) \tristan{why is data written to cache with "addtocache" here and with "write" at line 10?}
-                \State sm.write(cs - mem\_amt) \tristan{According to Fig 1 the IO Controller is not supposed to interact directly with sm}
+                \State sm.write(cs - mem\_amt) \tristan{According to Fig 1 the IOController is not supposed to interact directly with sm}
             \EndIf
@@
     Algorithm~\ref{alg:write} describes our simulation of chunk writes in 
-    the IO Controller. 
+    the IOController. 
     % Data is written to cache until the amount of dirty data in the cache 
@@
     If this amount is greater than 0, the Memory Manager is requested to evict 
-    data from cache if necessary (line 8) \tristan{the difference between available and free memory should be explained the first time it is referred}.
+    data from cache if necessary (line 8) \tristan{the difference between available and free memory should be explained somewhere}.
     After eviction, the amount of data that can be written to 
@@
             We also implemented our model in WRENCH, separately from existing WRENCH components. 
-            We exposed page cache simulation through a user API, activated with a command-line argument.
+            Page cache simulation is available through a user API and can be activated using a command-line argument.
             We used SimGrid 3.25 and WRENCH 1.6. SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid},
@@
 
-        Our experiments compare real executions with our Python prototype,
-        with the original WRENCH, and with our WRENCH-cache extension. We
+        Our experiments compare real executions with the original WRENCH,
+        with our WRENCH-cache extension, and with our prototype Python simulator. We
         evaluate our page cache simulation in single-threaded and
@@
 
-        We simulate an application that consists of three single-core, sequential tasks
+        We simulate an application with three single-core, sequential tasks
         where each task reads the file produced by the previous task,
@@
         memory, local disk, remote disk (NFS), and network
-        (Table~\ref{table:benchmark}). Since WRENCH currently only supports
+        (Table~\ref{table:benchmark}). Since SimGrid only supports
         symmetrical bandwidths, we use the mean of the read and write
@@
             sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed
-            to the same local disk. The goal is to validate the simulation model 
-            in absence of any concurrent disk access.
+            to the same local disk.
 
@@
             Our third experiment occurs in the same configuration as the
-            previous one, albeit reading and writing on a 400-GB remote NFS disk
+            previous one, albeit on a 400-GB NFS-mounted remote disk
             mounted from another compute node. \tristan{As is commonly
@@
             writing data directly to disk and adding them to the server
-            cache \tristan{it it part of the IO Controller? it would deserve
+            cache \tristan{it it part of the IOController? it would deserve
             a few sentences in the methods.}.
@@
         \subsection{Single-threaded experiment}
-
-        Figure~\ref{fig:single_error} shows the relative error of simulated
-        read and write times throughout the three application tasks. The
-        page cache simulation model drastically reduces simulation errors
-        for all I/O operations except the first read as it only involves
-        uncached data. Errors are reduced from an average of \texttt{xxx\%}
-        in WRENCH to \texttt{yyy \%} in the Python prototype and
-        \texttt{zzz \%} in WRENCH-cache \tristan{Dzung, update average percentages}.
-        Unsurprisingly, the original WRENCH simulator significantly overestimates
-        read and write times, due to its lack of page cache simulation
-        model.
-
-        % In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
-        % output of \textit{Write $<$i-1$>$}.
-        With the input size of 20 GB, in both real execution and WRENCH-cache
-        simulator, there is no difference in the page cache since all files in the
-        pipeline can be entirely cached \tristan{what does this refer to? 
-        For 20 GB there is still some error in WRENCH-cache, what do you mean ``there is no difference in the page cache''?}.
-
-        With the input of 100 GB, the difference between real execution and WRENCH-cache
-        simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
-        In real execution, after being written by \textit{Write 2},
-        \textit{file 3} is entirely in cache.
-        Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
-        However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
-        cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
-        \textit{file 3} from disk.
-        The reason for the difference is in our WRENCH-cache simulator,
-        with 100 GB of input, while \textit{file 3} is being written,
-        the page cache is saturated, eviction is then evoked to evict data
-        from page cache and make space available for \textit{file 3}.
-        However, because the evicted amount is less than required,
-        some \textit{file 3} data is evicted after being written, make \textit{file 3}
-        partially cached.
-        As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
-        with memory bandwidth, a part of it is read from disk with disk read bandwidth,
-        which is many times slower, leading to longer read time.
 
@@
 
+			The I/O time of tasks in simulators and real execution is also tracked and compared.
+			Fig.~\ref{fig:single_error} shows the relative error of simulated I/O time
+			of individual I/O activities.
+			Our pipeline consists of 3 tasks, and each task reads an input file,
+			performs some computation, and writes output file.
+			Thus, there are 6 sequential I/O activities, \textit{Read 1} and \textit{Write 1}
+			from the first task, \textit{Read 2} and \textit{Write 2} the second task,
+			and \textit{Read 3}, \textit{Write3} from the last task.
+
 			\begin{figure*}
 			\centering
-            \begin{subfigure}{\linewidth}
-                \centering
-                   \includegraphics[width=\linewidth]{result/single/figures/single_errors.pdf}
-            \end{subfigure}
-            \caption{Simulation errors with different input file sizes}
-            \label{fig:single_error}
 			\begin{subfigure}{\linewidth}
@@
 				\centering
+   				\includegraphics[width=\linewidth]{result/single/figures/single_errors.pdf}
+			\end{subfigure}
+			\caption{Simulation errors with different input file sizes}
+			\label{fig:single_error}
+			\begin{subfigure}{\linewidth}
+				\centering
    				\includegraphics[width=\linewidth]{result/single/figures/cached_files.pdf}
@@
 			\end{figure*}
+
+			In general, with all input sizes in the experiment, the simulation errors
+			of simulators with page cache model (Python and WRENCH-cache simulators)
+			are remarkably lower than the result with original WRENCH .
+			The reason is, in original WRENCH, because page cache is not simulated,
+			data is only read and written with disk bandwidth, which is much slower
+			than memory bandwidths, leading to simulated I/O time longer than in reality.
+			Especially, the simulation time errors of \textit{Read 3} with 100 GB of input
+			in the simulators with page cache are just under the error in original WRENCH.
+			This longer read can be explained with the amount of cached data of files
+			at the beginning of the read in reality and simulators shown in
+			Fig.~\ref{fig:single_cache}.
+			In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
+			output of \textit{Write $<$i-1$>$}.
+			With the input size of 20 GB, in both real execution and WRENCH-cache
+			simulator, there is no difference in the page cache since all files in the
+			pipeline can be entirely cached.
+			With the input of 100 GB, the difference between real execution and WRENCH-cache
+			simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
+			In real execution, after being written by \textit{Write 2},
+			\textit{file 3} is entirely in cache.
+			Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
+			However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
+			cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
+			\textit{file 3} from disk.
+			The reason for the difference is in our WRENCH-cache simulator,
+			with 100 GB of input, while \textit{file 3} is being written,
+			the page cache is saturated, eviction is then evoked to evict data
+			from page cache and make space available for \textit{file 3}.
+			However, because the evicted amount is less than required,
+			some \textit{file 3} data is evicted after being written, make \textit{file 3}
+			partially cached.
+			As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
+			with memory bandwidth, a part of it is read from disk with disk read bandwidth,
+			which is many times slower, leading to longer read time.
 

### 2020-12-09 paper_io_sim.tex
added comment

--- 2020-12-09 paper_io_sim.tex
@@
 
-            Linux implements a disk cache called \textit{page cache}. 
-            It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
-            The goal of the page cache is to avoid disk I/O by keeping data in memory 
-            for later accesses, minimize I/O time since disk accesses are considerably 
-            slower than memory accesses.
-            When the kernel starts a read operation, it checks if required data is in page cache.
-            If yes, data is read from memory. This is called a \textit{cache hit}. 
-            Otherwise, data is read from disk and placed in page cache.
-            For writing, page cache is implemented with three different strategies. 
-            The first is \textit{direct~I/O} or \textit{no-write}, which means the page cache 
-            is not involved in write operations. 
-            The second is \textit{write-through}, in which data goes \textit{through} 
-            page cache to disk. In this case, data is written to both page cache and disk. 
-            In the third strategy, called \textit{writeback}, data are written directly 
-            to disk as \textit{dirty} pages and these pages will be periodically flushed 
-            to disk later by a \textit{writeback} process and then marked as not dirty. 
+        To improve performance of disk accesses, the page cache was introduced to the Linux kernel. %It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
+            The goal of the page cache is to try ensure that most I/O operations occur entirely
+            within memory space rather than synchronously to and from the storage device.
+            All disk-backed data located in the page cache are partitioned into many small pages.
+            When a read operation is initiated, the kernel checks whether the required pages are in memory.
+            If it is the case, data is read entirely from memory (i.e. \textit{cache hit}). Otherwise,
+            the necessary pages must be loaded to memory from the disk (i.e. \textit{cache miss})
+            Conversely, for writing, one of three strategies can be applied: direct~I/O (or, ``no-write''),
+            writethrough and writeback.
+            If direct~I/O is enabled for a given file system, all I/O operations bypass the
+            page cache for writes, writing data directly to disk. In contrast, write-through enables the page cache
+            for file system writes, but the data is written synchronously from memory to disk. Writeback extends write-through
+            by enabling asynchronous writes to disk, meaning that a write call will return once
+            the data has been written to memory.
             The writeback strategy is considered to outperform write-through as well as 
-            direct I/O as it delays disk writes and perform a bulk write later.
+            direct I/O as it delays disk writes to perform a bulk write at a later time.
             \cite{linuxdev3rd2010}.
 
-            In writeback, one of the keys is the cache eviction mechanism, which is
-            triggered when available memory is insufficient to remove pages 
-            from page cache. 
-            The term \textit{available memory} here includes both free memory and 
-            the memory used but can be reclaimed by the kernel with flushing 
-            and eviction. 
-            Because dirty pages are data written to page cache but yet to be 
-            written to disk, dirty pages must be kept in page cache and only clean pages 
-            are available for eviction.
-            If the clean pages are inadequate, dirty pages are flushed to disk, 
-            marked as clean and available for eviction. 
-            There is another type of flushing mechanism, which is 
-            \textit{periodical flushing}, called periodically in predefined intervals. 
+            Cache eviction and flushing strategies are integral to proper page cache functioning.
+            Whenever space in memory becomes limited, either as a result of application memory
+            or page cache use, page cache data may be evicted. Only data that 
+            has been persisted to storage (clean pages) can be flagged for eviction and removed from
+            memory. Written data that has not yet been persisted to disk (dirty data) must first
+            be copied (flushed) to storage prior to eviction. When sufficient memory is
+            being occupied, the flushing process is synchronous. However, even when 
+            there is sufficient available memory, written data will be flushed to disk
+            at a predefined interval through a process known as \textit{periodical flushing}.
             Periodical flushing only flushes expired dirty pages, which remain dirty in 
@@
             not only to balance the trade-off between accuracy and scalability, 
-            but also to achieve versatility, which allows the use of the framework 
-            to simulate applications in multiple domains \cite{casanova2014simgrid}.
+            but also to achieve versatility, allowing the use of the framework 
+            to simulate applications in multiple domains. WRENCH, a workflow management
+            system simulator based on the SimGrid framework, provides a high-level API
+            for simulating workflows, thus facilitating the integration of new models.
+            \todo{say more about wrench}
 

### 2020-12-09 result/single/plot_error.py

@@ -19,34 +19,38 @@ def grouped_bar_chart(ax, labels, xlabel, ylabel, *argv):
     ax.set_ylim(bottom=0, top=700)
 
 
-def plot_task_error(ax, size):
+def plot_task_error(ax, size, ylabel):
     labels = ["Read 1", "Write 1", "Read 2", "Write 2", "Read 3", "Write 3"]
 
     # atop_file =     "log/cluster/100gb/atop_mem.log"
     real_time_log = "real/%dgb/timestamps.csv" % size
     sim_py_log = "pysim/%dgb_sim_time.csv" % size
-    simgrid_org_log = "wrench/original/%dgb_sim_time.csv" % size
-    simgrid_ext_log = "wrench/pagecache/%dgb_sim_time.csv" % size
+    wrench_org_log = "wrench/original/%dgb_sim_time.csv" % size
+    wrench_ext_log = "wrench/pagecache/%dgb_sim_time.csv" % size
 
     py_error = [item * 100 for item in evaluate.task_time_error(real_time_log, sim_py_log)]
-    simgrid_org_error = [item * 100 for item in evaluate.task_time_error(real_time_log, simgrid_org_log)]
-    simgrid_ext_error = [item * 100 for item in evaluate.task_time_error(real_time_log, simgrid_ext_log)]
+    wrench_org_error = [item * 100 for item in evaluate.task_time_error(real_time_log, wrench_org_log)]
+    wrench_ext_error = [item * 100 for item in evaluate.task_time_error(real_time_log, wrench_ext_log)]
 
-    grouped_bar_chart(ax, labels, "", "error (%)",
-                      ("Python simulator", py_error, "tab:pink"), ("WRENCH", simgrid_org_error, 'tab:orange'),
-                      ("WRENCH-cache", simgrid_ext_error, 'tab:cyan'))
+    if ylabel:
+        ylabel = "error (%)"
+    else:
+        ylabel = ""
+    grouped_bar_chart(ax, labels, "", ylabel,
+                      ("Python prototype", py_error, "#994F88"), ("WRENCH", wrench_org_error, '#7BAFDE'),
+                      ("WRENCH-cache", wrench_ext_error, '#1965B0'))
     ax.set_title("%d GB" % size)
 
-
 def plot_error():
-    plt.rcParams.update({'font.size': 7})
[…]

### 2020-12-09 paper_io_sim.tex

@@
         and network-based filesystems.
-		The results show that our simulator not only achieves higher accuracy than 
-		original WRENCH, but also effectively simulates memory contents and 
-		page cache during I/O operations. 
-		Through this study, our effort is to contribute an analytical model 
-		of the page cache, which can also be integrated to other simulators 
-		to improve I/O simulation accuracy.
 

### 2020-12-09 paper_io_sim.tex
added comment
edited results and conclusion

--- 2020-12-09 paper_io_sim.tex
@@
 
-        The Linux Page Cache plays an important role in
-        the reduction of filesystem data transfer costs. With the page cache, previously
+        The Linux Page Cache, introduced in kernel version 2.4, plays an important role in
+        the reduction of filesystem data transfer costs~\cite{pagecache}. With the page cache, previously
         read data can be reread directly from memory and written data can be written to
@@
         costs, unstable environments, and the difficulties in producing reproducible results.
-        Simulations address these concerns by providing abstractions for the performance of computer
-        hardware, such as CPU, network and storage. Consequently, simulations provide and
+        Simulations address these concerns by providing abstractions to the performance of computer
+        hardware, such as CPU, network and storage. As a result, simulations provide and
         cost-effective, fast, easy and reproducible  way evaluate application performance
@@
 
-        The amount of data-intensive applications has been on the rise as a result of
-        open-data and data sharing initiatives. Due to the sheer size of the data being
-        processed, these applications must be executed on large scale infrastructure
-        such as High Performance Computing (HPC) clusters or the cloud. Determining
-        which type of infrastructure is the most cost-effective for the particular
-        application can be tricky without the help of a simulator.
-
         Page cache simulation is crucial when simulating data intensive applications,
-        which rely on cache hits and writing to memory for additional performance.
-        Furthermore, simulating the page cache enables us to study important application-I/O
-        bottlenecks which can arise over the course of execution, such as page cache
-        starvation~\cite{zhuang2017}.
+        which would rely on cache hits and writing to memory for additional performance.
         While existing simulators can adequately characterize many different computing environments,
@@
 
-        In this paper, we present WRENCH-Cache, a simulation model of I/O with page
-        cache incorporated into WRENCH~\cite{casanova2018wrench}, a workflow management system simulator based on
-        the popular SimGrid framework. The goal of WRENCH-Cache is to provide users
+        In this paper, we present WRENCH-cache, a simulation model of I/O with page
+        cache incorporated into WRENCH, a workflow management system simulator based of
+        the popular SimGrid framework. The goal of WRENCH-cache is to provide users
         with an accurate representation of a workflow execution, when file system page
-        cache is used, while providing a good balance between accuracy and scalability.
-        WRENCH-Cache is able to simulate both writeback and writethrough
+        cache is used. WRENCH-cache is able to simulate both writeback and writethrough
         cache which would normally be available to local and NFS-based filesystems.
-        WRENCH-Cache can provide accurate simulations for large-scale parallel
-        applications using the page cache.
 
@@
 
-        To improve performance of disk accesses, the page cache was introduced to the Linux kernel. %It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
-            The goal of the page cache is to try ensure that most I/O operations occur entirely
-            within memory space rather than synchronously to and from the storage device.
-            All disk-backed data located in the page cache are partitioned into many small pages.
-            When a read operation is initiated, the kernel checks whether the required pages are in memory.
-            If it is the case, data is read entirely from memory (i.e. \textit{cache hit}). Otherwise,
-            the necessary pages must be loaded to memory from the disk (i.e. \textit{cache miss})
-            Conversely, for writing, one of three strategies can be applied: direct~I/O (or, ``no-write''),
-            writethrough and writeback.
-            If direct~I/O is enabled for a given file system, all I/O operations bypass the
-            page cache for writes, writing data directly to disk. In contrast, write-through enables the page cache
-            for file system writes, but the data is written synchronously from memory to disk. Writeback extends write-through
-            by enabling asynchronous writes to disk, meaning that a write call will return once
-            the data has been written to memory.
+            Linux implements a disk cache called \textit{page cache}. 
+            It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
+            The goal of the page cache is to avoid disk I/O by keeping data in memory 
+            for later accesses, minimize I/O time since disk accesses are considerably 
+            slower than memory accesses.
+            When the kernel starts a read operation, it checks if required data is in page cache.
+            If yes, data is read from memory. This is called a \textit{cache hit}. 
+            Otherwise, data is read from disk and placed in page cache.
+            For writing, page cache is implemented with three different strategies. 
+            The first is \textit{direct~I/O} or \textit{no-write}, which means the page cache 
+            is not involved in write operations. 
+            The second is \textit{write-through}, in which data goes \textit{through} 
+            page cache to disk. In this case, data is written to both page cache and disk. 
+            In the third strategy, called \textit{writeback}, data are written directly 
+            to disk as \textit{dirty} pages and these pages will be periodically flushed 
+            to disk later by a \textit{writeback} process and then marked as not dirty. 
             The writeback strategy is considered to outperform write-through as well as 
-            direct I/O as it delays disk writes to perform a bulk write at a later time.
+            direct I/O as it delays disk writes and perform a bulk write later.
             \cite{linuxdev3rd2010}.
 
-            Cache eviction and flushing strategies are integral to proper page cache functioning.
-            Whenever space in memory becomes limited, either as a result of application memory
-            or page cache use, page cache data may be evicted. Only data that 
-            has been persisted to storage (clean pages) can be flagged for eviction and removed from
-            memory. Written data that has not yet been persisted to disk (dirty data) must first
-            be copied (flushed) to storage prior to eviction. When sufficient memory is
-            being occupied, the flushing process is synchronous. However, even when 
-            there is sufficient available memory, written data will be flushed to disk
-            at a predefined interval through a process known as \textit{periodical flushing}.
+            In writeback, one of the keys is the cache eviction mechanism, which is
+            triggered when available memory is insufficient to remove pages 
+            from page cache. 
+            The term \textit{available memory} here includes both free memory and 
+            the memory used but can be reclaimed by the kernel with flushing 
+            and eviction. 
+            Because dirty pages are data written to page cache but yet to be 
+            written to disk, dirty pages must be kept in page cache and only clean pages 
+            are available for eviction.
+            If the clean pages are inadequate, dirty pages are flushed to disk, 
+            marked as clean and available for eviction. 
+            There is another type of flushing mechanism, which is 
+            \textit{periodical flushing}, called periodically in predefined intervals. 
             Periodical flushing only flushes expired dirty pages, which remain dirty in 
@@
             \cite{owda2014comparison}.
-            The Linux kernel uses a two-list strategy to flag pages for eviction.
-            The two-list strategy is based on a least recently used (LRU) policy
-            and uses an active and inactive list in its implementation.
+            In Linux kernel, a \textit{two lists strategy (LRU/2)} is implemented based on 
+            the least-recently-used (LRU) scheme, with an \textit{active list} and 
+            an \textit{inactive list}. 
             If accessed pages are not in the page cache, they are added to the inactive list. 
-            Should pages located on the inactive list be accessed, they will be moved from 
-            the inactive to the active list. 
+            If pages are accessed when residing in the inactive list, they will be moved from 
+            inactive list to active list. 
             The lists are also kept balanced by moving pages from the active list
@@
             and not evictable, while the inactive list includes pages accessed once only, 
-            or pages that have be accessed more than once but moved from the active list. 
-            Both lists operate using LRU eviction policies meaning the data that has
-            not be accessed recently will be moved first.
+            or pages accessed more than once but moved from the active list. 
+            The lists work in LRU manner, pages are added to the tail and removed 
+            from the top.
 
@@
 
-            Many simulation frameworks exist to describe the behaviours of applications
-            running on the various infrastructures available to users, such as personal computers,
-            clouds and clusters.
-            These simulators are built with resource models such CPU, network and 
-            storage, using various different approaches to describe the hardware.
-            For example, there is \textit{CloudSim} and \textit{iCanCloud} for packet-level network models 
-            and \textit{DiskSim} the simulates storage hardware functioning.
-
-            Simulators must choose between scalability and accuracy when describing the
-            hardware.\cite{casanova2014simgrid}. 
+            Many simulation frameworks and applications have been developed 
+            in recent years to support studies in grids, cloud, HPC, peer-to-peer 
+            networks, volunteer computers.                     
+            These simulators are built with models of resources such as CPU, network, 
+            storage with different approaches and levels of details in different simulators. 
+            The detail level of the model causes a conflict between accuracy and scalability, 
+            and these simulators usually trade of one for the other 
+            \cite{casanova2014simgrid}. 
+            For example, some simulators are built with packet-level network models 
+            such as \textit{CloudSim} and \textit{iCanCloud}, or storage model that 
+            simulates the operation of storage hardware in \textit{DiskSim} 
+            \cite{casanova2014simgrid}. 
             Fine-grained models allow simulators to achieve high accuracy, 
-            but lead to scalability issues due to long simulation time 
+            but at the same time lead to scalability issues due to long simulation time 
             when simulated platforms and applications grow. 
@@
             not only to balance the trade-off between accuracy and scalability, 
-            but also to achieve versatility, allowing the use of the framework 
-            to simulate applications in multiple domains. WRENCH, a workflow management
-            system simulator based on the SimGrid framework, provides a high-level API
-            for simulating workflows, thus facilitating the integration of new models.
-            \todo{say more about wrench}
+            but also to achieve versatility, which allows the use of the framework 
+            to simulate applications in multiple domains \cite{casanova2014simgrid}.
 
@@
             accessed on disk in a block cache \cite{nunez2012simcan}. 
-            Page cache is also modeled in the iCanCloud simulator through a component which
-            manages memory accesses and cached data \cite{nunez2012icancloud}. However,
-            the scalability of the iCanCloud simulator is limited by its sophisticated model.
-            Ultimately, none of these simulators provide any writeback cache simulator nor
-            cache eviction policys through LRU lists.
-            Although cache replacement policies are applied in \cite{xu2018saving} to simulate 
-            in-memory caching, their simulator is specific to energy consumption of multi-tier 
-            heterogenous networks.
+            Page cache is also modeled in iCanCloud simulator with a component in charge 
+            of managing memory accesses and cached data \cite{nunez2012icancloud}, 
+            but one of the issues is scalability due to its sophisticated model.
+            Ultimately, non of these simulators provide any cache write strategy, 
+            cache eviction mechanism with dirty/clean data and the page cache LRU lists. 
+            Cache replacement policies is applied in \cite{xu2018saving} to simulate 
+            in-memory caching, but for the study in energy consumption of multi-tier 
+            heterogenous network.
 
-            In this study, we use WRENCH 
-            to implement our page cache model. We chose WRENCH due to accurate yet scalable implementation, resulting
-            from its SimGrid backend, as well as its high-level API.
-            We have integrated our model in WRENCH and use the SimGrid API to interact with 
-            simulated storage resources.
-            This approach allows us to leverage existing features in WRENCH and SimGrid such as process control and synchronization of 
-            resource consumption.
-            We also leverage the existing SimGrid storage model with capacity 
+            In this study, we chose SimGrid to implement our model of I/O with page cache 
+            due to the balancing between accuracy and scalibility, the versatility allowing it 
+            to be adopted in different domains. 
+            In addition, it is an actively maintained open-scource framework with 
+            modular design, which enables the framework to be extended with new models. 
+            The main issue users have to face when using SimGrid is everything has to be 
+            implemented from scratch due to its low-level API. 
+            To ease this challenge, WRENCH, another open-source library, is built on top 
+            of SimGrid to abstract the low-level logic and expose a higher abstraction 
+            level API \cite{casanova2018wrench}. 
+            We integrate our model in WRENCH and use the SimGrid API to interact with 
+            simulated storage resource.
+            This approach allows us to leverage existing features in WRENCH and SimGrid.
+            Firstly, we can take advantage of the process control, synchronization of 
+            resource consumption by concurrent processes provided by the simulation core 
+            in SimGrid. 
+            Secondly, we can reuse the existing SimGrid storage model with capacity 
             and bandwidth properties, which are similar to the characteristics of memory.
+            And finally, WRENCH provides a high level abstraction, which help us drop 
+            the burden of implementation.
 
@@
 
-    \begin{algorithm}\caption{Periodical flushing simulation}\label{alg:pdflush}
+    \begin{algorithm}\caption{Periodical flushing simulation in IO Controller}\label{alg:pdflush}
         \small
@@
 
-    \begin{algorithm}\caption{File chunk read simulation of IO Controller}
+    \begin{algorithm}\caption{File chunk read simulation in IO Controller}
     \label{alg:read}
@@
 
-    \begin{algorithm}\caption{File chunk write simulation of IO Controller}
+    \begin{algorithm}\caption{File chunk write simulation in IO Controller}
     \label{alg:write}
@@
 
-            Our first experiment (\textbf{Exp single-threaded}) simply runs one instance of the
+            Our first experiment (\textit{Exp single-threaded}) simply runs one instance of the
             application on a single cluster node, with different input file
@@
 
-            Our second experiment (\textbf{Exp concurrent}) runs concurrent instances of the
+            Our second experiment (\textit{Exp concurrent}) runs concurrent instances of the
             application on a single node, all application instances
@@
 
-            Our third experiment (\textbf{Exp NFS}) occurs in the same configuration as the
+            Our third experiment (\textit{Exp NFS}) occurs in the same configuration as the
             previous one, albeit reading and writing on a 400-GB NFS-mounted remote disk
@@
 
-        Figure~\ref{fig:single_error} shows the relative error of simulated
-        read and write times for the three successive application tasks in \textbf{Exp single-threaded}. The
-        page cache simulation model drastically reduces simulation errors
-        for all I/O operations except the first read as it involves
-        uncached data. Errors are reduced from an average of 345\%
-        in the original WRENCH to 46\% in the Python prototype and 39\% in WRENCH-cache.
-        Unsurprisingly, the original WRENCH simulator significantly overestimates
-        read and write times, due to its lack of page cache simulation
-        model.
+        The page cache simulation model drastically reduces I/O simulation
+        errors in the three successive application tasks (Figure~\ref{fig:single_error}). The first read is not impacted 
+        as it only involves uncached data. Errors are reduced from an average
+        of 345\% in the original WRENCH to 46\% in the Python prototype and
+        39\% in WRENCH-cache. Unsurprisingly, the original WRENCH simulator
+        significantly overestimates read and write times, due to its lack
+        of page cache simulation model. Results with files of 50~GB and 75~GBs
+        show similar behaviors and are not reported due to space limitations.
 
-        Especially, the simulation time errors of \textit{Read 3} with 100 GB of input
-        in the simulators with page cache are just under the error in original WRENCH.
-        This longer read can be explained with the amount of cached data of files
-        at the beginning of the read in reality and simulators shown in
-        Fig.~\ref{fig:single_cache}.
-        % In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
-        % output of \textit{Write $<$i-1$>$}.
-        With the input size of 20 GB, the amount of cached data of each files after 
-        every I/O operation is similar since the page cache can keep all these files.
-%             \tristan{what does this refer to?
-%             For 20 GB there is still some error in WRENCH-cache, what do you mean 
-%``there is no difference in the page cache''?}.
-%   Dzung: Some text was lost in a merge, I added it again.
-        With the input of 100 GB, the difference between real execution and WRENCH-cache
-        simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
-        In real execution, after being written by \textit{Write 2},
-        \textit{file 3} is entirely in cache.
-        Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
-        However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
-        cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
-        \textit{file 3} from disk.
-        The reason for the difference is in our WRENCH-cache simulator,
-        with 100 GB of input, while \textit{file 3} is being written,
-        the page cache is saturated, eviction is then evoked to evict data
-        from page cache and make space available for \textit{file 3}.
-        However, because the evicted amount is less than required,
-        some \textit{file 3} data is evicted after being written, make \textit{file 3}
-        partially cached.
-        As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
-        with memory bandwidth, a part of it is read from disk with disk read bandwidth,
-        which is many times slower, leading to longer read time.
+        WRENCH simulation errors are substantially lower with 100~GB
+        files than with 20~GB files, due to the fact that part of the
+        100~GB file needs to be read and written to disk, the only storage
+        device in WRENCH, as it does not fit in cache. Conversely,
+        simulation errors of the Python prototype and WRENCH-cache are higher with
+        100~GB files than with 20~GB files, due to idiosyncracies in the kernel 
+        flushing and eviction strategies.
 
-            Memory profiling results of real pipeline execution and simulators 
-            with 20GB and 100GB input files are illustrated in Fig.~\ref{fig:single_memprof}.
-            Because the results with 20Gb, 50GB, 100GB shows a similar pattern,
-            we only include the results with 20GB and 100GB in this paper.
+        Simulated memory profiles are highly consistent with the real ones
+        (Figure~\ref{fig:single_memprof}). With 20~GB files, memory
+        profiles almost exactly match the real ones, although dirty data
+        seems to be flushing faster in real life than in simulation
+        \tristan{any idea why?}, a behavior also observed with 100~GB files. With
+        100~GB files, used memory reaches total memory during the first write,
+        triggering dirty data flushing. Used memory drops to cached memory
+        when application tasks release anonymous memory. Simulated cached
+        memory is highly consistent with real values, except toward the end
+        of Read 3 where it slightly increases in simulation but not in
+        reality \tristan{why?}. As expected, dirty data remains under the
+        dirty ratio. The Python prototype and WRENCH-cache exhibit nearly
+        identical memory profiles, which reinforces the confidence in our
+        implementation.
 
-            As shown in the figures, the makespans of the simulated pipeline are
-            close to the makespan of the real execution with similar patterns of I/O time
-            real execution and simulation.
-            But more importantly, with all input sizes, similar trends can also be found
-            in the lines describing memory status in simulators and the real execution.
-            The similar evolutions of amount of cache used (purple lines) in reality
-            and simulation show that the use of page cache is well reproduced by the simulators.
-            The amount of dirty data (red lines) also follow similar trends,
-            which means not only the amount dirty data, but also the flushing and
-            periodical flushing mechanisms are accurately modeled in our simulators.
-            This can be seen in in the increase of the amount of dirty data during the writes
-            and the decrease during the reads and computations.
-            According to the results achieved with all input sizes in the experiment,
-            we can observe identical trends in reality and simulation with 20GB,
-            50GB and 75GB input files. However, with 100GB input file,
-            the third read is not correctly simulated, leading to longer simulated time
-            than in real execution. This will be explained later in this section.
+        The content of the simulated memory cache is also highly consistent
+        with reality (Figure~\ref{fig:single_cache}). With 20~GB files, the
+        simulated cache content exactly matches reality, owing to the fact
+        that all files fit in page cache. With 100~GB files, a slight
+        discrepancy is observed after Write 2, which explains the
+        simulation error previously mentioned in Read 3. In the real execution indeed, 
+        File 3 is entirely cached after Write 2, whereas in the simulated execution, part 
+        of it is on disk. This is due to the fact that the simulated flushing and eviction 
+        was triggered after File 3 was already written, while in reality these operations 
+        happened before File 3 was written \tristan{Dzung, pls check that. Also, why doesn't it happen irl?}.
 
-             Figure~\ref{fig:single_error} shows the relative error of simulated
-             read and write times throughout the three application tasks. The
-             page cache simulation model drastically reduces simulation errors
-             for all I/O operations except the first read that deals with
-             uncached data. Errors are reduced from an average of 345\%
-             in WRENCH to 46\% in the Python prototype and 39\% in WRENCH-cache.
-             The original WRENCH simulator significantly overestimates
-             read and write times, due to its lack of page cache simulation
-             model.
-             Especially, the simulation time errors of \textit{Read 3} with 100 GB of input
-             in the simulators with page cache are just under the error in original WRENCH.
-             This longer read can be explained with the amount of cached data of files
-             at the beginning of the read in reality and simulators shown in
-             Fig.~\ref{fig:single_cache}.
-             % In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
-             % output of \textit{Write $<$i-1$>$}.
-             With the input size of 20 GB, the amount of cached data of each files after 
-             every I/O operation is similar since the page cache can keep all these files.
-%             \tristan{what does this refer to?
-%             For 20 GB there is still some error in WRENCH-cache, what do you mean 
-%``there is no difference in the page cache''?}.
-%   Dzung: Some text was lost in a merge, I added it again.
-             With the input of 100 GB, the difference between real execution and WRENCH-cache
-             simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
-             In real execution, after being written by \textit{Write 2},
-             \textit{file 3} is entirely in cache.
-             Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
-             However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
-             cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
-             \textit{file 3} from disk.
-             The reason for the difference is in our WRENCH-cache simulator,
-             with 100 GB of input, while \textit{file 3} is being written,
-             the page cache is saturated, eviction is then evoked to evict data
-             from page cache and make space available for \textit{file 3}.
-             However, because the evicted amount is less than required,
-             some \textit{file 3} data is evicted after being written, make \textit{file 3}
-             partially cached.
-             As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
-             with memory bandwidth, a part of it is read from disk with disk read bandwidth,
-             which is many times slower, leading to longer read time.
+        % We conclude from this experiment that our model can accurately
+        % simulate cached I/Os in a sequential context. The next experiment 
+        % evaluates our simulation with concurrent I/Os.
 
@@
                    \vspace*{-0.7cm}
-                   \caption{Simulation errors}
+                   \caption{Absolute relative simulation errors}
                    \vspace*{0.5cm}
@@
                    \vspace*{-0.7cm}
-                   \caption{Memory profiles}
+                   \caption{Memory profiles \tristan{Remove trailing whitespace for 20GB graphs (xmax should be about 160s, not 200s)}}
                    \vspace*{0.5cm}
@@
                    \includegraphics[width=1.05\linewidth]{result/single/figures/cached_files.pdf}
-                   \caption{Cache contents}
+                   \caption{Cache contents \emph{after} application I/O operations}
                 %    \textcolor{red}{Update real results of 20GB}}
                    \label{fig:single_cache}
-            \end{subfigure}\\
+            \end{subfigure}
             \caption{Single-threaded results}
@@
 
-            In summary, in this single-threaded experiment, our model can accurately
-            simulate I/O time as well as the internal memory behavior.
-            This can help confirm that our model is on a right track and
-            it can be further improved to achieve higher accuracy.
-
-        \subsection{Multi-threaded experiment}
+        \subsection{Concurrent applications}
 
@@
             \end{subfigure}
-            \caption{I/O time of concurrent pipelines with local storage.}
+            \caption{Concurrent results with 3~GB files (averages on 5 repetitions)}
             \label{fig:multi_local}
@@
 
-            In this experiment, we run the real execution in 5 repetitions to
-            capture the variability on the cluster.
-            The results are analyzed in terms of average I/O time
-            of concurrent pipelines shown in Fig.~\ref{fig:multi_local}.
-
-            As is shown in the figure, when the number of pipelines increases,
-            the average read time raises accordingly because the amount of data
-            increases but memory and disk bandwidths are fixed and shared between pipelines.
-            In comparison with original WRENCH, the results from WRENCH-cache are closer
-            to reality.
-            Especially, when it comes to the average write time,
-            although the trend in reality is more complex, our WRENCH-cache
-            simulator can still capture the pattern in real execution.
-            In both, the average write time gradually increases before surging with the
-            same slopes after the number of pipelines reaches 10.
-            This can be explained that with less than 10 concurrent pipelines,
-            the page cache is not saturated, so all files can be written entirely to
-            cache with memory bandwidth in a short time.
-            After the page cache is saturated with dirty data (at around
-            10 concurrent pipelines in this experiment), this dirty data needs
-            to be flushed in order to make space available for writing new data to cache.
-            This dirty data is flushed to disk with disk bandwidth, which is much
-            slower than memory bandwidth.
-            The more pipelines we have, the more data needs to be written, the more data
-            needs to be flushed, leading to a sharper increase in the average write time.
-
-            To conclude, the results from WRENCH-cache simulator show that
-            our model can also simulate I/O with page cache accurately in a
-            multi-threaded application.
+            The page cache model notably reduce WRENCH's simulation error
+            for concurrent applications and local I/Os
+            (Figure~\ref{fig:multi_local}). For reads, WRENCH-cache still
+            slightly overestimates reality, due to the discrepancy between
+            simulated and real read bandwidths mentioned before
+            (\tristan{check it's indeed mentioned. Dzung: is it correct?}). For writes, 
+            WRENCH-cache slightly underestimates reality for the same reason. The plateau 
+            ending around 7 concurrent pipelines in the WRENCH-cache curve
+            marks the limit beyond which the page cache is saturated with dirty data 
+            and needs flushing \tristan{why is there no plateay irl?}.
 
@@
             \end{subfigure}
-            \caption{I/O time of concurrent pipelines with NFS}
+            \caption{NFS results with 3~GB files (averages on 5 repetitions)}
             \label{fig:multi_nfs}
@@
 
-            Similar to the previous experiment, the real execution with NFS is also
-            repeated for 5 times and the results are described with the average I/O time
-            shown in Fig.~\ref{fig:multi_nfs}.
-            In general, the average I/O time reflect a very similar level of accuracy
-            to the previous experiment results when the I/O time of pipelines
-            is very close to the real execution.
-            The simulated read time if a bit off when the number of pipelines
-            surpasses 22, with 264 GB of data in total.
-            This is due to the impact of simulated cache eviction when
-            the page cache is saturated, which has been seen in
-            Fig.~\ref{fig:single_error} with 100 GB of input.
-            When it comes to write time, original WRENCH and WRENCH-cache
-            results are identical when both are writing directly to disk.
+            Page cache simulation importantly reduces the simulation error
+            on remote storage as well (Figure~\ref{fig:multi_nfs}). This
+            manifests only for reads, as only write-through cache was
+            enabled in the NFS server. Both WRENCH and WRENCH-cache
+            underestimate write times due to the discrepancy between
+            simulated and real bandwidths mentioned previously. For reads,
+            this discrepancy only impacts the results beyond 22
+            applications since before this threshold, most reads result in cache
+            hits \tristan{Dzung, pls check that.}.
 
         \subsection{Real application}
+        \subsection{Overhead comparison}
+        \tristan{Dzung, add your overhead measures here.}
 
-    \section{Discussion and Future Work}
+    \section{Conclusion}
     \label{discussion}
-        In computing infrastructures, especially those use for data-intensive applications,
-        page cache undoubtedly has certain impacts on performance.
-        Conducting simulation experiments is an effective approach to predict,
-        evaluate the performance of not only application, scheduler but also
-        infrastructure settings.
-        Unfortunately, page cache is missing in many simulation tools for HPC systems.
-        In this study, we proposed a simulation model of I/O with page cache,
-        implement it with SimGrid using WRENCH, and evaluate the results
-        with different experiments.
-        The results shows that our model can result in better simulation accuracy than
-        the simulators in which page cache is not modeled.
-        Furthermore, this level of accuracy can be achieved without a fine-grained
-        level of details.
-        This may create opportunities for the model to be applied in simulation of
-        big data applications on HPC infrastructures, which leverage memory to
-        employ page cache as a disk cache.
-        Nevertheless, the model exposed some limitations that can be
-        further improved.
+        Page cache undoubtedly has important impacts on performance, in
+        particular for data-intensive applications. Numerical simulation is
+        an effective approach to predict and compare the performance of
+        applications, schedulers and infrastructures. However, page cache
+        is missing in many simulation tools for distributed systems. In
+        this study, we proposed a simulation model of I/Os with page cache,
+        we implemented it in the SimGrid-based WRENCH library, and we
+        evaluated its benefits in different conditions. Results show that
+        our model very substantially improves simulation accuracy,
+        decreasing absolute relative errors between simulated and real
+        performance by a factor close to 10 (see results of the
+        single-threaded experiment). Furthermore, this level of accuracy
+        can be achieved without a fine-grained level of details
+        \tristan{not sure what you mean}. Our page cache model is available
+        in WRENCH from commit \tristan{abcd}, ready to be used for the
+        simulation of Big Data applications.
 
-        Our short term future work will be re-evaluate the model with asymmetrical
-        disk bandwidths that will be released in the next SimGrid version.
-        The next step is to study the detail in cache eviction mechanism to make
-        the our simulation results closer to reality.
-        Another point that needs to be further improved is the ability to configure
-        different anonymous memory usage levels of applications since it affects
-        the amount of cache used.
-        We also plan to support random I/O and file readahead, as well as
-        saving and restoring disk state when hosts are turned on/off.
+        Simulation results could be made even more accurate by a deeper
+        investigation of cache eviction mechanisms \tristan{be more
+        specific}. The availability of asymmetrical bandwidths in the
+        forthcoming SimGrid release will also directly improve our
+        simulation accuracy. Relevant future work includes simulating
+        application memory consumption, random I/O, file readahead and
+        persistent disk storage.
 
+        \tristan{discussion could be fleshed out}
 \bibliographystyle{plain}

### 2020-12-09 result/single/plot_error.py

@@ -37,7 +37,7 @@ def plot_task_error(ax, size, ylabel):
     else:
         ylabel = ""
     grouped_bar_chart(ax, labels, "", ylabel,
-                      ("Python prototype", py_error, "#994F88"), ("WRENCH", wrench_org_error, '#7BAFDE'),
+                      ("Python prototype", py_error, "#7BAFDE"), ("WRENCH", wrench_org_error, '#994F88'),
                       ("WRENCH-cache", wrench_ext_error, '#1965B0'))
     ax.set_title("%d GB" % size)
 

### 2020-12-09 result/single/plot_memprof.py

@@ -158,7 +158,7 @@ def compare_plot(input_sizes=[20, 100], makespan=[200, 1300]):
     fig.text(0.77, 0.06, 'time (s)', ha='center', fontsize=12)
     fig.text(0.08, 0.45, 'memory (GB)', va='center', rotation='vertical', fontsize=12)
     fig.text(0.91, 0.69, 'Real execution', fontsize=14, color='white', bbox={'pad': 5, 'color': '#90C987'}, rotation=-90)
-    fig.text(0.91, 0.405, 'Python prototype', fontsize=14, color='white', bbox={'pad': 5, 'color': '#994F88'}, rotation=-90)
+    fig.text(0.91, 0.405, 'Python prototype', fontsize=14, color='white', bbox={'pad': 5, 'color': '#7BAFDE'}, rotation=-90)
     fig.text(0.91, 0.145, 'WRENCH-cache', fontsize=14, color='white', bbox={'pad': 5, 'color': '#1965B0'}, rotation=-90)
     
     

### 2020-12-09 paper_io_sim.tex

@@
         and network-based filesystems.
-        The results show that our simulator not only achieves higher accuracy than 
-        original WRENCH, but also effectively simulates memory contents and 
-        page cache during I/O operations. 
-        Through this study, our effort is to contribute an analytical model 
-        of the page cache, which can also be integrated to other simulators 
-        to improve I/O simulation accuracy.
 
@@
             not only to balance the trade-off between accuracy and scalability, 
-            but also to achieve versatility, which allows the use of the framework 
-            to simulate applications in multiple domains \cite{casanova2014simgrid}.
+            but also to achieve versatility, allowing the use of the framework 
+            to simulate applications in multiple domains. WRENCH, a workflow management
+            system simulator based on the SimGrid framework, provides a high-level API
+            for simulating workflows, thus facilitating the integration of new models.
+            \valerie{say more about wrench}
 

### 2020-12-09 paper_io_sim.tex

@@
 
-\title{Simulating the Linux page cache in distributed applications with the WRENCH library}
+\title{Simulating the Linux page cache in distributed applications with WRENCH}
 
@@
         The inability to simulate I/O with page cache leads to inaccurate 
-        simulations in many situations, especially when dealing with data intensive applications.
+        simulations in many situations, especially when dealing with data-intensive applications.
 
-        In this paper, we propose a I/O simulation model that characterizes the key features of the page cache.
-        We have implemented the model with SimGrid, a simulation framework for 
-        distributed systems, using WRENCH, a SimGrid-based workflow simulator. 
-        Our simulator is compatible with both single-threaded and multithreaded applications,
-        and can simulate both writeback and writethrough cache that can factor into local
-        and network-based filesystems.
+        In this paper, we propose a I/O simulation model that characterizes
+        the key features of the page cache. We have implemented the model
+        with SimGrid, a simulation framework for distributed systems, using
+        WRENCH, a SimGrid-based workflow simulator. Our simulator is
+        compatible with both single-threaded and multithreaded
+        applications, and can simulate both writeback and writethrough
+        cache that can factor into local and network-based filesystems. We
+        evaluate it in different conditions, including sequential and
+        concurrent applications, as well as local and remote I/Os.
+        The page cache model reduces the simulation
+        error by up to an order of magnitude compared to cacheless simulations.
 
@@
         Simulations address these concerns by providing abstractions for the performance of computer
-        hardware, such as CPU, network and storage. Consequently, simulations provide and
+        hardware, such as CPU, network and storage. Consequently, simulations provide a
         cost-effective, fast, easy and reproducible  way evaluate application performance
@@
         they lack much of the ability to adequately describe the page cache. Some
-        simulators try to simulator I/O with page cache, but fail to capture the key features such
+        simulators try to simulate I/O with page cache, but fail to capture the key features such
         as dirty data and cache eviction policies~\cite{nunez2012simcan,nunez2012icancloud}. 
-        Other simulators, such as~\cite{xu2018}, capture the important feature of the
+        Other simulators, such as~\cite{xu2018saving}, capture the important features of the
         page cache, but are domain-specific.
@@
 
-        In this paper, we present WRENCH-Cache, a simulation model of I/O with page
+        In this paper, we present WRENCH-cache, a simulation model of I/O with page
         cache incorporated into WRENCH~\cite{casanova2018wrench}, a workflow management system simulator based on
-        the popular SimGrid framework. The goal of WRENCH-Cache is to provide users
+        the popular SimGrid framework. The goal of WRENCH-cache is to provide users
         with an accurate representation of a workflow execution, when file system page
         cache is used, while providing a good balance between accuracy and scalability.
-        WRENCH-Cache is able to simulate both writeback and writethrough
+        WRENCH-cache is able to simulate both writeback and writethrough
         cache which would normally be available to local and NFS-based filesystems.
-        WRENCH-Cache can provide accurate simulations for large-scale parallel
+        WRENCH-cache can provide accurate simulations for large-scale parallel
         applications using the page cache.
@@
             The writeback strategy is considered to outperform write-through as well as 
-            direct I/O as it delays disk writes to perform a bulk write at a later time.
+            direct I/O as it delays disk writes to perform a bulk write at a later time
             \cite{linuxdev3rd2010}.
@@
 
-            Many simulation frameworks exist to describe the behaviours of applications
+            Many simulation frameworks exist to describe the behaviour of applications
             running on the various infrastructures available to users, such as personal computers,
@@
             For example, there is \textit{CloudSim} and \textit{iCanCloud} for packet-level network models 
-            and \textit{DiskSim} the simulates storage hardware functioning.
+            and \textit{DiskSim} the simulates storage hardware functioning \tristan{add references}.
 
             Simulators must choose between scalability and accuracy when describing the
-            hardware.\cite{casanova2014simgrid}. 
+            hardware~\cite{casanova2014simgrid}. 
             Fine-grained models allow simulators to achieve high accuracy, 
@@
     \label{method}
-    In this section, we present our page cache simulation model for file I/O,
-    flusing, and eviction. We implement this model in two simulators:
-    (1) a standalone Python prototype, and (2) WRENCH-cache, an extension of the WRENCH library.
-    We also describe experimental scenarios to evaluate our simulators in various configurations.
-
     We separate our simulation model in two components, the IO
     Controller and the Memory Manager, which together simulate 
-    file reads and writes in Linux (Figure~\ref{fig:interaction}). 
+    file reads and writes (Figure~\ref{fig:interaction}). 
     To read or write a file chunk, a simulated application sends a
     request to the IO Controller. The IO Controller interacts as needed with
-    the Memory Manager to free memory through flushing or eviction,
-    and to read or write to cache. The Memory Manager
+    the Memory Manager to free memory by triggering flushing or eviction,
+    and interact with cached data. The Memory Manager
     implements these operations, simulates periodical flushing
     and eviction from cache, and reads or writes to disk when necessary.
-    In case write-through strategy is used, IO Controller can also interact directly 
+    In case the write-through strategy is used, the IO Controller can also interact directly 
     with disk model to simulate disk writes.
@@
     The Memory Manager simulates two parallel threads: the main one
-    implements flushing, eviction, and cache I/Os synchronously, whereas
+    implements flushing, eviction, and cached I/Os synchronously, whereas
     the second one, operating in the background, periodically searches for
@@
         \tristan{discussion could be fleshed out}
+
+        \section{Acknowledgments}
+The computing platform used in the experiments was obtained with funding
+from the Canada Foundation for Innovation.
 \bibliographystyle{plain}
+\tristan{Check references, in particular capitalization}
 \bibliography{citation}

### 2020-12-09 paper_io_sim.tex
Henri's pass on abstract and Section 1

@@
 
+%\title{Simulating the Linux page cache in distributed applications with WRENCH}
 \title{Simulating the Linux page cache in distributed applications with WRENCH}
@@
     $^1$Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada\\
-     $^2$Department of Information and Computer Sciences, University of Hawai‘i at Manoa, USA\\
+     $^2$Department of Information and Computer Sciences, University of Hawai`i at M\=anoa, USA\\
      $^3$Department of Computer Science, University of Southern California, Los Angeles, USA
@@
 
-        The rapid emergence of Big Data seen in recent years has resulted in a 
-        growing need for efficient data storing and optimized data analysis techniques. 
-        Despite relying on large computing infrastructures with powerful computing capability 
-        to process these large datasets, an I/O bottleneck remains. 
-        One widely-used approach to minimize the costs of data transfers during
-        regular use is through the employment of the Linux page cache.
-        While many simulation tools have been developed to 
-        support studies in computing systems optimization, page cache simulation 
-        has not been fully captured or remains absent from these simulators. 
-        The inability to simulate I/O with page cache leads to inaccurate 
-        simulations in many situations, especially when dealing with data-intensive applications.
+        % Henri Rewrote quite  a bit in the next paragraph here to spell out a bit
+        % more why we need simulation
+	The emergence of Big Data in recent years has resulted in a growing
+	need for efficient data storage management techniques.  While
+	infrastructures with sufficient compute power are available to
+	process large datasets, the I/O bottleneck remains.  A
+	well-known and ubiquitous approach for reducing the overhead of
+	I/O operations is to use the Linux page cache.  It is thus crucial
+	to study experimentally the impact of the Linux page cache on the
+	performance Big Data applications. Given the difficulties with and shortcomings of
+	real-world experiments, a popular approach for application
+	performance studies is to use simulation.  While many simulation
+	frameworks have been developed for this purpose, they do not 
+        simulate page caching fully, or even at all.  As a result, using
+	these frameworks for studying the performance of data-intensive
+	applications leads inaccurate simulation results.
 
+        % Some cosmetic rewrites below
 	In this paper, we propose a I/O simulation model that characterizes
-        the key features of the page cache. We have implemented the model
-        with SimGrid, a simulation framework for distributed systems, using
-        WRENCH, a SimGrid-based workflow simulator. Our simulator is
-        compatible with both single-threaded and multithreaded
-        applications, and can simulate both writeback and writethrough
-        cache that can factor into local and network-based filesystems. We
-        evaluate it in different conditions, including sequential and
-        concurrent applications, as well as local and remote I/Os.
-        The page cache model reduces the simulation
-        error by up to an order of magnitude compared to cacheless simulations.
+	the key features of the page cache. We have implemented this model
+	as  part of the WRENCH workflow simulation framework, which itself
+	builds on the popular SimGrid distributed systems simulation
+	framework.  Our model and its implementation enable the simulation
+	of both single-threaded and multithreaded applications, and of both
+	writeback and writethrough caches for local or network-based
+	filesystems. We evaluate the accuracy of our model in different
+	conditions, including sequential and concurrent applications, as
+	well as local and remote I/Os.  We find that our page cache model
+	reduces the simulation error by up to an order of magnitude when
+	compared to state-of-the-art, cacheless simulations.
 
@@
         The Linux Page Cache plays an important role in
-        the reduction of filesystem data transfer costs. With the page cache, previously
-        read data can be reread directly from memory and written data can be written to
+        reducing filesystem data transfer times. With the page cache, previously
+        read data can be re-read directly from memory and written data can be written to
         memory before being asynchronously flushed to disk, resulting in improved I/O performance
-        on slower storage devices. The performance benefits that can be obtained from using the page
+        on slower storage devices. The magnitude of the performance benefits that can be obtained from using the page
         cache are determined by many factors including the total amount of memory,
-        the amount of data being written (i.e. dirty data), and the percentage of available memory available for
-        written data. All these factors are important when determining the impact of data transfers on
-        application performance, particularly in data intensive applications.
+        the amount of data being written (i.e., dirty data), and the percentage of available memory available for
+        written data. All these factors are important when determining the impact of I/O on
+        application performance, particularly in data-intensive applications.
 
-        Characterizing the performance of applications directly on computing infrastructure
-        has various shortcomings, which include, limited and shared resources, operational
-        costs, unstable environments, and the difficulties in producing reproducible results.
-        Simulations address these concerns by providing abstractions for the performance of computer
-        hardware, such as CPU, network and storage. Consequently, simulations provide a
-        cost-effective, fast, easy and reproducible  way evaluate application performance
-        on different infrastructures without having to execute the code.
-
-        The amount of data-intensive applications has been on the rise as a result of
+        % Swapped the order the two paragraphs and merged them into one
+        The number of data-intensive applications has been steadily rising as a result of
         open-data and data sharing initiatives. Due to the sheer size of the data being
         processed, these applications must be executed on large scale infrastructure
-        such as High Performance Computing (HPC) clusters or the cloud. Determining
-        which type of infrastructure is the most cost-effective for the particular
-        application can be tricky without the help of a simulator.
+	such as High Performance Computing (HPC) clusters or the cloud.  It
+	is thus crucial to quantify the performance of these applications
+	on these platforms. The goals include determining which type of hardware/software
+	stacks are best suited to different application classes, as well as
+	understanding the shortcomings of current algorithms, designs and
+	technologies so as to propose new solutions. Unfortunately, application performance
+        studies via experiments on real-world compute platforms
+        faces several difficulties (high operational costs, labor-intensive experimental setups,
+        shared platforms with dynamic loads that hinder reproductibilty of results) and shortcomings
+        (experiments are limited to the available platform/softare configurations, which precludes
+        the exploration of hypothetical scenarios). 
+	Simulations address these concerns by providing simulation models
+	and abstractions for the performance of computer hardware, such as
+	CPU, network and storage. Consequently, simulations provide a
+	cost-effective, fast, easy and reproducible way to evaluate
+	application performance on arbitrary platform configurations. It thus comes
+        as no surprise that, considering here only the field of parallel and distributed
+        computing, a large number of simulation frameworks have been developed and used
+        to support research and development endeavors~\cite{LAUNDRYLISTOFREFS_THAT_HENRI_WILL_PUT_IN}.
 
-        Page cache simulation is crucial when simulating data intensive applications,
-        which rely on cache hits and writing to memory for additional performance.
-        Furthermore, simulating the page cache enables us to study important application-I/O
-        bottlenecks which can arise over the course of execution, such as page cache
+        Page caching is a well-known and ubiquitous technique for mitigating the I/O bottleneck
+        faced by data-intensive applications. As such, it is necessary to simulate page caching
+        when performing performance studies in simulation.
+        Furthermore, simulating the page cache makes it possible to study relevant application-I/O
+        bottlenecks that can arise over the course of execution, such as page cache
         starvation~\cite{zhuang2017}.
-        While existing simulators can adequately characterize many different computing environments,
-        they lack much of the ability to adequately describe the page cache. Some
-        simulators try to simulate I/O with page cache, but fail to capture the key features such
+        While existing simulation frameworks of parallel and distributed computing
+        systems  capture many relevant features of hardware/software stacks of interest, 
+        they lack the ability to simulate page caches adequately. Those frameworks
+        that implement some page cache simulation
+        fail to capture key features such
         as dirty data and cache eviction policies~\cite{nunez2012simcan,nunez2012icancloud}. 
-        Other simulators, such as~\cite{xu2018saving}, capture the important features of the
-        page cache, but are domain-specific.
+        Some simulators, such as that in~\cite{xu2018saving}, do capture such features,
+        but are domain-specific.
 
@@
 
-        In this paper, we present WRENCH-cache, a simulation model of I/O with page
-        cache incorporated into WRENCH~\cite{casanova2018wrench}, a workflow management system simulator based on
-        the popular SimGrid framework. The goal of WRENCH-cache is to provide users
-        with an accurate representation of a workflow execution, when file system page
-        cache is used, while providing a good balance between accuracy and scalability.
-        WRENCH-cache is able to simulate both writeback and writethrough
-        cache which would normally be available to local and NFS-based filesystems.
-        WRENCH-cache can provide accurate simulations for large-scale parallel
-        applications using the page cache.
+	In this work, we present WRENCH-cache, a simulation model of I/O
+	with page cache incorporated into the WRENCH simulation framework~\cite{casanova2018wrench},
+	which is  based on the popular SimGrid distributed simulation framework~\cite{XXX}.
+	WRENCH-cache is to make it possible to simulate 
+        application execution accurately in the presence of page caching. Our specific 
+        contributions are as follows:
+    \begin{compactitem}
+        \item A page cache simulation model that supports 
+	both single-threaded and multithreaded applications, and of both
+	writeback and writethrough caches for local or network-based
+	filesystems;
+        \item An implementation of this model as part of the WRENCH framework; 
+        \item An evaluation and validation of the accuracy of our model, and of its implementation,
+              for a range of applications, platforms, and page cache configurations. 
+    \end{compactitem}
 

### 2020-12-09 paper_io_sim.tex
fixing paper structure and usage of latest IEEE template

--- 2020-12-09 paper_io_sim.tex
@@
 
-\author{\IEEEauthorblockN{Hoang-Dung Do$^1$, Val\'erie Hayot-Sasson$^1$, Rafael Ferreira da Silva$^3$, Henri Casanova$^2$, Tristan Glatard$^1$
+\author{
+  \IEEEauthorblockN{
+    Hoang-Dung Do\IEEEauthorrefmark{1},
+    Val\'erie Hayot-Sasson\IEEEauthorrefmark{1},
+    Rafael Ferreira da Silva\IEEEauthorrefmark{3},
+    Henri Casanova\IEEEauthorrefmark{2},
+    Tristan Glatard\IEEEauthorrefmark{1}
   }\\
   \IEEEauthorblockA{
-    $^1$Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada\\
-     $^2$Department of Information and Computer Sciences, University of Hawai‘i at Manoa, USA\\
-     $^3$Department of Computer Science, University of Southern California, Los Angeles, USA
+    \IEEEauthorrefmark{1}Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada\\
+    \IEEEauthorrefmark{2}Department of Information and Computer Sciences, University of Hawai‘i at Manoa, USA\\
+    \IEEEauthorrefmark{3}Information Sciences Institute, University of Southern California, Marina Del Rey, CA, USA
   }
@@
 from the Canada Foundation for Innovation.
-\bibliographystyle{plain}
+
+\bibliographystyle{IEEEtran}
 \tristan{Check references, in particular capitalization}

### 2020-12-10 paper_io_sim.tex
Typo

--- 2020-12-09 paper_io_sim.tex
@@
 \usepackage{graphicx}
+\usepackage{paralist}
 \usepackage{algorithm}
@@
 
-\author{\IEEEauthorblockN{Hoang-Dung Do$^1$, Val\'erie Hayot-Sasson$^1$, Rafael Ferreira da Silva$^3$, Henri Casanova$^2$, Tristan Glatard$^1$
+\author{
+  \IEEEauthorblockN{
+    Hoang-Dung Do\IEEEauthorrefmark{1},
+    Val\'erie Hayot-Sasson\IEEEauthorrefmark{1},
+    Rafael Ferreira da Silva\IEEEauthorrefmark{3},
+    Henri Casanova\IEEEauthorrefmark{2},
+    Tristan Glatard\IEEEauthorrefmark{1}
   }\\
   \IEEEauthorblockA{
-    $^1$Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada\\
-     $^2$Department of Information and Computer Sciences, University of Hawai`i at M\=anoa, USA\\
-     $^3$Department of Computer Science, University of Southern California, Los Angeles, USA
+    \IEEEauthorrefmark{1}Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada\\
+    \IEEEauthorrefmark{2}Department of Information and Computer Sciences, University of Hawai`i at M\=anoa, USA\\
+    \IEEEauthorrefmark{3}Information Sciences Institute, University of Southern California, Marina Del Rey, CA, USA
   }
@@
 	which is  based on the popular SimGrid distributed simulation framework~\cite{XXX}.
-	WRENCH-cache is to make it possible to simulate 
+	The added value of WRENCH-cache is to make it possible to simulate 
         application execution accurately in the presence of page caching. Our specific 

### 2020-12-10 result/multi/process_result.py
Add simulation time results

@@ -168,4 +168,5 @@ def run_time():
     plt.ylabel("time (seconds)")
     plt.legend()
 
-    plt.show()
+    plt.savefig("figures/simulation_time.pdf", format="pdf")
+    plt.savefig("figures/simulation_time.svg", format="svg")

### 2020-12-10 paper_io_sim.tex
replaced tabs with spaces
Small edits

--- 2020-12-09 paper_io_sim.tex
@@
 
-            We also implemented our model in WRENCH, separately from existing WRENCH components. 
-            We exposed page cache simulation through a user API, activated using a command-line argument.
-            We used SimGrid 3.25 which is available at \url{https://framagit.org/simgrid/simgrid}.
-            To implement the simulator, we used the WRENCH 1.6 version at this commit \url{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}.
+            We also implemented our model as part of WRENCH, enhancing its internal implementation
+            and APIs with a page cache abstraction, acticated on-demand via a {\tt --pagecache}
+            command-line argument.
+For the experiments in this paper we use WRENCH 1.6 at this commit \url{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}, which uses
+            SimGrid 3.25 (available at \url{https://framagit.org/simgrid/simgrid}).
+However, our implementation is now part of WRENCH's master branch and will be available to users with the next WRENCH release.
 
@@
         memory, local disk, remote disk (NFS), and network
-        (Table~\ref{table:benchmark}). Since WRENCH currently only supports
+        (Table~\ref{table:benchmark}). Since SimGrid, and thus WRENCH, currently only supports
         symmetrical bandwidths, we use the mean of the read and write

### 2020-12-10 paper_io_sim.tex

--- 2020-12-10 paper_io_sim.tex
@@
 \newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:}#1\color{black}}
+\newcommand{\henri}[1]{\color{green}\textbf{From Henri:}#1\color{black}}
 \newcommand{\english}[1]{\uwave{#1}}
@@
 
-            We also implemented our model in WRENCH, separately from existing WRENCH components.
-            We exposed page cache simulation through a user API, activated using a command-line argument.
-            We used SimGrid 3.25 which is available at \url{https://framagit.org/simgrid/simgrid}.
-            To implement the simulator, we used the WRENCH 1.6 version at this commit \url{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}.
+            We also implemented our model as part of WRENCH, enhancing its internal implementation
+            and APIs with a page cache abstraction, and allowing users to activate the feature via a
+            command-line argument.
+For the experiments in this paper we use WRENCH 1.6 at this commit \url{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}, which uses
+            SimGrid 3.25 (available at \url{https://framagit.org/simgrid/simgrid}).
+However, our implementation is now part of WRENCH's master branch and will be available to users with the next WRENCH release.
 
 \tristan{anything else interesting to mention about the implementation?}
+\henri{I mentioned that the implementaion is now part of the WRENCH master }
 

### 2020-12-10 paper_io_sim.tex
updated title
Added simulation time results

--- 2020-12-10 paper_io_sim.tex
@@
 
-%\title{Simulating the Linux page cache in distributed applications with WRENCH}
-\title{Simulating the Linux page cache in distributed applications with WRENCH}
+\title{Modeling the Linux page cache for accurate simulation of data-intensive applications}
 
@@
         \subsection{Real application}
+
         \subsection{Overhead comparison}
-        \tristan{Dzung, add your overhead measures here.}
+        As is the case for WRENCH, simulation time with WRENCH-cache scales
+        linearly with the number of concurrent applications
+        (Figure~\ref{fig:multi_time}, p \textless $10^{-24}$). However, the page
+        cache model substantially increases the simulation time by
+        application as can be seen by comparing regression slopes in
+        Figure~\ref{fig:multi_time}. Interestingly, WRENCH-cache is faster with 
+        NFS I/Os than with local I/Os, most likely due to the use of writethrough
+        cache in NFS, which by-passes flushing operations.
+        \begin{figure}
+            \begin{subfigure}{\columnwidth}
+                \centering
+                \includegraphics[width=\linewidth]{result/multi/figures/simulation_time.pdf}
+            \end{subfigure}
+            \caption{Simulation time comparison. WRENCH-cache scales
+            linearly with the number of concurrent applications, albeit
+            with a higher overhead than WRENCH.}
+            \label{fig:multi_time}
+            \end{figure}
 

### 2020-12-10 gen_fig.py

@@ -25,5 +25,5 @@ import result.multi.process_result as multi_result
 multi_result.result_local(rep_no=5)
 # Generate figures of experiment with NFS
 multi_result.result_nfs(rep_no=5)
-# Generate figures of simulation time
+# runtime plots
 multi_result.run_time()

### 2020-12-10 result/multi/process_result.py

@@ -6,6 +6,10 @@ import json
 import seaborn as sns
 
 
+real_color = "#90C987"
+wrench_color = "#994F88"
+wrench_cache_color = "#1965B0"
+
 def parse_single_pipeline(filename):
     """
     :param filename:
@@ -86,14 +90,14 @@ def suplot_prop(ax, real_dir, wrench_dir, propname, title, ylabel, rep_no=1):
 
     ax.set_title(title)
 
-    ax.plot(no_pipeline_df, mean_df / no_pipeline_df, color="#90C987", linewidth=2, label="Real execution mean")
+    ax.plot(no_pipeline_df, mean_df / no_pipeline_df, color=real_color, linewidth=2, label="Real execution mean")
     # ax.plot(no_pipeline_df, max_df / no_pipeline_df, color="green", linewidth=1, alpha=0.5)
     # ax.plot(no_pipeline_df, min_df / no_pipeline_df, color="green", linewidth=1, alpha=0.5)
-    ax.fill_between(no_pipeline_df, min_df / no_pipeline_df, max_df / no_pipeline_df, facecolor='#CAE0AB', alpha=0.5,
+    ax.fill_between(no_pipeline_df, min_df / no_pipeline_df, max_df / no_pipeline_df, facecolor=real_color, alpha=0.5,
                     label="Real execution min-max interval")
 
-    ax.plot(no_pipeline_df, simg_org_df[propname] / no_pipeline_df, linewidth=2, color="#994F88", label="WRENCH")
-    ax.plot(no_pipeline_df, simg_ext_df[propname] / no_pipeline_df, linewidth=2, color="#1965B0", label="WRENCH-cache")
+    ax.plot(no_pipeline_df, simg_org_df[propname] / no_pipeline_df, linewidth=2, color=wrench_color, label="WRENCH")
+    ax.plot(no_pipeline_df, simg_ext_df[propname] / no_pipeline_df, linewidth=2, color=wrench_cache_color, label="WRENCH-cache")
 
     ax.set_xlabel("Concurrent applications")
     if ylabel:
@@ -147,25 +151,57 @@ def result_nfs(rep_no=1):
 
 
 def run_time():
+    # WRENCH data
     local_org_df = pd.read_csv("local/wrench/run_time_original.csv")
-    local_cache_df = pd.read_csv("local/wrench/run_time_pagecache.csv")
     nfs_org_df = pd.read_csv("nfs/wrench/run_time_original.csv")
[…]

### 2020-12-10 paper_io_sim.tex
Added bib references

@@
         computing, a large number of simulation frameworks have been developed and used
-        to support research and development endeavors~\cite{LAUNDRYLISTOFREFS_THAT_HENRI_WILL_PUT_IN}.
+        to support research and development endeavors~\cite{ optorsim, gridsim, groudsim, cloudsim, nunez2012simcan,nunez2012icancloud, dissect_cf, cloudnetsimplusplus, fognetsimplusplus, casanova2014simgrid, ROSS, casanova2020fgcs}.
 
@@
     In this work, we present WRENCH-cache, a simulation model of I/O
-    with page cache incorporated into the WRENCH simulation framework~\cite{casanova2018wrench},
-    which is  based on the popular SimGrid distributed simulation framework~\cite{XXX}.
+    with page cache incorporated into the WRENCH simulation framework~\cite{casanova2020fgcs},
+    which is  based on the popular SimGrid distributed simulation framework~\cite{casanova2014simgrid}.
     The added value of WRENCH-cache is to make it possible to simulate 

### 2020-12-10 paper_io_sim.tex
Tweaks for Section 1 to remove "orphan words" that caused line wastage.

@@
         memory before being asynchronously flushed to disk, resulting in improved I/O performance
-        on slower storage devices. The magnitude of the performance benefits that can be obtained from using the page
-        cache are determined by many factors including the total amount of memory,
-        the amount of data being written (i.e., dirty data), and the percentage of available memory available for
+        on slower storage devices. The performance improvements
+        depend on many factors including the total amount of memory,
+        the amount of data being written (i.e., dirty data), and the amount of memory available for
         written data. All these factors are important when determining the impact of I/O on
@@
         While existing simulation frameworks of parallel and distributed computing
-        systems  capture many relevant features of hardware/software stacks of interest, 
+        systems  capture many relevant features of hardware/software stacks, 
         they lack the ability to simulate page caches adequately. Those frameworks
@@
         as dirty data and cache eviction policies~\cite{nunez2012simcan,nunez2012icancloud}. 
-        Some simulators, such as that in~\cite{xu2018saving}, do capture such features,
+        Some simulators, such as that in~\cite{xu2018saving}, do capture such features
         but are domain-specific.
@@
     filesystems;
-        \item An implementation of this model as part of the WRENCH framework; 
+        \item An implementation of this model in WRENCH; 
         \item An evaluation and validation of the accuracy of our model, and of its implementation,

### 2020-12-10 paper_io_sim.tex
edited methods

--- 2020-12-10 paper_io_sim.tex
@@
             If direct~I/O is enabled for a given file system, all I/O operations bypass the
-            page cache for writes, writing data directly to disk. In contrast, write-through enables the page cache
-            for file system writes, but the data is written synchronously from memory to disk. Writeback extends write-through
+            page cache for writes, writing data directly to disk. In contrast, writethrough enables the page cache
+            for file system writes, but the data is written synchronously from memory to disk. Writeback extends writethrough
             by enabling asynchronous writes to disk, meaning that a write call will return once
             the data has been written to memory.
-            The writeback strategy is considered to outperform write-through as well as
+            The writeback strategy is considered to outperform writethrough as well as
             direct I/O as it delays disk writes to perform a bulk write at a later time
@@
 
-    \section{Method}
+    \section{Methods}
     \label{method}
-    We separate our simulation model in two components, the IO
+    We separate our simulation model in two components, the I/O
     Controller and the Memory Manager, which together simulate
@@
     To read or write a file chunk, a simulated application sends a
-    request to the IO Controller. The IO Controller interacts as needed with
-    the Memory Manager to free memory by triggering flushing or eviction,
-    and interact with cached data. The Memory Manager
+    request to the I/O Controller. The I/O Controller interacts as needed with
+    the Memory Manager to free memory through flushing or eviction,
+    and to read or write cached data. The Memory Manager
     implements these operations, simulates periodical flushing
-    and eviction from cache, and reads or writes to disk when necessary.
-    In case the write-through strategy is used, the IO Controller can also interact directly
-    with disk model to simulate disk writes.
+    and eviction, and reads or writes to disk when necessary.
+    In case the writethrough strategy is used, the I/O Controller directly
+    writes to disk \tristan{Then how is the written data added to cache?}.
 
@@
            Applications send file read or write requests to the
-           IO Controller that orchestrates flushing, eviction, cache
+           I/O Controller that orchestrates flushing, eviction, cache
            and disk accesses with the Memory Manager. Concurrent accesses to storage
@@
     use existing storage simulation models~\cite{lebre2015} to simulate disk and
-    memory, characterized by its storage capacity, read and write
+    memory, characterized by their storage capacity, read and write
     bandwidths, and latency. Reusing existing storage models allows us to
@@
     In the Linux kernel, page cache LRU lists contain file pages. However,
-    due to the large number of file pages, managing lists of pages
+    due to the large number of file pages, simulating lists of pages
     induces substantial overheads.
@@
     page cache that were accessed in the same I/O operation.
-    A data block has information about file name, block size, last access
+    A data block has information about the file name, block size, last access
     time, a dirty flag that represents whether the data is clean (0)
@@
            \includegraphics[width=\columnwidth]{figures/lru_lists.pdf}
-           \caption{Model of page cache LRU lists with data blocks.}    \label{fig:lrulist}
+           \caption{Model of page cache LRU lists with data blocks.}
+           \label{fig:lrulist}
     \end{figure}
@@
     At any given time, a file can be partially cached, completely cached,
-    or uncached at all. A cached data block can only reside in one of two
-    page cache LRU lists. The first time they are accessed, blocks are
+    or uncached at all. A cached data block can only reside in one of the two
+    LRU lists. The first time they are accessed, blocks are
     added to the inactive list. On subsequent accesses, blocks of the
@@
 
-    \begin{algorithm}\caption{Periodical flushing simulation in IO Controller}\label{alg:pdflush}
+    \begin{algorithm}\caption{Periodical flushing simulation in I/O Controller}\label{alg:pdflush}
         \small
@@
                 \State flushing\_time = sm.write(blocks)
-                \State for blk in blocks: blk.dirty = 0
+                \State for blk in blocks: blk.dirty = 0 \tristan{with this logic a flushed block can remain dirty for a long time. Is it what you implemented?
+                it would be wise to set the dirty flag to 0 right after the block is written}
                 \If{flushing\_time $<$ t}
@@
     background thread. As in the Linux kernel, a dirty block
-    in our model is considered expired if (1) it is dirty, and (2)
+    in our model is considered expired if
     the duration since its entry time is longer than a
@@
     Periodical flushing is simulated as an infinite loop in which
-    the Memory Manager searches and flushes these expired dirty blocks to disk (Algorithm~\ref{alg:pdflush}).
+    the Memory Manager searches for dirty blocks and flushes them to disk (Algorithm~\ref{alg:pdflush}).
     % From Tristan: the algorithm is quite straightforward, I don't think this is
@@
 
-    \begin{algorithm}\caption{File chunk read simulation in IO Controller}
+    \begin{algorithm}\caption{File chunk read simulation in the I/O Controller}
     \label{alg:read}
@@
                \State required\_mem = cs + disk\_read
-               \State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn)
-               \State mm.evict(cs + disk\_read - mm.free\_mem, fn)
+               \State mm.flush(required\_mem - mm.free\_mem - mm.evictable, fn)
+               \State mm.evict(required\_mem - mm.free\_mem, fn)
                \If {disk\_read $>$ 0}  \Comment{Read uncached data}
@@
     round-robin fashion. To read a file chunk, simulated applications send
-    chunk read requests to the IO Controller which processes them using
+    chunk read requests to the I/O Controller which processes them using
     Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
     data that needs to be read from disk, and the remaining amount is read
-    from cache (line 6-7). The amount of memory required to read the chunk
+    from cache (line 7-8). The amount of memory required to read the chunk
     is calculated, corresponding to a copy of the chunk in anonymous memory
-    and a copy of the chunk in cache (line 8).
+    and a copy of the chunk in cache (line 9).
     If there is not enough available memory, the Memory Manager is called
-    to flush dirty data (line 9). If necessary, flushing is complemented by
-    eviction (line 10). Note that, when called with negative arguments, functions
-    \texttt{flush} and \texttt{evict} simply return and do not do anything. Then, if there is
-    uncached data, memory manager is called to read data from disk and add this
-    data to cache (line 12).
+    to flush dirty data (line 10). If necessary, flushing is complemented by
+    eviction (line 11). Note that, when called with negative arguments, functions
+    \texttt{flush} and \texttt{evict} simply return and do not do anything. Then,
+    if the block requires
+    uncached data, the memory manager is called to read data from disk and to add this
+    data to cache (line 14).
     If cached data needs to be read, the Memory Manager is called to simulate
-    a cache read  and update the corresponding data blocks accordingly (line 15).
+    a cache read  and update the corresponding data blocks accordingly (line 17).
     Finally, the memory manager is called to reduce the amount of free memory
-    corresponding to the amount of anonymous memory used by application (line 17).
+    corresponding to the amount of anonymous memory used by application (line 19).
 
-    \begin{algorithm}\caption{File chunk write simulation in IO Controller}
+    \begin{algorithm}\caption{File chunk write simulation in the I/O Controller}
     \label{alg:write}
@@
             \State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
-            \If {remain\_dirty $>$ 0} \Comment{Write with memory bandwidth}
+            \If {remain\_dirty $>$ 0} \Comment{Write to memory}
                 \State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
@@
             \State remaining = cs - mem\_amt
-            \While {remaining $>$ 0}  \Comment{Write with disk bandwidth}
+            \While {remaining $>$ 0}  \Comment{Flush to disk, then write to cache}
                 \State mm.flush(cs - mem\_amt)
@@
     Algorithm~\ref{alg:write} describes our simulation of chunk writes in
-    the IO Controller.
+    the I/O Controller.
     % Data is written to cache until the amount of dirty data in the cache
     % exceeds the dirty ratio times the amount of available memory. Data is then written to disk.
-    Our algorithm initially checks the  amount of data that
-    can be written as dirty data given the dirty ratio (line 5).
+    Our algorithm initially checks the  amount of dirty data that
+    can be written given the dirty ratio (line 5).
     If this amount is greater than 0, the Memory Manager is requested to evict
@@
     page cache is calculated (line 8), and a cache write is simulated (line 9).
-    If there is remaining data, this data is written to cache in a loop (line 12 - line 18).
-    In this loop, we flush and evict data as possible so that there will be
-    available space for writing (line 13-14).
-    After flushing and eviction, the amount of data that can be written
-    is calculated (line 15) and a cache write is simulated (line 16).
-    We update the remaining amount (line 17) and continue the loop until all data
-    is written.
+    If the dirty threshold is reached and there is remaining data to be written,
+    the remaining data is written to cache in a loop
+    where we repeatedly flush and evict to free space in the cache (line 12-18).
 
-    In addition to the chunk write simulation with writeback page cache above,
-    we also add a chunk write function with write-through strategy to the IOController.
-    This function simply simulates a disk write with the amount of data passed in,
-    then evicts cache if needed and adds this data to cache.
+    The write simulation described above models page cache in writeback
+    mode. Our model also includes a write function in writethrough mode,
+    which simply simulates a disk write with the amount of data passed in,
+    then evicts cache if needed and adds the written data to the cache.
 
@@
 
-            To validate our simulation model, we created a simple prototype
-            simulator independent of existing simulation frameworks and libraries.
-            This allowed us to evaluate the accuracy and correctness of our
-            model in a simple scenario before integrating it to the more complex
+            We first created a standalone prototype
+            simulator to evaluate the accuracy and correctness of our
+            model in a simple scenario before integrating it in the more complex
             WRENCH framework.
-            In this prototype we used the following basic storage model for
+            The prototype uses the following basic storage model for
             both memory and disk:
@@
 
-            Since bandwidth sharing was not simulated, this prototype does not support
+            Since bandwidth sharing is not simulated, the prototype does not support
             concurrency: it is limited to single-threaded applications running on systems
@@
 
-            We also implemented our model as part of WRENCH, enhancing its internal implementation
-            and APIs with a page cache abstraction, and allowing users to activate the feature via a
-            command-line argument.
-For the experiments in this paper we use WRENCH 1.6 at this commit \url{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}, which uses
-            SimGrid 3.25 (available at \url{https://framagit.org/simgrid/simgrid}).
-However, our implementation is now part of WRENCH's master branch and will be available to users with the next WRENCH release.
-
-\tristan{anything else interesting to mention about the implementation?}
-\henri{I mentioned that the implementaion is now part of the WRENCH master }
+            We also implemented our model as part of WRENCH, enhancing its
+            internal implementation and APIs with a page cache abstraction,
+            and allowing users to activate the feature via a command-line
+            argument. For the experiments in this paper we use WRENCH 1.6
+            at commit
+            \href{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}{6718537433},
+            which uses SimGrid 3.25, available at
+            \url{https://framagit.org/simgrid/simgrid}. Our implementation
+            is now part of WRENCH's master branch and will be available to
+            users with the next WRENCH release. WRENCH provides a full SimGrid-based simulation 
+            environment that supports, among other features, concurrent accesses to storage devices, 
+            applications distributed on multiple hosts, network transfers, 
+            and multi-threading.
 
@@
         multi-threaded applications, accessing data on local and remote
-        file systems.
+        file systems. We use two applications: a synthetic one created to
+        evaluate the simulation model, and a real one representative of
+        neuroimaging data processing.
 
-        We simulate an application that consists of three single-core, sequential tasks
-        where each task reads the file produced by the previous task,
-        increments every byte of this file to emulate real processing, and
-        writes the resulting data to disk.
-        In real execution, we clear cache before starting the application to make sure
-        the conditions are consistent, the cache is not in use and it is only used by
-        our application.
-        In addition, the anonymous memory used by application
-        is released after each task in the pipeline.
-        Thus, we also release the memory used at the end of each task in our simulator.
-        We also use a real application
-        \textcolor{red}{[Maybe briefly describe this application's purpose
-        (e.g. a standard fMRI preprocessing piepeline)]}, to evaluate the
-        applicability of our simulation model.
+        The synthetic application, implemented in C, consists of three single-core,
+        sequential tasks where each task reads the file produced by the
+        previous task, increments every byte of this file to emulate real
+        processing, and writes the resulting data to disk. In real
+        executions, we clear the page cache before starting the application,
+        to ensure comparable conditions among the runs.
+        The anonymous memory used by application
+        is released after each task, which we also simulate in the Python prototype and in WRENCH-cache.
 
-        The real application runs on a dedicated cluster hosted at
+        The real application is a workflow of the Nighres
+        toolbox~\cite{huntenburg2018nighres} that implements skull
+        stripping in high-resolution brain images, an operation common to
+        many analysis pipelines. The workflow is implemented as a Python
+        script that calls Java image-processing routines. We used Python
+        3.6, Java 8, and Nighres \tristan{add version}. We patched the
+        skull stripping workflow to monitor task CPU times and inject them
+        in the simulation, to focus our analysis on I/O times \tristan{add
+        link to patched code}.
+
+        The applications were executed on a dedicated cluster at
         Concordia University, with one login node, 9 compute nodes, and 4
@@
         compute node has 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
-        @ 2.10GHz, 275~GB (256~GiB) \tristan{are you sure it's not 256GB?
-        pls check.} of RAM, 6 $\times$ SSDs of 450~GB each with the XFS
-        file system, 378~GB of tmpfs, and 126~GB of devtmpfs file system.
+        @ 2.10GHz, 250~GiB of RAM, 6 $\times$ SSDs of 450~GiB each with the XFS
+        file system, 378~GiB of tmpfs, and 126~GiB of devtmpfs file system.
         Nodes run CentOS~8.1 with NFS version 4. We use the \texttt{atop}
@@
 
-        To parametrize our simulators, we benchmarked the bandwidths of
-        memory, local disk, remote disk (NFS), and network
+        To parametrize our simulators, we benchmarked the
+        memory, local disk, remote disk (NFS), and network bandwidths
         (Table~\ref{table:benchmark}). Since SimGrid, and thus WRENCH, currently only supports
@@
             As our focus is on I/O rather than compute, we measured
-            application task durations on a cluster node and used these
-            durations in our simulation. For the Python simulator,
-            execution times were directly used in the simulation. For
-            WRENCH and WRENCH-cache, we determined the corresponding number
-            of flops on a 1~Gflops CPU and used these values in the
-            simulation (see Table~\ref{table:cputime}). The final simulated
-            platform and application used in our experiment are avaiable at
+            application task CPU times on a cluster node
+            (Table~\ref{table:cputime}), and used these durations in our
+            simulations. For the Python simulator, we used CPU times
+            directly in the simulation. For WRENCH and WRENCH-cache, we
+            determined the corresponding number of flops on a 1~Gflops CPU
+            and used these values in the simulation. The final simulated
+            platform and application used in our experiments are avaiable at
             \url{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}.
@@
             \end{tabularx}
-            \caption{Task CPU time corresponding to input file size.}
+            \caption{Task CPU time corresponding to input file size. \tristan{are the input files in GB or in GiB? If GiB, then change the abbreviation everywhere.}
+            \tristan{add nighres CPU times when available}}
             \label{table:cputime}
@@
 
-            Our first experiment (\textit{Exp single-threaded}) simply runs one instance of the
-            application on a single cluster node, with different input file
-            sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed
-            to the same local disk. The goal is to validate the simulation model
-            in absence of any concurrent disk access.
+            Our first experiment (\textit{Exp single-threaded}) runs a
+            single instance of the synthetic application on a single
+            cluster node, with different input file sizes (20~GB, 50~GB,
+            75~GB, 100~GB), and with all I/Os directed to the same local
+            disk. The goal is to validate the simulation model in absence
+            of any concurrent disk access.
 
@@
             operating on different files of size 3~GB stored in the same
-            local disk, sharing the disk bandwidth. We vary the number of
+            local disk. We vary the number of
             concurrent application instances from 1 to 32 since cluster
@@
 
-            Our third experiment (\textit{Exp NFS}) occurs in the same configuration as the
-            previous one, albeit reading and writing on a 400-GB NFS-mounted remote disk
-            mounted from another compute node. As is commonly
-            configured in HPC environments to avoid data loss, in our simulator,
-            NFS client and server read caches are enabled, but there is no cache
-            for writing on client side and server writes are configured with
-            write-through cache instead of writeback.
+            Our third experiment (\textit{Exp NFS}) occurs in the same
+            configuration as the previous one, albeit reading and writing
+            on a 50-GiB NFS-mounted partition of a 450-GiB remote disk of
+            another compute node. As is commonly configured in HPC
+            environments to avoid data loss, there is no client write cache
+            and the server cache is configured as writethrough instead of
+            writeback. NFS client and server read caches are enabled. 
             Therefore, all the writes happen at disk bandwidth, but
-            reads can benefit from cache hits. We simulate write-through by
-            writing data directly to disk and adding that data to the server
-            cache.
+            reads can benefit from cache hits.
 
-            Finally, we simulate a real application from the neuroimaging
-            domain. \textcolor{red}{[Description of the real pipeline with
-            nighres (including the workflow engine if it's Dask)]}.
-            We simulate this pipeline with simulators using original WRENCH
-            and WRENCH-cache. Because our work focuses on I/O time, we
-            assume that CPU time is correctly modeled and use the CPU time
-            measured in the real pipelines to setup our simulated
-            application.
+            Our fourth and final experiment (\textit{Exp Nighres})
+            simulates the Nighres skull stripping workflow
+            \tristan{describe data, and if the execution was single node or
+            distributed.}
 
@@
         with reality (Figure~\ref{fig:single_cache}). With 20~GB files, the
-        simulated cache content exactly matches reality, owing to the fact
-        that all files fit in page cache. With 100~GB files, a slight
+        simulated cache content exactly matches reality, since all files fit in page cache. With 100~GB files, a slight
         discrepancy is observed after Write 2, which explains the
@@
 
-            The page cache model notably reduce WRENCH's simulation error
+            The page cache model notably reduces WRENCH's simulation error
             for concurrent applications and local I/Os
@@
             slightly overestimates reality, due to the discrepancy between
-            simulated and real read bandwidths mentioned before
-            (\tristan{check it's indeed mentioned. Dzung: is it correct?}). For writes,
+            simulated and real read bandwidths mentioned before. For writes,
             WRENCH-cache slightly underestimates reality for the same reason. The plateau
-            ending around 7 concurrent pipelines in the WRENCH-cache curve
+            ending around 7 concurrent pipelines in the WRENCH-cache writes
             marks the limit beyond which the page cache is saturated with dirty data
-            and needs flushing \tristan{why is there no plateay irl?}.
+            and needs flushing \tristan{why is there no plateau in real life?}.
 
@@
             on remote storage as well (Figure~\ref{fig:multi_nfs}). This
-            manifests only for reads, as only write-through cache was
+            manifests only for reads, as only writethrough cache was
             enabled in the NFS server. Both WRENCH and WRENCH-cache
@@
 
-        \tristan{discussion could be fleshed out}
+        \tristan{discussion could be fleshed out. cgroups}
 
@@
 \bibliographystyle{IEEEtran}
-\tristan{Check references, in particular capitalization}
 \bibliography{citation}

### 2020-12-10 paper_io_sim.tex

--- 2020-12-10 paper_io_sim.tex
@@
         \item An implementation of this model in WRENCH; 
-        \item An evaluation and validation of the accuracy of our model, and of its implementation,
+        \item An evaluation of the accuracy and scalability of our model, and of its implementation,
               for a range of applications, platforms, and page cache configurations. 

### 2020-12-10 paper_io_sim.tex

--- 2020-12-10 paper_io_sim.tex
@@
     filesystems;
-        \item An implementation of this model as part of the WRENCH framework; 
-        \item An evaluation and validation of the accuracy of our model, and of its implementation,
+        \item An implementation of this model in WRENCH; 
+        \item An evaluation of the accuracy and scalability of our model, and of its implementation,
               for a range of applications, platforms, and page cache configurations. 

### 2020-12-10 paper_io_sim.tex
Pass on the "simulation" part of the Related work section Added many bib references

--- 2020-12-10 paper_io_sim.tex
@@
 \usepackage{ulem}
+\usepackage{xspace}
 \usepackage{tabularx}
@@
 \newcommand{\english}[1]{\uwave{#1}}
+
+\newcommand{\simgrid}{SimGrid\xspace}
+\newcommand{\wrench}{WRENCH\xspace}
 
@@
     the key features of the page cache. We have implemented this model
-    as  part of the WRENCH workflow simulation framework, which itself
-    builds on the popular SimGrid distributed systems simulation
+    as  part of the \wrench workflow simulation framework, which itself
+    builds on the popular \simgrid distributed systems simulation
     framework.  Our model and its implementation enable the simulation
@@
         faces several difficulties (high operational costs, labor-intensive experimental setups,
-        shared platforms with dynamic loads that hinder reproductibilty of results) and shortcomings
-        (experiments are limited to the available platform/softare configurations, which precludes
+        shared platforms with dynamic loads that hinder reproducibility of results) and shortcomings
+        (experiments are limited to the available platform/software configurations, which precludes
         the exploration of hypothetical scenarios). 
@@
         computing, a large number of simulation frameworks have been developed and used
-        to support research and development endeavors~\cite{ optorsim, gridsim, groudsim, cloudsim, nunez2012simcan,nunez2012icancloud, dissect_cf, cloudnetsimplusplus, fognetsimplusplus, casanova2014simgrid, ROSS, casanova2020fgcs}.
+        to support research and development endeavors~\cite{ optorsim, gridsim, groudsim, cloudsim, nunez2012simcan,nunez2012icancloud, mdcsim, dissect_cf, cloudnetsimplusplus, fognetsimplusplus, casanova2014simgrid, ROSS, casanova2020fgcs}.
 
@@
 
-    In this work, we present WRENCH-cache, a simulation model of I/O
-    with page cache incorporated into the WRENCH simulation framework~\cite{casanova2020fgcs},
-    which is  based on the popular SimGrid distributed simulation framework~\cite{casanova2014simgrid}.
-    The added value of WRENCH-cache is to make it possible to simulate 
+    In this work, we present \wrench-cache, a simulation model of I/O
+    with page cache incorporated into the \wrench simulation framework~\cite{casanova2020fgcs},
+    which is  based on the popular \simgrid distributed simulation framework~\cite{casanova2014simgrid}.
+    The added value of \wrench-cache is to make it possible to simulate 
         application execution accurately in the presence of page caching. Our specific 
@@
     filesystems;
-        \item An implementation of this model in WRENCH; 
+        \item An implementation of this model in \wrench; 
         \item An evaluation of the accuracy and scalability of our model, and of its implementation,
@@
 
-            Many simulation frameworks exist to describe the behaviour of applications
-            running on the various infrastructures available to users, such as personal computers,
-            clouds and clusters.
-            These simulators are built with resource models such CPU, network and
-            storage, using various different approaches to describe the hardware.
-            For example, there is \textit{CloudSim} and \textit{iCanCloud} for packet-level network models
-            and \textit{DiskSim} the simulates storage hardware functioning \tristan{add references}.
+	    Many simulation frameworks have been developed to enable the
+	    simulation of parallel and distributed
+	    applications~\cite{optorsim, gridsim, groudsim, cloudsim,
+	    nunez2012simcan,nunez2012icancloud, mdcsim, dissect_cf,
+	    cloudnetsimplusplus, fognetsimplusplus, casanova2014simgrid,
+	    ROSS, casanova2020fgcs}. These frameworks implement simulation
+	    models and abstractions to aid the development of simulators
+	    for studying the functional and performance behaviors of
+	    application workloads executed on various hardware/software
+	    infrastructures.
 
-            Simulators must choose between scalability and accuracy when describing the
-            hardware~\cite{casanova2014simgrid}.
-            Fine-grained models allow simulators to achieve high accuracy,
-            but lead to scalability issues due to long simulation time
-            when simulated platforms and applications grow.
-            SimGrid is an open source simulation platform that has been developed
-            not only to balance the trade-off between accuracy and scalability,
-            but also to achieve versatility, allowing the use of the framework
-            to simulate applications in multiple domains. WRENCH, a workflow management
-            system simulator based on the SimGrid framework, provides a high-level API
-            for simulating workflows, thus facilitating the integration of new models.
-            \valerie{say more about wrench}
+	    The two main concerns for simulation are \emph{accuracy}
+	    (the ability to faithfully reproduce real-world executions) and
+	    \emph{scalability}  (the ability to simulate large/long real-world
+	    executions quickly and with low RAM footprint), and the above
+	    frameworks achieve different compromises between the two.  At
+	    one extreme are discrete-event models that capture
+	    ``microscopic'' behaviors of hardware/software systems (e.g.,
+	    packet-level network simulation, block-level disk simulation,
+	    cycle-accurate CPU simulation), which favor accuracy over
+	    speed.  At the other extreme are analytical models that capture
+	    ``macroscopic'' behaviors via mathematical models.  While these
+	    models lead to fast simulation, they must be developed
+	    carefully if high levels of accuracy are to be
+	    achieved~\cite{velhoTOMACS2013}.
 
-            Despite the fact that many sophisticated models for resources have been
-            integrated in simulators, and the page cache shows apparent impacts on I/O
-            performance, the simulation model of page cache is still missing in most
-            of simulators.
-            For example, SimGrid only simulates I/O operations with storage bandwidths
-            and capacity.
-            SIMCAN is a simulation framework that models page cache by storing data
-            accessed on disk in a block cache \cite{nunez2012simcan}.
-            Page cache is also modeled in the iCanCloud simulator through a component which
-            manages memory accesses and cached data \cite{nunez2012icancloud}. However,
-            the scalability of the iCanCloud simulator is limited by its sophisticated model.
-            Ultimately, none of these simulators provide any writeback cache simulator nor
-            cache eviction policys through LRU lists.
-            Although cache replacement policies are applied in \cite{xu2018saving} to simulate
-            in-memory caching, their simulator is specific to energy consumption of multi-tier
-            heterogenous networks.
+	    In this  work, we use the \simgrid and \wrench simulation
+	    frameworks.  The years of research and development invested in
+	    the popular \simgrid simulation framework~\cite{casanova2014simgrid}, have
+	    culminated in a set of state-of-the-art macroscopic simulation
+	    models that yield high accuracy, as demonstrated by
+	    (in)validation studies and comparisons to competing
+	    frameworks~\cite{smpi_validity, velhoTOMACS2013, simutool_09,
+	    nstools_07, lebre2015, pouilloux:hal-01197274,
+	    smpi_tpds2017,  7885814, 8048921, 7384330}.  But one
+	    significant drawback of \simgrid is that its simulation
+	    abstractions are low-level, meaning that implementing a
+	    simulator of complex systems can be
+	    labor-intensive~\cite{kecskemeti_2014}. To remedy this problem,
+	    the \wrench simulation framework~\cite{casanova2020fgcs}
+	    builds on top of \simgrid to provide higher-level simulation
+	    abstractions, so that simulators of complex applications and
+	    systems can be implemented with a few hundred lines.
 
-            In this study, we use WRENCH
-            to implement our page cache model. We chose WRENCH due to accurate yet scalable implementation, resulting
-            from its SimGrid backend, as well as its high-level API.
-            We have integrated our model in WRENCH and use the SimGrid API to interact with
-            simulated storage resources.
-            This approach allows us to leverage existing features in WRENCH and SimGrid such as process control and synchronization of
-            resource consumption.
-            We also leverage the existing SimGrid storage model with capacity
-            and bandwidth properties, which are similar to the characteristics of memory.
+	    Although the Linux page cache has a large impact on I/O
+	    performance, and thus on the execution of data-intensive
+	    applications, its simulation is rarely considered in the above
+	    frameworks.  Most frameworks merely simulate I/O operations
+	    based on storage bandwidths and capacities.  The SIMCAN
+	    framework does models page caching by storing data accessed on
+	    disk in a block cache~\cite{nunez2012simcan}.  Page cache is
+	    also modeled in iCanCloud through a component that manages
+	    memory accesses and cached data~\cite{nunez2012icancloud}.
+	    However, the scalability of the iCanCloud simulator is limited
+	    due to it using  microscopic simulation models.  Besides, none
+	    of these simulators provide any writeback cache simulator nor
+	    cache eviction policies through LRU lists.  Although cache
+	    replacement policies are applied in~\cite{xu2018saving} to
+	    simulate in-memory caching, their simulator is specific to
+	    energy consumption of multi-tier heterogeneous networks.
+
+            In this study, we develop a page cache simulation model which we implement
+            as part of the \wrench framework. We targeted \wrench because it is a
+            recent, actively developed framework that provides convenient simulation
+            abstractions, because it is easy to incorporate new simulation models
+            into it, and because it builds on \simgrid, whose scalable and accurate
+            simulation models we can thus reuse for our purpose.
 
@@
     At any given time, a file can be partially cached, completely cached,
-    or uncached at all. A cached data block can only reside in one of two
+    or not cached at all. A cached data block can only reside in one of two
     page cache LRU lists. The first time they are accessed, blocks are
@@
             model in a simple scenario before integrating it to the more complex
-            WRENCH framework.
+            \wrench framework.
             In this prototype we used the following basic storage model for
@@
 
-            We also implemented our model as part of WRENCH, enhancing its internal implementation
+            We also implemented our model as part of \wrench, enhancing its internal implementation
             and APIs with a page cache abstraction, and allowing users to activate the feature via a
             command-line argument.
-For the experiments in this paper we use WRENCH 1.6 at this commit \url{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}, which uses
-            SimGrid 3.25 (available at \url{https://framagit.org/simgrid/simgrid}).
-However, our implementation is now part of WRENCH's master branch and will be available to users with the next WRENCH release.
+For the experiments in this paper we use \wrench 1.6 at this commit \url{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}, which uses
+            \simgrid 3.25 (available at \url{https://framagit.org/simgrid/simgrid}).
+However, our implementation is now part of \wrench's master branch and will be available to users with the next \wrench release.
 
 \tristan{anything else interesting to mention about the implementation?}
-\henri{I mentioned that the implementaion is now part of the WRENCH master }
+\henri{I mentioned that the implementaion is now part of the \wrench master }
 
@@
         Our experiments compare real executions with our Python prototype,
-        with the original WRENCH, and with our WRENCH-cache extension. We
+        with the original \wrench, and with our \wrench-cache extension. We
         evaluate our page cache simulation in single-threaded and
@@
         memory, local disk, remote disk (NFS), and network
-        (Table~\ref{table:benchmark}). Since SimGrid, and thus WRENCH, currently only supports
+        (Table~\ref{table:benchmark}). Since \simgrid, and thus \wrench, currently only supports
         symmetrical bandwidths, we use the mean of the read and write
@@
             \hline
-                Bandwidths  & Cluster (real) & Python simulator & WRENCH simulator\\
+                Bandwidths  & Cluster (real) & Python simulator & \wrench simulator\\
             \hline
@@
             execution times were directly used in the simulation. For
-            WRENCH and WRENCH-cache, we determined the corresponding number
+            \wrench and \wrench-cache, we determined the corresponding number
             of flops on a 1~Gflops CPU and used these values in the
             simulation (see Table~\ref{table:cputime}). The final simulated
-            platform and application used in our experiment are avaiable at
+            platform and application used in our experiment are available at
             \url{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}.
@@
 
-            % Finally, because WRENCH simulates applications base on network-communication,
+            % Finally, because \wrench simulates applications base on network-communication,
             % we use an infinite bandwidth to eliminate network latency for local I/Os.
@@
             nighres (including the workflow engine if it's Dask)]}.
-            We simulate this pipeline with simulators using original WRENCH
-            and WRENCH-cache. Because our work focuses on I/O time, we
+            We simulate this pipeline with simulators using original \wrench
+            and \wrench-cache. Because our work focuses on I/O time, we
             assume that CPU time is correctly modeled and use the CPU time
@@
         as it only involves uncached data. Errors are reduced from an average
-        of 345\% in the original WRENCH to 46\% in the Python prototype and
-        39\% in WRENCH-cache. Unsurprisingly, the original WRENCH simulator
+        of 345\% in the original \wrench to 46\% in the Python prototype and
+        39\% in \wrench-cache. Unsurprisingly, the original \wrench simulator
         significantly overestimates read and write times, due to its lack
@@
 
-        WRENCH simulation errors are substantially lower with 100~GB
+        \wrench simulation errors are substantially lower with 100~GB
         files than with 20~GB files, due to the fact that part of the
         100~GB file needs to be read and written to disk, the only storage
-        device in WRENCH, as it does not fit in cache. Conversely,
-        simulation errors of the Python prototype and WRENCH-cache are higher with
-        100~GB files than with 20~GB files, due to idiosyncracies in the kernel
+        device in \wrench, as it does not fit in cache. Conversely,
+        simulation errors of the Python prototype and \wrench-cache are higher with
+        100~GB files than with 20~GB files, due to idiosyncrasies in the kernel
         flushing and eviction strategies.
@@
         reality \tristan{why?}. As expected, dirty data remains under the
-        dirty ratio. The Python prototype and WRENCH-cache exhibit nearly
+        dirty ratio. The Python prototype and \wrench-cache exhibit nearly
         identical memory profiles, which reinforces the confidence in our
@@
 
-            The page cache model notably reduce WRENCH's simulation error
+            The page cache model notably reduce \wrench's simulation error
             for concurrent applications and local I/Os
-            (Figure~\ref{fig:multi_local}). For reads, WRENCH-cache still
+            (Figure~\ref{fig:multi_local}). For reads, \wrench-cache still
             slightly overestimates reality, due to the discrepancy between
@@
             (\tristan{check it's indeed mentioned. Dzung: is it correct?}). For writes,
-            WRENCH-cache slightly underestimates reality for the same reason. The plateau
-            ending around 7 concurrent pipelines in the WRENCH-cache curve
+            \wrench-cache slightly underestimates reality for the same reason. The plateau
+            ending around 7 concurrent pipelines in the \wrench-cache curve
             marks the limit beyond which the page cache is saturated with dirty data
@@
             manifests only for reads, as only write-through cache was
-            enabled in the NFS server. Both WRENCH and WRENCH-cache
+            enabled in the NFS server. Both \wrench and \wrench-cache
             underestimate write times due to the discrepancy between
@@
         \subsection{Overhead comparison}
-        As is the case for WRENCH, simulation time with WRENCH-cache scales
+        As is the case for \wrench, simulation time with \wrench-cache scales
         linearly with the number of concurrent applications
@@
         application as can be seen by comparing regression slopes in
-        Figure~\ref{fig:multi_time}. Interestingly, WRENCH-cache is faster with 
+        Figure~\ref{fig:multi_time}. Interestingly, \wrench-cache is faster with 
         NFS I/Os than with local I/Os, most likely due to the use of writethrough
@@
             \end{subfigure}
-            \caption{Simulation time comparison. WRENCH-cache scales
+            \caption{Simulation time comparison. \wrench-cache scales
             linearly with the number of concurrent applications, albeit
-            with a higher overhead than WRENCH.}
+            with a higher overhead than \wrench.}
             \label{fig:multi_time}
@@
         this study, we proposed a simulation model of I/Os with page cache,
-        we implemented it in the SimGrid-based WRENCH library, and we
+        we implemented it in the \simgrid-based \wrench library, and we
         evaluated its benefits in different conditions. Results show that
@@
         \tristan{not sure what you mean}. Our page cache model is available
-        in WRENCH from commit \tristan{abcd}, ready to be used for the
+        in \wrench from commit \tristan{abcd}, ready to be used for the
         simulation of Big Data applications.
@@
         specific}. The availability of asymmetrical bandwidths in the
-        forthcoming SimGrid release will also directly improve our
+        forthcoming \simgrid release will also directly improve our
         simulation accuracy. Relevant future work includes simulating

### 2020-12-10 paper_io_sim.tex

--- 2020-12-10 paper_io_sim.tex
@@
     At any given time, a file can be partially cached, completely cached,
-    or uncached at all. A cached data block can only reside in one of the two
+    or not cached at all. A cached data block can only reside in one of two
     LRU lists. The first time they are accessed, blocks are
@@
             model in a simple scenario before integrating it in the more complex
-            WRENCH framework.
+            \wrench framework.
             The prototype uses the following basic storage model for
@@
 
-            We also implemented our model as part of WRENCH, enhancing its
+            We also implemented our model as part of \wrench, enhancing its
             internal implementation and APIs with a page cache abstraction,
             and allowing users to activate the feature via a command-line
-            argument. For the experiments in this paper we use WRENCH 1.6
+            argument. For the experiments in this paper we use \wrench 1.6
             at commit
             \href{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}{6718537433},
-            which uses SimGrid 3.25, available at
+            which uses \simgrid 3.25, available at
             \url{https://framagit.org/simgrid/simgrid}. Our implementation
-            is now part of WRENCH's master branch and will be available to
-            users with the next WRENCH release. WRENCH provides a full SimGrid-based simulation 
+            is now part of \wrench's master branch and will be available to
+            users with the next \wrench release. W\wrench provides a full \simgrid-based simulation 
             environment that supports, among other features, concurrent accesses to storage devices, 
@@
         memory, local disk, remote disk (NFS), and network bandwidths
-        (Table~\ref{table:benchmark}). Since SimGrid, and thus WRENCH, currently only supports
+        (Table~\ref{table:benchmark}). Since \simgrid, and thus \wrench, currently only supports
         symmetrical bandwidths, we use the mean of the read and write
@@
             simulations. For the Python simulator, we used CPU times
-            directly in the simulation. For WRENCH and WRENCH-cache, we
+            directly in the simulation. For \wrench and \wrench-cache, we
             determined the corresponding number of flops on a 1~Gflops CPU
@@
 
-            The page cache model notably reduces WRENCH's simulation error
+            The page cache model notably reduces \wrench's simulation error
             for concurrent applications and local I/Os
-            (Figure~\ref{fig:multi_local}). For reads, WRENCH-cache still
+            (Figure~\ref{fig:multi_local}). For reads, \wrench-cache still
             slightly overestimates reality, due to the discrepancy between
             simulated and real read bandwidths mentioned before. For writes,
-            WRENCH-cache slightly underestimates reality for the same reason. The plateau
-            ending around 7 concurrent pipelines in the WRENCH-cache writes
+            \wrench-cache slightly underestimates reality for the same reason. The plateau
+            ending around 7 concurrent pipelines in the \wrench-cache writes
             marks the limit beyond which the page cache is saturated with dirty data
@@
             manifests only for reads, as only writethrough cache was
-            enabled in the NFS server. Both WRENCH and WRENCH-cache
+            enabled in the NFS server. Both \wrench and \wrench-cache
             underestimate write times due to the discrepancy between

### 2020-12-10 paper_io_sim.tex
replaced tabs with spaces
Added Chris as co-author (nighres application)

@@
     Rafael Ferreira da Silva\IEEEauthorrefmark{3},
+    Christopher Steele\IEEEauthorrefmark{4}
     Henri Casanova\IEEEauthorrefmark{2},
@@
     \IEEEauthorrefmark{2}Department of Information and Computer Sciences, University of Hawai`i at M\=anoa, USA\\
-    \IEEEauthorrefmark{3}Information Sciences Institute, University of Southern California, Marina Del Rey, CA, USA
+    \IEEEauthorrefmark{3}Information Sciences Institute, University of Southern California, Marina Del Rey, CA, USA\\
+    \IEEEauthorrefmark{4}Department of Psychology, Concordia University, Montreal, Canada
   }
@@
         \cite{linuxdev3rd2010}.
-
         Cache eviction and flushing strategies are integral to proper page cache functioning.

### 2020-12-10 paper_io_sim.tex
small things

@@
     Rafael Ferreira da Silva\IEEEauthorrefmark{3},
-    Christopher Steele\IEEEauthorrefmark{4}
+    Christopher Steele\IEEEauthorrefmark{4},
     Henri Casanova\IEEEauthorrefmark{2},
@@
         stripping in high-resolution brain images, an operation common to
-        many analysis pipelines. The workflow is implemented as a Python
+        many neuroimage analysis pipelines. The workflow is implemented as a Python
         script that calls Java image-processing routines. We used Python

### 2020-12-11 paper_io_sim.tex
prettified tables

@@
 \usepackage{tabularx}
+\usepackage{booktabs}
+\usepackage{multirow}
 
@@
     process large datasets, the I/O bottleneck remains.  A
-    well-known and ubiquitous approach for reducing the overhead of
+    well-known approach for reducing the overhead of
     I/O operations is to use the Linux page cache.  It is thus crucial
     to study experimentally the impact of the Linux page cache on the
-    performance Big Data applications. Given the difficulties with and shortcomings of
+    performance of Big Data applications. Given the difficulties with and shortcomings of
     real-world experiments, a popular approach for application
@@
     these frameworks for studying the performance of data-intensive
-    applications leads inaccurate simulation results.
+    applications leads to inaccurate simulation results.
 
-        % Some cosmetic rewrites below
     In this paper, we propose a I/O simulation model that characterizes
@@
         open-data and data sharing initiatives. Due to the sheer size of the data being
-        processed, these applications must be executed on large scale infrastructure
+        processed, these applications must be executed on large-scale infrastructures
     such as High Performance Computing (HPC) clusters or the cloud.  It
@@
         studies via experiments on real-world compute platforms
-        faces several difficulties (high operational costs, labor-intensive experimental setups,
+        face several difficulties (high operational costs, labor-intensive experimental setups,
         shared platforms with dynamic loads that hinder reproducibility of results) and shortcomings
@@
         the exploration of hypothetical scenarios). 
-    Simulations address these concerns by providing simulation models
+    Simulations address these concerns by providing models
     and abstractions for the performance of computer hardware, such as
-    CPU, network and storage. Consequently, simulations provide a
+    CPU, network and storage. As a result, simulations provide a
     cost-effective, fast, easy and reproducible way to evaluate
@@
 
-        Page caching is a well-known and ubiquitous technique for mitigating the I/O bottleneck
+        Page caching is an ubiquitous technique for mitigating the I/O bottleneck
         faced by data-intensive applications. As such, it is necessary to simulate page caching
@@
         systems  capture many relevant features of hardware/software stacks, 
-        they lack the ability to simulate page caches adequately. Those frameworks
+        they lack the ability to simulate page cache adequately. Those frameworks
         that implement some page cache simulation
@@
         To improve performance of disk accesses, the page cache was introduced to the Linux kernel. %It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
-        The goal of the page cache is to try ensure that most I/O operations occur entirely
-        within memory space rather than synchronously to and from the storage device.
+        The goal of the page cache is to ensure that most I/O operations occur entirely
+        within memory space rather than synchronously to and from storage devices.
         All disk-backed data located in the page cache are partitioned into many small pages.
@@
         If it is the case, data is read entirely from memory (i.e. \textit{cache hit}). Otherwise,
-        the necessary pages must be loaded to memory from the disk (i.e. \textit{cache miss})
+        the necessary pages must be loaded to memory from the disk (i.e. \textit{cache miss}).
         Conversely, for writing, one of three strategies can be applied: direct~I/O (or, ``no-write''),
@@
         \emph{scalability}  (the ability to simulate large/long real-world
-        executions quickly and with low RAM footprint), and the above
+        executions quickly and with low RAM footprint). The above
         frameworks achieve different compromises between the two.  At
@@
         bandwidth values in our experiments.
-            \begin{table}[htbp]
+            \begin{table}
             \centering
-            \begin{tabularx}{\columnwidth}{|l
-            |>{\centering\arraybackslash}X
-            |>{\centering\arraybackslash}X
-            |>{\centering\arraybackslash}X|}
-            \hline
-                Bandwidths  & Cluster (real) & Python simulator & \wrench simulator\\
-            \hline
-                Memory read  & 6860    & 4812     & 4812\\
-                Memory write & 2764    & 4812 & 4812\\
-                Local disk read & 510 & 465 & 465\\
-                Local disk write & 420 & 465     & 465\\
-                Remote disk read & 515 & - & 445\\
-                Remote disk write & 375 & - & 445\\
-                Network & 3000 & - & 3000\\
-            \hline
+            \begin{tabularx}{\columnwidth}{ll
+            >{\centering\arraybackslash}X
+            >{\centering\arraybackslash}X
+            >{\centering\arraybackslash}X}
+            \toprule
+                \multicolumn{2}{c}{Bandwidths}  & Cluster (real) & Python prototype & \wrench simulator\\
+            \midrule
+            \multirow{2}{*}{Memory}      & read  & 6860 & 4812 & 4812\\
+                                         & write & 2764 & 4812 & 4812\\
+            \multirow{2}{*}{Local disk}  & read  & 510  & 465  & 465\\
+                                         & write & 420  & 465  & 465\\
+            \multirow{2}{*}{Remote disk} & read  & 515  & -    & 445\\
+                                         & write & 375  & -    & 445\\
+            \multicolumn{2}{l}{Network}  & 3000  & -    & 3000\\
+            \bottomrule
             \end{tabularx}
@@
             measured in the cluster.
-            Remote disk and network accesses are not simulated in the Python simulator. \tristan{table could be prettified}}
+            Remote disk and network accesses are not simulated in the Python simulator.}
             \label{table:benchmark}
@@
 
-            \begin{table}[htbp]
+            \begin{table}[b]
             \centering
-            \begin{tabularx}{0.8\columnwidth}{|l|>{\centering\arraybackslash}X|}
-            \hline
+            \begin{tabularx}{0.8\columnwidth}{c>{\centering\arraybackslash}X}
+            \toprule
                 Input size (GB)  & CPU time (second)\\
-            \hline
+            \midrule
                 3      & 4.4 \\
@@
                 100  & 155 \\
-            \hline
+            \bottomrule
             \end{tabularx}
-            \caption{Task CPU time corresponding to input file size. \tristan{are the input files in GB or in GiB? If GiB, then change the abbreviation everywhere.}
+            \caption{Task CPU times used in the simulation. \tristan{are the input files in GB or in GiB? If GiB, then change the abbreviation everywhere.}
             \tristan{add nighres CPU times when available}}

### 2020-12-11 paper_io_sim.tex
Update some text according to comments

@@
     and eviction, and reads or writes to disk when necessary.
-    In case the writethrough strategy is used, the I/O Controller directly
-    writes to disk \tristan{Then how is the written data added to cache?}.
+    In case the writethrough strategy is used, the I/O Controller directly writes to disk, 
+    cache is flushed if needed and written data is added to page cache.
 
@@
                 \State blocks = expired\_blocks(exp, in) + expired\_blocks(exp, ac)
-                \State flushing\_time = sm.write(blocks)
-                \State for blk in blocks: blk.dirty = 0 \tristan{with this logic a flushed block can remain dirty for a long time. Is it what you implemented?
-                it would be wise to set the dirty flag to 0 right after the block is written}
+                \State flushing\_time = 0
+                \For{blk in blocks} 
+                  \State blk.dirty = 0 
+                  \State flushing\_time = flushing\_time + sm.write(blocks)
+                \EndFor
                 \If{flushing\_time $<$ t}
                     \State sleep(t - flushing\_time)
-                \EndIf  % Else: do nothing, continue the loop
+                \EndIf  
             \EndWhile
@@
         script that calls Java image-processing routines. We used Python
-        3.6, Java 8, and Nighres \tristan{add version}. We patched the
+        3.6, Java 8, and Nighres 1.3.0. We patched the
         skull stripping workflow to monitor task CPU times and inject them
-        in the simulation, to focus our analysis on I/O times \tristan{add
-        link to patched code}.
+        in the simulation, to focus our analysis on I/O times, the patched code is 
+        available at \url{https://github.com/dohoangdzung/nighres}.
 
@@
             \end{tabularx}
-            \caption{Task CPU times used in the simulation. \tristan{are the input files in GB or in GiB? If GiB, then change the abbreviation everywhere.}
+            \caption{Task CPU times used in the simulation.
+%            \tristan{are the input files in GB or in GiB? If GiB, then change the abbreviation everywhere.}
+%  Dzung: it's in GB
             \tristan{add nighres CPU times when available}}
@@
             Our fourth and final experiment (\textit{Exp Nighres})
-            simulates the Nighres skull stripping workflow
+            simulates the Nighres CRUISE cortical reconstruction workflow
             \tristan{describe data, and if the execution was single node or
@@
         of Read 3 where it slightly increases in simulation but not in
-        reality \tristan{why?}. As expected, dirty data remains under the
+        reality. Our explanation could be that we allocate anonymous memory 
+        chunk-by-chunk to application, while anonymous memory is allocated 
+        when as soon as task starts.
+        As expected, dirty data remains under the
         dirty ratio. The Python prototype and \wrench-cache exhibit nearly
@@
         simulation error previously mentioned in Read 3. In the real execution indeed,
-        File 3 is entirely cached after Write 2, whereas in the simulated execution, part
-        of it is on disk. This is due to the fact that the simulated flushing and eviction
-        was triggered after File 3 was already written, while in reality these operations
-        happened before File 3 was written \tristan{Dzung, pls check that. Also, why doesn't it happen irl?}.
+        File 3 is entirely cached after Write 2, whereas in the simulated execution, 
+        only a part of it is cached. 
+        This is due to the fact that the in simulation, data is flushed and evicted 
+        from the inactive list first, then from the active list. Thus, File 2, which is 
+        accessed again by Read 2 and is cached in the active list, is not flushed and evicted, 
+        while File 3 is flushed and evicted since it is accessed once only and cached 
+        in the inactive list. This may not correctly simulate what really happens 
+        in reality.
 

### 2020-12-11 result/nighres/parse_real_log.py

@@ -1,0 +1,52 @@
+import json
+
+
+def parse_time(logfile, sum_file):
+    with open(logfile, "r") as logfile:
+        log = json.load(logfile)
+
+        summary = {}
+        for task in log:
+            makespan = log[task]["makespan"][0]["duration"]
+
+            read_total = 0
+            write_total = 0
+            for fileread in log[task]["read"]:
+                read_total += fileread["duration"]
+            for filewrite in log[task]["write"]:
+                write_total += filewrite["duration"]
+
+            summary[task] = {
+                "cputime": makespan - read_total - write_total,
+                "read": read_total,
+                "write": write_total
+            }
+
+    with open(sum_file, "w") as outfile:
+        json.dump(summary, outfile)
+
+
+def parse_task(logfile, sum_file):
+    with open(logfile, "r") as logfile:
+        log = json.load(logfile)
+
+        summary = {}
+        for task in log:
+            _input = []
+            _output = []
+            for fileread in log[task]["read"]:
+                _input.append(fileread["filename"])
+            for filewrite in log[task]["write"]:
+                _output.append(filewrite["filename"])
+
+            summary[task] = {
+                "input": _input,
+                "output": _output
+            }
+
+    with open(sum_file, "w") as outfile:
+        json.dump(summary, outfile)
+
+
+parse_time("real/log/timelog.json", "real/log/summary_time.json")
+parse_task("real/log/timelog.json", "real/log/summary_file.json")

### 2020-12-11 paper_io_sim.tex

--- 2020-12-11 paper_io_sim.tex
@@
 
-        \subsection{Simulators}
+        \subsection{Simulation}
 
@@
     implements flushing, eviction, and cached I/Os synchronously, whereas
-    the second one, operating in the background, periodically searches for
-    expired dirty data in page cache LRU lists and flushes them to disk. We
+    the second one, which operates in the background, periodically searches for
+    expired dirty data in page cache LRU lists and flushes data to disk. We
     use existing storage simulation models~\cite{lebre2015} to simulate disk and
     memory, characterized by their storage capacity, read and write
-    bandwidths, and latency. Reusing existing storage models allows us to
-    simulate bandwidth sharing between concurrent memory accesses and concurrent disk accesses.
+    bandwidths, and latency. These models account for
+    bandwidth sharing between concurrent memory accesses and concurrent disk accesses.
 
@@
     due to the large number of file pages, simulating lists of pages
-    induces substantial overheads.
+    induces substantial overhead.
     Therefore, we introduce the concept of a data block as a unit to represent data
@@
     Because the chunk and block sizes may be different, there are situations
-    where a block is not entirely read.
+    in a block is not entirely read.
     In this case, the block is split in two smaller blocks and one of them is re-accessed.
@@
     block be partially flushed, the block is split in two blocks,
-    one that is flushed and one that remains dirty. Flushing time
-    to disk is simulated by the storage model.
+    one that is flushed and one that remains dirty. The time needed
+    to flush data to disk is simulated by the storage model.
 
@@
     and only one of them is evicted.
-    The cache eviction simulation does not contribute to the simulated time
+    The overhead of the cache eviction algorithm is not part of the simulated time
     since cache eviction time is negligible in real systems \tristan{a reference would be welcome}.
@@
 
-    The write simulation described above models page cache in writeback
+    The above models page cache in writeback
     mode. Our model also includes a write function in writethrough mode,
@@
 
-            Since bandwidth sharing is not simulated, the prototype does not support
+            This prototype does not simulate  bandwidth sharing and thus does not support
             concurrency: it is limited to single-threaded applications running on systems
@@
             is now part of \wrench's master branch and will be available to
-            users with the next \wrench release. W\wrench provides a full \simgrid-based simulation 
+            users with the upcoming \wrench 1.8 release. \wrench provides a full \simgrid-based simulation 
             environment that supports, among other features, concurrent accesses to storage devices, 

### 2020-12-11 paper_io_sim.tex
typos

@@
     the second one, which operates in the background, periodically searches for
-    expired dirty data in page cache LRU lists and flushes data to disk. We
+    expired dirty data in page cache LRU lists and flushes this data to disk. We
     use existing storage simulation models~\cite{lebre2015} to simulate disk and
@@
     Because the chunk and block sizes may be different, there are situations
-    in a block is not entirely read.
+    where a block is not entirely read.
     In this case, the block is split in two smaller blocks and one of them is re-accessed.

### 2020-12-12 paper_io_sim.tex

--- 2020-12-11 paper_io_sim.tex
@@
     implements flushing, eviction, and cached I/Os synchronously, whereas
-    the second one, operating in the background, periodically searches for
-    expired dirty data in page cache LRU lists and flushes them to disk. We
+    the second one, which operates in the background, periodically searches for
+    expired dirty data in page cache LRU lists and flushes this data to disk. We
     use existing storage simulation models~\cite{lebre2015} to simulate disk and
     memory, characterized by their storage capacity, read and write
-    bandwidths, and latency. Reusing existing storage models allows us to
-    simulate bandwidth sharing between concurrent memory accesses and concurrent disk accesses.
+    bandwidths, and latency. These models account for
+    bandwidth sharing between concurrent memory accesses and concurrent disk accesses.
 

### 2020-12-12 paper_io_sim.tex
Add links to nighres examples

@@
             Our fourth and final experiment (\textit{Exp Nighres})
-            simulates the Nighres CRUISE cortical reconstruction workflow
+            simulates the Nighres CRUISE cortical reconstruction workflow, which consists 
+            of four tasks. The first two tasks are skull stripping and MGDM segmentation, 
+            which are in this Nighres example: 
+            \url{https://nighres.readthedocs.io/en/latest/auto_examples/example_01_tissue_classification.html}. 
+            The last two steps are region extraction and CRUISE cortical reconstruction 
+            in this example: 
+            \url{https://nighres.readthedocs.io/en/latest/auto_examples/example_02_cortical_depth_estimation.html}.
             \tristan{describe data, and if the execution was single node or

### 2020-12-12 result/nighres/parse_real_log.py

@@ -32,12 +32,12 @@ def parse_task(logfile, sum_file):
 
         summary = {}
         for task in log:
-            _input = []
-            _output = []
+            _input = {}
+            _output = {}
             for fileread in log[task]["read"]:
-                _input.append(fileread["filename"])
+                _input[fileread["filename"]] = fileread["filesize"]
             for filewrite in log[task]["write"]:
-                _output.append(filewrite["filename"])
+                _output[filewrite["filename"]] = filewrite["filesize"]
 
             summary[task] = {
                 "input": _input,
@@ -48,5 +48,5 @@ def parse_task(logfile, sum_file):
         json.dump(summary, outfile)
 
 
-parse_time("real/log/timelog.json", "real/log/summary_time.json")
-parse_task("real/log/timelog.json", "real/log/summary_file.json")
+parse_time("real/log/timelog_sub001_sess1_.json", "real/log/__summary_time_sub001_sess1.json")
+parse_task("real/log/timelog_sub001_sess1_.json", "real/log/__summary_file_sub001_sess1.json")

### 2020-12-12 paper_io_sim.tex
Fix typo

--- 2020-12-12 paper_io_sim.tex
@@
             Our fourth and final experiment (\textit{Exp Nighres})
-            simulates the Nighres CRUISE cortical reconstruction workflow
+            simulates the Nighres CRUISE cortical reconstruction workflow, which consists 
+            of four steps. The first two steps are skull stripping and MGDM segmentation, 
+            which are in this Nighres example: 
+            \url{https://nighres.readthedocs.io/en/latest/auto_examples/example_01_tissue_classification.html}. 
+            The last two tasks are region extraction and CRUISE cortical reconstruction 
+            in this example: 
+            \url{https://nighres.readthedocs.io/en/latest/auto_examples/example_02_cortical_depth_estimation.html}.
             \tristan{describe data, and if the execution was single node or

### 2020-12-12 paper_io_sim.tex
Update commented text

--- 2020-12-12 paper_io_sim.tex
@@
             and used these values in the simulation. The final simulated
-            platform and application used in our experiments are avaiable at
-            \url{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}.
+            platform and application used in our experiments are available at
+            commit \href{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}{ec6b43561b}.
 
@@
             Our fourth and final experiment (\textit{Exp Nighres})
-            simulates the Nighres CRUISE cortical reconstruction workflow
+            simulates the Nighres CRUISE cortical reconstruction workflow, which consists 
+            of four steps. The first two steps are skull stripping and MGDM segmentation, 
+            which are in this Nighres \href{{https://nighres.readthedocs.io/en/latest/auto_examples/example_01_tissue_classification.html}}{tissue classification example}. 
+            The last two tasks are region extraction and CRUISE cortical reconstruction 
+            in the   
+            \href{https://nighres.readthedocs.io/en/latest/auto_examples/example_02_cortical_depth_estimation.html}{cortical depth estimation example}.
             \tristan{describe data, and if the execution was single node or
@@
         of Read 3 where it slightly increases in simulation but not in
-        reality. Our explanation could be that we allocate anonymous memory 
-        chunk-by-chunk to application, while anonymous memory is allocated 
-        when as soon as task starts.
-        As expected, dirty data remains under the
-        dirty ratio. The Python prototype and \wrench-cache exhibit nearly
+        reality. Our explanation could be that in reality, because File 3 is
+        entirely in cache, the cache amount does not increase when File 3 is read.
+        While in simulation, a part of File 3 is read from disk, leading to an increase
+        in cache amount followed by a decrease when eviction is called.
+        Apart from that, dirty data remains under the dirty ratio as expected. 
+        The Python prototype and \wrench-cache exhibit nearly
         identical memory profiles, which reinforces the confidence in our
@@
         while File 3 is flushed and evicted since it is accessed once only and cached 
-        in the inactive list. This may not correctly simulate what really happens 
-        in reality.
+        in the inactive list. 
+        This flushing and eviction order in simulation may not correctly reflect the 
+        order in reality and lead to the difference.
 
@@
             applications since before this threshold, most reads result in cache
-            hits \tristan{Dzung, pls check that.}.
+            hits.
+            % \tristan{Dzung, pls check that.}.
+            % Yes, disk reads appear from 23 apps indeed.
 
@@
         performance by a factor close to 10 (see results of the
-        single-threaded experiment). Furthermore, this level of accuracy
-        can be achieved without a fine-grained level of details
-        \tristan{not sure what you mean}. Our page cache model is available
-        in \wrench from commit \tristan{abcd}, ready to be used for the
-        simulation of Big Data applications.
+        single-threaded experiment). 
+        %Furthermore, this level of accuracy
+        %can be achieved without a fine-grained level of details
+        %\tristan{not sure what you mean}. 
+        % Dzung: I meant the results are accurate without a very detailed model. But the overhead results dot not seem to show that.
+        Our page cache model is available
+        in \wrench from commit \href{https://github.com/wrench-project/wrench/tree/d4e188d2f0cf69a19ae0a6b5e41530eb77d12cd8}{d4e188d2f0}, ready to be used 
+        for the simulation of Big Data applications.
 
         Simulation results could be made even more accurate by a deeper
-        investigation of cache eviction mechanisms \tristan{be more
-        specific}. The availability of asymmetrical bandwidths in the
+        investigation of page cache flushing and eviction order. 
+        The availability of asymmetrical bandwidths in the
         forthcoming \simgrid release will also directly improve our

### 2020-12-14 paper_io_sim.tex
Edits in results, started editing conclusion

@@
             and allowing users to activate the feature via a command-line
-            argument. For the experiments in this paper we use \wrench 1.6
-            at commit
+            argument. We used SimGrid's locking mechanism to handle
+            concurrent accesses to page cache LRU lists by the two Memory
+            Manager threads. For the experiments in this paper we use
+            \wrench 1.6 at commit
             \href{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}{6718537433},
@@
         previous task, increments every byte of this file to emulate real
-        processing, and writes the resulting data to disk. In real
-        executions, we clear the page cache before starting the application,
-        to ensure comparable conditions among the runs.
-        The anonymous memory used by application
-        is released after each task, which we also simulate in the Python prototype and in WRENCH-cache.
+        processing, and writes the resulting data to disk. Files are
+        numbered by ascending creation times: File 1 is file read by Task 1, and File \texttt{i+1} is the file
+        written by task \texttt{i}. In real executions, we clear the page
+        cache before starting the application, to ensure comparable
+        conditions among the runs. The anonymous memory used by application
+        is released after each task, which we also simulate in the Python
+        prototype and in WRENCH-cache.
 
@@
         The page cache simulation model drastically reduces I/O simulation
-        errors in the three successive application tasks (Figure~\ref{fig:single_error}). The first read is not impacted
+        errors in each application task (Figure~\ref{fig:single_error}). The first read is not impacted
         as it only involves uncached data. Errors are reduced from an average
@@
         39\% in \wrench-cache. Unsurprisingly, the original \wrench simulator
-        significantly overestimates read and write times, due to its lack
-        of page cache simulation model. Results with files of 50~GB and 75~GBs
-        show similar behaviors and are not reported due to space limitations.
+        significantly overestimates read and write times, due to the lack
+        of page cache simulation. Results with files of 50~GB and 75~GB
+        show similar behaviors and are not reported for brevity.
 
@@
         Simulated memory profiles are highly consistent with the real ones
-        (Figure~\ref{fig:single_memprof}). With 20~GB files, memory
-        profiles almost exactly match the real ones, although dirty data
-        seems to be flushing faster in real life than in simulation
-        \tristan{any idea why?}, a behavior also observed with 100~GB files. With
-        100~GB files, used memory reaches total memory during the first write,
-        triggering dirty data flushing. Used memory drops to cached memory
-        when application tasks release anonymous memory. Simulated cached
-        memory is highly consistent with real values, except toward the end
-        of Read 3 where it slightly increases in simulation but not in
-        reality. Our explanation could be that in reality, because File 3 is
-        entirely in cache, the cache amount does not increase when File 3 is read.
-        While in simulation, a part of File 3 is read from disk, leading to an increase
-        in cache amount followed by a decrease when eviction is called.
-        Apart from that, dirty data remains under the dirty ratio as expected. 
-        The Python prototype and \wrench-cache exhibit nearly
+        (Figure~\ref{fig:single_memprof}). Used memory consists of both
+        cached and anonymous memory that cannot be evicted at the current
+        time. With 20~GB files, memory profiles almost exactly match the
+        real ones, although dirty data seems to be flushing faster in real
+        life than in simulation \tristan{any idea why?}, a behavior also
+        observed with 100~GB files. With 100~GB files, used memory reaches
+        total memory during the first write, triggering dirty data
+        flushing. Used memory drops to cached memory when application tasks
+        release anonymous memory. Simulated cached memory is highly
+        consistent with real values, except toward the end of Read 3 where
+        it slightly increases in simulation but not in reality. This
+        occurred due to the fact that after Write 2, File 3 was only partly
+        cached in simulation whereas it was entirely cached in the real
+        system. In all cases, dirty data remains under the dirty ratio as
+        expected. The Python prototype and \wrench-cache exhibit nearly
         identical memory profiles, which reinforces the confidence in our
@@
         This is due to the fact that the in simulation, data is flushed and evicted 
-        from the inactive list first, then from the active list. Thus, File 2, which is 
+        from the inactive list first, then from the active list \tristan{whereas in the real execution ...?}. Thus, File 2, which is 
         accessed again by Read 2 and is cached in the active list, is not flushed and evicted, 
@@
         in the inactive list.
-        This flushing and eviction order in simulation may not correctly reflect the 
-        order in reality and lead to the difference.
 
@@
                    \vspace*{-0.7cm}
-                   \caption{Memory profiles \tristan{Remove trailing whitespace for 20GB graphs (xmax should be about 160s, not 200s)}}
+                   \caption{Memory profiles \tristan{Remove trailing
+                   whitespace for 20GB graphs (xmax should be about 160s,
+                   not 200s)} \tristan{Find a way to add Read 1, Write 1,
+                   Read 2, etc on the graphs, perhaps at the top of the
+                   real execution graphs}}
                    \vspace*{0.5cm}
@@
 
-        \subsection{Concurrent applications}
+        \subsection{Concurrent application (local storage)}
 
@@
             The page cache model notably reduces \wrench's simulation error
-            for concurrent applications and local I/Os
+            for concurrent applications executed with local I/Os
             (Figure~\ref{fig:multi_local}). For reads, \wrench-cache still
@@
             \wrench-cache slightly underestimates reality for the same reason. The plateau
-            ending around 7 concurrent pipelines in the \wrench-cache writes
+            ending around 7 concurrent applications in the \wrench-cache writes
             marks the limit beyond which the page cache is saturated with dirty data
@@
 
-        \subsection{Remote storage}
+        \subsection{Remote storage (NFS)}
 
@@
         (Figure~\ref{fig:multi_time}, p \textless $10^{-24}$). However, the page
-        cache model substantially increases the simulation time by
-        application as can be seen by comparing regression slopes in
+        cache model substantially increases simulation time by
+        application, as can be seen by comparing regression slopes in
         Figure~\ref{fig:multi_time}. Interestingly, \wrench-cache is faster with 
         NFS I/Os than with local I/Os, most likely due to the use of writethrough
-        cache in NFS, which by-passes flushing operations.
+        cache in NFS, which bypasses flushing operations.
         \begin{figure}
@@
     \label{discussion}
-        Page cache undoubtedly has important impacts on performance, in
-        particular for data-intensive applications. Numerical simulation is
-        an effective approach to predict and compare the performance of
-        applications, schedulers and infrastructures. However, page cache
-        is missing in many simulation tools for distributed systems. In
-        this study, we proposed a simulation model of I/Os with page cache,
-        we implemented it in the \simgrid-based \wrench library, and we
-        evaluated its benefits in different conditions. Results show that
-        our model very substantially improves simulation accuracy,
-        decreasing absolute relative errors between simulated and real
-        performance by a factor close to 10 (see results of the
-        single-threaded experiment). 
-        %Furthermore, this level of accuracy
-        %can be achieved without a fine-grained level of details
-        %\tristan{not sure what you mean}. 
-        % Dzung: I meant the results are accurate without a very detailed model. But the overhead results dot not seem to show that.
-        Our page cache model is available
-        in \wrench from commit \href{https://github.com/wrench-project/wrench/tree/d4e188d2f0cf69a19ae0a6b5e41530eb77d12cd8}{d4e188d2f0}, ready to be used 
-        for the simulation of Big Data applications.
+    We designed a model of the Linux page cache and implemented it in
+    the \simgrid-based \wrench library to simulate distributed
+    applications. Evaluation results show that our model very
+    substantially improves simulation accuracy, decreasing absolute
+    relative simulation errors up to 9$\times$ (see results of
+    the single-threaded experiment).
+    Our page cache model is publicly available
+    in the \wrench GitHub repository.
+
+    Page cache simulation would be instrumental in a number of studies.
+    For instance, it is now common for HPC clusters to run 
+    application jobs in Linux control groups (cgroups), where resource consumption 
+    including memory is controlled.
+
+    With the page cache, I/O times are strongly dependent on the amount of
+    available memory, which is itself impacted by the use of anonymous
+    memory by applications. A more accurate simulation of application
+    memory consumption would therefore be a relevant future direction to
+    further improve I/O simulation.
+
+    I/Os remain much more complex than is currently available in simulation
+    models.
 
@@
         forthcoming \simgrid release will also directly improve our
-        simulation accuracy. Relevant future work includes simulating
+        simulation accuracy.
+
+        Relevant future work includes simulating
         application memory consumption, random I/O, file readahead and

### 2020-12-14 paper_io_sim.tex
more in conclusion

@@
 
-    Page cache simulation would be instrumental in a number of studies.
-    For instance, it is now common for HPC clusters to run 
-    application jobs in Linux control groups (cgroups), where resource consumption 
-    including memory is controlled.
+    Page cache simulation would be instrumental in a number of studies. For
+    instance, it is now common for HPC clusters to run application jobs in
+    Linux control groups (cgroups), where resource consumption, including
+    memory and therefore page cache usage, is limited. Using our simulator,
+    it would be possible to model the interaction between memory allocation
+    and I/O performance, with the goal to improve scheduling algorithms.
+    Our simulator could also be utilized to evaluate solutions that
+    reduce the impact of network file transfers on distributed
+    applications, such as burst
+    buffers~\cite{ferreiradasilva-fgcs-bb-2019}, hierarchical file
+    systems~\cite{islam2015triple}, active storage, or specific
+    hardware architectures~\cite{hayot2020performance}.
 
-    With the page cache, I/O times are strongly dependent on the amount of
-    available memory, which is itself impacted by the use of anonymous
-    memory by applications. A more accurate simulation of application
-    memory consumption would therefore be a relevant future direction to
-    further improve I/O simulation.
+    However, I/Os remain much more complex than is currently available in
+    simulation models, which could drastically limit their accuracy. A more
+    accurate description of the use of anonymous memory by applications
+    would be an important extension directly related to our page
+    cache model. With the page cache indeed, I/O times are indeed strongly
+    dependent on the amount of available memory, which is itself impacted
+    by the use of anonymous memory by applications. File access patterns, 
+    including random I/O, might also be worth describing, as they directly impact the content of the page cache.
 
-    I/Os remain much more complex than is currently available in simulation
-    models.
+        % Simulation results could be made even more accurate by a deeper
+        % investigation of page cache flushing and eviction order. 
+        % The availability of asymmetrical bandwidths in the
+        % forthcoming \simgrid release will also directly improve our
+        % simulation accuracy.
 
-        Simulation results could be made even more accurate by a deeper
-        investigation of page cache flushing and eviction order. 
-        The availability of asymmetrical bandwidths in the
-        forthcoming \simgrid release will also directly improve our
-        simulation accuracy.
+        % Relevant future work includes simulating
+        % application memory consumption, random I/O, file readahead and
+        % persistent disk storage.
 
-        Relevant future work includes simulating
-        application memory consumption, random I/O, file readahead and
-        persistent disk storage.
-
-        \tristan{discussion could be fleshed out. cgroups}
+        % \tristan{discussion could be fleshed out. cgroups}
 

### 2020-12-15 paper_io_sim.tex
Fixed conclusion and references

--- 2020-12-14 paper_io_sim.tex
@@
     \label{discussion}
-    We designed a model of the Linux page cache and implemented it in
-    the \simgrid-based \wrench library to simulate distributed
-    applications. Evaluation results show that our model very
-    substantially improves simulation accuracy, decreasing absolute
-    relative simulation errors up to 9$\times$ (see results of
-    the single-threaded experiment).
-    Our page cache model is publicly available
-    in the \wrench GitHub repository.
+    We designed a model of the Linux page cache and implemented it in the
+    \simgrid-based \wrench library to simulate distributed applications.
+    Evaluation results show that our model improves simulation accuracy
+    substantially, decreasing absolute relative simulation errors up to
+    9$\times$ (see results of the single-threaded experiment). The
+    availability of asymmetrical disk bandwidths in the forthcoming
+    \simgrid release will further improve the accuracy of our simulations.
+    Our page cache model is publicly available in the \wrench GitHub
+    repository.
 
-    Page cache simulation would be instrumental in a number of studies.
-    For instance, it is now common for HPC clusters to run 
-    application jobs in Linux control groups (cgroups), where resource consumption 
-    including memory is controlled.
+    Page cache simulation would be instrumental in a number of studies. For
+    instance, it is now common for HPC clusters to run applications in
+    Linux control groups (cgroups), where resource consumption is limited, including
+    memory and therefore page cache usage. Using our simulator,
+    it would be possible to study the interaction between memory allocation
+    and I/O performance, for instance to improve scheduling algorithms.
+    Our simulator could also be utilized to evaluate solutions that
+    reduce the impact of network file transfers on distributed
+    applications, such as burst
+    buffers~\cite{ferreiradasilva-fgcs-bb-2019}, hierarchical file
+    systems~\cite{islam2015triple}, active storage~\cite{5496981}, or specific
+    hardware architectures~\cite{hayot2020performance}.
 
-    With the page cache, I/O times are strongly dependent on the amount of
-    available memory, which is itself impacted by the use of anonymous
-    memory by applications. A more accurate simulation of application
-    memory consumption would therefore be a relevant future direction to
-    further improve I/O simulation.
-
-    I/Os remain much more complex than is currently available in simulation
-    models.
-
-        Simulation results could be made even more accurate by a deeper
-        investigation of page cache flushing and eviction order. 
-        The availability of asymmetrical bandwidths in the
-        forthcoming \simgrid release will also directly improve our
-        simulation accuracy.
-
-        Relevant future work includes simulating
-        application memory consumption, random I/O, file readahead and
-        persistent disk storage.
-
-        \tristan{discussion could be fleshed out. cgroups}
+    However, I/Os remain much more complex than is currently available in
+    simulation models, which could drastically limit the accuracy of simulations.
+    Relevant extensions to this work could include more
+    accurate descriptions of anonymous memory usage by applications. 
+    With the page cache, I/O times are indeed strongly
+    dependent on the amount of available memory, which is itself impacted
+    by the use of anonymous memory. File access patterns, 
+    including random I/O, might also be worth including in application models,
+    as they directly impact page cache content.
+    % readahead and persistent storage could also be adeed
+    % Simulation results could be made even more accurate by a deeper
+    % investigation of page cache flushing and eviction order.
 

### 2020-12-15 paper_io_sim.tex
Edited abstract

@@
 
-    % Henri Rewrote quite  a bit in the next paragraph here to spell out a bit
-    % more why we need simulation
     The emergence of Big Data in recent years has resulted in a growing
-    need for efficient data storage management techniques.  While
-    infrastructures with sufficient compute power are available to
-    process large datasets, the I/O bottleneck remains.  A
-    well-known approach for reducing the overhead of
-    I/O operations is to use the Linux page cache.  It is thus crucial
-    to study experimentally the impact of the Linux page cache on the
-    performance of Big Data applications. Given the difficulties with and shortcomings of
-    real-world experiments, a popular approach for application
-    performance studies is to use simulation.  While many simulation
-    frameworks have been developed for this purpose, they do not 
-    simulate page caching fully, or even at all.  As a result, using
-    these frameworks for studying the performance of data-intensive
-    applications leads to inaccurate simulation results.
+    need for efficient data processing solutions. While infrastructures
+    with sufficient compute power are available,
+    the I/O bottleneck remains. The Linux page cache is an efficient
+    approach to reduce I/O overheads, but few
+    experimental studies of its interactions with Big Data applications exist,
+    partly due to limitations associated with
+    real-world experiments. Simulation is a popular approach to address
+    these issues, however, existing simulation frameworks do not simulate
+    page caching fully, or even at all.  As a result, simulation-based
+    performance studies of data-intensive applications leads to inaccurate
+    results.
 
-    In this paper, we propose a I/O simulation model that characterizes
+    In this paper, we propose an I/O simulation model that includes
     the key features of the page cache. We have implemented this model
@@
 
-        The Linux Page Cache plays an important role in
+        The Linux page cache plays an important role in
         reducing filesystem data transfer times. With the page cache, previously

### 2020-12-15 paper_io_sim.tex
edits in intro

@@
     experimental studies of its interactions with Big Data applications exist,
-    partly due to limitations associated with
+    partly due to limitations of
     real-world experiments. Simulation is a popular approach to address
@@
     In this paper, we propose an I/O simulation model that includes
-    the key features of the page cache. We have implemented this model
+    the key features of the Linux page cache. We have implemented this model
     as part of the \wrench workflow simulation framework, which itself
@@
         reducing filesystem data transfer times. With the page cache, previously
-        read data can be re-read directly from memory and written data can be written to
+        read data can be re-read directly from memory, and written data can be written to
         memory before being asynchronously flushed to disk, resulting in improved I/O performance
@@
     stacks are best suited to different application classes, as well as
-    understanding the shortcomings of current algorithms, designs and
-    technologies so as to propose new solutions. Unfortunately, application performance
-        studies via experiments on real-world compute platforms
+    understanding the limitations of current algorithms, designs and
+    technologies. Unfortunately, performance
+        studies relying on real-world experiments on compute platforms
         face several difficulties (high operational costs, labor-intensive experimental setups,
@@
     application performance on arbitrary platform configurations. It thus comes
-        as no surprise that, considering here only the field of parallel and distributed
-        computing, a large number of simulation frameworks have been developed and used
-        to support research and development endeavors~\cite{ optorsim, gridsim, groudsim, cloudsim, nunez2012simcan,nunez2012icancloud, mdcsim, dissect_cf, cloudnetsimplusplus, fognetsimplusplus, casanova2014simgrid, ROSS, casanova2020fgcs}.
+        as no surprise that a large number of simulation frameworks have been developed and used 
+        for research and development~\cite{ optorsim, gridsim, groudsim, cloudsim, nunez2012simcan,nunez2012icancloud, mdcsim, dissect_cf, cloudnetsimplusplus, fognetsimplusplus, casanova2014simgrid, ROSS, casanova2020fgcs}.
 
-        Page caching is an ubiquitous technique for mitigating the I/O bottleneck
-        faced by data-intensive applications. As such, it is necessary to simulate page caching
-        when performing performance studies in simulation.
-        Furthermore, simulating the page cache makes it possible to study relevant application-I/O
-        bottlenecks that can arise over the course of execution, such as page cache
-        starvation~\cite{zhuang2017}.
+        Page caching is an ubiquitous technique for mitigating the I/O bottleneck.
+        As such, it is necessary to model it
+        when simulating data-intensive applications.
         While existing simulation frameworks of parallel and distributed computing
         systems  capture many relevant features of hardware/software stacks, 
-        they lack the ability to simulate page cache adequately. Those frameworks
-        that implement some page cache simulation
-        fail to capture key features such
+        they lack the ability to simulate page cache with enough details to capture key features such
         as dirty data and cache eviction policies~\cite{nunez2012simcan,nunez2012icancloud}. 
-        Some simulators, such as that in~\cite{xu2018saving}, do capture such features
+        Some simulators, such as the one in~\cite{xu2018saving}, do capture such features,
         but are domain-specific.
@@
 
-    In this work, we present \wrench-cache, a simulation model of I/O
-    with page cache incorporated into the \wrench simulation framework~\cite{casanova2020fgcs},
-    which is  based on the popular \simgrid distributed simulation framework~\cite{casanova2014simgrid}.
-    The added value of \wrench-cache is to make it possible to simulate 
-        application execution accurately in the presence of page caching. Our specific 
-        contributions are as follows:
+    In this work, we present \wrench-cache, a page cache simulation model
+    implemented in \wrench~\cite{casanova2020fgcs}, a workflow simulation
+    framework based on the popular \simgrid distributed simulation
+    toolkit~\cite{casanova2014simgrid}. Our contributions are:
     \begin{compactitem}
@@
         \item An evaluation of the accuracy and scalability of our model, and of its implementation,
-              for a range of applications, platforms, and page cache configurations. 
+              for multiple applications, execution scenarios, and page cache configurations. 
     \end{compactitem}
@@
     instance, it is now common for HPC clusters to run applications in
-    Linux control groups (cgroups), where resource consumption is limited, including
-    memory and therefore page cache usage. Using our simulator,
+    Linux control groups (cgroups), where resource consumption is limited,
+    including memory and therefore page cache usage. Using our simulator,
     it would be possible to study the interaction between memory allocation
-    and I/O performance, for instance to improve scheduling algorithms.
-    Our simulator could also be utilized to evaluate solutions that
-    reduce the impact of network file transfers on distributed
-    applications, such as burst
+    and I/O performance, for instance to improve scheduling algorithms or
+    avoid page cache starvation~\cite{zhuang2017}. Our simulator could also
+    be utilized to evaluate solutions that reduce the impact of network
+    file transfers on distributed applications, such as burst
     buffers~\cite{ferreiradasilva-fgcs-bb-2019}, hierarchical file
-    systems~\cite{islam2015triple}, active storage~\cite{5496981}, or specific
-    hardware architectures~\cite{hayot2020performance}.
+    systems~\cite{islam2015triple}, active storage~\cite{5496981}, or
+    specific hardware architectures~\cite{hayot2020performance}.
 

### 2020-12-15 paper_io_sim.tex
Edited methods

@@
     two lists of data blocks, an active list and an inactive list, both ordered by
-    last access time (earliest first, see Figure~\ref{fig:lrulist}).
+    last access time (earliest first, Figure~\ref{fig:lrulist}).
     As in the kernel, our simulator limits the size of the active list to
@@
             model against a real sequential application running on a real system.
-            The source code this prototype is available at
-            \url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}. We used Python 3.7.
+            The Python 3.7 source code this prototype is available at
+            \url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}.
 
@@
         multi-threaded applications, accessing data on local and remote
-        file systems. We use two applications: a synthetic one created to
-        evaluate the simulation model, and a real one representative of
+        file systems. We use two applications: a synthetic one, designed to
+        evaluate the simulation model, and a real one, representative of
         neuroimaging data processing.
@@
         processing, and writes the resulting data to disk. Files are
-        numbered by ascending creation times: File 1 is file read by Task 1, and File \texttt{i+1} is the file
+        numbered by ascending creation times: File 1 is the file read by Task 1, and file \texttt{i+1} is the file
         written by task \texttt{i}. In real executions, we clear the page
@@
         The real application is a workflow of the Nighres
-        toolbox~\cite{huntenburg2018nighres} that implements skull
-        stripping in high-resolution brain images, an operation common to
-        many neuroimage analysis pipelines. The workflow is implemented as a Python
+        toolbox~\cite{huntenburg2018nighres} that implements
+        cortical reconstruction from brain images in four steps: skull stripping, tissue classification, 
+        region extraction, and cortical reconstruction. More information 
+        on this applicatoin is available in the Nighres documentation at \url{https://nighres.readthedocs.io/en/latest/auto_examples/example_02_cortical_depth_estimation.html}.
+        The workflow is implemented as a Python
         script that calls Java image-processing routines. We used Python
         3.6, Java 8, and Nighres 1.3.0. We patched the
-        skull stripping workflow to monitor task CPU times and inject them
-        in the simulation, to focus our analysis on I/O times, the patched code is 
+        application to remove lazy data loading, which made CPU time difficult to separate from I/O time, 
+        and to monitor task CPU times, to inject them in the simulators. The patched code is 
         available at \url{https://github.com/dohoangdzung/nighres}.
@@
         file system, 378~GiB of tmpfs, and 126~GiB of devtmpfs file system.
-        Nodes run CentOS~8.1 with NFS version 4. We use the \texttt{atop}
+        Nodes run CentOS~8.1 with NFS version 4. We used the \texttt{atop}
         and \texttt{collectl} tools to monitor and collect memory status
@@
 
-        To parametrize our simulators, we benchmarked the
+        To parametrize the simulators, we benchmarked the
         memory, local disk, remote disk (NFS), and network bandwidths
@@
             (Table~\ref{table:cputime}), and used these durations in our
-            simulations. For the Python simulator, we used CPU times
+            simulations. For the Python prototype, we injected CPU times
             directly in the simulation. For \wrench and \wrench-cache, we
@@
 
-            Our first experiment (\textit{Exp single-threaded}) runs a
+            Our first experiment runs a
             single instance of the synthetic application on a single
@@
 
-            Our second experiment (\textit{Exp concurrent}) runs concurrent instances of the
+            Our second experiment runs concurrent instances of the
             application on a single node, all application instances
@@
 
-            Our third experiment (\textit{Exp NFS}) occurs in the same
+            Our third experiment occurs in the same
             configuration as the previous one, albeit reading and writing
@@
 
-            Our fourth and final experiment (\textit{Exp Nighres})
-            simulates the Nighres CRUISE cortical reconstruction workflow, which consists 
-            of four steps. The first two steps are skull stripping and MGDM segmentation, 
-            which are in this Nighres \href{{https://nighres.readthedocs.io/en/latest/auto_examples/example_01_tissue_classification.html}}{tissue classification example}. 
-            The last two tasks are region extraction and CRUISE cortical reconstruction 
-            in the   
-            \href{https://nighres.readthedocs.io/en/latest/auto_examples/example_02_cortical_depth_estimation.html}{cortical depth estimation example}.
-            \tristan{describe data, and if the execution was single node or
-            distributed.}
+            Our fourth and final experiment runs the Nighres workflow on
+            data from participant 0027430 in the dataset of the Max Planck Institute for
+            Human Cognitive and Brain Sciences
+            available at
+            \url{http://dx.doi.org/10.15387/fcp_indi.corr.mpg1}. The data
+            amounts to 2.3~GB and was downloaded from
+            \url{http://datasets.datalad.org/?dir=/corr/RawDataBIDS/MPG_1/sub-0027430}.
+            The workflow was run on a single node of the cluster, using a
+            single local disk.
 
@@
     Evaluation results show that our model improves simulation accuracy
-    substantially, decreasing absolute relative simulation errors up to
+    substantially, reducing absolute relative simulation errors up to
     9$\times$ (see results of the single-threaded experiment). The
     availability of asymmetrical disk bandwidths in the forthcoming
-    \simgrid release will further improve the accuracy of our simulations.
+    \simgrid release will further improve these results.
     Our page cache model is publicly available in the \wrench GitHub
@@
 
-    Page cache simulation would be instrumental in a number of studies. For
+    Page cache simulation can be instrumental in a number of studies. For
     instance, it is now common for HPC clusters to run applications in
@@
     including random I/O, might also be worth including in application models,
-    as they directly impact page cache content.
+    as they directly affect page cache content.
     % readahead and persistent storage could also be adeed
@@
 The computing platform used in the experiments was obtained with funding
-from the Canada Foundation for Innovation.
+from the Canada Foundation for Innovation. This work was partially supported
+by NSF contracts \#1923539 and \#1923621.
 

### 2020-12-15 paper_io_sim.tex
Edits in results and conclusion

@@
 
-        \subsection{Single-threaded experiment}
-
-        The page cache simulation model drastically reduces I/O simulation
-        errors in each application task (Figure~\ref{fig:single_error}). The first read is not impacted
-        as it only involves uncached data. Errors are reduced from an average
-        of 345\% in the original \wrench to 46\% in the Python prototype and
-        39\% in \wrench-cache. Unsurprisingly, the original \wrench simulator
-        significantly overestimates read and write times, due to the lack
-        of page cache simulation. Results with files of 50~GB and 75~GB
-        show similar behaviors and are not reported for brevity.
-
-        \wrench simulation errors are substantially lower with 100~GB
-        files than with 20~GB files, due to the fact that part of the
-        100~GB file needs to be read and written to disk, the only storage
-        device in \wrench, as it does not fit in cache. Conversely,
-        simulation errors of the Python prototype and \wrench-cache are higher with
-        100~GB files than with 20~GB files, due to idiosyncrasies in the kernel
-        flushing and eviction strategies.
-
-        Simulated memory profiles are highly consistent with the real ones
-        (Figure~\ref{fig:single_memprof}). Used memory consists of both
-        cached and anonymous memory that cannot be evicted at the current
-        time. With 20~GB files, memory profiles almost exactly match the
-        real ones, although dirty data seems to be flushing faster in real
-        life than in simulation \tristan{any idea why?}, a behavior also
-        observed with 100~GB files. With 100~GB files, used memory reaches
-        total memory during the first write, triggering dirty data
-        flushing. Used memory drops to cached memory when application tasks
-        release anonymous memory. Simulated cached memory is highly
-        consistent with real values, except toward the end of Read 3 where
-        it slightly increases in simulation but not in reality. This
-        occurred due to the fact that after Write 2, File 3 was only partly
-        cached in simulation whereas it was entirely cached in the real
-        system. In all cases, dirty data remains under the dirty ratio as
-        expected. The Python prototype and \wrench-cache exhibit nearly
-        identical memory profiles, which reinforces the confidence in our
-        implementation.
-
-        The content of the simulated memory cache is also highly consistent
-        with reality (Figure~\ref{fig:single_cache}). With 20~GB files, the
-        simulated cache content exactly matches reality, since all files fit in page cache. With 100~GB files, a slight
-        discrepancy is observed after Write 2, which explains the
-        simulation error previously mentioned in Read 3. In the real execution indeed,
-        File 3 is entirely cached after Write 2, whereas in the simulated execution, 
-        only a part of it is cached. 
-        This is due to the fact that the in simulation, data is flushed and evicted 
-        from the inactive list first, then from the active list \tristan{whereas in the real execution ...?}. Thus, File 2, which is 
-        accessed again by Read 2 and is cached in the active list, is not flushed and evicted, 
-        while File 3 is flushed and evicted since it is accessed once only and cached 
-        in the inactive list.
-
-        % We conclude from this experiment that our model can accurately
-        % simulate cached I/Os in a sequential context. The next experiment
-        % evaluates our simulation with concurrent I/Os.
-
     \begin{figure*}
@@
 
-        \subsection{Concurrent application (local storage)}
+        \subsection{Single-threaded experiment}
+
+        The page cache simulation model drastically reduces I/O simulation
+        errors in each application task (Figure~\ref{fig:single_error}). The first read is not impacted
+        as it only involves uncached data. Errors are reduced from an average
+        of 345\% in the original \wrench to 46\% in the Python prototype and
+        39\% in \wrench-cache. Unsurprisingly, the original \wrench simulator
+        significantly overestimates read and write times, due to the lack
+        of page cache simulation. Results with files of 50~GB and 75~GB
+        show similar behaviors and are not reported for brevity.
+
+        \wrench simulation errors are substantially lower with 100~GB
+        files than with 20~GB files, due to the fact that part of the
+        100~GB file needs to be read and written to disk, the only storage
+        device in \wrench, as it does not fit in cache. Conversely,
+        simulation errors of the Python prototype and \wrench-cache are higher with
+        100~GB files than with 20~GB files, due to idiosyncrasies in the kernel
+        flushing and eviction strategies that could not be easily modeled.
+
+        Simulated memory profiles are highly consistent with the real ones
+        (Figure~\ref{fig:single_memprof}). Used memory consists of both
+        cached and anonymous memory that cannot be evicted at the current
+        time. With 20~GB files, memory profiles almost exactly match the
+        real ones, although dirty data seems to be flushing faster in real
+        life than in simulation \tristan{any idea why?}, a behavior also
+        observed with 100~GB files. With 100~GB files, used memory reaches
+        total memory during the first write, triggering dirty data
+        flushing, and drops back to cached memory when application tasks
+        release anonymous memory. Simulated cached memory is highly
+        consistent with real values, except toward the end of Read 3 where
+        it slightly increases in simulation but not in reality. This
+        occurred due to the fact that after Write 2, File 3 was only partially
+        cached in simulation whereas it was entirely cached in the real
+        system. In all cases, dirty data remains under the dirty ratio as
+        expected. The Python prototype and \wrench-cache exhibit nearly
+        identical memory profiles, which reinforces the confidence in our
+        implementations.
 
@@
         \end{figure*}
+
+        The content of the simulated memory cache is also highly consistent
+        with reality (Figure~\ref{fig:single_cache}). With 20~GB files, the
+        simulated cache content exactly matches reality, since all files fit in page cache. With 100~GB files, a slight
+        discrepancy is observed after Write 2, which explains the
+        simulation error previously mentioned in Read 3. In the real execution indeed,
+        File 3 is entirely cached after Write 2, whereas in the simulated execution, 
+        only a part of it is cached. 
+        This is due to the fact that the Linux kernel tend to not evict 
+        pages that belong to files being currently written (File 3 in this case), which we 
+        could not easily reproduce in our model.
+        \begin{figure*}
+            \begin{subfigure}{\linewidth}
+                \centering
+                \includegraphics[width=\linewidth]{result/multi/figures/multi_nfs.pdf}
+            \end{subfigure}
+            \caption{NFS results with 3~GB files (averages on 5 repetitions)}
+            \label{fig:multi_nfs}
+            \end{figure*}
+
+        \subsection{Concurrent applications}
 
@@
 
-            \begin{figure*}
-            \begin{subfigure}{\linewidth}
-                \centering
-                \includegraphics[width=\linewidth]{result/multi/figures/multi_nfs.pdf}
-            \end{subfigure}
-            \caption{NFS results with 3~GB files (averages on 5 repetitions)}
-            \label{fig:multi_nfs}
-            \end{figure*}
-
             Page cache simulation importantly reduces the simulation error
             on remote storage as well (Figure~\ref{fig:multi_nfs}). This
-            manifests only for reads, as only writethrough cache was
-            enabled in the NFS server. Both \wrench and \wrench-cache
+            manifests only for reads, as the NFS server used writethrough rather than writeback cache.
+            Both \wrench and \wrench-cache
             underestimate write times due to the discrepancy between
@@
     avoid page cache starvation~\cite{zhuang2017}. Our simulator could also
-    be utilized to evaluate solutions that reduce the impact of network
+    be leveraged to evaluate solutions that reduce the impact of network
     file transfers on distributed applications, such as burst
@@
 
-    However, I/Os remain much more complex than is currently available in
-    simulation models, which could drastically limit the accuracy of simulations.
+    I/Os remain much more complex than is currently available in
+    simulation models, which could substantially limit the accuracy of simulations.
     Relevant extensions to this work could include more
-    accurate descriptions of anonymous memory usage by applications. 
-    With the page cache, I/O times are indeed strongly
-    dependent on the amount of available memory, which is itself impacted
-    by the use of anonymous memory. File access patterns, 
+    accurate descriptions of anonymous memory usage in applications, 
+    which strongly affects I/O times through writeback cache. File access patterns, 
     including random I/O, might also be worth including in application models,

### 2020-12-15 gen_fig.py
Fix mem_prof xmax

@@ -11,8 +11,8 @@ import result.single.plot_error as plot_error
 import result.single.plot_cache as plot_cache
 
 # Generate memory profiling comparision between 20GB and 100GB inputs.
-# xmax is set to 200 and 1300
-plot_memprof.compare_plot([20, 100], [200, 1300])
+# xmax is set to 165 and 1300
+plot_memprof.compare_plot([20, 100], [165, 1300])
 # Generate simulation errors with 20GB and 100GB inputs
 plot_error.plot_error()
 # Generate amount of cached data with 20GB and 100GB inputs

### 2020-12-15 paper_io_sim.tex
tiny fix in related work
figure edits

@@
 
-        To improve performance of disk accesses, the page cache was introduced to the Linux kernel. %It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
+        %To improve performance of disk accesses, the page cache was introduced to the Linux kernel. %It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
         The goal of the page cache is to ensure that most I/O operations occur entirely
@@
         \cite{linuxdev3rd2010}.
+
         Cache eviction and flushing strategies are integral to proper page cache functioning.
@@
                \vspace*{-0.7cm}
-               \caption{Memory profiles \tristan{Remove trailing
-               whitespace for 20GB graphs (xmax should be about 160s,
-               not 200s)} \tristan{Find a way to add Read 1, Write 1,
-               Read 2, etc on the graphs, perhaps at the top of the
-               real execution graphs}}
+               \caption{Memory profiles}
                \vspace*{0.5cm}

### 2020-12-15 result/single/plot_memprof.py

@@ -39,7 +39,7 @@ def real_subplot(subplot_ax, real_time_file, real_mem_file, xmin, xmax, ymin, ym
                                label="Read" if i == 0 else "")
         else:
             subplot_ax.axvspan(xmin=timestamps[i - 1][2] - start, xmax=timestamps[i][1] - start, color="k", alpha=0.25,
-                               label="Computation" if i == 1 else "")
+                               label="Compute" if i == 1 else "")
             subplot_ax.axvspan(xmin=timestamps[i][1] - start, xmax=timestamps[i][2] - start, color="k", alpha=0.4,
                                label="write" if i == 1 else "")
 
@@ -99,7 +99,7 @@ def sim_subplot(subplot_ax, sim_time_file, sim_mem_file, title, bar_alpha=0.4, l
         if idx == 0:
             subplot_ax.axvspan(xmin=0, xmax=read_ends[idx] - start, color="k", alpha=0.1, label="Read")
             subplot_ax.axvspan(xmin=read_ends[idx] - start, xmax=write_starts[idx] - start, color="k",
-                               alpha=0.25, label="Computation")
+                               alpha=0.25, label="Compute")
             subplot_ax.axvspan(xmin=write_starts[idx] - start, xmax=write_ends[idx] - start, color="k", alpha=0.4,
                                label="Write")
         else:
@@ -154,8 +154,8 @@ def compare_plot(input_sizes=[20, 100], makespan=[200, 1300]):
 
     lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.05, 3.75), ncol=8)
 
-    fig.text(0.27, 0.06, 'time (s)', ha='center', fontsize=12)
-    fig.text(0.77, 0.06, 'time (s)', ha='center', fontsize=12)
+    fig.text(0.32, 0.06, 'time (s)', ha='center', fontsize=12)
+    fig.text(0.73, 0.06, 'time (s)', ha='center', fontsize=12)
     fig.text(0.08, 0.45, 'memory (GB)', va='center', rotation='vertical', fontsize=12)
     fig.text(0.91, 0.69, 'Real execution', fontsize=14, color='white', bbox={'pad': 5, 'color': '#90C987'}, rotation=-90)
     fig.text(0.91, 0.405, 'Python prototype', fontsize=14, color='white', bbox={'pad': 5, 'color': '#7BAFDE'}, rotation=-90)

### 2020-12-15 result/multi/local/gantt_plot.py
Delete redundant logs, fix plots

@@ -33,8 +33,8 @@ def parse_pipeline(filename):
     return [task_1, task_2, task_3]
 
 
-def gantt_plot(no_pipelines=20):
-    pipelines = [parse_pipeline("indv_logs/time_pipeline_%d_%d.csv" % (no_pipelines, i + 1)) for i in
+def gantt_plot(no_pipelines=32):
+    pipelines = [parse_pipeline("real/only_5/time_pipeline_%d_%d.csv" % (no_pipelines, i + 1)) for i in
                  range(no_pipelines)]
 
     fig, ax = plt.subplots()
@@ -58,7 +58,9 @@ def gantt_plot(no_pipelines=20):
     plt.xlabel("Time (s)")
     plt.ylabel("Tasks")
     plt.title("no. pipelines = %d" % no_pipelines)
+    plt.legend()
+    plt.tight_layout()
     plt.show()
 
 
-gantt_plot(32)
+# gantt_plot(5)

### 2020-12-15 result/nighres/parse_real_log.py

@@ -48,5 +48,5 @@ def parse_task(logfile, sum_file):
         json.dump(summary, outfile)
 
 
-parse_time("real/log/timelog_sub001_sess1_.json", "real/log/__summary_time_sub001_sess1.json")
-parse_task("real/log/timelog_sub001_sess1_.json", "real/log/__summary_file_sub001_sess1.json")
+parse_time("real/log/timelog_sub001_sess1.json", "real/log/summary_time_sub001_sess1.json")
+parse_task("real/log/timelog_sub001_sess1.json", "real/log/summary_file_sub001_sess1.json")

### 2020-12-15 paper_io_sim.tex
rewording related works

@@
 
-        %To improve performance of disk accesses, the page cache was introduced to the Linux kernel. %It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
-        The goal of the page cache is to ensure that most I/O operations occur entirely
-        within memory space rather than synchronously to and from storage devices.
-        All disk-backed data located in the page cache are partitioned into many small pages.
-        When a read operation is initiated, the kernel checks whether the required pages are in memory.
-        If it is the case, data is read entirely from memory (i.e. \textit{cache hit}). Otherwise,
-        the necessary pages must be loaded to memory from the disk (i.e. \textit{cache miss}).
-        Conversely, for writing, one of three strategies can be applied: direct~I/O (or, ``no-write''),
-        writethrough and writeback.
-        If direct~I/O is enabled for a given file system, all I/O operations bypass the
-        page cache for writes, writing data directly to disk. In contrast, writethrough enables the page cache
-        for file system writes, but the data is written synchronously from memory to disk. Writeback extends writethrough
-        by enabling asynchronous writes to disk, meaning that a write call will return once
-        the data has been written to memory.
+        %To improve performance of disk accesses, the page cache was introduced to the Linux kernel. %It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}
+        Page cache offsets the cost of disk I/O by enabling I/O to occur directly from memory.
+        When a file is first loaded into memory, the file is read from disk and loaded into the page cache as
+        a series of pages. Subsequent reads to any of the files pages located in memory will result in
+        a \textit{cache hit}, meaning the I/O can occur directly from without disk involvement.
+        Any accessed page not loaded in memory results in a \textit{cache miss}, resulting in
+        the page being read directly from disk.
+
+        Written pages can also contribute to future application cache hits. When use of page cache
+        is enabled for a given filesystem, through the enabling of writeback or writethrough cache,
+        all written pages are written first to page cache, prior to being written to disk.
+        Accessing of these written pages may result in cache hits, should the pages remain in memory.
+
+        The kernel may also provide improved write performance if writeback cache is enabled. Unlike writethrough,
+        where the data is synchronously written from memory to disk, writeback enables asynchronous writes.
+        In other words, with writeback, the application may proceed with execution once the data has been
+        written to memory, even if it has not yet been materialized to disk.  
         The writeback strategy is considered to outperform writethrough as well as
-        direct I/O as it delays disk writes to perform a bulk write at a later time
+        direct I/O (page cache bypassed for I/O) as it delays disk writes to perform a bulk write at a later time
         \cite{linuxdev3rd2010}.
@@
         \cite{owda2014comparison}.
+
         The Linux kernel uses a two-list strategy to flag pages for eviction.

### 2020-12-15 paper_io_sim.tex

--- 2020-12-15 paper_io_sim.tex
@@
 
-        To parametrize the simulators, we benchmarked the
+        To parameterize the simulators, we benchmarked the
         memory, local disk, remote disk (NFS), and network bandwidths
@@
     We designed a model of the Linux page cache and implemented it in the
-    \simgrid-based \wrench library to simulate distributed applications.
+    \simgrid-based \wrench simulation framework to simulate the execution
+    of distributed applications.
     Evaluation results show that our model improves simulation accuracy
-    substantially, reducing absolute relative simulation errors up to
+    substantially, reducing absolute relative simulation errors by up to
     9$\times$ (see results of the single-threaded experiment). The
@@
 
-    I/Os remain much more complex than is currently available in
-    simulation models, which could substantially limit the accuracy of simulations.
-    Relevant extensions to this work could include more
+    Not all I/O behaviors are captured by currently available simulation models,
+    including the one developed in this work, 
+    which could substantially limit the accuracy of simulations.
+    Relevant extensions to this work include more
     accurate descriptions of anonymous memory usage in applications, 
     which strongly affects I/O times through writeback cache. File access patterns, 
-    including random I/O, might also be worth including in application models,
+    including random I/O, might also be worth including in the simulation models,
     as they directly affect page cache content.

### 2020-12-15 gen_fig.py
Add nighres results

@@ -27,3 +27,8 @@ multi_result.result_local(rep_no=5)
 multi_result.result_nfs(rep_no=5)
 # runtime plots
 multi_result.run_time()
+
+# ======== Nighres experiment results ==========
+os.chdir("../nighres")
+import result.nighres.process_result as nighres_result
+nighres_result.plot_error()

### 2020-12-15 result/nighres/parse_real_log.py

@@ -1,52 +1,0 @@
-import json
-
-
-def parse_time(logfile, sum_file):
-    with open(logfile, "r") as logfile:
-        log = json.load(logfile)
-
-        summary = {}
-        for task in log:
-            makespan = log[task]["makespan"][0]["duration"]
-
-            read_total = 0
-            write_total = 0
-            for fileread in log[task]["read"]:
-                read_total += fileread["duration"]
-            for filewrite in log[task]["write"]:
-                write_total += filewrite["duration"]
-
-            summary[task] = {
-                "cputime": makespan - read_total - write_total,
-                "read": read_total,
-                "write": write_total
-            }
-
-    with open(sum_file, "w") as outfile:
-        json.dump(summary, outfile)
-
-
-def parse_task(logfile, sum_file):
-    with open(logfile, "r") as logfile:
-        log = json.load(logfile)
-
-        summary = {}
-        for task in log:
-            _input = {}
-            _output = {}
-            for fileread in log[task]["read"]:
-                _input[fileread["filename"]] = fileread["filesize"]
-            for filewrite in log[task]["write"]:
-                _output[filewrite["filename"]] = filewrite["filesize"]
-
-            summary[task] = {
-                "input": _input,
-                "output": _output
-            }
-
-    with open(sum_file, "w") as outfile:
-        json.dump(summary, outfile)
-
-
-parse_time("real/log/timelog_sub001_sess1.json", "real/log/summary_time_sub001_sess1.json")
-parse_task("real/log/timelog_sub001_sess1.json", "real/log/summary_file_sub001_sess1.json")

### 2020-12-15 result/nighres/process_result.py

@@ -1,0 +1,174 @@
+import json
+import matplotlib.pyplot as plt
+import numpy as np
+
+def parse_time_real(logfile, sum_file):
+    with open(logfile, "r") as logfile:
+        log = json.load(logfile)
+
+        summary = {}
+        for task in log:
+            makespan = log[task]["makespan"][0]["duration"]
+
+            read_total = 0
+            write_total = 0
+            for fileread in log[task]["read"]:
+                read_total += fileread["duration"]
+            for filewrite in log[task]["write"]:
+                write_total += filewrite["duration"]
+
+            summary[task] = {
+                "cputime": makespan - read_total - write_total,
+                "read": read_total,
+                "write": write_total
+            }
+
+    with open(sum_file, "w") as outfile:
+        json.dump(summary, outfile)
+
+
+def parse_time_wrench(dumpfile, sum_file):
+    with open(dumpfile, "r") as logfile:
+        log = json.load(logfile)
+
+        summary = {}
+        for task in log["workflow_execution"]["tasks"]:
+            makespan = task["whole_task"]["end"] - task["whole_task"]["start"]
+
+            read_start = min(read["start"] for read in task["read"])
+            read_end = max(read["end"] for read in task["read"])
+            read_duration = read_end - read_start
+
+            write_start = min(write["start"] for write in task["write"])
+            write_end = max(write["end"] for write in task["write"])
+            write_duration = write_end - write_start
+
+            summary[task["task_id"]] = {
+                "cputime": makespan - write_duration - read_duration,
+                "read": read_duration,
+                "write": write_duration
+            }
+
+    with open(sum_file, "w") as outfile:
+        json.dump(summary, outfile)
+
+
+def parse_task_real(logfile, sum_file):
+    with open(logfile, "r") as logfile:
+        log = json.load(logfile)
+
+        summary = {}
+        for task in log:
+            _input = {}
+            _output = {}
+            for fileread in log[task]["read"]:
+                _input[fileread["filename"]] = fileread["filesize"]
+            for filewrite in log[task]["write"]:
[…]

### 2020-12-15 result/nighres/process_result.py
Fix nighres error plot function

@@ -101,12 +101,11 @@ def plot_error():
     wrench_cache_log_file = "wrench/pagecache/dump_nighres_pagecache_sim_time.json"
     wrench_cache_time_file = "wrench/pagecache/summary_wrench_nighres_pagecache.json"
 
+    parse_time_real(real_log_file, real_time_file)
+    parse_task_real(real_log_file, "real/summary_file_sub001_sess1.json")
 
-    # parse_time_real(real_log_file, real_time_file)
-    # parse_task_real(real_log_file, "real/summary_file_sub001_sess1.json")
-    #
-    # parse_time_wrench(wrench_log_file, wrench_time_file)
-    # parse_time_wrench(wrench_cache_log_file, wrench_cache_time_file)
+    parse_time_wrench(wrench_log_file, wrench_time_file)
+    parse_time_wrench(wrench_cache_log_file, wrench_cache_time_file)
 
     tasks = ["mp2rage_skullstripping", "mgdm_segmentation", "extract_brain_region", "cruise_cortex_extraction"]
 

### 2020-12-16 paper_io_sim.tex
integrated real application results in text

@@
 
-        Our experiments compare real executions with our Python prototype,
-        with the original \wrench, and with our \wrench-cache extension. We
-        evaluate our page cache simulation in single-threaded and
-        multi-threaded applications, accessing data on local and remote
-        file systems. We use two applications: a synthetic one, designed to
-        evaluate the simulation model, and a real one, representative of
-        neuroimaging data processing.
+        Our experiments compared real executions with our Python prototype,
+        with the original \wrench simulator, and with our \wrench-cache
+        extension. Executions included single-threaded and multi-threaded
+        applications, accessing data on local and remote file systems. We
+        used two applications: a synthetic one, created to evaluate the
+        simulation model, and a real one, representative of neuroimaging
+        data processing.
 
-        The synthetic application, implemented in C, consists of three single-core,
-        sequential tasks where each task reads the file produced by the
-        previous task, increments every byte of this file to emulate real
-        processing, and writes the resulting data to disk. Files are
-        numbered by ascending creation times: File 1 is the file read by Task 1, and file \texttt{i+1} is the file
-        written by task \texttt{i}. In real executions, we clear the page
-        cache before starting the application, to ensure comparable
-        conditions among the runs. The anonymous memory used by application
-        is released after each task, which we also simulate in the Python
-        prototype and in WRENCH-cache.
-
-        The real application is a workflow of the Nighres
-        toolbox~\cite{huntenburg2018nighres} that implements
-        cortical reconstruction from brain images in four steps: skull stripping, tissue classification, 
-        region extraction, and cortical reconstruction. More information 
-        on this applicatoin is available in the Nighres documentation at \url{https://nighres.readthedocs.io/en/latest/auto_examples/example_02_cortical_depth_estimation.html}.
-        The workflow is implemented as a Python
-        script that calls Java image-processing routines. We used Python
-        3.6, Java 8, and Nighres 1.3.0. We patched the
-        application to remove lazy data loading, which made CPU time difficult to separate from I/O time, 
-        and to monitor task CPU times, to inject them in the simulators. The patched code is 
-        available at \url{https://github.com/dohoangdzung/nighres}.
-
-        The applications were executed on a dedicated cluster at
+        Experiments were conducted on a dedicated cluster at
         Concordia University, with one login node, 9 compute nodes, and 4
         storage nodes connected with a network bandwidth of 25 Gbps. Each
-        compute node has 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
+        compute node had 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
         @ 2.10GHz, 250~GiB of RAM, 6 $\times$ SSDs of 450~GiB each with the XFS
         file system, 378~GiB of tmpfs, and 126~GiB of devtmpfs file system.
-        Nodes run CentOS~8.1 with NFS version 4. We used the \texttt{atop}
+        Nodes ran CentOS~8.1 with NFS version 4. We used the \texttt{atop}
         and \texttt{collectl} tools to monitor and collect memory status
-        and disk throughput.
+        and disk throughput. We cleared the page
+        cache before starting the application, to ensure comparable
+        conditions among runs.
+
+        \begin{table}[b]
+            \centering
+            \begin{tabularx}{0.8\columnwidth}{c>{\centering\arraybackslash}X}
+            \toprule
+                Input size (GB)  & CPU time (s)\\
+            \midrule
+                3      & 4.4 \\
+                20  & 28 \\
+                50  & 75 \\
+                75  & 110 \\
+                100  & 155 \\
+            \bottomrule
+            \end{tabularx}
+            \caption{Synthetic application parameters}
+            \label{table:cputime}
+            \end{table}
+        The synthetic application, implemented in C, consisted of three single-core,
+        sequential tasks where each task read the file produced by the
+        previous task, incremented every byte of this file to emulate real
+        processing, and wrote the resulting data to disk. Files were
+        numbered by ascending creation times: File 1 was the file read by Task 1, and file \texttt{i+1} was the file
+        written by task \texttt{i}. The anonymous memory used by the application
+        was released after each task, which we also simulated in the Python
+        prototype and in WRENCH-cache. As our focus is on I/O rather than compute, we measured
+        application task CPU times on a cluster node
+        (Table~\ref{table:cputime}), and used these durations in our
+        simulations. For the Python prototype, we injected CPU times
+        directly in the simulation. For \wrench and \wrench-cache, we
+        determined the corresponding number of flops on a 1~Gflops CPU
+        and used these values in the simulation. The final simulated
+        platform and application are available at
+        commit \href{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}{ec6b43561b}.
+
+        We used the synthetic application in three experiments. In the
+        first one (\textit{Exp 1}), we executed \emph{a single instance} of
+        the application on a single cluster node, with different input file
+        sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed to
+        the same local disk.
+        In the second experiment (\textit{Exp 2}), we executed
+        \emph{concurrent} instances of the application on a single node,
+        all application instances operating on different files of size 3~GB
+        stored in the same local disk. We varied the number of concurrent
+        application instances from 1 to 32 since cluster nodes had 32 CPU
+        cores.
+        In the third experiment (\textit{Exp 3}), we used the same
+        configuration as the previous one, albeit reading and writing
+        on a 50-GiB NFS-mounted partition of a 450-GiB remote disk of
+        another compute node. As is commonly configured in HPC
+        environments to avoid data loss, there was no client write cache
+        and the server cache was configured as writethrough instead of
+        writeback. NFS client and server read caches were enabled. 
+        Therefore, all the writes happened at disk bandwidth, but
+        reads can benefit from cache hits.
+
+        The real application was a workflow of the Nighres
+        toolbox~\cite{huntenburg2018nighres}, implementing cortical
+        reconstruction from brain images in four steps: skull stripping,
+        tissue classification, region extraction, and cortical
+        reconstruction. Each step read
+        files produced by the previous step, and wrote files that were or were not read by the subsequent step. 
+        More information on this
+        application is available in the Nighres documentation at
+        \url{https://nighres.readthedocs.io/en/latest/auto_examples/example_02_cortical_depth_estimation.html}.
+        The workflow is implemented as a Python script that calls Java
+        image-processing routines. We used Python 3.6, Java 8, and Nighres
+        1.3.0. We patched the application to remove lazy data loading,
+        which made CPU time difficult to separate from I/O time, and to
+        monitor task CPU times, to inject them in the simulators. The
+        patched code is available at
+        \url{https://github.com/dohoangdzung/nighres}.
+        \begin{table}
+            \centering
+            \begin{tabular}{lccc}
+            \toprule
+                \multicolumn{1}{c}{Workflow step}& Input size       & Output size      & CPU time\\
+                                       & (MB)             & (MB)             & (s)\\
+            \midrule
+               Skull stripping         &  111             & 30               & 137 \\
+               Tissue classification   &  22              & 175              & 624 \\
+               Region extraction       &  175             & 121              & 81 \\
+               Cortical reconstruction &  47              & 250              & 270\\
+            \bottomrule
+            \end{tabular} 
+            \caption{Nighres application parameters}
+            \label{table:nighres_stats}
+            \end{table}
+
+        We used the real application in the fourth and final experiment
+        (\textit{Exp 4}). The workflow was run on a single node of the
+        cluster, using a single local disk. We processed data from
+        participant 0027430 in the dataset of the Max Planck Institute for
+        Human Cognitive and Brain Sciences available at
+        \url{http://dx.doi.org/10.15387/fcp_indi.corr.mpg1}, leading to the
+        workflow parameters in Table~\ref{table:nighres_stats}.
 
@@
         bandwidth values in our experiments.
-            \begin{table}
+            \begin{table}[b]
             \centering
@@
 
-            As our focus is on I/O rather than compute, we measured
-            application task CPU times on a cluster node
-            (Table~\ref{table:cputime}), and used these durations in our
-            simulations. For the Python prototype, we injected CPU times
-            directly in the simulation. For \wrench and \wrench-cache, we
-            determined the corresponding number of flops on a 1~Gflops CPU
-            and used these values in the simulation. The final simulated
-            platform and application used in our experiments are available at
-            commit \href{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}{ec6b43561b}.
-
-            \begin{table}[b]
-            \centering
-            \begin{tabularx}{0.8\columnwidth}{c>{\centering\arraybackslash}X}
-            \toprule
-                Input size (GB)  & CPU time (second)\\
-            \midrule
-                3      & 4.4 \\
-                20  & 28 \\
-                50  & 75 \\
-                75  & 110 \\
-                100  & 155 \\
-            \bottomrule
-            \end{tabularx}
-            \caption{Task CPU times used in the simulation.
-%            \tristan{are the input files in GB or in GiB? If GiB, then change the abbreviation everywhere.}
-%  Dzung: it's in GB
-            \tristan{add nighres CPU times when available}}
-            \label{table:cputime}
-            \end{table}
-
             % Finally, because \wrench simulates applications base on network-communication,
             % we use an infinite bandwidth to eliminate network latency for local I/Os.
-
-            Our first experiment runs a
-            single instance of the synthetic application on a single
-            cluster node, with different input file sizes (20~GB, 50~GB,
-            75~GB, 100~GB), and with all I/Os directed to the same local
-            disk. The goal is to validate the simulation model in absence
-            of any concurrent disk access.
-
-            Our second experiment runs concurrent instances of the
-            application on a single node, all application instances
-            operating on different files of size 3~GB stored in the same
-            local disk. We vary the number of
-            concurrent application instances from 1 to 32 since cluster
-            nodes have 32 CPU cores.
-
-            Our third experiment occurs in the same
-            configuration as the previous one, albeit reading and writing
-            on a 50-GiB NFS-mounted partition of a 450-GiB remote disk of
-            another compute node. As is commonly configured in HPC
-            environments to avoid data loss, there is no client write cache
-            and the server cache is configured as writethrough instead of
-            writeback. NFS client and server read caches are enabled. 
-            Therefore, all the writes happen at disk bandwidth, but
-            reads can benefit from cache hits.
-
-            Our fourth and final experiment runs the Nighres workflow on
-            data from participant 0027430 in the dataset of the Max Planck Institute for
-            Human Cognitive and Brain Sciences
-            available at
-            \url{http://dx.doi.org/10.15387/fcp_indi.corr.mpg1}. The data
-            amounts to 2.3~GB and was downloaded from
-            \url{http://datasets.datalad.org/?dir=/corr/RawDataBIDS/MPG_1/sub-0027430}.
-            The workflow was run on a single node of the cluster, using a
-            single local disk.
 
@@
         \end{subfigure}
-        \caption{Single-threaded results}
+        \caption{Single-threaded results (\textit{Exp 1})}
         \end{figure*}
 
-        \subsection{Single-threaded experiment}
+        \subsection{Single-threaded execution (Exp 1)}
 
@@
         show similar behaviors and are not reported for brevity.
+
+        \begin{figure*}
+            \begin{subfigure}{\linewidth}
+                \centering
+                \includegraphics[width=\linewidth]{result/multi/figures/multi_local.pdf}
+            \end{subfigure}
+            \caption{Concurrent results with 3~GB files (\textit{Exp 2}). Averages on 5 repetitions.}
+            \label{fig:multi_local}
+        \end{figure*}
 
@@
 
-        \begin{figure*}
-            \begin{subfigure}{\linewidth}
-                \centering
-                \includegraphics[width=\linewidth]{result/multi/figures/multi_local.pdf}
-            \end{subfigure}
-            \caption{Concurrent results with 3~GB files (averages on 5 repetitions)}
-            \label{fig:multi_local}
-        \end{figure*}
-
         The content of the simulated memory cache is also highly consistent
@@
             \end{subfigure}
-            \caption{NFS results with 3~GB files (averages on 5 repetitions)}
+            \caption{NFS results with 3~GB files (\textit{Exp 3}). Averages on 5 repetitions.}
             \label{fig:multi_nfs}
@@
 
-        \subsection{Concurrent applications}
+        \subsection{Concurrent applications (Exp 2)}
 
@@
 
-        \subsection{Remote storage (NFS)}
+        \subsection{Remote storage (Exp 3)}
 
@@
 
-        \subsection{Real application}
+        \subsection{Real application (Exp 4)}
+        As for the synthetic application, simulation errors are again
+        substantially reduced by the WRENCH-cache simulator compared to
+        WRENCH (Figure~\ref{fig:nighres}). On average, errors are reduced
+        from 212~\% in WRENCH to 68~\% in WRENCH-cache. The error
+        reduction is mostly due to more accurate read times, since write
+        times remain generally identical between both implementations, presumably due to 
+        the dominance of file compression time in write time.
+        Similar to
+        the synthetic application, the first read happens entirely from disk and is therefore 
+        very accurately simulated by both WRENCH and WRENCH-cache.
+        \begin{figure}
+            \begin{subfigure}{\columnwidth}
+                \centering
+                \includegraphics[width=\linewidth]{result/nighres/figures/nighres_errors.pdf}
+            \end{subfigure}
+            \caption{Real application results (\textit{Exp 4})}
+            \label{fig:nighres}
+            \end{figure}
 
-        \subsection{Overhead comparison}
+        \subsection{Simulation time}
         As is the case for \wrench, simulation time with \wrench-cache scales

### 2020-12-16 result/multi/process_result.py

@@ -120,7 +120,7 @@ def result_local(rep_no=1):
     ax2.set_ylim(bottom=0, top=1500)
 
     plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.18), ncol=4)
-    plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.3)
+    plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.4)
     plt.savefig("figures/multi_local.pdf", format="pdf")
     plt.savefig("figures/multi_local.svg", format="svg")
 
@@ -143,7 +143,7 @@ def result_nfs(rep_no=1):
     ax2.set_ylim(bottom=0, top=1500)
 
     plt.legend(loc='upper center', bbox_to_anchor=(-0.1, 1.2), ncol=4)
-    plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.1)
+    plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.4)
     plt.savefig("figures/multi_nfs.pdf", format="pdf")
     plt.savefig("figures/multi_nfs.svg", format="svg")
 

### 2020-12-16 paper_io_sim.tex
Cleaned up version

@@
 
-\newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:}#1\color{black}}
+\newcommand{\tristan}[1]{}%\color{orange}\textbf{From Tristan:}#1\color{black}}
 \newcommand{\henri}[1]{\color{green}\textbf{From Henri:}#1\color{black}}
@@
     page caching fully, or even at all.  As a result, simulation-based
-    performance studies of data-intensive applications leads to inaccurate
+    performance studies of data-intensive applications lead to inaccurate
     results.
@@
         where the data is synchronously written from memory to disk, writeback enables asynchronous writes.
-        In other words, with writeback, the application may proceed with execution once the data has been
+        With writeback, the application may proceed with execution once the data has been
         written to memory, even if it has not yet been materialized to disk.  
@@
         page cache longer than an expiration time configured in the kernel.
-        Different cache eviction algorithms have also been developed and proposed
+        Different cache eviction algorithms have also been proposed
         \cite{owda2014comparison}.
@@
 
-        The two main concerns for simulation are \emph{accuracy}
-        (the ability to faithfully reproduce real-world executions) and
-        \emph{scalability}  (the ability to simulate large/long real-world
-        executions quickly and with low RAM footprint). The above
+        The two main concerns for simulation are accuracy,
+        the ability to faithfully reproduce real-world executions, and
+        scalability, the ability to simulate large/long real-world
+        executions quickly and with low RAM footprint. The above
         frameworks achieve different compromises between the two.  At
@@
         However, the scalability of the iCanCloud simulator is limited
-        due to it using  microscopic simulation models.  Besides, none
+        as it uses microscopic models.  Besides, none
         of these simulators provide any writeback cache simulator nor
@@
         replacement policies are applied in~\cite{xu2018saving} to
-        simulate in-memory caching, their simulator is specific to
+        simulate in-memory caching, this simulator is specific to
         energy consumption of multi-tier heterogeneous networks.
 
-        In this study, we develop a page cache simulation model which we implement
-        as part of the \wrench framework. We targeted \wrench because it is a
-        recent, actively developed framework that provides convenient simulation
-        abstractions, because it is easy to incorporate new simulation models
-        into it, and because it builds on \simgrid, whose scalable and accurate
-        simulation models we can thus reuse for our purpose.
+        In this study, we implement a page cache simulation model in the
+        \wrench framework. We targeted \wrench because it is a recent,
+        actively developed framework that provides convenient simulation
+        abstractions, because it is extensible, and because it reuses
+        \simgrid's scalable and accurate models.
 
@@
     the second one, which operates in the background, periodically searches for
-    expired dirty data in page cache LRU lists and flushes this data to disk. We
+    expired dirty data in LRU lists and flushes this data to disk. We
     use existing storage simulation models~\cite{lebre2015} to simulate disk and
@@
     bandwidths, and latency. These models account for
-    bandwidth sharing between concurrent memory accesses and concurrent disk accesses.
+    bandwidth sharing between concurrent memory or disk accesses.
 
@@
     page cache that were accessed in the same I/O operation.
-    A data block has information about the file name, block size, last access
+    A data block stores the file name, block size, last access
     time, a dirty flag that represents whether the data is clean (0)
@@
     added to the inactive list. On subsequent accesses, blocks of the
-    inactive list are moved to the top of the active list. Cached blocks
-    written to cache are marked as dirty until they are flushed to disk.
+    inactive list are moved to the top of the active list. Blocks
+    written to cache are marked dirty until flushed.
 
@@
     function takes the amount of data to flush as parameter. While
-    the amount of data to flush is not reached and there are dirty
-    blocks remaining in cache, this function traverses the sorted
+    this amount is not reached and dirty
+    blocks remain in cache, this function traverses the sorted
     inactive list, then the sorted active list, and writes the
@@
 
-    \begin{algorithm}\caption{Periodical flushing simulation in I/O Controller}\label{alg:pdflush}
+    \begin{algorithm}[b]\caption{Periodical flush simulation in Memory Manager}\label{alg:pdflush}
         \small
@@
 
-    \begin{algorithm}\caption{File chunk read simulation in the I/O Controller}
+    \begin{algorithm}\caption{File chunk read simulation in I/O Controller}
     \label{alg:read}
@@
     a cache read  and update the corresponding data blocks accordingly (line 17).
-    Finally, the memory manager is called to reduce the amount of free memory
-    corresponding to the amount of anonymous memory used by application (line 19).
+    Finally, the memory manager is called to deallocate the amount of anonymous memory used by the application (line 19).
 
-    \begin{algorithm}\caption{File chunk write simulation in the I/O Controller}
+    \begin{algorithm}\caption{File chunk write simulation in I/O Controller}
     \label{alg:write}
@@
     page cache is calculated (line 8), and a cache write is simulated (line 9).
-    If the dirty threshold is reached and there is remaining data to be written,
+    If the dirty threshold is reached and there is still data to write,
     the remaining data is written to cache in a loop
-    where we repeatedly flush and evict to free space in the cache (line 12-18).
+    where we repeatedly flush and evict from the cache (line 12-18).
 
-    The above models page cache in writeback
+    The above model describes page cache in writeback
     mode. Our model also includes a write function in writethrough mode,
@@
             model against a real sequential application running on a real system.
-            The Python 3.7 source code this prototype is available at
+            The Python 3.7 source code of this prototype is available at
             \url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}.
@@
             concurrent accesses to page cache LRU lists by the two Memory
-            Manager threads. For the experiments in this paper we use
+            Manager threads. For the experiments, we used
             \wrench 1.6 at commit
@@
             is now part of \wrench's master branch and will be available to
-            users with the upcoming \wrench 1.8 release. \wrench provides a full \simgrid-based simulation 
+            users with the upcoming 1.8 release. \wrench provides a full \simgrid-based simulation 
             environment that supports, among other features, concurrent accesses to storage devices, 
@@
         extension. Executions included single-threaded and multi-threaded
-        applications, accessing data on local and remote file systems. We
+        applications, accessing data on local and network file systems. We
         used two applications: a synthetic one, created to evaluate the
@@
 
-        Experiments were conducted on a dedicated cluster at
+        Experiments were run on a dedicated cluster at
         Concordia University, with one login node, 9 compute nodes, and 4
-        storage nodes connected with a network bandwidth of 25 Gbps. Each
+        storage nodes connected with a 25 Gbps network. Each
         compute node had 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
         @ 2.10GHz, 250~GiB of RAM, 6 $\times$ SSDs of 450~GiB each with the XFS
-        file system, 378~GiB of tmpfs, and 126~GiB of devtmpfs file system.
-        Nodes ran CentOS~8.1 with NFS version 4. We used the \texttt{atop}
+        file system, 378~GiB of tmpfs, 126~GiB of devtmpfs file system,
+        CentOS~8.1 and NFS version 4. We used the \texttt{atop}
         and \texttt{collectl} tools to monitor and collect memory status
         and disk throughput. We cleared the page
-        cache before starting the application, to ensure comparable
-        conditions among runs.
+        cache before each application run to ensure comparable
+        conditions.
 
@@
         processing, and wrote the resulting data to disk. Files were
-        numbered by ascending creation times: File 1 was the file read by Task 1, and file \texttt{i+1} was the file
-        written by task \texttt{i}. The anonymous memory used by the application
+        numbered by ascending access times (File 1 was the file read by Task 1, etc).
+         The anonymous memory used by the application
         was released after each task, which we also simulated in the Python
-        prototype and in WRENCH-cache. As our focus is on I/O rather than compute, we measured
+        prototype and in WRENCH-cache. As our focus was on I/O rather than compute, we measured
         application task CPU times on a cluster node
@@
         determined the corresponding number of flops on a 1~Gflops CPU
-        and used these values in the simulation. The final simulated
+        and used these values in the simulation. The simulated
         platform and application are available at
@@
         We used the synthetic application in three experiments. In the
-        first one (\textit{Exp 1}), we executed \emph{a single instance} of
+        first one (\textit{Exp~1}), we ran \emph{a single instance} of
         the application on a single cluster node, with different input file
@@
         the same local disk.
-        In the second experiment (\textit{Exp 2}), we executed
+        In the second experiment (\textit{Exp~2}), we ran
         \emph{concurrent} instances of the application on a single node,
@@
         cores.
-        In the third experiment (\textit{Exp 3}), we used the same
+        In the third experiment (\textit{Exp~3}), we used the same
         configuration as the previous one, albeit reading and writing
@@
         \url{https://nighres.readthedocs.io/en/latest/auto_examples/example_02_cortical_depth_estimation.html}.
-        The workflow is implemented as a Python script that calls Java
+        The application is implemented as a Python script that calls Java
         image-processing routines. We used Python 3.6, Java 8, and Nighres
@@
         \url{https://github.com/dohoangdzung/nighres}.
-        \begin{table}
+        \begin{table}[t]
             \centering
@@
             \end{table}
-
-        We used the real application in the fourth and final experiment
-        (\textit{Exp 4}). The workflow was run on a single node of the
-        cluster, using a single local disk. We processed data from
-        participant 0027430 in the dataset of the Max Planck Institute for
-        Human Cognitive and Brain Sciences available at
-        \url{http://dx.doi.org/10.15387/fcp_indi.corr.mpg1}, leading to the
-        workflow parameters in Table~\ref{table:nighres_stats}.
-
-        To parameterize the simulators, we benchmarked the
-        memory, local disk, remote disk (NFS), and network bandwidths
-        (Table~\ref{table:benchmark}). Since \simgrid, and thus \wrench, currently only supports
-        symmetrical bandwidths, we use the mean of the read and write
-        bandwidth values in our experiments.
             \begin{table}[b]
@@
                 \caption{Bandwidth benchmarks (MBps) and simulator configurations.
-            The bandwidths used in the simulations are the average of the read and write bandwidths
-            measured in the cluster.
-            Remote disk and network accesses are not simulated in the Python simulator.}
+                The bandwidths used in the simulations were the average of the measured read and write bandwidths.
+                Network accesses were not simulated in the Python prototype.}
                 \label{table:benchmark}
                 \end{table}
+
+        We used the real application in the fourth experiment
+        (\textit{Exp~4}), run on a single cluster node 
+        using a single local disk. We processed data from
+        participant 0027430 in the dataset of the Max Planck Institute for
+        Human Cognitive and Brain Sciences available at
+        \url{http://dx.doi.org/10.15387/fcp_indi.corr.mpg1}, leading to the
+        parameters in Table~\ref{table:nighres_stats}.
+
+        To parameterize the simulators, we benchmarked the
+        memory, local disk, remote disk (NFS), and network bandwidths
+        (Table~\ref{table:benchmark}). Since \simgrid, and thus \wrench, currently only supports
+        symmetrical bandwidths, we use the mean of the read and write
+        bandwidth values in our experiments.
 
@@
 
-        The page cache simulation model drastically reduces I/O simulation
-        errors in each application task (Figure~\ref{fig:single_error}). The first read is not impacted
-        as it only involves uncached data. Errors are reduced from an average
+        The page cache simulation model drastically reduced I/O simulation
+        errors in each application task (Figure~\ref{fig:single_error}). The first read was not impacted
+        as it only involved uncached data. Errors were reduced from an average
         of 345\% in the original \wrench to 46\% in the Python prototype and
         39\% in \wrench-cache. Unsurprisingly, the original \wrench simulator
-        significantly overestimates read and write times, due to the lack
+        significantly overestimated read and write times, due to the lack
         of page cache simulation. Results with files of 50~GB and 75~GB
-        show similar behaviors and are not reported for brevity.
+        showed similar behaviors and were not reported for brevity.
+
+        \wrench simulation errors were substantially lower with 100~GB
+        files than with 20~GB files, due to the fact that part of the
+        100~GB file needed to be read and written to disk, the only storage
+        device in \wrench, as it did not fit in cache. Conversely,
+        simulation errors of the Python prototype and \wrench-cache were higher with
+        100~GB files than with 20~GB files, due to idiosyncrasies in the kernel
+        flushing and eviction strategies that could not be easily modeled.
+
+        Simulated memory profiles were highly consistent with the real ones
+        (Figure~\ref{fig:single_memprof}). With 20~GB files, memory profiles almost exactly matched the
+        real ones, although dirty data seemed to be flushing faster in real
+        life than in simulation \tristan{any idea why?}, a behavior also
+        observed with 100~GB files. With 100~GB files, used memory reached
+        total memory during the first write, triggering dirty data
+        flushing, and droped back to cached memory when application tasks
+        released anonymous memory. Simulated cached memory was highly
+        consistent with real values, except toward the end of Read 3 where
+        it slightly increased in simulation but not in reality. This
+        occurred due to the fact that after Write 2, File 3 was only partially
+        cached in simulation whereas it was entirely cached in the real
+        system. In all cases, dirty data remained under the dirty ratio as
+        expected. The Python prototype and \wrench-cache exhibited nearly
+        identical memory profiles, which reinforces the confidence in our
+        implementations.
+
+        The content of the simulated memory cache was also highly
+        consistent with reality (Figure~\ref{fig:single_cache}). With 20~GB
+        files, the simulated cache content exactly matched reality, since
+        all files fit in page cache. With 100~GB files, a slight
+        discrepancy was observed after Write 2, which explains the
+        simulation error previously mentioned in Read 3. In the real
+        execution indeed, File 3 was entirely cached after Write 2, whereas
+        in the simulated execution, only a part of it was cached. This was
+        due to the fact that the Linux kernel tends to not evict pages that
+        belong to files being currently written (File 3 in this case),
+        which we could not easily reproduce in our model.
 
@@
             \end{figure*}
+        \subsection{Concurrent applications (Exp 2)}
 
-        \wrench simulation errors are substantially lower with 100~GB
-        files than with 20~GB files, due to the fact that part of the
-        100~GB file needs to be read and written to disk, the only storage
-        device in \wrench, as it does not fit in cache. Conversely,
-        simulation errors of the Python prototype and \wrench-cache are higher with
-        100~GB files than with 20~GB files, due to idiosyncrasies in the kernel
-        flushing and eviction strategies that could not be easily modeled.
+            The page cache model notably reduced \wrench's simulation error
+            for concurrent applications executed with local I/Os
+            (Figure~\ref{fig:multi_local}). For reads, \wrench-cache still
+            slightly overestimated reality, due to the discrepancy between
+            simulated and real read bandwidths mentioned before. For writes,
+            \wrench-cache slightly underestimated reality for the same reason. The plateau
+            ending around 7 concurrent applications in the \wrench-cache writes
+            marks the limit beyond which the page cache was saturated with dirty data
+            and needed flushing. \tristan{why is there no plateau in real life?}
 
-        Simulated memory profiles are highly consistent with the real ones
-        (Figure~\ref{fig:single_memprof}). Used memory consists of both
-        cached and anonymous memory that cannot be evicted at the current
-        time. With 20~GB files, memory profiles almost exactly match the
-        real ones, although dirty data seems to be flushing faster in real
-        life than in simulation \tristan{any idea why?}, a behavior also
-        observed with 100~GB files. With 100~GB files, used memory reaches
-        total memory during the first write, triggering dirty data
-        flushing, and drops back to cached memory when application tasks
-        release anonymous memory. Simulated cached memory is highly
-        consistent with real values, except toward the end of Read 3 where
-        it slightly increases in simulation but not in reality. This
-        occurred due to the fact that after Write 2, File 3 was only partially
-        cached in simulation whereas it was entirely cached in the real
-        system. In all cases, dirty data remains under the dirty ratio as
-        expected. The Python prototype and \wrench-cache exhibit nearly
-        identical memory profiles, which reinforces the confidence in our
-        implementations.
+            \begin{figure}[b]
+                \begin{subfigure}{0.95\linewidth}
+                    \centering
+                    \includegraphics[width=\linewidth]{result/nighres/figures/nighres_errors.pdf}
+                \end{subfigure}
+                \caption{Real application results (\textit{Exp 4}) \tristan{increase font size}}
+                \label{fig:nighres}
+                \end{figure}
+        \subsection{Remote storage (Exp 3)}
 
-        The content of the simulated memory cache is also highly consistent
-        with reality (Figure~\ref{fig:single_cache}). With 20~GB files, the
-        simulated cache content exactly matches reality, since all files fit in page cache. With 100~GB files, a slight
-        discrepancy is observed after Write 2, which explains the
-        simulation error previously mentioned in Read 3. In the real execution indeed,
-        File 3 is entirely cached after Write 2, whereas in the simulated execution, 
-        only a part of it is cached. 
-        This is due to the fact that the Linux kernel tend to not evict 
-        pages that belong to files being currently written (File 3 in this case), which we 
-        could not easily reproduce in our model.
+            Page cache simulation importantly reduced simulation error
+            on NFS storage as well (Figure~\ref{fig:multi_nfs}). This
+            manifested only for reads, as the NFS server used writethrough rather than writeback cache.
+            Both \wrench and \wrench-cache
+            underestimated write times due to the discrepancy between
+            simulated and real bandwidths mentioned previously. For reads,
+            this discrepancy only impacted the results beyond 22
+            applications since before this threshold, most reads resulted in cache
+            hits.
+
+        \subsection{Real application (Exp 4)}
+        As for the synthetic application, simulation errors were
+        substantially reduced by the WRENCH-cache simulator compared to
+        WRENCH (Figure~\ref{fig:nighres}). On average, errors were reduced
+        from 212~\% in WRENCH to 68~\% in WRENCH-cache. Error
+        reduction was mostly due to read times, since write
+        times remained generally identical between both implementations, presumably due to 
+        the dominance of file compression time in write time.
+        Similar to
+        the synthetic application, the first read happened entirely from disk and was therefore 
+        very accurately simulated by both WRENCH and WRENCH-cache.
+
             \begin{figure*}
@@
 
-        \subsection{Concurrent applications (Exp 2)}
-
-            The page cache model notably reduces \wrench's simulation error
-            for concurrent applications executed with local I/Os
-            (Figure~\ref{fig:multi_local}). For reads, \wrench-cache still
-            slightly overestimates reality, due to the discrepancy between
-            simulated and real read bandwidths mentioned before. For writes,
-            \wrench-cache slightly underestimates reality for the same reason. The plateau
-            ending around 7 concurrent applications in the \wrench-cache writes
-            marks the limit beyond which the page cache is saturated with dirty data
-            and needs flushing \tristan{why is there no plateau in real life?}.
-
-        \subsection{Remote storage (Exp 3)}
-
-            Page cache simulation importantly reduces the simulation error
-            on remote storage as well (Figure~\ref{fig:multi_nfs}). This
-            manifests only for reads, as the NFS server used writethrough rather than writeback cache.
-            Both \wrench and \wrench-cache
-            underestimate write times due to the discrepancy between
-            simulated and real bandwidths mentioned previously. For reads,
-            this discrepancy only impacts the results beyond 22
-            applications since before this threshold, most reads result in cache
-            hits.
-            % \tristan{Dzung, pls check that.}.
-            % Yes, disk reads appear from 23 apps indeed.
-
-        \subsection{Real application (Exp 4)}
-        As for the synthetic application, simulation errors are again
-        substantially reduced by the WRENCH-cache simulator compared to
-        WRENCH (Figure~\ref{fig:nighres}). On average, errors are reduced
-        from 212~\% in WRENCH to 68~\% in WRENCH-cache. The error
-        reduction is mostly due to more accurate read times, since write
-        times remain generally identical between both implementations, presumably due to 
-        the dominance of file compression time in write time.
-        Similar to
-        the synthetic application, the first read happens entirely from disk and is therefore 
-        very accurately simulated by both WRENCH and WRENCH-cache.
             \begin{figure}
@@
                     \centering
-                \includegraphics[width=\linewidth]{result/nighres/figures/nighres_errors.pdf}
+                    \includegraphics[width=\linewidth]{result/multi/figures/simulation_time.pdf}
                 \end{subfigure}
-            \caption{Real application results (\textit{Exp 4})}
-            \label{fig:nighres}
+                \caption{Simulation time comparison. \wrench-cache scales
+                linearly with the number of concurrent applications, albeit
+                with a higher overhead than \wrench.}
+                \label{fig:multi_time}
                 \end{figure}
@@
         cache in NFS, which bypasses flushing operations.
-        \begin{figure}
-            \begin{subfigure}{\columnwidth}
-                \centering
-                \includegraphics[width=\linewidth]{result/multi/figures/simulation_time.pdf}
-            \end{subfigure}
-            \caption{Simulation time comparison. \wrench-cache scales
-            linearly with the number of concurrent applications, albeit
-            with a higher overhead than \wrench.}
-            \label{fig:multi_time}
-            \end{figure}
 
@@
     accurate descriptions of anonymous memory usage in applications, 
-    which strongly affects I/O times through writeback cache. File access patterns, 
-    including random I/O, might also be worth including in the simulation models,
+    which strongly affects I/O times through writeback cache. File access patterns
+     might also be worth including in the simulation models,
     as they directly affect page cache content.

### 2020-12-17 paper_io_sim.tex
Update nighres results

@@
             \midrule
-               Skull stripping         &  111             & 30               & 137 \\
-               Tissue classification   &  22              & 175              & 624 \\
-               Region extraction       &  175             & 121              & 81 \\
-               Cortical reconstruction &  47              & 250              & 270\\
+               Skull stripping         &  295             & 393               & 137 \\
+               Tissue classification   &  197              & 1376              & 614 \\
+               Region extraction       &  1376             & 885              & 76 \\
+               Cortical reconstruction &  393              & 786              & 272\\
             \bottomrule
@@
         WRENCH (Figure~\ref{fig:nighres}). On average, errors were reduced
-        from 212~\% in WRENCH to 68~\% in WRENCH-cache. Error
-        reduction was mostly due to read times, since write
-        times remained generally identical between both implementations, presumably due to 
-        the dominance of file compression time in write time.
+        from 337~\% in WRENCH to 47~\% in WRENCH-cache. 
+%        Error reduction was mostly due to read times, since write
+%        times remained generally identical between both implementations, presumably due to 
+%        the dominance of file compression time in write time.
         Similar to

### 2020-12-17 paper_io_sim.tex
minor fixes

@@
         \item A page cache simulation model that supports 
-    both single-threaded and multithreaded applications, and of both
+    both single-threaded and multithreaded applications, and both
     writeback and writethrough caches for local or network-based
@@
         When a file is first loaded into memory, the file is read from disk and loaded into the page cache as
-        a series of pages. Subsequent reads to any of the files pages located in memory will result in
+        a series of pages. Subsequent reads to any of the file pages located in memory will result in
         a \textit{cache hit}, meaning the I/O can occur directly from without disk involvement.
@@
         and not evictable, while the inactive list includes pages accessed once only,
-        or pages that have be accessed more than once but moved from the active list.
-        Both lists operate using LRU eviction policies meaning the data that has
+        or pages that have been accessed more than once but moved from the active list.
+        Both lists operate using LRU eviction policies, meaning that data that has
         not be accessed recently will be moved first.
@@
 
-        The two main concerns for simulation are accuracy,
+        The two main concerns of simulation are accuracy,
         the ability to faithfully reproduce real-world executions, and
@@
         tissue classification, region extraction, and cortical
-        reconstruction. Each step read
-        files produced by the previous step, and wrote files that were or were not read by the subsequent step. 
-        More information on this
-        application is available in the Nighres documentation at
+        reconstruction. Each step read files produced by the previous step,
+        and wrote files that were or were not read by the subsequent step.
+        More information on this application is available in the Nighres
+        documentation at
         \url{https://nighres.readthedocs.io/en/latest/auto_examples/example_02_cortical_depth_estimation.html}.
@@
         image-processing routines. We used Python 3.6, Java 8, and Nighres
-        1.3.0. We patched the application to remove lazy data loading,
-        which made CPU time difficult to separate from I/O time, and to
-        monitor task CPU times, to inject them in the simulators. The
-        patched code is available at
+        1.3.0. We patched the application to remove lazy data loading and
+        data compression, which made CPU time difficult to separate from
+        I/O time, and to capture task CPU times to inject them in the
+        simulation. The patched code is available at
         \url{https://github.com/dohoangdzung/nighres}.
@@
         \subsection{Real application (Exp 4)}
-        As for the synthetic application, simulation errors were
+        Similar to the synthetic application, simulation errors were
         substantially reduced by the WRENCH-cache simulator compared to
@@
         from 337~\% in WRENCH to 47~\% in WRENCH-cache. 
-%        Error reduction was mostly due to read times, since write
-%        times remained generally identical between both implementations, presumably due to 
-%        the dominance of file compression time in write time.
-        Similar to
-        the synthetic application, the first read happened entirely from disk and was therefore 
+        The first read happened entirely from disk and was therefore 
         very accurately simulated by both WRENCH and WRENCH-cache.

### 2020-12-18 paper_io_sim.tex
minor edits

@@
 
-        The two main concerns of simulation are accuracy,
+        The two main concerns for simulation are accuracy,
         the ability to faithfully reproduce real-world executions, and
@@
             model against a real sequential application running on a real system.
-            The Python 3.7 source code of this prototype is available at
+            The Python 3.7 source code is available at
             \url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}.
@@
         Therefore, all the writes happened at disk bandwidth, but
-        reads can benefit from cache hits.
+        reads could benefit from cache hits.
 
@@
         of page cache simulation. Results with files of 50~GB and 75~GB
-        showed similar behaviors and were not reported for brevity.
+        showed similar behaviors and are not reported for brevity.
 
@@
         real ones, although dirty data seemed to be flushing faster in real
-        life than in simulation \tristan{any idea why?}, a behavior also
+        life than in simulation, \tristan{any idea why?} a behavior also
         observed with 100~GB files. With 100~GB files, used memory reached
@@
         files, the simulated cache content exactly matched reality, since
-        all files fit in page cache. With 100~GB files, a slight
+        all files fitted in page cache. With 100~GB files, a slight
         discrepancy was observed after Write 2, which explains the
@@
                 \end{subfigure}
-                \caption{Concurrent results with 3~GB files (\textit{Exp 2}). Averages on 5 repetitions.}
+                \caption{Concurrent results with 3~GB files (\textit{Exp 2})}
                 \label{fig:multi_local}
@@
                 \end{subfigure}
-                \caption{NFS results with 3~GB files (\textit{Exp 3}). Averages on 5 repetitions.}
+                \caption{NFS results with 3~GB files (\textit{Exp 3})}
                 \label{fig:multi_nfs}

### 2020-12-18 result/multi/process_result.py

@@ -94,7 +94,7 @@ def suplot_prop(ax, real_dir, wrench_dir, propname, title, ylabel, rep_no=1):
     # ax.plot(no_pipeline_df, max_df / no_pipeline_df, color="green", linewidth=1, alpha=0.5)
     # ax.plot(no_pipeline_df, min_df / no_pipeline_df, color="green", linewidth=1, alpha=0.5)
     ax.fill_between(no_pipeline_df, min_df / no_pipeline_df, max_df / no_pipeline_df, facecolor=real_color, alpha=0.5,
-                    label="Real execution min-max interval")
+                    label="Real execution min-max interval (5 repetitions)")
 
     ax.plot(no_pipeline_df, simg_org_df[propname] / no_pipeline_df, linewidth=2, color=wrench_color, label="WRENCH")
     ax.plot(no_pipeline_df, simg_ext_df[propname] / no_pipeline_df, linewidth=2, color=wrench_cache_color, label="WRENCH-cache")

### 2020-12-18 result/single/plot_real.py
Update plot style

@@ -31,32 +31,41 @@ def timestamp_readonly_plot(fig, time_stamps):
             fig.axvspan(xmin=read_end[idx] - start, xmax=read_start[idx + 1] - start, color="k", alpha=0.2)
 
 
-def mem_plot(fig, atoplog, time_stamp, input_size, readonly=False):
-    dirty_data = np.array(atoplog["total"])
+def mem_plot(fig, atop_log, time_stamp, input_size, readonly=False):
+    dirty_data = np.array(atop_log["total"])
     intervals = len(dirty_data)
     time = np.arange(0, intervals)
 
     fig.minorticks_on()
-    fig.set_title("memory profiling (input size = %s GB)" % input_size)
+    # fig.set_title("memory profiling (input size = %s GB)" % input_size)
+    start = time_stamp[0][1]
 
-    if time_stamp is not None:
-        if readonly:
-            timestamp_readonly_plot(fig, time_stamp)
+    for i in range(len(time_stamp)):
+        if time_stamp[i][0] == "read":
+            fig.axvspan(xmin=time_stamp[i][1] - start, xmax=time_stamp[i][2] - start, color="k", alpha=0.1,
+                        label="Read" if i == 0 else "")
         else:
-            timestamp_plot(fig, time_stamp)
+            fig.axvspan(xmin=time_stamp[i - 1][2] - start, xmax=time_stamp[i][1] - start, color="k", alpha=0.25,
+                        label="Compute" if i == 1 else "")
+            fig.axvspan(xmin=time_stamp[i][1] - start, xmax=time_stamp[i][2] - start, color="k", alpha=0.4,
+                        label="write" if i == 1 else "")
+
+    linewidth = 1.5
+    linestyle = "-"
+    line_alpha = 0.9
 
     # app_cache = list(np.array(app_mem) + np.array(cache_used))
 
-    fig.plot(time, atoplog["total"], color='k', linewidth=1, linestyle=":", label="total mem")
-    # ax1.plot(time, free_mem, color='g', linewidth=1.5, linestyle="-.", label="free memory")
-    fig.plot(time, atoplog["used_mem"], color='g', linewidth=1, label="used mem")
-    # ax1.plot(time, app_mem, color='c', linewidth=1.5, label="app memory")
-    fig.plot(time, atoplog["cache"], color='m', linewidth=1, label="cache used")
[…]

### 2020-12-18 paper_io_sim.tex
Edits in concurrent experiment

@@
 
-\newcommand{\tristan}[1]{}%\color{orange}\textbf{From Tristan:}#1\color{black}}
+\newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:}#1\color{black}}
 \newcommand{\henri}[1]{\color{green}\textbf{From Henri:}#1\color{black}}
@@
             slightly overestimated reality, due to the discrepancy between
-            simulated and real read bandwidths mentioned before. For writes,
-            \wrench-cache slightly underestimated reality for the same reason. The plateau
-            ending around 7 concurrent applications in the \wrench-cache writes
-            marks the limit beyond which the page cache was saturated with dirty data
-            and needed flushing. \tristan{why is there no plateau in real life?}
+            simulated and real read bandwidths mentioned before. For
+            writes, \wrench-cache correctly retrieves the plateau ending at
+            \tristan{x} concurrent applications in real life, corresponding
+            to the limit beyond which the page cache was saturated with
+            dirty data and needed flushing.
 

### 2020-12-18 paper_io_sim.tex
removed trailing comments

@@
     The overhead of the cache eviction algorithm is not part of the simulated time
-    since cache eviction time is negligible in real systems \tristan{a reference would be welcome}.
+    since cache eviction time is negligible in real systems. % \tristan{a reference would be welcome}.
 
@@
         real ones, although dirty data seemed to be flushing faster in real
-        life than in simulation, \tristan{any idea why?} a behavior also
+        life than in simulation, a behavior also
         observed with 100~GB files. With 100~GB files, used memory reached

### 2020-12-18 gen_fig.py
increase fontsize

@@ -12,11 +12,11 @@ import result.single.plot_cache as plot_cache
 
 # Generate memory profiling comparision between 20GB and 100GB inputs.
 # xmax is set to 165 and 1300
-# plot_memprof.compare_plot([20, 100], [165, 1300])
-# # Generate simulation errors with 20GB and 100GB inputs
-# plot_error.plot_error()
-# # Generate amount of cached data with 20GB and 100GB inputs
-# plot_cache.plot_cache_v2()
+plot_memprof.compare_plot([20, 100], [165, 1300])
+# Generate simulation errors with 20GB and 100GB inputs
+plot_error.plot_error()
+# Generate amount of cached data with 20GB and 100GB inputs
+plot_cache.plot_cache_v2()
 
 # ======== Multi-threaded experiment results ==========
 os.chdir("../multi")
@@ -24,11 +24,11 @@ import result.multi.process_result as multi_result
 # Generate figures of experiment with local file system
 multi_result.result_local(rep_no=2, step=5)
 # Generate figures of experiment with NFS
-# multi_result.result_nfs(rep_no=5)
-# # runtime plots
-# multi_result.run_time()
+multi_result.result_nfs(rep_no=5)
+# runtime plots
+multi_result.run_time()
 
 # ======== Nighres experiment results ==========
 os.chdir("../nighres")
-# import result.nighres.process_result as nighres_result
-# nighres_result.plot_error()
+import result.nighres.process_result as nighres_result
+nighres_result.plot_error()

### 2020-12-18 result/multi/process_result.py

@@ -129,10 +129,10 @@ def result_local(rep_no=1, step=1):
 
     plt.legend(loc='upper center', bbox_to_anchor=(-0.2, 1.18), ncol=4)
     plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.4)
-    # plt.savefig("figures/multi_local.pdf", format="pdf")
-    # plt.savefig("figures/multi_local.svg", format="svg")
+    plt.savefig("figures/multi_local.pdf", format="pdf")
+    plt.savefig("figures/multi_local.svg", format="svg")
 
-    plt.show()
+    # plt.show()
 
 
 def result_nfs(rep_no=1):

### 2020-12-18 result/nighres/process_result.py

@@ -155,8 +155,8 @@ def plot_error():
     # print("WRENCH-cache Write errors:")
     # print(abs(np.array(writes_wrench_cache) - np.array(writes_real)) / np.array(writes_real)% 100)
 
-    fig, ax = plt.subplots(figsize=(11, 5))
-    plt.rcParams.update({'font.size': 9})
+    plt.rcParams.update({'font.size': 12})
+    fig, ax = plt.subplots(figsize=(9, 4))
     wrench_error = abs(np.array(wrench_arr) - np.array(real_arr)) / np.array(real_arr) * 100
     wrench_cache_error = abs(np.array(wrench_cache_arr) - np.array(real_arr)) / np.array(real_arr) * 100
 
@@ -165,9 +165,10 @@ def plot_error():
     # labels = ["Write 1", "Write 2", "Write 3",  "Write 4"]
     grouped_bar_chart(ax, labels, "", "error (%)",  ("WRENCH", wrench_error, '#994F88'),
                       ("WRENCH-cache", wrench_cache_error, '#1965B0'))
-    lgd = plt.legend(loc='upper center', bbox_to_anchor=(0.5, 1), ncol=3)
+    lgd = plt.legend(loc='upper center', bbox_to_anchor=(0.55, 1), ncol=2)
+
+    plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.85, wspace=0.3)
 
     plt.savefig("figures/nighres_errors.svg", format="svg", bbox_extra_artists=(lgd,), bbox_inches='tight')
     plt.savefig("figures/nighres_errors.pdf", format="pdf", bbox_extra_artists=(lgd,), bbox_inches='tight')
-    plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.85, wspace=0.3)
     # plt.show()

### 2020-12-19 gen_fig.py
Update exp2 results

@@ -1,14 +1,16 @@
 import os
 import sys
-sys.path.append(os.path.abspath("result/single/"))
-sys.path.append(os.path.abspath("result/multi/"))
-
-
-# ======== Single-threaded experiment results ==========
-os.chdir("result/single")
 import result.single.plot_memprof as plot_memprof
 import result.single.plot_error as plot_error
 import result.single.plot_cache as plot_cache
+import result.multi.process_result as multi_result
+import result.nighres.process_result as nighres_result
+
+sys.path.append(os.path.abspath("result/single/"))
+sys.path.append(os.path.abspath("result/multi/"))
+
+# ======== Single-threaded experiment results ==========
+os.chdir("result/single")
 
 # Generate memory profiling comparision between 20GB and 100GB inputs.
 # xmax is set to 165 and 1300
@@ -20,9 +22,8 @@ plot_cache.plot_cache_v2()
 
 # ======== Multi-threaded experiment results ==========
 os.chdir("../multi")
-import result.multi.process_result as multi_result
 # Generate figures of experiment with local file system
-multi_result.result_local(rep_no=2, step=5)
+multi_result.result_local(rep_no=5, step=5)
 # Generate figures of experiment with NFS
 multi_result.result_nfs(rep_no=5)
 # runtime plots
@@ -30,5 +31,4 @@ multi_result.run_time()
 
 # ======== Nighres experiment results ==========
 os.chdir("../nighres")
-import result.nighres.process_result as nighres_result
 nighres_result.plot_error()

### 2020-12-19 result/multi/process_result.py

@@ -81,12 +81,9 @@ def parse_simgrid_result(filename, no_pipeline):
     return no_pipeline, makespan, read, write
 
 
-def suplot_prop(ax, real_dir, wrench_dir, propname, title, ylabel, rep_no=1, step=1):
-    simg_org_df = pd.read_csv("%s/original/aggregated.csv" % wrench_dir)
-    simg_ext_df = pd.read_csv("%s/pagecache/aggregated.csv" % wrench_dir)
-    no_pipeline_df = simg_org_df["no_pipeline"]
-    no_pipeline_real = list(range(1, len(no_pipeline_df) + 1, step))
+def suplot_prop(ax, real_dir, wrench_dir, propname, title, ylabel, rep_no=1, step=1, xmax=32):
 
+    # Real execution
     real_df = pd.DataFrame()
     for i in range(rep_no):
         df = pd.read_csv("%s/%d/aggregated.csv" % (real_dir, i + 1))
@@ -95,18 +92,27 @@ def suplot_prop(ax, real_dir, wrench_dir, propname, title, ylabel, rep_no=1, ste
     mean_df = real_df.mean(axis=1)
     max_df = real_df.max(axis=1)
     min_df = real_df.min(axis=1)
+    no_pipeline_real = list(range(1, xmax + 1, step))
 
-    ax.set_title(title)
+    # Simulation
+    simg_org_df = pd.read_csv("%s/original/aggregated.csv" % wrench_dir)
+    simg_ext_df = pd.read_csv("%s/pagecache/aggregated.csv" % wrench_dir)
+    simg_org_show = simg_org_df[simg_org_df["no_pipeline"] <= xmax]
+    simg_ext_show = simg_ext_df[simg_ext_df["no_pipeline"] <= xmax]
+    no_pipeline_sim = simg_org_df["no_pipeline"][simg_org_df["no_pipeline"] <= xmax]
 
+    # Plot real
     ax.plot(no_pipeline_real, mean_df / no_pipeline_real, color=real_color, linewidth=2, label="Real execution mean")
     # ax.plot(no_pipeline_df, max_df / no_pipeline_df, color="green", linewidth=1, alpha=0.5)
     # ax.plot(no_pipeline_df, min_df / no_pipeline_df, color="green", linewidth=1, alpha=0.5)
     ax.fill_between(no_pipeline_real, min_df / no_pipeline_real, max_df / no_pipeline_real, facecolor=real_color, alpha=0.5,
[…]

### 2020-12-19 gen_fig.py
Fix generating figures script

@@ -1,16 +1,14 @@
 import os
 import sys
-import result.single.plot_memprof as plot_memprof
-import result.single.plot_error as plot_error
-import result.single.plot_cache as plot_cache
-import result.multi.process_result as multi_result
-import result.nighres.process_result as nighres_result
-
 sys.path.append(os.path.abspath("result/single/"))
 sys.path.append(os.path.abspath("result/multi/"))
 
+
 # ======== Single-threaded experiment results ==========
 os.chdir("result/single")
+import result.single.plot_memprof as plot_memprof
+import result.single.plot_error as plot_error
+import result.single.plot_cache as plot_cache
 
 # Generate memory profiling comparision between 20GB and 100GB inputs.
 # xmax is set to 165 and 1300
@@ -22,6 +20,7 @@ plot_cache.plot_cache_v2()
 
 # ======== Multi-threaded experiment results ==========
 os.chdir("../multi")
+import result.multi.process_result as multi_result
 # Generate figures of experiment with local file system
 multi_result.result_local(rep_no=5, step=5)
 # Generate figures of experiment with NFS
@@ -31,4 +30,5 @@ multi_result.run_time()
 
 # ======== Nighres experiment results ==========
 os.chdir("../nighres")
+import result.nighres.process_result as nighres_result
 nighres_result.plot_error()

### 2020-12-19 paper_io_sim.tex
Fixed concurrent experiment text

@@
             for concurrent applications executed with local I/Os
-            (Figure~\ref{fig:multi_local}). For reads, \wrench-cache still
-            slightly overestimated reality, due to the discrepancy between
-            simulated and real read bandwidths mentioned before. For
-            writes, \wrench-cache correctly retrieves the plateau ending at
-            \tristan{x} concurrent applications in real life, corresponding
-            to the limit beyond which the page cache was saturated with
-            dirty data and needed flushing.
+            (Figure~\ref{fig:multi_local}). For reads, \wrench-cache
+            slightly overestimated runtime, due to the discrepancy between
+            simulated and real read bandwidths mentioned before. 
+            For writes, \wrench-cache
+            retrieved a plateau similar to the one observed in the real
+            execution, marking the limit beyond which the page cache was
+            saturated with dirty data and needed flushing.
 
@@
                 \end{subfigure}
-                \caption{Real application results (\textit{Exp 4}) \tristan{increase font size}}
+                \caption{Real application results (\textit{Exp 4})}
                 \label{fig:nighres}

### 2020-12-19 gen_fig.py
Fix plotting of exp2 results

@@ -22,7 +22,7 @@ plot_cache.plot_cache_v2()
 os.chdir("../multi")
 import result.multi.process_result as multi_result
 # Generate figures of experiment with local file system
-multi_result.result_local(rep_no=5, step=5)
+multi_result.result_local(rep_no=5)
 # Generate figures of experiment with NFS
 multi_result.result_nfs(rep_no=5)
 # runtime plots

### 2020-12-19 result/multi/process_result.py

@@ -5,10 +5,10 @@ import csv
 import json
 import seaborn as sns
 
-
 real_color = "#90C987"
 wrench_color = "#994F88"
 wrench_cache_color = "#1965B0"
+
 
 def parse_single_pipeline(filename):
     """
@@ -43,7 +43,7 @@ def aggregate_result(folder, no_pipeline):
     return no_pipeline, makespan, readtime, writetime
 
 
-def export_real_results(folder, filename, step=1):
+def export_real_results(folder, filename):
     """
     Export aggregated results for each repetition of read execution
     :param folder:
@@ -55,8 +55,9 @@ def export_real_results(folder, filename, step=1):
     with open(file, 'w', newline='') as csvfile:
         writer = csv.writer(csvfile)
         writer.writerow(["no_pipeline", "makespan", "readtime", "writetime"])
-        for i in range(0, 32, step):
-            writer.writerow(list(aggregate_result(folder, i + 1)))
+        no_pipelines = [1] + list(range(6, 17)) + [21, 26, 31]
+        for i in no_pipelines:
+            writer.writerow(list(aggregate_result(folder, i)))
 
 
 def export_simgrid_result(folder, filename):
@@ -81,18 +82,19 @@ def parse_simgrid_result(filename, no_pipeline):
     return no_pipeline, makespan, read, write
 
 
-def suplot_prop(ax, real_dir, wrench_dir, propname, title, ylabel, rep_no=1, step=1, xmax=32):
-
+def suplot_prop(ax, real_dir, wrench_dir, propname, title, ylabel, rep_no=1, xmax=32):
     # Real execution
     real_df = pd.DataFrame()
+    no_pipeline_real = pd.Series()
     for i in range(rep_no):
         df = pd.read_csv("%s/%d/aggregated.csv" % (real_dir, i + 1))
+        if no_pipeline_real.empty:
+            no_pipeline_real = df["no_pipeline"]
         real_df[i + 1] = df[propname]
 
     mean_df = real_df.mean(axis=1)
     max_df = real_df.max(axis=1)
     min_df = real_df.min(axis=1)
-    no_pipeline_real = list(range(1, xmax + 1, step))
 
     # Simulation
     simg_org_df = pd.read_csv("%s/original/aggregated.csv" % wrench_dir)
[…]

### 2021-05-11 paper_io_sim.tex

@@
            \State required\_mem = cs + disk\_read
-               \State mm.flush(required\_mem - mm.free\_mem - mm.evictable, fn)
-               \State mm.evict(required\_mem - mm.free\_mem, fn)
+           \State mm.flush(required\_mem - mm.free\_mem - mm.evictable)
+           \State mm.evict(required\_mem - mm.free\_mem)
            \If {disk\_read $>$ 0}  \Comment{Read uncached data}
@@
             \State mem\_amt = min(cs, mm.free\_mem)
-                \State mm.write\_to\_cache(fn, mem\_amt)
+            \State mm.write\_to\_cache( mem\_amt, fn)
         \EndIf
@@
             \State to\_cache = min(remaining, mm.free\_mem)
-                \State mm.write\_to\_cache(fn, to\_cache)
+            \State mm.write\_to\_cache(to\_cache, fn)
             \State remaining = remaining - to\_cache

### 2021-05-19 paper_io_sim.tex
Pass on the whole paper with mostly minor stuff Swapped the order of two paragraphs in Section II Added \todo for comments/questions/thoughts

@@
 \usepackage[dvipsnames]{xcolor}
+\usepackage{todonotes}
 \usepackage{amsmath}
@@
     page caching fully, or even at all.  As a result, simulation-based
-    performance studies of data-intensive applications lead to inaccurate
+    performance studies of data-intensive applications can lead to misleading
     results.
 
-    In this paper, we propose an I/O simulation model that includes
+    In this paper, we propose an I/O simulation model that captures
     the key features of the Linux page cache. We have implemented this model
@@
         written data. All these factors are important when determining the impact of I/O on
-        application performance, particularly in data-intensive applications.
+        application performance, particularly for data-intensive applications.
 
@@
     technologies. Unfortunately, performance
-        studies relying on real-world experiments on compute platforms
+        studies that rely on real-world experiments on compute platforms
         face several difficulties (high operational costs, labor-intensive experimental setups,
-        shared platforms with dynamic loads that hinder reproducibility of results) and shortcomings
+        shared platforms with dynamic loads that hinder reproducibility of results) and have several shortcomings
         (experiments are limited to the available platform/software configurations, which precludes
         the exploration of hypothetical scenarios). 
-    Simulations address these concerns by providing models
+    Simulation addresses these concerns by providing models
     and abstractions for the performance of computer hardware, such as
     CPU, network and storage. As a result, simulations provide a
-    cost-effective, fast, easy and reproducible way to evaluate
+    cost-effective, fast, convenient and reproducible way to evaluate
     application performance on arbitrary platform configurations. It thus comes
         as no surprise that a large number of simulation frameworks have been developed and used 
-        for research and development~\cite{ optorsim, gridsim, groudsim, cloudsim, nunez2012simcan,nunez2012icancloud, mdcsim, dissect_cf, cloudnetsimplusplus, fognetsimplusplus, casanova2014simgrid, ROSS, casanova2020fgcs}.
+        for research and development in the field of parallel and distributed computing~\cite{ optorsim, gridsim, groudsim, cloudsim, nunez2012simcan,nunez2012icancloud, mdcsim, dissect_cf, cloudnetsimplusplus, fognetsimplusplus, casanova2014simgrid, ROSS, casanova2020fgcs}.
 
@@
         systems  capture many relevant features of hardware/software stacks, 
-        they lack the ability to simulate page cache with enough details to capture key features such
+        most lack the ability to simulate page cache with enough details to capture key features such
         as dirty data and cache eviction policies~\cite{nunez2012simcan,nunez2012icancloud}. 
-        Some simulators, such as the one in~\cite{xu2018saving}, do capture such features,
+        Some simulators, such as the one in~\cite{xu2018saving}, do capture such features
         but are domain-specific.
@@
     implemented in \wrench~\cite{casanova2020fgcs}, a workflow simulation
-    framework based on the popular \simgrid distributed simulation
-    toolkit~\cite{casanova2014simgrid}. Our contributions are:
+    framework that builds on the \simgrid distributed systems simulation
+    framework~\cite{casanova2014simgrid}. Our contributions are:
     \begin{compactitem}
@@
 
-    \section{Related Work}
+    \section{Background and Related Work}
     \label{relatedwork}
@@
         at a predefined interval through a process known as \textit{periodical flushing}.
-        Periodical flushing only flushes expired dirty pages, which remain dirty in
+        Periodical flushing only flushes ``expired" dirty pages, that is, pages that
+        have remained dirty in the
         page cache longer than an expiration time configured in the kernel.
@@
 
-        In this  work, we use the \simgrid and \wrench simulation
-        frameworks.  The years of research and development invested in
-        the popular \simgrid simulation framework~\cite{casanova2014simgrid}, have
-        culminated in a set of state-of-the-art macroscopic simulation
-        models that yield high accuracy, as demonstrated by
-        (in)validation studies and comparisons to competing
-        frameworks~\cite{smpi_validity, velhoTOMACS2013, simutool_09,
-        nstools_07, lebre2015, pouilloux:hal-01197274,
-        smpi_tpds2017,  7885814, 8048921, 7384330}.  But one
-        significant drawback of \simgrid is that its simulation
-        abstractions are low-level, meaning that implementing a
-        simulator of complex systems can be
-        labor-intensive~\cite{kecskemeti_2014}. To remedy this problem,
-        the \wrench simulation framework~\cite{casanova2020fgcs}
-        builds on top of \simgrid to provide higher-level simulation
-        abstractions, so that simulators of complex applications and
-        systems can be implemented with a few hundred lines.
-
         Although the Linux page cache has a large impact on I/O
@@
         memory accesses and cached data~\cite{nunez2012icancloud}.
-        However, the scalability of the iCanCloud simulator is limited
-        as it uses microscopic models.  Besides, none
+        However, the scalability of these simulators is limited
+        as they rely on microscopic simulation models.  Besides, none
         of these simulators provide any writeback cache simulator nor
@@
 
+        In this  work, we use the \simgrid and \wrench simulation
+        frameworks.  The years of research and development invested in
+        \simgrid~\cite{casanova2014simgrid}, have
+        culminated in a set of state-of-the-art macroscopic simulation
+        models that yield high accuracy, as demonstrated by
+        (in)validation studies and comparisons to competing
+        frameworks~\cite{smpi_validity, velhoTOMACS2013, simutool_09,
+        nstools_07, lebre2015, pouilloux:hal-01197274,
+        smpi_tpds2017,  7885814, 8048921, 7384330}.  One
+        significant drawback of \simgrid is that its simulation
+        abstractions are low-level, meaning that implementing a
+        simulator of complex systems can be
+        labor-intensive~\cite{kecskemeti_2014}. To remedy this problem,
+        the \wrench simulation framework~\cite{casanova2020fgcs}
+        builds on top of \simgrid to provide higher-level simulation
+        abstractions, so that simulators of complex applications and
+        systems can be implemented with a few hundred lines of code. 
+        Neither \wrench nor \simgrid simulate page cache simulation.
+
 	In this study, we implement a page cache simulation model in the
-        \wrench framework. We targeted \wrench because it is a recent,
-        actively developed framework that provides convenient simulation
-        abstractions, because it is extensible, and because it reuses
-        \simgrid's scalable and accurate models.
+	\wrench framework. We target \wrench because it is a recent,
+	actively developed framework that provides convenient and
+	high-level simulation abstractions, because it is straightforward
+	to extend, and because it builds on \simgrid's scalable and
+	validated simulation models.
 
@@
     \label{method}
+    \todo[inline]{HC: Personally, I would actually make this subsection a section, and rename the previous section from "Methods" to "Page Cache Simulation Model".
+I've never loved the "Methods" title for section, although I know it is standard practice for many authors.
+}
+
     We separate our simulation model in two components, the I/O
@@
     where a block is not entirely read.
-    In this case, the block is split in two smaller blocks and one of them is re-accessed.
+    In this case, the block is split into two smaller blocks and one of them is re-accessed.
     % From Tristan: not sure where to put this nor if it's necessary:
@@
     flag to 0. In case the amount of data to flush requires that a
-    block be partially flushed, the block is split in two blocks,
+    block be partially flushed, the block is split into two blocks,
     one that is flushed and one that remains dirty. The time needed
@@
     or until there is no clean block left in the list.
-    If the last evicted block does not have to be entirely evicted, the block is split in two blocks,
+    If the last evicted block does not have to be entirely evicted, the block is split into two blocks,
     and only one of them is evicted.
@@
     \end{algorithm}
-    Algorithm~\ref{alg:write} describes our simulation of chunk writes in
-    the I/O Controller.
+    Algorithm~\ref{alg:write} describes how chunk writes in
+    the I/O Controller are simulated.
     % Data is written to cache until the amount of dirty data in the cache
     % exceeds the dirty ratio times the amount of available memory. Data is then written to disk.
-    Our algorithm initially checks the  amount of dirty data that
+    The algorithm initially checks the  amount of dirty data that
     can be written given the dirty ratio (line 5).
@@
     the remaining data is written to cache in a loop
-    where we repeatedly flush and evict from the cache (line 12-18).
+    where we repeatedly flush and evict data from the cache (line 12-18).
 
@@
     which simply simulates a disk write with the amount of data passed in,
-    then evicts cache if needed and adds the written data to the cache.
+    performs cache eviction if needed and adds the written data to the cache.
 
@@
             simulator to evaluate the accuracy and correctness of our
-            model in a simple scenario before integrating it in the more complex
+            model in a simple scenario before integrating it into the
             \wrench framework.
@@
 
-        \subsection{Experiments}
+        \subsection{Experimental Methodology}
 
-        Our experiments compared real executions with our Python prototype,
+        \todo[inline]{HC: I've renamed this subsection from "Experiments" to "Experimental Methodology", as otherwise I was expecting results in there.  Not a big deal.}
+        \todo[inline]{HC: I find it strange to have the experimental methodology section be a subsection of the "Methods" section, and then have the results in a new section.
+So I would actually make the "Experimental Methodology" a section, instead of a subsection.
+The flow would be: introduction; background and related work; methods; experimental methodology; results; conclusion.
+Again not a big deal, just a thought.}
+
+        Our experiments compare real executions with our Python prototype,
         with the original \wrench simulator, and with our \wrench-cache
@@
 
-        Experiments were run on a dedicated cluster at
+        Experiments are run on a dedicated cluster at
         Concordia University, with one login node, 9 compute nodes, and 4
         storage nodes connected with a 25 Gbps network. Each
-        compute node had 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
+        compute node has 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
         @ 2.10GHz, 250~GiB of RAM, 6 $\times$ SSDs of 450~GiB each with the XFS
         file system, 378~GiB of tmpfs, 126~GiB of devtmpfs file system,
-        CentOS~8.1 and NFS version 4. We used the \texttt{atop}
+        CentOS~8.1 and NFS version 4. We use the \texttt{atop}
         and \texttt{collectl} tools to monitor and collect memory status
-        and disk throughput. We cleared the page
+        and disk throughput. We clear the page
         cache before each application run to ensure comparable
@@
             \end{table}
-        The synthetic application, implemented in C, consisted of three single-core,
-        sequential tasks where each task read the file produced by the
-        previous task, incremented every byte of this file to emulate real
-        processing, and wrote the resulting data to disk. Files were
-        numbered by ascending access times (File 1 was the file read by Task 1, etc).
+        The synthetic application, implemented in C, consists of three single-core,
+        sequential tasks where each task reads the file produced by the
+        previous task, increments every byte of this file to emulate real
+        processing, and writes the resulting data to disk. Files are
+        numbered by ascending access times (File 1 is the file read by Task 1, etc).
          The anonymous memory used by the application
-        was released after each task, which we also simulated in the Python
-        prototype and in WRENCH-cache. As our focus was on I/O rather than compute, we measured
+        is released after each task, which we also simulate in the Python
+        prototype and in \wrench-cache. As our focus is on I/O rather than compute, we measure
         application task CPU times on a cluster node
-        (Table~\ref{table:cputime}), and used these durations in our
-        simulations. For the Python prototype, we injected CPU times
-        directly in the simulation. For \wrench and \wrench-cache, we
+        (Table~\ref{table:cputime}), and use these durations in our
+        simulations. For the Python prototype, we inject CPU times
+        directly into the simulation. For \wrench and \wrench-cache, we
         determined the corresponding number of flops on a 1~Gflops CPU
@@
         We used the synthetic application in three experiments. In the
-        first one (\textit{Exp~1}), we ran \emph{a single instance} of
+        first one (\textit{Exp~1}), we run \emph{a single instance} of
         the application on a single cluster node, with different input file
@@
         the same local disk.
-        In the second experiment (\textit{Exp~2}), we ran
+        In the second experiment (\textit{Exp~2}), we run
         \emph{concurrent} instances of the application on a single node,
         all application instances operating on different files of size 3~GB
-        stored in the same local disk. We varied the number of concurrent
-        application instances from 1 to 32 since cluster nodes had 32 CPU
+        stored in the same local disk. We vary the number of concurrent
+        application instances from 1 to 32 since cluster nodes have 32 CPU
         cores.
-        In the third experiment (\textit{Exp~3}), we used the same
+        In the third experiment (\textit{Exp~3}), we use the same
         configuration as the previous one, albeit reading and writing
@@
         another compute node. As is commonly configured in HPC
-        environments to avoid data loss, there was no client write cache
-        and the server cache was configured as writethrough instead of
-        writeback. NFS client and server read caches were enabled. 
-        Therefore, all the writes happened at disk bandwidth, but
-        reads could benefit from cache hits.
+        environments to avoid data loss, there is no client write cache
+        and the server cache is configured as writethrough instead of
+        writeback. NFS client and server read caches are enabled. 
+        Therefore, all the writes happen at disk bandwidth, but
+        reads can benefit from cache hits.
 
-        The real application was a workflow of the Nighres
-        toolbox~\cite{huntenburg2018nighres}, implementing cortical
+        The real application is a workflow of the Nighres
+        toolbox~\cite{huntenburg2018nighres}, which implements cortical
         reconstruction from brain images in four steps: skull stripping,
         tissue classification, region extraction, and cortical
-        reconstruction. Each step read files produced by the previous step,
-        and wrote files that were or were not read by the subsequent step.
+        reconstruction. Each step reads files produced by the previous step,
+        and writes files that are or are not read by the subsequent step.
         More information on this application is available in the Nighres
@@
         The application is implemented as a Python script that calls Java
-        image-processing routines. We used Python 3.6, Java 8, and Nighres
+        image-processing routines. We use Python 3.6, Java 8, and Nighres
         1.3.0. We patched the application to remove lazy data loading and
         data compression, which made CPU time difficult to separate from
-        I/O time, and to capture task CPU times to inject them in the
+        I/O time, and to capture task CPU times to inject in the
         simulation. The patched code is available at
@@
 
-        We used the real application in the fourth experiment
-        (\textit{Exp~4}), run on a single cluster node 
-        using a single local disk. We processed data from
+        We use the real application in the fourth experiment
+        (\textit{Exp~4}), which runs on a single cluster node 
+        using a single local disk, to process data from
         participant 0027430 in the dataset of the Max Planck Institute for
@@
 
-        To parameterize the simulators, we benchmarked the
+        To calibrate the simulators, we benchmark the
         memory, local disk, remote disk (NFS), and network bandwidths
@@
     \label{results}
+
+    \todo[inline]{HC: Not sure what "Absolute relative simulation errors" means in the caption of Figure~\ref{fig:single_error}. I assume it's a typo and you just mean "Relative"?}
 
@@
 
-        The page cache simulation model drastically reduced I/O simulation
+        \todo[inline]{Henri: In the paragraph below I would remove the ``The first read was ... data." sentence.}
+
+        The page cache simulation model drastically reduces I/O simulation
         errors in each application task (Figure~\ref{fig:single_error}). The first read was not impacted
@@
         files than with 20~GB files, due to the fact that part of the
-        100~GB file needed to be read and written to disk, the only storage
-        device in \wrench, as it did not fit in cache. Conversely,
-        simulation errors of the Python prototype and \wrench-cache were higher with
+        100~GB file needs to be read and written to disk, the only storage
+        device in \wrench, as it does not fit in cache. Conversely,
+        simulation errors of the Python prototype and \wrench-cache are higher with
         100~GB files than with 20~GB files, due to idiosyncrasies in the kernel
@@
 
-        Simulated memory profiles were highly consistent with the real ones
-        (Figure~\ref{fig:single_memprof}). With 20~GB files, memory profiles almost exactly matched the
-        real ones, although dirty data seemed to be flushing faster in real
+        Simulated memory profiles are highly consistent with the real ones
+        (Figure~\ref{fig:single_memprof}). With 20~GB files, memory profiles almost exactly match the
+        real ones, although dirty data seems to be flushing faster in real
         life than in simulation, a behavior also
-        observed with 100~GB files. With 100~GB files, used memory reached
+        observed with 100~GB files. With 100~GB files, used memory reaches
         total memory during the first write, triggering dirty data
-        flushing, and droped back to cached memory when application tasks
-        released anonymous memory. Simulated cached memory was highly
+        flushing, and drops back to cached memory when application tasks
+        release anonymous memory. Simulated cached memory is highly
         consistent with real values, except toward the end of Read 3 where
-        it slightly increased in simulation but not in reality. This
-        occurred due to the fact that after Write 2, File 3 was only partially
-        cached in simulation whereas it was entirely cached in the real
-        system. In all cases, dirty data remained under the dirty ratio as
-        expected. The Python prototype and \wrench-cache exhibited nearly
+        it slightly increases in simulation but not in reality. This
+        occurs because after Write 2 File 3 is only partially
+        cached in simulation whereas it is entirely cached in the real
+        system. In all cases, dirty data remains under the dirty ratio as
+        expected. The Python prototype and \wrench-cache exhibit nearly
         identical memory profiles, which reinforces the confidence in our
@@
 
-        The content of the simulated memory cache was also highly
+        The content of the simulated memory cache is also highly
         consistent with reality (Figure~\ref{fig:single_cache}). With 20~GB
-        files, the simulated cache content exactly matched reality, since
-        all files fitted in page cache. With 100~GB files, a slight
-        discrepancy was observed after Write 2, which explains the
-        simulation error previously mentioned in Read 3. In the real
-        execution indeed, File 3 was entirely cached after Write 2, whereas
-        in the simulated execution, only a part of it was cached. This was
-        due to the fact that the Linux kernel tends to not evict pages that
+        files, the simulated cache content exactly match reality, since
+        all files fit in page cache. With 100~GB files, a slight
+        discrepancy is observed after Write 2, which explains the
+        simulation error previously mentioned in Read 3. Indeed, in the real
+        execution File 3 is entirely cached after Write 2, whereas
+        in the simulated execution only part of it is cached. This is because
+        the Linux kernel tends to not evict pages that
         belong to files being currently written (File 3 in this case),
-        which we could not easily reproduce in our model.
+        which we cannot easily reproduce in our model.
 
@@
 
-            The page cache model notably reduced \wrench's simulation error
+            \todo[inline]{HC: I don't really see a plateau in Figure~\ref{fig:multi_local}}
+
+            The page cache model notably reduces \wrench's simulation error
             for concurrent applications executed with local I/Os
             (Figure~\ref{fig:multi_local}). For reads, \wrench-cache
-            slightly overestimated runtime, due to the discrepancy between
-            simulated and real read bandwidths mentioned before. 
-            For writes, \wrench-cache
-            retrieved a plateau similar to the one observed in the real
-            execution, marking the limit beyond which the page cache was
+            slightly overestimates runtime, due to the aforementioned discrepancy between
+            simulated and real read bandwidths.
+            For writes, \wrench-cache results show a
+            plateau similar to the one observed in the real
+            execution, marking the limit beyond which the page cache is
             saturated with dirty data and needed flushing.
@@
                 \end{figure}
+
         \subsection{Remote storage (Exp 3)}
 
-            Page cache simulation importantly reduced simulation error
-            on NFS storage as well (Figure~\ref{fig:multi_nfs}). This
-            manifested only for reads, as the NFS server used writethrough rather than writeback cache.
+            Page cache simulation  reduces simulation error significantly for
+            NFS storage as well (Figure~\ref{fig:multi_nfs}). This
+            is manifested only for reads, as the NFS server uses writethrough rather than writeback caching.
             Both \wrench and \wrench-cache
-            underestimated write times due to the discrepancy between
-            simulated and real bandwidths mentioned previously. For reads,
-            this discrepancy only impacted the results beyond 22
-            applications since before this threshold, most reads resulted in cache
+            underestimate write times due to the aforementioned discrepancy between
+            simulated and real bandwidths. For reads,
+            this discrepancy only impacts the results beyond 22
+            applications since before this threshold most reads result in cache
             hits.
@@
         \subsection{Real application (Exp 4)}
-        Similar to the synthetic application, simulation errors were
-        substantially reduced by the WRENCH-cache simulator compared to
-        WRENCH (Figure~\ref{fig:nighres}). On average, errors were reduced
-        from 337~\% in WRENCH to 47~\% in WRENCH-cache. 
-        The first read happened entirely from disk and was therefore 
-        very accurately simulated by both WRENCH and WRENCH-cache.
+
+        Similar to the synthetic application, simulation error for the real application is
+        substantially reduced by the \wrench-cache simulator compared to
+        \wrench (Figure~\ref{fig:nighres}). On average, errors are reduced
+        from 337~\% in \wrench to 47~\% in \wrench-cache. 
+        The first read happens entirely from disk and is thus 
+        very accurately simulated by both \wrench and \wrench-cache.
 
@@
         \subsection{Simulation time}
+
         As is the case for \wrench, simulation time with \wrench-cache scales
@@
     including the one developed in this work, 
-    which could substantially limit the accuracy of simulations.
+    which limits simulation accuracy.
     Relevant extensions to this work include more
     accurate descriptions of anonymous memory usage in applications, 
-    which strongly affects I/O times through writeback cache. File access patterns
+    which strongly affects I/O times through writeback caching. File access patterns
     might also be worth including in the simulation models,
     as they directly affect page cache content.
+
     % readahead and persistent storage could also be adeed

### 2021-05-20 paper_io_sim.tex
Fixes following @henricasanova's review

@@
 \usepackage{multirow}
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\usepackage{cite}
 
@@
 
-    \section{Methods}
+    \section{Page Cache Simulation Model}
     \label{method}
-    \todo[inline]{HC: Personally, I would actually make this subsection a section, and rename the previous section from "Methods" to "Page Cache Simulation Model".
-I've never loved the "Methods" title for section, although I know it is standard practice for many authors.
-}
 
@@
             \url{https://framagit.org/simgrid/simgrid}. Our implementation
-            is now part of \wrench's master branch and will be available to
-            users with the upcoming 1.8 release. \wrench provides a full \simgrid-based simulation 
+            is now part of \wrench's master branch and available to
+            users in the 1.8 release. \wrench provides a full \simgrid-based simulation 
             environment that supports, among other features, concurrent accesses to storage devices, 
@@
 
-        \subsection{Experimental Methodology}
+        \section{Experimental Methodology}
 
-        \todo[inline]{HC: I've renamed this subsection from "Experiments" to "Experimental Methodology", as otherwise I was expecting results in there.  Not a big deal.}
-        \todo[inline]{HC: I find it strange to have the experimental methodology section be a subsection of the "Methods" section, and then have the results in a new section.
+        % \todo[inline]{HC: I've renamed this subsection from "Experiments" to "Experimental Methodology", as otherwise I was expecting results in there.  Not a big deal.}
+        % \todo[inline]{HC: I find it strange to have the experimental methodology section be a subsection of the "Methods" section, and then have the results in a new section.
 So I would actually make the "Experimental Methodology" a section, instead of a subsection.
@@
 
-    \todo[inline]{HC: Not sure what "Absolute relative simulation errors" means in the caption of Figure~\ref{fig:single_error}. I assume it's a typo and you just mean "Relative"?}
-
     \begin{figure*}
@@
                \vspace*{-0.7cm}
-               \caption{Absolute relative simulation errors}
+               \caption{Relative simulation errors (absolute values)}
                \vspace*{0.5cm}
@@
 
-        \todo[inline]{Henri: In the paragraph below I would remove the ``The first read was ... data." sentence.}
+        % \todo[inline]{Henri: In the paragraph below I would remove the ``The first read was ... data." sentence.}
 
         The page cache simulation model drastically reduces I/O simulation
-        errors in each application task (Figure~\ref{fig:single_error}). The first read was not impacted
-        as it only involved uncached data. Errors were reduced from an average
+        errors in each application task (Figure~\ref{fig:single_error}). Errors were reduced from an average
         of 345\% in the original \wrench to 46\% in the Python prototype and
@@
         significantly overestimated read and write times, due to the lack
-        of page cache simulation. Results with files of 50~GB and 75~GB
+        of page cache simulation. Consistently, the first read was not impacted
+        as it only involved uncached data. Results with files of 50~GB and 75~GB
         showed similar behaviors and are not reported for brevity.
@@
 
-            \todo[inline]{HC: I don't really see a plateau in Figure~\ref{fig:multi_local}}
-
             The page cache model notably reduces \wrench's simulation error
             for concurrent applications executed with local I/Os
-            (Figure~\ref{fig:multi_local}). For reads, \wrench-cache
+            (Figure~\ref{fig:multi_local}). For reads (Figure~\ref{fig:multi_local}-left), \wrench-cache
             slightly overestimates runtime, due to the aforementioned discrepancy between
             simulated and real read bandwidths.
-            For writes, \wrench-cache results show a
+            For writes (Figure~\ref{fig:multi_local}-right), \wrench-cache results show a
             plateau similar to the one observed in the real
-            execution, marking the limit beyond which the page cache is
+            execution between 1 and 10 concurrent executions. The plateau marks the limit beyond which the page cache is
             saturated with dirty data and needed flushing.

### 2021-05-20 result/single/plot_error.py
Fix run time logs

@@ -46,10 +46,8 @@ def plot_error():
     fig, (ax1, ax2) = plt.subplots(figsize=(11, 3), ncols=2, nrows=1)
     plt.rcParams.update({'font.size': 10})
 
-    plot_task_error(ax1[0], 20, ylabel=False)
-    plot_task_error(ax1[1], 50, ylabel=False)
-    plot_task_error(ax2[0], 75, ylabel=False)
-    plot_task_error(ax2[1], 100, ylabel=False)
+    plot_task_error(ax1, 20, ylabel=False)
+    plot_task_error(ax2, 100, ylabel=False)
 
     fig.text(0.01, 0.5, 'memory (GB)', va='center', rotation='vertical')
 

### 2021-05-20 paper_io_sim.tex
A first pass, not the last one

@@
         a series of pages. Subsequent reads to any of the file pages located in memory will result in
-        a \textit{cache hit}, meaning the I/O can occur directly from without disk involvement.
-        Any accessed page not loaded in memory results in a \textit{cache miss}, resulting in
+        a cache hit, meaning the I/O can occur directly from without disk involvement.
+        Any accessed page not loaded in memory results in a cache miss, resulting in
         the page being read directly from disk.
@@
         there is sufficient available memory, written data will be flushed to disk
-        at a predefined interval through a process known as \textit{periodical flushing}.
+        at a predefined interval through a process known as periodical flushing.
         Periodical flushing only flushes ``expired" dirty pages, that is, pages that
@@
         based on storage bandwidths and capacities.  The SIMCAN
-        framework does models page caching by storing data accessed on
+        framework does model page caching by storing data accessed on
         disk in a block cache~\cite{nunez2012simcan}.  Page cache is
@@
            and disk accesses with the Memory Manager. Concurrent accesses to storage
-           devices (memory and disk) are simulated using existing models.}
+           devices (memory and disk) are simulated using the existing SimGrid models.}
            \label{fig:interaction}
@@
     expired dirty data in LRU lists and flushes this data to disk. We
-    use existing storage simulation models~\cite{lebre2015} to simulate disk and
+    rely on the existing SimGrid storage simulation models~\cite{lebre2015} to simulate disk and
     memory, characterized by their storage capacity, read and write
@@
     Our simulation model supports chunk-by-chunk file accesses
-    with a user-defined chunk size. However, for simplicity, we assume that file pages are
+    with a user-defined chunk size. However, since file access pattern models are currently not available in SimGrid, we assume that file pages are
     accessed in a round-robin fashion rather than fully randomly.
     Therefore, when a file is read, cached data is read only after all uncached data was read, and data from the inactive list is read
-    before data from the active list
-    (data reads occur from left to right in Figure~\ref{fig:read_order}).
+    before data from the active list. Such an access pattern is commonly encountered in scientific data processing, where 
+    applications often read data files entirely and possibly multiple times. 
     When a chunk of \emph{uncached} data is read, a new clean block is created
@@
     which simply simulates a disk write with the amount of data passed in,
-    performs cache eviction if needed and adds the written data to the cache.
+    performs cache eviction if needed and adds the written data to the
+    cache. The read function remains unchanged between the writeback and
+    writethrough modes.
 
@@
 
-            This prototype does not simulate  bandwidth sharing and thus does not support
-            concurrency: it is limited to single-threaded applications running on systems
-            with a single-core CPU. We used this prototype for a first validation of our simulation
-            model against a real sequential application running on a real system.
+            This prototype does not simulate  bandwidth sharing and thus
+            does not support concurrency: it is limited to single-threaded
+            applications running on systems with a single-core CPU. We used
+            this prototype for a first validation of our simulation model
+            against a real sequential application running on a real system,
+            and as a way to detect potential bugs in our implementation.
             The Python 3.7 source code is available at
@@
 
-        Experiments are run on a dedicated cluster at
+        Experiments were run on a dedicated cluster at
         Concordia University, with one login node, 9 compute nodes, and 4
@@
         file system, 378~GiB of tmpfs, 126~GiB of devtmpfs file system,
-        CentOS~8.1 and NFS version 4. We use the \texttt{atop}
+        CentOS~8.1 and NFS version 4. We used the \texttt{atop}
         and \texttt{collectl} tools to monitor and collect memory status
-        and disk throughput. We clear the page
+        and disk throughput. We cleared the page
         cache before each application run to ensure comparable
@@
         is released after each task, which we also simulate in the Python
-        prototype and in \wrench-cache. As our focus is on I/O rather than compute, we measure
+        prototype and in \wrench-cache. As our focus is on I/O rather than compute, we measured
         application task CPU times on a cluster node
-        (Table~\ref{table:cputime}), and use these durations in our
-        simulations. For the Python prototype, we inject CPU times
+        (Table~\ref{table:cputime}), and used these durations in our
+        simulations. For the Python prototype, we injected CPU times
         directly into the simulation. For \wrench and \wrench-cache, we
@@
         all application instances operating on different files of size 3~GB
-        stored in the same local disk. We vary the number of concurrent
+        stored in the same local disk. We varied the number of concurrent
         application instances from 1 to 32 since cluster nodes have 32 CPU
@@
         on a 50-GiB NFS-mounted partition of a 450-GiB remote disk of
-        another compute node. As is commonly configured in HPC
-        environments to avoid data loss, there is no client write cache
-        and the server cache is configured as writethrough instead of
-        writeback. NFS client and server read caches are enabled. 
-        Therefore, all the writes happen at disk bandwidth, but
-        reads can benefit from cache hits.
+        another compute node. NFS client and server read caches were enabled, however, as is commonly configured in HPC
+        environments to avoid data loss, there was no client write cache
+        and the server cache was configured as writethrough instead of
+        writeback.
+        Therefore, all the writes happened at disk bandwidth, but
+        reads could benefit from cache hits.
 
@@
         The application is implemented as a Python script that calls Java
-        image-processing routines. We use Python 3.6, Java 8, and Nighres
+        image-processing routines. We used Python 3.6, Java 8, and Nighres
         1.3.0. We patched the application to remove lazy data loading and
@@
 
-        We use the real application in the fourth experiment
+        We used the real application in the fourth experiment
         (\textit{Exp~4}), which runs on a single cluster node 
@@
 
-        To calibrate the simulators, we benchmark the
+        To calibrate the simulators, we benchmarked the
         memory, local disk, remote disk (NFS), and network bandwidths
@@
             \centering
-               \includegraphics[width=\linewidth]{result/single/figures/single_errors.pdf}
+               \includegraphics[width=\linewidth]{result/single/figures/single_errors_v2.pdf}
                \vspace*{-0.7cm}
@@
             %    Lines represent memory usage along pipeline execution time.}
-               \includegraphics[width=\linewidth]{result/single/figures/single_memprof.pdf}
+               \includegraphics[width=\linewidth]{result/single/figures/single_memprof_v2.pdf}
                \vspace*{-0.7cm}
@@
             \centering
-               \includegraphics[width=1.05\linewidth]{result/single/figures/cached_files.pdf}
-               \caption{Cache contents \emph{after} application I/O operations}
+               \includegraphics[width=1.05\linewidth]{result/single/figures/cached_files_v2.pdf}
+               \caption{Page cache content after each I/O operation}
             %    \textcolor{red}{Update real results of 20GB}}
@@
 
-        \wrench simulation errors were substantially lower with 100~GB
+        \wrench simulation errors are substantially lower with 100~GB
         files than with 20~GB files, due to the fact that part of the
-        100~GB file needs to be read and written to disk, the only storage
-        device in \wrench, as it does not fit in cache. Conversely,
-        simulation errors of the Python prototype and \wrench-cache are higher with
-        100~GB files than with 20~GB files, due to idiosyncrasies in the kernel
-        flushing and eviction strategies that could not be easily modeled.
+        100~GB file needs to be read and written to disk as it does not fit in cache. Disk models are
+        identical in WRENCH and WRENCH-cache.
 
@@
         (Figure~\ref{fig:single_memprof}). With 20~GB files, memory profiles almost exactly match the
-        real ones, although dirty data seems to be flushing faster in real
+        real ones, although for some reason dirty data seems to be flushing faster in real
         life than in simulation, a behavior also
@@
         release anonymous memory. Simulated cached memory is highly
-        consistent with real values, except toward the end of Read 3 where
-        it slightly increases in simulation but not in reality. This
-        occurs because after Write 2 File 3 is only partially
-        cached in simulation whereas it is entirely cached in the real
-        system. In all cases, dirty data remains under the dirty ratio as
+        consistent with real values. In all cases, dirty data remains under the dirty ratio as
         expected. The Python prototype and \wrench-cache exhibit nearly
@@
 
-        The content of the simulated memory cache is also highly
-        consistent with reality (Figure~\ref{fig:single_cache}). With 20~GB
-        files, the simulated cache content exactly match reality, since
-        all files fit in page cache. With 100~GB files, a slight
-        discrepancy is observed after Write 2, which explains the
-        simulation error previously mentioned in Read 3. Indeed, in the real
-        execution File 3 is entirely cached after Write 2, whereas
-        in the simulated execution only part of it is cached. This is because
-        the Linux kernel tends to not evict pages that
-        belong to files being currently written (File 3 in this case),
-        which we cannot easily reproduce in our model.
+        The content of the simulated memory cache is also highly consistent
+        with reality (Figure~\ref{fig:single_cache}), showing that the
+        previous observations were not incidental.
 
@@
                     \centering
-                    \includegraphics[width=\linewidth]{result/multi/figures/multi_local.pdf}
+                    \includegraphics[width=0.8\linewidth]{result/multi/figures/multi_local.pdf}
                 \end{subfigure}
@@
                 \label{fig:multi_local}
+            \end{figure*}
+            \begin{figure*}
+                \begin{subfigure}{\linewidth}
+                    \centering
+                    \includegraphics[width=0.8\linewidth]{result/multi/figures/multi_nfs.pdf}
+                \end{subfigure}
+                \caption{NFS results with 3~GB files (\textit{Exp 3})}
+                \label{fig:multi_nfs}
                 \end{figure*}
@@
             saturated with dirty data and needed flushing.
-
-            \begin{figure}[b]
-                \begin{subfigure}{0.95\linewidth}
-                    \centering
-                    \includegraphics[width=\linewidth]{result/nighres/figures/nighres_errors.pdf}
-                \end{subfigure}
-                \caption{Real application results (\textit{Exp 4})}
-                \label{fig:nighres}
-                \end{figure}
 
@@
 
-            \begin{figure*}
-                \begin{subfigure}{\linewidth}
+        \begin{figure}[b]
                 \centering
-                    \includegraphics[width=\linewidth]{result/multi/figures/multi_nfs.pdf}
-                \end{subfigure}
-                \caption{NFS results with 3~GB files (\textit{Exp 3})}
-                \label{fig:multi_nfs}
-                \end{figure*}
-
-            \begin{figure}
-                \begin{subfigure}{\columnwidth}
-                    \centering
-                    \includegraphics[width=\linewidth]{result/multi/figures/simulation_time.pdf}
-                \end{subfigure}
-                \caption{Simulation time comparison. \wrench-cache scales
-                linearly with the number of concurrent applications, albeit
-                with a higher overhead than \wrench.}
-                \label{fig:multi_time}
+                \includegraphics[width=0.95\linewidth]{result/nighres/figures/nighres_errors.pdf}
+            \caption{Real application results (\textit{Exp 4})}
+            \label{fig:nighres}
             \end{figure}
@@
         linearly with the number of concurrent applications
-        (Figure~\ref{fig:multi_time}, p \textless $10^{-24}$). However, the page
+        (Figure~\ref{fig:multi_time}, p \textless $10^{-17}$). However, the page
         cache model substantially increases simulation time by
@@
         cache in NFS, which bypasses flushing operations.
+
+        \begin{figure}
+            \centering
+            \includegraphics[width=0.95\linewidth]{result/multi/figures/simulation_time.pdf}
+            \caption{Simulation time comparison. \wrench-cache scales
+            linearly with the number of concurrent applications, albeit
+            with a higher overhead than \wrench.}
+            \label{fig:multi_time}
+            \end{figure}
 
@@
 
+\clearpage
 \bibliographystyle{IEEEtran}

### 2021-05-20 result/multi/process_result.py

@@ -186,10 +186,11 @@ def run_time():
 
     from scipy import stats
 
-    ax1.set_ylim(top=2)
+    ax1.set_ylim(top=0.4)
     s = 10
 
     # WRENCH original
+    print(org_df)
     slope, intercept, r_value, p_value, std_err = stats.linregress(org_df["no_pipeline"], org_df["run_time"])
     print(f'WRENCH original: p={p_value}')
     ax1.scatter(local_org_df["no_pipeline"], local_org_df["run_time"],
@@ -197,7 +198,7 @@ def run_time():
     ax1.scatter(nfs_org_df["no_pipeline"], nfs_org_df["run_time"],
                 label="WRENCH (NFS)", s=s, color=wrench_color, facecolor="none")
     ax1.plot(org_df["no_pipeline"], intercept + slope * org_df["no_pipeline"], lw=1, color=wrench_color)
-    plt.text(20, 0.1, f'y={round(slope,2)}x+{round(intercept,2)}', color=wrench_color)
+    plt.text(20, 0.025, f'y={round(slope,3)}x+{round(intercept,3)}', color=wrench_color)
 
     # WRENCH-cache Local I/Os
     slope, intercept, r_value, p_value, std_err = stats.linregress(local_cache_df["no_pipeline"],
@@ -207,7 +208,7 @@ def run_time():
              color=wrench_cache_color, lw=1)
     ax1.scatter(local_cache_df["no_pipeline"], local_cache_df["run_time"],
                 label="WRENCH-cache (local)", s=s, color=wrench_cache_color)
-    plt.text(17, 1, f'y={round(slope,2)}x{round(intercept,2)}', color=wrench_cache_color)
+    plt.text(15, 0.2, f'y={round(slope,3)}x{round(intercept,3)}', color=wrench_cache_color)
 
     # WRENCH-cache NFS
     slope, intercept, r_value, p_value, std_err = stats.linregress(nfs_cache_df["no_pipeline"],
@@ -217,7 +218,7 @@ def run_time():
              color=wrench_cache_color, lw=1, linestyle='dotted')
     ax1.scatter(nfs_cache_df["no_pipeline"], nfs_cache_df["run_time"],
                 label="WRENCH-cache (NFS)", s=s, color=wrench_cache_color, facecolor="none")
-    plt.text(27, 0.8, f'y={round(slope,2)}x{round(intercept,2)}', color=wrench_cache_color)
[…]

### 2021-05-20 paper_io_sim.tex
Fixed typo in yaxis label in Fig 4-a, text edits

@@
             This prototype does not simulate  bandwidth sharing and thus
-            does not support concurrency: it is limited to single-threaded
-            applications running on systems with a single-core CPU. We used
+            does not support concurrency. We used
             this prototype for a first validation of our simulation model
@@
         @ 2.10GHz, 250~GiB of RAM, 6 $\times$ SSDs of 450~GiB each with the XFS
-        file system, 378~GiB of tmpfs, 126~GiB of devtmpfs file system,
+        file system (no RAID enabled), 378~GiB of tmpfs, 126~GiB of devtmpfs file system,
         CentOS~8.1 and NFS version 4. We used the \texttt{atop}
@@
                     \centering
-                    \includegraphics[width=0.8\linewidth]{result/multi/figures/multi_local.pdf}
+                    \includegraphics[width=\linewidth]{result/multi/figures/multi_local.pdf}
                 \end{subfigure}
@@
                     \centering
-                    \includegraphics[width=0.8\linewidth]{result/multi/figures/multi_nfs.pdf}
+                    \includegraphics[width=\linewidth]{result/multi/figures/multi_nfs.pdf}
                 \end{subfigure}

### 2021-05-20 result/single/plot_error.py

@@ -49,7 +49,7 @@ def plot_error():
     plot_task_error(ax1, 20, ylabel=False)
     plot_task_error(ax2, 100, ylabel=False)
 
-    fig.text(0.01, 0.5, 'memory (GB)', va='center', rotation='vertical')
+    fig.text(0.01, 0.5, 'error (%)', va='center', rotation='vertical')
 
     lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.12, 2.6), ncol=3)
     plt.subplots_adjust(left=0.07, bottom=0.075, right=0.97, top=0.85, wspace=0.2, hspace=0.3)

### 2021-05-20 result/single/plot_error.py
Fixed error figure: log scale, layout

@@ -21,6 +21,7 @@ def grouped_bar_chart(ax, labels, xlabel, ylabel, *argv):
 
 def plot_task_error(ax, size, ylabel):
     labels = ["Read 1", "Write 1", "Read 2", "Write 2", "Read 3", "Write 3"]
+    ax.set_yscale('log')
 
     # atop_file =     "log/cluster/100gb/atop_mem.log"
     real_time_log = "real/%dgb/timestamps.csv" % size
@@ -44,16 +45,15 @@ def plot_task_error(ax, size, ylabel):
 
 def plot_error():
     fig, (ax1, ax2) = plt.subplots(figsize=(11, 3), ncols=2, nrows=1)
-    plt.rcParams.update({'font.size': 10})
+    plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.8, wspace=0.1)
+    plt.rcParams.update({'font.size': 9})
 
-    plot_task_error(ax1, 20, ylabel=False)
+    plot_task_error(ax1, 20, ylabel=True)
     plot_task_error(ax2, 100, ylabel=False)
 
-    fig.text(0.01, 0.5, 'error (%)', va='center', rotation='vertical')
-
-    lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.12, 2.6), ncol=3)
-    plt.subplots_adjust(left=0.07, bottom=0.075, right=0.97, top=0.85, wspace=0.2, hspace=0.3)
+    lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.12, 1.25), ncol=3)
 
     plt.savefig("figures/single_errors_v2.svg", format="svg", bbox_extra_artists=(lgd,))
     plt.savefig("figures/single_errors_v2.pdf", format="pdf", bbox_extra_artists=(lgd,))
+    plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.7, wspace=0.3)
     # plt.show()

### 2021-05-20 paper_io_sim.tex
adjusted layout in Fig 4

@@
                \vspace*{-0.7cm}
-               \caption{Relative simulation errors (absolute values)}
-               \vspace*{0.5cm}
+               \caption{Relative simulation errors (absolute values)\\ Left: 20-GB files. Right: 100-GB files.}
+                \vspace*{0.2cm}
                \label{fig:single_error}
@@
                \vspace*{-0.7cm}
-               \caption{Memory profiles}
-               \vspace*{0.5cm}
+               \caption{Memory profiles\\ From top to bottom: real execution, Python prototype, WRENCH cache. Left: 20-GB files. Right: 100-GB files.}
+               \vspace*{0.2cm}
                \label{fig:single_memprof}
@@
             \centering
-               \includegraphics[width=1.05\linewidth]{result/single/figures/cached_files_v2.pdf}
-               \caption{Page cache content after each I/O operation}
+               \includegraphics[width=\linewidth]{result/single/figures/cached_files_v2.pdf}
+               \caption{Page cache content after each I/O operation\\Left: 20-GB files. Right: 100-GB files.}
             %    \textcolor{red}{Update real results of 20GB}}

### 2021-05-20 result/single/plot_error.py

@@ -46,12 +46,12 @@ def plot_task_error(ax, size, ylabel):
 def plot_error():
     fig, (ax1, ax2) = plt.subplots(figsize=(11, 3), ncols=2, nrows=1)
     plt.subplots_adjust(left=0.1, bottom=0.1, right=0.95, top=0.8, wspace=0.1)
-    plt.rcParams.update({'font.size': 9})
+    plt.rcParams.update({'font.size': 10})
 
     plot_task_error(ax1, 20, ylabel=True)
     plot_task_error(ax2, 100, ylabel=False)
 
-    lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.12, 1.25), ncol=3)
+    lgd = plt.legend(loc='upper center', bbox_to_anchor=(-0.12, 1.3), ncol=3)
 
     plt.savefig("figures/single_errors_v2.svg", format="svg", bbox_extra_artists=(lgd,))
     plt.savefig("figures/single_errors_v2.pdf", format="pdf", bbox_extra_artists=(lgd,))

### 2021-05-21 result/nighres/process_result.py
Remove _v2 figures, remove old logs

@@ -172,6 +172,6 @@ def plot_error():
 
     plt.title("Nighres workflow simulation errors")
 
-    plt.savefig("figures/nighres_errors_v2.svg", format="svg", bbox_extra_artists=(lgd,))
-    plt.savefig("figures/nighres_errors_v2.pdf", format="pdf", bbox_extra_artists=(lgd,))
+    plt.savefig("figures/nighres_errors.svg", format="svg", bbox_extra_artists=(lgd,))
+    plt.savefig("figures/nighres_errors.pdf", format="pdf", bbox_extra_artists=(lgd,))
     # plt.show()

### 2021-05-21 result/single/plot_cache.py

@@ -63,6 +63,6 @@ def plot_cache_v2():
 
     fig.text(0.015, 0.45, 'memory (GB)', va='center', rotation='vertical')
     plt.legend(legend_elements, envs + files, ncol=6, loc='upper center', bbox_to_anchor=(-0.1, 1.3))
-    plt.savefig("figures/cached_files_v2.svg", format="svg")
-    plt.savefig("figures/cached_files_v2.pdf", format="pdf")
-    plt.show()
+    plt.savefig("figures/cached_files.svg", format="svg")
+    plt.savefig("figures/cached_files.pdf", format="pdf")
+    # plt.show()

### 2021-05-21 paper_io_sim.tex
Update cache eviction simulation, WRENCH commmit used

@@
     least recently used clean data blocks in the inactive list.
+    In case the amount evictable memory in the inactive list is insufficient 
+    (the inactive list is empty or a file being opened is residing in the inactive list), 
+    blocks in the active list are considered for eviction. 
     The amount of data to evict is passed as a parameter and data blocks are deleted
-    from the inactive list until the evicted data reaches the required amount,
+    from the lists until the evicted data reaches the required amount,
     or until there is no clean block left in the list.
@@
             \wrench 1.6 at commit
-            \href{https://github.com/wrench-project/wrench/tree/67185374330d2c4bf274fce222c937e838df5b03}{6718537433},
-            which uses \simgrid 3.25, available at
+            \href{https://github.com/dohoangdzung/wrench/tree/cefc0bf739d015b2329a62750a9d70085b65a3ae}{cefc0bf739},
+            which uses \simgrid 3.26, available at
             \url{https://framagit.org/simgrid/simgrid}. Our implementation

### 2021-05-21 paper_io_sim.tex
New average errors

@@
             \centering
-               \includegraphics[width=\linewidth]{result/single/figures/single_errors_v2.pdf}
+               \includegraphics[width=\linewidth]{result/single/figures/single_errors.pdf}
                \vspace*{-0.7cm}
@@
             %    Lines represent memory usage along pipeline execution time.}
-               \includegraphics[width=\linewidth]{result/single/figures/single_memprof_v2.pdf}
+               \includegraphics[width=\linewidth]{result/single/figures/single_memprof.pdf}
                \vspace*{-0.7cm}
@@
             \centering
-               \includegraphics[width=\linewidth]{result/single/figures/cached_files_v2.pdf}
+               \includegraphics[width=\linewidth]{result/single/figures/cached_files.pdf}
                \caption{Page cache content after each I/O operation\\Left: 20-GB files. Right: 100-GB files.}
@@
         errors in each application task (Figure~\ref{fig:single_error}). Errors were reduced from an average
-        of 345\% in the original \wrench to 46\% in the Python prototype and
-        39\% in \wrench-cache. Unsurprisingly, the original \wrench simulator
+        of 328\% in the original \wrench to 36\% in the Python prototype and
+        32\% in \wrench-cache. Unsurprisingly, the original \wrench simulator
         significantly overestimated read and write times, due to the lack

### 2021-05-24 result/single/plot_memprof.py
font size

@@ -146,7 +146,7 @@ def sim_subplot(subplot_ax, sim_time_file, sim_mem_file, title, bar_alpha=0.4, l
 def single_plot(input_size=20, makespan=165):
     fig, axes = plt.subplots(figsize=(9, 9), ncols=1, nrows=3)
     plt.subplots_adjust(top=0.85, hspace=0.3)
-    plt.rcParams.update({'font.size': 9})
+    plt.rcParams.update({'font.size': 10})
 
     compare_size([axes[0], axes[1], axes[2]],
                  real_time_file="real/%dgb/timestamps.csv" % input_size,
@@ -169,7 +169,7 @@ def single_plot(input_size=20, makespan=165):
 def compare_plot(input_sizes=[20, 100], makespan=[165, 1300]):
     fig, axes = plt.subplots(figsize=(15, 9), ncols=2, nrows=3)
     plt.subplots_adjust(hspace=0.3, wspace=0.1)
-    plt.rcParams.update({'font.size': 9})
+    plt.rcParams.update({'font.size': 10})
 
     real_condition_desc = "memory read bw = 6860 MBps\nmemory write bw = 2764 MBps\ndisk read bw = 510 MBps\ndisk write bw = 420 MBps"
     py_condition_desc = "memory bw = 4812MBps\ndisk bw = 465 MBps"
@@ -217,7 +217,7 @@ def compare_full():
 
     fig = plt.figure(figsize=(12, 16))
     outer = gridspec.GridSpec(ncols=2, nrows=2, wspace=0.15, hspace=0.2, left=0.05, right=0.95, bottom=0.1, top=0.9, figure=fig)
-    plt.rcParams.update({'font.size': 11})
+    plt.rcParams.update({'font.size': 10})
 
     # real_condition_desc = "memory read bw = 6860 MBps\nmemory write bw = 2764 MBps\ndisk read bw = 510 MBps\ndisk write bw = 420 MBps"
     # py_condition_desc = "memory bw = 4812MBps\ndisk bw = 465 MBps"

### 2021-05-24 paper_io_sim.tex
Text edits

@@
     performance studies of data-intensive applications can lead to misleading
-    results.
+    results and inaccurate conclusions.
 
@@
     reduces the simulation error by up to an order of magnitude when
-    compared to state-of-the-art, cacheless simulations.
+    compared to state-of-the-art, cacheless simulations. Our model is publicly available in the 
+    \wrench framework, making it usable in a wide range of simulation studies.
 
@@
     application performance, particularly for data-intensive applications.
-
  % Swapped the order the two paragraphs and merged them into one
@@
 
-        Page caching is an ubiquitous technique for mitigating the I/O bottleneck.
-        As such, it is necessary to model it
+    Page caching is an ubiquitous technique for mitigating the I/O
+    bottleneck, reducing disk read and write times by up to an order of
+    magnitude. As such, it is necessary to model it
     when simulating data-intensive applications.
@@
     Some simulators, such as the one in~\cite{xu2018saving}, do capture such features
-        but are domain-specific.
+    but are domain-specific. In addition, analytical models of the page
+    cache have been proposed ~\cite{DBLP:journals/concurrency/InacioD18,
+    DBLP:journals/pacmpl/BaoKPS18, DBLP:conf/pldi/GysiGBH19} but their
+    integration in simulation frameworks is challenging. A scalable, general and accurate
+    page cache simulation model is required to enable reliable simulations of
+    data-intensive applications.
 
@@
     and eviction, and reads or writes to disk when necessary.
-    In case the writethrough strategy is used, the I/O Controller directly writes to disk, 
-    cache is flushed if needed and written data is added to page cache.
 
@@
     expired dirty data in LRU lists and flushes this data to disk. We
-    rely on the existing SimGrid storage simulation models~\cite{lebre2015} to simulate disk and
+    rely on the existing SimGrid storage simulation models to simulate disk and
     memory, characterized by their storage capacity, read and write
-    bandwidths, and latency. These models account for
-    bandwidth sharing between concurrent memory or disk accesses.
+    bandwidths, and latency. These models account for bandwidth sharing
+    between concurrent memory or disk accesses. We refer the reader to~\cite{lebre2015} for a
+    detailed description and validation of these models.
 
@@
     one that is flushed and one that remains dirty. The time needed
-    to flush data to disk is simulated by the storage model.
+    to flush data to disk is simulated by the SimGrid storage model.
 
@@
     least recently used clean data blocks in the inactive list.
-    In case the amount evictable memory in the inactive list is insufficient 
-    (the inactive list is empty or a file being opened is residing in the inactive list), 
+    In case the amount of evictable memory in the inactive list is insufficient 
+    (the inactive list is empty or only contains blocks that belong to open files), 
     blocks in the active list are considered for eviction. 
@@
     from the lists until the evicted data reaches the required amount,
-    or until there is no clean block left in the list.
+    or until there is no clean block left in the lists.
     If the last evicted block does not have to be entirely evicted, the block is split into two blocks,
@@
     If this amount is greater than 0, the Memory Manager is requested to evict
-    data from cache if necessary (line 7).
+    data from cache if necessary to accomodate the chunk size (line 7).
     After eviction, the amount of data that can be written to
@@
 
-        Our experiments compare real executions with our Python prototype,
+        Our experiments compared real executions with our Python prototype,
         with the original \wrench simulator, and with our \wrench-cache
@@
         and used these values in the simulation. The simulated
-        platform and application are available at
+        platform and application are available in WRENCH at
         commit \href{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}{ec6b43561b}.
@@
         We used the synthetic application in three experiments. In the
-        first one (\textit{Exp~1}), we run \emph{a single instance} of
+        first one (\textit{Exp~1}), we ran \emph{a single instance} of
         the application on a single cluster node, with different input file
@@
         the same local disk.
-        In the second experiment (\textit{Exp~2}), we run
+        In the second experiment (\textit{Exp~2}), we ran
         \emph{concurrent} instances of the application on a single node,
@@
         cores.
-        In the third experiment (\textit{Exp~3}), we use the same
+        In the third experiment (\textit{Exp~3}), we used the same
         configuration as the previous one, albeit reading and writing
@@
         another compute node. NFS client and server read caches were enabled, however, as is commonly configured in HPC
-        environments to avoid data loss, there was no client write cache
+        environments to avoid data loss in netework file systems, there was no client write cache
         and the server cache was configured as writethrough instead of
@@
 
-        We used the real application in the fourth experiment
-        (\textit{Exp~4}), which runs on a single cluster node 
+        In the fourth experiment
+        (\textit{Exp~4}), we ran the real application on a single cluster node, 
         using a single local disk, to process data from
@@
         Human Cognitive and Brain Sciences available at
-        \url{http://dx.doi.org/10.15387/fcp_indi.corr.mpg1}, leading to the
-        parameters in Table~\ref{table:nighres_stats}.
+        \url{http://dx.doi.org/10.15387/fcp_indi.corr.mpg1}. The execution parameters are reported
+        in Table~\ref{table:nighres_stats}.
 
@@
 
-        The page cache simulation model drastically reduces I/O simulation
+        The page cache simulation model drastically reduced I/O simulation
         errors in each application task (Figure~\ref{fig:single_error}). Errors were reduced from an average
@@
 
-        \wrench simulation errors are substantially lower with 100~GB
+        \wrench simulation errors were substantially lower with 100~GB
         files than with 20~GB files, due to the fact that part of the
-        100~GB file needs to be read and written to disk as it does not fit in cache. Disk models are
+        100~GB file needed to be read and written to disk as it did not fit in cache. Disk models are
         identical in WRENCH and WRENCH-cache.
 
-        Simulated memory profiles are highly consistent with the real ones
-        (Figure~\ref{fig:single_memprof}). With 20~GB files, memory profiles almost exactly match the
-        real ones, although for some reason dirty data seems to be flushing faster in real
+        Simulated memory profiles were highly consistent with the real ones
+        (Figure~\ref{fig:single_memprof}). With 20~GB files, memory profiles almost exactly matched the
+        real ones, although for some reason dirty data seemed to be flushing faster in real
         life than in simulation, a behavior also
-        observed with 100~GB files. With 100~GB files, used memory reaches
+        observed with 100~GB files. With 100~GB files, used memory reached
         total memory during the first write, triggering dirty data
-        flushing, and drops back to cached memory when application tasks
-        release anonymous memory. Simulated cached memory is highly
-        consistent with real values. In all cases, dirty data remains under the dirty ratio as
-        expected. The Python prototype and \wrench-cache exhibit nearly
+        flushing, and droped back to cached memory when application tasks
+        released anonymous memory. Simulated cached memory was highly
+        consistent with real values. In all cases, dirty data remained under the dirty ratio as
+        expected. The Python prototype and \wrench-cache exhibited nearly
         identical memory profiles, which reinforces the confidence in our
@@
 
-        The content of the simulated memory cache is also highly consistent
+        The content of the simulated memory cache was also highly consistent
         with reality (Figure~\ref{fig:single_cache}), showing that the
@@
 
-            The page cache model notably reduces \wrench's simulation error
+            The page cache model notably reduced \wrench's simulation error
             for concurrent applications executed with local I/Os
-            (Figure~\ref{fig:multi_local}). For reads (Figure~\ref{fig:multi_local}-left), \wrench-cache
+            (Figure~\ref{fig:multi_local}). For reads (Figure~\ref{fig:multi_local}-left), \wrench-cache still
             slightly overestimates runtime, due to the aforementioned discrepancy between
@@
             plateau similar to the one observed in the real
-            execution between 1 and 10 concurrent executions. The plateau marks the limit beyond which the page cache is
+            execution between 1 and 10 concurrent executions. The plateau marks the limit beyond which the page cache was
             saturated with dirty data and needed flushing.
@@
 
-            Page cache simulation  reduces simulation error significantly for
+            Page cache simulation  reduced simulation error significantly for
             NFS storage as well (Figure~\ref{fig:multi_nfs}). This
-            is manifested only for reads, as the NFS server uses writethrough rather than writeback caching.
+            was manifested only for reads, as the NFS server used writethrough rather than writeback caching.
             Both \wrench and \wrench-cache
-            underestimate write times due to the aforementioned discrepancy between
+            underestimated write times due to the aforementioned discrepancy between
             simulated and real bandwidths. For reads,
-            this discrepancy only impacts the results beyond 22
-            applications since before this threshold most reads result in cache
+            this discrepancy only impacted the results beyond 22
+            applications since before this threshold most reads resulted in cache
             hits.
@@
 
-        Similar to the synthetic application, simulation error for the real application is
+        Similar to the synthetic application, simulation error for the real application was
         substantially reduced by the \wrench-cache simulator compared to
-        \wrench (Figure~\ref{fig:nighres}). On average, errors are reduced
+        \wrench (Figure~\ref{fig:nighres}). On average, errors were reduced
         from 337~\% in \wrench to 47~\% in \wrench-cache. 
-        The first read happens entirely from disk and is thus 
+        The first read happened entirely from disk and was thus 
         very accurately simulated by both \wrench and \wrench-cache.
@@
 
-        As is the case for \wrench, simulation time with \wrench-cache scales
-        linearly with the number of concurrent applications
-        (Figure~\ref{fig:multi_time}, p \textless $10^{-17}$). However, the page
-        cache model substantially increases simulation time by
+        Simulation time with \wrench-cache scaled
+        linearly with the number of concurrent applications, as was the case for \wrench
+        (Figure~\ref{fig:multi_time}, p \textless $10^{-16}$). However, the page
+        cache model substantially increased simulation time by
         application, as can be seen by comparing regression slopes in
-        Figure~\ref{fig:multi_time}. Interestingly, \wrench-cache is faster with 
+        Figure~\ref{fig:multi_time}. Interestingly, \wrench-cache was faster with 
         NFS I/Os than with local I/Os, most likely due to the use of writethrough
@@
             \includegraphics[width=0.95\linewidth]{result/multi/figures/simulation_time.pdf}
-            \caption{Simulation time comparison. \wrench-cache scales
+            \caption{Simulation time comparison. \wrench-cache scaled
             linearly with the number of concurrent applications, albeit
@@
     Evaluation results show that our model improves simulation accuracy
-    substantially, reducing absolute relative simulation errors by up to
-    9$\times$ (see results of the single-threaded experiment). The
+    substantially, reducing relative simulation errors by up to
+    10$\times$. The
     availability of asymmetrical disk bandwidths in the forthcoming
@@
     Our page cache model is publicly available in the \wrench GitHub
-    repository.
+    repository and can therefore be applied transparently to simulate data-intensive applications.
 
@@
     systems~\cite{islam2015triple}, active storage~\cite{5496981}, or
-    specific hardware architectures~\cite{hayot2020performance}.
+    specific hardware architectures~\cite{hayot2020performance}. We expect
+    page cache simulation to drastically reduce the simulation error in all
+    these scenarios, with substantial consequences on study conclusions.
 
-    Not all I/O behaviors are captured by currently available simulation models,
-    including the one developed in this work, 
-    which limits simulation accuracy.
     Relevant extensions to this work include more
@@
 
-\clearpage
 \bibliographystyle{IEEEtran}

### 2021-05-24 result/single/plot_memprof.py
added caption

@@ -199,7 +199,7 @@ def compare_plot(input_sizes=[20, 100], makespan=[165, 1300]):
 
     # fig.text(0.32, 0.06, 'time (s)', ha='center', fontsize=12)
     # fig.text(0.73, 0.06, 'time (s)', ha='center', fontsize=12)
-    # fig.text(0.08, 0.45, 'memory (GB)', va='center', rotation='vertical', fontsize=12)
+    fig.text(0.08, 0.5, 'memory (GB)', va='center', rotation='vertical', fontsize=12)
     # fig.text(0.91, 0.7, 'Real execution', fontsize=14, color='white', bbox={'pad': 5, 'color': '#90C987'},
     #          rotation=-90)
     # fig.text(0.91, 0.41, 'Python prototype', fontsize=14, color='white', bbox={'pad': 5, 'color': '#7BAFDE'},

### 2021-07-27 paper_io_sim.tex
fixed number of reps
Minor edits for camera-ready version

@@
         a series of pages. Subsequent reads to any of the file pages located in memory will result in
-        a cache hit, meaning the I/O can occur directly from without disk involvement.
+        a cache hit, meaning the I/O can occur directly from memory without disk involvement.
         Any accessed page not loaded in memory results in a cache miss, resulting in
@@
         systems can be implemented with a few hundred lines of code. 
-        Neither \wrench nor \simgrid simulate page cache simulation.
+        Neither \wrench nor \simgrid simulate page cache mechanisms.
 
@@
            \includegraphics[width=\columnwidth]{figures/lru_lists.pdf}
-           \caption{Model of page cache LRU lists with data blocks.}
+           \caption{Model of page cache LRU lists with data blocks. Grey blocks identify dirty data.}
            \label{fig:lrulist}
@@
     from cache (line 7-8). The amount of memory required to read the chunk
-    is calculated, corresponding to a copy of the chunk in anonymous memory
+    is calculated, corresponding to a copy of the chunk in anonymous memory --- the memory that is not backed by any file ---
     and a copy of the chunk in cache (line 9).
@@
         symmetrical bandwidths, we use the mean of the read and write
-        bandwidth values in our experiments.
+        bandwidth values in our experiments. We aso used symmetrical bandwidths in the Python prototype 
+        to facilitate comparison with \wrench-cache.
 
@@
                 \end{subfigure}
-                \caption{Concurrent results with 3~GB files (\textit{Exp 2})}
+                \caption{Concurrent results with 3~GB files, 5 repetitions (\textit{Exp 2})}
                 \label{fig:multi_local}
@@
                 \end{subfigure}
-                \caption{NFS results with 3~GB files (\textit{Exp 3})}
+                \caption{NFS results with 3~GB files, 5 repetitions (\textit{Exp 3})}
                 \label{fig:multi_nfs}

### 2021-07-27 paper_io_sim.tex
Fixed figure placement

@@
 
+        \begin{figure*}[h!]
+            \centering
+            \begin{subfigure}{0.95\linewidth}
+                \centering
+                \includegraphics[width=\linewidth]{result/multi/figures/multi_local.pdf}
+            \end{subfigure}
+            \caption{Concurrent results with 3~GB files, 5 repetitions (\textit{Exp 2})}
+            \label{fig:multi_local}
+        \end{figure*}
         \subsection{Single-threaded execution (Exp 1)}
-
+        \begin{figure*}[h!]
+            \centering
+            \begin{subfigure}{0.95\linewidth}
+                \centering
+                \includegraphics[width=\linewidth]{result/multi/figures/multi_nfs.pdf}
+            \end{subfigure}
+            \caption{NFS results with 3~GB files, 5 repetitions (\textit{Exp 3})}
+            \label{fig:multi_nfs}
+            \end{figure*}
         % \todo[inline]{Henri: In the paragraph below I would remove the ``The first read was ... data." sentence.}
@@
 
-            \begin{figure*}
-                \begin{subfigure}{\linewidth}
-                    \centering
-                    \includegraphics[width=\linewidth]{result/multi/figures/multi_local.pdf}
-                \end{subfigure}
-                \caption{Concurrent results with 3~GB files, 5 repetitions (\textit{Exp 2})}
-                \label{fig:multi_local}
-            \end{figure*}
-            \begin{figure*}
-                \begin{subfigure}{\linewidth}
-                    \centering
-                    \includegraphics[width=\linewidth]{result/multi/figures/multi_nfs.pdf}
-                \end{subfigure}
-                \caption{NFS results with 3~GB files, 5 repetitions (\textit{Exp 3})}
-                \label{fig:multi_nfs}
-                \end{figure*}
         \subsection{Concurrent applications (Exp 2)}
@@
 
-        \begin{figure}[b]
+        \begin{figure}[h!]
                 \centering
@@
         cache in NFS, which bypasses flushing operations.
-
-        \begin{figure}
-            \centering
-            \includegraphics[width=0.95\linewidth]{result/multi/figures/simulation_time.pdf}
-            \caption{Simulation time comparison. \wrench-cache scaled
-            linearly with the number of concurrent applications, albeit
-            with a higher overhead than \wrench.}
-            \label{fig:multi_time}
-            \end{figure}
 
@@
     these scenarios, with substantial consequences on study conclusions.
-
+    \begin{figure}[t]
+        \centering
+        \includegraphics[width=0.95\linewidth]{result/multi/figures/simulation_time.pdf}
+        \caption{Simulation time comparison. \wrench-cache scaled
+        linearly with the number of concurrent applications, albeit
+        with a higher overhead than \wrench.}
+        \label{fig:multi_time}
+        \end{figure}
     Relevant extensions to this work include more
